### 2023-07-18 ms.tex
[showyourwork] automatic showyourwork update

% Define document class
\documentclass[twocolumn]{aastex631}
\usepackage{showyourwork}

% Begin!
\begin{document}

% Title
\title{An open source scientific article}

% Author list
\author{@matiscke}

% Abstract with filler text
\begin{abstract}
    Lorem ipsum dolor sit amet, consectetuer adipiscing elit.
    Ut purus elit, vestibulum ut, placerat ac, adipiscing vitae, felis.
    Curabitur dictum gravida mauris, consectetuer id, vulputate a, magna.
    Donec vehicula augue eu neque, morbi tristique senectus et netus et.
    Mauris ut leo, cras viverra metus rhoncus sem, nulla et lectus vestibulum.
    Phasellus eu tellus sit amet tortor gravida placerat.
    Integer sapien est, iaculis in, pretium quis, viverra ac, nunc.
    Praesent eget sem vel leo ultrices bibendum.
    Aenean faucibus, morbi dolor nulla, malesuada eu, pulvinar at, mollis ac.
    Curabitur auctor semper nulla donec varius orci eget risus.
    Duis nibh mi, congue eu, accumsan eleifend, sagittis quis, diam.
    Duis eget orci sit amet orci dignissim rutrum.
\end{abstract}

% Main body with filler text
\section{Introduction}
\label{sec:intro}

Lorem ipsum dolor sit amet, consectetuer adipiscing elit.
Ut purus elit, vestibulum ut, placerat ac, adipiscing vitae, felis.
Curabitur dictum gravida mauris, consectetuer id, vulputate a, magna.
Donec vehicula augue eu neque, morbi tristique senectus et netus et.
Mauris ut leo, cras viverra metus rhoncus sem, nulla et lectus vestibulum.
Phasellus eu tellus sit amet tortor gravida placerat.
Integer sapien est, iaculis in, pretium quis, viverra ac, nunc.
Praesent eget sem vel leo ultrices bibendum.
Aenean faucibus, morbi dolor nulla, malesuada eu, pulvinar at, mollis ac.
Curabitur auctor semper nulla donec varius orci eget risus.
Duis nibh mi, congue eu, accumsan eleifend, sagittis quis, diam.
Duis eget orci sit amet orci dignissim rutrum.

Nam dui ligula, fringilla a, euismod sodales, sollici- tudin vel, wisi.
Morbi auctor lorem non justo, nam lacus libero, pretium at, lobortis vitae.
Donec aliquet, tortor sed accumsan bibendum, erat ligula aliquet magna.
Morbi ac orci et nisl hendrerit mollis, suspendisse ut massa, cras nec ante.
Pellentesque a nulla cum sociis natoque penatibus et magnis dis parturient.
Aliquam tincidunt urna, nulla ullamcorper vestibulum turpis.
Pellentesque cursus luctus mauris \citep{Luger2021}.

\bibliography{bib}

\end{document}

### 2023-07-18 ms.tex
Update on Overleaf.

@@
 % Title
-\title{An open source scientific article}
+\title{Bioverse: Origins of Life}
 
 % Author list
-\author{@matiscke}
+\author[0000-0001-8355-2107]{Martin Schlecker}
+\affiliation{Steward Observatory, The University of Arizona, Tucson, AZ 85721, USA; \href{mailto:schlecker@arizona.edu}{schlecker@arizona.edu}}
+\author[]{et al.}
 
@@
 
-Lorem ipsum dolor sit amet, consectetuer adipiscing elit.
-Ut purus elit, vestibulum ut, placerat ac, adipiscing vitae, felis.
-Curabitur dictum gravida mauris, consectetuer id, vulputate a, magna.
-Donec vehicula augue eu neque, morbi tristique senectus et netus et.
-Mauris ut leo, cras viverra metus rhoncus sem, nulla et lectus vestibulum.
-Phasellus eu tellus sit amet tortor gravida placerat.
-Integer sapien est, iaculis in, pretium quis, viverra ac, nunc.
-Praesent eget sem vel leo ultrices bibendum.
-Aenean faucibus, morbi dolor nulla, malesuada eu, pulvinar at, mollis ac.
-Curabitur auctor semper nulla donec varius orci eget risus.
-Duis nibh mi, congue eu, accumsan eleifend, sagittis quis, diam.
-Duis eget orci sit amet orci dignissim rutrum.
-
-Nam dui ligula, fringilla a, euismod sodales, sollici- tudin vel, wisi.
-Morbi auctor lorem non justo, nam lacus libero, pretium at, lobortis vitae.
-Donec aliquet, tortor sed accumsan bibendum, erat ligula aliquet magna.
-Morbi ac orci et nisl hendrerit mollis, suspendisse ut massa, cras nec ante.
-Pellentesque a nulla cum sociis natoque penatibus et magnis dis parturient.
-Aliquam tincidunt urna, nulla ullamcorper vestibulum turpis.
-Pellentesque cursus luctus mauris \citep{Luger2021}.
-
 \bibliography{bib}

### 2023-08-03 ms.tex
Update on Overleaf.

@@
 \documentclass[twocolumn]{aastex631}
-\usepackage{showyourwork}
-
-% Begin!
+\input{preamble}
 \begin{document}
@@
 \affiliation{Steward Observatory, The University of Arizona, Tucson, AZ 85721, USA; \href{mailto:schlecker@arizona.edu}{schlecker@arizona.edu}}
-\author[]{et al.}
+\author{et al.}
 
-% Abstract with filler text
 \begin{abstract}
-    Lorem ipsum dolor sit amet, consectetuer adipiscing elit.
-    Ut purus elit, vestibulum ut, placerat ac, adipiscing vitae, felis.
-    Curabitur dictum gravida mauris, consectetuer id, vulputate a, magna.
-    Donec vehicula augue eu neque, morbi tristique senectus et netus et.
-    Mauris ut leo, cras viverra metus rhoncus sem, nulla et lectus vestibulum.
-    Phasellus eu tellus sit amet tortor gravida placerat.
-    Integer sapien est, iaculis in, pretium quis, viverra ac, nunc.
-    Praesent eget sem vel leo ultrices bibendum.
-    Aenean faucibus, morbi dolor nulla, malesuada eu, pulvinar at, mollis ac.
-    Curabitur auctor semper nulla donec varius orci eget risus.
-    Duis nibh mi, congue eu, accumsan eleifend, sagittis quis, diam.
-    Duis eget orci sit amet orci dignissim rutrum.
+    tbd.
 \end{abstract}
 
-% Main body with filler text
 \section{Introduction}
 \label{sec:intro}
+
+\section{Origins of Life Scenarios}
+\label{sec:ool_scenarios}
+
+\subsection{Hydrothermal vents}
+e.g., Russell+2010
+
+\subsection{Subaerial ponds}
+
+\subsection{Planetary redox state and evolution}
+The synthesis of prebiotic compounds requires moderately to highly reduced chemical environments (Kitadai \& Maruyama 2018, Benner+2020, Sasselov+2020, Lichtenberg \& Clement 2022).
+
+\subsection{Cyanosulfidic scenario}
+
+\subsection{Impact trigger}
+Iron-rich impactors have been suggested to intermittently provide the reduced environments favored by prebiotic chemistry (e.g., Sekine+2003, Hashimoto+2007, Kuwahara \& Sugita 2015, Genda+2017, Wogan+2023).
+
+\section{Required conditions and predicted observables}
+\label{sec:observables}
+
+What are the environmental conditions necessary for the processes and reactions inherent to an OoL scenario to take place?
+What distinct observables connected to these conditions are accessible via present and near-future remote sensing techniques?
+
+\subsection{Hydrothermal vents}
+The hydrothermal vents scenario requires a direct contact of an ocean and the planetary mantle/crust. 
+This requirement is not met on an ocean world with large amounts of water, where the water pressure on the ocean floor is high enough to form high-pressure ices (Noack+2016).
+\todo[inline]{see discussion in Kite \& Ford 2018 Sect. 6.4}
+
+\subsection{Subaerial ponds}
+By its nature, the subaerial ponds scenario relies on rock surfaces exposed to the planetary atmosphere.
+Water worlds that have their entire planetary surface covered with water contradict this requirement and do not allow for the wet-dry cycling inherent to this origin of life scenario.
+The competition of tectonic stress with gravitational crustal spreading (Melosh 2011) sets the maximum possible height of mountains, which in the solar system does not exceed $\sim \SI{20}{\kilo\meter}$.
+Such mountains will be permanently under water on water worlds.
+Another impediment to wet-dry cycles is tidal locking of the planet as it stalls stellar tide-induced water movement and diurnal irradiation variability.
+
+\subsection{Redox state window}
+Surficial origins of life chemistries are dependent on the redox state of a planet being $\sim$neutral (not too reduced or oxidized) to allow the presence of precursor molecules like HCN.
+The planetary redox state leaves an imprint on its atmospheric composition and thus planet size (very reduced atmospheres are large) and spectral signatures.
+Connected to the cyanosulfidic scenario, the pond scenario, and the impact trigger.
+
+\subsection{Cyanosulfidic scenario}
+A strong requirement for the cyanosulfidic scenario is a sufficient Ultraviolet (UV) flux incident on the planet.
+On planets that have never received significant UV fluxes, the relevant photochemistry is limited.
+
+\subsection{Impact trigger}
+Prebiotic synthesis triggered through reduced impactors that stochastically create transiently reducing or neutral atmospheres requires a certain composition of the impactors, the planet to not be in a magma ocean state (???) (Lichtenberg \& Clement 2022), and, related to this requirement, occurrence of impact events during early planetary evolution.
+
+Observables: stochastic increase in brightness temperature, transient increase of planet size, and change of planet composition (decreasing with decreasing impact rate, ie stellar age).
+
+\section{Exoplanet survey simulations}
+\todo[inline]{Focus on Oxygen as a biomarker(?)}
+\subsection{Habitable Worlds Observatory}
+\subsubsection{target list}
+Provisional stellar target List for the habitable Worlds Observatory~\citep{Mamajek2023}.
+
+\subsection{Nautilus}
+\subsection{LIFE} % optional
+
+\section{Hypothesis tests}
+\label{sec:hypotests}
+\subsection{Possibility of hydrothermal vents}
+
+\subsection{Availability of rock-air interfaces}
+
+\subsection{Planetary redox state evolution}
+
+\subsection{Planetary UV flux and its evolution}
+
+\section{Results}
+\label{sec:results}
+
+\section{Discussion and Conclusions}
+\label{sec:discussion}
 

### 2023-08-05 ms.tex
Update on Overleaf.

@@
 
+\section{Bayesian Analysis}
+It is instructive to consider the constraining power of a successful biosignature detection for competing OoL scenarios, which we here attempt with an analytical approach.
+We may first consider the probability $\pdf(bio|H_i)$ of detecting a convincing biosignature on a planet, given a specific OoL hypothesis $H_i, \, i \in 1, 2, 3$ is true. 
+This can be decomposed into the probability of abiogenesis $\pdf_\mathrm{abio}$, the occurrence rate of the planet type required by the hypothesis $\eta_i$, and the probability of detecting the biosignature on this planet type with current technology $\pdf_\mathrm{det, i}$, yielding
+
+\begin{align}
+\label{eqn:pbio}
+\pdf(bio|H_i) = \pdf_\mathrm{abio} \times \eta_i \times \pdf_\mathrm{det, i}.
+\end{align}
+
+However, what we are actually interested in is the probability of a OoL hypothesis being true, given a specific biosignature detection, $\pdf(H_i|bio)$.
+To obtain this we can use Bayes' theorem, which yields 
+\begin{align}
+\pdf(H_i|bio) = \pdf(bio|H_i) \pdf(H_i) / \pdf(bio).
+\end{align}
+Putting everything together, we arrive at
+\begin{align}
+\label{eqn:posterior}
+\pdf(H_i|bio) = \frac{\pdf(H_i)}{\pdf(bio)} \times \pdf_\mathrm{abio} \times \eta_i \times \pdf_\mathrm{det, i}.
+\end{align}
+
+Here, $\pdf(bio)$ is the prior probability of detecting the biosignature.
+Following Eq.~\ref{eqn:pbio}, we can sum over all possible scenarios to obtain
+\begin{align}
+\pdf(bio) = \sum_{i=1}^{N} \pdf_\mathrm{abio} \times \eta_i \times \pdf_\mathrm{det, i}.
+\end{align}
+$\pdf(H_i)$ is the prior probability of the OoL hypothesis $H_i$.
+If we assume that all OoL hypotheses are a priori equally probable, we can treat $\frac{\pdf(H_i)}{\pdf(bio)}$ in Eq.~\ref{eqn:posterior} as a normalization constant.
+
+In the following, we discuss the impact of the remaining variables $\pdf_\mathrm{abio}, \eta_i, $ and $\pdf_\mathrm{det, i}$ on $\pdf(H_i|bio)$.
+
+\subsection{Probability of Abiogenesis $\pdf_\mathrm{abio}$}
+Following previous work~\citep{Spiegel2012,Chen2018,Kipping2021}, we may adopt a uniform rate model for abiogenesis, i.e., assume that OoL events occur at a uniform rate.
+This corresponds to a Poisson process with a rate parameter $lambda$, where we make the implicit assumption that abiogenesis occurs only via a single, unique mechanism, only once, and instantaneous.
+If this event occurs within a limited time window $t$, say, between planets form around a star and when it leaves the main sequence, we have
+\begin{align}
+\pdf_\mathrm{abio} = 1 - \exp(-\lambda t).
+\end{align}
+~\\
+...
+
+\subsection{Planet occurrence rate $\eta_i$}
+\subsection{Detection probability $\pdf_\mathrm{det, i}$}
+
 \section{Exoplanet survey simulations}
@@
 
+''Which natural processes best explain how living matter spontaneously appears from nonliving matter?''~\citep{Malaterre2022}
+%WE consider life ``as we know it''
+
 \bibliography{bib}

### 2023-08-05 src/tex/ms.tex
sync with structural draft on overleaf
add bibliography
draft Bayesian analysis

--- 2023-08-03 ms.tex
@@
 
+\section{Bayesian Analysis}
+It is instructive to consider the constraining power of a successful biosignature detection for competing OoL scenarios, which we here attempt with an analytical approach.
+We may first consider the probability $\pdf(bio|H_i)$ of detecting a convincing biosignature on a planet, given a specific OoL hypothesis $H_i, \, i \in 1, 2, 3$ is true.
+This can be decomposed into the probability of abiogenesis $\pdf_\mathrm{abio}$, the fraction of life-hosting worlds that develop atmospheric biosignatures $\pdf_\mathrm{sig}$, the probability of the planet type required by the OoL hypothesis occurring in the surveyed sample $\pdf_\mathrm{\eta, i}$, and the probability of detecting the biosignature on this planet type with current technology $\pdf_\mathrm{det, i}$, yielding
+
+\begin{align}
+\label{eqn:pbio}
+\pdf(bio|H_i) = \pdf_\mathrm{abio} \times \pdf_\mathrm{sig} \times \pdf_\mathrm{\eta, i} \times \pdf_\mathrm{det, i}.
+\end{align}
+
+However, what we are actually interested in is the probability of a OoL hypothesis being true, given a specific biosignature detection, $\pdf(H_i|bio)$.
+To obtain this we can use Bayes' theorem, which yields
+\begin{align}
+\pdf(H_i|bio) = \frac{\pdf(bio|H_i) \pdf(H_i)}{\pdf(bio)}.
+\end{align}
+Here, $\pdf(H_i)$ is the prior probability of the OoL hypothesis $H_i$, and $\pdf(bio)$ is the prior probability of detecting a biosignature.
+% Following Eq.~\ref{eqn:pbio}, we can sum over all possible scenarios to obtain
+% \begin{align}
+% \pdf(bio) = \sum_{i=1}^{N} \pdf_\mathrm{abio} \times \pdf_\mathrm{sig} \times \pdf_\mathrm{\eta, i} \times \pdf_\mathrm{det, i}.
+% \end{align}
+
+If the hypotheses $H_i$ are adjunct, i.e., their joint occurrence is impossible, one can show that
+\begin{align}
+\pdf(H_i|bio) = \frac{\pdf(bio|H_i) \pdf(H_i)}{\sum_{i=1}^{N} \pdf(bio|H_i) \pdf(H_i)}.
+\end{align}
+Then the parameters in Eq.~\ref{eqn:pbio} that are independent of the chosen hypothesis $H_i$ eliminate and we obtain
+\begin{align}
+\pdf(H_i|bio) = &\frac{\pdf_\mathrm{\eta, i} \pdf_\mathrm{det, i} \pdf(H_i)}{\sum_{i=1}^{N} \pdf_\mathrm{\eta, i} \pdf_\mathrm{det, i} \pdf(H_i)} \\
+\overset{\pdf(H_i) = \pdf(H_j)}{ \underset{\forall i,j \in \{1, 2, 3\}}{=}} &\frac{\pdf_\mathrm{\eta, i} \pdf_\mathrm{det, i}}{\sum_{i=1}^{N} \pdf_\mathrm{\eta, i} \pdf_\mathrm{det, i}},
+\end{align}
+where in the last step we made the implicit assumption that all OoL hypotheses are a priori equally probable.
+
+% Putting everything together, we arrive at
+% \begin{align}
+% \label{eqn:posterior}
+% \pdf(H_i|bio) = \frac{\pdf(H_i)}{\pdf(bio)} \times \pdf_\mathrm{abio} \times \pdf_\mathrm{sig} \times \pdf_\mathrm{\eta, i} \times \pdf_\mathrm{det, i}.
+% \end{align}
+% If we assume that all OoL hypotheses are a priori equally probable, we can treat $\frac{\pdf(H_i)}{\pdf(bio)}$ in Eq.~\ref{eqn:posterior} as a normalization constant.
+In the following, we discuss the impact of the remaining variables $\pdf_\mathrm{\eta, i}$ and $\pdf_\mathrm{det, i}$ on $\pdf(H_i|bio)$.
+
+% \subsection{Probability of Abiogenesis $\pdf_\mathrm{abio}$}
+% Following previous work~\citep{Spiegel2012,Chen2018,Kipping2021}, we may adopt a uniform rate model for abiogenesis, i.e., assume that OoL events occur at a uniform rate.
+% This corresponds to a Poisson process with a rate parameter $lambda$, where we make the implicit assumption that abiogenesis occurs only via a single, unique mechanism, only once, and instantaneous.
+% If this event occurs within a limited time window $t$, say, between planets form around a star and when it leaves the main sequence, we have
+% \begin{align}
+% \pdf_\mathrm{abio} = 1 - \exp(-\lambda t).
+% \end{align}
+% ~\\
+% ...
+
+% \subsection{biosignature fraction $\pdf_\mathrm{abio}$}
+% Limiting ourselves to the search for \textit{life-as-we-know-it}, we may assume that there are atmospheric biomarkers present on a planet after the abiogenesis event and until global extinction.
+
+\subsection{Fractional planet occurrence rate $\pdf_\mathrm{\eta, i}$}
+As the different OoL hypotheses do not all work on the same planet types, we may study the impact of the fractional occurrence rates of different planet types on the posterior probability $\pdf(H_i|bio)$.
+For simplicity, let's consider only planets that allow at least one of the scenarios.
+We also ignore any influence of other planets in the same system, e.g. an outer gas giant that itself does not develop life~\citep{Schlecker2021a} or panspermia scenarios CITE.
+We may then distinguish between:
+\begin{enumerate}
+    \item Earths, i.e., limited-water terrestrial planets in the liquid water habitable zone ($H1, H2, H3$).
+These planets with roughly Earth-like water mass fractions support both the existence of submarine hydrothermal vents ($H1$) and hydrothermal fields with wet/dry cycles ($H2$).
+Limits in exoplanet observables to this planet type are their orbital distance (both scenarios require liquid water; at least for FGK stars this requirement also puts the planet outside of the tidal locking zone ($H2$)), and bulk density ($H1$, $H2$, and $H3$ require limited water fractions).
+    \item ``shallow-water'' water worlds ($H1$). These planets have no land surface exposed to the atmosphere, thus excluding the subaerial pond scenario.
+    \item warm terrestrial planets ($H3$). These planets have smaller orbital distances, which serves as a proxy for high past UV fluxes (as required by $H3$).
+    \item ``deep ocean'' water worlds. Through the development of high-pressure ices, these planets do not support any of the considered OoL scenarios.
+\end{enumerate}
+
+\subsection{Detection probability $\pdf_\mathrm{det, i}$}
+
 \section{Exoplanet survey simulations}
-\todo[inline]{Focus on Oxygen as a biomarker(?)}
+\todo[inline]{Focus on molecular Oxygen as a biomarker(?)}
+A commonly discussed biomarker is molecular Oxygen (O$_2$), which on Earth emerged as a byproduct of photosynthesis during the Proterozoic era.
+While not the only atmospheric species discussed as a signature for planetary life CITE, we focused on O$_2$ as a representative biomarker for our spectroscopic survey simulations.
+
 \subsection{Habitable Worlds Observatory}

### 2023-08-09 ms.tex
Update on Overleaf.

--- 2023-08-03 ms.tex
@@
 
+\section{Bayesian Analysis}
+It is instructive to consider the constraining power of a successful biosignature detection for competing OoL scenarios, which we here attempt with an analytical approach.
+We may first consider the probability $\pdf(bio|H_i)$ of detecting a convincing biosignature on a planet, given a specific OoL hypothesis $H_i, \, i \in 1, 2, 3$ is true. 
+This can be decomposed into the probability of abiogenesis $\pdf_\mathrm{abio}$, the fraction of life-hosting worlds that develop atmospheric biosignatures $\pdf_\mathrm{sig}$, the probability of the planet type required by the OoL hypothesis occurring in the surveyed sample $\pdf_\mathrm{\eta, i}$, and the probability of detecting the biosignature on this planet type with current technology $\pdf_\mathrm{det, i}$, yielding
+
+\begin{align}
+\label{eqn:pbio}
+\pdf(bio|H_i) = \pdf_\mathrm{abio} \times \pdf_\mathrm{sig} \times \pdf_\mathrm{\eta, i} \times \pdf_\mathrm{det, i}.
+\end{align}
+
+However, what we are actually interested in is the probability of a OoL hypothesis being true, given a specific biosignature detection, $\pdf(H_i|bio)$.
+To obtain this we can use Bayes' theorem, which yields 
+\begin{align}
+\pdf(H_i|bio) = \frac{\pdf(bio|H_i) \pdf(H_i)}{\pdf(bio)}.
+\end{align}
+Here, $\pdf(H_i)$ is the prior probability of the OoL hypothesis $H_i$, and $\pdf(bio)$ is the prior probability of detecting a biosignature.
+% Following Eq.~\ref{eqn:pbio}, we can sum over all possible scenarios to obtain
+% \begin{align}
+% \pdf(bio) = \sum_{i=1}^{N} \pdf_\mathrm{abio} \times \pdf_\mathrm{sig} \times \pdf_\mathrm{\eta, i} \times \pdf_\mathrm{det, i}.
+% \end{align}
+
+If the hypotheses $H_i$ are adjunct, i.e., their joint occurrence is impossible, one can show that
+\begin{align}
+\pdf(H_i|bio) = \frac{\pdf(bio|H_i) \pdf(H_i)}{\sum_{i=1}^{N} \pdf(bio|H_i) \pdf(H_i)}.
+\end{align}
+Then the parameters in Eq.~\ref{eqn:pbio} that are independent of the chosen hypothesis $H_i$ eliminate and we obtain
+\begin{align}
+\pdf(H_i|bio) = &\frac{\pdf_\mathrm{\eta, i} \pdf_\mathrm{det, i} \pdf(H_i)}{\sum_{i=1}^{N} \pdf_\mathrm{\eta, i} \pdf_\mathrm{det, i} \pdf(H_i)} \\
+\overset{\pdf(H_i) = \pdf(H_j)}{ \underset{\forall i,j \in \{1, 2, 3\}}{=}} &\frac{\pdf_\mathrm{\eta, i} \pdf_\mathrm{det, i}}{\sum_{i=1}^{N} \pdf_\mathrm{\eta, i} \pdf_\mathrm{det, i}},
+\end{align}
+where in the last step we made the implicit assumption that all OoL hypotheses are a priori equally probable.
+
+% Putting everything together, we arrive at
+% \begin{align}
+% \label{eqn:posterior}
+% \pdf(H_i|bio) = \frac{\pdf(H_i)}{\pdf(bio)} \times \pdf_\mathrm{abio} \times \pdf_\mathrm{sig} \times \pdf_\mathrm{\eta, i} \times \pdf_\mathrm{det, i}.
+% \end{align}
+% If we assume that all OoL hypotheses are a priori equally probable, we can treat $\frac{\pdf(H_i)}{\pdf(bio)}$ in Eq.~\ref{eqn:posterior} as a normalization constant. 
+In the following, we discuss the impact of the remaining variables $\pdf_\mathrm{\eta, i}$ and $\pdf_\mathrm{det, i}$ on $\pdf(H_i|bio)$.
+
+% \subsection{Probability of Abiogenesis $\pdf_\mathrm{abio}$}
+% Following previous work~\citep{Spiegel2012,Chen2018,Kipping2021}, we may adopt a uniform rate model for abiogenesis, i.e., assume that OoL events occur at a uniform rate.
+% This corresponds to a Poisson process with a rate parameter $lambda$, where we make the implicit assumption that abiogenesis occurs only via a single, unique mechanism, only once, and instantaneous.
+% If this event occurs within a limited time window $t$, say, between planets form around a star and when it leaves the main sequence, we have
+% \begin{align}
+% \pdf_\mathrm{abio} = 1 - \exp(-\lambda t).
+% \end{align}
+% ~\\
+% ...
+
+% \subsection{biosignature fraction $\pdf_\mathrm{abio}$}
+% Limiting ourselves to the search for \textit{life-as-we-know-it}, we may assume that there are atmospheric biomarkers present on a planet after the abiogenesis event and until global extinction.
+
+\subsection{Fractional planet occurrence rate $\pdf_\mathrm{\eta, i}$}
+As the different OoL hypotheses do not all work on the same planet types, we may study the impact of the fractional occurrence rates of different planet types on the posterior probability $\pdf(H_i|bio)$.
+For simplicity, let's consider only planets that allow at least one of the scenarios.
+We also ignore any influence of other planets in the same system, e.g. an outer gas giant that itself does not develop life~\citep{Schlecker2021a} or panspermia scenarios CITE.
+We may then distinguish between:
+\begin{enumerate}
+    \item Earths, i.e., limited-water terrestrial planets in the liquid water habitable zone ($H1, H2, H3$).
+These planets with roughly Earth-like water mass fractions support both the existence of submarine hydrothermal vents ($H1$) and hydrothermal fields with wet/dry cycles ($H2$).
+Limits in exoplanet observables to this planet type are their orbital distance (both scenarios require liquid water; at least for FGK stars this requirement also puts the planet outside of the tidal locking zone ($H2$)), and bulk density ($H1$, $H2$, and $H3$ require limited water fractions).
+    \item ``shallow-water'' water worlds ($H1$). These planets have no land surface exposed to the atmosphere, thus excluding the subaerial pond scenario. 
+    \item warm terrestrial planets ($H3$). These planets have smaller orbital distances, which serves as a proxy for high past UV fluxes (as required by $H3$). 
+    \item ``deep ocean'' water worlds. Through the development of high-pressure ices, these planets do not support any of the considered OoL scenarios.
+\end{enumerate}
+
+\subsection{Detection probability $\pdf_\mathrm{det, i}$}
+
 \section{Exoplanet survey simulations}
-\todo[inline]{Focus on Oxygen as a biomarker(?)}
+\todo[inline]{Focus on molecular Oxygen as a biomarker(?)}
+A commonly discussed biomarker is molecular Oxygen (O$_2$), which on Earth emerged as a byproduct of photosynthesis during the Proterozoic era. 
+While not the only atmospheric species discussed as a signature for planetary life CITE, we focused on O$_2$ as a representative biomarker for our spectroscopic survey simulations.
+
 \subsection{Habitable Worlds Observatory}

### 2023-08-11 src/tex/ms.tex
remove ORCID to debug syw
showyourwork seems to complain when my ORCID is included in \author:
"
Preprocess: Setting up the workflow...
Traceback (most recent call last):
  File "/Users/schlecker/repos/originsoflife/.snakemake/scripts/tmps6etcd0c.preprocess.py", line 495, in <module>
    config["tree"] = get_json_tree(snakemake.input[0])
  File "/Users/schlecker/repos/originsoflife/.snakemake/scripts/tmps6etcd0c.preprocess.py", line 411, in get_json_tree
    free_floating_graphics = [
  File "/Users/schlecker/repos/originsoflife/.snakemake/scripts/tmps6etcd0c.preprocess.py", line 414, in <list

@@
 % Author list
-\author[0000-0001-8355-2107]{Martin Schlecker}
+\author{Martin Schlecker}
+%\author[0000-0001-8355-2107]{Martin Schlecker}
 \affiliation{Steward Observatory, The University of Arizona, Tucson, AZ 85721, USA; \href{mailto:schlecker@arizona.edu}{schlecker@arizona.edu}}
@@
 \section{Bayesian Analysis}
+\todo[inline]{consider Fisher matrix (Kendall \& Stuart 1977; Tegmark 1997) analysis to determine sensitivity of a survey to a set of parameters?}
 It is instructive to consider the constraining power of a successful biosignature detection for competing OoL scenarios, which we here attempt with an analytical approach.
-We may first consider the probability $\pdf(bio|H_i)$ of detecting a convincing biosignature on a planet, given a specific OoL hypothesis $H_i, \, i \in 1, 2, 3$ is true.
+The hydrothermal vents scenario ($H_1$) and the subaereal pond scenario ($H_2$) can be considered as mutually exclusive models, and we may study how a particular future observation of biosignatures impacts our beliefs about the relative model probabilities.
+
+We may first consider the probability $\pdf(bio|H_i)$ of detecting a convincing biosignature on a planet, given a particular OoL hypothesis $H_i, \, i \in 1, 2, 3$ is true.
 This can be decomposed into the probability of abiogenesis $\pdf_\mathrm{abio}$, the fraction of life-hosting worlds that develop atmospheric biosignatures $\pdf_\mathrm{sig}$, the probability of the planet type required by the OoL hypothesis occurring in the surveyed sample $\pdf_\mathrm{\eta, i}$, and the probability of detecting the biosignature on this planet type with current technology $\pdf_\mathrm{det, i}$, yielding
@@
 
-However, what we are actually interested in is the probability of a OoL hypothesis being true, given a specific biosignature detection, $\pdf(H_i|bio)$.
+However, what we are actually interested in is the probability of a OoL hypothesis being true, given a particular biosignature detection, $\pdf(H_i|bio)$.
 To obtain this we can use Bayes' theorem, which yields
@@
 Limits in exoplanet observables to this planet type are their orbital distance (both scenarios require liquid water; at least for FGK stars this requirement also puts the planet outside of the tidal locking zone ($H2$)), and bulk density ($H1$, $H2$, and $H3$ require limited water fractions).
-    \item ``shallow-water'' water worlds ($H1$). These planets have no land surface exposed to the atmosphere, thus excluding the subaerial pond scenario.
+    \item ``shallow ocean'' water worlds ($H1$). These planets have no land surface exposed to the atmosphere, thus excluding the subaerial pond scenario.
     \item warm terrestrial planets ($H3$). These planets have smaller orbital distances, which serves as a proxy for high past UV fluxes (as required by $H3$).
@@
 \subsection{Detection probability $\pdf_\mathrm{det, i}$}
+
+\subsection{Discussion of Bayesian analysis}
+
+\begin{note}
+    shallow ocean planets are vulnerable to water loss through high-energy radiation, limiting the time window for habitability especially if no geochemical feedbacks exist~\citep{Kite2018}.
+\end{note}
+\todo[inline]{discuss host star spectral type dependencies, e.g., abiogenesis time window for G stars ($\lessapprox 10 Gyr$, MS lifetime) or M dwarfs. \citep[e.g.,][]{Spiegel2012}}
 

### 2023-08-11 src/tex/ms.tex
remove H3 from Bayesian analysis

@@
 
-We may first consider the probability $\pdf(bio|H_i)$ of detecting a convincing biosignature on a planet, given a particular OoL hypothesis $H_i, \, i \in 1, 2, 3$ is true.
+We may first consider the probability $\pdf(bio|H_i)$ of detecting a convincing biosignature on a planet, given a particular OoL hypothesis $H_i, \, i \in 1, 2$ is true.
 This can be decomposed into the probability of abiogenesis $\pdf_\mathrm{abio}$, the fraction of life-hosting worlds that develop atmospheric biosignatures $\pdf_\mathrm{sig}$, the probability of the planet type required by the OoL hypothesis occurring in the surveyed sample $\pdf_\mathrm{\eta, i}$, and the probability of detecting the biosignature on this planet type with current technology $\pdf_\mathrm{det, i}$, yielding
@@
 \pdf(H_i|bio) = &\frac{\pdf_\mathrm{\eta, i} \pdf_\mathrm{det, i} \pdf(H_i)}{\sum_{i=1}^{N} \pdf_\mathrm{\eta, i} \pdf_\mathrm{det, i} \pdf(H_i)} \\
-\overset{\pdf(H_i) = \pdf(H_j)}{ \underset{\forall i,j \in \{1, 2, 3\}}{=}} &\frac{\pdf_\mathrm{\eta, i} \pdf_\mathrm{det, i}}{\sum_{i=1}^{N} \pdf_\mathrm{\eta, i} \pdf_\mathrm{det, i}},
+\overset{\pdf(H_i) = \pdf(H_j)}{ \underset{\forall i,j \in \{1, 2\}}{=}} &\frac{\pdf_\mathrm{\eta, i} \pdf_\mathrm{det, i}}{\sum_{i=1}^{N} \pdf_\mathrm{\eta, i} \pdf_\mathrm{det, i}},
 \end{align}
@@
 \begin{enumerate}
-    \item Earths, i.e., limited-water terrestrial planets in the liquid water habitable zone ($H1, H2, H3$).
+    \item Earths, i.e., limited-water terrestrial planets in the liquid water habitable zone ($H1, H2$).
 These planets with roughly Earth-like water mass fractions support both the existence of submarine hydrothermal vents ($H1$) and hydrothermal fields with wet/dry cycles ($H2$).
-Limits in exoplanet observables to this planet type are their orbital distance (both scenarios require liquid water; at least for FGK stars this requirement also puts the planet outside of the tidal locking zone ($H2$)), and bulk density ($H1$, $H2$, and $H3$ require limited water fractions).
+Limits in exoplanet observables to this planet type are their orbital distance (both scenarios require liquid water; at least for FGK stars this requirement also puts the planet outside of the tidal locking zone ($H2$)), and bulk density ($H1$ and $H2$ require limited water fractions).
     \item ``shallow ocean'' water worlds ($H1$). These planets have no land surface exposed to the atmosphere, thus excluding the subaerial pond scenario.
-    \item warm terrestrial planets ($H3$). These planets have smaller orbital distances, which serves as a proxy for high past UV fluxes (as required by $H3$).
     \item ``deep ocean'' water worlds. Through the development of high-pressure ices, these planets do not support any of the considered OoL scenarios.

### 2023-08-11 src/tex/ms.tex
combine OoL scenarios and their predictions into one section

@@
 
-\section{Origins of Life Scenarios}
+\section{Origins of Life Scenarios and their predictions}
 \label{sec:ool_scenarios}
+In this section, we present some of the most prominent origins of life scenarios and their observational predictions.
+We focus on the necessary environmental conditions for the processes and reactions inherent to each scenario, and aim to identify distinct observables that are accessible via present and near-future remote sensing techniques.
 
@@
 e.g., Russell+2010
-
-\subsection{Subaerial ponds}
-
-\subsection{Planetary redox state and evolution}
-The synthesis of prebiotic compounds requires moderately to highly reduced chemical environments (Kitadai \& Maruyama 2018, Benner+2020, Sasselov+2020, Lichtenberg \& Clement 2022).
-
-\subsection{Cyanosulfidic scenario}
-
-\subsection{Impact trigger}
-Iron-rich impactors have been suggested to intermittently provide the reduced environments favored by prebiotic chemistry (e.g., Sekine+2003, Hashimoto+2007, Kuwahara \& Sugita 2015, Genda+2017, Wogan+2023).
-
-\section{Required conditions and predicted observables}
-\label{sec:observables}
-
-What are the environmental conditions necessary for the processes and reactions inherent to an OoL scenario to take place?
-What distinct observables connected to these conditions are accessible via present and near-future remote sensing techniques?
-
-\subsection{Hydrothermal vents}
+...
 The hydrothermal vents scenario requires a direct contact of an ocean and the planetary mantle/crust.
@@
 \subsection{Subaerial ponds}
+...
 By its nature, the subaerial ponds scenario relies on rock surfaces exposed to the planetary atmosphere.
@@
 
-\subsection{Redox state window}
+\subsection{Cyanosulfidic scenario}
+...
+A strong requirement for the cyanosulfidic scenario is a sufficient Ultraviolet (UV) flux incident on the planet.
+On planets that have never received significant UV fluxes, the relevant photochemistry is limited.
+
+\subsection{Other Processes related to the Origins of Life}
+\subsubsection{Planetary redox state and evolution}
+The synthesis of prebiotic compounds requires moderately to highly reduced chemical environments (Kitadai \& Maruyama 2018, Benner+2020, Sasselov+2020, Lichtenberg \& Clement 2022).
+...
 Surficial origins of life chemistries are dependent on the redox state of a planet being $\sim$neutral (not too reduced or oxidized) to allow the presence of precursor molecules like HCN.
@@
 
-\subsection{Cyanosulfidic scenario}
-A strong requirement for the cyanosulfidic scenario is a sufficient Ultraviolet (UV) flux incident on the planet.
-On planets that have never received significant UV fluxes, the relevant photochemistry is limited.
-
-\subsection{Impact trigger}
+\subsubsection{Impact trigger}
+Iron-rich impactors have been suggested to intermittently provide the reduced environments favored by prebiotic chemistry (e.g., Sekine+2003, Hashimoto+2007, Kuwahara \& Sugita 2015, Genda+2017, Wogan+2023).
+...
 Prebiotic synthesis triggered through reduced impactors that stochastically create transiently reducing or neutral atmospheres requires a certain composition of the impactors, the planet to not be in a magma ocean state (???) (Lichtenberg \& Clement 2022), and, related to this requirement, occurrence of impact events during early planetary evolution.
-
-Observables: stochastic increase in brightness temperature, transient increase of planet size, and change of planet composition (decreasing with decreasing impact rate, ie stellar age).
+Suggested observables are stochastic increases in brightness temperature, transient increases of planet size, and change of planet composition (decreasing with decreasing impact rate, i.e., stellar age).
 

### 2023-08-11 src/tex/ms.tex
introduce Bayes factors

@@
 
+If we take the ratio of these posteriors for our two independent hypotheses $H_1$ and $H_2$, we get the \textit{Bayes Factor}
+\begin{align}
+    \frac{\pdf(H_1|bio)}{\pdf(H_2|bio)} = \frac{\pdf_\mathrm{\eta, 1} \pdf_\mathrm{det, 1}}{\pdf_\mathrm{\eta, 2} \pdf_\mathrm{det, 2}},
+\end{align}
+which quantifies the evidence of the data arising from $H_1$ versus $H_2$.
+
 % Putting everything together, we arrive at
@@
 % If we assume that all OoL hypotheses are a priori equally probable, we can treat $\frac{\pdf(H_i)}{\pdf(bio)}$ in Eq.~\ref{eqn:posterior} as a normalization constant.
-In the following, we discuss the impact of the remaining variables $\pdf_\mathrm{\eta, i}$ and $\pdf_\mathrm{det, i}$ on $\pdf(H_i|bio)$.
+In the following, we discuss the impact of the remaining variables $\pdf_\mathrm{\eta, i}$ and $\pdf_\mathrm{det, i}$ on the Bayes factor.
 

### 2023-08-12 src/tex/ms.tex
rename P_abio, develop motivation

@@
 We may first consider the probability $\pdf(bio|H_i)$ of detecting a convincing biosignature on a planet, given a particular OoL hypothesis $H_i, \, i \in 1, 2$ is true.
-This can be decomposed into the probability of abiogenesis $\pdf_\mathrm{abio}$, the fraction of life-hosting worlds that develop atmospheric biosignatures $\pdf_\mathrm{sig}$, the probability of the planet type required by the OoL hypothesis occurring in the surveyed sample $\pdf_\mathrm{\eta, i}$, and the probability of detecting the biosignature on this planet type with current technology $\pdf_\mathrm{det, i}$, yielding
+This can be decomposed into the probability of abiogenesis in a particular environment $\pdf_\mathrm{env, i}$, the fraction of life-hosting worlds that develop atmospheric biosignatures $\pdf_\mathrm{sig}$, the probability of the planet type required by the OoL hypothesis occurring in the surveyed sample $\pdf_\mathrm{\eta, i}$, and the probability of detecting the biosignature on this planet type with current technology $\pdf_\mathrm{det, i}$, yielding
 
@@
 \label{eqn:pbio}
-\pdf(bio|H_i) = \pdf_\mathrm{abio} \times \pdf_\mathrm{sig} \times \pdf_\mathrm{\eta, i} \times \pdf_\mathrm{det, i}.
+\pdf(bio|H_i) = \pdf_\mathrm{env, i} \times \pdf_\mathrm{sig} \times \pdf_\mathrm{\eta, i} \times \pdf_\mathrm{det, i}.
 \end{align}
@@
 % \begin{align}
-% \pdf(bio) = \sum_{i=1}^{N} \pdf_\mathrm{abio} \times \pdf_\mathrm{sig} \times \pdf_\mathrm{\eta, i} \times \pdf_\mathrm{det, i}.
+% \pdf(bio) = \sum_{i=1}^{N} \pdf_\mathrm{env, i} \times \pdf_\mathrm{sig} \times \pdf_\mathrm{\eta, i} \times \pdf_\mathrm{det, i}.
 % \end{align}
@@
 \begin{align}
-\pdf(H_i|bio) = &\frac{\pdf_\mathrm{\eta, i} \pdf_\mathrm{det, i} \pdf(H_i)}{\sum_{i=1}^{N} \pdf_\mathrm{\eta, i} \pdf_\mathrm{det, i} \pdf(H_i)} \\
-\overset{\pdf(H_i) = \pdf(H_j)}{ \underset{\forall i,j \in \{1, 2\}}{=}} &\frac{\pdf_\mathrm{\eta, i} \pdf_\mathrm{det, i}}{\sum_{i=1}^{N} \pdf_\mathrm{\eta, i} \pdf_\mathrm{det, i}},
+\pdf(H_i|bio) = &\frac{\pdf_\mathrm{env, i} \pdf_\mathrm{\eta, i} \pdf_\mathrm{det, i} \pdf(H_i)}{\sum_{i=1}^{N} \pdf_\mathrm{env, i} \pdf_\mathrm{\eta, i} \pdf_\mathrm{det, i} \pdf(H_i)} \\
+\overset{\pdf(H_i) = \pdf(H_j)}{ \underset{\forall i,j \in \{1, 2\}}{=}} &\frac{\pdf_\mathrm{env, i} \pdf_\mathrm{\eta, i} \pdf_\mathrm{det, i}}{\sum_{i=1}^{N} \pdf_\mathrm{env, i} \pdf_\mathrm{\eta, i} \pdf_\mathrm{det, i}},
 \end{align}
@@
 \begin{align}
-    \frac{\pdf(H_1|bio)}{\pdf(H_2|bio)} = \frac{\pdf_\mathrm{\eta, 1} \pdf_\mathrm{det, 1}}{\pdf_\mathrm{\eta, 2} \pdf_\mathrm{det, 2}},
+    \frac{\pdf(H_1|bio)}{\pdf(H_2|bio)} = \frac{\pdf_\mathrm{env, 1} \pdf_\mathrm{\eta, 1} \pdf_\mathrm{det, 1}}{\pdf_\mathrm{env, 2} \pdf_\mathrm{\eta, 2} \pdf_\mathrm{det, 2}},
 \end{align}
@@
 % \label{eqn:posterior}
-% \pdf(H_i|bio) = \frac{\pdf(H_i)}{\pdf(bio)} \times \pdf_\mathrm{abio} \times \pdf_\mathrm{sig} \times \pdf_\mathrm{\eta, i} \times \pdf_\mathrm{det, i}.
+% \pdf(H_i|bio) = \frac{\pdf(H_i)}{\pdf(bio)} \times \pdf_\mathrm{env, i} \times \pdf_\mathrm{sig} \times \pdf_\mathrm{\eta, i} \times \pdf_\mathrm{det, i}.
 % \end{align}
 % If we assume that all OoL hypotheses are a priori equally probable, we can treat $\frac{\pdf(H_i)}{\pdf(bio)}$ in Eq.~\ref{eqn:posterior} as a normalization constant.
-In the following, we discuss the impact of the remaining variables $\pdf_\mathrm{\eta, i}$ and $\pdf_\mathrm{det, i}$ on the Bayes factor.
+In the following, we discuss the impact of the remaining variables $\pdf_\mathrm{env, i}$, $\pdf_\mathrm{\eta, i}$ and $\pdf_\mathrm{det, i}$ on the Bayes factor.
 
-% \subsection{Probability of Abiogenesis $\pdf_\mathrm{abio}$}
+ \subsection{Required environment $\pdf_\mathrm{env, i}$}
 % Following previous work~\citep{Spiegel2012,Chen2018,Kipping2021}, we may adopt a uniform rate model for abiogenesis, i.e., assume that OoL events occur at a uniform rate.
@@
 % \begin{align}
-% \pdf_\mathrm{abio} = 1 - \exp(-\lambda t).
+% \pdf_\mathrm{env, i} = 1 - \exp(-\lambda t).
 % \end{align}
 % ~\\
-% ...
+\todo[inline]{The planet should be in the liquid-water HZ}
+...
+Let us assume that $H_1$ only requires a minimum bulk density $\rho_1$, such that $\pdf_\mathrm{env, 1} = 1$ if $\rho > \rho_1$ and zero otherwise.
+On the other hand, $H_2$ requires exposed land and thus a small water mass fraction.
+We may implement this in the same way as above but with a minimum bulk density $\rho_2 > \rho_1$.
+Furthermore, there is a requirement that the tidal locking timescale may not be so small that the planet is likely tidally locked at observation.
+This translates to imposing a minimum semimajor axis $a_2$.
+...
+...
 
-% \subsection{biosignature fraction $\pdf_\mathrm{abio}$}
+% \subsection{biosignature fraction $\pdf_\mathrm{env, i}$}
 % Limiting ourselves to the search for \textit{life-as-we-know-it}, we may assume that there are atmospheric biomarkers present on a planet after the abiogenesis event and until global extinction.

### 2023-08-12 src/scripts/plotstyle.py
add plot styles

@@ -1,0 +1,99 @@
+""" Config file with my personal preference of default matplotlibrc settings.
+"""
+from matplotlib import rc, rcParams
+from cycler import cycler
+
+
+def get_color_palette():
+    """ return a custom made color palette."""
+    return ['#029e73', '#fc4f30', '#008fd5',
+            '#e5ae38', '#810f7c', '#00035b', '#005249',
+            '#8b8b8b', '#fe828c']
+
+
+def styleplots():
+    # Set plot style
+    spinewidth = 1.8
+
+    font = {'family': 'sans',
+            'weight': 'normal',
+            'size': 14}
+    legend = {'handlelength': 0.75,
+              'handletextpad': 0.5}
+    figure = {'dpi': 100,
+              'subplot.left': 0.16,  # the left side of the subplots of the figure
+              'subplot.bottom': 0.21,  # the bottom of the subplots of the figure
+              'subplot.right': 0.98,  # the right side of the subplots of the figure
+              'subplot.top': 0.97,  # the top of the subplots of the figure
+              'subplot.hspace': 0.0}  # height reserved for space between subplots
+
+    rc('font', **font)
+    rc('legend', **legend)
+    rc('figure', **figure)
+    rc('lines', linewidth=2.5)
+
+    rc('axes', linewidth=spinewidth)
+    rc('xtick.major', width=spinewidth)
+    rc('ytick.major', width=spinewidth)
+    rc('xtick.minor', width=.5 * spinewidth)
+    rc('ytick.minor', width=.5 * spinewidth)
+    rc('image', cmap='inferno')
+    rc('hist', bins=20)
+    rc('patch', edgecolor='black')
+    rc('savefig', bbox='tight', dpi=200)
+
+    rcParams['axes.prop_cycle'] = cycler(color=get_color_palette())
+
+
+def set_size(width='aa', subplot=[1,1], scale=1):
+    """ Set aesthetic figure dimensions to avoid scaling in latex.
+
+    Parameters
+    ----------
+    width: float or string
+        either width in pts or one of the following strings:
+        'aa' : A&A column width
+        'aaDouble' : A&A total text width
+    subplot : list
+        subplot dimensions in [rows, columns]
+    scale: float
+        fraction of the width which you wish the figure to occupy
+
+    Returns
+    -------
+    fig_dim: tuple
+        Dimensions of figure in inches
+
+    Credits
[…]

### 2023-08-16 src/tex/ms.tex
model P_env and show overlaps in plot

@@
 ...
-Let us assume that $H_1$ only requires a minimum bulk density $\rho_1$, such that $\pdf_\mathrm{env, 1} = 1$ if $\rho > \rho_1$ and zero otherwise.
+Let us assume that $H_1$ only requires a minimum bulk density $\rho_1$, such that $\pdf_\mathrm{env, 1} \rightarrow 1$ for $\rho >> \rho_1$ and $\pdf_\mathrm{env, 1} \rightarrow 0$ for $\rho << \rho_1$.
 On the other hand, $H_2$ requires exposed land and thus a small water mass fraction.
@@
 This translates to imposing a minimum semimajor axis $a_2$.
-...
+
+To approximate these thresholds including their expected intrinsic fuzziness, we model them with logistic sigmoid functions
+\begin{align}
+    \pdf_\mathrm{env, 1} &= \frac{1}{1+\exp[-(C  (\rho - \rho_1))]} \quad \mathrm{and}\\
+    \pdf_\mathrm{env, 2} &= \frac{1}{1+\exp[-(C  (\rho - \rho_2))]} \times \frac{1}{1+\exp[-(C  (a - a_2))]},
+\end{align}
+where $C$ is a compression factor characterizing the steepness of the sigmoid function.
+\begin{figure}[ht!]
+    \script{bayes_rho-a.py}
+    \begin{centering}
+        \includegraphics[width=\linewidth]{figures/analytic/Penv.pdf}
+        \caption{
+        Probability of abiogenesis for the OoL hypotheses $H_1$ and $H_2$ as a function of planet bulk density and semi-major axis.
+        $H_1$ only requires large enough densities to exclude deep global water oceans.
+            $H_2$ requires a higher minimum density due to the exposed land requirement, and small semi-major axes are excluded to prevent tidal locking.
+        }
+        \label{fig:Penv}
+    \end{centering}
+\end{figure}
+Figure~\ref{fig:Penv} shows the corresponding $\pdf_\mathrm{env}$ factors and where their regions of high probability overlap.
 ...

### 2023-08-16 src/scripts/bayes_rho-a.py

@@ -1,0 +1,135 @@
+import paths
+import numpy as np
+import matplotlib.pyplot as plt
+import plotstyle
+import astropy.units as u
+from astropy.constants import G, M_sun
+from scipy.special import expit, log_expit
+
+plotstyle.styleplots()
+
+rhoE = 3 / 4 * u.earthMass / (np.pi * u.earthRad**3)
+
+
+def p_h1(rrho, aa, rho1=0.6, f_compress=20):
+    """Environment factor for H1: above threshold density rho1"""
+    # return np.array([[1 if (rho > rho1) else 0 for rho in rrho] for a in aa])
+    return np.array([expit(f_compress * (rrho - rho1)) for a in aa])
+
+
+def p_h2(rrho, aa, rho2=0.7, a2=0.01, f_compress=20):
+    """Environment factor for H2: above threshold density rho1, sma a2"""
+    # return np.array([[1 if ((rho > rho2) & (a > a2)) else 0 for rho in rrho] for a in aa])
+    return np.array(
+        [
+            [
+                expit(f_compress * (rho - rho2)) * expit(f_compress * (a - a2))
+                for rho in rrho
+            ]
+            for a in aa
+        ]
+    )
+
+
+def eval_Penv():
+    """Evaluate environment factors on a grid in rho, a."""
+    aa = np.geomspace(1e-3, 2.0, 30)
+    rrho = np.linspace(0.2, 1.6, 30)
+    P1 = p_h1(rrho, aa)
+    P2 = p_h2(rrho, aa)
+    return aa, rrho, P1, P2
+
+
+def eval_Peta(aa, rrho):
+    """Evaluate planet occurrence rate density on a grid in rho, a."""
+    # SAG13 power law parameters
+    R_break = 3.4
+    gamma = [0.38, 0.73]
+    alpha = [-0.19, -1.18]
+    beta = [0.26, 0.59]
+
+    def sma2period(a):
+        """Convert semi-major axis to period."""
+        return np.sqrt(4 * np.pi**2 * a * u.au**3 / (M_sun * G)).to(u.day).value
+
+    def rho2radius(rho):
+        """Convert bulk density to radius, assuming Earth density."""
[…]

### 2023-08-16 src/tex/ms.tex
include occurrence rates, show Bayes factor grid

@@
 \begin{align}
+    \label{eq:bayesfactor}
     \frac{\pdf(H_1|bio)}{\pdf(H_2|bio)} = \frac{\pdf_\mathrm{env, 1} \pdf_\mathrm{\eta, 1} \pdf_\mathrm{det, 1}}{\pdf_\mathrm{env, 2} \pdf_\mathrm{\eta, 2} \pdf_\mathrm{det, 2}},
@@
 Figure~\ref{fig:Penv} shows the corresponding $\pdf_\mathrm{env}$ factors and where their regions of high probability overlap.
-...
+
+\subsection{Planet occurrence rate $\pdf_{\eta}$}
+We model $\pdf_{\eta, i} (a, \rho)$ following the suggested broken power-law occurrence rates from NASA’s Exoplanet Program Analysis Group chartered Science Analysis Group 13 (SAG 13)~\citep[see][]{Bixel2021} and converting between semi-major axis and period, and between bulk density and radius assuming Earth-like orbits and compositions.
+\todo[inline]{this is an oversimplification.}
+\begin{figure}[ht!]
+    \script{bayes_rho-a.py}
+    \begin{centering}
+        \includegraphics[width=\linewidth]{figures/analytic/Peta.pdf}
+        \caption{
+        Planet occurrence rate density assuming SAG~13 occurrence rates.
+        }
+        \label{fig:Peta}
+    \end{centering}
+\end{figure}
+The resulting occurrence rate density has a strong preference for low-density planets on short orbits (see Fig.~\ref{fig:Peta}).
+\todo[inline]{This assumes the same occurrence rate for both hypotheses.
+Is this sensible?
+If yes, it does not impact the Bayes factor (Eq.
+\ref{eq:bayesfactor}); we could bring it only later to see where high Bayes factor and planet occurrence overlap.}
+
+\subsection{Information content of a biosignature detection}
+We may now evaluate Eqn.~\ref{eq:bayesfactor} to measure the information content with respect to favoring $H_1$ versus $H_2$ depending on the position of a planet with a confirmed biosignature detection in density-orbital distance space.
+\begin{figure}[ht!]
+    \script{bayes_rho-a.py}
+    \begin{centering}
+        \includegraphics[width=\linewidth]{figures/analytic/bayes_rho-a.pdf}
+        \caption{
+        Bayes factor (Eqn.~\ref{eq:bayesfactor}) evaluated at different bulk densities and orbital distances.
+        Contour levels reflect the empirical scale for strength of evidence suggested by Jeffreys et al. 19XX.
+        There, $\ln(\mathcal{B}) = 2.5$ corresponds to ``moderate'' evidence, and $\ln(\mathcal{B}) = 5 $ corresponds to ``strong'' evidence.
+        Only short orbits ($a \lessapprox \SI{0.03}{\au}$) and low bulk densities ($\rho \lessapprox \SI{0.6}{\rho_\oplus}$) allow a selection between the proposed models; they lead to a strong preference for hypothesis $H_1$.
+        There is no region in this parameter space that provides strong evidence for $H_2$.
+        }
+        \label{fig:bayes_rho-a}
+    \end{centering}
+\end{figure}
+Figure~\ref{fig:bayes_rho-a} shows the logarithm of the Bayes factor in this space, providing a scale for evaluating the strength of evidence to prefer one of the proposed models.
+Only a small region allows for a significant model selection: While short orbits ($a \lessapprox \SI{0.03}{\au}$) and low bulk densities ($\rho \lessapprox \SI{0.6}{\rho_\oplus}$) strongly support $H_1$, no combination of bulk density and orbital distance provide strong evidence for $H_2$ over $H_1$ without additional information.
+\todo[inline]{factor in Detection probability P_det}
+\todo[inline]{TODO: test sensitivity of this result on the assumed function and thresholds for P_env}
 
@@
 
-\subsection{Fractional planet occurrence rate $\pdf_\mathrm{\eta, i}$}
-As the different OoL hypotheses do not all work on the same planet types, we may study the impact of the fractional occurrence rates of different planet types on the posterior probability $\pdf(H_i|bio)$.
-For simplicity, let's consider only planets that allow at least one of the scenarios.
-We also ignore any influence of other planets in the same system, e.g. an outer gas giant that itself does not develop life~\citep{Schlecker2021a} or panspermia scenarios CITE.
-We may then distinguish between:
-\begin{enumerate}
-    \item Earths, i.e., limited-water terrestrial planets in the liquid water habitable zone ($H1, H2$).
+%\subsection{Fractional planet occurrence rate $\pdf_\mathrm{\eta, i}$}
+%As the different OoL hypotheses do not all work on the same planet types, we may study the impact of the fractional occurrence rates of different planet types on the posterior probability $\pdf(H_i|bio)$.
+%For simplicity, let's consider only planets that allow at least one of the scenarios.
+%We also ignore any influence of other planets in the same system, e.g. an outer gas giant that itself does not develop life~\citep{Schlecker2021a} or panspermia scenarios CITE.
+%We may then distinguish between:
+%\begin{enumerate}
+%    \item Earths, i.e., limited-water terrestrial planets in the liquid water habitable zone ($H1, H2$).
 These planets with roughly Earth-like water mass fractions support both the existence of submarine hydrothermal vents ($H1$) and hydrothermal fields with wet/dry cycles ($H2$).
 Limits in exoplanet observables to this planet type are their orbital distance (both scenarios require liquid water; at least for FGK stars this requirement also puts the planet outside of the tidal locking zone ($H2$)), and bulk density ($H1$ and $H2$ require limited water fractions).
-    \item ``shallow ocean'' water worlds ($H1$). These planets have no land surface exposed to the atmosphere, thus excluding the subaerial pond scenario.
-    \item ``deep ocean'' water worlds. Through the development of high-pressure ices, these planets do not support any of the considered OoL scenarios.
-\end{enumerate}
+%    \item ``shallow ocean'' water worlds ($H1$). These planets have no land surface exposed to the atmosphere, thus excluding the subaerial pond scenario.
+%    \item ``deep ocean'' water worlds. Through the development of high-pressure ices, these planets do not support any of the considered OoL scenarios.
+%\end{enumerate}
+%
+%
+%
 

### 2023-08-16 src/scripts/bayes_rho-a.py

@@ -85,8 +85,13 @@ def plot_factorgrid_dual(aa, rrho, P1, P2, title=None):
     """Plot evaluated environment factors on a grid in rho, a."""
     RRHO, AA = np.meshgrid(rrho, aa)
     fig, ax = plt.subplots()
+
+    # dummy plot for colorbar
+    dummy_cs = ax.pcolormesh(AA, RRHO, P2, cmap="gray_r", vmin=0.0, vmax=1.0)
+    dummy_cs.set_visible(False)
+
     ax.contourf(AA, RRHO, P1, cmap="Blues", alpha=0.5, vmin=0.3)
-    ax.contourf(AA, RRHO, P2, cmap="Oranges", alpha=0.5, vmin=0.3)
+    csf2 = ax.contourf(AA, RRHO, P2, cmap="Oranges", alpha=0.5, vmin=0.3)
     cs1 = ax.contour(AA, RRHO, P1, cmap="Blues", vmax=1.0)
     cs2 = ax.contour(AA, RRHO, P2, cmap="Oranges", vmax=1.0)
     h1, _ = cs1.legend_elements()
@@ -96,7 +101,7 @@ def plot_factorgrid_dual(aa, rrho, P1, P2, title=None):
     ax.set_xlabel("semi-major axis [au]")
     ax.set_ylabel(r"bulk density [$\rho_\oplus$]")
     ax.set_title(title)
-    fig.colorbar(cs2, label="$P_\mathrm{env}$")
+    fig.colorbar(dummy_cs, label="$P_\mathrm{env}$")
     return fig, ax
 
 

### 2023-08-17 src/tex/bayesian_analysis.tex
outsource bayesian analysis

--- 2023-08-16 src/tex/ms.tex
@@
-% Define document class
-\documentclass[twocolumn]{aastex631}
-\input{preamble}
-\begin{document}
-
-% Title
-\title{Bioverse: Origins of Life}
-
-% Author list
-\author{Martin Schlecker}
-%\author[0000-0001-8355-2107]{Martin Schlecker}
-\affiliation{Steward Observatory, The University of Arizona, Tucson, AZ 85721, USA; \href{mailto:schlecker@arizona.edu}{schlecker@arizona.edu}}
-\author{et al.}
-
-\begin{abstract}
-    tbd.
-\end{abstract}
-
-\section{Introduction}
-\label{sec:intro}
-
-\section{Origins of Life Scenarios and their predictions}
-\label{sec:ool_scenarios}
-In this section, we present some of the most prominent origins of life scenarios and their observational predictions.
-We focus on the necessary environmental conditions for the processes and reactions inherent to each scenario, and aim to identify distinct observables that are accessible via present and near-future remote sensing techniques.
-
-\subsection{Hydrothermal vents}
-e.g., Russell+2010
-...
-The hydrothermal vents scenario requires a direct contact of an ocean and the planetary mantle/crust.
-This requirement is not met on an ocean world with large amounts of water, where the water pressure on the ocean floor is high enough to form high-pressure ices (Noack+2016).
-\todo[inline]{see discussion in Kite \& Ford 2018 Sect. 6.4}
-
-\subsection{Subaerial ponds}
-...
-By its nature, the subaerial ponds scenario relies on rock surfaces exposed to the planetary atmosphere.
-Water worlds that have their entire planetary surface covered with water contradict this requirement and do not allow for the wet-dry cycling inherent to this origin of life scenario.
-The competition of tectonic stress with gravitational crustal spreading (Melosh 2011) sets the maximum possible height of mountains, which in the solar system does not exceed $\sim \SI{20}{\kilo\meter}$.
-Such mountains will be permanently under water on water worlds.
-Another impediment to wet-dry cycles is tidal locking of the planet as it stalls stellar tide-induced water movement and diurnal irradiation variability.
-
-\subsection{Cyanosulfidic scenario}
-...
-A strong requirement for the cyanosulfidic scenario is a sufficient Ultraviolet (UV) flux incident on the planet.
-On planets that have never received significant UV fluxes, the relevant photochemistry is limited.
-
-\subsection{Other Processes related to the Origins of Life}
-\subsubsection{Planetary redox state and evolution}
-The synthesis of prebiotic compounds requires moderately to highly reduced chemical environments (Kitadai \& Maruyama 2018, Benner+2020, Sasselov+2020, Lichtenberg \& Clement 2022).
-...
-Surficial origins of life chemistries are dependent on the redox state of a planet being $\sim$neutral (not too reduced or oxidized) to allow the presence of precursor molecules like HCN.
-The planetary redox state leaves an imprint on its atmospheric composition and thus planet size (very reduced atmospheres are large) and spectral signatures.
-Connected to the cyanosulfidic scenario, the pond scenario, and the impact trigger.
-
-\subsubsection{Impact trigger}
-Iron-rich impactors have been suggested to intermittently provide the reduced environments favored by prebiotic chemistry (e.g., Sekine+2003, Hashimoto+2007, Kuwahara \& Sugita 2015, Genda+2017, Wogan+2023).
-...
-Prebiotic synthesis triggered through reduced impactors that stochastically create transiently reducing or neutral atmospheres requires a certain composition of the impactors, the planet to not be in a magma ocean state (???) (Lichtenberg \& Clement 2022), and, related to this requirement, occurrence of impact events during early planetary evolution.
-Suggested observables are stochastic increases in brightness temperature, transient increases of planet size, and change of planet composition (decreasing with decreasing impact rate, i.e., stellar age).
-
 \section{Bayesian Analysis}
-\todo[inline]{consider Fisher matrix (Kendall \& Stuart 1977; Tegmark 1997) analysis to determine sensitivity of a survey to a set of parameters?}
+%\todo[inline]{consider Fisher matrix (Kendall \& Stuart 1977; Tegmark 1997) analysis to determine sensitivity of a survey to a set of parameters?}
 It is instructive to consider the constraining power of a successful biosignature detection for competing OoL scenarios, which we here attempt with an analytical approach.
@@
 Only a small region allows for a significant model selection: While short orbits ($a \lessapprox \SI{0.03}{\au}$) and low bulk densities ($\rho \lessapprox \SI{0.6}{\rho_\oplus}$) strongly support $H_1$, no combination of bulk density and orbital distance provide strong evidence for $H_2$ over $H_1$ without additional information.
-\todo[inline]{factor in Detection probability P_det}
-\todo[inline]{TODO: test sensitivity of this result on the assumed function and thresholds for P_env}
+ \todo[inline]{factor in Detection probability P\_det}
+ \todo[inline]{TODO: test sensitivity of this result on the assumed function and thresholds for P\_env}
 
@@
 \todo[inline]{discuss host star spectral type dependencies, e.g., abiogenesis time window for G stars ($\lessapprox 10 Gyr$, MS lifetime) or M dwarfs. \citep[e.g.,][]{Spiegel2012}}
-
-\section{Exoplanet survey simulations}
-\todo[inline]{Focus on molecular Oxygen as a biomarker(?)}
-A commonly discussed biomarker is molecular Oxygen (O$_2$), which on Earth emerged as a byproduct of photosynthesis during the Proterozoic era.
-While not the only atmospheric species discussed as a signature for planetary life CITE, we focused on O$_2$ as a representative biomarker for our spectroscopic survey simulations.
-
-\subsection{Habitable Worlds Observatory}
-\subsubsection{target list}
-Provisional stellar target List for the habitable Worlds Observatory~\citep{Mamajek2023}.
-
-\subsection{Nautilus}
-\subsection{LIFE} % optional
-
-\section{Hypothesis tests}
-\label{sec:hypotests}
-\subsection{Possibility of hydrothermal vents}
-
-\subsection{Availability of rock-air interfaces}
-
-\subsection{Planetary redox state evolution}
-
-\subsection{Planetary UV flux and its evolution}
-
-\section{Results}
-\label{sec:results}
-
-\section{Discussion and Conclusions}
-\label{sec:discussion}
-
-''Which natural processes best explain how living matter spontaneously appears from nonliving matter?''~\citep{Malaterre2022}
-%WE consider life ``as we know it''
-
-\bibliography{bib}
-
-\end{document}

### 2023-08-17 src/tex/ms.tex
outsource bayesian analysis

--- 2023-08-12 src/tex/ms.tex
@@
 
-\section{Bayesian Analysis}
-\todo[inline]{consider Fisher matrix (Kendall \& Stuart 1977; Tegmark 1997) analysis to determine sensitivity of a survey to a set of parameters?}
-It is instructive to consider the constraining power of a successful biosignature detection for competing OoL scenarios, which we here attempt with an analytical approach.
-The hydrothermal vents scenario ($H_1$) and the subaereal pond scenario ($H_2$) can be considered as mutually exclusive models, and we may study how a particular future observation of biosignatures impacts our beliefs about the relative model probabilities.
-
-We may first consider the probability $\pdf(bio|H_i)$ of detecting a convincing biosignature on a planet, given a particular OoL hypothesis $H_i, \, i \in 1, 2$ is true.
-This can be decomposed into the probability of abiogenesis in a particular environment $\pdf_\mathrm{env, i}$, the fraction of life-hosting worlds that develop atmospheric biosignatures $\pdf_\mathrm{sig}$, the probability of the planet type required by the OoL hypothesis occurring in the surveyed sample $\pdf_\mathrm{\eta, i}$, and the probability of detecting the biosignature on this planet type with current technology $\pdf_\mathrm{det, i}$, yielding
-
-\begin{align}
-\label{eqn:pbio}
-\pdf(bio|H_i) = \pdf_\mathrm{env, i} \times \pdf_\mathrm{sig} \times \pdf_\mathrm{\eta, i} \times \pdf_\mathrm{det, i}.
-\end{align}
-
-However, what we are actually interested in is the probability of a OoL hypothesis being true, given a particular biosignature detection, $\pdf(H_i|bio)$.
-To obtain this we can use Bayes' theorem, which yields
-\begin{align}
-\pdf(H_i|bio) = \frac{\pdf(bio|H_i) \pdf(H_i)}{\pdf(bio)}.
-\end{align}
-Here, $\pdf(H_i)$ is the prior probability of the OoL hypothesis $H_i$, and $\pdf(bio)$ is the prior probability of detecting a biosignature.
-% Following Eq.~\ref{eqn:pbio}, we can sum over all possible scenarios to obtain
-% \begin{align}
-% \pdf(bio) = \sum_{i=1}^{N} \pdf_\mathrm{env, i} \times \pdf_\mathrm{sig} \times \pdf_\mathrm{\eta, i} \times \pdf_\mathrm{det, i}.
-% \end{align}
-
-If the hypotheses $H_i$ are adjunct, i.e., their joint occurrence is impossible, one can show that
-\begin{align}
-\pdf(H_i|bio) = \frac{\pdf(bio|H_i) \pdf(H_i)}{\sum_{i=1}^{N} \pdf(bio|H_i) \pdf(H_i)}.
-\end{align}
-Then the parameters in Eq.~\ref{eqn:pbio} that are independent of the chosen hypothesis $H_i$ eliminate and we obtain
-\begin{align}
-\pdf(H_i|bio) = &\frac{\pdf_\mathrm{env, i} \pdf_\mathrm{\eta, i} \pdf_\mathrm{det, i} \pdf(H_i)}{\sum_{i=1}^{N} \pdf_\mathrm{env, i} \pdf_\mathrm{\eta, i} \pdf_\mathrm{det, i} \pdf(H_i)} \\
-\overset{\pdf(H_i) = \pdf(H_j)}{ \underset{\forall i,j \in \{1, 2\}}{=}} &\frac{\pdf_\mathrm{env, i} \pdf_\mathrm{\eta, i} \pdf_\mathrm{det, i}}{\sum_{i=1}^{N} \pdf_\mathrm{env, i} \pdf_\mathrm{\eta, i} \pdf_\mathrm{det, i}},
-\end{align}
-where in the last step we made the implicit assumption that all OoL hypotheses are a priori equally probable.
-
-If we take the ratio of these posteriors for our two independent hypotheses $H_1$ and $H_2$, we get the \textit{Bayes Factor}
-\begin{align}
-    \frac{\pdf(H_1|bio)}{\pdf(H_2|bio)} = \frac{\pdf_\mathrm{env, 1} \pdf_\mathrm{\eta, 1} \pdf_\mathrm{det, 1}}{\pdf_\mathrm{env, 2} \pdf_\mathrm{\eta, 2} \pdf_\mathrm{det, 2}},
-\end{align}
-which quantifies the evidence of the data arising from $H_1$ versus $H_2$.
-
-% Putting everything together, we arrive at
-% \begin{align}
-% \label{eqn:posterior}
-% \pdf(H_i|bio) = \frac{\pdf(H_i)}{\pdf(bio)} \times \pdf_\mathrm{env, i} \times \pdf_\mathrm{sig} \times \pdf_\mathrm{\eta, i} \times \pdf_\mathrm{det, i}.
-% \end{align}
-% If we assume that all OoL hypotheses are a priori equally probable, we can treat $\frac{\pdf(H_i)}{\pdf(bio)}$ in Eq.~\ref{eqn:posterior} as a normalization constant.
-In the following, we discuss the impact of the remaining variables $\pdf_\mathrm{env, i}$, $\pdf_\mathrm{\eta, i}$ and $\pdf_\mathrm{det, i}$ on the Bayes factor.
-
- \subsection{Required environment $\pdf_\mathrm{env, i}$}
-% Following previous work~\citep{Spiegel2012,Chen2018,Kipping2021}, we may adopt a uniform rate model for abiogenesis, i.e., assume that OoL events occur at a uniform rate.
-% This corresponds to a Poisson process with a rate parameter $lambda$, where we make the implicit assumption that abiogenesis occurs only via a single, unique mechanism, only once, and instantaneous.
-% If this event occurs within a limited time window $t$, say, between planets form around a star and when it leaves the main sequence, we have
-% \begin{align}
-% \pdf_\mathrm{env, i} = 1 - \exp(-\lambda t).
-% \end{align}
-% ~\\
-\todo[inline]{The planet should be in the liquid-water HZ}
-...
-Let us assume that $H_1$ only requires a minimum bulk density $\rho_1$, such that $\pdf_\mathrm{env, 1} = 1$ if $\rho > \rho_1$ and zero otherwise.
-On the other hand, $H_2$ requires exposed land and thus a small water mass fraction.
-We may implement this in the same way as above but with a minimum bulk density $\rho_2 > \rho_1$.
-Furthermore, there is a requirement that the tidal locking timescale may not be so small that the planet is likely tidally locked at observation.
-This translates to imposing a minimum semimajor axis $a_2$.
-...
-...
-
-% \subsection{biosignature fraction $\pdf_\mathrm{env, i}$}
-% Limiting ourselves to the search for \textit{life-as-we-know-it}, we may assume that there are atmospheric biomarkers present on a planet after the abiogenesis event and until global extinction.
-
-\subsection{Fractional planet occurrence rate $\pdf_\mathrm{\eta, i}$}
-As the different OoL hypotheses do not all work on the same planet types, we may study the impact of the fractional occurrence rates of different planet types on the posterior probability $\pdf(H_i|bio)$.
-For simplicity, let's consider only planets that allow at least one of the scenarios.
-We also ignore any influence of other planets in the same system, e.g. an outer gas giant that itself does not develop life~\citep{Schlecker2021a} or panspermia scenarios CITE.
-We may then distinguish between:
-\begin{enumerate}
-    \item Earths, i.e., limited-water terrestrial planets in the liquid water habitable zone ($H1, H2$).
-These planets with roughly Earth-like water mass fractions support both the existence of submarine hydrothermal vents ($H1$) and hydrothermal fields with wet/dry cycles ($H2$).
-Limits in exoplanet observables to this planet type are their orbital distance (both scenarios require liquid water; at least for FGK stars this requirement also puts the planet outside of the tidal locking zone ($H2$)), and bulk density ($H1$ and $H2$ require limited water fractions).
-    \item ``shallow ocean'' water worlds ($H1$). These planets have no land surface exposed to the atmosphere, thus excluding the subaerial pond scenario.
-    \item ``deep ocean'' water worlds. Through the development of high-pressure ices, these planets do not support any of the considered OoL scenarios.
-\end{enumerate}
-
-\subsection{Detection probability $\pdf_\mathrm{det, i}$}
-
-\subsection{Discussion of Bayesian analysis}
-
-\begin{note}
-    shallow ocean planets are vulnerable to water loss through high-energy radiation, limiting the time window for habitability especially if no geochemical feedbacks exist~\citep{Kite2018}.
-\end{note}
-\todo[inline]{discuss host star spectral type dependencies, e.g., abiogenesis time window for G stars ($\lessapprox 10 Gyr$, MS lifetime) or M dwarfs. \citep[e.g.,][]{Spiegel2012}}
+\include{bayesian_analysis}
 

### 2023-08-25 src/tex/ms.tex
add schematic figure with OOL scenarios
structure paper outline

@@
 \label{sec:intro}
+\todo[inline]{Introduce OOL, the importance of planetary contexts}
 
 \section{Origins of Life Scenarios and their predictions}
+\todo[inline]{Present widely discussed OOL scenarios and their predictions on exoplanet observables; derive testable hypotheses.}
 \label{sec:ool_scenarios}
@@
 
-\include{bayesian_analysis}
+%\include{bayesian_analysis}
+
+\subsubsection{Origins of Life Hypotheses}
+...
+\begin{figure*}
+    \begin{centering}
+        \includegraphics[width=\hsize]{figures/hypotheses.pdf}
+        \caption{Origins of Life scenarios, their predictions on exoplanet observables, and derived population-level hypotheses.}
+        \label{fig:hypotheses}
+    \end{centering}
+\end{figure*}
+Figure~\ref{fig:hypotheses} lists the hypotheses derived from the predictions of each OOL scenario.
+...
 
@@
 \subsubsection{target list}
-Provisional stellar target List for the habitable Worlds Observatory~\citep{Mamajek2023}.
+\todo[inline]{Provisional stellar target List for the habitable Worlds Observatory:~\citep{Mamajek2023}}
 
@@
 \label{sec:hypotests}
-\subsection{Possibility of hydrothermal vents}
+\todo[inline]{Briefly introduce Bayesian model comparison, then present the particular hypotheses in their mathematical form.}
+\subsection{H1: Life only originates in hydrothermal vents}
 
-\subsection{Availability of rock-air interfaces}
+\subsection{H2: Life only originates in subaerial ponds with wet/dry cycles}
 
-\subsection{Planetary redox state evolution}
-
-\subsection{Planetary UV flux and its evolution}
+\subsection{H3: Life only originates on planets with particular UV irradiance}
 
@@
 \label{sec:results}
+\subsection{Information content in mass-density space}
+
+\subsection{Information content in tidal locking timescale-density space}
+
+\subsection{Correlation of past UV flux and biosignature occurrence}
 
@@
 
-''Which natural processes best explain how living matter spontaneously appears from nonliving matter?''~\citep{Malaterre2022}
+%''Which natural processes best explain how living matter spontaneously appears from nonliving matter?''~\citep{Malaterre2022}
 %WE consider life ``as we know it''

### 2023-08-30 src/tex/ms.tex
Provide a .bib file for co-authors
explicitly mention predictions
develop structure

@@
 % Define document class
-\documentclass[twocolumn]{aastex631}
+%\documentclass[twocolumn,twocolappendix,linenumbers]{aastex631}
+\documentclass[modern,linenumbers]{aastex631}
 \input{preamble}
@@
 
-\section{Origins of Life Scenarios and their predictions}
+\section{Origins of Life Scenarios and their Predictions}
 \todo[inline]{Present widely discussed OOL scenarios and their predictions on exoplanet observables; derive testable hypotheses.}
@@
 
+\textbf{Prediction:} Planets with high-pressure ices do not show biosignatures.
+
 \subsection{Subaerial ponds}
@@
 
+\textbf{Prediction:} Biosignatures occur outside the tidal locking zone and at bulk densities consistent with exposed rock..
+
 \subsection{Cyanosulfidic scenario}
@@
 On planets that have never received significant UV fluxes, the relevant photochemistry is limited.
+
+\textbf{Prediction:} Past UV flux and the occurrence of biosignatures are correlated.
 
@@
 
+\section{Hypothesis Tests}
+\label{sec:hypotests}
+\todo[inline]{Briefly introduce Bayesian model comparison, then present the particular hypotheses in their mathematical form.}
+
+\subsection{Bayesian hypothesis testing with \bioverse}
+
+\subsection{H1: Life only originates in hydrothermal vents}
+The hydrothermal vent scenario does not allow oceans deep enough to form an impenetrable layer of high-pressure ice on its floor.
+The resulting allowed parameter space is described by a lower limit on the bulk density. \todo[inline]{+ exclude atmospheric signature for water worlds?}
+...
+
+\subsection{H2: Life only originates in subaerial ponds with wet/dry cycles}
+We parametrize the required exposed land surface in this scenario as a lower limit in bulk density that is higher than for H1.
+Further, the tidal locking timescale of the planet may not be smaller than the age of the system.
+...
+
+\subsection{H3: Life only originates on planets with particular UV irradiance}
+We test a UV irradiance requirement by relating the occurrence of life on a planet with a minimum past UV flux, which is a function of ...? \todo[inline]{SUKRIT}
+
+\todo[inline]{OPTIONAL: ``We further test the scenario of a linear correlation of past UV flux and biosignature occurrence rate.
+This test requires the detection of multiple biosignatures.''
+~\\
+Test for negative correlations as well?}
+
 \section{Exoplanet survey simulations}
@@
 
-\section{Hypothesis tests}
-\label{sec:hypotests}
-\todo[inline]{Briefly introduce Bayesian model comparison, then present the particular hypotheses in their mathematical form.}
-\subsection{H1: Life only originates in hydrothermal vents}
-
-\subsection{H2: Life only originates in subaerial ponds with wet/dry cycles}
-
-\subsection{H3: Life only originates on planets with particular UV irradiance}
-
 \section{Results}
@@
 \label{sec:discussion}
-
 %''Which natural processes best explain how living matter spontaneously appears from nonliving matter?''~\citep{Malaterre2022}
@@
 
-\bibliography{bib}
+\subsection{What do we learn from a single biosignature detection?}
+\todo[inline]{discuss constraining power on OOL of a convincing biosignature detection on a single planet, depending on the position of the planet in the parameter space we explored.}
+
+\subsection{Constraining power for the origins of life as a function of biosignature location}
+\todo[inline]{How does the location of biosignature detections impact the credibility of OOL scenarios?}
+\todo[inline]{biosignature on M~dwarf planet vs. FGK: Impact on UV flux requirement?}
+
+\subsection{Conclusions and future work}
+
+\bibliography{bib,coauthors}
 

### 2023-09-06 src/tex/ms.tex
SR comments
.gitignore, outsource acknowledgments & contributions

@@
 This requirement is not met on an ocean world with large amounts of water, where the water pressure on the ocean floor is high enough to form high-pressure ices (Noack+2016).
+% TODO Sleep et al., 2011?
+Sobotta et al., 2020?
 \todo[inline]{see discussion in Kite \& Ford 2018 Sect. 6.4}
@@
 \section{Exoplanet survey simulations}
-\todo[inline]{Focus on molecular Oxygen as a biomarker(?)}
+\todo[inline]{biomarkers: Focus on molecular Oxygen? Excess Methane? \citep[e.g.,][]{Seeburger2023}}
 A commonly discussed biomarker is molecular Oxygen (O$_2$), which on Earth emerged as a byproduct of photosynthesis during the Proterozoic era.
@@
 
+%\input{acknowledgments_contributions}
+
+\section*{Reproducibility}
+
 \bibliography{bib,coauthors}

### 2023-09-06 src/tex/ms.tex
mention caveats on atm. transmission and quiescent flux vs. flares

@@
 
+\subsection{Caveats}
+\subsubsection{Atmosphere transmission}
+Theoretical work suggests that the atmosphere of prebiotic Earth was largely transparent at near-UV wavelengths with the only known source of attenuation being Rayleigh scattering~\citep{Ranjan2017,Ranjan2017c}.
+We thus approximated surface UV flux using top-of-atmosphere fluxes.
+This represents a conservative approach, since any planet that fails to meet the irradiance criterion receives even lower near-UV radiation at its surface.
+\todo[inline]{SUKRIT please specify as neccessary}
+
+\subsubsection{Stellar flares}
+Our assumptions on past UV flux neglect the contribution of stellar flares, which may be hypothesized as an alternative source of UV light~\citep{Ranjan2017}.
+This concerns mainly ultracool dwarfs, due to their low quiescent emission and high pre-main sequence stellar activity~\citep{Buccino2007,West2008}.
+Recent work indicates that the majority of stars show inadequate activity levels for a sufficient contribution through flares~\citep{Glazier2020,Ducrot2020,Günther2020}.
+The biosignature surveys we simulated here may test the hypothesis of sufficient UV radiation via stellar flares.
+
 \subsection{Conclusions and future work}

### 2023-09-07 src/tex/ms.tex
Define the habitable zone

@@
 
+\subsection{Habitable Zone}
+All Origins of Life scenarios considered here require water as a solvent; we thus consider only planets that may sustain liquid water on their surface.
+Different formulations of habitable zones as regions around a star where a planet with Earth's atmospheric composition can maintain liquid water on its surface exist~\citep[e.g.,][]{MolLous2022,Spinelli2023,Tuchow2023}.
+CITE!
+Ramirez \& Kaltenegger 2017, 2018
+Here, we adopt the popular estimates of \citet{Kasting1993} and \citet{Kopparapu2013,Kopparapu2014} that define a temperate zone between the runaway greenhouse transition CITE!
+and the maximum greenhouse limit CITE!.
+We use the parametrization in \citet{Kopparapu2014} to derive luminosity and planetary mass-dependent edges of the habitable zone $a_\mathrm{inner}$ and $a_\mathrm{outer}$.
+
 \subsection{Bayesian hypothesis testing with \bioverse}

### 2023-09-09 src/tex/ms.tex
add comments on UV flux and cyanosulfidic scenario

@@
 \todo[inline]{see discussion in Kite \& Ford 2018 Sect. 6.4}
+\todo[inline]{SR: The sealing away of the planetary interior from the ocean due to high-pressure ice layers is a common assumption for water world exoplanets (in addition to the references above, see e.g.
+Hu et al.
+2021).
+I'm not convinced it is correct, because of relatively recent evidence showing the possibility of molecular assimitation into such ices and subsequent transport, e.g., https://iopscience.iop.org/article/10.1088/0004-637X/769/1/29/meta, https://iopscience.iop.org/article/10.3847/1538-4357/acb49a/meta, https://iopscience.iop.org/article/10.3847/1538-4357/aa5cfe/meta (I'm sure there are other workers in this area, this is just the group with which I am familiar).
+Exoplaneteers mostly uniformly accept this proposition, so it's not an unreasonable assumption if you want to run with it so long as its acknowledged and caveated reasonably; I'm just highlighting this for your attention so that you can make an informed decision.
+}
 
@@
 \textbf{Prediction:} Biosignatures occur outside the tidal locking zone and at bulk densities consistent with exposed rock..
+\todo[inline]{SR: I'm not a priori sold that tidal locking means that no wet-dry cycles occur.
+you can still have cycling driven by transient changes in instellation due to flares, for example (e.g., https://iopscience.iop.org/article/10.3847/1538-4357/aadfd1/meta).
+Similarly, I wonder if 3D effects might not give rise to variability (https://iopscience.iop.org/article/10.3847/PSJ/acc9c4/meta).
+I argue that it is more robust to establish a correlation between biosignatures and planets which show evidence of continents/land.
+I think that Ty Robinson in our department has done some work in this area, his papers might be a good starting point.
+Other papers which look relevant (but with which I am not familiar, as this is not my area): https://academic.oup.com/mnras/article/511/1/440/6501216, https://academic.oup.com/mnras/article/495/1/1/5733176, https://iopscience.iop.org/article/10.3847/1538-3881/aad775/meta, https://iopscience.iop.org/article/10.3847/1538-3881/aaed3a/meta, https://iopscience.iop.org/article/10.3847/1538-3881/ab2df3/meta}
 
-\subsection{Cyanosulfidic scenario}
-...
-A strong requirement for the cyanosulfidic scenario is a sufficient Ultraviolet (UV) flux incident on the planet.
-On planets that have never received significant UV fluxes, the relevant photochemistry is limited.
+\subsection{UV flux}\label{sec:predictions:uv}
+A major hypothesis in the origin of life is that UV light played a constructive role in getting life started on Earth (see Ranjan et al.
+2016, 2017c; Rimmer et al.
+2018; Rapf \& Vaida 2016; Pascal et al.
+2012; Green et al.
+2021; and sources therein).
+
+If UV light is required to get life started, then there is a minimum planetary UV flux requirement to have an inhabited world.
+This requirement is set by competitor thermal processes; if the photo-reaction does not move forward at a rate faster than the competitor thermal process(es), then the abiogenesis scenario cannot function.
+On the other hand, abundant UV light vastly in excess of this threshold does not increase the probability of abiogenesis, since once the UV photochemistry is no longer limiting, some other thermal process in the reaction network will be rate-limiting process instead.
+Therefore, a putative dependence of life on UV light is best encoded as a step function (see, e.g., Ranjan et al. 2017c; Rimmer et al. 2018; Rimmer, Ranjan \& Rugheimer 2021).
+
+One origin-of-life scenario has been refined to the point where the threshold flux has been measured.
+The cyanosulfidic scenario has been shown to require a mean flux of at least $ F_\mathrm{NUV, min} = \SI{6.8\pm3.6 e10}{\photons\per\centi\meter\squared\per\second\per\nano\meter}$ integrated from \SIrange{200}{280}{\nano\meter} at the surface in order to function (Rimmer et al.
+2018; Rimmer et al.
+2021 Astrobiology; Rimmer et al.
+2023).
+\todo[inline]{SR: This is an interesting number, because it is below what was available on early Earth (so this scenario could have worked on early Earth) but until recently it was below what was thought to be available on habitable zone M-dwarf exoplanets.
+So it was thought that identification of biosignatures on M-dwarf planets could therefore falsify the cyanosulfidic scenario, with a potential caveat for transient UV from flares.
+Two recent developments have complicated the picture.
+First, Rimmer et al.
+2018 had an error in their radiative transfer routines.
+Correcting for this error, early M-dwarfs and highly active M-dwarfs emit enough UV to meet the Rimmer et al.
+2018 criterion (Ranjan et al.
+2023).
+Second, a recent publication argues that /all/ estimates of M-dwarf UV are underestimates, and that late M-dwarf stars have similar emission to Sunlike stars (Rekhi et al.
+2023).
+I suspect this is incorrect, because it contradicts a lot of work from e.g.
+the MUSCLES collaboration and the HAZMAT project, but it's worth keeping an eye on in case it is correct after all.}
+We use this threshold value as our baseline case.
 
@@
 
-%\include{bayesian_analysis}
+%\input{bayesian_analysis}
 
-\subsubsection{Origins of Life Hypotheses}
+\subsection{Origins of Life Hypotheses}
 ...

### 2023-09-09 src/tex/ms.tex
add planet radius requirement

@@
 
-\subsection{Habitable Zone}
-All Origins of Life scenarios considered here require water as a solvent; we thus consider only planets that may sustain liquid water on their surface.
+\subsection{Considered planets and their orbits}
+All Origins of Life scenarios considered here require water as a solvent; we thus consider only rocky planets that may sustain liquid water on their surface.
 Different formulations of habitable zones as regions around a star where a planet with Earth's atmospheric composition can maintain liquid water on its surface exist~\citep[e.g.,][]{MolLous2022,Spinelli2023,Tuchow2023}.
@@
 We use the parametrization in \citet{Kopparapu2014} to derive luminosity and planetary mass-dependent edges of the habitable zone $a_\mathrm{inner}$ and $a_\mathrm{outer}$.
+We further follow \citet{Bixel2021} and classify as ``exo-Earth candidates'' all planets with radii $0.8 S^{0.25} < R < 1.4 $ that orbit within the above boundaries.
+The lower limit was suggested as a minimum planet size to retain an atmosphere~\citep{Zahnle2017}.
 

### 2023-09-09 src/tex/ms.tex
formulate hypothesis H3

@@
 
+\subsection{Fraction of inhabited planets with detectable biosignatures}
+Presumably, not all habitable worlds are inhabited and not all inhabited worlds develop detectable biosignatures.
+The fraction of exo-Earth candidates that are both inhabited and harbor detectable biomarkers at the time when we observe them remains speculative; we aggregate them in the unitless parameter $f_\mathrm{life}$.
+
 \subsection{H1: Life only originates in hydrothermal vents}
@@
 \subsection{H3: Life only originates on planets with particular UV irradiance}
-We test a UV irradiance requirement by relating the occurrence of life on a planet with a minimum past UV flux, which is a function of ...? \todo[inline]{SUKRIT}
+Our third theoretical experiment is to test the UV irradiance requirement (Sect.~\ref{sec:predictions:uv}) by relating the occurrence of life on a planet with a minimum past UV flux.
+Here, we focus on the prebiotically interesting \SIrange{200}{280}{\nano\meter} range CITE and the quiescent stellar flux.  %~\citep{RanjanSasselov2016}.
+Our concrete hypothesis shall be that life only occurs on planets that at some point in their history have received UV radiation in this wavelength range that exceeded a minimum NUV flux $F_\mathrm{NUV, min}$.
+We further require this flux to have lasted for $\geq \SI{100}{\mega\year}$ to allow for a sufficient ``origins timescale''~\citep{Rimmer2023}.
+We then have
+\begin{equation}\label{eq:hypothesis-uv}
+    H_{3} = f_\mathrm{life} (\theta, F_\mathrm{NUV}) =
+        \begin{cases}
+            0 , & F_\mathrm{NUV} < F_\mathrm{NUV, min}\\
+            f_\mathrm{life}, & F_\mathrm{NUV} \geq F_\mathrm{NUV, min} \, \mathrm{for} \, \Delta t \geq \SI{100}{\mega\year}
+        \end{cases}
+\end{equation}
+and the corresponding null hypothesis
+%\begin{equation}
+$H_{3, null} = f_\mathrm{life} (\theta),$
+%\end{equation}
+i.e., no correlation with UV flux.
+
+\subsubsection{past UV flux}
+\todo[inline]{\citet{Günther2020} relate U-band energy to bolometric flux.}
 
@@
 This concerns mainly ultracool dwarfs, due to their low quiescent emission and high pre-main sequence stellar activity~\citep{Buccino2007,West2008}.
-Recent work indicates that the majority of stars show inadequate activity levels for a sufficient contribution through flares~\citep{Glazier2020,Ducrot2020,Günther2020}.
+Recent work indicates that the majority of stars show inadequate activity levels for a sufficient contribution through flares~\citep{Glazier2020,Ducrot2020,Guenther2020}.
 The biosignature surveys we simulated here may test the hypothesis of sufficient UV radiation via stellar flares.

### 2023-09-16 src/tex/ms.tex
distinguish between injected signal and hypothesis

@@
 We further require this flux to have lasted for $\geq \SI{100}{\mega\year}$ to allow for a sufficient ``origins timescale''~\citep{Rimmer2023}.
-We then have
+We assign the development of life to a fraction $f_\mathrm{life}$ of all temperate planets fulfilling these requirements.
+For a population-level hypothesis, we then have
 \begin{equation}\label{eq:hypothesis-uv}
@@
 \todo[inline]{biomarkers: Focus on molecular Oxygen? Excess Methane? \citep[e.g.,][]{Seeburger2023}}
+% detecting methanogenic life: Sandora2020, Wang2018,
 A commonly discussed biomarker is molecular Oxygen (O$_2$), which on Earth emerged as a byproduct of photosynthesis during the Proterozoic era.

### 2023-10-03 src/tex/ms.tex
add a note
describe modeling of H3 (NUV correlation)

@@
 \subsection{H3: Life only originates on planets with particular UV irradiance}
-Our third theoretical experiment is to test the UV irradiance requirement (Sect.~\ref{sec:predictions:uv}) by relating the occurrence of life on a planet with a minimum past UV flux.
-Here, we focus on the prebiotically interesting \SIrange{200}{280}{\nano\meter} range CITE and the quiescent stellar flux.  %~\citep{RanjanSasselov2016}.
-Our concrete hypothesis shall be that life only occurs on planets that at some point in their history have received UV radiation in this wavelength range that exceeded a minimum NUV flux $F_\mathrm{NUV, min}$.
-We further require this flux to have lasted for $\geq \SI{100}{\mega\year}$ to allow for a sufficient ``origins timescale''~\citep{Rimmer2023}.
-We assign the development of life to a fraction $f_\mathrm{life}$ of all temperate planets fulfilling these requirements.
+Our third theoretical experiment is to test the UV irradiance requirement (Sect.~\ref{sec:predictions:uv}) by relating the occurrence of life on an exo-earth candidate with a minimum past quiescent stellar UV flux, focusing on the prebiotically interesting NUV range from \SIrange{200}{280}{\nano\meter}.
+\todo[inline]{SUKRIT: got a good reference for this?}  %~\citep{RanjanSasselov2016}.?
+Our concrete hypothesis shall be that life only occurs on planets that at some point in their history have received such radiation exceeded a minimum flux $F_\mathrm{NUV, min}$.
+We further require this flux to have lasted for a minimum duration $\Delta T_\mathrm{min}$ to allow for a sufficient ``origins timescale''~\citep{Rimmer2023}, and the planet to be within its momentary habitable zone during this entire period.
+
+For each exo-earth candidate, we interpolated the age- and stellar mass-dependent habitable zone fluxes obtained by \citet{Richey-Yowell2023} and computed the evolution of the inner (runaway greenhouse) and outer (maximum greenhouse) edges as a function of planet mass, stellar spectral type using the results of \citet{Kopparapu2014} and the stellar luminosity evolution of \citet{Baraffe1998}.
+We then determined which planets were both in the HZ and had NUV fluxes above $F_\mathrm{NUV, min}$ for $\Delta T_\mathrm{min} \geq \SI{100}{\mega\year}$.
+We assigned the development of life to a fraction $f_\mathrm{life}$ of all temperate planets fulfilling these requirements.
 For a population-level hypothesis, we then have
@@
             0 , & F_\mathrm{NUV} < F_\mathrm{NUV, min}\\
-            f_\mathrm{life}, & F_\mathrm{NUV} \geq F_\mathrm{NUV, min} \, \mathrm{for} \, \Delta t \geq \SI{100}{\mega\year}
+            f_\mathrm{life}, & F_\mathrm{NUV} \geq F_\mathrm{NUV, min} \,\mathrm{and\, in\, HZ} \, \mathrm{for} \, \Delta t \geq \SI{100}{\mega\year}
         \end{cases}
@@
 \subsubsection{target list}
-\todo[inline]{Provisional stellar target List for the habitable Worlds Observatory:~\citep{Mamajek2023}}
+\todo[inline]{Provisional stellar target List for the habitable Worlds Observatory:~\citep{Mamajek2023}} % read like in https://github.com/mkenworthy/HWObows/tree/40b1e5f2bc2bbbe8c999b097c80ca5a8cda0aeae/src/scripts
 
@@
 
+\subsection{Contextual support for potential biosignature detections}
+\todo[inline]{discuss how planetary context may impact credibility of tentative biosignature detections (e.g., if there is a good fit with a predicted OOL pattern; or the opposite: the planetary context does not fit well to any OOL scenarios)}
+
 \subsection{Caveats}

### 2023-10-04 src/tex/ms.tex
Some structural comments

--- 2023-09-09 src/tex/ms.tex
@@
 We use this threshold value as our baseline case.
+\todo[inline]{TL: this scenario, the impact trigger, and the redox state are all interrelated and not easily to disentangle (UV scenarios depend on the right redox range at the surface, as do impact scenarios); however we can try to build this into the general debate.
+instead of having an ``other scnearios'' seciton i would suggest discussing/introducing the redox constraints in the introdcution together iwht planet formation and primordial differentiation processes.}
 
@@
 \subsection{LIFE} % optional
+\todo[inline]{TL: please not optional... :) happy to help support this section}
 
@@
 \subsection{Caveats}
+
+\subsubsection{Planetary composition}
+\todo[inline]{TL: the three scenarios are very specific in terms of planet composition (== redox state :-) and thermal state, we need to discuss this here }
+
 \subsubsection{Atmosphere transmission}

### 2023-11-01 src/tex/glossary.tex
add glossary for acronyms, own section for Conclusions

\newglossaryentry{HST}{name={HST}, description={Hubble Space Telescope}, first={Hubble Space Telescope (HST)}}
\newglossaryentry{NIR}{name={NIR}, description={Near-Infrared}, first={near-infrared (NIR)}}
\newglossaryentry{MIR}{name={MIR}, description={Mid-Infrared}, first={mid-infrared (MIR)}}
\newglossaryentry{FIR}{name={FIR}, description={Far-Infrared}, first={far-infrared (FIR)}}
\newglossaryentry{FUV}{name={FUV}, description={Far-Ultraviolet}, first={far-Ultraviolet (FUV)}}
\newglossaryentry{NUV}{name={NUV}, description={Near-Ultraviolet}, first={near-Ultraviolet (NUV)}}
\newglossaryentry{EUV}{name={EUV}, description={extreme-Ultraviolet}, first={extreme-Ultraviolet (EUV)}}
\newglossaryentry{SED}{name={SED}, description={Spectral energy distribution}, first={spectral energy distribution (SED)}, firstplural={spectral energy distributions (SEDs)}}
\newglossaryentry{ppm}{name={ppm}, description={parts per million}, first={parts per million (ppm)}}
\newglossaryentry{PDF}{name={PDF}, description={Probability density function}, first={probability density function (PDF)}, firstplural={probability density functions (PDF)}}
\newglossaryentry{PLATO}{name={PLATO}, description={PLAnetary Transits and Oscillations of stars}, first={PLATO~\citep[][]{Rauer2014}}}
\newglossaryentry{PSF}{name={PSF},description={Point Spread Function}, first={point spread function (PSF)}, firstplural={point spread functions (PSF)}}
\newglossaryentry{SFD}{name={SFD}, description={Size frequency distribution}, first={size frequency distribution (SFD)}, firstplural={Size frequency distributions (SFD)}}
\newglossaryentry{SNR}{name={SNR}, description={Signal-to-noise ratio}, first={signal-to-noise ratio (SNR)}}
\newglossaryentry{RV}{name={RV}, description={Radial Velocity}, first={Radial Velocity (RV)}, firstplural={Radial Velocities (RV)}}
\newglossaryentry{TTV}{name={TTV}, description={Transit Timing Variation}, first={transit timing variation (TTV)}, firstplural={transit timing variations (TTVs)}}
\newglossaryentry{TESS}{name={TESS}, description={Transiting Exoplanet Survey Satellite}, first={Transiting Exoplanet Survey Satellite~\citep[\textit{TESS},][]{Ricker2014}}}
\newglossaryentry{TIC}{name={TIC}, description={\textit{TESS} Input Catalog}, first={\textit{TESS} Input Catalog~\citep[TIC,][]{Stassun2018}}}
\newglossaryentry{JWST}{name={JWST}, description={James Webb Space Telescope}, first={James Webb Space Telescope~\citep[\textit{JWST},][]{Beichman2014}}}
\newglossaryentry{HZ}{name={HZ}, description={habitable zone}, first={habitable zone (HZ)}}
%\newglossaryentry{}{name={}, description={}, first={}, firstplural={}}

### 2023-11-01 src/tex/ms.tex
add glossary for acronyms, own section for Conclusions

--- 2023-10-03 src/tex/ms.tex
@@
 and the maximum greenhouse limit CITE!.
-We use the parametrization in \citet{Kopparapu2014} to derive luminosity and planetary mass-dependent edges of the habitable zone $a_\mathrm{inner}$ and $a_\mathrm{outer}$.
+We use the parametrization in \citet{Kopparapu2014} to derive luminosity and planetary mass-dependent edges of the \gls{HZ} $a_\mathrm{inner}$ and $a_\mathrm{outer}$.
 We further follow \citet{Bixel2021} and classify as ``exo-Earth candidates'' all planets with radii $0.8 S^{0.25} < R < 1.4 $ that orbit within the above boundaries.
@@
 \subsection{H3: Life only originates on planets with particular UV irradiance}
-Our third theoretical experiment is to test the UV irradiance requirement (Sect.~\ref{sec:predictions:uv}) by relating the occurrence of life on an exo-earth candidate with a minimum past quiescent stellar UV flux, focusing on the prebiotically interesting NUV range from \SIrange{200}{280}{\nano\meter}.
+Our third theoretical experiment is to test the UV irradiance requirement (Sect.~\ref{sec:predictions:uv}) by relating the occurrence of life on an exo-earth candidate with a minimum past quiescent stellar UV flux, focusing on the prebiotically interesting \gls{NUV} range from \SIrange{200}{280}{\nano\meter}.
 \todo[inline]{SUKRIT: got a good reference for this?}  %~\citep{RanjanSasselov2016}.?
 Our concrete hypothesis shall be that life only occurs on planets that at some point in their history have received such radiation exceeded a minimum flux $F_\mathrm{NUV, min}$.
-We further require this flux to have lasted for a minimum duration $\Delta T_\mathrm{min}$ to allow for a sufficient ``origins timescale''~\citep{Rimmer2023}, and the planet to be within its momentary habitable zone during this entire period.
+We further require this flux to have lasted for a minimum duration $\Delta T_\mathrm{min}$ to allow for a sufficient ``origins timescale''~\citep{Rimmer2023}, and the planet to be within its momentary \gls{HZ} during this entire period.
 
-For each exo-earth candidate, we interpolated the age- and stellar mass-dependent habitable zone fluxes obtained by \citet{Richey-Yowell2023} and computed the evolution of the inner (runaway greenhouse) and outer (maximum greenhouse) edges as a function of planet mass, stellar spectral type using the results of \citet{Kopparapu2014} and the stellar luminosity evolution of \citet{Baraffe1998}.
-We then determined which planets were both in the HZ and had NUV fluxes above $F_\mathrm{NUV, min}$ for $\Delta T_\mathrm{min} \geq \SI{100}{\mega\year}$.
+For each exo-earth candidate, we interpolated the age- and stellar mass-dependent \gls{HZ} fluxes obtained by \citet{Richey-Yowell2023} and computed the evolution of the inner (runaway greenhouse) and outer (maximum greenhouse) edges as a function of planet mass, stellar spectral type using the results of \citet{Kopparapu2014} and the stellar luminosity evolution of \citet{Baraffe1998}.
+We then determined which planets were both in the \gls{HZ} and had \gls{NUV} fluxes above $F_\mathrm{NUV, min}$ for $\Delta T_\mathrm{min} \geq \SI{100}{\mega\year}$.
 We assigned the development of life to a fraction $f_\mathrm{life}$ of all temperate planets fulfilling these requirements.
@@
 
-\section{Discussion and Conclusions}
+\section{Discussion}
 \label{sec:discussion}
@@
 
-\subsection{Conclusions and future work}
+\section{Conclusions and future work}
 

### 2023-11-01 src/tex/ms.tex
"signatures", not "markers".

@@
 Presumably, not all habitable worlds are inhabited and not all inhabited worlds develop detectable biosignatures.
-The fraction of exo-Earth candidates that are both inhabited and harbor detectable biomarkers at the time when we observe them remains speculative; we aggregate them in the unitless parameter $f_\mathrm{life}$.
+The fraction of exo-Earth candidates that are both inhabited and harbor detectable biosignatures at the time when we observe them remains speculative; we aggregate them in the unitless parameter $f_\mathrm{life}$.
 
@@
 \section{Exoplanet survey simulations}
-\todo[inline]{biomarkers: Focus on molecular Oxygen? Excess Methane? \citep[e.g.,][]{Seeburger2023}}
+\todo[inline]{biosignatures: Focus on molecular Oxygen? Excess Methane? \citep[e.g.,][]{Seeburger2023}}
 % detecting methanogenic life: Sandora2020, Wang2018,
-A commonly discussed biomarker is molecular Oxygen (O$_2$), which on Earth emerged as a byproduct of photosynthesis during the Proterozoic era.
-While not the only atmospheric species discussed as a signature for planetary life CITE, we focused on O$_2$ as a representative biomarker for our spectroscopic survey simulations.
+A commonly discussed biosignature is molecular Oxygen (O$_2$), which on Earth emerged as a byproduct of photosynthesis during the Proterozoic era.
+While not the only atmospheric species discussed as a signature for planetary life CITE, we focused on O$_2$ as a representative biosignature for our spectroscopic survey simulations.
 
@@
 
+\subsubsection{Diurnal cycles}
+\todo[inline]{We assume that tidal locking implies that there are no wet-dry cycles. Challenge this assumption.}
+
 \section{Conclusions and future work}

### 2023-11-03 src/tex/ms.tex
include variables, new gls entries

@@
 We use the parametrization in \citet{Kopparapu2014} to derive luminosity and planetary mass-dependent edges of the \gls{HZ} $a_\mathrm{inner}$ and $a_\mathrm{outer}$.
-We further follow \citet{Bixel2021} and classify as ``exo-Earth candidates'' all planets with radii $0.8 S^{0.25} < R < 1.4 $ that orbit within the above boundaries.
+We further follow \citet{Bixel2021} and classify as \glspl{EEC} all planets with radii $0.8 S^{0.25} < R < 1.4 $ that orbit within the above boundaries.
 The lower limit was suggested as a minimum planet size to retain an atmosphere~\citep{Zahnle2017}.
@@
 Presumably, not all habitable worlds are inhabited and not all inhabited worlds develop detectable biosignatures.
-The fraction of exo-Earth candidates that are both inhabited and harbor detectable biosignatures at the time when we observe them remains speculative; we aggregate them in the unitless parameter $f_\mathrm{life}$.
+The fraction of \glspl{EEC} that are both inhabited and harbor detectable biosignatures at the time when we observe them remains speculative; we aggregate them in the unitless parameter $f_\mathrm{life}$.
 
@@
 
-For each exo-earth candidate, we interpolated the age- and stellar mass-dependent \gls{HZ} fluxes obtained by \citet{Richey-Yowell2023} and computed the evolution of the inner (runaway greenhouse) and outer (maximum greenhouse) edges as a function of planet mass, stellar spectral type using the results of \citet{Kopparapu2014} and the stellar luminosity evolution of \citet{Baraffe1998}.
-We then determined which planets were both in the \gls{HZ} and had \gls{NUV} fluxes above $F_\mathrm{NUV, min}$ for $\Delta T_\mathrm{min} \geq \SI{100}{\mega\year}$.
+For each exo-earth candidate, we interpolated the age- and stellar mass-dependent \gls{NUV} fluxes in the \gls{HZ} obtained by \citet{Richey-Yowell2023} and computed the evolution of the inner (runaway greenhouse) and outer (maximum greenhouse) edges as a function of planet mass, stellar spectral type using the results of \citet{Kopparapu2014} and the stellar luminosity evolution of \citet{Baraffe1998}.
+We then determined which planets were both in the \gls{HZ} and had \gls{NUV} fluxes above $F_\mathrm{NUV, min}$ for $\Delta T_\mathrm{min} \geq \var{deltaT_min}\,\SI{}{\mega\year}$.
 We assigned the development of life to a fraction $f_\mathrm{life}$ of all temperate planets fulfilling these requirements.
@@
             0 , & F_\mathrm{NUV} < F_\mathrm{NUV, min}\\
-            f_\mathrm{life}, & F_\mathrm{NUV} \geq F_\mathrm{NUV, min} \,\mathrm{and\, in\, HZ} \, \mathrm{for} \, \Delta t \geq \SI{100}{\mega\year}
+            f_\mathrm{life}, & F_\mathrm{NUV} \geq F_\mathrm{NUV, min} \,\mathrm{and\, in\, HZ} \, \mathrm{for} \, \Delta t \geq \var{deltaT_min} \,\SI{}{\mega\year}
         \end{cases}

### 2023-11-03 src/scripts/bioverse_pipeline.py
prepare a bioverse pipeline
showyourwork workflow doesn't seem to work at this point.

@@ -1,0 +1,223 @@
+"""
+This file contains the code to generate the datasets for the paper.
+It heavily relies on Bioverse and its auxiliary functions.
+"""
+import pickle
+
+import paths
+from utils import *
+
+# import pickle
+import numpy as np
+
+# Import the Generator class
+from bioverse.generator import Generator
+from bioverse.survey import TransitSurvey
+
+# from bioverse.constants import CONST, DATA_DIR
+
+
+def inject_nuv_life_correlation(d, f_life=0.1):
+    """Inject a correlation between past NUV flux and the occurrence of life.
+
+    Uses the 'hz_and_uv' column created with Bioverse's `Table.evolve()` method that
+    indicates whether a planet fulfills the following criteria:
+     1. it is in the habitable zone now
+     2. In its past, it simultaneously received a NUV flux above a threshold and was
+     in the habitable zone for a minimum amount of time.
+
+    Parameters
+    ----------
+    d : Table
+        The table of planets to inject the correlation into.
+    f_life : float
+        The fraction of planets fulfilling the above criteria
+        that are inhabited.
+
+    Returns
+    -------
+    d : Table
+        The table of planets with the injected correlation
+        and including a new column `inhabited`.
+    """
+
+    d["inhabited"] = np.logical_and(
+        d["hz_and_uv"], np.random.random(size=len(d)) <= f_life
+    )
+    return d
+
+
+def inject_biosignature(d, P=1.0):
+    # Randomly assign O2 to some inhabited planets
+    d["has_O2"] = np.logical_and(d["inhabited"], np.random.random(size=len(d)) <= P)
+
+    return d
+
+
+def get_generator_args():
+    """define generator parameters"""
+    # Parameters for star generation
+    stars_args = {
+        "d_max": 30,  # max. dist to stars (pc)
+        "M_st_max": 1.5,  # Maximum stellar mass to consider, in solar units.
+        # 'M_G_max' : 11,           # Maximum gaia magnitude of stars
+        "M_G_max": 16,  # Maximum gaia magnitude of stars
+        "seed": 42,  # seed for random number generators
+        "lum_evo": True,  # luminosities from luminosity tracks (Baraffe et al. 1998), based on random ages
+    }
+
+    # Parameters for planet generation
+    planets_args = {
[…]

### 2023-11-03 src/tex/ms.tex
new plot showing spT distribution of different planets

@@
 
+\begin{figure}
+    \begin{centering}
+        \includegraphics[width=\hsize]{figures/inhabited_FGKM.pdf}
+        \caption{Host stars of all transiting planets, \glspl{EEC}, and inhabited planets.
+In a sample of \var{N_nautilus} transiting planets, most \glspl{EEC} and all inhabited planets orbit M~dwarfs.}
+        \label{fig:inhabited_FGKM}
+    \end{centering}
+\end{figure}
+
+As shown in Figure~\ref{fig:inhabited_FGKM}, the majority of \glspl{EEC} and all inhabited planets orbit M~dwarfs.
+
 \subsubsection{past UV flux}

### 2023-11-03 src/scripts/inhabited_FGKM.py

@@ -1,0 +1,35 @@
+import pickle
+import paths
+import plotstyle
+plotstyle.styleplots()
+
+import seaborn as sns
+
+def plot_inhabited_FGKM(d):
+
+    dd = d.to_pandas()
+
+    def cat(obj):
+        if obj.EEC:
+            cat = "EEC"
+            if obj.inhabited:
+                cat = "inhabited"
+        else:
+            cat = "non-EEC"
+        return cat
+
+    dd.loc[:, "Category"] = dd.apply(cat, axis=1)
+    
+    g = sns.catplot(
+        dd, x="SpT", kind="count", hue="Category", order=['F', 'G', 'K', 'M']
+    )  # , hue_order = [True, False])
+    g.ax.set_yscale("log")
+
+    g.savefig(paths.figures / "inhabited_FGKM.pdf")
+    return g
+
+
+with open(paths.data / 'pipeline/sample.pkl', 'rb') as f:
+    sample = pickle.load(f)
+
+g = plot_inhabited_FGKM(sample)

### 2023-11-03 src/scripts/hz_nuv_evo.py
new plot with lum, NUV flux interpolations

@@ -1,0 +1,57 @@
+import paths
+import pickle
+import plotstyle
+plotstyle.styleplots()
+
+import numpy as np
+import matplotlib.pyplot as plt
+import matplotlib.colors as colors
+
+from bioverse.util import interpolate_luminosity, interpolate_nuv
+
+def plot_interpolation(fig, ax, T,M,Z, d=None, input_points=None, log=True, cbarlabel=None, **kwargs):
+    if log:
+        norm = colors.LogNorm()
+    else:
+        norm = colors.Normalize()
+    q = ax.pcolormesh(T, M, Z, shading='auto',
+                   norm=norm, rasterized=True, **kwargs)
+    if input_points:
+        ax.plot(*input_points, "x", ms=1, label="input point")
+    if d:
+        ax.scatter(d['age'], d['M_st'], c=d['L_st_interp'], s=4,
+                norm=norm, edgecolors='white', linewidths=0.3, label="transiting planet")
+    # ax.legend()
+    fig.colorbar(q, label=cbarlabel)
+    ax.set_xscale('log')
+    ax.set_xlabel('Time (Gyr)')
+    ax.set_ylabel('Stellar Mass ($M_\odot$)')
+    return ax
+
+
+def plot_nuv_evo(fig, ax):
+    interp_nuv = interpolate_nuv()
+    T = np.geomspace(0.01, 9., num=200)
+    M = np.linspace(0.1, 0.9, num=200)
+    M, T = np.meshgrid(M, T)
+    Z = interp_nuv(M, T)
+    plot_interpolation(fig, ax, T,M,Z, log=True, cbarlabel='NUV flux [erg/s/cm$^2$]', cmap='viridis')
+
+
+
+with open(paths.data / 'pipeline/sample.pkl', 'rb') as f:
+    d = pickle.load(f)
+
+M = np.linspace(0.1, 1.4, num=200)
+A = np.geomspace(5e-4, 10., num=200)
+M, A = np.meshgrid(M, A)
+interp_lum, extrap_nn = interpolate_luminosity()
+
+Z = interp_lum(M, A)
+d['L_st_interp'] = interp_lum(d['M_st'], d['age'])
+
+fig, axs = plt.subplots(1, 2, figsize=(12, 3.5))
+axs[0] = plot_interpolation(fig, axs[0], A,M,Z, d,  cbarlabel='Bol. luminosity ($L/L_\odot$)')
[…]

### 2023-11-03 src/tex/ms.tex
new plot with lum, NUV flux interpolations
describe interpolations

@@
 \subsection{H3: Life only originates on planets with particular UV irradiance}
+\todo[inline]{\citet{Guenther2020} relate U-band energy to bolometric flux.}
+\todo[inline]{OPTIONAL: ``We further test the scenario of a linear correlation of past UV flux and biosignature occurrence rate.
+This test requires the detection of multiple biosignatures.''
+~\\
+Test for negative correlations as well?}
 Our third theoretical experiment is to test the UV irradiance requirement (Sect.~\ref{sec:predictions:uv}) by relating the occurrence of life on an exo-earth candidate with a minimum past quiescent stellar UV flux, focusing on the prebiotically interesting \gls{NUV} range from \SIrange{200}{280}{\nano\meter}.
@@
 
-For each exo-earth candidate, we interpolated the age- and stellar mass-dependent \gls{NUV} fluxes in the \gls{HZ} obtained by \citet{Richey-Yowell2023} and computed the evolution of the inner (runaway greenhouse) and outer (maximum greenhouse) edges as a function of planet mass, stellar spectral type using the results of \citet{Kopparapu2014} and the stellar luminosity evolution of \citet{Baraffe1998}.
+To determine \gls{HZ} occupancy, we interpolated the stellar luminosity evolution grid of \citet{Baraffe1998} using a Clough Tocher interpolant~\citep[][see left panel of Fig.~\ref{fig:hz_nuv_evo}]{Nielson1983,Alfeld1984} to compute the evolution of the inner (runaway greenhouse) and outer (maximum greenhouse) edges as a function of planet mass and stellar spectral type~\citep{Kopparapu2014}.
+This provides each planet's epochs within and outside the \gls{HZ}.
+For the \gls{NUV} flux, we use the age- and stellar mass-dependent \gls{NUV} fluxes in the \gls{HZ} obtained by \citet{Richey-Yowell2023}.
+We linearly interpolate in their measured grid, where we convert spectral type to stellar mass using the midpoints of their mass ranges (\SI{0.75}{\Msun} for K~stars, \SI{0.475}{\Msun} for early-type M~stars,  and \SI{0.215}}{\Msun} for late-type M~stars).
+Outside of the age and stellar mass range covered in \citet{Richey-Yowell2023}, we extrapolate using nearest simplex (see right panel of Fig.~\ref{fig:hz_nuv_evo}).
+
 We then determined which planets were both in the \gls{HZ} and had \gls{NUV} fluxes above $F_\mathrm{NUV, min}$ for $\Delta T_\mathrm{min} \geq \var{deltaT_min}\,\SI{}{\mega\year}$.
-We assigned the development of life to a fraction $f_\mathrm{life}$ of all temperate planets fulfilling these requirements.
+We assigned the development of life to a random fraction $f_\mathrm{life}$ of all temperate planets fulfilling these requirements.
 For a population-level hypothesis, we then have
@@
 
-\subsubsection{past UV flux}
-\todo[inline]{\citet{Günther2020} relate U-band energy to bolometric flux.}
-
-\todo[inline]{OPTIONAL: ``We further test the scenario of a linear correlation of past UV flux and biosignature occurrence rate.
-This test requires the detection of multiple biosignatures.''
-~\\
-Test for negative correlations as well?}
+\begin{figure*}
+    \begin{centering}
+        \includegraphics[width=\hsize]{figures/hz_nuv_evo.pdf}
+        \caption{Stellar luminosity evolution (left) and evolution of the \gls{NUV} flux in the \gls{HZ} as a function of host star mass.
+White markers show age and host star mass of the transiting planets in the synthetic sample.}
+        \label{fig:hz_nuv_evo}
+    \end{centering}
+\end{figure*}
 

### 2023-11-06 src/tex/ms.tex
add planet examples and their HZ&NUV evolution to plot

@@
 For the \gls{NUV} flux, we use the age- and stellar mass-dependent \gls{NUV} fluxes in the \gls{HZ} obtained by \citet{Richey-Yowell2023}.
-We linearly interpolate in their measured grid, where we convert spectral type to stellar mass using the midpoints of their mass ranges (\SI{0.75}{\Msun} for K~stars, \SI{0.475}{\Msun} for early-type M~stars,  and \SI{0.215}}{\Msun} for late-type M~stars).
+We linearly interpolate in their measured grid, where we convert spectral type to stellar mass using the midpoints of their mass ranges (\SI{0.75}{\Msun} for K~stars, \SI{0.475}{\Msun} for early-type M~stars,  and \SI{0.215}{\Msun} for late-type M~stars).
 Outside of the age and stellar mass range covered in \citet{Richey-Yowell2023}, we extrapolate using nearest simplex (see right panel of Fig.~\ref{fig:hz_nuv_evo}).
@@
         \includegraphics[width=\hsize]{figures/hz_nuv_evo.pdf}
-        \caption{Stellar luminosity evolution (left) and evolution of the \gls{NUV} flux in the \gls{HZ} as a function of host star mass.
-White markers show age and host star mass of the transiting planets in the synthetic sample.}
+        \caption{Interpolated stellar luminosity evolution (left) and evolution of the \gls{NUV} flux in the \gls{HZ} as a function of host star mass.
+White markers show age and host star mass of the transiting planets in the synthetic sample.
+        Two of the three planet trajectories shown have extended overlap of \gls{HZ} occupancy and high \gls{NUV} flux (green sections), our requirement for abiogenesis.}
         \label{fig:hz_nuv_evo}

### 2023-11-06 src/scripts/hz_nuv_evo.py

@@ -1,6 +1,7 @@
 import paths
 import pickle
 import plotstyle
+
 plotstyle.styleplots()
 
 import numpy as np
@@ -9,49 +10,98 @@ import matplotlib.colors as colors
 
 from bioverse.util import interpolate_luminosity, interpolate_nuv
 
-def plot_interpolation(fig, ax, T,M,Z, d=None, input_points=None, log=True, cbarlabel=None, **kwargs):
+
+def plot_interpolation(
+    fig, ax, T, M, Z, d=None, input_points=None, log=True, cbarlabel=None, **kwargs
+):
     if log:
         norm = colors.LogNorm()
     else:
         norm = colors.Normalize()
-    q = ax.pcolormesh(T, M, Z, shading='auto',
-                   norm=norm, rasterized=True, **kwargs)
+    q = ax.pcolormesh(T, M, Z, shading="auto", norm=norm, rasterized=True, **kwargs)
     if input_points:
         ax.plot(*input_points, "x", ms=1, label="input point")
     if d:
-        ax.scatter(d['age'], d['M_st'], c=d['L_st_interp'], s=4,
-                norm=norm, edgecolors='white', linewidths=0.3, label="transiting planet")
+        ax.scatter(
+            d["age"],
+            d["M_st"],
+            c=d["L_st_interp"],
+            s=4,
+            norm=norm,
+            edgecolors="white",
+            linewidths=0.3,
+            label="transiting planet",
+        )
     # ax.legend()
     fig.colorbar(q, label=cbarlabel)
-    ax.set_xscale('log')
-    ax.set_xlabel('Time (Gyr)')
-    ax.set_ylabel('Stellar Mass ($M_\odot$)')
+    ax.set_xscale("log")
+    ax.set_xlabel("Time (Gyr)")
+    ax.set_ylabel("Stellar Mass ($M_\odot$)")
     return ax
 
 
 def plot_nuv_evo(fig, ax):
     interp_nuv = interpolate_nuv()
-    T = np.geomspace(0.01, 9., num=200)
-    M = np.linspace(0.1, 0.9, num=200)
+    T = np.geomspace(5e-3, 10., num=200)
+    M = np.linspace(0.1, 1.0, num=200)
[…]

### 2023-11-06 src/scripts/hz_nuv_evo.py
highlight EECs

@@ -22,17 +22,20 @@ def plot_interpolation(
     if input_points:
         ax.plot(*input_points, "x", ms=1, label="input point")
     if d:
-        ax.scatter(
-            d["age"],
-            d["M_st"],
-            c=d["L_st_interp"],
-            s=4,
-            norm=norm,
-            edgecolors="white",
-            linewidths=0.3,
-            label="transiting planet",
-        )
-    # ax.legend()
+        # for (v, c) in [(False, 'white'), (True, 'C1')]:
+        for c, (EEC, group) in zip(['white', 'C0'], d.to_pandas().groupby('EEC')):
+            ax.scatter(
+                group["age"],
+                group["M_st"],
+                c=group["L_st_interp"],
+                s=4,
+                norm=norm,
+                # edgecolors="white",
+                edgecolors=c,
+                linewidths=0.6,
+                label=["EEC" if EEC else None][0],
+            )
+        ax.legend(loc='upper left')
     fig.colorbar(q, label=cbarlabel)
     ax.set_xscale("log")
     ax.set_xlabel("Time (Gyr)")

### 2023-11-06 src/tex/ms.tex
Restructure: one big Methods section.
No separate Survey Sims section.
highlight EECs
add short results, discussion, conclusion on spT dependency

@@
 
-\section{Hypothesis Tests}
+\section{Methods}
 \label{sec:hypotests}
 \todo[inline]{Briefly introduce Bayesian model comparison, then present the particular hypotheses in their mathematical form.}
+\subsection{Bayesian hypothesis testing with \bioverse}
 
@@
 
-\subsection{Bayesian hypothesis testing with \bioverse}
+\subsection{Exoplanet survey simulations}
+\todo[inline]{biosignatures: Focus on molecular Oxygen? Excess Methane? \citep[e.g.,][]{Seeburger2023}}
+% detecting methanogenic life: Sandora2020, Wang2018,
+A commonly discussed biosignature is molecular Oxygen (O$_2$), which on Earth emerged as a byproduct of photosynthesis during the Proterozoic era.
+While not the only atmospheric species discussed as a signature for planetary life CITE, we focused on O$_2$ as a representative biosignature for our spectroscopic survey simulations.
+
+\subsubsection{Habitable Worlds Observatory}
+\subsubsection{target list}
+\todo[inline]{Provisional stellar target List for the habitable Worlds Observatory:~\citep{Mamajek2023}} % read like in https://github.com/mkenworthy/HWObows/tree/40b1e5f2bc2bbbe8c999b097c80ca5a8cda0aeae/src/scripts
+
+\subsubsection{Nautilus}
+\subsubsection{LIFE} % optional
 
@@
         \caption{Interpolated stellar luminosity evolution (left) and evolution of the \gls{NUV} flux in the \gls{HZ} as a function of host star mass.
-White markers show age and host star mass of the transiting planets in the synthetic sample.
+The scatter plot shows age and host star mass of the transiting planets in the synthetic sample with \gls{EEC} highlighted in green.
         Two of the three planet trajectories shown have extended overlap of \gls{HZ} occupancy and high \gls{NUV} flux (green sections), our requirement for abiogenesis.}
@@
 \end{figure*}
-
-\section{Exoplanet survey simulations}
-\todo[inline]{biosignatures: Focus on molecular Oxygen? Excess Methane? \citep[e.g.,][]{Seeburger2023}}
-% detecting methanogenic life: Sandora2020, Wang2018,
-A commonly discussed biosignature is molecular Oxygen (O$_2$), which on Earth emerged as a byproduct of photosynthesis during the Proterozoic era.
-While not the only atmospheric species discussed as a signature for planetary life CITE, we focused on O$_2$ as a representative biosignature for our spectroscopic survey simulations.
-
-\subsection{Habitable Worlds Observatory}
-\subsubsection{target list}
-\todo[inline]{Provisional stellar target List for the habitable Worlds Observatory:~\citep{Mamajek2023}} % read like in https://github.com/mkenworthy/HWObows/tree/40b1e5f2bc2bbbe8c999b097c80ca5a8cda0aeae/src/scripts
-
-\subsection{Nautilus}
-\subsection{LIFE} % optional
 
@@
 
-\subsection{Correlation of past UV flux and biosignature occurrence}
+\subsection{Correlation of past UV flux and biosignature occurrence}\label{sec:results_uv}
+
+In a magnitude- and volume-limited sample of a transit survey, the host star distribution will be skewed toward later spectral types and dominated by M~dwarfs (see Fig.~\ref{fig:inhabited_FGKM}).
+Due to how the \gls{HZ} scales with spectral type, by far most transiting \glspl{EEC} occur around M~dwarfs.
+These host stars, in particular late subtypes, also provide extended periods of increased \gls{NUV} emission that overlap with times when some of these planets occupy the \gls{HZ} (see Fig.~\ref{fig:hz_nuv_evo}), our requirement for abiogenesis (compare Equation~\ref{eq:hypothesis-uv}).
+Because of that, all inhabited planets orbit M~dwarfs.
 
@@
 \todo[inline]{discuss how planetary context may impact credibility of tentative biosignature detections (e.g., if there is a good fit with a predicted OOL pattern; or the opposite: the planetary context does not fit well to any OOL scenarios)}
+As we have shown in Section~\ref{sec:results_uv}, the interplay of \gls{NUV} evolution and \gls{HZ} occupancy strongly favors late spectral types for abiogenesis via the cyanosulfidic scenario.
+This strong predicted correlation between stellar spectral type and the occurrence of life can be used to falsify \gls{ool} scenarios.
+The constraint of this context is particularly strong if a candidate biosignature is detected on a planet orbiting an earlier type star, i.e., where it is unexpected in the context of this \gls{ool} scenario.
 
@@
 
+Our main findings are:
+\begin{enumerate}
+    \item Abiogenesis in the cyanosulfidic scenario strongly prefers late host star spectral types in a sample of transiting planets.
+Any potential biosignature detection on a planet around a star other than an M~dwarf would disvafor this scenario, and any \gls{ool} theory requiring extended periods of high \gls{NUV} irradiation.
+\end{enumerate}
+
 %\input{acknowledgments_contributions}

### 2023-11-07 src/tex/ms.tex
write up prelim. conclusions

@@
 \begin{enumerate}
+    \item Models of the Origins of Life provide hypotheses that are testable with near-future exoplanet surveys.
+    \item The required \gls{nuv} radiation in the cyanosulfidic scenario should lead to a correlation between past \gls{nuv} flux and current occurrence of biosignatures, both of which will soon be observationally accessible.
+    \item The required sample size for detecting this correlation depends on the occurrence of life on temperate exoplanets, the distribution of host star spectral types in the sample, ...
+        A direct imaging mission with the \gls{hwo} will require characterizing XXX planets ...
     \item Abiogenesis in the cyanosulfidic scenario strongly prefers late host star spectral types in a sample of transiting planets.
-Any potential biosignature detection on a planet around a star other than an M~dwarf would disvafor this scenario, and any \gls{ool} theory requiring extended periods of high \gls{NUV} irradiation.
+Any potential biosignature detection on a planet around a star other than an M~dwarf would disvafor this scenario and any other \gls{ool} theory requiring extended periods of high \gls{nuv} irradiation.
 \end{enumerate}

### 2023-12-12 src/tex/ms.tex
split methods into semi-analytical analysis and Bioverse

@@
 Our concrete hypothesis shall be that life only occurs on planets that at some point in their history have received such radiation exceeded a minimum flux $F_\mathrm{NUV, min}$.
-We further require this flux to have lasted for a minimum duration $\Delta T_\mathrm{min}$ to allow for a sufficient ``origins timescale''~\citep{Rimmer2023}, and the planet to be within its momentary \gls{HZ} during this entire period.
+
+\subsubsection{Semi-analytical analysis}
+First, we apply a Bayes Factor Design Analysis~\citep{Schoenbrodt2018} to assess the expected probabilities of obtaining true negative or true positive evidence for the hypothesis above, as well as the probability for misleading or inconclusive evidence.
+
+\subsubsection{Survey simulations with Bioverse}
+To model the sensitivity of the information gain of a proposed mission to sample selection and survey strategy, we now conduct survey simulations with Bioverse using different sample sizes and survey strategies.
+We again require that life occurs only on planets with sufficient past UV irradiation exceeding $F_\mathrm{NUV, min}$.
+Further, we require this flux to have lasted for a minimum duration $\Delta T_\mathrm{min}$ to allow for a sufficient ``origins timescale''~\citep{Rimmer2023}, and the planet to be within its momentary \gls{HZ} during this entire period.
 

### 2023-12-12 src/tex/ms.tex
define likelihoods and Bayes factors

@@
 
+Let our observable be the inferred past \gls{NUV} flux of the planet $F_\mathrm{NUV}$.
+Under Hypothesis $H_3$ (Equation \ref{eq:hypothesis-uv}), there exists a special unknown value of $F_\mathrm{NUV}$, noted $F_\mathrm{NUV, min}$ such that
+\begin{array}{rcll}
+P(L|\theta,H_3) & = & f_\mathrm{life} &\text{if } \theta>F_\mathrm{NUV, min}\\
+P(L|\theta,H_3) & = & 0 &\text{otherwise}
+\end{array}
+where $f_\mathrm{life}$ is the unknown probability of abiogenesis.
+The corresponding null hypothesis is that there exists no such special value of $F_\mathrm{NUV}$ and that
+\begin{equation}
+P(L|\theta,H_\mathrm{3, null}) = f_\mathrm{life}.
+\end{equation}
+
+If we now define a sample of size $n$ as $X=\{F_\mathrm{NUV, i},L_i\}_{i \in [1,n]}$ where $L_i$ is equal to 1 if life is detected and 0 otherwise, we can calculate the evidence for hypothesis $H_i$ against $H_j$ through the Bayes factor
+\begin{equation}
+BF_{H_i,H_j} = \frac{P(X|H_i)}{P(X|H_j)},
+\end{equation}
+with $P(X|H_i)$ and $P(X|H_j)$ likelihoods of obtaining the sample $X$ under either hypothesis.
+
+If we define $k=\sum L_i$ and denote $Y$ the random variable that describes it, $H_\mathrm{3, null}$ represents the likelihood that the number of planets with life in the sample follows the binomial distribution
+\begin{equation}
+P(Y=k|H_\mathrm{3, null}) = \binom{n}{k}f_\mathrm{life}^k(1-f_\mathrm{life})^{n-k}.
+\end{equation}
+
+Under $H_3$, $Y$ also follows a binomial distribution, however it is conditioned by $n_{\lambda}=Card(\{F_\mathrm{NUV, i} \text{ if } F_\mathrm{NUV, i}>F_\mathrm{NUV, min}\}_{i \in [1,n]})$ the number of values of $F_\mathrm{NUV}$ in the experiment that exceed $F_\mathrm{NUV, min}$
+\begin{equation}
+P(Y=k|H_3) = \binom{n_{\lambda}}{k}f_\mathrm{life}^k(1-f_\mathrm{life})^{n_{\lambda}-k}.
+\end{equation}
+
+Hence,
+\begin{equation}
+BF_{H_3,H_\mathrm{3, null}} = \frac{P(Y=k|H_3)}{P(Y=k|H_\mathrm{3, null})} = \frac{\binom{n_\lambda}{k}}{\binom{n}{k}}(1-f_\mathrm{life})^{n_{\lambda}-n}.
+\end{equation}
+
 \subsubsection{Survey simulations with Bioverse}

### 2023-12-13 src/tex/ms.tex
add true evidence plot

@@
 Hence,
-\begin{equation}
+\begin{equation}\label{eq:bayes_factor}
 BF_{H_3,H_\mathrm{3, null}} = \frac{P(Y=k|H_3)}{P(Y=k|H_\mathrm{3, null})} = \frac{\binom{n_\lambda}{k}}{\binom{n}{k}}(1-f_\mathrm{life})^{n_{\lambda}-n}.
 \end{equation}
+
+Given a sample of planets, where for some of them we have convincing biosignature detections but remaining agnostic on $f_\mathrm{life}$: What evidence for $H_\mathrm{3}$ and $H_\mathrm{3, null}$ can we expect to get?
+Our Bayes factor (Equation~\ref{eq:bayes_factor}) is determined by the unknown variables $f_\mathrm{life}$ and $F_\mathrm{NUV, min}$, as well as the number of planets with biosignature detections in the sample $k$.
+To compute the distribution of evidences, we repeatedly generated samples under $H_\mathrm{3}$ and $H_\mathrm{3, null}$ and computed the Bayes factors $BF_{H_3,H_\mathrm{3, null}}$ and $BF_{H_\mathrm{3, null}, H_3}$.
+We then evaluated the fraction of Monte Carlo runs in which certain evidence thresholds~\citep{Jeffreys1939} were exceeded.
 
@@
 \begin{figure}
+    \script{inhabited_FGKM.py}
     \begin{centering}
@@
 \subsection{Correlation of past UV flux and biosignature occurrence}\label{sec:results_uv}
+
+\subsubsection{Semi-analytical approach}
+In Section~\ref{sec:met-semianalytical} We computed the probability for true positive evidence for $H_\mathrm{3}$ and $H_\mathrm{3, null}$ respectively.
+Figure~\ref{fig:semian_true_evidence} shows how these evidences are distributed for sample sizes \var{semian_Nsamp1} and \var{semian_Nsamp2}, and how likely we are to obtain strong evidence ($BF_{H_i, H_j}$ > 10).
+For $n = \var{semian_Nsamp1}$, strong true evidence for $H_\mathrm{3}$ ($H_\mathrm{3, null}$) can be expected in $\sim \SI{30}{\percent}$ ($\sim \SI{40}{\percent}$) of all random experiments.
+In the majority of cases, the outcome of the survey will be inconclusive.
+\begin{figure}
+    \script{semian_true_evidence.py}
+    \begin{centering}
+        \includegraphics[width=\hsize]{figures/semian_true_evidence.pdf}
+        \caption{Probability to obtain true strong evidence for $H_\mathrm{3}$ and $H_\mathrm{3, null}$ under sample sizes $n = \var{semian_Nsamp1}$ (solid) and $n = \var{semian_Nsamp2}$ (dashed).
+The vertical lines denote the thresholds for ``strong'' evidence, $BF_{H_i, H_j}$ > 10, and ``extreme'' evidence, $BF_{H_i, H_j}$ > 100.}
+        \label{fig:semian_true_evidence}
+    \end{centering}
+\end{figure}
+The situation improves for larger samples -- \SI{80}{\percent} of random experiments yield true strong evidence under either $H_\mathrm{3}$ or $H_\mathrm{3, null}$.
+
+\subsubsection{Survey simulations with Bioverse}
 

### 2023-12-13 src/scripts/semian_true_evidence.py

@@ -1,0 +1,131 @@
+import paths
+import pickle
+import plotstyle
+
+plotstyle.styleplots()
+
+import numpy as np
+import matplotlib.pyplot as plt
+import scipy.special as scp
+from scipy import stats
+
+rng = np.random.default_rng(seed=42)
+
+def BFH1(k, theta, pi, theta_l):
+    """
+    Evidence for H1 against H0
+    k number of life detection in the sample
+    theta : vector of theta values in the sample
+    pi : probability of life emergence
+    theta_l : threshold of theta for life emergence
+
+    returns bf the bayes ratio P(k|H1)/P(k|H0)
+    """
+    if len(np.shape(theta_l)) == 2:
+        nl = np.sum(theta > theta_l, axis=1)
+    elif len(np.shape(theta_l)) == 1:
+        nl = np.sum(theta > np.reshape(theta_l, (len(theta_l), 1)), axis=1)
+    else:
+        nl = np.sum(theta > theta_l)
+    bf = stats.binom.pmf(k, nl, pi) / stats.binom.pmf(k, len(theta), pi)
+    return bf
+
+def sample_H0(theta,pi,theta_l):
+    """
+    Generates a sample of values of k under H0
+    """
+    # theta is the vector of theta values
+    # pi is the probability of life
+    # theta_l is the value of the limit theta
+    return(np.random.binomial(len(theta),pi))
+
+
+def sample_H1(theta,pi,theta_l):
+    """
+    Generates a sample of values of k under H1
+    """
+    # theta is the vector of theta values
+    # pi is the probability of life
+    # theta_l is the value of the limit theta
+    if len(np.shape(theta_l))==2:
+        X = np.random.binomial(np.sum(theta>theta_l,axis=1),pi)
+    elif len(np.shape(theta_l))==1:
+        X = np.random.binomial(np.sum(theta>np.reshape(theta_l,(len(theta_l),1)),axis=1),pi)
+    else:
+        X = np.random.binomial(np.sum(theta>theta_l),pi)
+    return(X)
+
+def get_evidence(sampsize=100):
+    # For convenience, we redefine theta here so that one can play with sample size and such
+    theta = rng.uniform(
+        0, 1, sampsize
[…]

### 2023-12-13 src/tex/ms.tex
add evidence aafo sample size

@@
 In the majority of cases, the outcome of the survey will be inconclusive.
-\begin{figure}
+\begin{figure*}
     \script{semian_true_evidence.py}
@@
         \includegraphics[width=\hsize]{figures/semian_true_evidence.pdf}
-        \caption{Probability to obtain true strong evidence for $H_\mathrm{3}$ and $H_\mathrm{3, null}$ under sample sizes $n = \var{semian_Nsamp1}$ (solid) and $n = \var{semian_Nsamp2}$ (dashed).
-The vertical lines denote the thresholds for ``strong'' evidence, $BF_{H_i, H_j}$ > 10, and ``extreme'' evidence, $BF_{H_i, H_j}$ > 100.}
+        \caption{Probability to obtain true strong evidence.
+Left: evidence levels for $H_\mathrm{3}$ and $H_\mathrm{3, null}$ under sample sizes $n = \var{semian_Nsamp1}$ (solid) and $n = \var{semian_Nsamp2}$ (dashed).
+The vertical lines denote the thresholds for ``strong'' evidence, $BF_{H_i, H_j}$ > 10, and ``extreme'' evidence, $BF_{H_i, H_j}$ > 100.
+Right: Probability of true strong evidence for $H_\mathrm{3}$ or $H_\mathrm{3, null}$, whichever is smaller, as a function of sample size $n$.}
         \label{fig:semian_true_evidence}
     \end{centering}
-\end{figure}
+\end{figure*}
 The situation improves for larger samples -- \SI{80}{\percent} of random experiments yield true strong evidence under either $H_\mathrm{3}$ or $H_\mathrm{3, null}$.

### 2023-12-13 src/scripts/semian_true_evidence.py

@@ -79,10 +79,9 @@ def get_evidence(sampsize=100):
     return bfh1_kh1, bfh0_kh0
 
 
-def true_evidence():
+def plot_true_evidence(ax):
     gtvec = 10 ** np.linspace(-2, 3, 300)
-    fig, ax = plt.subplots()
-    
+
     for s, ls, showlabel in zip([10, 100], ["-", "--"], ["", "_"]):
         bfh1_kh1, bfh0_kh0 = get_evidence(s)
     
@@ -119,12 +118,69 @@ def true_evidence():
         columnspacing=1.6,
     )
     ax.set_xscale("log")
+    ax.set_ylim(-.05, 1.05)
     ax.set_ylabel("P(evidence>x)")
     ax.set_xlabel("x")
 
-    return fig, ax
+    return ax
 
-fig, ax = true_evidence()
+## calculate P(BF>eta)
+
+
+def P_true_evidence(sampsize, eta=10, nboot=int(1e4), pisamp=None, theta_lsamp=None):
+    theta = rng.uniform(0, 1, sampsize)
+
+    if pisamp is None:
+        pisamp = rng.uniform(0, 1, nboot)
+    elif np.isscalar(pisamp):
+        # try a fixed value for pi
+        pisamp = np.full(nboot, pisamp)
+    else:
+        # seems we're dealing with an array
+        pisamp = np.full((nboot, pisamp.shape[1]), pisamp)
+
+    if theta_lsamp is None:
+        # random threshold
+        theta_lsamp = rng.uniform(0, 1, nboot)
+    elif np.isscalar(theta_lsamp):
+        # try a fixed threshold
+        theta_lsamp = np.full(nboot, theta_lsamp)
+    else:
+        # seems we're dealing with an array
+        theta_lsamp = np.full((nboot, theta_lsamp.shape[1]), theta_lsamp)
+
+    kh1samp = sample_H1(theta, pisamp, theta_lsamp)
+    kh0samp = sample_H0(theta, pisamp, theta_lsamp)
+
+    bfh1_kh1 = BFH1(
+        kh1samp, theta, pisamp, theta_lsamp
+    )  # evidence for H1 when H1 is true (k is sampled under H1)
+    bfh1_kh0 = BFH1(
+        kh0samp, theta, pisamp, theta_lsamp
+    )  # evidence for H1 when H0 is true (k is sampled under H0)
[…]

### 2023-12-15 src/tex/glossary.tex
include variables, new gls entries
add short results, discussion, conclusion on spT dependency
fix compilation errors

@@
 \newglossaryentry{JWST}{name={JWST}, description={James Webb Space Telescope}, first={James Webb Space Telescope~\citep[\textit{JWST},][]{Beichman2014}}}
+\newglossaryentry{HWO}{name={HWO}, description={Habitable Worlds Observatory}, first={Habitable Worlds Observatory (HWO)}}
 \newglossaryentry{HZ}{name={HZ}, description={habitable zone}, first={habitable zone (HZ)}}
+\newglossaryentry{EEC}{name={EEC}, description={exo-Earth candidate}, first={exo-Earth candidate (EEC)}, firstplural={exo-Earth candidates (EEC)}}
+\newglossaryentry{ool}{name={OoL}, description={Origins of Life}, first={Origins of Life (OoL)}}
 %\newglossaryentry{}{name={}, description={}, first={}, firstplural={}}

### 2023-12-15 src/tex/ms.tex
fix compilation errors

@@
 Let our observable be the inferred past \gls{NUV} flux of the planet $F_\mathrm{NUV}$.
-Under Hypothesis $H_3$ (Equation \ref{eq:hypothesis-uv}), there exists a special unknown value of $F_\mathrm{NUV}$, noted $F_\mathrm{NUV, min}$ such that
-\begin{array}{rcll}
-P(L|\theta,H_3) & = & f_\mathrm{life} &\text{if } \theta>F_\mathrm{NUV, min}\\
-P(L|\theta,H_3) & = & 0 &\text{otherwise}
-\end{array}
+Under Hypothesis $H_3$ (Equation~\ref{eq:hypothesis-uv}), there exists a special unknown value of $F_\mathrm{NUV}$, noted $F_\mathrm{NUV, min}$ such that
+
+\begin{align}
+    P(L|\theta,H_3) & = & f_\mathrm{life} & \text{if } \theta>F_\mathrm{NUV, min}\\
+    P(L|\theta,H_3) & = & 0 & \text{otherwise}
+\end{align}
+
 where $f_\mathrm{life}$ is the unknown probability of abiogenesis.
@@
 \begin{figure*}
+    \script{hz_nuv_evo.py}
     \begin{centering}
@@
     \item Models of the Origins of Life provide hypotheses that are testable with near-future exoplanet surveys.
-    \item The required \gls{nuv} radiation in the cyanosulfidic scenario should lead to a correlation between past \gls{nuv} flux and current occurrence of biosignatures, both of which will soon be observationally accessible.
+    \item The required \gls{NUV} radiation in the cyanosulfidic scenario should lead to a correlation between past \gls{NUV} flux and current occurrence of biosignatures, both of which will soon be observationally accessible.
     \item The required sample size for detecting this correlation depends on the occurrence of life on temperate exoplanets, the distribution of host star spectral types in the sample, ...
-        A direct imaging mission with the \gls{hwo} will require characterizing XXX planets ...
+        A direct imaging mission with the \gls{HWO} will require characterizing XXX planets ...
     \item Abiogenesis in the cyanosulfidic scenario strongly prefers late host star spectral types in a sample of transiting planets.
-Any potential biosignature detection on a planet around a star other than an M~dwarf would disvafor this scenario and any other \gls{ool} theory requiring extended periods of high \gls{nuv} irradiation.
+Any potential biosignature detection on a planet around a star other than an M~dwarf would disvafor this scenario and any other \gls{ool} theory requiring extended periods of high \gls{NUV} irradiation.
 \end{enumerate}

### 2023-12-16 src/tex/ms.tex
that can't work
add sampling selectivity results and plots

@@
 \begin{align}
-    P(L|\theta,H_3) & = & f_\mathrm{life} & \text{if } \theta>F_\mathrm{NUV, min}\\
-    P(L|\theta,H_3) & = & 0 & \text{otherwise}
+    P(L|\theta,H_3) &=  f_\mathrm{life} \quad \text{if } \theta>F_\mathrm{NUV, min}\\
+    P(L|\theta,H_3) &=  0               \quad  \text{otherwise}
 \end{align}
@@
 
+So far, we have assumed random, uniform distributions of $f_\mathrm{life}$, $F_\mathrm{NUV, min}$, and $F_\mathrm{NUV}$.
+A high biosignature detection rate $f_\mathrm{life}$ increases the evidence (cmp. Equation~\ref{eq:bayes_factor}) but we cannot influence it.
+The same is true for $F_\mathrm{NUV, min}$, where again higher values increase the evidence as the binomial distribution for $H_\mathrm{3}$ gets increasingly skewed and shifted away from the one for $H_\mathrm{3, null}$.
+The distribution of $F_\mathrm{NUV}$ in the planet sample, on the other hand, can be influenced by the survey strategy.
+A targeted sampling approach could be to favor extreme values of $F_\mathrm{NUV}$ in the sample selection.
+Figure~\ref{fig:semian_selectivity} shows how the probability of obtaining true strong evidence for $H_\mathrm{3}$ scales with selectivity $s$, where $s\in]-1,1[$ such that $F_\mathrm{NUV} \sim Beta(1/10^s,1/10^s)$.
+Here, $s=0$ corresponds to a random uniform distribution.
+Compared to this case, a high selectivity can increase the probability of obtaining true strong evidence to $\gtrsim \SI{85}{\percent}$ for large samples.
+
+\begin{figure}
+    \script{semian_true_evidence.py}
+    \begin{centering}
+        \includegraphics[width=\hsize]{figures/semian_selectivity.pdf}
+        \caption{Scaling of the probability of obtaining true strong evidence with sample selectivity.
+Left: Sampling distribution for different selectivity parameters $s$.
+Right: Resulting \mbox{P(true strong evidence)}.
+Sampling more extreme values of $F_\mathrm{NUV}$ is more likely to yield strong evidence.}
+        \label{fig:semian_selectivity}
+    \end{centering}
+\end{figure}
+
 \subsubsection{Survey simulations with Bioverse}

### 2023-12-16 src/scripts/semian_true_evidence.py

@@ -1,6 +1,7 @@
 import paths
 import pickle
 import plotstyle
+import cmocean
 
 plotstyle.styleplots()
 
@@ -177,11 +178,116 @@ def plot_evidence_sampsize(ax, s_max=500):
     return ax
 
 
+def plot_beta(ax):
+    # show Beta function for different parameters
+    x = np.linspace(0, 1, 100)
+    ax.plot(x, stats.beta.pdf(x, 0.1, 0.1), label=r"$s$=1, $Beta (0.1, 0.1)$")
+    ax.plot(x, stats.beta.pdf(x, 1, 1), label=r"$s$=0, $Beta (1, 1)$")
+    ax.plot(x, stats.beta.pdf(x, 10, 10), label=r"$s$=-1, $Beta (10,10)$")
+    ax.set_xlabel("x")
+    ax.set_ylabel("Probability density")
+    ax.legend()
+    return ax
+
+
+def P_true_evidence_beta(
+    sampsize, selectivity, eta=10, nboot=int(1e4), pisamp=None, theta_lsamp=None
+):
+    s = (
+        10**selectivity
+    )  # selectivity is log such that -1 is the center case, 0 is the neutral case and +1 is the border case
+    theta = rng.beta(1 / s, 1 / s, sampsize)
+
+    if pisamp is None:
+        pisamp = rng.uniform(0, 1, nboot)
+    elif np.isscalar(pisamp):
+        # try a fixed value for pi
+        pisamp = np.full(nboot, pisamp)
+    else:
+        # seems we're dealing with an array
+        pisamp = np.full((nboot, pisamp.shape[1]), pisamp)
+
+    if theta_lsamp is None:
+        # random threshold
+        theta_lsamp = rng.uniform(0, 1, nboot)
+    elif np.isscalar(theta_lsamp):
+        # try a fixed threshold
+        theta_lsamp = np.full(nboot, theta_lsamp)
+    else:
+        # seems we're dealing with an array
+        theta_lsamp = np.full((nboot, theta_lsamp.shape[1]), theta_lsamp)
+
+    kh1samp = sample_H1(theta, pisamp, theta_lsamp)
+    kh0samp = sample_H0(theta, pisamp, theta_lsamp)
+
+    bfh1_kh1 = BFH1(
+        kh1samp, theta, pisamp, theta_lsamp
[…]

### 2023-12-16 src/tex/ms.tex
add discussion, conclusion for sampling selectivity

@@
 
-\subsubsection{Semi-analytical analysis}
+\subsubsection{Semi-analytical analysis}\label{sec:met-semianalytical}
 First, we apply a Bayes Factor Design Analysis~\citep{Schoenbrodt2018} to assess the expected probabilities of obtaining true negative or true positive evidence for the hypothesis above, as well as the probability for misleading or inconclusive evidence.
@@
 \subsubsection{Semi-analytical approach}
-In Section~\ref{sec:met-semianalytical} We computed the probability for true positive evidence for $H_\mathrm{3}$ and $H_\mathrm{3, null}$ respectively.
+In Section~\ref{sec:met-semianalytical} we computed the probability for true positive evidence for $H_\mathrm{3}$ and $H_\mathrm{3, null}$ respectively.
 Figure~\ref{fig:semian_true_evidence} shows how these evidences are distributed for sample sizes \var{semian_Nsamp1} and \var{semian_Nsamp2}, and how likely we are to obtain strong evidence ($BF_{H_i, H_j}$ > 10).
@@
 \todo[inline]{biosignature on M~dwarf planet vs. FGK: Impact on UV flux requirement?}
+\subsubsection{Sampling strategy for testing a predicted minimum \gls{NUV} flux}
+We show in Sect.~\ref{sec:results_uv} that the constraining power for testing the hypothesis of a minimum past \gls{NUV} flux required for abiogenesis is sensitive on the occurrence of life, the value of this threshold flux, the sample size, and the distribution of sampled past \gls{NUV} fluxes.
+In particular the last parameter offers an opportunity to optimize the survey strategy: Although constraints on a planet's UV history have generally large uncertainties~\citep[e.g.,][]{Richey-Yowell2023}, the likely maximum flux of a planet can be inferred and used as a proxy.
+Sampling more extreme (low and high) values for the maximum flux increases the probability of obtaining true strong evidence for the hypothesis.
+For large sample sizes $\gtrsim 200$, this strategy can push this probability into the \SI{90}{\percent} range.
+\todo[inline]{and what about the probability of getting true strong evidence for H0?}
 
@@
         A direct imaging mission with the \gls{HWO} will require characterizing XXX planets ...
+    \item If the predicted \gls{NUV} correlation exists, yielding strong evidence for it is likely ($> \SI{90}{\percent}$) for sample sizes $\geq 100$, and a survey strategy that targets extreme values of inferred past \gls{NUV} irradiation.
     \item Abiogenesis in the cyanosulfidic scenario strongly prefers late host star spectral types in a sample of transiting planets.

### 2023-12-16 src/tex/ms.tex
add 3-param evidence grid

@@
 
+The expected resulting evidence further depends on the a priori unknown abiogenesis rate $f_\mathrm{life}$ and on the \gls{NUV} flux threshold.
+\begin{figure*}
+    \script{semian_true_evidence.py}
+    \begin{centering}
+        \includegraphics[width=\hsize]{figures/semian_evidence-grid.pdf}
+        \caption{Probability of obtaining true strong evidence for different abiogenesis rates, \gls{NUV} flux thresholds, and sample sizes.
+For each of these parameters, higher values increase the probability of yielding strong evidence.}
+        \label{fig:semian_evidence-grid}
+    \end{centering}
+\end{figure*}
+Figure~\ref{fig:semian_evidence-grid} illustrates this dependency: For very low values of either parameter, samples drawn under the null or alternative hypotheses are indistinguishable and the Bayesian evidence is always low.
+Both higher $f_\mathrm{life}$ and higher \gls{NUV} flux thresholds increase the probability of obtaining strong evidence.
+Larger sample sizes enable this at lower values of these parameters.
+
 So far, we have assumed random, uniform distributions of $f_\mathrm{life}$, $F_\mathrm{NUV, min}$, and $F_\mathrm{NUV}$.

### 2023-12-16 src/scripts/semian_true_evidence.py

@@ -177,6 +177,51 @@ def plot_evidence_sampsize(ax, s_max=500):
 
     return ax
 
+def get_meshplot(smooth_sigma, sampsize, ax):
+    x = np.linspace(0.0, 1.0, 10)  # theta_l
+    y = np.linspace(0.0, 1.0, 10)  # pi
+    z = np.zeros((len(x), len(y)))
+    for i, xi in enumerate(x):
+        for j, yj in enumerate(y):
+            z[i, j] = P_true_evidence(
+                sampsize=sampsize, eta=10, nboot=int(1e5), theta_lsamp=xi, pisamp=yj
+            )
+
+    # Smooth the data
+    if smooth_sigma:
+        from scipy.ndimage import gaussian_filter
+
+        z = gaussian_filter(z, smooth_sigma)
+
+    x, y = np.meshgrid(x, y, indexing="ij")
+    im = ax.pcolormesh(
+        x, y, z, cmap=cmocean.cm.ice
+    )  # , vmin=vmin, vmax=vmax, cmap=cmap, lw=0, rasterized=True, shading='auto', edgecolors='k', linewidths=4)
+
+    # ax.set_xlabel('$\\theta_{\lambda}$')
+    ax.set_xlabel("NUV flux threshold (arb. units)")
+    ax.set_ylabel("$f_\mathrm{life}$")
+
+    return ax, im
+
+
+def plot_evidence_grid():
+    steps = 4
+    sampsizes = np.rint(np.geomspace(10, 500, steps) / 10).astype(np.int64) * 10
+    smooth_sigma = None
+
+    fig, axs = plt.subplots(1, steps, figsize=[13, 3.3], sharey=True)
+
+    for i, sampsize in enumerate(sampsizes):
+        axs[i], im = get_meshplot(smooth_sigma, sampsize, axs[i])
+        axs[i].text(0.8, 0.95, "$n = {}$".format(sampsize))
+
+    cbar_ax = fig.add_axes([1.015, 0.18, 0.02, 0.775])
+    cbar = fig.colorbar(im, label="P(true strong evidence)", cax=cbar_ax)
+    fig.tight_layout()
+
+    return fig, axs
+
 
 def plot_beta(ax):
     # show Beta function for different parameters
@@ -280,6 +325,9 @@ axs[1] = plot_evidence_sampsize(axs[1])
 # fig.tight_layout(
[…]

### 2024-03-01 src/tex/ms.tex
argue for O2

@@
 A commonly discussed biosignature is molecular Oxygen (O$_2$), which on Earth emerged as a byproduct of photosynthesis during the Proterozoic era.
-While not the only atmospheric species discussed as a signature for planetary life CITE, we focused on O$_2$ as a representative biosignature for our spectroscopic survey simulations.
+No individual component of an atmosphere has been identified as a reliable biosignature in isolation~\citep{Seager2016}, and the detection of Oxygen alone will certainly not be sufficient to confirm the presence of life.
+Nevertheless, we focus here on detecting this key absorber as it represents the general inherent observational challenges and trends.
 

### 2024-03-06 src/scripts/bioverse_pipeline.py
fix hypothesis grid
still sensitive on number of planets. n_inhabited cannot get zero, otherwise grid gets stuck!

@@ -20,6 +20,7 @@ from bioverse.hypothesis import Hypothesis
 
 # from bioverse.constants import CONST, DATA_DIR
 
+
 def eec_filter(d):
     """Filter the planets to only include EECs."""
     return d[d["EEC"].astype(bool)]
@@ -48,7 +49,6 @@ def inject_nuv_life_correlation(d, f_life=0.1):
         The table of planets with the injected correlation
         and including a new column `inhabited`.
     """
-
 
     d["inhabited"] = np.logical_and(
         d["hz_and_uv"], np.random.random(size=len(d)) <= f_life
@@ -192,9 +192,9 @@ def survey_nautilus(sample, t_total=10 * 365.25):
     # compute yield, conduct survey
     detected = nautilus.compute_yield(sample)
     save_var_latex("N_nautilus", len(detected))
-    data = nautilus.observe(detected) # commented out for now because of Bioverse's issue #45: , t_total=t_total)
-
-
+    data = nautilus.observe(
+        detected
+    )  # commented out for now because of Bioverse's issue #45: , t_total=t_total)
 
     # print(data['max_nuv'][:10])
 
@@ -204,13 +204,15 @@ def survey_nautilus(sample, t_total=10 * 365.25):
 # hypothesis tests
 def h1(theta, X):
     f_life, NUV_thresh = theta
-    return  f_life*(X >=  NUV_thresh)
+    return f_life * (X >= NUV_thresh)
+
 
 def h_null(theta, X):
     shape = (np.shape(X)[0], 1)
     return np.full(shape, theta)
 
-def hypothesis_test(data, method='dynesty'):
+
+def hypothesis_test(data, method="dynesty"):
     """Perform a single hypothesis test on the data."""
     params = ("f_life", "NUV_thresh")
     log = (True, True)
@@ -240,7 +242,8 @@ def hypothesis_test(data, method='dynesty'):
     )
     return results
 
-def hypotest_grid(generator, survey, N_grid=4):
+
+def hypotest_grid(generator, survey, N_grid):
     params = ("f_life", "NUV_thresh")
     log = (True, True)
     bounds = np.array([[1e-3, 1.0], [10.0, 1e5]])
@@ -260,10 +263,18 @@ def hypotest_grid(generator, survey, N_grid=4):
         log=(True,),
     )
 
-    f_life = np.logspace(-3, 0, N_grid)
[…]

### 2024-03-06 src/tex/ms.tex
add biosignature distribution figure

@@
 \subsubsection{Habitable Worlds Observatory}
-\subsubsection{target list}
-\todo[inline]{Provisional stellar target List for the habitable Worlds Observatory:~\citep{Mamajek2023}} % read like in https://github.com/mkenworthy/HWObows/tree/40b1e5f2bc2bbbe8c999b097c80ca5a8cda0aeae/src/scripts
+%\subsubsubsection{target list}
+%\todo[inline]{Provisional stellar target List for the habitable Worlds Observatory:~\citep{Mamajek2023}} % read like in https://github.com/mkenworthy/HWObows/tree/40b1e5f2bc2bbbe8c999b097c80ca5a8cda0aeae/src/scripts
 
@@
 
+A representative recovery of the injected biosignature pattern is shown in Figure~\ref{fig:detections_uv}.
+There, we assumed an abiogenesis rate of $f_\mathrm{life} = \var{f_life}$ and a minimum \gls{NUV} flux of $F_\mathrm{NUV, min} = \var{NUV_thresh}\SI{}{\photons\per\centi\meter\squared\per\second\per\nano\meter}$.
+This leads to a distribution of biosignature detections with detections increasingly occurring above a threshold inferred NUV flux.
+\begin{figure}
+    \script{detections_uv.py}
+    \begin{centering}
+        \includegraphics[width=\hsize]{figures/detections_uv.pdf}
+        \caption{Recovered biosignature detections in the \gls{NUV} flux -- biosignature occurrence space. The dashed line denotes the threshold \gls{NUV} flux $F_\mathrm{NUV, min}$.}
+        \label{fig:detections_uv}
+    \end{centering}
+\end{figure}
+
 \section{Discussion}

### 2024-03-06 src/scripts/bioverse_pipeline.py

@@ -267,9 +267,8 @@ def hypotest_grid(generator, survey, N_grid):
     f_life = np.geomspace(0.1, 1., N_grid)
     # f_life = (0.9,)  # test 1D hypothesis grid test
     # f_life = 0.99  # test 1D hypothesis grid test
-    # NUV_thresh = np.logspace(1, 3, N_grid)
     # NUV_thresh = np.logspace(1, 2.5, N_grid)
-    NUV_thresh = np.array([300., 380.])
+    NUV_thresh = np.geomspace(300., 380., N_grid)
 
 
 
@@ -339,6 +338,7 @@ def past_uv(grid=True, N_grid=4, **kwargs):
     save_var_latex("M_G_max", g_args["M_G_max"])
 
     save_var_latex("f_life", params_past_uv["f_life"])
+    save_var_latex("NUV_thresh", params_past_uv["NUV_thresh"])
     save_var_latex("deltaT_min", int(params_past_uv["deltaT_min"]))
     save_var_latex("uv_inhabited", len(dd[dd.inhabited]))
 

### 2024-03-06 src/scripts/detections_uv.py

@@ -1,0 +1,29 @@
+import pickle
+import paths
+import plotstyle
+plotstyle.styleplots()
+
+import matplotlib.pyplot as plt
+
+def plot_detections_uv(eec):
+    fig, ax = plt.subplots(figsize=(15, 1.5))
+    ax.scatter(eec['max_nuv'], eec['has_O2'], s=9.)
+    ax.set_yticks([0, 1])
+    # ax.set_xlim([0,10])
+    ax.set_yticklabels(['no biosignature', 'biosignature'],fontsize=24)
+    ax.set_xlabel('max. NUV irradiance [erg/s/$cm^2$]',fontsize=24)
+
+    ax.spines['right'].set_visible(False)
+    ax.spines['top'].set_visible(False)
+    # To turn off the bottom or left
+    #ax.spines['bottom'].set_visible(False)
+    ax.spines['left'].set_visible(False)
+    return fig, ax
+
+
+with open(paths.data / 'pipeline/sample.pkl', 'rb') as f:
+    sample = pickle.load(f)
+
+fig, ax = plot_detections_uv(sample)
+
+fig.savefig(paths.figures / "detections_uv.pdf")

### 2024-03-06 src/tex/ms.tex
include powergrid figure

@@
 
+We now investigate the sensitivity of the achieved statistical power of our default transit survey to the a priori unconstrained threshold \gls{NUV} flux $F_\mathrm{NUV, min}$ and the abiogenesis rate $f_\mathrm{life}$.
+Figure~\ref{fig:powergrid} shows the statistical power as a function of these parameters for a sample size of $N=\var{N_nautilus}$.
+Higher values of $F_\mathrm{NUV, min}$ increase the achieved statistical power of the survey, as the dataset under the alternative hypothesis $H_3$ differs more from the null hypothesis with increasing threshold flux.
+The same is true for the abiogenesis rate $f_\mathrm{life}$, where higher values increase the evidence for $H_3$.
+
+\begin{figure}
+    \script{powergrid.py}
+    \begin{centering}
+        \includegraphics[width=\hsize]{figures/powergrid.pdf}
+        \caption{Statistical power as a function of threshold \gls{NUV} flux and abiogenesis rate.
+For a given sample size (here: $N=\var{N_nautilus}$), the achieved statistical power of the survey is enhanced for higher values of $F_\mathrm{NUV, min}$ and $f_\mathrm{life}$.}
+        \label{fig:powergrid}
+    \end{centering}
+\end{figure}
+
 \section{Discussion}

### 2024-03-06 src/scripts/powergrid.py

@@ -1,0 +1,30 @@
+import dill
+import paths
+import plotstyle
+import matplotlib.pyplot as plt
+import cmocean
+from bioverse import plots
+
+plotstyle.styleplots()
+
+def plot_powergrid(grid):
+    """Plot the power grid."""
+    labels = ('S_thresh', 'f_life')
+
+    fig, ax = plt.subplots()
+    fig, ax = plots.plot_power_grid(grid,
+                                        # method='p',
+                                        method='dlnZ',
+                                        axes=('NUV_thresh', 'f_life'), labels=labels,
+                                        log=(True, False), show=False, fig=fig, ax=ax,
+                                        # zoom_factor=2, smooth_sigma=.5,
+                                        zoom_factor=0, smooth_sigma=.2,
+                                        levels=[50, 95], cmap=cmocean.cm.dense_r)
+    return fig, ax
+
+
+with open(paths.data / 'pipeline/grid_flife_nuv.dll', 'rb') as f:
+    grid = dill.load(f)
+
+fig, ax = plot_powergrid(grid)
+fig.savefig(paths.figures / "powergrid.pdf")

### 2024-03-07 src/tex/ms.tex
correct statement on ideal NUV threshold

@@
 Figure~\ref{fig:powergrid} shows the statistical power as a function of these parameters for a sample size of $N=\var{N_nautilus}$.
-Higher values of $F_\mathrm{NUV, min}$ increase the achieved statistical power of the survey, as the dataset under the alternative hypothesis $H_3$ differs more from the null hypothesis with increasing threshold flux.
+Values of $F_\mathrm{NUV, min}$ that lie between the extrema of the inferred maximum \gls{NUV} flux increase the achieved statistical power of the survey, as in this case the dataset under the alternative hypothesis $H_3$ differs more from the null hypothesis.
 The same is true for the abiogenesis rate $f_\mathrm{life}$, where higher values increase the evidence for $H_3$.
@@
         \caption{Statistical power as a function of threshold \gls{NUV} flux and abiogenesis rate.
-For a given sample size (here: $N=\var{N_nautilus}$), the achieved statistical power of the survey is enhanced for higher values of $F_\mathrm{NUV, min}$ and $f_\mathrm{life}$.}
+For a given sample size (here: $N=\var{N_nautilus}$), the achieved statistical power of the survey is enhanced for higher values of $f_\mathrm{life}$ and intermediate values of $F_\mathrm{NUV, min}$.}
         \label{fig:powergrid}

### 2024-03-07 src/scripts/bioverse_pipeline.py
implement a switch for fast pipeline test runs

@@ -248,7 +248,7 @@ def hypothesis_test(data, method="dynesty"):
     return results
 
 
-def hypotest_grid(generator, survey, N_grid):
+def hypotest_grid(generator, survey, N_grid, fast):
     params = ("f_life", "NUV_thresh")
     log = (True, True)
     bounds = np.array([[1e-3, 1.0], [10.0, 1e5]])
@@ -268,6 +268,10 @@ def hypotest_grid(generator, survey, N_grid):
         log=(True,),
     )
 
+    if fast:
+        N_iter = 2
+    else:
+        N_iter = 6
     # f_life = np.geomspace(0.1, 1.0, N_grid)
     f_life = np.geomspace(0.2, 1.0, N_grid)
     # f_life = (0.9,)  # test 1D hypothesis grid test
@@ -287,15 +291,14 @@ def hypotest_grid(generator, survey, N_grid):
         method="dynesty",
         f_life=f_life,
         NUV_thresh=NUV_thresh,
-        # N=5,
-        N=2,
+        N=N_iter,
         processes=8,
         t_total=10 * 365.25,
     )
     return results
 
 
-def past_uv(grid=True, N_grid=3, **kwargs):
+def past_uv(grid=True, N_grid=None, fast=False, **kwargs):
     """Test the hypothesis that life only originates on planets with a minimum past UV irradiance."""
 
     # default parameters for planet generation
@@ -324,8 +327,14 @@ def past_uv(grid=True, N_grid=3, **kwargs):
 
     if grid:
         # perform a grid of hypothesis tests
+        if fast:
+            N_grid = 2
+        elif N_grid:
+            N_grid = N_grid
+        else:
+            N_grid = 6
         nautilus = create_survey_nautilus()
-        grid = hypotest_grid(g, nautilus, N_grid=N_grid)
+        grid = hypotest_grid(g, nautilus, N_grid=N_grid, fast=fast)
         detected = None
         data = None
     else:
@@ -354,12 +363,12 @@ def past_uv(grid=True, N_grid=3, **kwargs):
     return d, grid, detected, data, nautilus
 
 
-def main():
+def main(fast=True):
     """Run the Bioverse pipeline."""
     print("RUNNING BIOVERSE PIPELINE")
 
     d, _grid, detected, data, nautilus = past_uv(grid=False)
-    _d, grid, _detected, _data, _nautilus = past_uv()
+    _d, grid, _detected, _data, _nautilus = past_uv(fast=fast)
 
 
     # save Bioverse objects

### 2024-03-07 src/scripts/bioverse_pipeline.py
fast mode=False

@@ -383,7 +383,7 @@ def main(fast=True):
 
 
 if __name__ == "__main__":
-    # result = timeit.timeit("main()", number=1)
-    result = main()
+    """Run the Bioverse pipeline, either doing the full grid or fast mode."""
+    result = main(fast=False)
 
     # wait = input("PRESS ENTER TO CONTINUE.")

### 2024-05-16 src/tex/ms.tex
fix evidence grid plot text insets

@@
 \end{figure*}
-The situation improves for larger samples -- \SI{80}{\percent} of random experiments yield true strong evidence under either $H_\mathrm{3}$ or $H_\mathrm{3, null}$.
+The situation improves with larger samples: for $n = \var{semian_Nsamp2}$, \SI{80}{\percent} of random experiments yield true strong evidence under either $H_\mathrm{3}$ or $H_\mathrm{3, null}$.
 

### 2024-05-16 src/scripts/semian_true_evidence.py

@@ -214,7 +214,7 @@ def plot_evidence_grid():
 
     for i, sampsize in enumerate(sampsizes):
         axs[i], im = get_meshplot(smooth_sigma, sampsize, axs[i])
-        axs[i].text(0.8, 0.95, "$n = {}$".format(sampsize))
+        axs[i].text(0.99, 0.95, "$n = {}$".format(sampsize), horizontalalignment='right')
 
     cbar_ax = fig.add_axes([1.015, 0.18, 0.02, 0.775])
     cbar = fig.colorbar(im, label="P(true strong evidence)", cax=cbar_ax)

### 2024-05-28 src/scripts/bioverse_pipeline.py
add comment

@@ -64,7 +64,7 @@ def inject_biosignature(d, P=1.0):
 
 
 def get_generator_args():
-    """define generator parameters"""
+    """define default generator parameters. These may be overwritten by other functions"""
     # Parameters for star generation
     stars_args = {
         "d_max": 30,  # max. dist to stars (pc)

### 2024-05-28 src/tex/ms.tex
specify distributions in caption
Fig. -> Figure

@@
 
-To determine \gls{HZ} occupancy, we interpolated the stellar luminosity evolution grid of \citet{Baraffe1998} using a Clough Tocher interpolant~\citep[][see left panel of Fig.~\ref{fig:hz_nuv_evo}]{Nielson1983,Alfeld1984} to compute the evolution of the inner (runaway greenhouse) and outer (maximum greenhouse) edges as a function of planet mass and stellar spectral type~\citep{Kopparapu2014}.
+To determine \gls{HZ} occupancy, we interpolated the stellar luminosity evolution grid of \citet{Baraffe1998} using a Clough Tocher interpolant~\citep[][see left panel of Figure~\ref{fig:hz_nuv_evo}]{Nielson1983,Alfeld1984} to compute the evolution of the inner (runaway greenhouse) and outer (maximum greenhouse) edges as a function of planet mass and stellar spectral type~\citep{Kopparapu2014}.
 This provides each planet's epochs within and outside the \gls{HZ}.
@@
 We linearly interpolate in their measured grid, where we convert spectral type to stellar mass using the midpoints of their mass ranges (\SI{0.75}{\Msun} for K~stars, \SI{0.475}{\Msun} for early-type M~stars,  and \SI{0.215}{\Msun} for late-type M~stars).
-Outside of the age and stellar mass range covered in \citet{Richey-Yowell2023}, we extrapolate using nearest simplex (see right panel of Fig.~\ref{fig:hz_nuv_evo}).
+Outside of the age and stellar mass range covered in \citet{Richey-Yowell2023}, we extrapolate using nearest simplex (see right panel of Figure~\ref{fig:hz_nuv_evo}).
 
@@
 Left: Sampling distribution for different selectivity parameters $s$.
-Right: Resulting \mbox{P(true strong evidence)}.
+Right: Resulting \mbox{P(true strong evidence)} (random, uniform $f_\mathrm{life}$, $F_\mathrm{NUV, min}$).
 Sampling more extreme values of $F_\mathrm{NUV}$ is more likely to yield strong evidence.}
@@
 
-In a magnitude- and volume-limited sample of a transit survey, the host star distribution will be skewed toward later spectral types and dominated by M~dwarfs (see Fig.~\ref{fig:inhabited_FGKM}).
+In a magnitude- and volume-limited sample of a transit survey, the host star distribution will be skewed toward later spectral types and dominated by M~dwarfs (see Figure~\ref{fig:inhabited_FGKM}).
 Due to how the \gls{HZ} scales with spectral type, by far most transiting \glspl{EEC} occur around M~dwarfs.
-These host stars, in particular late subtypes, also provide extended periods of increased \gls{NUV} emission that overlap with times when some of these planets occupy the \gls{HZ} (see Fig.~\ref{fig:hz_nuv_evo}), our requirement for abiogenesis (compare Equation~\ref{eq:hypothesis-uv}).
+These host stars, in particular late subtypes, also provide extended periods of increased \gls{NUV} emission that overlap with times when some of these planets occupy the \gls{HZ} (see Figure~\ref{fig:hz_nuv_evo}), our requirement for abiogenesis (compare Equation~\ref{eq:hypothesis-uv}).
 Because of that, all inhabited planets orbit M~dwarfs.

### 2024-05-28 src/tex/ms.tex
combine individual missions
adapt intro

@@
 
-\section{Origins of Life Scenarios and their Predictions}
+\subsection{Origins of Life Scenarios and their Predictions}
 \todo[inline]{Present widely discussed OOL scenarios and their predictions on exoplanet observables; derive testable hypotheses.}
@@
 
-\subsection{Subaerial ponds}
+%\subsection{Subaerial ponds}
+A different scenario is the emergence of life in hot springs or ponds that are exposed to the planet's atmosphere~\citep[e.g.,][]{Deamer2019}.
 ...
@@
 
-\subsection{UV flux}\label{sec:predictions:uv}
+%\subsection{UV flux}\label{sec:predictions:uv}
 A major hypothesis in the origin of life is that UV light played a constructive role in getting life started on Earth (see Ranjan et al.
@@
 
-\subsection{Other Processes related to the Origins of Life}
-\subsubsection{Planetary redox state and evolution}
-The synthesis of prebiotic compounds requires moderately to highly reduced chemical environments (Kitadai \& Maruyama 2018, Benner+2020, Sasselov+2020, Lichtenberg \& Clement 2022).
-...
-Surficial origins of life chemistries are dependent on the redox state of a planet being $\sim$neutral (not too reduced or oxidized) to allow the presence of precursor molecules like HCN.
+%\subsection{Other Processes related to the Origins of Life}
+%\subsubsection{Planetary redox state and evolution}
+%The synthesis of prebiotic compounds requires moderately to highly reduced chemical environments (Kitadai \& Maruyama 2018, Benner+2020, Sasselov+2020, Lichtenberg \& Clement 2022).
+%...
+%Surficial origins of life chemistries are dependent on the redox state of a planet being $\sim$neutral (not too reduced or oxidized) to allow the presence of precursor molecules like HCN.
 The planetary redox state leaves an imprint on its atmospheric composition and thus planet size (very reduced atmospheres are large) and spectral signatures.
 Connected to the cyanosulfidic scenario, the pond scenario, and the impact trigger.
-
-\subsubsection{Impact trigger}
-Iron-rich impactors have been suggested to intermittently provide the reduced environments favored by prebiotic chemistry (e.g., Sekine+2003, Hashimoto+2007, Kuwahara \& Sugita 2015, Genda+2017, Wogan+2023).
-...
-Prebiotic synthesis triggered through reduced impactors that stochastically create transiently reducing or neutral atmospheres requires a certain composition of the impactors, the planet to not be in a magma ocean state (???) (Lichtenberg \& Clement 2022), and, related to this requirement, occurrence of impact events during early planetary evolution.
-Suggested observables are stochastic increases in brightness temperature, transient increases of planet size, and change of planet composition (decreasing with decreasing impact rate, i.e., stellar age).
+%
+%\subsubsection{Impact trigger}
+%Iron-rich impactors have been suggested to intermittently provide the reduced environments favored by prebiotic chemistry (e.g., Sekine+2003, Hashimoto+2007, Kuwahara \& Sugita 2015, Genda+2017, Wogan+2023).
+%...
+%Prebiotic synthesis triggered through reduced impactors that stochastically create transiently reducing or neutral atmospheres requires a certain composition of the impactors, the planet to not be in a magma ocean state (???) (Lichtenberg \& Clement 2022), and, related to this requirement, occurrence of impact events during early planetary evolution.
+%Suggested observables are stochastic increases in brightness temperature, transient increases of planet size, and change of planet composition (decreasing with decreasing impact rate, i.e., stellar age).
 
@@
 
-\subsection{Origins of Life Hypotheses}
+%\subsection{Origins of Life Hypotheses}
 ...
@@
 
-\subsubsection{Habitable Worlds Observatory}
+\subsubsection{Transit survey (e.g., Nautilus)}
+
+\subsubsection{Direct imaging survey (e.g., Habitable Worlds Observatory, LIFE)}
 %\subsubsubsection{target list}
 %\todo[inline]{Provisional stellar target List for the habitable Worlds Observatory:~\citep{Mamajek2023}} % read like in https://github.com/mkenworthy/HWObows/tree/40b1e5f2bc2bbbe8c999b097c80ca5a8cda0aeae/src/scripts
-
-\subsubsection{Nautilus}
-\subsubsection{LIFE} % optional
 

### 2024-05-28 src/tex/ms.tex
adapt figure for only one hypothesis

@@
 
-\subsection{Hydrothermal vents}
-e.g., Russell+2010
+%\subsection{Hydrothermal vents}
+A widely regarded origins-of-life scenario is that abiogenesis happens in hydrothermal vents~\citep[e.g.,][]{Russel2010}.
 ...
@@
         \includegraphics[width=\hsize]{figures/hypotheses.pdf}
-        \caption{Origins of Life scenarios, their predictions on exoplanet observables, and derived population-level hypotheses.}
+        \caption{Population-level hypothesis and null hypothesis on UV irradiance derived from the cyanosulfidic scenario.}
         \label{fig:hypotheses}
@@
 \end{figure*}
-Figure~\ref{fig:hypotheses} lists the hypotheses derived from the predictions of each OOL scenario.
+Figure~\ref{fig:hypotheses} shows the hypothesis and null hypothesis derived from the predictions of the cyanosulfidic scenario.
 ...

### 2024-05-28 src/tex/ms.tex
restructure methods

@@
 \todo[inline]{Briefly introduce Bayesian model comparison, then present the particular hypotheses in their mathematical form.}
-\subsection{Bayesian hypothesis testing with \bioverse}
-
-\subsection{Considered planets and their orbits}
-All Origins of Life scenarios considered here require water as a solvent; we thus consider only rocky planets that may sustain liquid water on their surface.
-Different formulations of habitable zones as regions around a star where a planet with Earth's atmospheric composition can maintain liquid water on its surface exist~\citep[e.g.,][]{MolLous2022,Spinelli2023,Tuchow2023}.
-CITE!
-Ramirez \& Kaltenegger 2017, 2018
-Here, we adopt the popular estimates of \citet{Kasting1993} and \citet{Kopparapu2013,Kopparapu2014} that define a temperate zone between the runaway greenhouse transition CITE!
-and the maximum greenhouse limit CITE!.
-We use the parametrization in \citet{Kopparapu2014} to derive luminosity and planetary mass-dependent edges of the \gls{HZ} $a_\mathrm{inner}$ and $a_\mathrm{outer}$.
-We further follow \citet{Bixel2021} and classify as \glspl{EEC} all planets with radii $0.8 S^{0.25} < R < 1.4 $ that orbit within the above boundaries.
-The lower limit was suggested as a minimum planet size to retain an atmosphere~\citep{Zahnle2017}.
-
-\subsection{Exoplanet survey simulations}
-\todo[inline]{biosignatures: Focus on molecular Oxygen? Excess Methane? \citep[e.g.,][]{Seeburger2023}}
-% detecting methanogenic life: Sandora2020, Wang2018,
-A commonly discussed biosignature is molecular Oxygen (O$_2$), which on Earth emerged as a byproduct of photosynthesis during the Proterozoic era.
-No individual component of an atmosphere has been identified as a reliable biosignature in isolation~\citep{Seager2016}, and the detection of Oxygen alone will certainly not be sufficient to confirm the presence of life.
-Nevertheless, we focus here on detecting this key absorber as it represents the general inherent observational challenges and trends.
-
-\subsubsection{Transit survey (e.g., Nautilus)}
-
-\subsubsection{Direct imaging survey (e.g., Habitable Worlds Observatory, LIFE)}
-%\subsubsubsection{target list}
-%\todo[inline]{Provisional stellar target List for the habitable Worlds Observatory:~\citep{Mamajek2023}} % read like in https://github.com/mkenworthy/HWObows/tree/40b1e5f2bc2bbbe8c999b097c80ca5a8cda0aeae/src/scripts
 
@@
 
-\subsection{H1: Life only originates in hydrothermal vents}
-The hydrothermal vent scenario does not allow oceans deep enough to form an impenetrable layer of high-pressure ice on its floor.
-The resulting allowed parameter space is described by a lower limit on the bulk density. \todo[inline]{+ exclude atmospheric signature for water worlds?}
-...
-
-\subsection{H2: Life only originates in subaerial ponds with wet/dry cycles}
-We parametrize the required exposed land surface in this scenario as a lower limit in bulk density that is higher than for H1.
-Further, the tidal locking timescale of the planet may not be smaller than the age of the system.
-...
-
-\subsection{H3: Life only originates on planets with particular UV irradiance}
+%\subsection{H1: Life only originates in hydrothermal vents}
+%The hydrothermal vent scenario does not allow oceans deep enough to form an impenetrable layer of high-pressure ice on its floor.
+%The resulting allowed parameter space is described by a lower limit on the bulk density. \todo[inline]{+ exclude atmospheric signature for water worlds?}
+%...
+%
+%\subsection{H2: Life only originates in subaerial ponds with wet/dry cycles}
+%We parametrize the required exposed land surface in this scenario as a lower limit in bulk density that is higher than for H1.
+%Further, the tidal locking timescale of the planet may not be smaller than the age of the system.
+%...
+%
+%\subsection{H3: Life only originates on planets with particular UV irradiance}
 \todo[inline]{\citet{Guenther2020} relate U-band energy to bolometric flux.}
@@
 
-\subsubsection{Semi-analytical analysis}\label{sec:met-semianalytical}
+\subsection{Semi-analytical analysis}\label{sec:met-semianalytical}
 First, we apply a Bayes Factor Design Analysis~\citep{Schoenbrodt2018} to assess the expected probabilities of obtaining true negative or true positive evidence for the hypothesis above, as well as the probability for misleading or inconclusive evidence.
@@
 
-\subsubsection{Survey simulations with Bioverse}
+%\subsection{Bayesian hypothesis testing with \bioverse}
+
+\subsection{Considered planets and their orbits}
+All Origins of Life scenarios considered here require water as a solvent; we thus consider only rocky planets that may sustain liquid water on their surface.
+Different formulations of habitable zones as regions around a star where a planet with Earth's atmospheric composition can maintain liquid water on its surface exist~\citep[e.g.,][]{MolLous2022,Spinelli2023,Tuchow2023}.
+CITE!
+Ramirez \& Kaltenegger 2017, 2018
+Here, we adopt the popular estimates of \citet{Kasting1993} and \citet{Kopparapu2013,Kopparapu2014} that define a temperate zone between the runaway greenhouse transition CITE!
+and the maximum greenhouse limit CITE!.
+We use the parametrization in \citet{Kopparapu2014} to derive luminosity and planetary mass-dependent edges of the \gls{HZ} $a_\mathrm{inner}$ and $a_\mathrm{outer}$.
+We further follow \citet{Bixel2021} and classify as \glspl{EEC} all planets with radii $0.8 S^{0.25} < R < 1.4 $ that orbit within the above boundaries.
+The lower limit was suggested as a minimum planet size to retain an atmosphere~\citep{Zahnle2017}.
+
+\todo[inline]{biosignatures: Focus on molecular Oxygen? Excess Methane? \citep[e.g.,][]{Seeburger2023}}
+% detecting methanogenic life: Sandora2020, Wang2018,
+A commonly discussed biosignature is molecular Oxygen (O$_2$), which on Earth emerged as a byproduct of photosynthesis during the Proterozoic era.
+No individual component of an atmosphere has been identified as a reliable biosignature in isolation~\citep{Seager2016}, and the detection of Oxygen alone will certainly not be sufficient to confirm the presence of life.
+Nevertheless, we focus here on detecting this key absorber as it represents the general inherent observational challenges and trends.
+
+\subsection{Exoplanet survey simulations with Bioverse}
 To model the sensitivity of the information gain of a proposed mission to sample selection and survey strategy, we now conduct survey simulations with Bioverse using different sample sizes and survey strategies.
@@
 \end{figure*}
+
+\subsubsection{Transit survey (e.g., Nautilus)}
+
+\subsubsection{Direct imaging survey (e.g., Habitable Worlds Observatory, LIFE)}
+%\subsubsubsection{target list}
+%\todo[inline]{Provisional stellar target List for the habitable Worlds Observatory:~\citep{Mamajek2023}} % read like in https://github.com/mkenworthy/HWObows/tree/40b1e5f2bc2bbbe8c999b097c80ca5a8cda0aeae/src/scripts
 

### 2024-05-28 src/tex/ms.tex
change equations to only one hypothesis
All the `H_3` become `H_1`, and all the `H_3, null` becomes simply `H_null`.

@@
 Let our observable be the inferred past \gls{NUV} flux of the planet $F_\mathrm{NUV}$.
-Under Hypothesis $H_3$ (Equation~\ref{eq:hypothesis-uv}), there exists a special unknown value of $F_\mathrm{NUV}$, noted $F_\mathrm{NUV, min}$ such that
+Under Hypothesis $H_1$ (Equation~\ref{eq:hypothesis-uv}), there exists a special unknown value of $F_\mathrm{NUV}$, noted $F_\mathrm{NUV, min}$ such that
 
 \begin{align}
-    P(L|\theta,H_3) &=  f_\mathrm{life} \quad \text{if } \theta>F_\mathrm{NUV, min}\\
-    P(L|\theta,H_3) &=  0               \quad  \text{otherwise}
+    P(L|\theta,H_1) &=  f_\mathrm{life} \quad \text{if } \theta>F_\mathrm{NUV, min}\\
+    P(L|\theta,H_1) &=  0               \quad  \text{otherwise}
 \end{align}
@@
 \begin{equation}
-P(L|\theta,H_\mathrm{3, null}) = f_\mathrm{life}.
+P(L|\theta,H_\mathrm{null}) = f_\mathrm{life}.
 \end{equation}
@@
 
-If we define $k=\sum L_i$ and denote $Y$ the random variable that describes it, $H_\mathrm{3, null}$ represents the likelihood that the number of planets with life in the sample follows the binomial distribution
+If we define $k=\sum L_i$ and denote $Y$ the random variable that describes it, $H_\mathrm{null}$ represents the likelihood that the number of planets with life in the sample follows the binomial distribution
 \begin{equation}
-P(Y=k|H_\mathrm{3, null}) = \binom{n}{k}f_\mathrm{life}^k(1-f_\mathrm{life})^{n-k}.
+P(Y=k|H_\mathrm{null}) = \binom{n}{k}f_\mathrm{life}^k(1-f_\mathrm{life})^{n-k}.
 \end{equation}
 
-Under $H_3$, $Y$ also follows a binomial distribution, however it is conditioned by $n_{\lambda}=Card(\{F_\mathrm{NUV, i} \text{ if } F_\mathrm{NUV, i}>F_\mathrm{NUV, min}\}_{i \in [1,n]})$ the number of values of $F_\mathrm{NUV}$ in the experiment that exceed $F_\mathrm{NUV, min}$
+Under $H_1$, $Y$ also follows a binomial distribution, however it is conditioned by $n_{\lambda}=Card(\{F_\mathrm{NUV, i} \text{ if } F_\mathrm{NUV, i}>F_\mathrm{NUV, min}\}_{i \in [1,n]})$ the number of values of $F_\mathrm{NUV}$ in the experiment that exceed $F_\mathrm{NUV, min}$
 \begin{equation}
-P(Y=k|H_3) = \binom{n_{\lambda}}{k}f_\mathrm{life}^k(1-f_\mathrm{life})^{n_{\lambda}-k}.
+P(Y=k|H_1) = \binom{n_{\lambda}}{k}f_\mathrm{life}^k(1-f_\mathrm{life})^{n_{\lambda}-k}.
 \end{equation}
@@
 \begin{equation}\label{eq:bayes_factor}
-BF_{H_3,H_\mathrm{3, null}} = \frac{P(Y=k|H_3)}{P(Y=k|H_\mathrm{3, null})} = \frac{\binom{n_\lambda}{k}}{\binom{n}{k}}(1-f_\mathrm{life})^{n_{\lambda}-n}.
+BF_{H_1,H_\mathrm{null}} = \frac{P(Y=k|H_1)}{P(Y=k|H_\mathrm{null})} = \frac{\binom{n_\lambda}{k}}{\binom{n}{k}}(1-f_\mathrm{life})^{n_{\lambda}-n}.
 \end{equation}
 
-Given a sample of planets, where for some of them we have convincing biosignature detections but remaining agnostic on $f_\mathrm{life}$: What evidence for $H_\mathrm{3}$ and $H_\mathrm{3, null}$ can we expect to get?
+Given a sample of planets, where for some of them we have convincing biosignature detections but remaining agnostic on $f_\mathrm{life}$: What evidence for $H_\mathrm{1}$ and $H_\mathrm{null}$ can we expect to get?
 Our Bayes factor (Equation~\ref{eq:bayes_factor}) is determined by the unknown variables $f_\mathrm{life}$ and $F_\mathrm{NUV, min}$, as well as the number of planets with biosignature detections in the sample $k$.
-To compute the distribution of evidences, we repeatedly generated samples under $H_\mathrm{3}$ and $H_\mathrm{3, null}$ and computed the Bayes factors $BF_{H_3,H_\mathrm{3, null}}$ and $BF_{H_\mathrm{3, null}, H_3}$.
+To compute the distribution of evidences, we repeatedly generated samples under $H_\mathrm{1}$ and $H_\mathrm{null}$ and computed the Bayes factors $BF_{H_1,H_\mathrm{null}}$ and $BF_{H_\mathrm{null}, H_1}$.
 We then evaluated the fraction of Monte Carlo runs in which certain evidence thresholds~\citep{Jeffreys1939} were exceeded.
@@
 \begin{equation}\label{eq:hypothesis-uv}
-    H_{3} = f_\mathrm{life} (\theta, F_\mathrm{NUV}) =
+    H_{1} = f_\mathrm{life} (\theta, F_\mathrm{NUV}) =
         \begin{cases}
@@
 %\begin{equation}
-$H_{3, null} = f_\mathrm{life} (\theta),$
+$H_{null} = f_\mathrm{life} (\theta),$
 %\end{equation}
@@
 \subsubsection{Semi-analytical approach}
-In Section~\ref{sec:met-semianalytical} we computed the probability for true positive evidence for $H_\mathrm{3}$ and $H_\mathrm{3, null}$ respectively.
+In Section~\ref{sec:met-semianalytical} we computed the probability for true positive evidence for $H_\mathrm{1}$ and $H_\mathrm{null}$ respectively.
 Figure~\ref{fig:semian_true_evidence} shows how these evidences are distributed for sample sizes \var{semian_Nsamp1} and \var{semian_Nsamp2}, and how likely we are to obtain strong evidence ($BF_{H_i, H_j}$ > 10).
-For $n = \var{semian_Nsamp1}$, strong true evidence for $H_\mathrm{3}$ ($H_\mathrm{3, null}$) can be expected in $\sim \SI{30}{\percent}$ ($\sim \SI{40}{\percent}$) of all random experiments.
+For $n = \var{semian_Nsamp1}$, strong true evidence for $H_\mathrm{1}$ ($H_\mathrm{null}$) can be expected in $\sim \SI{30}{\percent}$ ($\sim \SI{40}{\percent}$) of all random experiments.
 In the majority of cases, the outcome of the survey will be inconclusive.
@@
         \caption{Probability to obtain true strong evidence.
-Left: evidence levels for $H_\mathrm{3}$ and $H_\mathrm{3, null}$ under sample sizes $n = \var{semian_Nsamp1}$ (solid) and $n = \var{semian_Nsamp2}$ (dashed).
+Left: evidence levels for $H_\mathrm{1}$ and $H_\mathrm{null}$ under sample sizes $n = \var{semian_Nsamp1}$ (solid) and $n = \var{semian_Nsamp2}$ (dashed).
 The vertical lines denote the thresholds for ``strong'' evidence, $BF_{H_i, H_j}$ > 10, and ``extreme'' evidence, $BF_{H_i, H_j}$ > 100.
-Right: Probability of true strong evidence for $H_\mathrm{3}$ or $H_\mathrm{3, null}$, whichever is smaller, as a function of sample size $n$.}
+Right: Probability of true strong evidence for $H_\mathrm{1}$ or $H_\mathrm{null}$, whichever is smaller, as a function of sample size $n$.}
         \label{fig:semian_true_evidence}
@@
 \end{figure*}
-The situation improves with larger samples: for $n = \var{semian_Nsamp2}$, \SI{80}{\percent} of random experiments yield true strong evidence under either $H_\mathrm{3}$ or $H_\mathrm{3, null}$.
+The situation improves with larger samples: for $n = \var{semian_Nsamp2}$, \SI{80}{\percent} of random experiments yield true strong evidence under either $H_\mathrm{1}$ or $H_\mathrm{null}$.
 
@@
 A high biosignature detection rate $f_\mathrm{life}$ increases the evidence (cmp. Equation~\ref{eq:bayes_factor}) but we cannot influence it.
-The same is true for $F_\mathrm{NUV, min}$, where again higher values increase the evidence as the binomial distribution for $H_\mathrm{3}$ gets increasingly skewed and shifted away from the one for $H_\mathrm{3, null}$.
+The same is true for $F_\mathrm{NUV, min}$, where again higher values increase the evidence as the binomial distribution for $H_\mathrm{1}$ gets increasingly skewed and shifted away from the one for $H_\mathrm{null}$.
 The distribution of $F_\mathrm{NUV}$ in the planet sample, on the other hand, can be influenced by the survey strategy.
 A targeted sampling approach could be to favor extreme values of $F_\mathrm{NUV}$ in the sample selection.
-Figure~\ref{fig:semian_selectivity} shows how the probability of obtaining true strong evidence for $H_\mathrm{3}$ scales with selectivity $s$, where $s\in]-1,1[$ such that $F_\mathrm{NUV} \sim Beta(1/10^s,1/10^s)$.
+Figure~\ref{fig:semian_selectivity} shows how the probability of obtaining true strong evidence for $H_\mathrm{1}$ scales with selectivity $s$, where $s\in]-1,1[$ such that $F_\mathrm{NUV} \sim Beta(1/10^s,1/10^s)$.
 Here, $s=0$ corresponds to a random uniform distribution.
@@
 Figure~\ref{fig:powergrid} shows the statistical power as a function of these parameters for a sample size of $N=\var{N_nautilus}$.
-Values of $F_\mathrm{NUV, min}$ that lie between the extrema of the inferred maximum \gls{NUV} flux increase the achieved statistical power of the survey, as in this case the dataset under the alternative hypothesis $H_3$ differs more from the null hypothesis.
-The same is true for the abiogenesis rate $f_\mathrm{life}$, where higher values increase the evidence for $H_3$.
+Values of $F_\mathrm{NUV, min}$ that lie between the extrema of the inferred maximum \gls{NUV} flux increase the achieved statistical power of the survey, as in this case the dataset under the alternative hypothesis $H_1$ differs more from the null hypothesis.
+The same is true for the abiogenesis rate $f_\mathrm{life}$, where higher values increase the evidence for $H_1$.
 

### 2024-05-28 src/tex/ms.tex
fix file ending problem
fix most detected problems

@@
+%! suppress = GroupedSubSupScript
+%! suppress = Ellipsis
 % Define document class
@@
 \input{preamble}
+%! suppress = LineBreak
 \begin{document}
@@
 
-\subsection{Origins of Life Scenarios and their Predictions}
+\subsection{Origins of Life Scenarios and their Predictions}\label{sec:predictions}
 \todo[inline]{Present widely discussed OOL scenarios and their predictions on exoplanet observables; derive testable hypotheses.}
@@
 
-\textbf{Prediction:} Biosignatures occur outside the tidal locking zone and at bulk densities consistent with exposed rock..
+\textbf{Prediction:} Biosignatures occur outside the tidal locking zone and at bulk densities consistent with exposed rock.
 \todo[inline]{SR: I'm not a priori sold that tidal locking means that no wet-dry cycles occur.
@@
     \begin{centering}
-        \includegraphics[width=\hsize]{figures/hypotheses.pdf}
+        \includegraphics[width=\hsize]{figures/hypotheses}
         \caption{Population-level hypothesis and null hypothesis on UV irradiance derived from the cyanosulfidic scenario.}
@@
 Test for negative correlations as well?}
-Our third theoretical experiment is to test the UV irradiance requirement (Sect.~\ref{sec:predictions:uv}) by relating the occurrence of life on an exo-earth candidate with a minimum past quiescent stellar UV flux, focusing on the prebiotically interesting \gls{NUV} range from \SIrange{200}{280}{\nano\meter}.
+Our third theoretical experiment is to test the UV irradiance requirement (Sect.~\ref{sec:predictions}) by relating the occurrence of life on an exo-earth candidate with a minimum past quiescent stellar UV flux, focusing on the prebiotically interesting \gls{NUV} range from \SIrange{200}{280}{\nano\meter}.
 \todo[inline]{SUKRIT: got a good reference for this?}  %~\citep{RanjanSasselov2016}.?
@@
     \begin{centering}
-        \includegraphics[width=\hsize]{figures/inhabited_FGKM.pdf}
+        \includegraphics[width=\hsize]{figures/inhabited_FGKM}
         \caption{Host stars of all transiting planets, \glspl{EEC}, and inhabited planets.
@@
     \begin{centering}
-        \includegraphics[width=\hsize]{figures/hz_nuv_evo.pdf}
+        \includegraphics[width=\hsize]{figures/hz_nuv_evo}
         \caption{Interpolated stellar luminosity evolution (left) and evolution of the \gls{NUV} flux in the \gls{HZ} as a function of host star mass.
@@
 \label{sec:results}
-\subsection{Information content in mass-density space}
+%\subsection{Information content in mass-density space}
+%\subsection{Information content in tidal locking timescale-density space}
 
-\subsection{Information content in tidal locking timescale-density space}
+%\subsection{Correlation of past UV flux and biosignature occurrence}\label{sec:results_uv}
 
-\subsection{Correlation of past UV flux and biosignature occurrence}\label{sec:results_uv}
-
-\subsubsection{Semi-analytical approach}
+\subsection{Semi-analytical approach}
 In Section~\ref{sec:met-semianalytical} we computed the probability for true positive evidence for $H_\mathrm{1}$ and $H_\mathrm{null}$ respectively.
@@
     \begin{centering}
-        \includegraphics[width=\hsize]{figures/semian_true_evidence.pdf}
+        \includegraphics[width=\hsize]{figures/semian_true_evidence}
         \caption{Probability to obtain true strong evidence.
@@
     \begin{centering}
-        \includegraphics[width=\hsize]{figures/semian_evidence-grid.pdf}
+        \includegraphics[width=\hsize]{figures/semian_evidence-grid}
         \caption{Probability of obtaining true strong evidence for different abiogenesis rates, \gls{NUV} flux thresholds, and sample sizes.
@@
     \begin{centering}
-        \includegraphics[width=\hsize]{figures/semian_selectivity.pdf}
+        \includegraphics[width=\hsize]{figures/semian_selectivity}
         \caption{Scaling of the probability of obtaining true strong evidence with sample selectivity.
@@
 
-\subsubsection{Survey simulations with Bioverse}
+\subsection{Survey simulations with Bioverse}
 
@@
     \begin{centering}
-        \includegraphics[width=\hsize]{figures/detections_uv.pdf}
+        \includegraphics[width=\hsize]{figures/detections_uv}
         \caption{Recovered biosignature detections in the \gls{NUV} flux -- biosignature occurrence space. The dashed line denotes the threshold \gls{NUV} flux $F_\mathrm{NUV, min}$.}
@@
     \begin{centering}
-        \includegraphics[width=\hsize]{figures/powergrid.pdf}
+        \includegraphics[width=\hsize]{figures/powergrid}
         \caption{Statistical power as a function of threshold \gls{NUV} flux and abiogenesis rate.
@@
 \subsubsection{Sampling strategy for testing a predicted minimum \gls{NUV} flux}
-We show in Sect.~\ref{sec:results_uv} that the constraining power for testing the hypothesis of a minimum past \gls{NUV} flux required for abiogenesis is sensitive on the occurrence of life, the value of this threshold flux, the sample size, and the distribution of sampled past \gls{NUV} fluxes.
+We show in Sect.~\ref{sec:results} that the constraining power for testing the hypothesis of a minimum past \gls{NUV} flux required for abiogenesis is sensitive on the occurrence of life, the value of this threshold flux, the sample size, and the distribution of sampled past \gls{NUV} fluxes.
 In particular the last parameter offers an opportunity to optimize the survey strategy: Although constraints on a planet's UV history have generally large uncertainties~\citep[e.g.,][]{Richey-Yowell2023}, the likely maximum flux of a planet can be inferred and used as a proxy.
@@
 \todo[inline]{discuss how planetary context may impact credibility of tentative biosignature detections (e.g., if there is a good fit with a predicted OOL pattern; or the opposite: the planetary context does not fit well to any OOL scenarios)}
-As we have shown in Section~\ref{sec:results_uv}, the interplay of \gls{NUV} evolution and \gls{HZ} occupancy strongly favors late spectral types for abiogenesis via the cyanosulfidic scenario.
+As we have shown in Section~\ref{sec:results}, the interplay of \gls{NUV} evolution and \gls{HZ} occupancy strongly favors late spectral types for abiogenesis via the cyanosulfidic scenario.
 This strong predicted correlation between stellar spectral type and the occurrence of life can be used to falsify \gls{ool} scenarios.
@@
 
-\subsubsection{Diurnal cycles}
-\todo[inline]{We assume that tidal locking implies that there are no wet-dry cycles. Challenge this assumption.}
+%\subsubsection{Diurnal cycles}
+%\todo[inline]{We assume that tidal locking implies that there are no wet-dry cycles. Challenge this assumption.}
 
@@
 
+%! suppress = MissingBibliographystyle
 \bibliography{bib,coauthors}

### 2024-05-29 src/tex/ms.tex
extend planet tracks

@@
         \caption{Interpolated stellar luminosity evolution (left) and evolution of the \gls{NUV} flux in the \gls{HZ} as a function of host star mass.
-The scatter plot shows age and host star mass of the transiting planets in the synthetic sample with \gls{EEC} highlighted in green.
-        Two of the three planet trajectories shown have extended overlap of \gls{HZ} occupancy and high \gls{NUV} flux (green sections), our requirement for abiogenesis.}
+The scatter plot shows age and host star mass of the transiting planets in the synthetic sample.% with \gls{EEC} highlighted in green.
+        We show the evolutionary tracks of a few randomly selected planets.
+        Tracks with extended overlap of \gls{HZ} occupancy and high \gls{NUV} flux (green sections) fulfill our requirement for abiogenesis.}
         \label{fig:hz_nuv_evo}
@@
 A representative recovery of the injected biosignature pattern is shown in Figure~\ref{fig:detections_uv}.
-There, we assumed an abiogenesis rate of $f_\mathrm{life} = \var{f_life}$ and a minimum \gls{NUV} flux of $F_\mathrm{NUV, min} = \var{NUV_thresh}\SI{}{\photons\per\centi\meter\squared\per\second\per\nano\meter}$.
+There, we assumed an abiogenesis rate of $f_\mathrm{life} = \var{f_life}$ and a minimum \gls{NUV} flux of $F_\mathrm{NUV, min} = \var{NUV_thresh}\,\SI{}{\photons\per\centi\meter\squared\per\second\per\nano\meter}$.
 This leads to a distribution of biosignature detections with detections increasingly occurring above a threshold inferred NUV flux.

### 2024-05-29 src/scripts/hz_nuv_evo.py

@@ -23,7 +23,7 @@ def plot_interpolation(
         ax.plot(*input_points, "x", ms=1, label="input point")
     if d:
         # for (v, c) in [(False, 'white'), (True, 'C1')]:
-        for c, (EEC, group) in zip(['white', 'C0'], d.to_pandas().groupby('EEC')):
+        for c, (EEC, group) in zip(["white", "C0"], d.to_pandas().groupby("EEC")):
             ax.scatter(
                 group["age"],
                 group["M_st"],
@@ -35,7 +35,7 @@ def plot_interpolation(
                 linewidths=0.6,
                 label=["EEC" if EEC else None][0],
             )
-        ax.legend(loc='upper left')
+        ax.legend(loc="upper left")
     fig.colorbar(q, label=cbarlabel)
     ax.set_xscale("log")
     ax.set_xlabel("Time (Gyr)")
@@ -65,8 +65,12 @@ def plot_hz_and_nuv(fig, ax, sample, NUV_thresh=100.0, N_sample=2, random_state=
     for id in inhabited.sample(N_sample, random_state=random_state).planetID:
         Mst = eecdf[eecdf.planetID == id]["M_st"]
         t = eec.evolution[id]["time"]
-        ax.plot(t, [Mst for tt in t], lw=1, c="gray")
-
+        ax.plot(
+            np.concatenate((np.array([1e-3]), t)),
+            [Mst] + [Mst for tt in t],
+            lw=1,
+            c="gray",
+        )
         in_hz = eec.evolution[id]["in_hz"]
         nuv = eec.evolution[id]["nuv"]
         # check if planet ever was in the HZ and had NUV fluxes above the threshold value
@@ -79,9 +83,10 @@ def plot_hz_and_nuv(fig, ax, sample, NUV_thresh=100.0, N_sample=2, random_state=
     id = eecdf[~eecdf.hz_and_uv].sample(1, random_state=48).planetID.values[0]
     Mst = eecdf[eecdf.planetID == id]["M_st"]
     t = eec.evolution[id]["time"]
-    ax.plot(t, [Mst for tt in t], lw=1, c="gray")
+    ax.plot(np.concatenate((np.array([1e-3]), t)), [Mst] + [Mst for tt in t], lw=1, c="gray")
 
     return ax
+
 
 if __name__ == "__main__":
     with open(paths.data / "pipeline/sample.dll", "rb") as f:

### 2024-05-29 src/scripts/hz_nuv_evo.py
beautify legend

@@ -35,7 +35,7 @@ def plot_interpolation(
                 linewidths=0.6,
                 label=["EEC" if EEC else None][0],
             )
-        ax.legend(loc="upper left")
+        ax.legend(loc="upper left", framealpha=.0)
     fig.colorbar(q, label=cbarlabel)
     ax.set_xscale("log")
     ax.set_xlabel("Time (Gyr)")

### 2024-06-05 src/tex/ms.tex
adapt headings

@@
 
-\subsection{Semi-analytical approach}
+\subsection{Semi-analytical approach}\label{sec:results-semianalytical}
 In Section~\ref{sec:met-semianalytical} we computed the probability for true positive evidence for $H_\mathrm{1}$ and $H_\mathrm{null}$ respectively.
@@
 Due to how the \gls{HZ} scales with spectral type, by far most transiting \glspl{EEC} occur around M~dwarfs.
+Their \gls{NUV} fluxes are generally highest in at early times \SI{\lesssim 100}{\mega\year}.
 These host stars, in particular late subtypes, also provide extended periods of increased \gls{NUV} emission that overlap with times when some of these planets occupy the \gls{HZ} (see Figure~\ref{fig:hz_nuv_evo}), our requirement for abiogenesis (compare Equation~\ref{eq:hypothesis-uv}).
@@
 
+\subsubsection{expected biosignature pattern in a transit survey}
 A representative recovery of the injected biosignature pattern is shown in Figure~\ref{fig:detections_uv}.
@@
 
+\subsubsection{Statistical power for a transit survey and sensitivity on astrophysical parameters}
 We now investigate the sensitivity of the achieved statistical power of our default transit survey to the a priori unconstrained threshold \gls{NUV} flux $F_\mathrm{NUV, min}$ and the abiogenesis rate $f_\mathrm{life}$.

### 2024-06-11 src/scripts/utils.py
save variables with two decimal places

@@ -19,6 +19,10 @@ def save_var_latex(key, value, datafile=paths.tex/'variables.dat'):
     except FileNotFoundError:
         pass
 
+    # Check if the value is a float, if so format it to have a maximum of two decimal places
+    if isinstance(value, float):
+        value = round(value, 2)
+
     dict_var[key] = value
 
     with open(file_path, "w") as f:

### 2024-06-11 src/scripts/nuv_distribution.py
plot max. NUV distribution for generic transit survey

@@ -1,0 +1,56 @@
+import dill
+import paths
+import plotstyle
+from src.scripts.utils import save_var_latex
+
+plotstyle.styleplots()
+
+import matplotlib.pyplot as plt
+
+from scipy import stats
+import numpy as np
+
+
+def plot_nuv_distribution(data):
+    dataa = data.to_pandas()
+    fig, ax = plt.subplots(figsize=(15, 1.5))
+
+    # force-fit a beta distribution to our NUV_max
+    # max_likeli = stats.beta.fit(dataa.max_nuv, method='MLE')
+    max_likeli = stats.beta.fit(dataa.max_nuv, method="MM")
+    print(f"fit parameters: {max_likeli}")
+
+    # estimate selectivity from average of fitted beta function parameters
+    selectivity = np.log10(1 / (np.mean(max_likeli[:2])))
+    save_var_latex("selectivity_transit_volume-lim", round(selectivity, 1))
+    print(f"selectivity ~ {selectivity:.2f}")
+
+    x = np.arange(0.0, 1000.0, 5)
+    fig, ax = plt.subplots()
+    ax.hist(dataa.max_nuv, density=True, color="C0")
+    ax.plot(
+        x,
+        stats.beta.pdf(x, *max_likeli[:2], loc=max_likeli[2], scale=max_likeli[3]),
+        c="0.4",
+    )
+
+    ax.set_xlabel("max. NUV irradiance $F_\mathrm{NUV, max}$ [erg/s/$cm^2$]")
+    ax.set_ylabel("Probability density")
+    ax.text(
+        0.97,
+        0.9,
+        transform=ax.transAxes,
+        s="selectivity s = {:.1f}".format(selectivity),
+        horizontalalignment="right",
+        color="0.2",
+    )
+
+    return fig, ax
+
+
+with open(paths.data / "pipeline/data.dll", "rb") as f:
+    data = dill.load(f)
+
+fig, ax = plot_nuv_distribution(data)
+
+fig.savefig(paths.figures / "nuv_distribution.pdf")

### 2024-06-11 src/tex/ms.tex
include NUV distribution plot

@@
 
+\subsubsection{selectivity of a generic transit survey}
+In Section~\ref{sec:results-semianalytical}, we demonstrated that the probability of obtaining true strong evidence for the hypothesis that life only originates on planets with a minimum past \gls{NUV} flux is sensitive to the distribution of sampled past \gls{NUV} fluxes, i.e., the selectivity of the survey (compare Figure~\ref{fig:semian_selectivity}).
+The maximum \gls{NUV} distribution of our generic transit survey is rather unimodal with a median around \SI{500}{\erg\per\second\per\centi\meter\squared} (see Figure~\ref{fig:nuv_distribution}), leading to a selectivity of roughly \var{selectivity_transit_volume-lim}.
+\begin{figure}
+    \script{nuv_distribution.py}
+    \begin{centering}
+        \includegraphics[width=\hsize]{figures/nuv_distribution}
+        \caption{Distribution of maximum past \gls{NUV} flux in a generic volume- and magnitude-limited transit survey.
+The best-fit beta distribution (gray) corresponds to a selectivity of \var{selectivity_transit_volume-lim}.}
+        \label{fig:nuv_distribution}
+    \end{centering}
+\end{figure}
+
 \subsubsection{expected biosignature pattern in a transit survey}

### 2024-06-11 src/tex/ms.tex
fix units

@@
 A representative recovery of the injected biosignature pattern is shown in Figure~\ref{fig:detections_uv}.
-There, we assumed an abiogenesis rate of $f_\mathrm{life} = \var{f_life}$ and a minimum \gls{NUV} flux of $F_\mathrm{NUV, min} = \var{NUV_thresh}\,\SI{}{\photons\per\centi\meter\squared\per\second\per\nano\meter}$.
+There, we assumed an abiogenesis rate of $f_\mathrm{life} = \var{f_life}$ and a minimum \gls{NUV} flux of $F_\mathrm{NUV, min} = \var{NUV_thresh}\,\SI{}{\erg\per\second\per\centi\meter\squared}$.
 This leads to a distribution of biosignature detections with detections increasingly occurring above a threshold inferred NUV flux.

### 2024-06-13 src/scripts/bioverse_pipeline.py
distinguish between M and FGK

@@ -199,7 +199,9 @@ def run_survey_nautilus(sample, t_total=10 * 365.25):
     # compute yield, conduct survey
     detected = nautilus.compute_yield(sample)
     save_var_latex("N_nautilus", len(detected))
-    data = nautilus.observe(detected)  # commented out for now because of Bioverse's issue #45: , t_total=t_total)
+    data = nautilus.observe(
+        detected
+    )  # commented out for now because of Bioverse's issue #45: , t_total=t_total)
 
     # print(data['max_nuv'][:10])
 
@@ -298,8 +300,36 @@ def hypotest_grid(generator, survey, N_grid, fast):
     return results
 
 
-def past_uv(grid=True, N_grid=None, fast=False, **kwargs):
-    """Test the hypothesis that life only originates on planets with a minimum past UV irradiance."""
+def past_uv(hoststars="all", grid=True, N_grid=None, fast=False, **kwargs):
+    """Test the hypothesis that life only originates on planets with a minimum past UV irradiance.
+
+    Parameters
+    ----------
+    hoststars : str
+        The spectral type of the host stars to consider.
+        Options are 'all', 'FGK', 'M'.
+    grid : bool
+        Whether to perform a grid of hypothesis tests.
+    N_grid : int
+        The number of grid points to use.
+    fast : bool
+        Whether to run a fast grid for testing (less grid points)
+    kwargs : dict
+        Additional parameters to pass to the generator.
+
+    Returns
+    -------
+    d : Table
+        The sample of planets.
+    grid : Table
+        The grid of hypothesis tests.
+    detected : Table
+        The detected planets.
+    data : Table
+        The measurements obtained.
+    nautilus : TransitSurvey object
+        The transit survey.
+    """
 
     # default parameters for planet generation
     params_past_uv = {
@@ -316,7 +346,23 @@ def past_uv(grid=True, N_grid=None, fast=False, **kwargs):
     for key, value in kwargs.items():
         params_past_uv[key] = value
 
-    g, g_args = generate_generator(label=None, **params_past_uv, **kwargs)
+    if hoststars == "FGK":
+        # exclude other spectral types
+        params_past_uv["SpT"] = ["F", "G", "K"]
+        params_past_uv["d_max"] = 125  # Gaia GCNS doesn't go further than 119 pc
[…]

### 2024-06-13 src/scripts/bioverse_pipeline.py
change default threshold

@@ -336,10 +336,11 @@ def past_uv(hoststars="all", grid=True, N_grid=None, fast=False, **kwargs):
         # "d_max": 60,        # TOO SMALL SAMPLE AND THE HYPOTHESIS TESTING GRID GETS STUCK WITHOUT AN ERROR MESSAGE
         "d_max": 75,  # TOO SMALL SAMPLE AND THE HYPOTHESIS TESTING GRID GETS STUCK WITHOUT AN ERROR MESSAGE
         "deltaT_min": 10.0,  # Myr
-        "NUV_thresh": 350.0,  # choose such that n_inhabited can't be zero
+        # "NUV_thresh": 350.0,  # choose such that n_inhabited can't be zero
         # "NUV_thresh": 200.0,
+        "NUV_thresh": 150.0,
         "f_life": 0.8,
-        "f_eta": 5.0,  # Occurrence rate scaling factor (MAKE SURE SAMPLE IS LARGE ENOUGH (see above))
+        # "f_eta": 5.0,  # Occurrence rate scaling factor (MAKE SURE SAMPLE IS LARGE ENOUGH (see above))
     }
 
     # replace parameters with kwargs, if any

### 2024-06-13 src/tex/ms.tex
fix a siunitx issue and a typo

@@
 We linearly interpolate in their measured grid, where we convert spectral type to stellar mass using the midpoints of their mass ranges (\SI{0.75}{\Msun} for K~stars, \SI{0.475}{\Msun} for early-type M~stars,  and \SI{0.215}{\Msun} for late-type M~stars).
-Outside of the age and stellar mass range covered in \citet{Richey-Yowell2023}, we extrapolate using nearest simplex (see right panel of Figure~\ref{fig:hz_nuv_evo}).
+Outside the age and stellar mass range covered in \citet{Richey-Yowell2023}, we extrapolate using nearest simplex (see right panel of Figure~\ref{fig:hz_nuv_evo}).
 
@@
 Due to how the \gls{HZ} scales with spectral type, by far most transiting \glspl{EEC} occur around M~dwarfs.
-Their \gls{NUV} fluxes are generally highest in at early times \SI{\lesssim 100}{\mega\year}.
+Their \gls{NUV} fluxes are generally highest at early times $\lesssim\SI{100}{\mega\year}$.
 These host stars, in particular late subtypes, also provide extended periods of increased \gls{NUV} emission that overlap with times when some of these planets occupy the \gls{HZ} (see Figure~\ref{fig:hz_nuv_evo}), our requirement for abiogenesis (compare Equation~\ref{eq:hypothesis-uv}).

### 2024-06-18 src/scripts/powergrid.py
double plot FGK/M for power grid

@@ -7,24 +7,58 @@ from bioverse import plots
 
 plotstyle.styleplots()
 
-def plot_powergrid(grid):
-    """Plot the power grid."""
-    labels = ('S_thresh', 'f_life')
 
-    fig, ax = plt.subplots()
-    fig, ax = plots.plot_power_grid(grid,
-                                        # method='p',
-                                        method='dlnZ',
-                                        axes=('NUV_thresh', 'f_life'), labels=labels,
-                                        log=(True, False), show=False, fig=fig, ax=ax,
-                                        # zoom_factor=2, smooth_sigma=.5,
-                                        zoom_factor=0, smooth_sigma=.2,
-                                        levels=[50, 95], cmap=cmocean.cm.dense_r)
+def plot_powergrid(grid, fig, ax, **kwargs):
+    """Plot the power grid."""
+    labels = ("S_thresh", "f_life")
+
+    fig, ax = plots.plot_power_grid(
+        grid,
+        # method='p',
+        method="dlnZ",
+        axes=("NUV_thresh", "f_life"),
+        labels=labels,
+        log=(True, False),
+        show=False,
+        fig=fig,
+        ax=ax,
+        # zoom_factor=2, smooth_sigma=.5,
+        zoom_factor=0,
+        smooth_sigma=0.2,
+        levels=[50, 95],
+        cmap=cmocean.cm.dense_r,
+        **kwargs,
+    )
     return fig, ax
 
 
-with open(paths.data / 'pipeline/grid_flife_nuv.dll', 'rb') as f:
-    grid = dill.load(f)
+fig, axs = plt.subplots(1, 2, figsize=[13, 4.5])
 
-fig, ax = plot_powergrid(grid)
+for spt, ax in zip(["FGK", "M"], axs):
+    # with open(
+    #     paths.data / "pipeline/grid_flife_nuv_{}.dll".format(spt), "rb"
+    # ) as f, open(paths.data / "pipeline/data_{}.dll".format(spt), "rb") as d:
+    #     grid = dill.load(f)
+    #     data = dill.load(d)  # do we need this?
+
+
+
+
+    # WHILE DEBUGGING
+    import pickle
+    with open(
+            paths.data / "pipeline/grid_flife_nuv_{}.pkl".format(spt), "rb"
+    ) as f, open(paths.data / "pipeline/data_{}.dll".format(spt), "rb") as d:
+        grid = pickle.load(f)
[…]

### 2024-06-19 src/scripts/bioverse_pipeline.py
add dlnZ values

@@ -396,7 +396,8 @@ def past_uv(hoststars="all", grid=True, N_grid=None, fast=False, **kwargs):
         # perform a single hypothesis test
         grid = None
         d, detected, data, nautilus = run_survey_nautilus(d)
-        _ = hypothesis_test(data)
+        results = hypothesis_test(data)
+        save_var_latex("dlnZ_{}".format(hoststars), results["dlnZ"])
 
     print("Number of planets in the sample: {}".format(len(d)))
 

### 2024-06-19 src/scripts/detections_uv.py

@@ -1,6 +1,7 @@
 import dill
 import paths
 import plotstyle
+import numpy as np
 from src.scripts.utils import read_var_latex
 
 plotstyle.styleplots()
@@ -10,10 +11,10 @@ import matplotlib.pyplot as plt
 
 def plot_detections_uv(eec, fig, ax, NUV_thresh):
     ax.scatter(eec["max_nuv"], eec["has_O2"], s=9.0)
-    ax.axvline(x=NUV_thresh, linestyle="--", color="grey")
+    ax.axvline(x=float(NUV_thresh), linestyle="--", color="grey")
     ax.set_yticks([0, 1])
     # ax.set_xlim([0,10])
-    ax.set_yticklabels(["no biosignature", "biosignature"])  # , fontsize=24)
+    ax.set_yticklabels(["no biosignature", "biosignature"], fontsize=16)
 
     ax.spines["right"].set_visible(False)
     ax.spines["top"].set_visible(False)
@@ -31,6 +32,8 @@ for spt, ax in zip(["FGK", "M"], axs):
 
     fig, ax = plot_detections_uv(data, fig, ax, NUV_thresh=read_var_latex("NUV_thresh"))
     ax.set_title(f"{spt}-type host stars")
+    ax.text(0.97, 0.85, "dlnZ = {:.0f}".format(np.round(float(read_var_latex("dlnZ_{}".format(spt)))) + 0.),
+            transform=ax.transAxes, horizontalalignment="right", color="0.2")
 
 ax.set_xlabel("max. NUV irradiance [erg/s/$cm^2$]")  # , fontsize=24)
 

### 2024-06-19 src/scripts/bioverse_pipeline.py
change grid range

@@ -279,7 +279,8 @@ def hypotest_grid(generator, survey, N_grid, fast):
     # f_life = (0.9,)  # test 1D hypothesis grid test
     # f_life = 0.99  # test 1D hypothesis grid test
     # NUV_thresh = np.logspace(1, 2.5, N_grid)
-    NUV_thresh = np.geomspace(300.0, 380.0, N_grid)
+    # NUV_thresh = np.geomspace(300.0, 380.0, N_grid)
+    NUV_thresh = np.geomspace(30.0, 3000.0, N_grid)
 
     # NUV_thresh=350.,
 

### 2024-06-19 src/scripts/bioverse_pipeline.py
change default NUV thresh

@@ -297,6 +297,7 @@ def hypotest_grid(generator, survey, N_grid, fast):
         N=N_iter,
         processes=8,
         t_total=10 * 365.25,
+        error_dump_filename=paths.root / "out/error_dump.txt",
     )
     return results
 
@@ -337,9 +338,9 @@ def past_uv(hoststars="all", grid=True, N_grid=None, fast=False, **kwargs):
         # "d_max": 60,        # TOO SMALL SAMPLE AND THE HYPOTHESIS TESTING GRID GETS STUCK WITHOUT AN ERROR MESSAGE
         "d_max": 75,  # TOO SMALL SAMPLE AND THE HYPOTHESIS TESTING GRID GETS STUCK WITHOUT AN ERROR MESSAGE
         "deltaT_min": 10.0,  # Myr
-        # "NUV_thresh": 350.0,  # choose such that n_inhabited can't be zero
+        "NUV_thresh": 350.0,  # choose such that n_inhabited can't be zero
         # "NUV_thresh": 200.0,
-        "NUV_thresh": 150.0,
+        # "NUV_thresh": 150.0,
         "f_life": 0.8,
         # "f_eta": 5.0,  # Occurrence rate scaling factor (MAKE SURE SAMPLE IS LARGE ENOUGH (see above))
     }

### 2024-06-19 src/scripts/bioverse_pipeline.py
save vars of FGK run as default

@@ -406,16 +406,26 @@ def past_uv(hoststars="all", grid=True, N_grid=None, fast=False, **kwargs):
     # save some variables for the manuscript
     if hoststars == "FGK":
         save_var_latex("uv_inhabited_FGK", len(dd[dd.inhabited]))
-    elif hoststars == "M":
-        save_var_latex("uv_inhabited_M", len(dd[dd.inhabited]))
-    elif hoststars == "all":
-        # general run
+
+        # general
         save_var_latex("d_max", g_args["d_max"])
         save_var_latex("M_G_max", g_args["M_G_max"])
         save_var_latex("f_life", params_past_uv["f_life"])
         save_var_latex("NUV_thresh", params_past_uv["NUV_thresh"])
         save_var_latex("deltaT_min", int(params_past_uv["deltaT_min"]))
         save_var_latex("uv_inhabited", len(dd[dd.inhabited]))
+
+    elif hoststars == "M":
+        save_var_latex("uv_inhabited_M", len(dd[dd.inhabited]))
+    elif hoststars == "all":
+        # general run (are we even still doing this)?
+        pass
+        # save_var_latex("d_max", g_args["d_max"])
+        # save_var_latex("M_G_max", g_args["M_G_max"])
+        # save_var_latex("f_life", params_past_uv["f_life"])
+        # save_var_latex("NUV_thresh", params_past_uv["NUV_thresh"])
+        # save_var_latex("deltaT_min", int(params_past_uv["deltaT_min"]))
+        # save_var_latex("uv_inhabited", len(dd[dd.inhabited]))
 
     # fixed variables from semianalytical analysis
     save_var_latex("semian_Nsamp1", 10)

### 2024-06-19 src/tex/ms.tex
specify file ending of static figure
double plot FGK/M for power grid
update selectivity plot

@@
 \author{Martin Schlecker}
-%\author[0000-0001-8355-2107]{Martin Schlecker}
+%\author[0000-0001-8355-2107]{Martin Schlecker} # issue with ORCID? https://github.com/showyourwork/showyourwork/issues/454
 \affiliation{Steward Observatory, The University of Arizona, Tucson, AZ 85721, USA; \href{mailto:schlecker@arizona.edu}{schlecker@arizona.edu}}
@@
     \begin{centering}
-        \includegraphics[width=\hsize]{figures/hypotheses}
+        \includegraphics[width=\hsize]{figures/hypotheses.pdf}
         \caption{Population-level hypothesis and null hypothesis on UV irradiance derived from the cyanosulfidic scenario.}
@@
 In Section~\ref{sec:results-semianalytical}, we demonstrated that the probability of obtaining true strong evidence for the hypothesis that life only originates on planets with a minimum past \gls{NUV} flux is sensitive to the distribution of sampled past \gls{NUV} fluxes, i.e., the selectivity of the survey (compare Figure~\ref{fig:semian_selectivity}).
-The maximum \gls{NUV} distribution of our generic transit survey is rather unimodal with a median around \SI{500}{\erg\per\second\per\centi\meter\squared} (see Figure~\ref{fig:nuv_distribution}), leading to a selectivity of roughly \var{selectivity_transit_volume-lim}.
+%The maximum \gls{NUV} distribution of our generic transit survey is rather unimodal with a median around \SI{500}{\erg\per\second\per\centi\meter\squared} (see Figure~\ref{fig:nuv_distribution}), leading to a selectivity of roughly \var{selectivity_transit_volumelim}.
+For both surveys targeting M dwarfs and those targeting FGK dwarfs, the maximum \gls{NUV} distribution is rather unimodal, leading to low selectivities (Figure~\ref{fig:nuv_distribution}).
 \begin{figure}
@@
         \includegraphics[width=\hsize]{figures/nuv_distribution}
-        \caption{Distribution of maximum past \gls{NUV} flux in a generic volume- and magnitude-limited transit survey.
-The best-fit beta distribution (gray) corresponds to a selectivity of \var{selectivity_transit_volume-lim}.}
+        \caption{Distribution of maximum past \gls{NUV} flux in transit surveys targeting FGK and M~stars, respectively.
+The best-fit beta distributions (gray) correspond to selectivities of $s_\mathrm{FGK} = $\var{selectivity_FGK} and $s_\mathrm{M} = $\var{selectivity_M}.}
         \label{fig:nuv_distribution}

### 2024-06-19 src/scripts/nuv_distribution.py

@@ -11,7 +11,7 @@ from scipy import stats
 import numpy as np
 
 
-def plot_nuv_distribution(data, fig, ax):
+def plot_nuv_distribution(data, fig, ax, spt):
     dataa = data.to_pandas()
 
     # force-fit a beta distribution to our NUV_max
@@ -21,7 +21,7 @@ def plot_nuv_distribution(data, fig, ax):
 
     # estimate selectivity from average of fitted beta function parameters
     selectivity = np.log10(1 / (np.mean(max_likeli[:2])))
-    save_var_latex("selectivity_transit_volume-lim", round(selectivity, 1))
+    save_var_latex("selectivity_{}".format(spt), round(selectivity, 1))
     print(f"selectivity ~ {selectivity:.2f}")
 
     x = np.arange(0.0, 1000.0, 5)
@@ -52,7 +52,7 @@ for spt, ax in zip(['FGK', 'M'], axs):
     with open(paths.data / "pipeline/data_{}.dll".format(spt), "rb") as f:
         data = dill.load(f)
 
-    fig, ax = plot_nuv_distribution(data, fig, ax)
+    fig, ax = plot_nuv_distribution(data, fig, ax, spt)
     ax.set_title(f"{spt}-type host stars")
 
 fig.savefig(paths.figures / "nuv_distribution.pdf")

### 2024-06-19 src/tex/ms.tex
bring this in the right order

@@
 
-\subsection{Semi-analytical analysis}\label{sec:met-semianalytical}
+\subsection{Semi-analytical assessment}\label{sec:met-semianalytical}
 First, we apply a Bayes Factor Design Analysis~\citep{Schoenbrodt2018} to assess the expected probabilities of obtaining true negative or true positive evidence for the hypothesis above, as well as the probability for misleading or inconclusive evidence.
@@
 
-\subsection{Considered planets and their orbits}
-All Origins of Life scenarios considered here require water as a solvent; we thus consider only rocky planets that may sustain liquid water on their surface.
+\subsection{Exoplanet survey simulations with Bioverse}
+To model the sensitivity of the information gain of a proposed mission to sample selection and survey strategy, we now conduct survey simulations with Bioverse using different sample sizes and survey strategies.
+We again require that life occurs only on planets with sufficient past UV irradiation exceeding $F_\mathrm{NUV, min}$.
+Further, we require this flux to have lasted for a minimum duration $\Delta T_\mathrm{min}$ to allow for a sufficient ``origins timescale''~\citep{Rimmer2023}.
+
+All common Origin of Life scenarios require water as a solvent.
+We thus consider only rocky planets that may sustain liquid water on their surface, i.e., that occupy their momentary \gls{HZ} during the above period.
+
+\subsubsection{Habitable zone occupancy and UV flux}
 Different formulations of habitable zones as regions around a star where a planet with Earth's atmospheric composition can maintain liquid water on its surface exist~\citep[e.g.,][]{MolLous2022,Spinelli2023,Tuchow2023}.
@@
 Nevertheless, we focus here on detecting this key absorber as it represents the general inherent observational challenges and trends.
-
-\subsection{Exoplanet survey simulations with Bioverse}
-To model the sensitivity of the information gain of a proposed mission to sample selection and survey strategy, we now conduct survey simulations with Bioverse using different sample sizes and survey strategies.
-We again require that life occurs only on planets with sufficient past UV irradiation exceeding $F_\mathrm{NUV, min}$.
-Further, we require this flux to have lasted for a minimum duration $\Delta T_\mathrm{min}$ to allow for a sufficient ``origins timescale''~\citep{Rimmer2023}, and the planet to be within its momentary \gls{HZ} during this entire period.
 
@@
 %The maximum \gls{NUV} distribution of our generic transit survey is rather unimodal with a median around \SI{500}{\erg\per\second\per\centi\meter\squared} (see Figure~\ref{fig:nuv_distribution}), leading to a selectivity of roughly \var{selectivity_transit_volumelim}.
-For both surveys targeting M dwarfs and those targeting FGK dwarfs, the maximum \gls{NUV} distribution is rather unimodal, leading to low selectivities (Figure~\ref{fig:nuv_distribution}).
+For both surveys targeting M dwarfs and those targeting FGK dwarfs, the maximum \gls{NUV} distribution is rather unimodal, leading to low selectivities (see Figure~\ref{fig:nuv_distribution}).
 \begin{figure}
@@
         \includegraphics[width=\hsize]{figures/detections_uv}
-        \caption{Recovered biosignature detections in the \gls{NUV} flux -- biosignature occurrence space. The dashed line denotes the threshold \gls{NUV} flux $F_\mathrm{NUV, min}$.}
+        \caption{Recovered biosignature detections in the \gls{NUV} flux -- biosignature occurrence space.
+The dashed line denotes a generic threshold \gls{NUV} flux $F_\mathrm{NUV, min} = \var{NUV_thresh}\,\SI{}{\erg\per\second\per\centi\meter\squared}$.}
         \label{fig:detections_uv}

### 2024-06-19 src/tex/ms.tex
direct imaging = optional
rewrite results for FGK and M surveys

@@
 
-\subsection{Semi-analytical assessment}\label{sec:met-semianalytical}
+\subsection{Semi-analytical approach}\label{sec:met-semianalytical}
 First, we apply a Bayes Factor Design Analysis~\citep{Schoenbrodt2018} to assess the expected probabilities of obtaining true negative or true positive evidence for the hypothesis above, as well as the probability for misleading or inconclusive evidence.
@@
 
-\subsubsection{Direct imaging survey (e.g., Habitable Worlds Observatory, LIFE)}
+%\subsubsection{Direct imaging survey (e.g., Habitable Worlds Observatory, LIFE)}
 %\subsubsubsection{target list}
@@
 
-\subsection{Semi-analytical approach}\label{sec:results-semianalytical}
+\subsection{Semi-analytical assessment}\label{sec:results-semianalytical}
 In Section~\ref{sec:met-semianalytical} we computed the probability for true positive evidence for $H_\mathrm{1}$ and $H_\mathrm{null}$ respectively.
@@
 These host stars, in particular late subtypes, also provide extended periods of increased \gls{NUV} emission that overlap with times when some of these planets occupy the \gls{HZ} (see Figure~\ref{fig:hz_nuv_evo}), our requirement for abiogenesis (compare Equation~\ref{eq:hypothesis-uv}).
-Because of that, all inhabited planets orbit M~dwarfs.
+Because of that, all inhabited planets in our magnitude- and volume-limited sample orbit M~dwarfs.
 
-\subsubsection{selectivity of a generic transit survey}
+Here, we are interested in the statistical power of a transit survey with a realistic sample selection and size.
+In the following, we fix the sample size to \var{N_nautilus} and consider two different survey strategies targeting FGK and M~dwarfs, respectively.
+We further investigate the sensitivity of the survey to the a priori unknown threshold \gls{NUV} flux $F_\mathrm{NUV, min}$ and the abiogenesis rate $f_\mathrm{life}$.
+
+\subsubsection{Selectivity of simulated transit surveys}
 In Section~\ref{sec:results-semianalytical}, we demonstrated that the probability of obtaining true strong evidence for the hypothesis that life only originates on planets with a minimum past \gls{NUV} flux is sensitive to the distribution of sampled past \gls{NUV} fluxes, i.e., the selectivity of the survey (compare Figure~\ref{fig:semian_selectivity}).
 %The maximum \gls{NUV} distribution of our generic transit survey is rather unimodal with a median around \SI{500}{\erg\per\second\per\centi\meter\squared} (see Figure~\ref{fig:nuv_distribution}), leading to a selectivity of roughly \var{selectivity_transit_volumelim}.
-For both surveys targeting M dwarfs and those targeting FGK dwarfs, the maximum \gls{NUV} distribution is rather unimodal, leading to low selectivities (see Figure~\ref{fig:nuv_distribution}).
+For both surveys targeting M dwarfs and those targeting FGK dwarfs, the maximum \gls{NUV} distribution is rather unimodal.
+Applying the approach from Sect.~\ref{sec:results-semianalytical} of fitting a Beta function to the distribution, we find rather low selectivities (see Figure~\ref{fig:nuv_distribution}).
 \begin{figure}
@@
 
-\subsubsection{expected biosignature pattern in a transit survey}
+\subsubsection{Expected biosignature pattern in a transit survey}
 A representative recovery of the injected biosignature pattern is shown in Figure~\ref{fig:detections_uv}.
 There, we assumed an abiogenesis rate of $f_\mathrm{life} = \var{f_life}$ and a minimum \gls{NUV} flux of $F_\mathrm{NUV, min} = \var{NUV_thresh}\,\SI{}{\erg\per\second\per\centi\meter\squared}$.
-This leads to a distribution of biosignature detections with detections increasingly occurring above a threshold inferred NUV flux.
+All injected biosignatures are assumed to be detected, and the maximum \gls{NUV} flux is estimated from the host star's spectral type and age with an uncertainty corresponding to the intrinsic scatter in the \gls{NUV} fluxes in \citet{Richey-Yowell2023}.
 \begin{figure}
@@
 \end{figure}
+This leads to a distribution of biosignature detections with detections increasingly occurring above a threshold inferred NUV flux.
+In this example case, the few biosignature detections in the FGK sample lead to a higher evidence ($dlnZ_\mathrm{FGK} = \var{dlnZ_FGK}$) than in the M dwarf sample ($dlnZ_\mathrm{M} = \var{dlnZ_M}$), where the majority of planets are above the threshold \gls{NUV} flux.
 
@@
     \item The required sample size for detecting this correlation depends on the occurrence of life on temperate exoplanets, the distribution of host star spectral types in the sample, ...
-        A direct imaging mission with the \gls{HWO} will require characterizing XXX planets ...
+%        A direct imaging mission with the \gls{HWO} will require characterizing XXX planets ...
     \item If the predicted \gls{NUV} correlation exists, yielding strong evidence for it is likely ($> \SI{90}{\percent}$) for sample sizes $\geq 100$, and a survey strategy that targets extreme values of inferred past \gls{NUV} irradiation.

### 2024-06-19 src/tex/ms.tex
update conclusions

@@
     \item The required \gls{NUV} radiation in the cyanosulfidic scenario should lead to a correlation between past \gls{NUV} flux and current occurrence of biosignatures, both of which will soon be observationally accessible.
-    \item The required sample size for detecting this correlation depends on the occurrence of life on temperate exoplanets, the distribution of host star spectral types in the sample, ...
+    \item The required sample size for detecting this correlation depends on the occurrence of life on temperate exoplanets and the distribution of host star spectral types in the sample; in particular their past maximum \gls{NUV} fluxes.
 %        A direct imaging mission with the \gls{HWO} will require characterizing XXX planets ...
     \item If the predicted \gls{NUV} correlation exists, yielding strong evidence for it is likely ($> \SI{90}{\percent}$) for sample sizes $\geq 100$, and a survey strategy that targets extreme values of inferred past \gls{NUV} irradiation.
-    \item Abiogenesis in the cyanosulfidic scenario strongly prefers late host star spectral types in a sample of transiting planets.
+    \item Abiogenesis in the cyanosulfidic scenario prefers late host star spectral types due to the higher \gls{NUV} fluxes in their habitable zones.
+However, despite the overall lower biosignature rate, the distribution of maximum past \gls{NUV} fluxes in a transit survey targeting planets around FGK stars is beneficial for Bayesian hypothesis testing.
+%     in a sample of transiting planets.
 Any potential biosignature detection on a planet around a star other than an M~dwarf would disvafor this scenario and any other \gls{ool} theory requiring extended periods of high \gls{NUV} irradiation.

### 2024-06-20 src/tex/ms.tex
only one theoretical experiment conducted

@@
 Test for negative correlations as well?}
-Our third theoretical experiment is to test the UV irradiance requirement (Sect.~\ref{sec:predictions}) by relating the occurrence of life on an exo-earth candidate with a minimum past quiescent stellar UV flux, focusing on the prebiotically interesting \gls{NUV} range from \SIrange{200}{280}{\nano\meter}.
+Here, we conduct a theoretical experiment on the UV irradiance requirement (Sect.~\ref{sec:predictions}) by relating the occurrence of life on an exo-earth candidate with a minimum past quiescent stellar UV flux, focusing on the prebiotically interesting \gls{NUV} range from \SIrange{200}{280}{\nano\meter}.
 \todo[inline]{SUKRIT: got a good reference for this?}  %~\citep{RanjanSasselov2016}.?

### 2024-06-20 src/tex/ms.tex
explain Bioverse

@@
 \subsection{Exoplanet survey simulations with Bioverse}
-To model the sensitivity of the information gain of a proposed mission to sample selection and survey strategy, we now conduct survey simulations with Bioverse using different sample sizes and survey strategies.
-We again require that life occurs only on planets with sufficient past UV irradiation exceeding $F_\mathrm{NUV, min}$.
+To assess the information gain of a real exoplanet survey, we employed Bioverse~\citep{Bixel2021}, a framework that integrates multiple components including statistically realistic simulations of exoplanet populations, a survey simulation module, and a Bayesian hypothesis testing module to evaluate the statistical power of different observational strategies.
+The general approach is as follows:
+\begin{enumerate}
+%    \item We simulate a synthetic population of exoplanets with known properties, including their host stars, orbital parameters, and physical characteristics.
+\item \textbf{Exoplanet population synthesis:} We populate the Gaia Nearby Star Catalog with synthetic exoplanets whose orbital parameters and planetary properties reflect our current understanding of exoplanet demographics.
+    \item \textbf{Survey Simulation:} We simulate the detection and characterization of these exoplanets with a hypothetical survey, taking into account the survey's sensitivity, target selection, and observational biases.
+    \item \textbf{Hypothesis Tests:} We employ Bayesian statistical methods to evaluate the likelihood that a given survey would detect a specified statistical trend in the exoplanet population and estimate the precision with which the survey could constrain the parameters of that trend.
+\end{enumerate}
+To model the sensitivity of the information gain of a proposed mission to sample selection and survey strategy, we conduct survey simulations with Bioverse using different sample sizes and survey strategies.
+
+To test the UV hypothesis, we require that life occurs only on planets with sufficient past UV irradiation exceeding $F_\mathrm{NUV, min}$.
 Further, we require this flux to have lasted for a minimum duration $\Delta T_\mathrm{min}$ to allow for a sufficient ``origins timescale''~\citep{Rimmer2023}.

### 2024-06-20 src/scripts/powergrid.py
format ticklabels

@@ -10,7 +10,7 @@ plotstyle.styleplots()
 
 def plot_powergrid(grid, fig, ax, **kwargs):
     """Plot the power grid."""
-    labels = ("S_thresh", "f_life")
+    labels = ("$F_\mathrm{NUV, min}$", "$f_\mathrm{life}$")
 
     fig, ax = plots.plot_power_grid(
         grid,
@@ -29,33 +29,32 @@ def plot_powergrid(grid, fig, ax, **kwargs):
         cmap=cmocean.cm.dense_r,
         **kwargs,
     )
+
+    ax.set_ylim(0.0, 1.0)
+
+
     return fig, ax
 
 
 fig, axs = plt.subplots(1, 2, figsize=[13, 4.5])
 
-for spt, ax in zip(["FGK", "M"], axs):
+for i, (spt, ax) in enumerate(zip(["FGK", "M"], axs)):
     # with open(
     #     paths.data / "pipeline/grid_flife_nuv_{}.dll".format(spt), "rb"
     # ) as f, open(paths.data / "pipeline/data_{}.dll".format(spt), "rb") as d:
     #     grid = dill.load(f)
     #     data = dill.load(d)  # do we need this?
 
-
-
-
-    # WHILE DEBUGGING
+    # WHILE DEBUGGING (and no .dll files are available)
     import pickle
+
     with open(
-            paths.data / "pipeline/grid_flife_nuv_{}.pkl".format(spt), "rb"
+        paths.data / "pipeline/grid_flife_nuv_{}.pkl".format(spt), "rb"
     ) as f, open(paths.data / "pipeline/data_{}.dll".format(spt), "rb") as d:
         grid = pickle.load(f)
         data = dill.load(d)  # do we need this?
 
-
-
-
-    cbar = False if spt == "FGK" else True
+    cbar = False if i == 0 else True
 
     fig, ax = plot_powergrid(grid, fig, ax, cbar=cbar)
     ax.set_title(f"{spt}-type host stars")

### 2024-06-20 src/tex/ms.tex
elaborate and specify Bioverse's application here

--- 2024-06-20 src/tex/ms.tex
@@
 \subsection{Exoplanet survey simulations with Bioverse}
-To model the sensitivity of the information gain of a proposed mission to sample selection and survey strategy, we now conduct survey simulations with Bioverse using different sample sizes and survey strategies.
-We again require that life occurs only on planets with sufficient past UV irradiation exceeding $F_\mathrm{NUV, min}$.
-Further, we require this flux to have lasted for a minimum duration $\Delta T_\mathrm{min}$ to allow for a sufficient ``origins timescale''~\citep{Rimmer2023}.
+To assess the information gain of a real exoplanet survey, we employed Bioverse~\citep{Bixel2021}, a framework that integrates multiple components including statistically realistic simulations of exoplanet populations, a survey simulation module, and a Bayesian hypothesis testing module to evaluate the statistical power of different observational strategies.
+The general approach is as follows:
+\begin{enumerate}
+\item \textbf{Exoplanet population synthesis:} We populate the Gaia Catalogue of Nearby Stars~\citep{Smart2021} with synthetic exoplanets whose orbital parameters and planetary properties reflect our current understanding of exoplanet demographics~\citep{Bergsten2022}.
+    \item \textbf{Survey simulation:} We simulate the detection and characterization of these exoplanets with a hypothetical survey, taking into account the survey's sensitivity, target selection, and observational biases.
+To model the sensitivity of the information gain of a proposed mission to sample selection and survey strategy, we conduct survey simulations with Bioverse using different sample sizes and survey strategies.
+    \item \textbf{Hypothesis testing:} We employ Bayesian statistical methods to evaluate the likelihood that a given survey would detect a specified statistical trend in the exoplanet population and estimate the precision with which the survey could constrain the parameters of that trend.
+    A common definition of the null hypothesis $H_0$, which is also applied here, is that there is no relationship between the independent variable (here: maximum NUV flux) and the dependent variable (here: biosignature occurrence).
+    The alternative hypothesis $H_1$ proposes a specific relationship parameterized by $\theta$.
+    If the dependent variable is of binary nature such as in the case of biosignature detections, we model the likelihood function as
+    \begin{equation}
+\mathcal{L}(y \mid \boldsymbol{\theta})=\prod_i^N\left[y_i h\left(\boldsymbol{\theta}, x_i\right)+\left(1-y_i\right)\left(1-h\left(\boldsymbol{\theta}, x_i\right)\right)\right],
+\end{equation}
+where here $y_i \in \{0,1\}$ is the biosignature detection variable, $x_i$ is the maximum NUV flux, and $h(\boldsymbol{\theta}, x_i)$ is the probability of detecting a biosignature given the maximum NUV flux and the model parameters $\boldsymbol{\theta}$.
+We then use a nested sampling method~\citep{Speagle2020} to compute the Bayesian evidence for the null and alternative hypotheses and estimate the strength of evidence for the alternative hypothesis.
 
-All common Origin of Life scenarios require water as a solvent.
-We thus consider only rocky planets that may sustain liquid water on their surface, i.e., that occupy their momentary \gls{HZ} during the above period.
+To determine the diagnostic capability of a given survey, Bioverse runs multiple iterations of the simulated survey and calculates the fraction of realizations that successfully reject the null hypothesis.
+We use this metric, known as the statistical power, to quantify the potential information content of the survey, identify critical design trades, and find strategies that maximize the survey's scientific return.
+The posterior samples obtained from the nested sampling runs further allows us to estimate the precision with which the survey could constrain the parameters of the hypothesized trend.
+\end{enumerate}
 
 \subsubsection{Habitable zone occupancy and UV flux}
+To test the UV hypothesis, we require that life occurs only on planets with sufficient past UV irradiation exceeding $F_\mathrm{NUV, min}$.
+Further, we require this flux to have lasted for a minimum duration $\Delta T_\mathrm{min}$ to allow for a sufficient ``origins timescale''~\citep{Rimmer2023}.
+All common Origin of Life scenarios require water as a solvent;
+we thus consider only rocky planets that may sustain liquid water on their surface, i.e., that occupy their momentary \gls{HZ} during the above period.
+
 Different formulations of habitable zones as regions around a star where a planet with Earth's atmospheric composition can maintain liquid water on its surface exist~\citep[e.g.,][]{MolLous2022,Spinelli2023,Tuchow2023}.

### 2024-06-21 src/tex/ms.tex
specify priors

@@
 We assigned the development of life to a random fraction $f_\mathrm{life}$ of all temperate planets fulfilling these requirements.
-For a population-level hypothesis, we then have
+
+For the hypothesis tests, we then define our alternative hypothesis as
 \begin{equation}\label{eq:hypothesis-uv}
@@
 i.e., no correlation with UV flux.
+We imposed log-uniform priors on the parameters $\boldsymbol{\theta}$, sampling $f_\mathrm{life}$ from a log-uniform distribution between $10^{-3}$ and $1$ and $F_\mathrm{NUV, min}$ from a log-uniform distribution between $10^{1}$ and $10^{5}$.
 

### 2024-06-21 src/scripts/inhabited_FGKM.py
revise histogram plot, separate spectral types

@@ -1,22 +1,28 @@
 import dill
+from matplotlib import pyplot as plt
+
 import paths
 import plotstyle
 plotstyle.styleplots()
 
 import seaborn as sns
 
-def plot_inhabited_FGKM(d):
+def plot_inhabited_FGKM(d, fig, ax):
     """Plot the distribution of inhabited planets in the FGKM diagram.
 
     Parameters
     ----------
     d : bioverse Table
         The table containing the sample.
+    fig : matplotlib figure
+        The figure to plot on
+    ax : matplotlib axis
+        The axis to plot on
 
     Returns
     -------
-    g : seaborn.FacetGrid
-        a seaborn FacetGrid object containing the plot.
+    ax : matplotlib axis
+        The axis containing the plot
     """
 
     dd = d.to_pandas()
@@ -32,16 +38,33 @@ def plot_inhabited_FGKM(d):
 
     dd.loc[:, "Category"] = dd.apply(cat, axis=1)
     
-    g = sns.catplot(
-        dd, x="SpT", kind="count", hue="Category", order=['F', 'G', 'K', 'M']
+    # g = sns.catplot(
+    ax = sns.countplot(
+        dd, ax=ax, x="SpT",
+        # kind="count",
+        hue="Category", order=['F', 'G', 'K', 'M']
     )  # , hue_order = [True, False])
-    g.ax.set_yscale("log")
+    # ax.set_yscale("log")
 
-    g.savefig(paths.figures / "inhabited_FGKM.pdf")
-    return g
+    return ax
 
 
-with open(paths.data / 'pipeline/sample.dll', 'rb') as f:
-    sample = dill.load(f)
+fig, axs = plt.subplots(2, 1, figsize=(5, 8), sharex=False, sharey=True, gridspec_kw={"hspace": 0.3})
+for spt, ax in zip(["FGK", "M"], axs):
+    with open(paths.data / "pipeline/sample_{}.dll".format(spt), "rb") as f:
+        sample = dill.load(f)
 
-g = plot_inhabited_FGKM(sample)
+    ax = plot_inhabited_FGKM(sample, fig, ax)
+    ax.set_title(f"{spt}-type host stars")
+
+    # add sample size to the upper right corner
+    ax.text(0.97, 0.9, f"N = {len(sample)}", transform=ax.transAxes, horizontalalignment="right", color="0.2")
+
+    if ax is axs[0]:
[…]

### 2024-06-21 src/tex/ms.tex
describe fractions of EEC/inhabited planets in the different samples

@@
         \includegraphics[width=\hsize]{figures/inhabited_FGKM}
-        \caption{Host stars of all transiting planets, \glspl{EEC}, and inhabited planets.
-In a sample of \var{N_nautilus} transiting planets, most \glspl{EEC} and all inhabited planets orbit M~dwarfs.}
+        \caption{Host stars of all transiting \glspl{EEC} and inhabited planets.
+        In the FGK sample, most \glspl{EEC} and all inhabited planets orbit K~dwarfs.
+        In an M~dwarf sample of the same size, the fraction of inhabited planets is larger.
+        }
         \label{fig:inhabited_FGKM}
@@
 
-As shown in Figure~\ref{fig:inhabited_FGKM}, the majority of \glspl{EEC} and all inhabited planets orbit M~dwarfs.
+As shown in Figure~\ref{fig:inhabited_FGKM}, the majority of \glspl{EEC} orbit lower-mass stars.
+The fraction of inhabited planets is highest in the M~dwarf sample due to the higher \gls{NUV} fluxes in the \gls{HZ} of these stars.
 

### 2024-06-22 src/tex/ms.tex
make multipanel figs wide
number and explain example planet tracks

@@
 \documentclass[modern,linenumbers]{aastex631}
+
 \input{preamble}
@@
         \caption{Interpolated stellar luminosity evolution (left) and evolution of the \gls{NUV} flux in the \gls{HZ} as a function of host star mass.
-The scatter plot shows age and host star mass of the transiting planets in the synthetic sample.% with \gls{EEC} highlighted in green.
-        We show the evolutionary tracks of a few randomly selected planets.
-        Tracks with extended overlap of \gls{HZ} occupancy and high \gls{NUV} flux (green sections) fulfill our requirement for abiogenesis.}
+The scatter plot shows age and host star mass of the transiting planets in the synthetic sample.
+% with \gls{EEC} highlighted in green.
+        A few example tracks are shown, where extended overlap of \gls{HZ} occupancy and high \gls{NUV} flux (green sections) fulfill our requirement for abiogenesis.
+        Planet~1 is an \gls{EEC} that never receives sufficient \gls{NUV} flux for abiogenesis.
+        Planet~2 and Planet~3 enter the \gls{HZ} at different times and receive sufficient \gls{NUV} flux for different durations until their respective host star evolves below the threshold.
+        }
         \label{fig:hz_nuv_evo}
@@
 
-\begin{figure}
+\begin{figure*}
     \script{semian_true_evidence.py}
@@
     \end{centering}
-\end{figure}
+\end{figure*}
 
@@
 Applying the approach from Sect.~\ref{sec:results-semianalytical} of fitting a Beta function to the distribution, we find rather low selectivities (see Figure~\ref{fig:nuv_distribution}).
-\begin{figure}
+\begin{figure*}
     \script{nuv_distribution.py}
@@
 All injected biosignatures are assumed to be detected, and the maximum \gls{NUV} flux is estimated from the host star's spectral type and age with an uncertainty corresponding to the intrinsic scatter in the \gls{NUV} fluxes in \citet{Richey-Yowell2023}.
-\begin{figure}
+\begin{figure*}
     \script{detections_uv.py}
@@
         \includegraphics[width=\hsize]{figures/detections_uv}
-        \caption{Recovered biosignature detections in the \gls{NUV} flux -- biosignature occurrence space.
+        \caption{Recovered biosignature detections in the \gls{NUV} flux-biosignature occurrence space.
 The dashed line denotes a generic threshold \gls{NUV} flux $F_\mathrm{NUV, min} = \var{NUV_thresh}\,\SI{}{\erg\per\second\per\centi\meter\squared}$.}
@@
     \end{centering}
-\end{figure}
+\end{figure*}
 This leads to a distribution of biosignature detections with detections increasingly occurring above a threshold inferred NUV flux.
@@
 
-\begin{figure}
+\begin{figure*}
     \script{powergrid.py}

### 2024-06-22 src/scripts/hz_nuv_evo.py

@@ -35,7 +35,7 @@ def plot_interpolation(
                 linewidths=0.6,
                 label=["EEC" if EEC else None][0],
             )
-        ax.legend(loc="upper left", framealpha=.0)
+        ax.legend(loc="upper left", framealpha=0.0)
     fig.colorbar(q, label=cbarlabel)
     ax.set_xscale("log")
     ax.set_xlabel("Time (Gyr)")
@@ -45,7 +45,7 @@ def plot_interpolation(
 
 def plot_nuv_evo(fig, ax):
     interp_nuv = interpolate_nuv()
-    T = np.geomspace(5e-3, 10., num=200)
+    T = np.geomspace(5e-3, 10.0, num=200)
     M = np.linspace(0.1, 1.0, num=200)
     M, T = np.meshgrid(M, T)
     Z = interp_nuv(M, T)
@@ -55,14 +55,17 @@ def plot_nuv_evo(fig, ax):
     return ax
 
 
-def plot_hz_and_nuv(fig, ax, sample, NUV_thresh=100.0, N_sample=2, random_state=42):
+def plot_hz_and_nuv(fig, ax, sample, NUV_thresh=100.0, N_sample=2, random_state=44):
     "plot where HZ and sufficient NUV overlap for a few planets."
     eec = sample[sample["EEC"].astype(bool)]
     eec.evolve()
     eecdf = eec.to_pandas()
     inhabited = eecdf[eecdf.hz_and_uv]
 
-    for id in inhabited.sample(N_sample, random_state=random_state).planetID:
+    exampleplanets = {}
+    for i, id in enumerate(
+        inhabited.sample(N_sample, random_state=random_state).planetID
+    ):
         Mst = eecdf[eecdf.planetID == id]["M_st"]
         t = eec.evolution[id]["time"]
         ax.plot(
@@ -79,17 +82,43 @@ def plot_hz_and_nuv(fig, ax, sample, NUV_thresh=100.0, N_sample=2, random_state=
         t_hzuv = t[hz_and_uv]
         ax.plot(t_hzuv, [Mst for tt in t_hzuv], lw=4, c="C0")
 
+        # save Mst and last time to dictionary
+        exampleplanets[Mst.values[0]] = t[-1]
+
+        # print out the planet's SpT
+        print(f"Planet {i+1} has SpT {eecdf[eecdf.planetID == id].SpT.values[0]}")
+
     # plot a not-inhabited planet, too
[…]

### 2024-06-24 src/tex/ms.tex
plot grid points from HAZMAT paper.

@@
 
-\todo[inline]{biosignatures: Focus on molecular Oxygen? Excess Methane? \citep[e.g.,][]{Seeburger2023}}
-% detecting methanogenic life: Sandora2020, Wang2018,
+% detecting methanogenic life: Sandora2020, Wang2018, Seeburger2023
 A commonly discussed biosignature is molecular Oxygen (O$_2$), which on Earth emerged as a byproduct of photosynthesis during the Proterozoic era.
@@
         \includegraphics[width=\hsize]{figures/hz_nuv_evo}
-        \caption{Interpolated stellar luminosity evolution (left) and evolution of the \gls{NUV} flux in the \gls{HZ} as a function of host star mass.
-The scatter plot shows age and host star mass of the transiting planets in the synthetic sample.
+        \caption{Interpolated stellar luminosity evolution (left) and evolution of the \gls{NUV} flux in the \gls{HZ} (right) as a function of host star mass.
+        The scatter plots show age and host star mass of the transiting planets in the synthetic FGK sample; crosses denote the estimated \gls{NUV} values in \citet{Richey-Yowell2019}.
 % with \gls{EEC} highlighted in green.

### 2024-06-24 src/scripts/hz_nuv_evo.py

@@ -1,3 +1,5 @@
+import pandas as pd
+
 import paths
 import dill
 import plotstyle
@@ -11,6 +13,26 @@ import matplotlib.colors as colors
 from bioverse.util import interpolate_luminosity, interpolate_nuv
 
 
+def plot_planets_scatter(ax, d, norm, col=None):
+    # for (v, c) in [(False, 'white'), (True, 'C1')]:
+    for c, (EEC, group) in zip(["white", "C0"], d.to_pandas().groupby("EEC")):
+        if col != "white":
+            col = group["L_st_interp"]
+        ax.scatter(
+            group["age"],
+            group["M_st"],
+            c=col,
+            s=4,
+            norm=norm,
+            # edgecolors="white",
+            edgecolors=c,
+            linewidths=0.6,
+            label=["EEC" if EEC else None][0],
+        )
+    ax.legend(loc="upper left", framealpha=0.0)
+    return ax
+
+
 def plot_interpolation(
     fig, ax, T, M, Z, d=None, input_points=None, log=True, cbarlabel=None, **kwargs
 ):
@@ -22,20 +44,7 @@ def plot_interpolation(
     if input_points:
         ax.plot(*input_points, "x", ms=1, label="input point")
     if d:
-        # for (v, c) in [(False, 'white'), (True, 'C1')]:
-        for c, (EEC, group) in zip(["white", "C0"], d.to_pandas().groupby("EEC")):
-            ax.scatter(
-                group["age"],
-                group["M_st"],
-                c=group["L_st_interp"],
-                s=4,
-                norm=norm,
-                # edgecolors="white",
-                edgecolors=c,
-                linewidths=0.6,
-                label=["EEC" if EEC else None][0],
-            )
-        ax.legend(loc="upper left", framealpha=0.0)
+        ax = plot_planets_scatter(ax, d, norm)
     fig.colorbar(q, label=cbarlabel)
     ax.set_xscale("log")
     ax.set_xlabel("Time (Gyr)")
@@ -43,7 +52,7 @@ def plot_interpolation(
     return ax
 
 
-def plot_nuv_evo(fig, ax):
+def plot_nuv_evo(fig, ax, log=True, d=None):
     interp_nuv = interpolate_nuv()
[…]

### 2024-06-25 src/scripts/semian_true_evidence.py
H3 -> H1

@@ -89,14 +89,14 @@ def plot_true_evidence(ax):
         ax.plot(
             gtvec,
             [np.sum(bfh1_kh1 > g) / len(bfh1_kh1) for g in gtvec],
-            label=showlabel + "True evidence for $H_\mathrm{3}$",
+            label=showlabel + "True evidence for $H_\mathrm{1}$",
             c="C0",
             ls=ls,
         )
         ax.plot(
             gtvec,
             [np.sum(bfh0_kh0 > g) / len(bfh1_kh1) for g in gtvec],
-            label=showlabel + "True evidence for $H_\mathrm{3, null}$",
+            label=showlabel + "True evidence for $H_\mathrm{null}$",
             c="C1",
             ls=ls,
         )

### 2024-06-25 src/scripts/semian_true_evidence.py
"whichever is smaller" was confusing

@@ -162,8 +162,10 @@ def P_true_evidence(sampsize, eta=10, nboot=int(1e4), pisamp=None, theta_lsamp=N
     bfh0_kh0 = 1 / bfh1_kh0
 
     trueeH1 = np.sum(bfh1_kh1 > eta) / len(bfh1_kh1)
-    trueeH0 = np.sum(bfh0_kh0 > eta) / len(bfh0_kh0)
-    return np.min([trueeH1, trueeH0])
+    # trueeH0 = np.sum(bfh0_kh0 > eta) / len(bfh0_kh0)
+    # return np.min([trueeH1, trueeH0])
+    return trueeH1
+
 
 def plot_evidence_sampsize(ax, s_max=500):
     """plot strong true evidence as a function of sample size"""
@@ -172,7 +174,7 @@ def plot_evidence_sampsize(ax, s_max=500):
 
     ax.plot(svec,true_strong_evidence_p)
     ax.set_ylim(-.05, 1.05)
-    ax.set_ylabel('Probability of strong true evidence')
+    ax.set_ylabel('P($BF_{H_1, H_\mathrm{null}}$ > 10)')
     ax.set_xlabel('Sample size')
 
     return ax

### 2024-06-26 src/scripts/bioverse_pipeline.py
cool kids use decorators

@@ -436,6 +436,7 @@ def past_uv(hoststars="all", grid=True, N_grid=None, fast=False, **kwargs):
     return d, grid, detected, data, nautilus
 
 
+@timeit
 def main(fast=True):
     """Run the Bioverse pipeline."""
     print("RUNNING BIOVERSE PIPELINE")

### 2024-06-26 src/scripts/utils.py

@@ -1,6 +1,6 @@
 # Code that we need that doesn't produce figures
-
 import paths
+import time
 
 
 def save_var_latex(key, value, datafile=paths.tex / "variables.dat"):
@@ -51,3 +51,18 @@ def read_var_latex(key, datafile=paths.tex / "variables.dat"):
         pass
 
     return dict_var[key]
+
+
+def timeit(func):
+    """
+    Decorator for measuring function's running time.
+    https://stackoverflow.com/questions/35656239/how-do-i-time-script-execution-time-in-pycharm-without-adding-code-every-time
+    """
+    def measure_time(*args, **kw):
+        start_time = time.time()
+        result = func(*args, **kw)
+        print("Processing time of %s(): %.2f seconds."
+              % (func.__qualname__, time.time() - start_time))
+        return result
+
+    return measure_time

### 2024-06-26 src/scripts/bioverse_pipeline.py
try Mann-Whitney U test instead of nested sampling

@@ -219,7 +219,8 @@ def h_null(theta, X):
     return np.full(shape, theta)
 
 
-def hypothesis_test(data, method="dynesty"):
+# def hypothesis_test(data, method="dynesty"):
+def hypothesis_test(data, method="mannwhitney"):
     """Perform a single hypothesis test on the data."""
     params = ("f_life", "NUV_thresh")
     log = (True, True)
@@ -242,11 +243,14 @@ def hypothesis_test(data, method="dynesty"):
 
     results = h_nuv.fit(data, method=method)
 
-    print(
-        "The evidence in favor of the hypothesis is: dlnZ = {:.1f} (corresponds to p = {:.1E})".format(
-            results["dlnZ"], np.exp(-results["dlnZ"])
+    if method in ["dynesty", "emcee"]:
+        print(
+            "The evidence in favor of the hypothesis is: dlnZ = {:.1f} (corresponds to p = {:.1E})".format(
+                results["dlnZ"], np.exp(-results["dlnZ"])
+            )
         )
-    )
+    elif method == "mannwhitney":
+        print("The p-value of the Mann-Whitney U test is {}".format(results["p"]))
     return results
 
 
@@ -273,7 +277,7 @@ def hypotest_grid(generator, survey, N_grid, fast):
     if fast:
         N_iter = 2
     else:
-        N_iter = 6
+        N_iter = 8
     # f_life = np.geomspace(0.1, 1.0, N_grid)
     f_life = np.geomspace(0.2, 1.0, N_grid)
     # f_life = (0.9,)  # test 1D hypothesis grid test
@@ -290,8 +294,8 @@ def hypotest_grid(generator, survey, N_grid, fast):
         h_nuv,
         generator,
         survey,
-        # method="mannwhitney",
-        method="dynesty",
+        method="mannwhitney",
+        # method="dynesty",
         f_life=f_life,
         NUV_thresh=NUV_thresh,
         N=N_iter,
@@ -389,7 +393,7 @@ def past_uv(hoststars="all", grid=True, N_grid=None, fast=False, **kwargs):
         elif N_grid:
             N_grid = N_grid
         else:
-            N_grid = 6
+            N_grid = 8
         nautilus = create_survey_nautilus()
         grid = hypotest_grid(g, nautilus, N_grid=N_grid, fast=fast)
         detected = None
@@ -399,7 +403,10 @@ def past_uv(hoststars="all", grid=True, N_grid=None, fast=False, **kwargs):
         grid = None
[…]

### 2024-06-26 src/tex/ms.tex
"whichever is smaller" was confusing
adapt conclusions

@@
 The vertical lines denote the thresholds for ``strong'' evidence, $BF_{H_i, H_j}$ > 10, and ``extreme'' evidence, $BF_{H_i, H_j}$ > 100.
-Right: Probability of true strong evidence for $H_\mathrm{1}$ or $H_\mathrm{null}$, whichever is smaller, as a function of sample size $n$.}
+Right: Probability of true strong evidence for $H_\mathrm{1}$ as a function of sample size $n$.}
         \label{fig:semian_true_evidence}
@@
     \item Models of the Origins of Life provide hypotheses that are testable with near-future exoplanet surveys.
-    \item The required \gls{NUV} radiation in the cyanosulfidic scenario should lead to a correlation between past \gls{NUV} flux and current occurrence of biosignatures, both of which will soon be observationally accessible.
+    \item The required \gls{NUV} radiation in the cyanosulfidic scenario should lead to a correlation between past \gls{NUV} flux and current occurrence of biosignatures that will be observationally testable.
     \item The required sample size for detecting this correlation depends on the occurrence of life on temperate exoplanets and the distribution of host star spectral types in the sample; in particular their past maximum \gls{NUV} fluxes.
@@
     \item Abiogenesis in the cyanosulfidic scenario prefers late host star spectral types due to the higher \gls{NUV} fluxes in their habitable zones.
-However, despite the overall lower biosignature rate, the distribution of maximum past \gls{NUV} fluxes in a transit survey targeting planets around FGK stars is beneficial for Bayesian hypothesis testing.
+However, due to a wider distribution of maximum past \gls{NUV} fluxes in the \gls{HZ} of FGK stars, a transit survey targeting FGK systems is more likely to obtain strong evidence for or against the \gls{NUV} hypothesis.
+
 %     in a sample of transiting planets.

### 2024-07-02 src/scripts/powergrid.py
plot grid depending on method

@@ -12,47 +12,54 @@ def plot_powergrid(grid, fig, ax, **kwargs):
     """Plot the power grid."""
     labels = ("$F_\mathrm{NUV, min}$", "$f_\mathrm{life}$")
 
+    # check if the dictionary grid contains 'p' or 'dlnZ' values
+    if "p" in grid:
+        method = "p"
+    elif grid["dlnZ"] is not None:
+        method = "dlnZ"
+    else:
+        method = "dBIC"
+
     fig, ax = plots.plot_power_grid(
         grid,
-        # method='p',
-        method="dlnZ",
+        method=method,
         axes=("NUV_thresh", "f_life"),
         labels=labels,
-        log=(True, False),
+        # log=(True, False),
+        log=(True, True),
         show=False,
         fig=fig,
         ax=ax,
         # zoom_factor=2, smooth_sigma=.5,
         zoom_factor=0,
-        smooth_sigma=0.2,
-        levels=[50, 95],
+        # smooth_sigma=0.2,
+        # levels=[50, 95],
+        levels=None,
         cmap=cmocean.cm.dense_r,
         **kwargs,
     )
 
     ax.set_ylim(0.0, 1.0)
-
-
     return fig, ax
 
 
 fig, axs = plt.subplots(1, 2, figsize=[13, 4.5])
 
 for i, (spt, ax) in enumerate(zip(["FGK", "M"], axs)):
-    # with open(
-    #     paths.data / "pipeline/grid_flife_nuv_{}.dll".format(spt), "rb"
-    # ) as f, open(paths.data / "pipeline/data_{}.dll".format(spt), "rb") as d:
-    #     grid = dill.load(f)
-    #     data = dill.load(d)  # do we need this?
-
-    # WHILE DEBUGGING (and no .dll files are available)
-    import pickle
-
     with open(
-        paths.data / "pipeline/grid_flife_nuv_{}.pkl".format(spt), "rb"
+        paths.data / "pipeline/grid_flife_nuv_{}.dll".format(spt), "rb"
     ) as f, open(paths.data / "pipeline/data_{}.dll".format(spt), "rb") as d:
-        grid = pickle.load(f)
+        grid = dill.load(f)
         data = dill.load(d)  # do we need this?
+
+    # # WHILE DEBUGGING (and no .dll files are available)
+    # import pickle
+    #
+    # with open(
[…]

### 2024-07-02 src/scripts/semian_true_evidence.py

@@ -60,9 +60,9 @@ def get_evidence(sampsize=100):
     # For convenience, we redefine theta here so that one can play with sample size and such
     theta = rng.uniform(
         0, 1, sampsize
-    )  # vector theta chosen at random for demonstrative purposes for simplicity we assume theta in 0,1
+    )  # vector theta chosen at random for demonstrative purposes; for simplicity we assume theta in 0,1
 
-    nboot = 100000  # it is advised to use larger nboot because we have increased the sampling space quite a bit (probably already too much)
+    nboot = 100000  # it is advised to use a large nboot because we have increased the sampling space quite a bit (probably already too much)
     pisamp = rng.uniform(0, 1, nboot)
     theta_lsamp = rng.uniform(0, 1, nboot)
     kh1samp = sample_H1(theta, pisamp, theta_lsamp)

### 2024-07-02 src/scripts/detections_uv.py
display sp. t.

@@ -32,7 +32,8 @@ for spt, ax in zip(["FGK", "M"], axs):
 
     fig, ax = plot_detections_uv(data, fig, ax, NUV_thresh=read_var_latex("NUV_thresh"))
     ax.set_title(f"{spt}-type host stars")
-    ax.text(0.97, 0.85, "dlnZ = {:.0f}".format(np.round(float(read_var_latex("dlnZ_{}".format(spt)))) + 0.),
+    # ax.text(0.97, 0.85, "dlnZ = {:.0f}".format(np.round(float(read_var_latex("dlnZ_{}".format(spt)))) + 0.),
+    ax.text(0.97, 0.85, "p = {:.2f}".format(float(read_var_latex("p_{}".format(spt)))),
             transform=ax.transAxes, horizontalalignment="right", color="0.2")
 
 ax.set_xlabel("max. NUV irradiance [erg/s/$cm^2$]")  # , fontsize=24)

### 2024-07-02 src/tex/ms.tex
reorder discussion
write first discussion section

@@
 % Define document class
-%\documentclass[twocolumn,twocolappendix,linenumbers]{aastex631}
-\documentclass[modern,linenumbers]{aastex631}
+\documentclass[twocolumn,twocolappendix,linenumbers]{aastex631}
+%\documentclass[modern,linenumbers]{aastex631}
 
@@
 \label{sec:hypotests}
-\todo[inline]{Briefly introduce Bayesian model comparison, then present the particular hypotheses in their mathematical form.}
+%\todo[inline]{Briefly introduce Bayesian model comparison, then present the particular hypotheses in their mathematical form.}
 
@@
 %WE consider life ``as we know it''
+\subsection{Constraining power for the origins of life as a function of biosignature location}
+\todo[inline]{How does the location of biosignature detections impact the credibility of OOL scenarios?}
+\todo[inline]{biosignature on M~dwarf planet vs. FGK: Impact on UV flux requirement?}
+The cyanosulfidic scenario, in particular its predicted existence of a minimum \gls{NUV} flux required for prebiotic chemistry, offers an opportunity to test an origins of life hypothesis with a statistical transit survey sampling planets with varying \gls{NUV} flux histories.
+It comes to no surprise that the success rate of such a test is sensitive to the sample size of the survey and to the occurrence of life on temperate exoplanets.
+As we have shown, the statistical power of this test also depends on the distribution of past \gls{NUV} fluxes in the sample and on the required threshold flux.
+Optimizing the survey to sample a wide range of NUV flux values, particularly at the extremes, can enhance the likelihood of obtaining strong evidence for or against the hypothesis.
+Intermediate values of the threshold \gls{NUV} flux are more likely to yield strong evidence than extreme values, as the dataset under the alternative hypothesis $H_1$ differs more from the null hypothesis in this case while still being sufficiently populated.
+The required threshold flux is, of course, a priori unknown and we cannot influence it.
+If, however, better theoretical predictions for the required \gls{NUV} flux for abiogenesis become available, the survey strategy can be further optimized, for instance by targeting planets that are estimated to have received a \gls{NUV} flux slightly below and above this threshold.
+
+An interesting  aspect lies in the distribution of host star spectral types:
+Under the \gls{NUV} hypothesis, the occurrence of life is expected to correlate with the host star's spectral type, with late-type M~dwarfs being favored due to their higher \gls{NUV} fluxes in the \gls{HZ}~\citep{Richey-Yowell2023}.
+However, the distribution of maximum past \gls{NUV} fluxes in this sample is narrow in this sample, which may limit the constraining power of the survey depending on the (unknown) threshold \gls{NUV} flux.
+An M~dwarf sample may help to test the high \gls{NUV} flux end of the hypothesis; a higher occurrence of biosignatures here would support the hypothesis that a higher NUV flux is beneficial or necessary for life.
+
+FGK stars, on the other hand, show a wider distribution of maximum past \gls{NUV} fluxes in the \gls{HZ}, which may increase the likelihood of obtaining strong evidence for or against the hypothesis.
+With this sample, the survey will be more sensitive to the low \gls{NUV} flux end of the hypothesis.
+A lack of biosignatures on these planets would support the \gls{NUV} hypothesis, whereas their presence might suggest that lower NUV fluxes are also sufficient for abiogenesis or indicate different abiogenic pathways.
+CITE
 
@@
 
-\subsection{Constraining power for the origins of life as a function of biosignature location}
-\todo[inline]{How does the location of biosignature detections impact the credibility of OOL scenarios?}
-\todo[inline]{biosignature on M~dwarf planet vs. FGK: Impact on UV flux requirement?}
 \subsubsection{Sampling strategy for testing a predicted minimum \gls{NUV} flux}

### 2024-07-02 src/scripts/bioverse_pipeline.py

@@ -12,14 +12,13 @@ from utils import *
 # import pickle
 import numpy as np
 
-# Import the Generator class
+# Import Bioverse objects
 from bioverse.generator import Generator
 from bioverse.survey import TransitSurvey
-
 from bioverse.hypothesis import Hypothesis
-
 # from bioverse.constants import CONST, DATA_DIR
 
+from semian_true_evidence import BFH1
 
 def eec_filter(d):
     """Filter the planets to only include EECs."""
@@ -220,8 +219,17 @@ def h_null(theta, X):
 
 
 # def hypothesis_test(data, method="dynesty"):
-def hypothesis_test(data, method="mannwhitney"):
+def hypothesis_test(data, testmethod):
     """Perform a single hypothesis test on the data."""
+
+    # if testmethod == "binomial":
+    #     bf = BFH1(k, theta, pi, theta_l)
+    #
+    #
+    #     print("The evidence in favor of the hypothesis is: dlnZ = {:.1f} (corresponds to p = {:.1E})".format()
+    #     return results
+    #
+
     params = ("f_life", "NUV_thresh")
     log = (True, True)
     bounds = np.array([[1e-3, 1.0], [10.0, 1e5]])
@@ -241,20 +249,30 @@ def hypothesis_test(data, method="mannwhitney"):
         log=(True,),
     )
 
-    results = h_nuv.fit(data, method=method)
+    results = h_nuv.fit(data, method=testmethod)
 
-    if method in ["dynesty", "emcee"]:
+    if testmethod == "dynesty":
         print(
             "The evidence in favor of the hypothesis is: dlnZ = {:.1f} (corresponds to p = {:.1E})".format(
                 results["dlnZ"], np.exp(-results["dlnZ"])
             )
         )
-    elif method == "mannwhitney":
+    elif testmethod == "emcee":
+        print(
+            "The differential Bayesian Information Criterion is dBIC={}".format(
+                results["dBIC"]
+            )
+        )
+    elif testmethod == "mannwhitney":
         print("The p-value of the Mann-Whitney U test is {}".format(results["p"]))
+
+    else:
+        raise KeyError("No recognized method for hypothesis testing was provided.")
+
     return results
 
 
-def hypotest_grid(generator, survey, N_grid, fast):
[…]

### 2024-07-05 src/tex/ms.tex
Motivate semianalytical and bioverse approaches

@@
 \subsection{Semi-analytical approach}\label{sec:met-semianalytical}
-First, we apply a Bayes Factor Design Analysis~\citep{Schoenbrodt2018} to assess the expected probabilities of obtaining true negative or true positive evidence for the hypothesis above, as well as the probability for misleading or inconclusive evidence.
+First, we apply a Bayes Factor Design Analysis~\citep{Schoenbrodt2018} to assess the expected probabilities of obtaining true negative or true positive evidence for the hypothesis above, as well as the probability for misleading or inconclusive evidence, under idealized conditions.
+This serves as a first-order estimate of the information content of a survey, before we take into account impacts from exoplanet demographics, sample selection, and survey strategy.
 
@@
 \subsection{Exoplanet survey simulations with Bioverse}
-To assess the information gain of a real exoplanet survey, we employed Bioverse~\citep{Bixel2021}, a framework that integrates multiple components including statistically realistic simulations of exoplanet populations, a survey simulation module, and a Bayesian hypothesis testing module to evaluate the statistical power of different observational strategies.
+A real exoplanet survey will be subject to observational biases, sample selection effects, and the underlying demographics of the planet sample.
+To assess the information gain of a realistic exoplanet survey, we employed Bioverse~\citep{Bixel2021}, a framework that integrates multiple components including statistically realistic simulations of exoplanet populations, a survey simulation module, and a Bayesian hypothesis testing module to evaluate the statistical power of different observational strategies.
 The general approach is as follows:
@@
         \includegraphics[width=\hsize]{figures/inhabited_FGKM}
-        \caption{Host stars of all transiting \glspl{EEC} and inhabited planets.
+        \caption{Host stars of all transiting \glspl{EEC} and inhabited planets in a simulated transit survey.
         In the FGK sample, most \glspl{EEC} and all inhabited planets orbit K~dwarfs.
@@
 % with \gls{EEC} highlighted in green.
-        A few example tracks are shown, where extended overlap of \gls{HZ} occupancy and high \gls{NUV} flux (green sections) fulfill our requirement for abiogenesis.
+        A few example tracks for an example threshold flux of $F_\mathrm{NUV, min} = \var{NUV_thresh}\,\SI{}{\erg\per\second\per\centi\meter\squared}$ are shown; extended overlap of \gls{HZ} occupancy and high \gls{NUV} flux (green sections) fulfills our requirement for abiogenesis.
         Planet~1 is an \gls{EEC} that never receives sufficient \gls{NUV} flux for abiogenesis.
@@
 
-\subsubsection{Sampling strategy for testing a predicted minimum \gls{NUV} flux}
+\subsection{Sampling strategy for testing a predicted minimum \gls{NUV} flux}
 We show in Sect.~\ref{sec:results} that the constraining power for testing the hypothesis of a minimum past \gls{NUV} flux required for abiogenesis is sensitive on the occurrence of life, the value of this threshold flux, the sample size, and the distribution of sampled past \gls{NUV} fluxes.

### 2024-07-05 src/tex/ms.tex
add summary at the very end

@@
 \section{Conclusions and future work}
+We have investigated the potential of upcoming exoplanet surveys to test the hypothesis that a minimum past \gls{NUV} flux is required for abiogenesis.
+To this end, we first employed a semi-analytical Bayesian analysis to estimate probabilities of obtaining strong evidence for or against this hypothesis.
+We then used the Bioverse framework to assess the diagnostic power of realistic transit surveys, taking into account exoplanet demographics, time-dependency of habitability and \gls{NUV} fluxes, observational biases, and target selection.
 
@@
 \begin{enumerate}
-    \item Models of the Origins of Life provide hypotheses that are testable with near-future exoplanet surveys.
     \item The required \gls{NUV} radiation in the cyanosulfidic scenario should lead to a correlation between past \gls{NUV} flux and current occurrence of biosignatures that will be observationally testable.
@@
 
+Overall, our work demonstrates that upcoming exoplanet surveys have the potential to test the hypothesis that a minimum past \gls{NUV} flux is required for abiogenesis.
+More generally, we found that models of the origins of life provide hypotheses that may be testable with near-future exoplanet surveys.
+Our work highlights the importance of understanding the context in which a biosignature detection is made, which can not only help to assess the credibility of the detection but also to test competing theories of the origins of life on Earth and beyond.
+
 %\input{acknowledgments_contributions}

### 2024-07-08 src/tex/ms.tex
describe star and planet generation

@@
 
-\subsubsection{Habitable zone occupancy and UV flux}
+\subsubsection{Simulated star and planet sample}
+We generated two sets of synthetic exoplanet populations, one for FGK-type stars and one for M-type stars.
+The stellar samples are drawn from the Gaia Catalogue of Nearby Stars~\citep{Smart2021} with a maximum Gaia magnitude of \var{M_G_max} and a maximum stellar mass of \var{M_st_max} \SI{}{\Msun}.
+We included stars out to a maximum distance $d_{\max}$ that depends on the required planet sample size.
+Planets were generated and assigned to the synthetic stars following the occurrence rates and size/orbit distributions of \citet{Bergsten2022}.
+Following \citet{Bixel2021}, we considered only transiting \glspl{EEC} with radii $0.8\, S^{0.25} < R < 1.4 $ that are within the habitable zone (see Section~\ref{sec:met-hz}).
+The lower limit was suggested as a minimum planet size to retain an atmosphere~\citep{Zahnle2017}.
+To generate planet samples larger than what the stellar catalog in combination with these occurrence rates yields, we scaled up the occurrence rates by a constant factor that yields the desired number of planets.
+This was in particular necessary for the FGK sample, where the rate of transiting planets that occupy the habitable zone is low.
+
+For all survey simulations and hypothesis tests, we repeated the above in a Monte Carlo fashion to generate randomized ensembles of synthetic star and planet populations~\citep[compare][]{Bixel2021}.
+
+\subsubsection{Habitable zone occupancy and UV flux}\label{sec:met-hz}
 To test the UV hypothesis, we require that life occurs only on planets with sufficient past UV irradiation exceeding $F_\mathrm{NUV, min}$.
@@
 We use the parametrization in \citet{Kopparapu2014} to derive luminosity and planetary mass-dependent edges of the \gls{HZ} $a_\mathrm{inner}$ and $a_\mathrm{outer}$.
-We further follow \citet{Bixel2021} and classify as \glspl{EEC} all planets with radii $0.8 S^{0.25} < R < 1.4 $ that orbit within the above boundaries.
-The lower limit was suggested as a minimum planet size to retain an atmosphere~\citep{Zahnle2017}.
 

### 2024-07-09 src/scripts/bioverse_pipeline.py
save and cite some transit survey parameters

@@ -538,6 +538,11 @@ def main(fast=False, testmethod='mannwhitney'):
                     paths.data / "pipeline/data_{}.dll".format(spt), "wb"
                 ) as file:
                     dill.dump(_data, file)
+
+    # save relevant Nautilus survey parameters to variables file
+    save_var_latex("nautilus_max_nuv", nautilus.measurements['max_nuv'].precision)
+    save_var_latex("nautilus_S", nautilus.measurements['S'].precision)
+
     #
     # # run statistical power grids
     # for spt in ["FGK", "M"]:

### 2024-07-09 src/scripts/inhabited-aafo-nuv_thresh.py
adapt fraction of inhabited plot

@@ -28,15 +28,19 @@ def run_simulation(nuv_thresh, n_planets, star_type, seed):
     params_past_uv = get_params_past_uv(star_type, NUV_thresh=nuv_thresh, seed=seed)
     g, g_args = generate_generator(label=None, **params_past_uv)
     d = g.generate()
+    d = d.to_pandas()
 
-    d = d[:n_planets]
+    # sample n_planets, avoiding error when the sample is smaller than n_planets
+    if len(d) < n_planets:
+        d = d.sample(len(d), random_state=seed, replace=True)
+    else:
+        d = d.sample(n_planets, random_state=seed)
 
     inhabited_planets = len(d[d["inhabited"]])
     return inhabited_planets / n_planets
 
 
 def simulate():
-
     # set_params
     n_planets = 100
     n_nuv_thresholds = 10
@@ -44,8 +48,10 @@ def simulate():
     n_simulations_per_threshold = 10
     # n_simulations_per_threshold = 2
     # nuv_thresholds = np.geomspace(30.0, 3000.0, n_nuv_thresholds)
-    nuv_thresholds = np.geomspace(200.0, 600.0, n_nuv_thresholds)
-    f_life = 1.0
+    # nuv_thresholds = np.geomspace(200.0, 600.0, n_nuv_thresholds)
+    nuv_thresholds = np.linspace(50.0, 500.0, n_nuv_thresholds)
+    # f_life = 1.0
+    f_life = .8
 
     # Storage for results
     results_fgk = np.zeros((n_nuv_thresholds, n_simulations_per_threshold))
@@ -102,22 +108,23 @@ def plot_results():
     lower_bound_m = np.percentile(results_m, 5, axis=1)
     upper_bound_m = np.percentile(results_m, 95, axis=1)
 
-    plt.figure(figsize=(10, 6))
+    fig, ax = plt.subplots(figsize=(10, 6))
 
     # Plot FGK stars
-    plt.plot(nuv_thresholds_fgk, mean_fractions_fgk, label="FGK Stars", marker="o")
-    plt.fill_between(nuv_thresholds_fgk, lower_bound_fgk, upper_bound_fgk, alpha=0.3)
+    ax.plot(nuv_thresholds_fgk, mean_fractions_fgk, label="FGK Stars", marker="o")
[…]

### 2024-07-09 src/scripts/nuv_distribution.py
show inhabited

@@ -15,7 +15,8 @@ def normalize_data(data):
     max_val = data.max()
     return (data - min_val) / (max_val - min_val)
 
-def plot_nuv_distribution(data, fig, ax, spt):
+def plot_nuv_distribution(sample, data, fig, ax, spt):
+    sample = sample.to_pandas()
     dataa = data.to_pandas()
     normalized_data = normalize_data(dataa.max_nuv)
 
@@ -32,7 +33,8 @@ def plot_nuv_distribution(data, fig, ax, spt):
     # plot histogram and beta distribution fitted on non-normalized data
     x = np.arange(0.0, 1000.0, 5)
     max_likeli = stats.beta.fit(dataa.max_nuv, method="MM")
-    ax.hist(dataa.max_nuv, density=True, color="C0")
+    # ax.hist(dataa.max_nuv, density=True, color="C0")
+    ax.hist([dataa.max_nuv[sample.inhabited], dataa.max_nuv[~sample.inhabited]], stacked=True, density=True, color=["C1", "C0"], label=["inhabited", "EEC"])
     ax.plot(
         x,
         stats.beta.pdf(x, *max_likeli[:2], loc=max_likeli[2], scale=max_likeli[3]),
@@ -41,6 +43,10 @@ def plot_nuv_distribution(data, fig, ax, spt):
 
     ax.set_xlabel("max. NUV irradiance $F_\mathrm{NUV, max}$ [erg/s/$cm^2$]")
     ax.set_ylabel("Probability density")
+    if spt == "FGK":
+        ax.legend(title=None)
+
+
     # ax.text(
     #     0.97,
     #     0.85,
@@ -56,10 +62,12 @@ def plot_nuv_distribution(data, fig, ax, spt):
 fig, axs = plt.subplots(1, 2, figsize=(15, 2.5))
 
 for spt, ax in zip(['FGK', 'M'], axs):
+    with open(paths.data / "pipeline/sample_{}.dll".format(spt), "rb") as f:
+        sample = dill.load(f)
     with open(paths.data / "pipeline/data_{}.dll".format(spt), "rb") as f:
         data = dill.load(f)
 
-    fig, ax = plot_nuv_distribution(data, fig, ax, spt)
+    fig, ax = plot_nuv_distribution(sample, data, fig, ax, spt)
     ax.set_title(f"{spt}-type host stars")
 
 fig.savefig(paths.figures / "nuv_distribution.pdf")

### 2024-07-09 src/scripts/bioverse_pipeline.py
change default NUV_thresh

@@ -335,7 +335,8 @@ def get_params_past_uv(hoststars="all", **kwargs):
         "d_max": 75,  # TOO SMALL SAMPLE AND THE HYPOTHESIS TESTING GRID GETS STUCK WITHOUT AN ERROR MESSAGE
         "deltaT_min": 10.0,  # Myr
         # "NUV_thresh": 350.0,  # choose such that n_inhabited can't be zero
-        "NUV_thresh": 380.0,  # choose such that n_inhabited can't be zero
+        # "NUV_thresh": 380.0,  # choose such that n_inhabited can't be zero
+        "NUV_thresh": 250.0,  # choose such that n_inhabited can't be zero
         # "f_life": 0.8,
         "f_life": 0.99,
         # "f_eta": 5.0,  # Occurrence rate scaling factor (MAKE SURE SAMPLE IS LARGE ENOUGH (see above))
@@ -451,13 +452,13 @@ def past_uv(
         # perform a single hypothesis test
         grid = None
         d, detected, data, nautilus = run_survey_nautilus(d)
+        print("Number of planets in the sample: {}".format(len(d)))
         results = hypothesis_test(data, testmethod)
         try:
             save_var_latex("dlnZ_{}".format(hoststars), results["dlnZ"])
         except KeyError:
             save_var_latex("p_{}".format(hoststars), results["p"])
 
-    print("Number of planets in the sample: {}".format(len(d)))
 
     # save some variables for the manuscript
     if hoststars == "FGK":
@@ -466,6 +467,7 @@ def past_uv(
         # general
         save_var_latex("d_max", g_args["d_max"])
         save_var_latex("M_G_max", g_args["M_G_max"])
+        save_var_latex("M_st_max", g_args["M_st_max"])
         try:
             save_var_latex("f_life", params_past_uv["f_life"])
         except KeyError:
@@ -525,7 +527,7 @@ def main(fast=False, testmethod='mannwhitney'):
 
             else:
                 # single hypothesis test
-                _d, _grid, _detected, _data, _nautilus = past_uv(
+                _d, _grid, _detected, _data, nautilus = past_uv(
                     hoststars=spt, testmethod=testmethod, grid=False
                 )
 

### 2024-07-09 src/scripts/inhabited-aafo-nuv_thresh.py
change labels

@@ -118,8 +118,8 @@ def plot_results():
     ax.plot(nuv_thresholds_m, mean_fractions_m, label="M Stars", marker="s")
     ax.fill_between(nuv_thresholds_m, lower_bound_m, upper_bound_m, alpha=0.3)
 
-    ax.set_xlabel("NUV Threshold")
-    ax.set_ylabel("Fraction of Inhabited Planets")
+    ax.set_xlabel("NUV Threshold [erg/s/$cm^2$]")
+    ax.set_ylabel("Fraction of inhabited planets")
     # ax.set_xscale("log")
     # ax.set_title("Fraction of Inhabited Planets vs. NUV Threshold")
     ax.legend()

### 2024-07-09 src/tex/ms.tex
save and cite some transit survey parameters
include new figure and describe it

@@
 
+... measure the maximum past \gls{NUV} flux with a precision of \var{nautilus_max_nuv} and the instellation that a planet receives with a precision of \var{nautilus_S}.
+
 \section{Results}
@@
 In this example case, the few biosignature detections in the FGK sample lead to a higher evidence ($dlnZ_\mathrm{FGK} = \var{dlnZ_FGK}$) than in the M dwarf sample ($dlnZ_\mathrm{M} = \var{dlnZ_M}$), where the majority of planets are above the threshold \gls{NUV} flux.
+
+\begin{figure}
+    \script{inhabited_aafo_nuv_thresh.py}
+    \begin{centering}
+        \includegraphics[width=\hsize]{figures/inhabited_aafo_nuv_thresh}
+        \caption{Fraction of inhabited planets for different threshold \gls{NUV} fluxes under the UV hypothesis if the abiogenesis rate is \var{f_life}.}
+        \label{fig:inhabited_aafo_nuv_thresh}
+    \end{centering}
+\end{figure}
+Figure \ref{fig:inhabited_aafo_nuv_thresh} shows the fraction of inhabited planets under the UV hypothesis for different threshold \gls{NUV} fluxes and for a high abiogenesis rate of \var{f_life}.
+This fraction decreases sharply with increasing threshold flux, as fewer planets receive sufficient \gls{NUV} flux for abiogenesis.
+For the FGK sample, the fraction of inhabited planets drops at lower threshold fluxes than for the M dwarf sample.
 

### 2024-07-10 src/tex/ms.tex
describe transit survey, measurements, precision

@@
 
-\subsubsection{Transit survey (e.g., Nautilus)}
+\subsubsection{Transit survey simulations}
+With the synthetic star and planet samples generated, we simulated observations of these planets with a hypothetical transit survey.
+Using Bioverse's survey module, we simulated noisy measurements of key observables assuming a capable transit survey that can characterize a large sample with high photometric precision.
+We roughly followed the mission parameters of the Nautilus mission concept~\citep{Apai2019,Apai2022} and measured planetary instellation (for \gls{HZ} occupancy) with a precision of \var{nautilus_S} and host star effective temperature with a precision of \var{nautilus_T_eff_st}~K.
+We assumed that the maximum past \gls{NUV} flux a planet received can be determined within a precision of \var{nautilus_max_nuv}.
+To marginalize over choices of biosignatures and their detectability, which are beyond the scope of this study, we assumed that any inhabited planet would show a biosignature detectable by the survey.
 
@@
 %\todo[inline]{Provisional stellar target List for the habitable Worlds Observatory:~\citep{Mamajek2023}} % read like in https://github.com/mkenworthy/HWObows/tree/40b1e5f2bc2bbbe8c999b097c80ca5a8cda0aeae/src/scripts
-
-... measure the maximum past \gls{NUV} flux with a precision of \var{nautilus_max_nuv} and the instellation that a planet receives with a precision of \var{nautilus_S}.
 

### 2024-07-10 src/tex/ms.tex
expand captions

@@
         \caption{Distribution of maximum past \gls{NUV} flux in transit surveys targeting FGK and M~stars, respectively.
-The best-fit beta distributions (gray) correspond to selectivities of $s_\mathrm{FGK} = $\var{selectivity_FGK} and $s_\mathrm{M} = $\var{selectivity_M}.}
+The best-fit beta distributions (gray) correspond to selectivities of $s_\mathrm{FGK} = $\var{selectivity_FGK} and $s_\mathrm{M} = $\var{selectivity_M}.
+Red areas show inhabited planets for a threshold \gls{NUV} flux of $F_\mathrm{NUV, min} = \var{NUV_thresh}\,\SI{}{\erg\per\second\per\centi\meter\squared}$ and an abiogenesis rate of $f_\mathrm{life} = \var{f_life}$.}
         \label{fig:nuv_distribution}
@@
         \includegraphics[width=\hsize]{figures/inhabited_aafo_nuv_thresh}
-        \caption{Fraction of inhabited planets for different threshold \gls{NUV} fluxes under the UV hypothesis if the abiogenesis rate is \var{f_life}.}
+        \caption{Fraction of inhabited planets for different threshold \gls{NUV} fluxes under the UV hypothesis if the abiogenesis rate is \var{f_life}.
+Shaded regions denote 90 percent confidence intervals of randomized sample generations.}
         \label{fig:inhabited_aafo_nuv_thresh}

### 2024-07-10 src/scripts/nuv_distribution.py
add description

@@ -1,3 +1,5 @@
+# Plot the distribution of measured maximum NUV irradiance for inhabited and non-inhabited planets in FGK and M stars, respectively.
+
 import dill
 import paths
 import plotstyle

### 2024-07-10 src/scripts/inhabited-aafo-nuv_thresh.py
plot transiting and non-transiting samples

@@ -22,10 +22,13 @@ np.random.seed(42)
 
 
 # Function to run a single simulation and return the fraction of inhabited planets
-def run_simulation(nuv_thresh, n_planets, star_type, seed):
+def run_simulation(nuv_thresh, n_planets, star_type, seed, generator_kwargs):
     np.random.seed(seed)
 
     params_past_uv = get_params_past_uv(star_type, NUV_thresh=nuv_thresh, seed=seed)
+    for key, value in generator_kwargs.items():
+        params_past_uv[key] = value
+
     g, g_args = generate_generator(label=None, **params_past_uv)
     d = g.generate()
     d = d.to_pandas()
@@ -34,31 +37,49 @@ def run_simulation(nuv_thresh, n_planets, star_type, seed):
     if len(d) < n_planets:
         d = d.sample(len(d), random_state=seed, replace=True)
     else:
-        d = d.sample(n_planets, random_state=seed)
+        d = d.sample(n_planets, random_state=seed, replace=True)
 
-    inhabited_planets = len(d[d["inhabited"]])
-    return inhabited_planets / n_planets
+    fraction = len(d[d["inhabited"]]) / len(d)
+    print(fraction)
+    return fraction
 
 
-def simulate():
+def simulate(transiting=True):
     # set_params
-    n_planets = 100
-    n_nuv_thresholds = 10
-    # n_nuv_thresholds = 2
-    n_simulations_per_threshold = 10
-    # n_simulations_per_threshold = 2
+    n_planets = 500   # maximum number of planets to sample
+    # n_nuv_thresholds = 10
+    n_nuv_thresholds = 7
+    # n_simulations_per_threshold = 10
+    n_simulations_per_threshold = 4
+
     # nuv_thresholds = np.geomspace(30.0, 3000.0, n_nuv_thresholds)
     # nuv_thresholds = np.geomspace(200.0, 600.0, n_nuv_thresholds)
-    nuv_thresholds = np.linspace(50.0, 500.0, n_nuv_thresholds)
+    nuv_thresholds = np.linspace(10.0, 800.0, n_nuv_thresholds)
     # f_life = 1.0
-    f_life = .8
+
+    # DEBUG
+    generator_kwargs = {
+    # 'f_life' : 0.8,
+    'f_life' : 1.0,
+    'deltaT_min' : 1., # Myr. Smaller values seems so slightly enhance #inhabited planets in the FGK sample.
+
+    # increase sample size
+    'f_eta' : 15,
[…]

### 2024-07-10 src/scripts/bioverse_pipeline.py
enable MW-U tests in pipeline run

@@ -4,7 +4,6 @@ It heavily relies on Bioverse and its auxiliary functions.
 """
 # import pickle
 import dill
-import timeit
 
 import paths
 from utils import *
@@ -368,8 +367,15 @@ def get_params_past_uv(hoststars="all", **kwargs):
 
     return params_past_uv
 
+
 def past_uv(
-    hoststars="all", grid=True, N_grid=None, testmethod='mannwhitney', powergrid=False, fast=False, **kwargs
+    hoststars="all",
+    grid=True,
+    N_grid=None,
+    testmethod="mannwhitney",
+    powergrid=False,
+    fast=False,
+    **kwargs
 ):
     """Test the hypothesis that life only originates on planets with a minimum past UV irradiance.
 
@@ -407,19 +413,6 @@ def past_uv(
     params_past_uv = get_params_past_uv(hoststars)
 
     g, g_args = generate_generator(label=None, **params_past_uv)  # , **kwargs)
-    d = g.generate()
-
-    # keep only 100 planets
-    # d = d[:100]
-
-    # DEBUG
-
-    dd = d.to_pandas()
-
-    print("Total number of planets: {}".format(len(d)))
-    print("Inhabited: {}".format(len(dd[dd.inhabited])))
-
-    # d.evolve()
 
     if grid:
         # perform a grid of hypothesis tests
@@ -450,6 +443,15 @@ def past_uv(
     #
     else:
         # perform a single hypothesis test
+
+        d = g.generate()
+        # keep only 100 planets
+        d = d[:100]
+
+        dd = d.to_pandas()
+        print("Total number of planets: {}".format(len(d)))
+        print("Inhabited: {}".format(len(dd[dd.inhabited])))
+
         grid = None
         d, detected, data, nautilus = run_survey_nautilus(d)
         print("Number of planets in the sample: {}".format(len(d)))
@@ -498,15 +500,12 @@ def past_uv(
 
 
 @timeit
-def main(fast=False, testmethod='mannwhitney'):
+def main(fast=False, testmethod="mannwhitney"):
     """Run the Bioverse pipeline."""
-    # for grid in [False, True]:
-
-    # for grid in [True]:
-    for grid in [False]:
-    # DEBUG
-
-
+    for grid in [False, True]:
+        # for grid in [True]:
+        # for grid in [False]:
+        # DEBUG
 
         # for spt in ["all", "FGK", "M"]:
[…]

### 2024-07-10 src/scripts/bioverse_pipeline.py
update grid parameters

@@ -15,9 +15,11 @@ import numpy as np
 from bioverse.generator import Generator
 from bioverse.survey import TransitSurvey
 from bioverse.hypothesis import Hypothesis
+
 # from bioverse.constants import CONST, DATA_DIR
 
 from semian_true_evidence import BFH1
+
 
 def eec_filter(d):
     """Filter the planets to only include EECs."""
@@ -293,20 +295,17 @@ def hypotest_grid(generator, survey, N_grid, fast, testmethod):
 
     if fast:
         N_iter = 3
-        # DEBUG
-
     else:
         N_iter = 8
     # f_life = np.geomspace(0.1, 1.0, N_grid)
-    # f_life = np.geomspace(1e-3, 1.0, N_grid)
-    f_life = np.geomspace(0.5, 1.0, N_grid)
+    f_life = np.geomspace(1e-3, 1.0, N_grid)
+    # f_life = np.geomspace(0.5, 1.0, N_grid)
     # f_life = (0.9,)  # test 1D hypothesis grid test
     # f_life = 0.99  # test 1D hypothesis grid test
     # NUV_thresh = np.geomspace(300.0, 380.0, N_grid)
     # NUV_thresh = np.geomspace(30.0, 3000.0, N_grid)
-    NUV_thresh = np.geomspace(200.0, 600.0, N_grid)
-    # DEBUG
-
+    # NUV_thresh = np.geomspace(200.0, 600.0, N_grid)
+    NUV_thresh = np.geomspace(20.0, 800.0, N_grid)
 
     from bioverse.analysis import test_hypothesis_grid
 
@@ -327,6 +326,7 @@ def hypotest_grid(generator, survey, N_grid, fast, testmethod):
     )
     return results
 
+
 def get_params_past_uv(hoststars="all", **kwargs):
     """define default parameters for the past UV hypothesis test."""
     params_past_uv = {
@@ -337,7 +337,7 @@ def get_params_past_uv(hoststars="all", **kwargs):
         # "NUV_thresh": 380.0,  # choose such that n_inhabited can't be zero
         "NUV_thresh": 250.0,  # choose such that n_inhabited can't be zero
         # "f_life": 0.8,
-        "f_life": 0.99,
+        "f_life": 1.0,
         # "f_eta": 5.0,  # Occurrence rate scaling factor (MAKE SURE SAMPLE IS LARGE ENOUGH (see above))
     }
 
@@ -351,7 +351,7 @@ def get_params_past_uv(hoststars="all", **kwargs):
[…]

### 2024-07-10 src/scripts/bioverse_pipeline.py
save survey precisions

@@ -542,8 +542,9 @@ def main(fast=False, testmethod="mannwhitney"):
                     dill.dump(_data, file)
 
     # save relevant Nautilus survey parameters to variables file
-    save_var_latex("nautilus_max_nuv", nautilus.measurements['max_nuv'].precision)
-    save_var_latex("nautilus_S", nautilus.measurements['S'].precision)
+    save_var_latex("nautilus_max_nuv", nautilus.measurements["max_nuv"].precision)
+    save_var_latex("nautilus_S", nautilus.measurements["S"].precision)
+    save_var_latex("nautilus_T_eff_st", nautilus.measurements["T_eff_st"].precision)
 
     #
     # # run statistical power grids

### 2024-07-11 src/scripts/bioverse_pipeline.py
scale up grid

@@ -296,7 +296,7 @@ def hypotest_grid(generator, survey, N_grid, fast, testmethod):
     if fast:
         N_iter = 3
     else:
-        N_iter = 8
+        N_iter = 16
     # f_life = np.geomspace(0.1, 1.0, N_grid)
     f_life = np.geomspace(1e-3, 1.0, N_grid)
     # f_life = np.geomspace(0.5, 1.0, N_grid)
@@ -305,7 +305,7 @@ def hypotest_grid(generator, survey, N_grid, fast, testmethod):
     # NUV_thresh = np.geomspace(300.0, 380.0, N_grid)
     # NUV_thresh = np.geomspace(30.0, 3000.0, N_grid)
     # NUV_thresh = np.geomspace(200.0, 600.0, N_grid)
-    NUV_thresh = np.geomspace(20.0, 800.0, N_grid)
+    NUV_thresh = np.geomspace(10.0, 1000.0, N_grid)
 
     from bioverse.analysis import test_hypothesis_grid
 
@@ -421,11 +421,12 @@ def past_uv(
         elif N_grid:
             N_grid = N_grid
         else:
-            N_grid = 8
+            N_grid = 10
         nautilus = create_survey_nautilus()
         grid = hypotest_grid(
             g, nautilus, N_grid=N_grid, fast=fast, testmethod=testmethod
         )
+        d = None
         detected = None
         data = None
     # elif powergrid:
@@ -463,33 +464,33 @@ def past_uv(
         except KeyError:
             save_var_latex("p_{}".format(hoststars), results["p"])
 
-    # save some variables for the manuscript
-    if hoststars == "FGK":
-        save_var_latex("uv_inhabited_FGK", len(dd[dd.inhabited]))
+        # save some variables for the manuscript
+        if hoststars == "FGK":
+            save_var_latex("uv_inhabited_FGK", len(dd[dd.inhabited]))
 
-        # general
-        save_var_latex("d_max", g_args["d_max"])
-        save_var_latex("M_G_max", g_args["M_G_max"])
-        save_var_latex("M_st_max", g_args["M_st_max"])
-        try:
-            save_var_latex("f_life", params_past_uv["f_life"])
-        except KeyError:
+            # general
+            save_var_latex("d_max", g_args["d_max"])
+            save_var_latex("M_G_max", g_args["M_G_max"])
+            save_var_latex("M_st_max", g_args["M_st_max"])
+            try:
+                save_var_latex("f_life", params_past_uv["f_life"])
[…]

### 2024-07-11 src/scripts/inhabited-aafo-nuv_thresh.py
change labels

@@ -147,7 +147,7 @@ def plot_results():
         ax.plot(nuv_thresholds_m, mean_fractions_m, label=f"M  host stars ({lbl})", color='C1', **plotkwargs[lbl])
         ax.fill_between(nuv_thresholds_m, lower_bound_m, upper_bound_m, color=f"{'C1' if transiting else 'gray'}", alpha=0.2)
 
-    ax.set_xlabel("NUV Threshold [erg/s/$cm^2$]")
+    ax.set_xlabel("$F_\mathrm{NUV, min}$ [erg/s/$cm^2$]")
     ax.set_ylabel("Fraction of inhabited planets")
     # ax.set_xscale("log")
     # ax.set_title("Fraction of Inhabited Planets vs. NUV Threshold")

### 2024-07-11 src/scripts/powergrid.py

@@ -10,7 +10,7 @@ plotstyle.styleplots()
 
 def plot_powergrid(grid, fig, ax, **kwargs):
     """Plot the power grid."""
-    labels = ("$F_\mathrm{NUV, min}$", "$f_\mathrm{life}$")
+    labels = ("$F_\mathrm{NUV, min}$ [erg/s/$cm^2$]", "$f_\mathrm{life}$")
 
     # check if the dictionary grid contains 'p' or 'dlnZ' values
     if "p" in grid:

### 2024-07-11 src/scripts/powergrid.py
polish plot

@@ -4,6 +4,8 @@ import plotstyle
 import matplotlib.pyplot as plt
 import cmocean
 from bioverse import plots
+import matplotlib.ticker as ticker
+
 
 plotstyle.styleplots()
 
@@ -26,20 +28,22 @@ def plot_powergrid(grid, fig, ax, **kwargs):
         axes=("NUV_thresh", "f_life"),
         labels=labels,
         # log=(True, False),
+        # log=(False, False),
         log=(True, True),
         show=False,
         fig=fig,
         ax=ax,
-        # zoom_factor=2, smooth_sigma=.5,
-        zoom_factor=0,
-        # smooth_sigma=0.2,
+        zoom_factor=2,
+        # zoom_factor=0,
+        # smooth_sigma=.5,
+        smooth_sigma=0.2,
         # levels=[50, 95],
         levels=None,
         cmap=cmocean.cm.dense_r,
         **kwargs,
     )
 
-    ax.set_ylim(0.0, 1.0)
+    ax.set_ylim(bottom=1e-3)
     return fig, ax
 
 
@@ -52,18 +56,17 @@ for i, (spt, ax) in enumerate(zip(["FGK", "M"], axs)):
         grid = dill.load(f)
         data = dill.load(d)  # do we need this?
 
-    # # WHILE DEBUGGING (and no .dll files are available)
-    # import pickle
-    #
-    # with open(
-    #     paths.data / "pipeline/grid_flife_nuv_{}.pkl".format(spt), "rb"
-    # ) as f, open(paths.data / "pipeline/data_{}.dll".format(spt), "rb") as d:
-    #     grid = pickle.load(f)
-    #     data = dill.load(d)  # do we need this?
-
     cbar = False if i == 0 else True
 
     fig, ax = plot_powergrid(grid, fig, ax, cbar=cbar)
+
+    if i == 1:
+        ax.set_ylabel("")
+
+    ax.xaxis.set_major_formatter(
+        ticker.FuncFormatter(lambda x, pos: "{:0.0f}".format(x))
+    )
+    ax.yaxis.set_major_formatter(ticker.StrMethodFormatter("{x:.3g}"))
     ax.set_title(f"{spt}-type host stars")
 
 

### 2024-07-11 src/scripts/inhabited-aafo-nuv_thresh.py
increase fontsize

@@ -7,6 +7,8 @@ import paths
 import dill
 import numpy as np
 import matplotlib.pyplot as plt
+
+import plotstyle
 from utils import timeit
 
 from tqdm import tqdm
@@ -16,6 +18,8 @@ from bioverse_pipeline import (
     get_params_past_uv,
     generate_generator,
 )
+
+plotstyle.styleplots()
 
 # Set a seed for reproducibility of the overall process
 np.random.seed(42)
@@ -114,11 +118,11 @@ def simulate(transiting=True):
 
 
 def plot_results():
-    fig, ax = plt.subplots(figsize=(10, 6))
+    fig, ax = plt.subplots(figsize=(7, 4.5))
 
     plotkwargs = {
     'transiting' : {'marker' : 'o'},
-    'all' : {'marker' : 's', 'linestyle' : '--', 'alpha' : 0.6},
+    'all' : {'marker' : 's', 'linestyle' : '--', 'alpha' : 0.6, 'linewidth' : 1.5},
     }
 
     for transiting, lbl in zip([True, False], ['transiting', 'all']):
@@ -141,11 +145,11 @@ def plot_results():
 
         # Plot FGK stars
         ax.plot(nuv_thresholds_fgk, mean_fractions_fgk, label=f"FGK host stars ({lbl})", color='C0', **plotkwargs[lbl])
-        ax.fill_between(nuv_thresholds_fgk, lower_bound_fgk, upper_bound_fgk, color=f"{'C0' if transiting else 'gray'}", alpha=0.2)
+        ax.fill_between(nuv_thresholds_fgk, lower_bound_fgk, upper_bound_fgk, color=f"{'C0' if transiting else 'white'}", alpha=0.2)
 
         # Plot M stars
         ax.plot(nuv_thresholds_m, mean_fractions_m, label=f"M  host stars ({lbl})", color='C1', **plotkwargs[lbl])
-        ax.fill_between(nuv_thresholds_m, lower_bound_m, upper_bound_m, color=f"{'C1' if transiting else 'gray'}", alpha=0.2)
+        ax.fill_between(nuv_thresholds_m, lower_bound_m, upper_bound_m, color=f"{'C1' if transiting else 'white'}", alpha=0.2)
 
     ax.set_xlabel("$F_\mathrm{NUV, min}$ [erg/s/$cm^2$]")
     ax.set_ylabel("Fraction of inhabited planets")
@@ -161,7 +165,7 @@ def main():
     for transiting in [True, False]:
[…]

### 2024-07-11 src/tex/ms.tex
describe new plots

@@
         \caption{Fraction of inhabited planets for different threshold \gls{NUV} fluxes under the UV hypothesis if the abiogenesis rate is \var{f_life}.
-Shaded regions denote 90 percent confidence intervals of randomized sample generations.}
+Shaded regions denote 90 percent confidence intervals of randomized sample generations, and dashed lines correspond to samples including non-transiting planets.
+        For all samples, the fraction of inhabited planets drops sharply with increasing threshold \gls{NUV} flux due to the combined effects of never receiving sufficient \gls{NUV} flux for abiogenesis or receiving it before entering the \gls{HZ}.}
         \label{fig:inhabited_aafo_nuv_thresh}
@@
 This fraction decreases sharply with increasing threshold flux, as fewer planets receive sufficient \gls{NUV} flux for abiogenesis.
+Another effect responsible for this drop is that some planets receive the required \gls{NUV} flux only before entering the \gls{HZ} -- this is especially likely for M~dwarfs.
 For the FGK sample, the fraction of inhabited planets drops at lower threshold fluxes than for the M dwarf sample.
@@
         \caption{Statistical power as a function of threshold \gls{NUV} flux and abiogenesis rate.
-For a given sample size (here: $N=\var{N_nautilus}$), the achieved statistical power of the survey is enhanced for higher values of $f_\mathrm{life}$ and intermediate values of $F_\mathrm{NUV, min}$.}
+Even for a large sample (here: $N=\var{N_nautilus}$), a high statistical power of the transit survey requires high abiogenesis rates $f_\mathrm{life}$.
+         Intermediate values of $F_\mathrm{NUV, min}$ are more likely to yield strong evidence than extreme values; the sensitivity of the M~dwarf sample extends into the low \gls{NUV} flux end.
+         }
         \label{fig:powergrid}

### 2024-07-11 src/tex/ms.tex
update conclusions

@@
 \section{Conclusions and future work}
-We have investigated the potential of upcoming exoplanet surveys to test the hypothesis that a minimum past \gls{NUV} flux is required for abiogenesis.
+We have investigated the potential of upcoming exoplanet surveys to test the hypothesis -- motivated by the cyanosulfidic origins-of-life scenario -- that a minimum past \gls{NUV} flux is required for abiogenesis.
 To this end, we first employed a semi-analytical Bayesian analysis to estimate probabilities of obtaining strong evidence for or against this hypothesis.
@@
 \begin{enumerate}
-    \item The required \gls{NUV} radiation in the cyanosulfidic scenario should lead to a correlation between past \gls{NUV} flux and current occurrence of biosignatures that will be observationally testable.
-    \item The required sample size for detecting this correlation depends on the occurrence of life on temperate exoplanets and the distribution of host star spectral types in the sample; in particular their past maximum \gls{NUV} fluxes.
+    \item Under the \gls{NUV} hypothesis, the fraction of inhabited planets in a transit survey is sensitive to the threshold \gls{NUV} flux and is expected to drop sharply for required fluxes above a few hundred~\SI{}{\erg\per\second\per\centi\meter\squared}.
+    \item The \gls{NUV} hypothesis should lead to a correlation between past \gls{NUV} flux and current occurrence of biosignatures that may be observationally testable.
+    \item The required sample size for detecting this correlation depends on the abiogenesis rate on temperate exoplanets and the distribution of host star spectral types in the sample; in particular their past maximum \gls{NUV} fluxes.
 %        A direct imaging mission with the \gls{HWO} will require characterizing XXX planets ...
-    \item If the predicted \gls{NUV} correlation exists, yielding strong evidence for it is likely ($> \SI{90}{\percent}$) for sample sizes $\geq 100$, and a survey strategy that targets extreme values of inferred past \gls{NUV} irradiation.
-    \item Abiogenesis in the cyanosulfidic scenario prefers late host star spectral types due to the higher \gls{NUV} fluxes in their habitable zones.
+    \item If the predicted \gls{NUV} correlation exists, obtaining strong evidence for the hypothesis is likely ($\gtrsim \SI{80}{\percent}$) for sample sizes $\geq 100$ if the abiogenesis rate is high ($\gtrsim \SI{50}{\percent}$) and if no very high \gls{NUV} fluxes are required.
+    A survey strategy that targets extreme values of inferred past \gls{NUV} irradiation increases the diagnostic power.
+    \item Samples of planets orbiting M~dwarfs overall yield higher chances of successfully testing the \gls{NUV} hypothesis and are more likely to yield biosignature detections under this hypothesis.
+%    \item Abiogenesis in the cyanosulfidic scenario prefers late host star spectral types due to the higher \gls{NUV} fluxes in their habitable zones.
 However, due to a wider distribution of maximum past \gls{NUV} fluxes in the \gls{HZ} of FGK stars, a transit survey targeting FGK systems is more likely to obtain strong evidence for or against the \gls{NUV} hypothesis.

### 2024-07-11 src/scripts/bioverse_pipeline.py
switch back to Mann-Whitney

@@ -502,8 +502,8 @@ def past_uv(
 
 
 @timeit
-# def main(fast=False, testmethod="mannwhitney"):
-def main(fast=False, testmethod="dynesty"):
+def main(fast=False, testmethod="mannwhitney"):
+# def main(fast=False, testmethod="dynesty"):
     """Run the Bioverse pipeline."""
     for grid in [False, True]:
     # for grid in [True]:

### 2024-07-11 src/tex/ms.tex
First draft of discussion 4.3

--- 2024-07-10 src/tex/ms.tex
@@
 \begin{equation}
+\label{eq:semian:likelihoodH1}
 P(Y=k|H_1) = \binom{n_{\lambda}}{k}f_\mathrm{life}^k(1-f_\mathrm{life})^{n_{\lambda}-k}.
@@
 \end{figure*}
-The situation improves with larger samples: for $n = \var{semian_Nsamp2}$, \SI{80}{\percent} of random experiments yield true strong evidence under either $H_\mathrm{1}$ or $H_\mathrm{null}$.
+The situation improves with larger samples: for $n = \var{semian_Nsamp2}$, \SI{80}{\percent} \textcolor{red}{of random samples permit conclusive inference (strong true evidence)} under either $H_\mathrm{1}$ or $H_\mathrm{null}$.
 
@@
 So far, we have assumed random, uniform distributions of $f_\mathrm{life}$, $F_\mathrm{NUV, min}$, and $F_\mathrm{NUV}$.
+% Actually it is more than just NUV, it is whatever index that integrates NUV and HZ requiremetns into a single scalar variable for which a threshold exist
 A high biosignature detection rate $f_\mathrm{life}$ increases the evidence (cmp. Equation~\ref{eq:bayes_factor}) but we cannot influence it.
@@
 The distribution of $F_\mathrm{NUV}$ in the planet sample, on the other hand, can be influenced by the survey strategy.
+\textcolor{red}{For instance, one might `cherry-pick' exoplanets for which a biosignature test is performed based on information that is available \textit{a priori} such as \gls{NUV} \citep[generally referred to as `planetary context';][]{catling2018exoplanet} as to maximize the science yield of performing additional observation of exoplanets.}
 A targeted sampling approach could be to favor extreme values of $F_\mathrm{NUV}$ in the sample selection.
@@
 \subsection{Sampling strategy for testing a predicted minimum \gls{NUV} flux}
-We show in Sect.~\ref{sec:results} that the constraining power for testing the hypothesis of a minimum past \gls{NUV} flux required for abiogenesis is sensitive on the occurrence of life, the value of this threshold flux, the sample size, and the distribution of sampled past \gls{NUV} fluxes.
-In particular the last parameter offers an opportunity to optimize the survey strategy: Although constraints on a planet's UV history have generally large uncertainties~\citep[e.g.,][]{Richey-Yowell2023}, the likely maximum flux of a planet can be inferred and used as a proxy.
-Sampling more extreme (low and high) values for the maximum flux increases the probability of obtaining true strong evidence for the hypothesis.
-For large sample sizes $\gtrsim 200$, this strategy can push this probability into the \SI{90}{\percent} range.
-\todo[inline]{and what about the probability of getting true strong evidence for H0?}
+In Sect.~\ref{sec:results}, we show that testing the hypothesis of a minimum past \gls{NUV} flux required for abiogenesis suffers from so-called `nuisance' parameters-- unspecified parameters entering the equation of a statistical test-- that render inference difficult.
+Here, these parameters are the unspecified value of the \gls{NUV} threshold hypothesized to exist under $H_1$, and the unknown probability of detection of biosignatures should life have emerged on a planet $f_{\mathrm{life}}$.
+While the history of a planet's received UV flux is difficult to infer~\citep[e.g.,][]{Richey-Yowell2023}, the maximum value of \gls{NUV} flux that a planet was exposed to during its existence might be used as a proxy.
+Indeed, the distribution of the number of planets with detected biosignature in a particular sample of planets with determined maximum \gls{NUV} values $F_{\mathrm{NUV}}$ depends on both the values of $F_{\mathrm{NUV},min}$ and $f_{\mathrm{life}}$ as shown in equation \ref{eq:semian:likelihoodH1}.
+Avoiding the loss in statistical power due to nuisance parameters is a common difficulty in statistical testing, and is the motivation behind elaborate test statistics.
+In the exercise that we do here, we project what could be a test performed by a future observer who disposes of a sample of exoplanets with known past maximum \gls{NUV} exposure and for which biosignature detection has been attempted.
+This is necessarily reductive as the future observer will have more knowledge about experimental conditions and will therefore be able to use this information to guide hypothesis testing.
+For instance, we have made the choice in Sect.~\ref{sec:results} to consider the total number of detected biosignatures as our summary statistic (equation \ref{eq:semian:likelihoodH1), which is not `sufficient' to infer $F_{\mathrm{NUV},min}$ and $f_{life}$ separately.
+However, by conditioning the Bayes factor to these variables (equation \ref{eq:bayesfactor}), we calculate the probability distribution of the Bayesian evidence in favor of $H_1$.
+In doing so, we may evaluate how evidence depends on the uncertainty over these unknown parameters in general terms, without assuming which particular test a future observer might choose to actually perform over real data when available.
+This rather unrefined analysis allows us to provide the intuitive insight that target selection may strongly affect the conclusiveness of one's future test aimed at assessing whether a planet's exposure to \gls{NUV} affects the probability of abiogenesis.
+In particular, a future observer might prefer prioritize sampling extreme values of this variable in order to maximize statistical power.
+This clashes with observational constraints, as the distribution of planets that we can observe and for which detection of biosignature can be attempted might not be independent from \gls{NUV} flux history as the latter is function of stellar mass.
+Hence, for the future observer, selectivity and sample size are in a trade-off conflict.
+Therefore, quantifying this trade-off conflict in terms of resulting expected evidence yield is relevant in order to design observational strategies.
+Our analysis shows that regardless of selectivity, sample sizes smaller than 50 likely result in inconclusive tests, and that increasing selectivity towards extreme $F_{\mathrm{NUV}}$ offers limited inference gains compared to the uniform case ($s=0$; Fig.
+\ref{fig:semian_selectivity}).
+However, our results show that samples of exoplanets with values of $F_{\mathrm{NUV}}$ even only weakly distributed around intermediate values may prevent inference entirely.
+Hence, we argue that selection of a sample of planets with uniformly or extreme-skewed (as opposed to center-skewed) distributed $F_{\mathrm{NUV}}$  should be considered a requirement of any future attempt at testing the hypothesis that \gls{NUV} flux contributes to determining the likelihood of abiogenesis.
+
+We stress that a future observer will likely be able to leverage standard frequentist testing methods which may be more likely to yield conclusive results (whether in favor or in disfavor of $H_1$).
+For instance, it would make sense when presented with a sample of planets with various values of $F_{\mathrm{NUV}}$ and for which biosigature detection is attempted to categorize them into two classes: those for which a biosignature is detected, and those for which this is not the case.
+Then, the distribution of $F_{\mathrm{NUV}}$ in each of those categories is compared under the null hypothesis that they are the same.
+In this case, a non-parametric test could be preferred, such as a two-sample Kolmogorov-Smirnov test or perhaps a Mann-Whitney $U$ test.
+% Is this what you end up looking at using BIOVERSE ? If so, then that should flow nicely with the rest.
+
+%\todo[inline]{Martin: is the given ref about inference difficulty only, or does it also support the statement that max NUV can be inferred?
+In any case, I beleive that there should be a statement here about whether max past NUV can be more easily inferred, how (star mass ?), and how it relates to a scalar variable pertaining to the probability ogf abiogenesis.
+e.g.
+if max past NUV is below required threshold, then no abiogenesis under H1, but if it is over, there might remain the uncertainty in the timing of this value of NUV ?
+Can we know if the timing matches the planet being in the HZ?}
+
+%We show in Sect.~\ref{sec:results} that the constraining power for testing the hypothesis of a minimum past \gls{NUV} flux required for abiogenesis is sensitive on the occurrence of life, the value of this threshold flux, the sample size, and the distribution of sampled past \gls{NUV} fluxes.
+%In particular the last parameter offers an opportunity to optimize the survey strategy: Although constraints on a planet's UV history have generally large uncertainties~\citep[e.g.,][]{Richey-Yowell2023}, the likely maximum flux of a planet can be inferred and used as a proxy.
+
+%Sampling more extreme (low and high) values for the maximum flux increases the probability of obtaining true strong evidence for the hypothesis.
+%For large sample sizes $\gtrsim 200$, this strategy can push this probability into the \SI{90}{\percent} range.
+%\todo[inline]{and what about the probability of getting true strong evidence for H0?}: AA reply to Martin.
+In BF analyses, the evidences for H0 and H1 are symmetric.
+We just look at the probability of strong true evidence, whether in favor of H0 or H1.
+If that is not what the code is doing, it is what it should be doing
 

### 2024-07-12 src/tex/ms.tex
edit semianalytical changes

@@
 \end{figure*}
-The situation improves with larger samples: for $n = \var{semian_Nsamp2}$, \SI{80}{\percent} \textcolor{red}{of random samples permit conclusive inference (strong true evidence)} under either $H_\mathrm{1}$ or $H_\mathrm{null}$.
+The situation improves with larger samples: for $n = \var{semian_Nsamp2}$, \SI{80}{\percent} of random samples permit conclusive inference (strong true evidence) under either $H_\mathrm{1}$ or $H_\mathrm{null}$.
 
@@
 The same is true for $F_\mathrm{NUV, min}$, where again higher values increase the evidence as the binomial distribution for $H_\mathrm{1}$ gets increasingly skewed and shifted away from the one for $H_\mathrm{null}$.
-The distribution of $F_\mathrm{NUV}$ in the planet sample, on the other hand, can be influenced by the survey strategy.
-\textcolor{red}{For instance, one might `cherry-pick' exoplanets for which a biosignature test is performed based on information that is available \textit{a priori} such as \gls{NUV} \citep[generally referred to as `planetary context';][]{catling2018exoplanet} as to maximize the science yield of performing additional observation of exoplanets.}
-A targeted sampling approach could be to favor extreme values of $F_\mathrm{NUV}$ in the sample selection.
+However, one might `cherry-pick' exoplanets for which a biosignature test is performed based on \textit{a priori} available contextual information~\citep{catling2018exoplanet} in order to maximize the science yield of investing additional resources.
+For instance, the distribution of $F_\mathrm{NUV}$ in the planet sample can be influenced by the survey strategy, and a targeted sampling approach could favor extreme values.% of $F_\mathrm{NUV}$ in the sample selection.
 Figure~\ref{fig:semian_selectivity} shows how the probability of obtaining true strong evidence for $H_\mathrm{1}$ scales with selectivity $s$, where $s\in]-1,1[$ such that $F_\mathrm{NUV} \sim Beta(1/10^s,1/10^s)$.

### 2024-07-12 src/tex/ms.tex
edit discussion section on semianalytical analysis

@@
 
-\subsection{What do we learn from a single biosignature detection?}
-\todo[inline]{discuss constraining power on OOL of a convincing biosignature detection on a single planet, depending on the position of the planet in the parameter space we explored.}
+%\subsection{What do we learn from a single biosignature detection?}
+%\todo[inline]{discuss constraining power on OOL of a convincing biosignature detection on a single planet, depending on the position of the planet in the parameter space we explored.}
 
 \subsection{Sampling strategy for testing a predicted minimum \gls{NUV} flux}
-In Sect.~\ref{sec:results}, we show that testing the hypothesis of a minimum past \gls{NUV} flux required for abiogenesis suffers from so-called `nuisance' parameters-- unspecified parameters entering the equation of a statistical test-- that render inference difficult.
-Here, these parameters are the unspecified value of the \gls{NUV} threshold hypothesized to exist under $H_1$, and the unknown probability of detection of biosignatures should life have emerged on a planet $f_{\mathrm{life}}$.
-While the history of a planet's received UV flux is difficult to infer~\citep[e.g.,][]{Richey-Yowell2023}, the maximum value of \gls{NUV} flux that a planet was exposed to during its existence might be used as a proxy.
+In Sect.~\ref{sec:results-semianalytical}, we show that testing the hypothesis of a minimum past \gls{NUV} flux required for abiogenesis suffers from `nuisance' parameters that render inference difficult.
+Here, these parameters are the unspecified value of the \gls{NUV} threshold hypothesized to exist under $H_1$, and the unknown probability of detectable life emerging on a habitable planet $f_{\mathrm{life}}$.
+While the history of a planet's received UV flux is difficult to infer~\citep[e.g.,][]{Richey-Yowell2023}, the estimated maximum \gls{NUV} flux that a planet was exposed to may be used as a proxy.
 Indeed, the distribution of the number of planets with detected biosignature in a particular sample of planets with determined maximum \gls{NUV} values $F_{\mathrm{NUV}}$ depends on both the values of $F_{\mathrm{NUV},min}$ and $f_{\mathrm{life}}$ as shown in equation \ref{eq:semian:likelihoodH1}.
-Avoiding the loss in statistical power due to nuisance parameters is a common difficulty in statistical testing, and is the motivation behind elaborate test statistics.
-In the exercise that we do here, we project what could be a test performed by a future observer who disposes of a sample of exoplanets with known past maximum \gls{NUV} exposure and for which biosignature detection has been attempted.
-This is necessarily reductive as the future observer will have more knowledge about experimental conditions and will therefore be able to use this information to guide hypothesis testing.
-For instance, we have made the choice in Sect.~\ref{sec:results} to consider the total number of detected biosignatures as our summary statistic (equation \ref{eq:semian:likelihoodH1), which is not `sufficient' to infer $F_{\mathrm{NUV},min}$ and $f_{life}$ separately.
-However, by conditioning the Bayes factor to these variables (equation \ref{eq:bayesfactor}), we calculate the probability distribution of the Bayesian evidence in favor of $H_1$.
+Avoiding a loss in statistical power due to nuisance parameters is a common difficulty in statistical testing, and is the motivation behind elaborate test statistics.
+
+In our semi-analytical analysis (Section~\ref{sec:results-semianalytical}), we project a possible test performed by a future observer equipped with a sample of exoplanets with known past maximum \gls{NUV} exposure for which biosignature detection has been attempted.
+This is necessarily reductive as this observer will have more knowledge about experimental conditions and will therefore be able to use this information to guide hypothesis testing.
+For instance, we have made the choice to consider the total number of detected biosignatures as our summary statistic (Equation~\ref{eq:semian:likelihoodH1}), which is not `sufficient' to infer $F_{\mathrm{NUV},min}$ and $f_{life}$ separately.
+However, by conditioning the Bayes factor to these variables (Equation \ref{eq:bayes_factor}), we calculate the probability distribution of the Bayesian evidence in favor of $H_1$.
 In doing so, we may evaluate how evidence depends on the uncertainty over these unknown parameters in general terms, without assuming which particular test a future observer might choose to actually perform over real data when available.
-This rather unrefined analysis allows us to provide the intuitive insight that target selection may strongly affect the conclusiveness of one's future test aimed at assessing whether a planet's exposure to \gls{NUV} affects the probability of abiogenesis.
-In particular, a future observer might prefer prioritize sampling extreme values of this variable in order to maximize statistical power.
-This clashes with observational constraints, as the distribution of planets that we can observe and for which detection of biosignature can be attempted might not be independent from \gls{NUV} flux history as the latter is function of stellar mass.
-Hence, for the future observer, selectivity and sample size are in a trade-off conflict.
-Therefore, quantifying this trade-off conflict in terms of resulting expected evidence yield is relevant in order to design observational strategies.
-Our analysis shows that regardless of selectivity, sample sizes smaller than 50 likely result in inconclusive tests, and that increasing selectivity towards extreme $F_{\mathrm{NUV}}$ offers limited inference gains compared to the uniform case ($s=0$; Fig.
-\ref{fig:semian_selectivity}).
-However, our results show that samples of exoplanets with values of $F_{\mathrm{NUV}}$ even only weakly distributed around intermediate values may prevent inference entirely.
-Hence, we argue that selection of a sample of planets with uniformly or extreme-skewed (as opposed to center-skewed) distributed $F_{\mathrm{NUV}}$  should be considered a requirement of any future attempt at testing the hypothesis that \gls{NUV} flux contributes to determining the likelihood of abiogenesis.
+From this, we can see that target selection can strongly affect the conclusiveness of a future test of the UV hypothesis.
 
-We stress that a future observer will likely be able to leverage standard frequentist testing methods which may be more likely to yield conclusive results (whether in favor or in disfavor of $H_1$).
-For instance, it would make sense when presented with a sample of planets with various values of $F_{\mathrm{NUV}}$ and for which biosigature detection is attempted to categorize them into two classes: those for which a biosignature is detected, and those for which this is not the case.
-Then, the distribution of $F_{\mathrm{NUV}}$ in each of those categories is compared under the null hypothesis that they are the same.
-In this case, a non-parametric test could be preferred, such as a two-sample Kolmogorov-Smirnov test or perhaps a Mann-Whitney $U$ test.
+%The particular finding that our future observer might prefer to prioritize extreme values of this variable in order to maximize statistical power.
+The particular finding that prioritizing extreme values of past \gls{NUV} flux can enhance statistical power clashes with observational constraints, as the composition of the subset of planets that we can observe and for which detection of biosignature can be attempted is not independent from their \gls{NUV} flux history.
+Hence, for our future observer, selectivity and sample size are in conflict.
+This trade-off can be quantified in terms of expected evidence yield, which we have done in Section~\ref{sec:results-semianalytical}.
+Our analysis shows that regardless of selectivity, sample sizes smaller than 50 likely result in inconclusive tests, and that increasing selectivity towards extreme $F_{\mathrm{NUV}}$ offers limited inference gains compared to the uniform case ($s=0$; Figure \ref{fig:semian_selectivity}).
+For larger samples, however, a narrow distribution of $F_{\mathrm{NUV}}$ may prevent inference entirely.
+We thus argue that selecting a sample with $F_{\mathrm{NUV}}$ distributed uniformly or emphasizing extreme values should -- barren any practical counterarguments -- be considered in any future attempt at testing the UV hypothesis.
+Since the practical implementation of an exoplanet survey can stand in the way of such a selection, the following discussion focuses on the results of our transit survey simulations with Bioverse.
+
+%We stress that a future observer will likely be able to leverage standard frequentist testing methods which may be more likely to yield conclusive results (whether in favor or in disfavor of $H_1$).
+%For instance, it would make sense when presented with a sample of planets with various values of $F_{\mathrm{NUV}}$ and for which biosigature detection is attempted to categorize them into two classes: those for which a biosignature is detected, and those for which this is not the case.
+%Then, the distribution of $F_{\mathrm{NUV}}$ in each of those categories is compared under the null hypothesis that they are the same.
+%In this case, a non-parametric test could be preferred, such as a two-sample Kolmogorov-Smirnov test or perhaps a Mann-Whitney $U$ test.
 % Is this what you end up looking at using BIOVERSE ? If so, then that should flow nicely with the rest.
-
-%\todo[inline]{Martin: is the given ref about inference difficulty only, or does it also support the statement that max NUV can be inferred?
-In any case, I beleive that there should be a statement here about whether max past NUV can be more easily inferred, how (star mass ?), and how it relates to a scalar variable pertaining to the probability ogf abiogenesis.
-e.g.
-if max past NUV is below required threshold, then no abiogenesis under H1, but if it is over, there might remain the uncertainty in the timing of this value of NUV ?
-Can we know if the timing matches the planet being in the HZ?}
-
-%We show in Sect.~\ref{sec:results} that the constraining power for testing the hypothesis of a minimum past \gls{NUV} flux required for abiogenesis is sensitive on the occurrence of life, the value of this threshold flux, the sample size, and the distribution of sampled past \gls{NUV} fluxes.
-%In particular the last parameter offers an opportunity to optimize the survey strategy: Although constraints on a planet's UV history have generally large uncertainties~\citep[e.g.,][]{Richey-Yowell2023}, the likely maximum flux of a planet can be inferred and used as a proxy.
-
-%Sampling more extreme (low and high) values for the maximum flux increases the probability of obtaining true strong evidence for the hypothesis.
-%For large sample sizes $\gtrsim 200$, this strategy can push this probability into the \SI{90}{\percent} range.
-%\todo[inline]{and what about the probability of getting true strong evidence for H0?}: AA reply to Martin.
-In BF analyses, the evidences for H0 and H1 are symmetric.
-We just look at the probability of strong true evidence, whether in favor of H0 or H1.
-If that is not what the code is doing, it is what it should be doing
 

### 2024-07-12 src/scripts/inhabited-aafo-nuv_thresh.py
launch production run for inhabited-aafo-nuv_thresh grid

@@ -51,21 +51,20 @@ def run_simulation(nuv_thresh, n_planets, star_type, seed, generator_kwargs):
 def simulate(transiting=True):
     # set_params
     n_planets = 500   # maximum number of planets to sample
-    # n_nuv_thresholds = 10
-    n_nuv_thresholds = 7
-    # n_simulations_per_threshold = 10
-    n_simulations_per_threshold = 4
+    n_nuv_thresholds = 13
+    # n_nuv_thresholds = 7
+    n_simulations_per_threshold = 10
+    # n_simulations_per_threshold = 4
 
     # nuv_thresholds = np.geomspace(30.0, 3000.0, n_nuv_thresholds)
     # nuv_thresholds = np.geomspace(200.0, 600.0, n_nuv_thresholds)
     nuv_thresholds = np.linspace(10.0, 800.0, n_nuv_thresholds)
     # f_life = 1.0
 
-    # DEBUG
     generator_kwargs = {
     # 'f_life' : 0.8,
     'f_life' : 1.0,
-    'deltaT_min' : 1., # Myr. Smaller values seems so slightly enhance #inhabited planets in the FGK sample.
+    'deltaT_min' : 10., # Myr. Smaller values seems so slightly enhance #inhabited planets in the FGK sample.
 
     # increase sample size
     'f_eta' : 15,
@@ -155,6 +154,8 @@ def plot_results():
     ax.set_ylabel("Fraction of inhabited planets")
     # ax.set_xscale("log")
     # ax.set_title("Fraction of Inhabited Planets vs. NUV Threshold")
+    ax.set_xlim(left=-8)
+    ax.set_ylim(bottom=-0.015)
     ax.legend()
     fig.show()
     fig.savefig(paths.figures / "inhabited_aafo_nuv_thresh.pdf")

### 2024-07-12 src/tex/ms.tex
swap conclusions, add survey sim sentence

@@
 
-\section{Conclusions and future work}
+\section{Conclusions}
 We have investigated the potential of upcoming exoplanet surveys to test the hypothesis -- motivated by the cyanosulfidic origins-of-life scenario -- that a minimum past \gls{NUV} flux is required for abiogenesis.
@@
 \begin{enumerate}
-    \item Under the \gls{NUV} hypothesis, the fraction of inhabited planets in a transit survey is sensitive to the threshold \gls{NUV} flux and is expected to drop sharply for required fluxes above a few hundred~\SI{}{\erg\per\second\per\centi\meter\squared}.
-    \item The \gls{NUV} hypothesis should lead to a correlation between past \gls{NUV} flux and current occurrence of biosignatures that may be observationally testable.
-    \item The required sample size for detecting this correlation depends on the abiogenesis rate on temperate exoplanets and the distribution of host star spectral types in the sample; in particular their past maximum \gls{NUV} fluxes.
+    \item The UV~flux requirement in the cyanosulfidic scenario for the origins of life should lead to a correlation between past \gls{NUV} flux and current occurrence of biosignatures that may be observationally testable.
+    \item The required sample size for detecting this correlation depends on the abiogenesis rate on temperate exoplanets and the distribution of host star properties in the sample; in particular their maximum past \gls{NUV} fluxes.
 %        A direct imaging mission with the \gls{HWO} will require characterizing XXX planets ...
-    \item If the predicted \gls{NUV} correlation exists, obtaining strong evidence for the hypothesis is likely ($\gtrsim \SI{80}{\percent}$) for sample sizes $\geq 100$ if the abiogenesis rate is high ($\gtrsim \SI{50}{\percent}$) and if no very high \gls{NUV} fluxes are required.
+    \item Under the UV hypothesis, the fraction of inhabited planets in a transit survey is sensitive to the threshold \gls{NUV} flux and is expected to drop sharply for required fluxes above a few hundred~\SI{}{\erg\per\second\per\centi\meter\squared}.
+    \item If the predicted UV~correlation exists, obtaining strong evidence for the hypothesis is likely ($\gtrsim \SI{80}{\percent}$) for sample sizes $\geq 100$ if the abiogenesis rate is high ($\gtrsim \SI{50}{\percent}$) and if no very high \gls{NUV} fluxes are required.
     A survey strategy that targets extreme values of inferred past \gls{NUV} irradiation increases the diagnostic power.
-    \item Samples of planets orbiting M~dwarfs overall yield higher chances of successfully testing the \gls{NUV} hypothesis and are more likely to yield biosignature detections under this hypothesis.
+    \item Samples of planets orbiting M~dwarfs overall yield higher chances of successfully testing the \gls{NUV} hypothesis.
+        Leaving other factors aside, they are also more likely to yield biosignature detections under this hypothesis.
+
 %    \item Abiogenesis in the cyanosulfidic scenario prefers late host star spectral types due to the higher \gls{NUV} fluxes in their habitable zones.
@@
 More generally, we found that models of the origins of life provide hypotheses that may be testable with near-future exoplanet surveys.
+Conducting realistic survey simulations with representative samples is key to making accurate predictions, developing the optimal survey strategy, and increasing science return.
 Our work highlights the importance of understanding the context in which a biosignature detection is made, which can not only help to assess the credibility of the detection but also to test competing theories of the origins of life on Earth and beyond.

### 2024-07-12 src/tex/ms.tex
rephrase H1, Hnull, priors

@@
 We assigned the development of life to a random fraction $f_\mathrm{life}$ of all temperate planets fulfilling these requirements.
-
-For the hypothesis tests, we then define our alternative hypothesis as
+For the probability of a planet having detectable biosignatures, $P(\mathrm{bio})$, the UV~hypothesis then states
 \begin{equation}\label{eq:hypothesis-uv}
-    H_{1} = f_\mathrm{life} (\theta, F_\mathrm{NUV}) =
+    H_{1} : P(\mathrm{bio}) =
         \begin{cases}
@@
 %\begin{equation}
-$H_{null} = f_\mathrm{life} (\theta),$
+$H_\mathrm{null} : P(\mathrm{bio}) = f_\mathrm{life},$
 %\end{equation}
 i.e., no correlation with UV flux.
-We imposed log-uniform priors on the parameters $\boldsymbol{\theta}$, sampling $f_\mathrm{life}$ from a log-uniform distribution between $10^{-3}$ and $1$ and $F_\mathrm{NUV, min}$ from a log-uniform distribution between $10^{1}$ and $10^{5}$.
+Our choice of priors was to sample $f_\mathrm{life}$ from a log-uniform distribution between $10^{-3}$ and $1$ and $F_\mathrm{NUV, min}$ from a log-uniform distribution between $10^{1}$ and $10^{5}$.
 

### 2024-07-12 src/tex/ms.tex
add intro for Discussion

@@
 \label{sec:discussion}
-%''Which natural processes best explain how living matter spontaneously appears from nonliving matter?''~\citep{Malaterre2022}
-%WE consider life ``as we know it''
-\subsection{Constraining power for the origins of life as a function of biosignature location}
-\todo[inline]{How does the location of biosignature detections impact the credibility of OOL scenarios?}
-\todo[inline]{biosignature on M~dwarf planet vs. FGK: Impact on UV flux requirement?}
-The cyanosulfidic scenario, in particular its predicted existence of a minimum \gls{NUV} flux required for prebiotic chemistry, offers an opportunity to test an origins of life hypothesis with a statistical transit survey sampling planets with varying \gls{NUV} flux histories.
-It comes to no surprise that the success rate of such a test is sensitive to the sample size of the survey and to the occurrence of life on temperate exoplanets.
-As we have shown, the statistical power of this test also depends on the distribution of past \gls{NUV} fluxes in the sample and on the required threshold flux.
-Optimizing the survey to sample a wide range of NUV flux values, particularly at the extremes, can enhance the likelihood of obtaining strong evidence for or against the hypothesis.
-Intermediate values of the threshold \gls{NUV} flux are more likely to yield strong evidence than extreme values, as the dataset under the alternative hypothesis $H_1$ differs more from the null hypothesis in this case while still being sufficiently populated.
-The required threshold flux is, of course, a priori unknown and we cannot influence it.
-If, however, better theoretical predictions for the required \gls{NUV} flux for abiogenesis become available, the survey strategy can be further optimized, for instance by targeting planets that are estimated to have received a \gls{NUV} flux slightly below and above this threshold.
-
-An interesting  aspect lies in the distribution of host star spectral types:
-Under the \gls{NUV} hypothesis, the occurrence of life is expected to correlate with the host star's spectral type, with late-type M~dwarfs being favored due to their higher \gls{NUV} fluxes in the \gls{HZ}~\citep{Richey-Yowell2023}.
-However, the distribution of maximum past \gls{NUV} fluxes in this sample is narrow in this sample, which may limit the constraining power of the survey depending on the (unknown) threshold \gls{NUV} flux.
-An M~dwarf sample may help to test the high \gls{NUV} flux end of the hypothesis; a higher occurrence of biosignatures here would support the hypothesis that a higher NUV flux is beneficial or necessary for life.
-
-FGK stars, on the other hand, show a wider distribution of maximum past \gls{NUV} fluxes in the \gls{HZ}, which may increase the likelihood of obtaining strong evidence for or against the hypothesis.
-With this sample, the survey will be more sensitive to the low \gls{NUV} flux end of the hypothesis.
-A lack of biosignatures on these planets would support the \gls{NUV} hypothesis, whereas their presence might suggest that lower NUV fluxes are also sufficient for abiogenesis or indicate different abiogenic pathways.
-CITE
-
-%\subsection{What do we learn from a single biosignature detection?}
-%\todo[inline]{discuss constraining power on OOL of a convincing biosignature detection on a single planet, depending on the position of the planet in the parameter space we explored.}
+A key question in the quest to understand the origins of life is which natural processes best explain how living matter spontaneously appears from nonliving matter~\citep{Malaterre2022}.
+The cyanosulfidic scenario~\citep{Patel2015}, in particular its predicted existence of a minimum \gls{NUV} flux required for prebiotic chemistry, seems to offer an opportunity to test an origins of life hypothesis with a statistical transit survey sampling planets with varying \gls{NUV} flux histories.
+In the following, we discuss the prospects of testing the UV hypothesis in light of our results.
 

### 2024-07-12 src/tex/ms.tex
rewrite spectral type dependency

@@
 
+\subsection{Constraining power for the origins of life as a function of biosignature location}
+\todo[inline]{How does the location of biosignature detections impact the credibility of OOL scenarios?}
+\todo[inline]{biosignature on M~dwarf planet vs. FGK: Impact on UV flux requirement?}
+It comes to no surprise that the success rate of such a test is sensitive to the sample size of the survey and to the occurrence of life on temperate exoplanets.
+As we have shown, the statistical power of this test also depends on the distribution of past \gls{NUV} fluxes in the sample and on the required threshold flux.
+Optimizing the survey to sample a wide range of NUV flux values, particularly at the extremes, can enhance the likelihood of obtaining strong evidence for or against the hypothesis.
+Intermediate values of the threshold \gls{NUV} flux are more likely to yield strong evidence than extreme values, as the dataset under the alternative hypothesis $H_1$ differs more from the null hypothesis in this case while still being sufficiently populated.
+The required threshold flux is, of course, a priori unknown and we cannot influence it.
+If, however, better theoretical predictions for the required \gls{NUV} flux for abiogenesis become available, the survey strategy can be further optimized, for instance by targeting planets that are estimated to have received a \gls{NUV} flux slightly below and above this threshold.
+
+An interesting aspect lies in the distribution of host star properties, as different spectral types probe different past \gls{NUV} flux regimes.
+Under the UV~hypothesis, late-type M~dwarfs are favored due to their higher \gls{NUV} fluxes in the \gls{HZ}~\citep{Richey-Yowell2023}.
+Furthermore, their wider distribution of maximum past \gls{NUV} fluxes may increase the likelihood of obtaining strong evidence for or against the hypothesis.
+An M~dwarf sample may help to test the high \gls{NUV} flux end of the hypothesis; a higher occurrence of biosignatures here would support the hypothesis that a higher NUV flux is beneficial or necessary for life.
+
+FGK stars, on the other hand, show a more narrow distribution of maximum past \gls{NUV} fluxes in the \gls{HZ}, which may limit the constraining power of the survey depending on the (unknown) threshold \gls{NUV} flux.
+With this sample, the survey will be more sensitive to the low \gls{NUV} flux end of the hypothesis.
+A lack of biosignatures on these planets would support the UV~hypothesis, whereas their presence might suggest that lower NUV fluxes are also sufficient for abiogenesis or indicate a different abiogenic pathway.
+CITE
+
+%\subsection{What do we learn from a single biosignature detection?}
+%\todo[inline]{discuss constraining power on OOL of a convincing biosignature detection on a single planet, depending on the position of the planet in the parameter space we explored.}
+
 \subsection{Contextual support for potential biosignature detections}

### 2024-07-12 src/tex/ms.tex
minor stuff

@@
         \includegraphics[width=\hsize]{figures/nuv_distribution}
-        \caption{Distribution of maximum past \gls{NUV} flux in transit surveys targeting FGK and M~stars, respectively.
+        \caption{Distribution of maximum past \gls{NUV} flux in transit surveys targeting \glspl{EEC} around FGK and M~stars, respectively.
 The best-fit beta distributions (gray) correspond to selectivities of $s_\mathrm{FGK} = $\var{selectivity_FGK} and $s_\mathrm{M} = $\var{selectivity_M}.
@@
 
-\subsection{Sampling strategy for testing a predicted minimum \gls{NUV} flux}
+\subsection{Sampling strategy for testing a predicted minimum \gls{NUV} flux} % corresponds to conclusion point 1
 In Sect.~\ref{sec:results-semianalytical}, we show that testing the hypothesis of a minimum past \gls{NUV} flux required for abiogenesis suffers from `nuisance' parameters that render inference difficult.
@@
 
-\subsection{Constraining power for the origins of life as a function of biosignature location}
+\subsection{Constraining power for the origins of life as a function of biosignature location} % corresponds to conclusion point 2
 \todo[inline]{How does the location of biosignature detections impact the credibility of OOL scenarios?}
-\todo[inline]{biosignature on M~dwarf planet vs. FGK: Impact on UV flux requirement?}
 It comes to no surprise that the success rate of such a test is sensitive to the sample size of the survey and to the occurrence of life on temperate exoplanets.
@@
 
-\subsection{Contextual support for potential biosignature detections}
+\subsection{Contextual support for potential biosignature detections} % corresponds to ... which conclusion?
 \todo[inline]{discuss how planetary context may impact credibility of tentative biosignature detections (e.g., if there is a good fit with a predicted OOL pattern; or the opposite: the planetary context does not fit well to any OOL scenarios)}
@@
     A survey strategy that targets extreme values of inferred past \gls{NUV} irradiation increases the diagnostic power.
-    \item Samples of planets orbiting M~dwarfs overall yield higher chances of successfully testing the \gls{NUV} hypothesis.
+    \item Samples of planets orbiting M~dwarfs overall yield higher chances of successfully testing the UV~hypothesis.
         Leaving other factors aside, they are also more likely to yield biosignature detections under this hypothesis.
@@
 %    \item Abiogenesis in the cyanosulfidic scenario prefers late host star spectral types due to the higher \gls{NUV} fluxes in their habitable zones.
-However, due to a wider distribution of maximum past \gls{NUV} fluxes in the \gls{HZ} of FGK stars, a transit survey targeting FGK systems is more likely to obtain strong evidence for or against the \gls{NUV} hypothesis.
+However, due to a wider distribution of maximum past \gls{NUV} fluxes in the \gls{HZ} of FGK stars, a transit survey targeting FGK systems is more likely to obtain strong evidence for or against the UV~hypothesis.
 

### 2024-07-13 src/tex/ms.tex
discuss sensitivity to astrophysical parameters

@@
 
-\subsection{Constraining power for the origins of life as a function of biosignature location} % corresponds to conclusion point 2
+\subsection{Constraining power for the origins of life as a function of biosignature location} % corresponds to conclusion points 2, 5
 \todo[inline]{How does the location of biosignature detections impact the credibility of OOL scenarios?}
@@
 CITE
+
+\subsection{Sensitivity to astrophysical parameters} % corresponds to conclusion point 3
+Our Bioverse simulations that take into account exoplanet demographics, the evolution of habitability and \gls{NUV} fluxes, and observational biases show that not only the likelihood of a conclusive test of the UV~hypothesis, but also the likelihood of successful biosignature detection itself is extremely sensitive to the threshold \gls{NUV} flux if the UV~hypothesis is true.
+Even if all biosignatures can be detected and the nominal abiogenesis rate is very high, say $f_\mathrm{life} = 1$, under the condition that prebiotic chemistry requires a minimum \gls{NUV} flux \textit{and} liquid water, the probability of finding life on a randomly selected planet may be very low.
+As we showed, for high required fluxes the two requirements of simultaneous \gls{HZ} occupancy and sufficient \gls{NUV} flux conspire to diminish the fraction of inhabited planets in the sample.
+Taking the infered fluxes from \citet{Richey-Yowell2023} at face value (but taking into account intrinsic scatter), a minimum required \gls{NUV} flux of $\gtrsim\SI{400}{\erg\per\second\per\centi\meter\squared}$ reduces the fraction of inhabited planets to below $\sim\SI{1}{\percent}$.
+This not only calls for a large sample size and a targeted sample selection prefering high expected past \gls{NUV} fluxes, but also highlight the necessity of continued theoretical and experimental research into the role of UV radiation in prebiotic chemistry~\citep{Ranjan2017b}.
 

### 2024-07-13 src/tex/ms.tex
discuss contextual support for biosignature detections

@@
 
-\subsection{Survey simulations with Bioverse}
+\subsection{Survey simulations with Bioverse}\label{sec:results-bioverse}
 
@@
 \todo[inline]{discuss how planetary context may impact credibility of tentative biosignature detections (e.g., if there is a good fit with a predicted OOL pattern; or the opposite: the planetary context does not fit well to any OOL scenarios)}
-As we have shown in Section~\ref{sec:results}, the interplay of \gls{NUV} evolution and \gls{HZ} occupancy strongly favors late spectral types for abiogenesis via the cyanosulfidic scenario.
-This strong predicted correlation between stellar spectral type and the occurrence of life can be used to falsify \gls{ool} scenarios.
-The constraint of this context is particularly strong if a candidate biosignature is detected on a planet orbiting an earlier type star, i.e., where it is unexpected in the context of this \gls{ool} scenario.
+The predicted interplay of \gls{NUV} flux and \gls{HZ} occupancy in enabling abiogenesis via the cyanosulfidic scenario could in principle be used to add or remove credibility from a tentative biosignature detection.
+For example, with a strong believe that this scenario is the only viable one for the origins of life, a biosignature detection on a planet orbiting a strongly UV-radiating star may add credibility to the detection.
+Conversely, a biosignature detection on a planet estimated to have received very little UV~radiation would increase the likelihood of a false positive detection.
+On the other hand, should the detection in the latter case be confirmed, it could be used to falsify the UV~hypothesis.
+
+Unfortunately, our simulations do not yield a clear criterion for the credibility of a biosignature detection based on spectral type of the host star, as both FGK and M~dwarf samples show similar maximum past \gls{NUV} flux distributions.
+The abiogenesis rate in both samples show similar trends with the threshold \gls{NUV} flux, and the fraction of inhabited planets drops at similar threshold fluxes (see Section~\ref{sec:results-bioverse}).
+A potentially inhabited planet's host star spectral type may thus not be a strong indicator for the credibility of a biosignature detection in the context of the UV~hypothesis.
 
@@
 \todo[inline]{SUKRIT please specify as neccessary}
-
-\subsubsection{Stellar flares}
-Our assumptions on past UV flux neglect the contribution of stellar flares, which may be hypothesized as an alternative source of UV light~\citep{Ranjan2017}.
-This concerns mainly ultracool dwarfs, due to their low quiescent emission and high pre-main sequence stellar activity~\citep{Buccino2007,West2008}.
-Recent work indicates that the majority of stars show inadequate activity levels for a sufficient contribution through flares~\citep{Glazier2020,Ducrot2020,Guenther2020}.
-The biosignature surveys we simulated here may test the hypothesis of sufficient UV radiation via stellar flares.
 

### 2024-07-13 src/tex/ms.tex
edit caveats

@@
 \subsection{Caveats}
+Our work is based on a number of assumptions and simplifications that may affect the results and conclusions.
+We discuss some of these caveats here.
+
+\subsubsection{Stellar flares}
+Our assumptions on past UV flux neglect the contribution of stellar flares, which may be hypothesized as an alternative source of UV light~\citep{Ranjan2017}.
+This concerns mainly ultracool dwarfs, due to their low quiescent emission and high pre-main sequence stellar activity~\citep{Buccino2007,West2008}.
+However, recent work indicates that the majority of stars show inadequate activity levels for a sufficient contribution through flares~\citep{Glazier2020,Ducrot2020,Guenther2020}.
+%The biosignature surveys we simulated here may test the hypothesis of sufficient UV radiation via stellar flares.
+
 \subsubsection{Atmosphere transmission}

### 2024-07-13 src/tex/ms.tex
discuss overall prospects

@@
 
-\subsubsection{Statistical power for a transit survey and sensitivity on astrophysical parameters}
+\subsubsection{Statistical power for a transit survey and sensitivity on astrophysical parameters}\label{sec:results-powergrid}
 We now investigate the sensitivity of the achieved statistical power of our default transit survey to the a priori unconstrained threshold \gls{NUV} flux $F_\mathrm{NUV, min}$ and the abiogenesis rate $f_\mathrm{life}$.
@@
 
+\subsection{Overall prospects for testing the UV hypothesis} % corresponds to conclusion point 4
+Our results show that the UV hypothesis is testable with upcoming exoplanet surveys, but that the success of such a test depends on the sample size, the distribution of past \gls{NUV} fluxes, and several unknown astrophysical nuisance parameters.
+Even under idealized conditions, obtaining strong evidence for or against the hypothesis likely requires sample sizes on the order of 100 (see Section~\ref{sec:results-semianalytical}).
+This is true for a future transit survey, the specifics of which we have reflected in our Bioverse simulations (see Section~\ref{sec:results-bioverse}).
+However, we have shown that the impacts from the combined requirements of the UV~hypothesis on the fraction of inhabited planets in a sample are comparable in the non-transiting case.
+The likely required sample sizes are larger than the projected sample sizes of most planned missions; for example the \gls{HWO} expects to characterize a sample of only $\sim 25$ Earth analogs~\citep{Mamajek2023,Tuchow2024}.
+
+Meaningful tests of the UV hypothesis may thus require a different survey strategy that increases the sample of characterized planets.
+Such concepts, like the Nautilus mission~\citep{Apai2019,Apai2022}, which aims to characterize up to $\sim 1000$ rocky planets via utilizing an innovative telescope design, are only appearing on the horizon.
+Constraints on the origins of life from exoplanet surveys may thus be a long-term endeavor, and the UV hypothesis may not be testable yet with the current and next generation of exoplanet surveys.
+In the meantime, it is important to refine predictions on the role of UV radiation in prebiotic chemistry with both theoretical and experimental work.
+
 \subsection{Caveats}
@@
 \todo[inline]{SUKRIT please specify as neccessary}
-
 %\subsubsection{Diurnal cycles}
@@
 
-Overall, our work demonstrates that upcoming exoplanet surveys have the potential to test the hypothesis that a minimum past \gls{NUV} flux is required for abiogenesis.
-More generally, we found that models of the origins of life provide hypotheses that may be testable with near-future exoplanet surveys.
+Overall, our work demonstrates that future exoplanet surveys have the potential to test the hypothesis that a minimum past \gls{NUV} flux is required for abiogenesis.
+More generally, we found that models of the origins of life provide hypotheses that may be testable with these surveys.
 Conducting realistic survey simulations with representative samples is key to making accurate predictions, developing the optimal survey strategy, and increasing science return.

### 2024-07-13 src/tex/ms.tex
rewrite powergrid results

@@
 Values of $F_\mathrm{NUV, min}$ that lie between the extrema of the inferred maximum \gls{NUV} flux increase the achieved statistical power of the survey, as in this case the dataset under the alternative hypothesis $H_1$ differs more from the null hypothesis.
-The same is true for the abiogenesis rate $f_\mathrm{life}$, where higher values increase the evidence for $H_1$.
+Furthermore, a higher abiogenesis rate $f_\mathrm{life}$ increases the evidence for $H_1$.
+
+Notably, the sensitivity of the M~dwarf sample extends into the low \gls{NUV} flux end due to the broader distribution of maximum past \gls{NUV} fluxes in this sample.
+Here, the FGK sample is less sensitive.
+The only parameter space region with statistical power above \SI{90}{\percent} lies at abiogenesis rates $f_\mathrm{life} > 0.9$ and threshold \gls{NUV} fluxes of $\sim \SIrange{30}{70}{\erg\per\second\per\centi\meter\squared}$.
 
@@
 %In this case, a non-parametric test could be preferred, such as a two-sample Kolmogorov-Smirnov test or perhaps a Mann-Whitney $U$ test.
-% Is this what you end up looking at using BIOVERSE ? If so, then that should flow nicely with the rest.
 
 \subsection{Constraining power for the origins of life as a function of biosignature location} % corresponds to conclusion points 2, 5
-\todo[inline]{How does the location of biosignature detections impact the credibility of OOL scenarios?}
 It comes to no surprise that the success rate of such a test is sensitive to the sample size of the survey and to the occurrence of life on temperate exoplanets.
@@
 \subsection{Contextual support for potential biosignature detections} % corresponds to ... which conclusion?
-\todo[inline]{discuss how planetary context may impact credibility of tentative biosignature detections (e.g., if there is a good fit with a predicted OOL pattern; or the opposite: the planetary context does not fit well to any OOL scenarios)}
 The predicted interplay of \gls{NUV} flux and \gls{HZ} occupancy in enabling abiogenesis via the cyanosulfidic scenario could in principle be used to add or remove credibility from a tentative biosignature detection.

### 2024-07-13 src/scripts/nuv_distribution.py
Center, reduce, then scale data

@@ -13,6 +13,8 @@ from scipy import stats
 import numpy as np
 
 def normalize_data(data):
+    """Center, reduce, and scale data to [0, 1] range."""
+    data = (data - data.mean()) / data.std()
     min_val = data.min()
     max_val = data.max()
     return (data - min_val) / (max_val - min_val)
@@ -24,11 +26,11 @@ def plot_nuv_distribution(sample, data, fig, ax, spt):
 
     # force-fit a beta distribution to our NUV_max
     # max_likeli = stats.beta.fit(dataa.max_nuv, method='MLE')
-    max_likeli = stats.beta.fit(normalized_data, method="MM")
-    print(f"fit parameters: {max_likeli}")
+    max_likeli_norm = stats.beta.fit(normalized_data, method="MM")
+    print(f"fit parameters: {max_likeli_norm}")
 
     # estimate selectivity from average of fitted beta function parameters
-    selectivity = np.log10(1 / (np.mean(max_likeli[:2])))
+    selectivity = np.log10(1 / (np.mean(max_likeli_norm[:2])))
     save_var_latex("selectivity_{}".format(spt), round(selectivity, 1))
     print(f"selectivity ~ {selectivity:.2f}")
 
@@ -72,4 +74,5 @@ for spt, ax in zip(['FGK', 'M'], axs):
     fig, ax = plot_nuv_distribution(sample, data, fig, ax, spt)
     ax.set_title(f"{spt}-type host stars")
 
+plt.show()
 fig.savefig(paths.figures / "nuv_distribution.pdf")

### 2024-07-13 src/tex/ms.tex
remove references to Bayesian tests/nested sampling

@@
 A real exoplanet survey will be subject to observational biases, sample selection effects, and the underlying demographics of the planet sample.
-To assess the information gain of a realistic exoplanet survey, we employed Bioverse~\citep{Bixel2021}, a framework that integrates multiple components including statistically realistic simulations of exoplanet populations, a survey simulation module, and a Bayesian hypothesis testing module to evaluate the statistical power of different observational strategies.
+To assess the information gain of a realistic exoplanet survey, we employed Bioverse~\citep{Bixel2021}, a framework that integrates multiple components including statistically realistic simulations of exoplanet populations, a survey simulation module, and a hypothesis testing module to evaluate the statistical power of different observational strategies.
 The general approach is as follows:
@@
 To model the sensitivity of the information gain of a proposed mission to sample selection and survey strategy, we conduct survey simulations with Bioverse using different sample sizes and survey strategies.
-    \item \textbf{Hypothesis testing:} We employ Bayesian statistical methods to evaluate the likelihood that a given survey would detect a specified statistical trend in the exoplanet population and estimate the precision with which the survey could constrain the parameters of that trend.
+    \item \textbf{Hypothesis testing:} We evaluate the likelihood that a given survey would detect a specified demographic trend in the exoplanet population and estimate the precision with which the survey could constrain the parameters of that trend.
     A common definition of the null hypothesis $H_0$, which is also applied here, is that there is no relationship between the independent variable (here: maximum NUV flux) and the dependent variable (here: biosignature occurrence).
-    The alternative hypothesis $H_1$ proposes a specific relationship parameterized by $\theta$.
-    If the dependent variable is of binary nature such as in the case of biosignature detections, we model the likelihood function as
-    \begin{equation}
-\mathcal{L}(y \mid \boldsymbol{\theta})=\prod_i^N\left[y_i h\left(\boldsymbol{\theta}, x_i\right)+\left(1-y_i\right)\left(1-h\left(\boldsymbol{\theta}, x_i\right)\right)\right],
-\end{equation}
-where here $y_i \in \{0,1\}$ is the biosignature detection variable, $x_i$ is the maximum NUV flux, and $h(\boldsymbol{\theta}, x_i)$ is the probability of detecting a biosignature given the maximum NUV flux and the model parameters $\boldsymbol{\theta}$.
-We then use a nested sampling method~\citep{Speagle2020} to compute the Bayesian evidence for the null and alternative hypotheses and estimate the strength of evidence for the alternative hypothesis.
+%    The alternative hypothesis $H_1$ proposes a specific relationship parameterized by $\theta$.
+    The alternative hypothesis $H_1$ proposes a specific relationship between the independent and dependent variables.
+    Bioverse offers either Bayesian model comparison or non-parametric tests to evaluate the evidence for or against the null hypothesis.
 
@@
 We use this metric, known as the statistical power, to quantify the potential information content of the survey, identify critical design trades, and find strategies that maximize the survey's scientific return.
-The posterior samples obtained from the nested sampling runs further allows us to estimate the precision with which the survey could constrain the parameters of the hypothesized trend.
+% TODO: IF INFERENCE: add:
+%The posterior samples obtained from the nested sampling runs further allows us to estimate the precision with which the survey could constrain the parameters of the hypothesized trend.
 \end{enumerate}
@@
 i.e., no correlation with UV flux.
-Our choice of priors was to sample $f_\mathrm{life}$ from a log-uniform distribution between $10^{-3}$ and $1$ and $F_\mathrm{NUV, min}$ from a log-uniform distribution between $10^{1}$ and $10^{5}$.
+% TODO: IF INFERENCE: add:
+%Our choice of priors was to sample $f_\mathrm{life}$ from a log-uniform distribution between $10^{-3}$ and $1$ and $F_\mathrm{NUV, min}$ from a log-uniform distribution between $10^{1}$ and $10^{5}$.
 
@@
 This leads to a distribution of biosignature detections with detections increasingly occurring above a threshold inferred NUV flux.
-In this example case, the few biosignature detections in the FGK sample lead to a higher evidence ($dlnZ_\mathrm{FGK} = \var{dlnZ_FGK}$) than in the M dwarf sample ($dlnZ_\mathrm{M} = \var{dlnZ_M}$), where the majority of planets are above the threshold \gls{NUV} flux.
+In this example case, the few biosignature detections in the FGK sample lead to a higher evidence %($dlnZ_\mathrm{FGK} = \var{dlnZ_FGK}$)
+than in the M dwarf sample%($dlnZ_\mathrm{M} = \var{dlnZ_M}$)
+, where the majority of planets are above the threshold \gls{NUV} flux.
 

### 2024-07-13 src/tex/ms.tex
Describe hypothesis testing with the Mann-Whitney U test

@@
 
+\subsubsection{Hypothesis testing}
+To evaluate the evidence for correctly rejecting the null hypothesis, we employed the Mann-Whitney U test~\citep{Mann1947}.
+This is a non-parametric test to determine whether two independent samples were drawn from a population with the same distribution and is in particular sensitive to one sample being stochastically greater than the other.
+We used the Mann-Whitney U test to compare the distributions of \gls{NUV} fluxes of planets with and without biosignatures.
+The implementation in Bioverse relies on the \texttt{scipy.stats.mannwhitneyu} function~\citep{Virtanen2020} and returns a two-tailed p-value, for which we set a significance level of $\alpha = 0.05$ to reject the null hypothesis.
+For every hypothesis test, we repeated randomized survey realizations to estimate the fraction of successful rejections of the null hypothesis, i.e., the statistical power of the survey.
+
 \section{Results}

### 2024-07-13 src/tex/ms.tex
remove O2 as biosignature text

@@
 % detecting methanogenic life: Sandora2020, Wang2018, Seeburger2023
-A commonly discussed biosignature is molecular Oxygen (O$_2$), which on Earth emerged as a byproduct of photosynthesis during the Proterozoic era.
-No individual component of an atmosphere has been identified as a reliable biosignature in isolation~\citep{Seager2016}, and the detection of Oxygen alone will certainly not be sufficient to confirm the presence of life.
-Nevertheless, we focus here on detecting this key absorber as it represents the general inherent observational challenges and trends.
+% TODO: IF SPECIFIC BIOSIGNATURE: use:
+%A commonly discussed biosignature is molecular Oxygen (O$_2$), which on Earth emerged as a byproduct of photosynthesis during the Proterozoic era.
+%No individual component of an atmosphere has been identified as a reliable biosignature in isolation~\citep{Seager2016}, and the detection of Oxygen alone will certainly not be sufficient to confirm the presence of life.
+%Nevertheless, we focus here on detecting this key absorber as it represents the general inherent observational challenges and trends.
 

### 2024-07-19 src/tex/ms.tex
upload current draft
move figure to results

@@
 \begin{abstract}
-    tbd.
+    ...
 \end{abstract}
@@
 \todo[inline]{Introduce OOL, the importance of planetary contexts}
-
+...
 \subsection{Origins of Life Scenarios and their Predictions}\label{sec:predictions}
@@
 
-\begin{figure}
-    \script{inhabited_FGKM.py}
-    \begin{centering}
-        \includegraphics[width=\hsize]{figures/inhabited_FGKM}
-        \caption{Host stars of all transiting \glspl{EEC} and inhabited planets in a simulated transit survey.
-        In the FGK sample, most \glspl{EEC} and all inhabited planets orbit K~dwarfs.
-        In an M~dwarf sample of the same size, the fraction of inhabited planets is larger.
-        }
-        \label{fig:inhabited_FGKM}
-    \end{centering}
-\end{figure}
-
-As shown in Figure~\ref{fig:inhabited_FGKM}, the majority of \glspl{EEC} orbit lower-mass stars.
-The fraction of inhabited planets is highest in the M~dwarf sample due to the higher \gls{NUV} fluxes in the \gls{HZ} of these stars.
-
 \begin{figure*}
@@
 \subsection{Survey simulations with Bioverse}\label{sec:results-bioverse}
-
-In a magnitude- and volume-limited sample of a transit survey, the host star distribution will be skewed toward later spectral types and dominated by M~dwarfs (see Figure~\ref{fig:inhabited_FGKM}).
-Due to how the \gls{HZ} scales with spectral type, by far most transiting \glspl{EEC} occur around M~dwarfs.
+\begin{figure}
+    \script{inhabited_FGKM.py}
+    \begin{centering}
+        \includegraphics[width=\hsize]{figures/inhabited_FGKM}
+        \caption{Host stars of all transiting \glspl{EEC} and inhabited planets in a simulated transit survey.
+        In the FGK sample, most \glspl{EEC} and all inhabited planets orbit K~dwarfs.
+        In an M~dwarf sample of the same size, the fraction of inhabited planets is larger.
+        }
+        \label{fig:inhabited_FGKM}
+    \end{centering}
+\end{figure}
+In a magnitude- and volume-limited sample of a transit survey, the host star distribution for \glspl{EEC} will be skewed toward later spectral types and dominated by M~dwarfs due to how the \gls{HZ} scales with spectral type.
+For a fixed planet sample size, the fraction of inhabited planets is highest in the M~dwarf sample due to the higher \gls{NUV} fluxes in the \gls{HZ} of these stars (see Figure~\ref{fig:inhabited_FGKM}).
 Their \gls{NUV} fluxes are generally highest at early times $\lesssim\SI{100}{\mega\year}$.

### 2024-07-20 src/scripts/surveys_FGKM.py
add detections plot

@@ -2,7 +2,7 @@ import dill
 import paths
 import plotstyle
 import seaborn as sns
-from src.scripts.utils import save_var_latex
+from src.scripts.utils import save_var_latex, read_var_latex
 import matplotlib.pyplot as plt
 import matplotlib.gridspec as gridspec
 
@@ -121,14 +121,37 @@ def plot_nuv_distribution(sample, data, fig, ax, spt):
     return fig, ax
 
 
+def plot_detections_uv(eec, fig, ax, NUV_thresh, ylabel=True):
+    eec = eec.to_pandas()
+    eec["has_O2"] = eec["has_O2"].astype('bool')
+    ax.scatter(eec[~eec.has_O2]["max_nuv"], eec[~eec.has_O2]["has_O2"], s=9.0, color="dimgray")
+    ax.scatter(eec[eec.has_O2]["max_nuv"], eec[eec.has_O2]["has_O2"], s=9.0, color="C1")
+    ax.axvline(x=float(NUV_thresh), linestyle="--", color="grey")
+    ax.set_yticks([0, 1])
+    # ax.set_xlim([0,10])
+    if ylabel:
+        ax.set_yticklabels(["$\oslash$", "$\checkmark$"], fontsize=16)
+    else:
+        ax.set_yticklabels(["", ""], fontsize=16)
+        ax.set_yticks([])
+
+    ax.spines["right"].set_visible(False)
+    ax.spines["top"].set_visible(False)
+    # To turn off the bottom or left
+    # ax.spines['bottom'].set_visible(False)
+    ax.spines["left"].set_visible(False)
+    ax.set_xlabel("max. NUV irradiance [erg/s/$cm^2$]")
+    return fig, ax
+
+
 fig = plt.figure(figsize=(10, 8))
 
 gs = gridspec.GridSpec(3, 2, height_ratios=[0.38, 0.38, 0.24])
 
 ax0 = fig.add_subplot(gs[0, 0])  # First row, first column
-ax1 = fig.add_subplot(gs[0, 1])  # First row, second column
+ax1 = fig.add_subplot(gs[0, 1], sharey=ax0)  # First row, second column
 ax2 = fig.add_subplot(gs[1, 0])  # Second row, first column
-ax3 = fig.add_subplot(gs[1, 1])  # Second row, second column
+ax3 = fig.add_subplot(gs[1, 1], sharey=ax2)  # Second row, second column
 ax4 = fig.add_subplot(gs[2, 0])  # Third row, first column
[…]

### 2024-07-20 src/scripts/surveys_FGKM.py
eye candy

@@ -106,8 +106,8 @@ def plot_nuv_distribution(sample, data, fig, ax, spt):
 
     ax.set_xlabel("max. NUV irradiance $F_\mathrm{NUV, max}$ [erg/s/$cm^2$]")
     ax.set_ylabel("Probability density")
-    if spt == "M":
-        ax.legend(title=None)
+    # if spt == "M":
+    #     ax.legend(title=None)
 
     # ax.text(
     #     0.97,
@@ -144,23 +144,25 @@ def plot_detections_uv(eec, fig, ax, NUV_thresh, ylabel=True):
     return fig, ax
 
 
-fig = plt.figure(figsize=(10, 8))
+fig = plt.figure(figsize=(10, 10))
 
-gs = gridspec.GridSpec(3, 2, height_ratios=[0.38, 0.38, 0.24])
+gs = gridspec.GridSpec(3, 2, height_ratios=[0.38, 0.38, 0.15])
 
 ax0 = fig.add_subplot(gs[0, 0])  # First row, first column
 ax1 = fig.add_subplot(gs[0, 1], sharey=ax0)  # First row, second column
 ax2 = fig.add_subplot(gs[1, 0])  # Second row, first column
 ax3 = fig.add_subplot(gs[1, 1], sharey=ax2)  # Second row, second column
-ax4 = fig.add_subplot(gs[2, 0])  # Third row, first column
-ax5 = fig.add_subplot(gs[2, 1])  # Third row, second column
+ax4 = fig.add_subplot(gs[2, 0], sharex=ax3)  # Third row, first column
+ax5 = fig.add_subplot(gs[2, 1], sharex=ax3)  # Third row, second column
 
 # axs = [ax0, ax1, ax2, ax3, ax4, ax5]
 axs_left = [ax0, ax2, ax4]
 axs_right = [ax1, ax3, ax5]
 
-[ax.set_title("FGK-type hosts") for ax in axs_left]
-[ax.set_title("M-type hosts") for ax in axs_right]
+# [ax.set_title("FGK-type hosts") for ax in axs_left]
+# [ax.set_title("M-type hosts") for ax in axs_right]
+ax0.set_title("FGK-type hosts")
+ax1.set_title("M-type hosts")
 
 
 for spt, axlr in zip(["FGK", "M"], [axs_left, axs_right]):
@@ -171,19 +173,21 @@ for spt, axlr in zip(["FGK", "M"], [axs_left, axs_right]):
 
[…]

### 2024-07-20 src/tex/ms.tex
remove remark on vol-limited
combine three figures

@@
 \subsection{Survey simulations with Bioverse}\label{sec:results-bioverse}
-\begin{figure}
-    \script{inhabited_FGKM.py}
-    \begin{centering}
-        \includegraphics[width=\hsize]{figures/inhabited_FGKM}
-        \caption{Host stars of all transiting \glspl{EEC} and inhabited planets in a simulated transit survey.
-        In the FGK sample, most \glspl{EEC} and all inhabited planets orbit K~dwarfs.
-        In an M~dwarf sample of the same size, the fraction of inhabited planets is larger.
-        }
-        \label{fig:inhabited_FGKM}
-    \end{centering}
-\end{figure}
+%\begin{figure}
+%    \script{inhabited_FGKM.py}
+%    \begin{centering}
+%        \includegraphics[width=\hsize]{figures/inhabited_FGKM}
+%        \caption{Host stars of all transiting \glspl{EEC} and inhabited planets in a simulated transit survey.
+%        In the FGK sample, most \glspl{EEC} and all inhabited planets orbit K~dwarfs.
+%        In an M~dwarf sample of the same size, the fraction of inhabited planets is larger.
+%        }
+%        \label{fig:inhabited_FGKM}
+%    \end{centering}
+%\end{figure}
 In a magnitude- and volume-limited sample of a transit survey, the host star distribution for \glspl{EEC} will be skewed toward later spectral types and dominated by M~dwarfs due to how the \gls{HZ} scales with spectral type.
-For a fixed planet sample size, the fraction of inhabited planets is highest in the M~dwarf sample due to the higher \gls{NUV} fluxes in the \gls{HZ} of these stars (see Figure~\ref{fig:inhabited_FGKM}).
+For a fixed planet sample size, the fraction of inhabited planets is highest in the M~dwarf sample due to the higher \gls{NUV} fluxes in the \gls{HZ} of these stars (see Figure~\ref{fig:surveys_FGKM}).
 Their \gls{NUV} fluxes are generally highest at early times $\lesssim\SI{100}{\mega\year}$.
 These host stars, in particular late subtypes, also provide extended periods of increased \gls{NUV} emission that overlap with times when some of these planets occupy the \gls{HZ} (see Figure~\ref{fig:hz_nuv_evo}), our requirement for abiogenesis (compare Equation~\ref{eq:hypothesis-uv}).
-Because of that, all inhabited planets in our magnitude- and volume-limited sample orbit M~dwarfs.
+%Because of that, most inhabited planets in our magnitude- and volume-limited sample orbit M~dwarfs.
 
@@
 In Section~\ref{sec:results-semianalytical}, we demonstrated that the probability of obtaining true strong evidence for the hypothesis that life only originates on planets with a minimum past \gls{NUV} flux is sensitive to the distribution of sampled past \gls{NUV} fluxes, i.e., the selectivity of the survey (compare Figure~\ref{fig:semian_selectivity}).
-%The maximum \gls{NUV} distribution of our generic transit survey is rather unimodal with a median around \SI{500}{\erg\per\second\per\centi\meter\squared} (see Figure~\ref{fig:nuv_distribution}), leading to a selectivity of roughly \var{selectivity_transit_volumelim}.
+%The maximum \gls{NUV} distribution of our generic transit survey is rather unimodal with a median around \SI{500}{\erg\per\second\per\centi\meter\squared} (see Figure~\ref{fig:surveys_FGKM}), leading to a selectivity of roughly \var{selectivity_transit_volumelim}.
 For both surveys targeting M dwarfs and those targeting FGK dwarfs, the maximum \gls{NUV} distribution is rather unimodal.
-Applying the approach from Sect.~\ref{sec:results-semianalytical} of fitting a Beta function to the distribution, we find rather low selectivities (see Figure~\ref{fig:nuv_distribution}).
-\begin{figure*}
-    \script{nuv_distribution.py}
-    \begin{centering}
-        \includegraphics[width=\hsize]{figures/nuv_distribution}
-        \caption{Distribution of maximum past \gls{NUV} flux in transit surveys targeting \glspl{EEC} around FGK and M~stars, respectively.
+Applying the approach from Sect.~\ref{sec:results-semianalytical} of fitting a Beta function to the distribution, we find rather low selectivities (see Figure~\ref{fig:surveys_FGKM}).
+%\begin{figure*}
+%    \script{nuv_distribution.py}
+%    \begin{centering}
+%        \includegraphics[width=\hsize]{figures/nuv_distribution}
+%        \caption{Distribution of maximum past \gls{NUV} flux in transit surveys targeting \glspl{EEC} around FGK and M~stars, respectively.
 The best-fit beta distributions (gray) correspond to selectivities of $s_\mathrm{FGK} = $\var{selectivity_FGK} and $s_\mathrm{M} = $\var{selectivity_M}.
 Red areas show inhabited planets for a threshold \gls{NUV} flux of $F_\mathrm{NUV, min} = \var{NUV_thresh}\,\SI{}{\erg\per\second\per\centi\meter\squared}$ and an abiogenesis rate of $f_\mathrm{life} = \var{f_life}$.}
-        \label{fig:nuv_distribution}
+%        \label{fig:nuv_distribution}
+%    \end{centering}
+%\end{figure*}
+
+\begin{figure*}
+    \script{surveys_FGKM.py}
+    \begin{centering}
+        \includegraphics[width=\hsize]{figures/surveys_FGKM}
+        \caption{
+        Simulated transit surveys targeting FGK and M~stars.\\
+        Top: Host stars of all transiting \glspl{EEC} and inhabited planets in a simulated transit survey.
+        In the FGK sample, most \glspl{EEC} and all inhabited planets orbit K~dwarfs.
+        In an M~dwarf sample of the same size, the fraction of inhabited planets is larger.\\
+        Center: Distribution of maximum past \gls{NUV} flux in transit surveys targeting \glspl{EEC} around FGK and M~stars, respectively.
+The best-fit beta distributions (gray) correspond to selectivities of $s_\mathrm{FGK} = $\var{selectivity_FGK} and $s_\mathrm{M} = $\var{selectivity_M}.
+Green areas show inhabited planets for an abiogenesis rate of $f_\mathrm{life} = \var{f_life}$ and a generic threshold \gls{NUV} flux $F_\mathrm{NUV, min} = \var{NUV_thresh}\,\SI{}{\erg\per\second\per\centi\meter\squared}$.\\
+        Bottom: Recovered biosignature detections ($\checkmark$) and non-detections ($\oslash$) of simulated transit surveys. The dashed line denotes $F_\mathrm{NUV, min}$.}
+        \label{fig:surveys_FGKM}
     \end{centering}
@@
 \subsubsection{Expected biosignature pattern in a transit survey}
-A representative recovery of the injected biosignature pattern is shown in Figure~\ref{fig:detections_uv}.
+A representative recovery of the injected biosignature pattern is shown in Figure~\ref{fig:surveys_FGKM}.
 There, we assumed an abiogenesis rate of $f_\mathrm{life} = \var{f_life}$ and a minimum \gls{NUV} flux of $F_\mathrm{NUV, min} = \var{NUV_thresh}\,\SI{}{\erg\per\second\per\centi\meter\squared}$.
 All injected biosignatures are assumed to be detected, and the maximum \gls{NUV} flux is estimated from the host star's spectral type and age with an uncertainty corresponding to the intrinsic scatter in the \gls{NUV} fluxes in \citet{Richey-Yowell2023}.
-\begin{figure*}
-    \script{detections_uv.py}
-    \begin{centering}
-        \includegraphics[width=\hsize]{figures/detections_uv}
-        \caption{Recovered biosignature detections in the \gls{NUV} flux-biosignature occurrence space.
+%\begin{figure*}
+%    \script{detections_uv.py}
+%    \begin{centering}
+%        \includegraphics[width=\hsize]{figures/detections_uv}
+%        \caption{Recovered biosignature detections in the \gls{NUV} flux-biosignature occurrence space.
 The dashed line denotes a generic threshold \gls{NUV} flux $F_\mathrm{NUV, min} = \var{NUV_thresh}\,\SI{}{\erg\per\second\per\centi\meter\squared}$.}
-        \label{fig:detections_uv}
-    \end{centering}
-\end{figure*}
+%        \label{fig:detections_uv}
+%    \end{centering}
+%\end{figure*}
 This leads to a distribution of biosignature detections with detections increasingly occurring above a threshold inferred NUV flux.

### 2024-07-26 src/tex/ms.tex
label the hypothesis "UV~Threshold Hypothesis" throughout

@@
 %\subsection{UV flux}\label{sec:predictions:uv}
-A major hypothesis in the origin of life is that UV light played a constructive role in getting life started on Earth (see Ranjan et al.
+A major hypothesis in the origin of life is the ``\textit{UV Threshold Hypothesis}'', which states that UV light played a constructive role in getting life started on Earth (see Ranjan et al.
 2016, 2017c; Rimmer et al.
@@
         \includegraphics[width=\hsize]{figures/hypotheses.pdf}
-        \caption{Population-level hypothesis and null hypothesis on UV irradiance derived from the cyanosulfidic scenario.}
+        \caption{UV~Threshold Hypothesis and null hypothesis derived from the cyanosulfidic scenario.}
         \label{fig:hypotheses}
@@
 Test for negative correlations as well?}
-Here, we conduct a theoretical experiment on the UV irradiance requirement (Sect.~\ref{sec:predictions}) by relating the occurrence of life on an exo-earth candidate with a minimum past quiescent stellar UV flux, focusing on the prebiotically interesting \gls{NUV} range from \SIrange{200}{280}{\nano\meter}.
+Here, we conduct a theoretical experiment on the UV~Threshold Hypothesis (Sect.~\ref{sec:predictions}) by relating the occurrence of life on an exo-earth candidate with a minimum past quiescent stellar UV flux, focusing on the prebiotically interesting \gls{NUV} range from \SIrange{200}{280}{\nano\meter}.
 \todo[inline]{SUKRIT: got a good reference for this?}  %~\citep{RanjanSasselov2016}.?
-Our concrete hypothesis shall be that life only occurs on planets that at some point in their history have received such radiation exceeded a minimum flux $F_\mathrm{NUV, min}$.
+Our concrete hypothesis shall be that life only occurs on planets that at some point in their history have received such radiation at a flux exceeding a threshold $F_\mathrm{NUV, min}$.
 
@@
 \subsubsection{Habitable zone occupancy and UV flux}\label{sec:met-hz}
-To test the UV hypothesis, we require that life occurs only on planets with sufficient past UV irradiation exceeding $F_\mathrm{NUV, min}$.
+To test the UV~Threshold Hypothesis, we require that life occurs only on planets with sufficient past UV irradiation exceeding $F_\mathrm{NUV, min}$.
 Further, we require this flux to have lasted for a minimum duration $\Delta T_\mathrm{min}$ to allow for a sufficient ``origins timescale''~\citep{Rimmer2023}.
@@
 We assigned the development of life to a random fraction $f_\mathrm{life}$ of all temperate planets fulfilling these requirements.
-For the probability of a planet having detectable biosignatures, $P(\mathrm{bio})$, the UV~hypothesis then states
+For the probability of a planet having detectable biosignatures, $P(\mathrm{bio})$, the UV~Threshold Hypothesis then states
 \begin{equation}\label{eq:hypothesis-uv}
@@
         \includegraphics[width=\hsize]{figures/inhabited_aafo_nuv_thresh}
-        \caption{Fraction of inhabited planets for different threshold \gls{NUV} fluxes under the UV hypothesis if the abiogenesis rate is \var{f_life}.
+        \caption{Fraction of inhabited planets for different threshold \gls{NUV} fluxes under the UV~Threshold Hypothesis if the abiogenesis rate is \var{f_life}.
 Shaded regions denote 90 percent confidence intervals of randomized sample generations, and dashed lines correspond to samples including non-transiting planets.
@@
 \end{figure}
-Figure \ref{fig:inhabited_aafo_nuv_thresh} shows the fraction of inhabited planets under the UV hypothesis for different threshold \gls{NUV} fluxes and for a high abiogenesis rate of \var{f_life}.
+Figure \ref{fig:inhabited_aafo_nuv_thresh} shows the fraction of inhabited planets under the UV~Threshold Hypothesis for different threshold \gls{NUV} fluxes and for a high abiogenesis rate of \var{f_life}.
 This fraction decreases sharply with increasing threshold flux, as fewer planets receive sufficient \gls{NUV} flux for abiogenesis.
@@
 The cyanosulfidic scenario~\citep{Patel2015}, in particular its predicted existence of a minimum \gls{NUV} flux required for prebiotic chemistry, seems to offer an opportunity to test an origins of life hypothesis with a statistical transit survey sampling planets with varying \gls{NUV} flux histories.
-In the following, we discuss the prospects of testing the UV hypothesis in light of our results.
+In the following, we discuss the prospects of testing the UV~Threshold Hypothesis in light of our results.
 
@@
 In doing so, we may evaluate how evidence depends on the uncertainty over these unknown parameters in general terms, without assuming which particular test a future observer might choose to actually perform over real data when available.
-From this, we can see that target selection can strongly affect the conclusiveness of a future test of the UV hypothesis.
+From this, we can see that target selection can strongly affect the conclusiveness of a future test of the UV~Threshold Hypothesis.
 
@@
 For larger samples, however, a narrow distribution of $F_{\mathrm{NUV}}$ may prevent inference entirely.
-We thus argue that selecting a sample with $F_{\mathrm{NUV}}$ distributed uniformly or emphasizing extreme values should -- barren any practical counterarguments -- be considered in any future attempt at testing the UV hypothesis.
+We thus argue that selecting a sample with $F_{\mathrm{NUV}}$ distributed uniformly or emphasizing extreme values should -- barren any practical counterarguments -- be considered in any future attempt at testing the UV~Threshold Hypothesis.
 Since the practical implementation of an exoplanet survey can stand in the way of such a selection, the following discussion focuses on the results of our transit survey simulations with Bioverse.
@@
 An interesting aspect lies in the distribution of host star properties, as different spectral types probe different past \gls{NUV} flux regimes.
-Under the UV~hypothesis, late-type M~dwarfs are favored due to their higher \gls{NUV} fluxes in the \gls{HZ}~\citep{Richey-Yowell2023}.
+Under the UV~Threshold Hypothesis, late-type M~dwarfs are favored due to their higher \gls{NUV} fluxes in the \gls{HZ}~\citep{Richey-Yowell2023}.
 Furthermore, their wider distribution of maximum past \gls{NUV} fluxes may increase the likelihood of obtaining strong evidence for or against the hypothesis.
@@
 With this sample, the survey will be more sensitive to the low \gls{NUV} flux end of the hypothesis.
-A lack of biosignatures on these planets would support the UV~hypothesis, whereas their presence might suggest that lower NUV fluxes are also sufficient for abiogenesis or indicate a different abiogenic pathway.
+A lack of biosignatures on these planets would support the UV~Threshold Hypothesis, whereas their presence might suggest that lower NUV fluxes are also sufficient for abiogenesis or indicate a different abiogenic pathway.
 CITE
@@
 \subsection{Sensitivity to astrophysical parameters} % corresponds to conclusion point 3
-Our Bioverse simulations that take into account exoplanet demographics, the evolution of habitability and \gls{NUV} fluxes, and observational biases show that not only the likelihood of a conclusive test of the UV~hypothesis, but also the likelihood of successful biosignature detection itself is extremely sensitive to the threshold \gls{NUV} flux if the UV~hypothesis is true.
+Our Bioverse simulations that take into account exoplanet demographics, the evolution of habitability and \gls{NUV} fluxes, and observational biases show that not only the likelihood of a conclusive test of the UV~Threshold Hypothesis, but also the likelihood of successful biosignature detection itself is extremely sensitive to the threshold \gls{NUV} flux if the hypothesis is true.
 Even if all biosignatures can be detected and the nominal abiogenesis rate is very high, say $f_\mathrm{life} = 1$, under the condition that prebiotic chemistry requires a minimum \gls{NUV} flux \textit{and} liquid water, the probability of finding life on a randomly selected planet may be very low.
@@
 Conversely, a biosignature detection on a planet estimated to have received very little UV~radiation would increase the likelihood of a false positive detection.
-On the other hand, should the detection in the latter case be confirmed, it could be used to falsify the UV~hypothesis.
+On the other hand, should the detection in the latter case be confirmed, it could be used to falsify the UV~Threshold Hypothesis.
 
@@
 The abiogenesis rate in both samples show similar trends with the threshold \gls{NUV} flux, and the fraction of inhabited planets drops at similar threshold fluxes (see Section~\ref{sec:results-bioverse}).
-A potentially inhabited planet's host star spectral type may thus not be a strong indicator for the credibility of a biosignature detection in the context of the UV~hypothesis.
+A potentially inhabited planet's host star spectral type may thus not be a strong indicator for the credibility of a biosignature detection in the context of the UV~Threshold Hypothesis.
 
-\subsection{Overall prospects for testing the UV hypothesis} % corresponds to conclusion point 4
-Our results show that the UV hypothesis is testable with upcoming exoplanet surveys, but that the success of such a test depends on the sample size, the distribution of past \gls{NUV} fluxes, and several unknown astrophysical nuisance parameters.
+\subsection{Overall prospects for testing the UV~Threshold Hypothesis} % corresponds to conclusion point 4
+Our results show that the UV~Threshold Hypothesis is testable with upcoming exoplanet surveys, but that the success of such a test depends on the sample size, the distribution of past \gls{NUV} fluxes, and several unknown astrophysical nuisance parameters.
 Even under idealized conditions, obtaining strong evidence for or against the hypothesis likely requires sample sizes on the order of 100 (see Section~\ref{sec:results-semianalytical}).
 This is true for a future transit survey, the specifics of which we have reflected in our Bioverse simulations (see Section~\ref{sec:results-bioverse}).
-However, we have shown that the impacts from the combined requirements of the UV~hypothesis on the fraction of inhabited planets in a sample are comparable in the non-transiting case.
+However, we have shown that the impacts from the combined requirements of the UV~Threshold Hypothesis on the fraction of inhabited planets in a sample are comparable in the non-transiting case.
 The likely required sample sizes are larger than the projected sample sizes of most planned missions; for example the \gls{HWO} expects to characterize a sample of only $\sim 25$ Earth analogs~\citep{Mamajek2023,Tuchow2024}.
 
-Meaningful tests of the UV hypothesis may thus require a different survey strategy that increases the sample of characterized planets.
+Meaningful tests of the UV~Threshold Hypothesis may thus require a different survey strategy that increases the sample of characterized planets.
 Such concepts, like the Nautilus mission~\citep{Apai2019,Apai2022}, which aims to characterize up to $\sim 1000$ rocky planets via utilizing an innovative telescope design, are only appearing on the horizon.
-Constraints on the origins of life from exoplanet surveys may thus be a long-term endeavor, and the UV hypothesis may not be testable yet with the current and next generation of exoplanet surveys.
+Constraints on the origins of life from exoplanet surveys may thus be a long-term endeavor, and the UV~Threshold Hypothesis may not be testable yet with the current and next generation of exoplanet surveys.
 In the meantime, it is important to refine predictions on the role of UV radiation in prebiotic chemistry with both theoretical and experimental work.
@@
 \subsubsection{Atmosphere transmission}
-Theoretical work suggests that the atmosphere of prebiotic Earth was largely transparent at near-UV wavelengths with the only known source of attenuation being Rayleigh scattering~\citep{Ranjan2017,Ranjan2017c}.
+Theoretical work suggests that the atmosphere of prebiotic Earth was largely transparent at \gls{NUV} wavelengths with the only known source of attenuation being Rayleigh scattering~\citep{Ranjan2017,Ranjan2017c}.
 We thus approximated surface UV flux using top-of-atmosphere fluxes.
-This represents a conservative approach, since any planet that fails to meet the irradiance criterion receives even lower near-UV radiation at its surface.
+This represents a conservative approach, since any planet that fails to meet the irradiance criterion receives even lower \gls{NUV} radiation at its surface.
 \todo[inline]{SUKRIT please specify as neccessary}
@@
 %        A direct imaging mission with the \gls{HWO} will require characterizing XXX planets ...
-    \item Under the UV hypothesis, the fraction of inhabited planets in a transit survey is sensitive to the threshold \gls{NUV} flux and is expected to drop sharply for required fluxes above a few hundred~\SI{}{\erg\per\second\per\centi\meter\squared}.
+    \item Under the UV~Threshold Hypothesis, the fraction of inhabited planets in a transit survey is sensitive to the threshold \gls{NUV} flux and is expected to drop sharply for required fluxes above a few hundred~\SI{}{\erg\per\second\per\centi\meter\squared}.
     \item If the predicted UV~correlation exists, obtaining strong evidence for the hypothesis is likely ($\gtrsim \SI{80}{\percent}$) for sample sizes $\geq 100$ if the abiogenesis rate is high ($\gtrsim \SI{50}{\percent}$) and if no very high \gls{NUV} fluxes are required.
     A survey strategy that targets extreme values of inferred past \gls{NUV} irradiation increases the diagnostic power.
-    \item Samples of planets orbiting M~dwarfs overall yield higher chances of successfully testing the UV~hypothesis.
+    \item Samples of planets orbiting M~dwarfs overall yield higher chances of successfully testing the UV~Threshold Hypothesis.
         Leaving other factors aside, they are also more likely to yield biosignature detections under this hypothesis.
@@
 %    \item Abiogenesis in the cyanosulfidic scenario prefers late host star spectral types due to the higher \gls{NUV} fluxes in their habitable zones.
-However, due to a wider distribution of maximum past \gls{NUV} fluxes in the \gls{HZ} of FGK stars, a transit survey targeting FGK systems is more likely to obtain strong evidence for or against the UV~hypothesis.
+However, due to a wider distribution of maximum past \gls{NUV} fluxes in the \gls{HZ} of FGK stars, a transit survey targeting FGK systems is more likely to obtain strong evidence for or against the UV~Threshold Hypothesis.
 

### 2024-07-26 src/scripts/semian_true_evidence.py
add better labels

@@ -127,9 +127,9 @@ def plot_true_evidence(ax):
         columnspacing=1.6,
     )
     ax.set_xscale("log")
-    ax.set_ylim(-.05, 1.05)
-    ax.set_ylabel("P(evidence>x)")
-    ax.set_xlabel("x")
+    ax.set_ylim(-0.05, 1.05)
+    ax.set_ylabel("$P(\mathrm{evidence} > x)$")
+    ax.set_xlabel("$x$")
 
     return ax
 
@@ -180,10 +180,10 @@ def plot_evidence_sampsize(ax, s_max=500):
     svec = np.arange(10,s_max)
     true_strong_evidence_p = [P_true_evidence(sampsize=s) for s in svec]
 
-    ax.plot(svec,true_strong_evidence_p)
-    ax.set_ylim(-.05, 1.05)
-    ax.set_ylabel('P($BF_{H_1, H_\mathrm{null}}$ > 10)')
-    ax.set_xlabel('Sample size')
+    ax.plot(svec, true_strong_evidence_p)
+    ax.set_ylim(-0.05, 1.05)
+    ax.set_ylabel("$P(BF_{H_1, H_\mathrm{null}} > 10$)")
+    ax.set_xlabel("Planet sample size")
 
     return ax
 

### 2024-07-26 src/tex/ms.tex
improve caption, and other edits

@@
 \subsubsection{Habitable zone occupancy and UV flux}\label{sec:met-hz}
-To test the UV~Threshold Hypothesis, we require that life occurs only on planets with sufficient past UV irradiation exceeding $F_\mathrm{NUV, min}$.
+To test the UV~Threshold Hypothesis, we require that life occurs only on planets with sufficient past UV irradiation exceeding the origins of life threshold $F_\mathrm{NUV, min}$.
 Further, we require this flux to have lasted for a minimum duration $\Delta T_\mathrm{min}$ to allow for a sufficient ``origins timescale''~\citep{Rimmer2023}.
@@
 To determine \gls{HZ} occupancy, we interpolated the stellar luminosity evolution grid of \citet{Baraffe1998} using a Clough Tocher interpolant~\citep[][see left panel of Figure~\ref{fig:hz_nuv_evo}]{Nielson1983,Alfeld1984} to compute the evolution of the inner (runaway greenhouse) and outer (maximum greenhouse) edges as a function of planet mass and stellar spectral type~\citep{Kopparapu2014}.
-This provides each planet's epochs within and outside the \gls{HZ}.
+Being a local interpolation method, Clough Tocher enables rapid processing while producing a smooth interpolating surface that highlights local trends.
+From this, we get each planet's epochs within and outside the \gls{HZ}.
 For the \gls{NUV} flux, we use the age- and stellar mass-dependent \gls{NUV} fluxes in the \gls{HZ} obtained by \citet{Richey-Yowell2023}.
@@
 \subsection{Semi-analytical assessment}\label{sec:results-semianalytical}
-In Section~\ref{sec:met-semianalytical} we computed the probability for true positive evidence for $H_\mathrm{1}$ and $H_\mathrm{null}$ respectively.
+In Section~\ref{sec:met-semianalytical} we computed the probability for true positive evidence for $H_\mathrm{1}$ and $H_\mathrm{null}$, respectively.
 Figure~\ref{fig:semian_true_evidence} shows how these evidences are distributed for sample sizes \var{semian_Nsamp1} and \var{semian_Nsamp2}, and how likely we are to obtain strong evidence ($BF_{H_i, H_j}$ > 10).
@@
         \includegraphics[width=\hsize]{figures/semian_true_evidence}
-        \caption{Probability to obtain true strong evidence.
-Left: evidence levels for $H_\mathrm{1}$ and $H_\mathrm{null}$ under sample sizes $n = \var{semian_Nsamp1}$ (solid) and $n = \var{semian_Nsamp2}$ (dashed).
-The vertical lines denote the thresholds for ``strong'' evidence, $BF_{H_i, H_j}$ > 10, and ``extreme'' evidence, $BF_{H_i, H_j}$ > 100.
-Right: Probability of true strong evidence for $H_\mathrm{1}$ as a function of sample size $n$.}
+        \caption{Obtaining true strong evidence with different sample sizes.
+Left: Probability to reach given evidence levels for $H_\mathrm{1}$ and $H_\mathrm{null}$ under sample sizes $n = \var{semian_Nsamp1}$ (solid) and $n = \var{semian_Nsamp2}$ (dashed).
+Vertical lines denote thresholds for ``strong'' evidence, $BF_{H_i, H_j}$ > 10, and ``extreme'' evidence, $BF_{H_i, H_j}$ > 100.
+Right: Probability of obtaining true strong evidence for $H_\mathrm{1}$ as a function of sample size $n$.}
         \label{fig:semian_true_evidence}

### 2024-07-26 src/scripts/semian_true_evidence.py
fix&improve figure

@@ -12,6 +12,7 @@ from scipy import stats
 
 rng = np.random.default_rng(seed=42)
 
+
 def BFH1(k, theta, pi, theta_l):
     """
     Evidence for H1 against H0
@@ -31,30 +32,34 @@ def BFH1(k, theta, pi, theta_l):
     bf = stats.binom.pmf(k, nl, pi) / stats.binom.pmf(k, len(theta), pi)
     return bf
 
-def sample_H0(theta,pi,theta_l):
+
+def sample_H0(theta, pi, theta_l):
     """
     Generates a sample of values of k under H0
     """
     # theta is the vector of theta values
     # pi is the probability of life
     # theta_l is the value of the limit theta
-    return(np.random.binomial(len(theta),pi))
+    return np.random.binomial(len(theta), pi)
 
 
-def sample_H1(theta,pi,theta_l):
+def sample_H1(theta, pi, theta_l):
     """
     Generates a sample of values of k under H1
     """
     # theta is the vector of theta values
     # pi is the probability of life
     # theta_l is the value of the limit theta
-    if len(np.shape(theta_l))==2:
-        X = np.random.binomial(np.sum(theta>theta_l,axis=1),pi)
-    elif len(np.shape(theta_l))==1:
-        X = np.random.binomial(np.sum(theta>np.reshape(theta_l,(len(theta_l),1)),axis=1),pi)
+    if len(np.shape(theta_l)) == 2:
+        X = np.random.binomial(np.sum(theta > theta_l, axis=1), pi)
+    elif len(np.shape(theta_l)) == 1:
+        X = np.random.binomial(
+            np.sum(theta > np.reshape(theta_l, (len(theta_l), 1)), axis=1), pi
+        )
     else:
-        X = np.random.binomial(np.sum(theta>theta_l),pi)
-    return(X)
+        X = np.random.binomial(np.sum(theta > theta_l), pi)
+    return X
+
 
 def get_evidence(sampsize=100):
     # For convenience, we redefine theta here so that one can play with sample size and such
@@ -85,7 +90,7 @@ def plot_true_evidence(ax):
 
     for s, ls, showlabel in zip([10, 100], ["-", "--"], ["", "_"]):
         bfh1_kh1, bfh0_kh0 = get_evidence(s)
-    
+
[…]

### 2024-07-26 src/scripts/semian_true_evidence.py
minor things

@@ -134,7 +134,8 @@ def plot_true_evidence(ax):
     )
     ax.set_xscale("log")
     ax.set_ylim(-0.05, 1.05)
-    ax.set_ylabel("$P(\mathrm{evidence} > x)$")
+    # ax.set_ylabel("$P(\mathrm{evidence} > x)$")
+    ax.set_ylabel("$P(BF_{H_i, H_j} > x$)")
     ax.set_xlabel("$x$")
 
     return ax

### 2024-07-26 src/scripts/bioverse_pipeline.py
minor things

@@ -86,7 +86,7 @@ def get_generator_args():
         # 'mr_relation' : 'Wolfgang2016',
         # 'mr_relation' : 'earthlike',
         # past HZ occupancy and NUV flux
-        "deltaT_min": 100.0,  # Myr
+        "deltaT_min": 10.0,  # Myr
         "NUV_thresh": 100.0,  # erg/s/cm2
     }
     return stars_args, planets_args
@@ -384,6 +384,8 @@ def past_uv(
         Whether to perform a grid of hypothesis tests.
     N_grid : int
         The number of grid points to use.
+    N_planets : int
+        The number of planets to generate.
     powergrid : bool
         Whether to run a grid of statistical power calculations.
     fast : bool
@@ -557,9 +559,8 @@ def main(fast=False, testmethod="mannwhitney"):
 
 
 if __name__ == "__main__":
-    # result = timeit.timeit("main()", number=1)
-    # main(fast=True)
-    main(fast=False)
+    # result = timeit(main())
+    result = timeit(main(fast=True))
 
 
 # # -----------------

### 2024-07-26 src/scripts/hz_nuv_evo.py

@@ -48,7 +48,7 @@ def plot_interpolation(
     fig.colorbar(q, label=cbarlabel)
     ax.set_xscale("log")
     ax.set_xlabel("Time (Gyr)")
-    ax.set_ylabel("Stellar Mass ($M_\odot$)")
+    ax.set_ylabel("Stellar mass ($M_\odot$)")
     return ax
 
 

### 2024-07-26 src/scripts/inhabited-aafo-nuv_thresh.py

@@ -44,7 +44,6 @@ def run_simulation(nuv_thresh, n_planets, star_type, seed, generator_kwargs):
         d = d.sample(n_planets, random_state=seed, replace=True)
 
     fraction = len(d[d["inhabited"]]) / len(d)
-    print(fraction)
     return fraction
 
 
@@ -163,7 +162,7 @@ def plot_results():
 
 
 @timeit
-def main(debug):
+def main(debug=False):
     for transiting in [True, False]:
     # for transiting in [False]:
     # for transiting in [True]:
@@ -175,3 +174,6 @@ def main(debug):
 
 if __name__ == "__main__":
     main()
+
+    # DEBUG
+    # main(debug=True)

### 2024-07-26 src/scripts/utils.py

@@ -72,7 +72,7 @@ def timeit(func):
 
         # Print formatted time
         print(
-            f"Processing time of {func.__qualname__}(): {int(hours):02}:{int(minutes):02}:{seconds:.0f}."
+            f"Processing time of {func.__qualname__}(): {int(hours):02}:{int(minutes):02}:{seconds:.0f}"
         )
         return result
 

### 2024-09-06 src/tex/ms.tex
add Regis' introduction draft

@@
 \todo[inline]{Introduce OOL, the importance of planetary contexts}
-...
+A wide variety of scenarios for the origin of life have been proposed~\citep[e.g.,][]{Dass2016,Westall2018}.
+While we are still far away from testing any of them, new prospects in the search for conditions favorable to life have opened up by thinking of the origin of life as a planetary phenomenon and identifying global-scale environmental properties that might support pathways to life~\citep{Sasselov2020}.
+In particular, specific planetary conditions are needed to create stockpiles of initial compounds for prebiotic chemistry; and planetary processes are required to trigger the prebiotic synthesis.
+Such planetary conditions can be hypothesized for exoplanets located in the habitable zone of their host star, with persistent liquid water on their surface.
+For example, the deep-sea hydrothermal (or hydrothermal-sedimentary) context for the origin of life requires a direct contact of an ocean and the planetary mantle/crust~\citep[e.g.,][]{Russel2010}.
+The alternate scenario of a surface locally subject to wet-dry cycles requires a planetary exposure to mid-range UV irradiation, as a source of energy and an agent of selection in chemical evolution~\citep[e.g.,][]{Deamer2019}.
+This “UV Threshold Hypothesis” states that UV light in a specific length range played a constructive role in getting life started on Earth (CITE see Ranjan et al.
+2016, 2017c; Rimmer et al.
+2018; Rapf \& Vaida 2016; Pascal et al.
+2012; Green et al.
+2021; and sources therein).
+
+The association of chemical pathways to life and planetary environmental conditions offers a new opportunity to test alternate scenarios for life emergence based on planetary-level data collected from the upcoming observations of populations of exoplanets.
+Deep-sea hydrothermal scenarios require planetary conditions that may not be met on ocean worlds with large amounts of water, where the water pressure on the ocean floor is high enough to form high-pressure ices~\citep{Noack2016}.
+In this case, a testable prediction would be that planets with high-pressure ices do not show biosignatures.
+Likewise, if UV light is required to get life started, then there is a minimum planetary UV flux requirement to have an inhabited world.
+This requirement is set by competitor thermal processes; if the photoreaction does not move forward at a rate faster than the competitor thermal process(es), then the abiogenesis scenario cannot function.
+On the other hand, abundant UV light vastly in excess of this threshold does not increase the probability of abiogenesis, since once the UV photochemistry is no longer limiting, some other thermal process in the reaction network will be the rate-limiting process instead.
+Therefore, a putative dependence of life on UV light is best encoded as a step function (see, e.g., Ranjan et al. 2017c; Rimmer et al. 2018; Rimmer, Ranjan \& Rugheimer 2021).
+
+The goal of this work is to evaluate the potential of upcoming exoplanet surveys to test the hypothesis that a minimum past NUV flux is required for abiogenesis.
+We focus on one version of the UV Threshold Hypothesis, the so-called cyanosulfidic scenario, which has been refined to the point where the required threshold flux has been measured to be $F_\mathrm{NUV, min} = \SI{6.8\pm3.6 e10}{\photons\per\centi\meter\squared\per\second\per\nano\meter}$ integrated from \SIrange{200}{280}{\nano\meter} at the surface (Rimmer et al.
+2018; Rimmer et al.
+2021 Astrobiology; Rimmer et al.
+2023).
+
+We first follow a semi-analytical Bayesian analysis to estimate probabilities of obtaining strong evidence for or against this hypothesis.
+Under the UV Threshold Hypothesis ($H_1$), the probability of an exoplanet having detectable biosignatures is zero if the near-ultraviolet (\gls{NUV}) irradiation is less than the threshold, and it is equal to the (unknown) probability of abiogenesis, $f_\mathrm{life}$ if \gls{NUV} exceeds the threshold for a sufficiently long period of time.
+Under the null hypothesis ($H_0$), that probability simply is $f_\mathrm{life}$, that is, it does not correlate with the UV flux.
+Given a sample of planets, where for some of them we have convincing biosignature detections but remain agnostic on $f_\mathrm{life}$, we ask what evidence for $H_1$ and $H_0$ we can expect to obtain.
+
+A real exoplanet survey will be subject to observational biases, sample selection effects, and the underlying demographics of the planet sample.
+To assess the information gain of a realistic exoplanet survey, we employed Bioverse~\citep{Bixel2021,Hardegree-Ullman2023,Schlecker2024,Hardegree-Ullman2024}, a framework that integrates multiple components including statistically realistic simulations of exoplanet populations, a survey simulation module, and a hypothesis testing module to evaluate the statistical power of different observational strategies.
+The general approach is as follows:
+After defining a population of synthetic exoplanets, we simulate the detection and characterization of these exoplanets with a hypothetical survey, taking into account the survey’s sensitivity, target selection, and observational biases.
+Then we evaluate the likelihood that a given survey would detect a specified demographic trend in the exoplanet population and estimate the precision with which the survey could constrain the parameters of that trend.
+A common definition of the null hypothesis $H_0$, which is also applied here, is that there is no relationship between the independent variable (here: maximum \gls{NUV} flux) and the dependent variable (here: biosignature occurrence).
+The alternative hypothesis $H_1$ proposes a specific relationship between the independent and dependent variables.
+To determine the diagnostic capability of a given survey, Bioverse runs multiple iterations of the simulated survey and calculates the fraction of realizations that successfully reject the null hypothesis.
+We use this metric, known as the statistical power, to quantify the potential information content of the survey, identify critical design trades, and find strategies that maximize the survey’s scientific return.
+
 \subsection{Origins of Life Scenarios and their Predictions}\label{sec:predictions}

### 2024-09-06 src/tex/ms.tex
move some old intro stuff into caveats section

@@
 Deep-sea hydrothermal scenarios require planetary conditions that may not be met on ocean worlds with large amounts of water, where the water pressure on the ocean floor is high enough to form high-pressure ices~\citep{Noack2016}.
+% TODO Sleep et al., 2011?
+Sobotta et al., 2020?
+\todo[inline]{see discussion in Kite \& Ford 2018 Sect. 6.4}
 In this case, a testable prediction would be that planets with high-pressure ices do not show biosignatures.
@@
 We use this metric, known as the statistical power, to quantify the potential information content of the survey, identify critical design trades, and find strategies that maximize the survey’s scientific return.
-
-\subsection{Origins of Life Scenarios and their Predictions}\label{sec:predictions}
-\todo[inline]{Present widely discussed OOL scenarios and their predictions on exoplanet observables; derive testable hypotheses.}
-\label{sec:ool_scenarios}
-In this section, we present some of the most prominent origins of life scenarios and their observational predictions.
-We focus on the necessary environmental conditions for the processes and reactions inherent to each scenario, and aim to identify distinct observables that are accessible via present and near-future remote sensing techniques.
-
-%\subsection{Hydrothermal vents}
-A widely regarded origins-of-life scenario is that abiogenesis happens in hydrothermal vents~\citep[e.g.,][]{Russel2010}.
-...
-The hydrothermal vents scenario requires a direct contact of an ocean and the planetary mantle/crust.
-This requirement is not met on an ocean world with large amounts of water, where the water pressure on the ocean floor is high enough to form high-pressure ices (Noack+2016).
-% TODO Sleep et al., 2011?
-Sobotta et al., 2020?
-\todo[inline]{see discussion in Kite \& Ford 2018 Sect. 6.4}
-\todo[inline]{SR: The sealing away of the planetary interior from the ocean due to high-pressure ice layers is a common assumption for water world exoplanets (in addition to the references above, see e.g.
-Hu et al.
-2021).
-I'm not convinced it is correct, because of relatively recent evidence showing the possibility of molecular assimitation into such ices and subsequent transport, e.g., https://iopscience.iop.org/article/10.1088/0004-637X/769/1/29/meta, https://iopscience.iop.org/article/10.3847/1538-4357/acb49a/meta, https://iopscience.iop.org/article/10.3847/1538-4357/aa5cfe/meta (I'm sure there are other workers in this area, this is just the group with which I am familiar).
-Exoplaneteers mostly uniformly accept this proposition, so it's not an unreasonable assumption if you want to run with it so long as its acknowledged and caveated reasonably; I'm just highlighting this for your attention so that you can make an informed decision.
-}
-
-\textbf{Prediction:} Planets with high-pressure ices do not show biosignatures.
-
-%\subsection{Subaerial ponds}
-A different scenario is the emergence of life in hot springs or ponds that are exposed to the planet's atmosphere~\citep[e.g.,][]{Deamer2019}.
-...
-By its nature, the subaerial ponds scenario relies on rock surfaces exposed to the planetary atmosphere.
-Water worlds that have their entire planetary surface covered with water contradict this requirement and do not allow for the wet-dry cycling inherent to this origin of life scenario.
-The competition of tectonic stress with gravitational crustal spreading (Melosh 2011) sets the maximum possible height of mountains, which in the solar system does not exceed $\sim \SI{20}{\kilo\meter}$.
-Such mountains will be permanently under water on water worlds.
-Another impediment to wet-dry cycles is tidal locking of the planet as it stalls stellar tide-induced water movement and diurnal irradiation variability.
 
@@
 
+\subsubsection{Existence of an atmosphere-crust interface}
+By its nature, cyanosulfidic scenario relies on rock surfaces exposed to the planetary atmosphere.
+Water worlds that have their entire planeary surface covered with water contradict this requireent and do not allow for the wet-dry cycling inherent to this origin of life scenario.
+The competition of tectonic stress with gravitational crustal spreading~\citep{Melosh2011} sets the maximum possible height of mountains, which in the solar system does not exceed $\SI{\sim 20}{\kilo\meter}$.
+Such mountains will be permanently underwater on water worlds.
+Another impediment to wet-dry cycles may be tidal locking of the planet as it stalls stellar tide-induced water movement and diurnal irradiation variability.
+
 \subsubsection{Stellar flares}

### 2024-09-06 src/tex/ms.tex
include some comments in semi-analytical part

@@
 2023).
+\todo[inline]{SR: This is an interesting number, because it is below what was available on early Earth (so this scenario could have worked on early Earth) but until recently it was below what was thought to be available on habitable zone M-dwarf exoplanets.
+So it was thought that identification of biosignatures on M-dwarf planets could therefore falsify the cyanosulfidic scenario, with a potential caveat for transient UV from flares.
+Two recent developments have complicated the picture.
+First, Rimmer et al.
+2018 had an error in their radiative transfer routines.
+Correcting for this error, early M-dwarfs and highly active M-dwarfs emit enough UV to meet the Rimmer et al.
+2018 criterion (Ranjan et al.
+2023).
+Second, a recent publication argues that /all/ estimates of M-dwarf UV are underestimates, and that late M-dwarf stars have similar emission to Sunlike stars (Rekhi et al.
+2023).
+I suspect this is incorrect, because it contradicts a lot of work from e.g.
+the MUSCLES collaboration and the HAZMAT project, but it's worth keeping an eye on in case it is correct after all.}
 
@@
 Under the null hypothesis ($H_0$), that probability simply is $f_\mathrm{life}$, that is, it does not correlate with the UV flux.
+\begin{figure*}
+    \begin{centering}
+        \includegraphics[width=\hsize]{figures/hypotheses.pdf}
+        \caption{UV~Threshold Hypothesis and null hypothesis derived from the cyanosulfidic scenario.}
+        \label{fig:hypotheses}
+    \end{centering}
+\end{figure*}
+Figure~\ref{fig:hypotheses} shows these hypotheses as derived from the predictions of the cyanosulfidic scenario.
 Given a sample of planets, where for some of them we have convincing biosignature detections but remain agnostic on $f_\mathrm{life}$, we ask what evidence for $H_1$ and $H_0$ we can expect to obtain.
@@
 
-\textbf{Prediction:} Biosignatures occur outside the tidal locking zone and at bulk densities consistent with exposed rock.
-\todo[inline]{SR: I'm not a priori sold that tidal locking means that no wet-dry cycles occur.
+%\textbf{Prediction:} Biosignatures occur outside the tidal locking zone and at bulk densities consistent with exposed rock.
+%\todo[inline]{SR: I'm not a priori sold that tidal locking means that no wet-dry cycles occur.
 you can still have cycling driven by transient changes in instellation due to flares, for example (e.g., https://iopscience.iop.org/article/10.3847/1538-4357/aadfd1/meta).
@@
 Other papers which look relevant (but with which I am not familiar, as this is not my area): https://academic.oup.com/mnras/article/511/1/440/6501216, https://academic.oup.com/mnras/article/495/1/1/5733176, https://iopscience.iop.org/article/10.3847/1538-3881/aad775/meta, https://iopscience.iop.org/article/10.3847/1538-3881/aaed3a/meta, https://iopscience.iop.org/article/10.3847/1538-3881/ab2df3/meta}
-
-%\subsection{UV flux}\label{sec:predictions:uv}
-A major hypothesis in the origin of life is the ``\textit{UV Threshold Hypothesis}'', which states that UV light played a constructive role in getting life started on Earth (see Ranjan et al.
-2016, 2017c; Rimmer et al.
-2018; Rapf \& Vaida 2016; Pascal et al.
-2012; Green et al.
-2021; and sources therein).
-
-If UV light is required to get life started, then there is a minimum planetary UV flux requirement to have an inhabited world.
-This requirement is set by competitor thermal processes; if the photo-reaction does not move forward at a rate faster than the competitor thermal process(es), then the abiogenesis scenario cannot function.
-On the other hand, abundant UV light vastly in excess of this threshold does not increase the probability of abiogenesis, since once the UV photochemistry is no longer limiting, some other thermal process in the reaction network will be rate-limiting process instead.
-Therefore, a putative dependence of life on UV light is best encoded as a step function (see, e.g., Ranjan et al. 2017c; Rimmer et al. 2018; Rimmer, Ranjan \& Rugheimer 2021).
-
-One origin-of-life scenario has been refined to the point where the threshold flux has been measured.
-The cyanosulfidic scenario has been shown to require a mean flux of at least $ F_\mathrm{NUV, min} = \SI{6.8\pm3.6 e10}{\photons\per\centi\meter\squared\per\second\per\nano\meter}$ integrated from \SIrange{200}{280}{\nano\meter} at the surface in order to function (Rimmer et al.
-2018; Rimmer et al.
-2021 Astrobiology; Rimmer et al.
-2023).
-\todo[inline]{SR: This is an interesting number, because it is below what was available on early Earth (so this scenario could have worked on early Earth) but until recently it was below what was thought to be available on habitable zone M-dwarf exoplanets.
-So it was thought that identification of biosignatures on M-dwarf planets could therefore falsify the cyanosulfidic scenario, with a potential caveat for transient UV from flares.
-Two recent developments have complicated the picture.
-First, Rimmer et al.
-2018 had an error in their radiative transfer routines.
-Correcting for this error, early M-dwarfs and highly active M-dwarfs emit enough UV to meet the Rimmer et al.
-2018 criterion (Ranjan et al.
-2023).
-Second, a recent publication argues that /all/ estimates of M-dwarf UV are underestimates, and that late M-dwarf stars have similar emission to Sunlike stars (Rekhi et al.
-2023).
-I suspect this is incorrect, because it contradicts a lot of work from e.g.
-the MUSCLES collaboration and the HAZMAT project, but it's worth keeping an eye on in case it is correct after all.}
-We use this threshold value as our baseline case.
-
-\textbf{Prediction:} Past UV flux and the occurrence of biosignatures are correlated.
 
@@
 
+This paper is organized as follows:
+In Section~\ref{sec:methods}, we introduce both our semi-analytical approach and Bioverse simulations for testing the UV Threshold Hypothesis.
+Section~\ref{sec:results} presents the results of these experiments for a generic survey as well as for a realistic transit survey.
+In Section~\ref{sec:discussion}, we discuss our findings before concluding with a summary in Section~\ref{sec:conclusions}.
+
 %\input{bayesian_analysis}
@@
 %\subsection{Origins of Life Hypotheses}
-...
-\begin{figure*}
-    \begin{centering}
-        \includegraphics[width=\hsize]{figures/hypotheses.pdf}
-        \caption{UV~Threshold Hypothesis and null hypothesis derived from the cyanosulfidic scenario.}
-        \label{fig:hypotheses}
-    \end{centering}
-\end{figure*}
-Figure~\ref{fig:hypotheses} shows the hypothesis and null hypothesis derived from the predictions of the cyanosulfidic scenario.
-...
 
 \section{Methods}
-\label{sec:hypotests}
+\label{sec:methods}
 %\todo[inline]{Briefly introduce Bayesian model comparison, then present the particular hypotheses in their mathematical form.}
@@
 \subsection{Fraction of inhabited planets with detectable biosignatures}
-Presumably, not all habitable worlds are inhabited and not all inhabited worlds develop detectable biosignatures.
-The fraction of \glspl{EEC} that are both inhabited and harbor detectable biosignatures at the time when we observe them remains speculative; we aggregate them in the unitless parameter $f_\mathrm{life}$.
 
@@
 
-Let our observable be the inferred past \gls{NUV} flux of the planet $F_\mathrm{NUV}$.
+Presumably, not all habitable worlds are inhabited and not all inhabited worlds develop detectable biosignatures.
+The fraction of \glspl{EEC} that are both inhabited and harbor detectable biosignatures at the time when we observe them remains speculative; we aggregate them in the unitless parameter $f_\mathrm{life}$.
+Let us consider the probability to detect a biosignature $P(L)$, and let our observable be the inferred past \gls{NUV} flux of the planet $F_\mathrm{NUV}$.
 Under Hypothesis $H_1$ (Equation~\ref{eq:hypothesis-uv}), there exists a special unknown value of $F_\mathrm{NUV}$, noted $F_\mathrm{NUV, min}$ such that
@@
 \begin{align}
-    P(L|\theta,H_1) &=  f_\mathrm{life} \quad \text{if } \theta>F_\mathrm{NUV, min}\\
-    P(L|\theta,H_1) &=  0               \quad  \text{otherwise}
+    P(L|F_\mathrm{NUV},H_1) &=  f_\mathrm{life} \quad \text{if } F_\mathrm{NUV}>F_\mathrm{NUV, min}\\
+    P(L|F_\mathrm{NUV},H_1) &=  0               \quad  \text{otherwise}
 \end{align}
@@
 \begin{equation}
-P(L|\theta,H_\mathrm{null}) = f_\mathrm{life}.
+P(L|F_\mathrm{NUV},H_\mathrm{null}) = f_\mathrm{life}.
 \end{equation}
@@
 Hence,
+\begin{subequations}
 \begin{equation}\label{eq:bayes_factor}
-BF_{H_1,H_\mathrm{null}} = \frac{P(Y=k|H_1)}{P(Y=k|H_\mathrm{null})} = \frac{\binom{n_\lambda}{k}}{\binom{n}{k}}(1-f_\mathrm{life})^{n_{\lambda}-n}.
+BF_{H_1,H_\mathrm{null}} = \frac{P(Y=k|H_1)}{P(Y=k|H_\mathrm{null})} = \frac{\binom{n_\lambda}{k}}{\binom{n}{k}}(1-f_\mathrm{life})^{n_{\lambda}-n},
 \end{equation}
+and
+\begin{equation}\label{eq:bayes_factor2}
+BF_{H_\mathrm{null},H_1} = \frac{P(Y=k|H_\mathrm{null})}{P(Y=k|H_1)} = \frac{\binom{n}{k}}{\binom{n_\lambda}{k}}(1-f_\mathrm{life})^{n-n_{\lambda}}.
+\end{equation}
+\end{subequations}
 
@@
     A common definition of the null hypothesis $H_0$, which is also applied here, is that there is no relationship between the independent variable (here: maximum NUV flux) and the dependent variable (here: biosignature occurrence).
-%    The alternative hypothesis $H_1$ proposes a specific relationship parameterized by $\theta$.
     The alternative hypothesis $H_1$ proposes a specific relationship between the independent and dependent variables.
@@
 \section{Conclusions}
+\label{sec:conclusions}
 We have investigated the potential of upcoming exoplanet surveys to test the hypothesis -- motivated by the cyanosulfidic origins-of-life scenario -- that a minimum past \gls{NUV} flux is required for abiogenesis.

### 2024-09-07 src/tex/ms.tex
remove duplicate text

@@
 To assess the information gain of a realistic exoplanet survey, we employed Bioverse~\citep{Bixel2021,Hardegree-Ullman2023,Schlecker2024,Hardegree-Ullman2024}, a framework that integrates multiple components including statistically realistic simulations of exoplanet populations, a survey simulation module, and a hypothesis testing module to evaluate the statistical power of different observational strategies.
-The general approach is as follows:
-After defining a population of synthetic exoplanets, we simulate the detection and characterization of these exoplanets with a hypothetical survey, taking into account the survey’s sensitivity, target selection, and observational biases.
-Then we evaluate the likelihood that a given survey would detect a specified demographic trend in the exoplanet population and estimate the precision with which the survey could constrain the parameters of that trend.
-A common definition of the null hypothesis $H_0$, which is also applied here, is that there is no relationship between the independent variable (here: maximum \gls{NUV} flux) and the dependent variable (here: biosignature occurrence).
-The alternative hypothesis $H_1$ proposes a specific relationship between the independent and dependent variables.
-To determine the diagnostic capability of a given survey, Bioverse runs multiple iterations of the simulated survey and calculates the fraction of realizations that successfully reject the null hypothesis.
-We use this metric, known as the statistical power, to quantify the potential information content of the survey, identify critical design trades, and find strategies that maximize the survey’s scientific return.
 
@@
 \subsection{Exoplanet survey simulations with Bioverse}
-A real exoplanet survey will be subject to observational biases, sample selection effects, and the underlying demographics of the planet sample.
-To assess the information gain of a realistic exoplanet survey, we employed Bioverse~\citep{Bixel2021}, a framework that integrates multiple components including statistically realistic simulations of exoplanet populations, a survey simulation module, and a hypothesis testing module to evaluate the statistical power of different observational strategies.
+To assess the diagnostic power of realistic exoplanet surveys, we employed our survey simulator and hypothesis testing framework Bioverse~\citep{Bixel2021}
 The general approach is as follows:

### 2024-09-17 src/tex/ms.tex
motivate semi-analytical approach

@@
 \subsection{Semi-analytical approach}\label{sec:met-semianalytical}
-First, we apply a Bayes Factor Design Analysis~\citep{Schoenbrodt2018} to assess the expected probabilities of obtaining true negative or true positive evidence for the hypothesis above, as well as the probability for misleading or inconclusive evidence, under idealized conditions.
+%First, we apply a Bayes Factor Design Analysis~\citep{Schoenbrodt2018}
+We first assessed the expected probabilities of obtaining true negative or true positive evidence for the UV Threshold Hypothesis ($H_1$) above, as well as the probability for misleading or inconclusive evidence, under idealized conditions.
 This serves as a first-order estimate of the information content of a survey, before we take into account impacts from exoplanet demographics, sample selection, and survey strategy.
@@
 
-For all survey simulations and hypothesis tests, we repeated the above in a Monte Carlo fashion to generate randomized ensembles of synthetic star and planet populations~\citep[compare][]{Bixel2021}.
+For all survey simulations and hypothesis tests, we repeated the above in a Monte Carlo fashion to generate randomized ensembles of synthetic star and planet populations~\citep[][]{Bixel2021}.
 
@@
 All common Origin of Life scenarios require water as a solvent;
-we thus consider only rocky planets that may sustain liquid water on their surface, i.e., that occupy their momentary \gls{HZ} during the above period.
+we thus consider only rocky planets that may sustain liquid water on their surface, i.e., that occupy their host star's momentary \gls{HZ} during the above period.
+This takes into account the evolution of the host star's luminosity and \gls{HZ} boundaries.
 

### 2024-09-20 src/tex/ms.tex
implement Sukrit's comments

@@
 2023).
-Second, a recent publication argues that /all/ estimates of M-dwarf UV are underestimates, and that late M-dwarf stars have similar emission to Sunlike stars (Rekhi et al.
+\sout{Second, a recent publication argues that /all/ estimates of M-dwarf UV are underestimates, and that late M-dwarf stars have similar emission to Sunlike stars (Rekhi et al.
 2023).
 I suspect this is incorrect, because it contradicts a lot of work from e.g.
-the MUSCLES collaboration and the HAZMAT project, but it's worth keeping an eye on in case it is correct after all.}
+the MUSCLES collaboration and the HAZMAT project, but it's worth keeping an eye on in case it is correct after all.}}
 
@@
 \subsection{Exoplanet survey simulations with Bioverse}
-To assess the diagnostic power of realistic exoplanet surveys, we employed our survey simulator and hypothesis testing framework Bioverse~\citep{Bixel2021}
+To assess the diagnostic power of realistic exoplanet surveys, we employed our survey simulator and hypothesis testing framework Bioverse~\citep{Bixel2021}.
 The general approach is as follows:
@@
 \item \textbf{Exoplanet population synthesis:} We populate the Gaia Catalogue of Nearby Stars~\citep{Smart2021} with synthetic exoplanets whose orbital parameters and planetary properties reflect our current understanding of exoplanet demographics~\citep{Bergsten2022}.
+Here, we also inject the demographic trend in question - in this case we assign biosignatures according to $H_1$, i.e., to planets in the \gls{HZ} that have received \gls{NUV} fluxes above a certain threshold.
     \item \textbf{Survey simulation:} We simulate the detection and characterization of these exoplanets with a hypothetical survey, taking into account the survey's sensitivity, target selection, and observational biases.
@@
     Bioverse offers either Bayesian model comparison or non-parametric tests to evaluate the evidence for or against the null hypothesis.
+\end{enumerate}
 
@@
 %The posterior samples obtained from the nested sampling runs further allows us to estimate the precision with which the survey could constrain the parameters of the hypothesized trend.
-\end{enumerate}
 
@@
 All common Origin of Life scenarios require water as a solvent;
-we thus consider only rocky planets that may sustain liquid water on their surface, i.e., that occupy their host star's momentary \gls{HZ} during the above period.
+we thus consider only rocky planets that may sustain liquid water on their surface, i.e., that occupy their host star's momentary \gls{HZ} during the above period, as well as at the time of observation.
+\todo[inline]{Does this need more detailed explanation?}
 This takes into account the evolution of the host star's luminosity and \gls{HZ} boundaries.
@@
 
-We then determined which planets were both in the \gls{HZ} and had \gls{NUV} fluxes above $F_\mathrm{NUV, min}$ for $\Delta T_\mathrm{min} \geq \var{deltaT_min}\,\SI{}{\mega\year}$.
+We then determined which planets were both in the \gls{HZ} and had \gls{NUV} fluxes above $F_\mathrm{NUV, min}$.
+To avoid considering short transitional phases, we require this situation to last for a minimum duration $\Delta T_\mathrm{min} \geq \var{deltaT_min}\,\SI{}{\mega\year}$.
 We assigned the development of life to a random fraction $f_\mathrm{life}$ of all temperate planets fulfilling these requirements.
@@
 \subsection{Semi-analytical assessment}\label{sec:results-semianalytical}
-In Section~\ref{sec:met-semianalytical} we computed the probability for true positive evidence for $H_\mathrm{1}$ and $H_\mathrm{null}$, respectively.
+In Section~\ref{sec:met-semianalytical} we computed the probability for true positive evidence for $H_\mathrm{1}$ and $H_\mathrm{null}$, respectively (Equations~\ref{eq:bayes_factor}, \ref{eq:bayes_factor2}).
 Figure~\ref{fig:semian_true_evidence} shows how these evidences are distributed for sample sizes \var{semian_Nsamp1} and \var{semian_Nsamp2}, and how likely we are to obtain strong evidence ($BF_{H_i, H_j}$ > 10).
@@
 However, one might `cherry-pick' exoplanets for which a biosignature test is performed based on \textit{a priori} available contextual information~\citep{catling2018exoplanet} in order to maximize the science yield of investing additional resources.
-For instance, the distribution of $F_\mathrm{NUV}$ in the planet sample can be influenced by the survey strategy, and a targeted sampling approach could favor extreme values.% of $F_\mathrm{NUV}$ in the sample selection.
-Figure~\ref{fig:semian_selectivity} shows how the probability of obtaining true strong evidence for $H_\mathrm{1}$ scales with selectivity $s$, where $s\in]-1,1[$ such that $F_\mathrm{NUV} \sim Beta(1/10^s,1/10^s)$.
-Here, $s=0$ corresponds to a random uniform distribution.
-Compared to this case, a high selectivity can increase the probability of obtaining true strong evidence to $\gtrsim \SI{85}{\percent}$ for large samples.
+For instance, the distribution of $F_\mathrm{NUV}$ in the planet sample can be influenced by the survey strategy, and a targeted sampling approach could favor extreme values.
+% of $F_\mathrm{NUV}$ in the sample selection.
+We model this by distributing $F_\mathrm{NUV}$ according to different $Beta$ functions and introduce a selectivity parameter $s\in]-1,1[$ such that $F_\mathrm{NUV} \sim Beta(1/10^s,1/10^s)$.
+Figure~\ref{fig:semian_selectivity} shows how the probability of obtaining true strong evidence for $H_\mathrm{1}$ scales with selectivity $s$.
+%Here, $s=0$ corresponds to a random uniform distribution.
+%Compared to this case, a high selectivity ($s \sim 1$) can increase the probability of obtaining true strong evidence to $\gtrsim \SI{85}{\percent}$ for large samples.
+For large samples, a high selectivity ($s \sim 1$) can increase the probability of obtaining true strong evidence from $\sim \SI{70}{\percent}$ for $s= 0$ (random uniform distribution) to $> \SI{90}{\percent}$.
 
@@
 Left: Sampling distribution for different selectivity parameters $s$.
-Right: Resulting \mbox{P(true strong evidence)} (random, uniform $f_\mathrm{life}$, $F_\mathrm{NUV, min}$).
+Right: Resulting \mbox{P(true strong evidence)}, where $f_\mathrm{life}$ and $F_\mathrm{NUV, min}$ are randomly drawn from a uniform distribution.
 Sampling more extreme values of $F_\mathrm{NUV}$ is more likely to yield strong evidence.}
@@
 There, we assumed an abiogenesis rate of $f_\mathrm{life} = \var{f_life}$ and a minimum \gls{NUV} flux of $F_\mathrm{NUV, min} = \var{NUV_thresh}\,\SI{}{\erg\per\second\per\centi\meter\squared}$.
-All injected biosignatures are assumed to be detected, and the maximum \gls{NUV} flux is estimated from the host star's spectral type and age with an uncertainty corresponding to the intrinsic scatter in the \gls{NUV} fluxes in \citet{Richey-Yowell2023}.
+All injected biosignatures are assumed to be detected without false positive ambiguity, and the maximum \gls{NUV} flux is estimated from the host star's spectral type and age with an uncertainty corresponding to the intrinsic scatter in the \gls{NUV} fluxes in \citet{Richey-Yowell2023}.
 %\begin{figure*}
@@
 The only parameter space region with statistical power above \SI{90}{\percent} lies at abiogenesis rates $f_\mathrm{life} > 0.9$ and threshold \gls{NUV} fluxes of $\sim \SIrange{30}{70}{\erg\per\second\per\centi\meter\squared}$.
+\todo[inline]{Update this with final results from stat. power grid}
 
@@
 \subsection{Sampling strategy for testing a predicted minimum \gls{NUV} flux} % corresponds to conclusion point 1
-In Sect.~\ref{sec:results-semianalytical}, we show that testing the hypothesis of a minimum past \gls{NUV} flux required for abiogenesis suffers from `nuisance' parameters that render inference difficult.
+In Sect.~\ref{sec:results-semianalytical}, we show that testing the hypothesis of a minimum past \gls{NUV} flux required for abiogenesis suffers from `nuisance' parameters that render inference through astronomical observations difficult.
 Here, these parameters are the unspecified value of the \gls{NUV} threshold hypothesized to exist under $H_1$, and the unknown probability of detectable life emerging on a habitable planet $f_{\mathrm{life}}$.
@@
 The required threshold flux is, of course, a priori unknown and we cannot influence it.
-If, however, better theoretical predictions for the required \gls{NUV} flux for abiogenesis become available, the survey strategy can be further optimized, for instance by targeting planets that are estimated to have received a \gls{NUV} flux slightly below and above this threshold.
+If, however, better theoretical predictions for the required \gls{NUV} flux for abiogenesis become available~\citep{Rimmer2021}, the survey strategy can be further optimized, for instance by targeting planets that are estimated to have received a \gls{NUV} flux slightly below and above this threshold.
 
@@
 Taking the infered fluxes from \citet{Richey-Yowell2023} at face value (but taking into account intrinsic scatter), a minimum required \gls{NUV} flux of $\gtrsim\SI{400}{\erg\per\second\per\centi\meter\squared}$ reduces the fraction of inhabited planets to below $\sim\SI{1}{\percent}$.
-This not only calls for a large sample size and a targeted sample selection prefering high expected past \gls{NUV} fluxes, but also highlight the necessity of continued theoretical and experimental research into the role of UV radiation in prebiotic chemistry~\citep{Ranjan2017b}.
+This not only calls for a large sample size and a targeted sample selection prefering high expected past \gls{NUV} fluxes, but also highlight the necessity of continued theoretical and experimental research into the role of UV radiation in prebiotic chemistry~\citep{Ranjan2017b,Rimmer2018,Rimmer2021}.
 
@@
 The predicted interplay of \gls{NUV} flux and \gls{HZ} occupancy in enabling abiogenesis via the cyanosulfidic scenario could in principle be used to add or remove credibility from a tentative biosignature detection.
-For example, with a strong believe that this scenario is the only viable one for the origins of life, a biosignature detection on a planet orbiting a strongly UV-radiating star may add credibility to the detection.
+For example, with a strong belief that this scenario is the only viable one for the origins of life, a biosignature detection on a planet orbiting a strongly UV-radiating star may add credibility to the detection.
 Conversely, a biosignature detection on a planet estimated to have received very little UV~radiation would increase the likelihood of a false positive detection.
@@
 \subsubsection{Stellar flares}
-Our assumptions on past UV flux neglect the contribution of stellar flares, which may be hypothesized as an alternative source of UV light~\citep{Ranjan2017}.
+Our assumptions on past UV flux neglect the contribution of stellar flares, which may be hypothesized as an alternative source of UV light~\citep{Buccino2007,Ranjan2017c}.
 This concerns mainly ultracool dwarfs, due to their low quiescent emission and high pre-main sequence stellar activity~\citep{Buccino2007,West2008}.
@@
 \subsubsection{Atmosphere transmission}
+We do not take into account absorption of UV radiation by the planetary atmosphere.
 Theoretical work suggests that the atmosphere of prebiotic Earth was largely transparent at \gls{NUV} wavelengths with the only known source of attenuation being Rayleigh scattering~\citep{Ranjan2017,Ranjan2017c}.
 We thus approximated surface UV flux using top-of-atmosphere fluxes.
-This represents a conservative approach, since any planet that fails to meet the irradiance criterion receives even lower \gls{NUV} radiation at its surface.
-\todo[inline]{SUKRIT please specify as neccessary}
+If there are planets in a sample that do not have a transparent atmosphere at \gls{NUV} wavelengths and require higher fluxes for abiogenesis, the fraction of inhabited planets in the sample will be lower.
+However, these planets will not pollute the below-threshold subsample, as they will not be able to host life under the UV~Threshold Hypothesis.
+
 %\subsubsection{Diurnal cycles}

### 2024-09-26 src/tex/ms.tex
draft a title
highlight M dwarfs in Results

@@
 % Title
-\title{Bioverse: Origins of Life}
+\title{Bioverse: Exoplanet Biosignature Detection Under the UV Threshold Hypothesis}
+% The Role of Ultraviolet Radiation in Abiogenesis: Insights from Exoplanet Survey Simulations
+% Bioverse: The Impact of UV Flux on Exoplanet Biosignatures under the Cyanosulfidic Origins of Life Scenario
 
@@
 In a magnitude- and volume-limited sample of a transit survey, the host star distribution for \glspl{EEC} will be skewed toward later spectral types and dominated by M~dwarfs due to how the \gls{HZ} scales with spectral type.
-For a fixed planet sample size, the fraction of inhabited planets is highest in the M~dwarf sample due to the higher \gls{NUV} fluxes in the \gls{HZ} of these stars (see Figure~\ref{fig:surveys_FGKM}).
+For a fixed planet sample size, the fraction of inhabited planets is highest in the M~dwarf sample due to the higher \gls{NUV} fluxes in the \gls{HZ} of these stars (see Figures~\ref{fig:hz_nuv_evo}, \ref{fig:surveys_FGKM}).
 Their \gls{NUV} fluxes are generally highest at early times $\lesssim\SI{100}{\mega\year}$.
 These host stars, in particular late subtypes, also provide extended periods of increased \gls{NUV} emission that overlap with times when some of these planets occupy the \gls{HZ} (see Figure~\ref{fig:hz_nuv_evo}), our requirement for abiogenesis (compare Equation~\ref{eq:hypothesis-uv}).
-%Because of that, most inhabited planets in our magnitude- and volume-limited sample orbit M~dwarfs.
+Thus -- under the UV Threshold Hypothesis, most inhabited transiting planets in a magnitude- and volume-limited sample orbit M~dwarfs.
 

### 2024-09-26 src/tex/ms.tex
highlight M dwarfs in Discussion

@@
 The required threshold flux is, of course, a priori unknown and we cannot influence it.
-If, however, better theoretical predictions for the required \gls{NUV} flux for abiogenesis become available~\citep{Rimmer2021}, the survey strategy can be further optimized, for instance by targeting planets that are estimated to have received a \gls{NUV} flux slightly below and above this threshold.
+If, however, better theoretical predictions for the required \gls{NUV} flux for abiogenesis become available~\citep{Rimmer2021}, the survey strategy can be further optimized, for instance by targeting planets that are estimated to have received a \gls{NUV} flux slightly below and above this threshold or by applying a bisection algorithm in a sequential survey~\citep{Fields2023}.
 
+\subsection{An M~dwarf opportunity}
 An interesting aspect lies in the distribution of host star properties, as different spectral types probe different past \gls{NUV} flux regimes.
-Under the UV~Threshold Hypothesis, late-type M~dwarfs are favored due to their higher \gls{NUV} fluxes in the \gls{HZ}~\citep{Richey-Yowell2023}.
-Furthermore, their wider distribution of maximum past \gls{NUV} fluxes may increase the likelihood of obtaining strong evidence for or against the hypothesis.
-An M~dwarf sample may help to test the high \gls{NUV} flux end of the hypothesis; a higher occurrence of biosignatures here would support the hypothesis that a higher NUV flux is beneficial or necessary for life.
-
-FGK stars, on the other hand, show a more narrow distribution of maximum past \gls{NUV} fluxes in the \gls{HZ}, which may limit the constraining power of the survey depending on the (unknown) threshold \gls{NUV} flux.
-With this sample, the survey will be more sensitive to the low \gls{NUV} flux end of the hypothesis.
+FGK stars show a narrow distribution of maximum past \gls{NUV} fluxes in the \gls{HZ}, which may limit the constraining power of a survey depending on the (unknown) threshold \gls{NUV} flux.
+With a pure FGK sample, a survey will be more sensitive to the low \gls{NUV} flux end of the hypothesis.
 A lack of biosignatures on these planets would support the UV~Threshold Hypothesis, whereas their presence might suggest that lower NUV fluxes are also sufficient for abiogenesis or indicate a different abiogenic pathway.
 CITE
+
+On the other hand, M~dwarfs show a wider distribution of maximum past \gls{NUV} fluxes in their \glspl{HZ}.
+While old M~dwarfs can be considered low-UV environments, a large fraction of them emit high \gls{NUV} fluxes into their \gls{HZ} during their early stages, in particular later subtypes~\citep{Richey-Yowell2023}.
+an M~dwarf sample will thus help to test the high \gls{NUV} flux end of the UV Threshold Hypothesis; a higher occurrence of biosignatures here would support the hypothesis that a higher NUV flux is beneficial or necessary for life.
+As our simulations show, the higher and more variable \gls{NUV} fluxes in M~dwarfs increase the likelihood of obtaining strong evidence for or against the hypothesis.
+The combination of a lack of UV radiation today, which makes biosignature gases more detectable~\citep{Segura2005}, and a UV-rich past that may have enabled abiogenesis could make M~dwarfs the preferred targets for biosignature searches.
 

### 2024-09-26 src/tex/ms.tex
adapt a few minor wordings

@@
 From this, we get each planet's epochs within and outside the \gls{HZ}.
-For the \gls{NUV} flux, we use the age- and stellar mass-dependent \gls{NUV} fluxes in the \gls{HZ} obtained by \citet{Richey-Yowell2023}.
+
+For the \gls{NUV} flux, we use the age- and stellar mass-dependent \gls{NUV} fluxes in the \gls{HZ} obtained by \citet{Richey-Yowell2023}, which considers GALEX UV data in the wavelength range of \SIrange{177}{283}{\nano\meter}.
 We linearly interpolate in their measured grid, where we convert spectral type to stellar mass using the midpoints of their mass ranges (\SI{0.75}{\Msun} for K~stars, \SI{0.475}{\Msun} for early-type M~stars,  and \SI{0.215}{\Msun} for late-type M~stars).
@@
         In an M~dwarf sample of the same size, the fraction of inhabited planets is larger.\\
-        Center: Distribution of maximum past \gls{NUV} flux in transit surveys targeting \glspl{EEC} around FGK and M~stars, respectively.
+        Center: Distribution of inferred maximum past \gls{NUV} flux in transit surveys targeting \glspl{EEC} around FGK and M~stars, respectively.
 The best-fit beta distributions (gray) correspond to selectivities of $s_\mathrm{FGK} = $\var{selectivity_FGK} and $s_\mathrm{M} = $\var{selectivity_M}.
@@
 \begin{enumerate}
-    \item The UV~flux requirement in the cyanosulfidic scenario for the origins of life should lead to a correlation between past \gls{NUV} flux and current occurrence of biosignatures that may be observationally testable.
+    \item The UV Threshold Hypothesis of the cyanosulfidic scenario for the origins of life should lead to a correlation between past \gls{NUV} flux and current occurrence of biosignatures that may be observationally testable.
     \item The required sample size for detecting this correlation depends on the abiogenesis rate on temperate exoplanets and the distribution of host star properties in the sample; in particular their maximum past \gls{NUV} fluxes.
@@
     \item Samples of planets orbiting M~dwarfs overall yield higher chances of successfully testing the UV~Threshold Hypothesis.
-        Leaving other factors aside, they are also more likely to yield biosignature detections under this hypothesis.
+          They may also be more likely to yield biosignature detections under this hypothesis.
 

### 2024-09-27 src/tex/ms.tex
clean up references in introduction

@@
 \label{sec:intro}
-\todo[inline]{Introduce OOL, the importance of planetary contexts}
-A wide variety of scenarios for the origin of life have been proposed~\citep[e.g.,][]{Dass2016,Westall2018}.
+A wide variety of scenarios for the origin of life have been proposed~\citep[e.g.,][]{Baross1985,Brasier2011,Mulkidjanian2012,Fox2013,Deamer2015,Westall2018}.
 While we are still far away from testing any of them, new prospects in the search for conditions favorable to life have opened up by thinking of the origin of life as a planetary phenomenon and identifying global-scale environmental properties that might support pathways to life~\citep{Sasselov2020}.
@@
 Such planetary conditions can be hypothesized for exoplanets located in the habitable zone of their host star, with persistent liquid water on their surface.
-For example, the deep-sea hydrothermal (or hydrothermal-sedimentary) context for the origin of life requires a direct contact of an ocean and the planetary mantle/crust~\citep[e.g.,][]{Russel2010}.
+For example, the deep-sea hydrothermal (or hydrothermal-sedimentary) context for the origin of life requires a direct contact of an ocean and the planetary mantle/crust~\citep[e.g.,][]{Baross1985}.
 The alternate scenario of a surface locally subject to wet-dry cycles requires a planetary exposure to mid-range UV irradiation, as a source of energy and an agent of selection in chemical evolution~\citep[e.g.,][]{Deamer2019}.
-This “UV Threshold Hypothesis” states that UV light in a specific length range played a constructive role in getting life started on Earth (CITE see Ranjan et al.
-2016, 2017c; Rimmer et al.
-2018; Rapf \& Vaida 2016; Pascal et al.
-2012; Green et al.
-2021; and sources therein).
+This “UV Threshold Hypothesis” states that UV light in a specific wavelength range played a constructive role in getting life started on Earth~\citep{Ranjan2016,Ranjan2017c,Rimmer2018,Rapf2016}.
 
@@
 On the other hand, abundant UV light vastly in excess of this threshold does not increase the probability of abiogenesis, since once the UV photochemistry is no longer limiting, some other thermal process in the reaction network will be the rate-limiting process instead.
-Therefore, a putative dependence of life on UV light is best encoded as a step function (see, e.g., Ranjan et al. 2017c; Rimmer et al. 2018; Rimmer, Ranjan \& Rugheimer 2021).
+Therefore, a putative dependence of life on UV light is best encoded as a step function~\citep[e.g.,][]{Ranjan2017c,Rimmer2018,Rimmer2021}.
 
 The goal of this work is to evaluate the potential of upcoming exoplanet surveys to test the hypothesis that a minimum past NUV flux is required for abiogenesis.
-We focus on one version of the UV Threshold Hypothesis, the so-called cyanosulfidic scenario, which has been refined to the point where the required threshold flux has been measured to be $F_\mathrm{NUV, min} = \SI{6.8\pm3.6 e10}{\photons\per\centi\meter\squared\per\second\per\nano\meter}$ integrated from \SIrange{200}{280}{\nano\meter} at the surface (Rimmer et al.
-2018; Rimmer et al.
-2021 Astrobiology; Rimmer et al.
-2023).
-\todo[inline]{SR: This is an interesting number, because it is below what was available on early Earth (so this scenario could have worked on early Earth) but until recently it was below what was thought to be available on habitable zone M-dwarf exoplanets.
+We focus on one version of the UV Threshold Hypothesis, the so-called cyanosulfidic scenario, which has been refined to the point where the required threshold flux has been measured to be $F_\mathrm{NUV, min} = \SI{6.8\pm3.6 e10}{\photons\per\centi\meter\squared\per\second\per\nano\meter}$ integrated from \SIrange{200}{280}{\nano\meter} at the surface~\citep{Rimmer2018,Rimmer2021a,Rimmer2023,Ranjan2023a}.
+%\todo[inline]{SR: This is an interesting number, because it is below what was available on early Earth (so this scenario could have worked on early Earth) but until recently it was below what was thought to be available on habitable zone M-dwarf exoplanets.
 So it was thought that identification of biosignatures on M-dwarf planets could therefore falsify the cyanosulfidic scenario, with a potential caveat for transient UV from flares.
@@
 2023).
-\sout{Second, a recent publication argues that /all/ estimates of M-dwarf UV are underestimates, and that late M-dwarf stars have similar emission to Sunlike stars (Rekhi et al.
+%\sout{Second, a recent publication argues that /all/ estimates of M-dwarf UV are underestimates, and that late M-dwarf stars have similar emission to Sunlike stars (Rekhi et al.
 2023).
@@
 With a pure FGK sample, a survey will be more sensitive to the low \gls{NUV} flux end of the hypothesis.
-A lack of biosignatures on these planets would support the UV~Threshold Hypothesis, whereas their presence might suggest that lower NUV fluxes are also sufficient for abiogenesis or indicate a different abiogenic pathway.
-CITE
+A lack of biosignatures on these planets would support the UV~Threshold Hypothesis.
+Their presence would have little constraining power; it would either suggest that low NUV fluxes are sufficient for abiogenesis or indicate a different abiogenic pathway~\citep[e.g.,][]{Westall2018}.
 
@@
 The combination of a lack of UV radiation today, which makes biosignature gases more detectable~\citep{Segura2005}, and a UV-rich past that may have enabled abiogenesis could make M~dwarfs the preferred targets for biosignature searches.
+
+These findings underscore the importance of constraining the UV emission profiles of \gls{EEC} host stars throughout their evolutionary stages to assess the viability of M-dwarf planets as testbeds for theories on the origins of life~\citep{Rimmer2021,Ranjan2023a}.
 

### 2024-11-07 src/tex/ms.tex
add a few citations

@@
 The alternate scenario of a surface locally subject to wet-dry cycles requires a planetary exposure to mid-range UV irradiation, as a source of energy and an agent of selection in chemical evolution~\citep[e.g.,][]{Deamer2019}.
-This “UV Threshold Hypothesis” states that UV light in a specific wavelength range played a constructive role in getting life started on Earth~\citep{Ranjan2016,Ranjan2017c,Rimmer2018,Rapf2016}.
+This “UV Threshold Hypothesis” states that UV light in a specific wavelength range played a constructive role in getting life started on Earth~\citep{Ranjan2016,Ranjan2017c,Rimmer2018,Rapf2016}, and it could provide a probabilistic approach to the interpretation of possible future biosignature detections~\citep[e.g.,][]{Catling2018,Walker2018}.
 
 The association of chemical pathways to life and planetary environmental conditions offers a new opportunity to test alternate scenarios for life emergence based on planetary-level data collected from the upcoming observations of populations of exoplanets.
-Deep-sea hydrothermal scenarios require planetary conditions that may not be met on ocean worlds with large amounts of water, where the water pressure on the ocean floor is high enough to form high-pressure ices~\citep{Noack2016}.
+Deep-sea hydrothermal scenarios require planetary conditions that may not be met on ocean worlds with large amounts of water, where the water pressure on the ocean floor is high enough to form high-pressure ices~\citep{Noack2016,Kite2018}.
 % TODO Sleep et al., 2011?
 Sobotta et al., 2020?
-\todo[inline]{see discussion in Kite \& Ford 2018 Sect. 6.4}
+%\todo[inline]{see discussion in Kite \& Ford 2018 Sect. 6.4}
 In this case, a testable prediction would be that planets with high-pressure ices do not show biosignatures.
@@
 \subsection{Fraction of inhabited planets with detectable biosignatures}
-
 %\subsection{H1: Life only originates in hydrothermal vents}
@@
 %\subsection{H3: Life only originates on planets with particular UV irradiance}
-\todo[inline]{\citet{Guenther2020} relate U-band energy to bolometric flux.}
-\todo[inline]{OPTIONAL: ``We further test the scenario of a linear correlation of past UV flux and biosignature occurrence rate.
-This test requires the detection of multiple biosignatures.''
-~\\
-Test for negative correlations as well?}
+%\todo[inline]{\citet{Guenther2020} relate U-band energy to bolometric flux.}
+%\todo[inline]{OPTIONAL: ``We further test the scenario of a linear correlation of past UV flux and biosignature occurrence rate.
+%This test requires the detection of multiple biosignatures.''
+%~\\
+%Test for negative correlations as well?}
 Here, we conduct a theoretical experiment on the UV~Threshold Hypothesis (Sect.~\ref{sec:predictions}) by relating the occurrence of life on an exo-earth candidate with a minimum past quiescent stellar UV flux, focusing on the prebiotically interesting \gls{NUV} range from \SIrange{200}{280}{\nano\meter}.
@@
         \includegraphics[width=\hsize]{figures/hz_nuv_evo}
+%        \includegraphics[width=\hsize]{figures/hz_nuv_evo.bak.png}
         \caption{Interpolated stellar luminosity evolution (left) and evolution of the \gls{NUV} flux in the \gls{HZ} (right) as a function of host star mass.
@@
         \includegraphics[width=\hsize]{figures/powergrid}
+%        \includegraphics[width=\hsize]{figures/powergrid.bak.png}
         \caption{Statistical power as a function of threshold \gls{NUV} flux and abiogenesis rate.

### 2024-11-16 src/tex/ms.tex
change title
edit methods, results
Daniel's comments

@@
 % Title
-\title{Bioverse: Exoplanet Biosignature Detection Under the UV Threshold Hypothesis}
+%\title{Bioverse: Exoplanet Biosignature Detection Under the UV Threshold Hypothesis}
+\title{Bioverse: Exoplanet Biosignature Patterns Under the UV Threshold Hypothesis for the Origin of Life}
 % The Role of Ultraviolet Radiation in Abiogenesis: Insights from Exoplanet Survey Simulations
@@
 We focus on one version of the UV Threshold Hypothesis, the so-called cyanosulfidic scenario, which has been refined to the point where the required threshold flux has been measured to be $F_\mathrm{NUV, min} = \SI{6.8\pm3.6 e10}{\photons\per\centi\meter\squared\per\second\per\nano\meter}$ integrated from \SIrange{200}{280}{\nano\meter} at the surface~\citep{Rimmer2018,Rimmer2021a,Rimmer2023,Ranjan2023a}.
-%\todo[inline]{SR: This is an interesting number, because it is below what was available on early Earth (so this scenario could have worked on early Earth) but until recently it was below what was thought to be available on habitable zone M-dwarf exoplanets.
+%\todo[inline]{SR: This is an interesting number, because it is below what was available on early Earth (so this scenario could have worked on early Earth) but until recently it was below what was thought to be available on \gls{HZ} M-dwarf exoplanets.
 So it was thought that identification of biosignatures on M-dwarf planets could therefore falsify the cyanosulfidic scenario, with a potential caveat for transient UV from flares.
@@
 \begin{enumerate}
-\item \textbf{Exoplanet population synthesis:} We populate the Gaia Catalogue of Nearby Stars~\citep{Smart2021} with synthetic exoplanets whose orbital parameters and planetary properties reflect our current understanding of exoplanet demographics~\citep{Bergsten2022}.
+\item \textbf{Exoplanet population synthesis:} We populate the Gaia Catalogue of Nearby Stars~\citep{Smart2021} with synthetic exoplanets whose orbital parameters and planetary properties reflect our current understanding of exoplanet demographics~\citep{Bergsten2022,Hardegree-Ullman2023}.
 Here, we also inject the demographic trend in question - in this case we assign biosignatures according to $H_1$, i.e., to planets in the \gls{HZ} that have received \gls{NUV} fluxes above a certain threshold.
@@
 Planets were generated and assigned to the synthetic stars following the occurrence rates and size/orbit distributions of \citet{Bergsten2022}.
-Following \citet{Bixel2021}, we considered only transiting \glspl{EEC} with radii $0.8\, S^{0.25} < R < 1.4 $ that are within the habitable zone (see Section~\ref{sec:met-hz}).
+Following \citet{Bixel2021}, we considered only transiting \glspl{EEC} with radii $0.8\, S^{0.25} < R < 1.4 $ that are within the \gls{HZ} (see Section~\ref{sec:met-hz}).
 The lower limit was suggested as a minimum planet size to retain an atmosphere~\citep{Zahnle2017}.
 To generate planet samples larger than what the stellar catalog in combination with these occurrence rates yields, we scaled up the occurrence rates by a constant factor that yields the desired number of planets.
-This was in particular necessary for the FGK sample, where the rate of transiting planets that occupy the habitable zone is low.
+This was in particular necessary for the FGK sample, where the rate of transiting planets that occupy the \gls{HZ} is low.
 
@@
 All common Origin of Life scenarios require water as a solvent;
-we thus consider only rocky planets that may sustain liquid water on their surface, i.e., that occupy their host star's momentary \gls{HZ} during the above period, as well as at the time of observation.
+we thus consider only rocky planets that may sustain liquid water on their surface, i.e., that occupy their host star's momentary \gls{HZ} during the above period, as well as at the epoch of observation.
 \todo[inline]{Does this need more detailed explanation?}
-This takes into account the evolution of the host star's luminosity and \gls{HZ} boundaries.
+To determine \gls{HZ} occupancy, we take into account the evolution of the host star's luminosity and \gls{HZ} boundaries.
 
-Different formulations of habitable zones as regions around a star where a planet with Earth's atmospheric composition can maintain liquid water on its surface exist~\citep[e.g.,][]{MolLous2022,Spinelli2023,Tuchow2023}.
-CITE!
-Ramirez \& Kaltenegger 2017, 2018
-Here, we adopt the popular estimates of \citet{Kasting1993} and \citet{Kopparapu2013,Kopparapu2014} that define a temperate zone between the runaway greenhouse transition CITE!
-and the maximum greenhouse limit CITE!.
+The \gls{HZ} describes a region around a star where a planet with Earth's atmospheric composition and climate feedbacks can maintain liquid water on its surface~\citep[e.g.,][]{Ramirez2017,Ramirez2018,MolLous2022,Spinelli2023,Tuchow2023}.
+Here, we adopt estimates that define the \gls{HZ} as the region between the runaway greenhouse transition, where the stellar instellation cannot anymore be balanced through infrared cooling to space~\citep{Ingersoll1969}, and the maximum greenhouse limit, corresponding to the maximum distance at which surface temperatures allowing liquid water can be maintained through a CO$_2$ greenhouse effect~\citep{Kasting1991,Kasting1993,Underwood2003,Kopparapu2013,Kopparapu2014}.
 We use the parametrization in \citet{Kopparapu2014} to derive luminosity and planetary mass-dependent edges of the \gls{HZ} $a_\mathrm{inner}$ and $a_\mathrm{outer}$.
@@
 \subsubsection{Transit survey simulations}
-With the synthetic star and planet samples generated, we simulated observations of these planets with a hypothetical transit survey.
-Using Bioverse's survey module, we simulated noisy measurements of key observables assuming a capable transit survey that can characterize a large sample with high photometric precision.
-We roughly followed the mission parameters of the Nautilus mission concept~\citep{Apai2019,Apai2022} and measured planetary instellation (for \gls{HZ} occupancy) with a precision of \var{nautilus_S} and host star effective temperature with a precision of \var{nautilus_T_eff_st}~K.
-We assumed that the maximum past \gls{NUV} flux a planet received can be determined within a precision of \var{nautilus_max_nuv}.
+With the synthetic star and planet samples generated, we used Bioverse's survey module to simulate noisy measurements of key observables with a transit survey.
+We assumed a hypothetical mission that can target a large planet sample with high photometric precision and conduct a biosignature search on these planets~\citep[e.g.,][]{Apai2019,Apai2022}.
+The simulated survey was designed to measure planetary instellation (for \gls{HZ} occupancy) with a precision of \var{nautilus_S} and host star effective temperature with a precision of \var{nautilus_T_eff_st}~K.
+The maximum past \gls{NUV} flux a planet received can be determined within a precision of \var{nautilus_max_nuv}.
 To marginalize over choices of biosignatures and their detectability, which are beyond the scope of this study, we assumed that any inhabited planet would show a biosignature detectable by the survey.
@@
 \subsection{Semi-analytical assessment}\label{sec:results-semianalytical}
-In Section~\ref{sec:met-semianalytical} we computed the probability for true positive evidence for $H_\mathrm{1}$ and $H_\mathrm{null}$, respectively (Equations~\ref{eq:bayes_factor}, \ref{eq:bayes_factor2}).
+In Section~\ref{sec:met-semianalytical}, we computed the probability for true positive evidence for $H_\mathrm{1}$ and $H_\mathrm{null}$, respectively (Equations~\ref{eq:bayes_factor}, \ref{eq:bayes_factor2}).
 Figure~\ref{fig:semian_true_evidence} shows how these evidences are distributed for sample sizes \var{semian_Nsamp1} and \var{semian_Nsamp2}, and how likely we are to obtain strong evidence ($BF_{H_i, H_j}$ > 10).
@@
 %\end{figure}
-In a magnitude- and volume-limited sample of a transit survey, the host star distribution for \glspl{EEC} will be skewed toward later spectral types and dominated by M~dwarfs due to how the \gls{HZ} scales with spectral type.
-For a fixed planet sample size, the fraction of inhabited planets is highest in the M~dwarf sample due to the higher \gls{NUV} fluxes in the \gls{HZ} of these stars (see Figures~\ref{fig:hz_nuv_evo}, \ref{fig:surveys_FGKM}).
+%In a magnitude- and volume-limited sample of a transit survey, the host star distribution for \glspl{EEC} will be skewed toward later spectral types and dominated by M~dwarfs due to how the \gls{HZ} scales with spectral type.
+With \gls{HZ} occupancy as a requirement for abiogenesis, and barring selection biases beyond stellar brightness, the host star distribution of inhabited planets in a simulated transit survey is skewed toward later spectral types.
+For a fixed planet sample size, the fraction of inhabited planets is highest for planets orbiting M~dwarfs due to the higher \gls{NUV} fluxes in the \gls{HZ} of these stars (see Figures~\ref{fig:hz_nuv_evo}, \ref{fig:surveys_FGKM}).
 Their \gls{NUV} fluxes are generally highest at early times $\lesssim\SI{100}{\mega\year}$.
 These host stars, in particular late subtypes, also provide extended periods of increased \gls{NUV} emission that overlap with times when some of these planets occupy the \gls{HZ} (see Figure~\ref{fig:hz_nuv_evo}), our requirement for abiogenesis (compare Equation~\ref{eq:hypothesis-uv}).
-Thus -- under the UV Threshold Hypothesis, most inhabited transiting planets in a magnitude- and volume-limited sample orbit M~dwarfs.
+Thus -- under the UV Threshold Hypothesis, most inhabited transiting planets in the sample orbit M~dwarfs.
 
-Here, we are interested in the statistical power of a transit survey with a realistic sample selection and size.
+Here, we are interested in the statistical power of a transit survey with a plausible sample selection and size.
 In the following, we fix the sample size to \var{N_nautilus} and consider two different survey strategies targeting FGK and M~dwarfs, respectively.
@@
 For both surveys targeting M dwarfs and those targeting FGK dwarfs, the maximum \gls{NUV} distribution is rather unimodal.
-Applying the approach from Sect.~\ref{sec:results-semianalytical} of fitting a Beta function to the distribution, we find rather low selectivities (see Figure~\ref{fig:surveys_FGKM}).
+Applying the approach from Sect.~\ref{sec:results-semianalytical} of fitting a Beta function to the distribution, we find rather low selectivities (see Figure~\ref{fig:surveys_FGKM}), which is likely a detriment for statistical hypothesis tests.
+
 %\begin{figure*}
@@
 
-\subsubsection{Expected biosignature pattern in a transit survey}
+\subsubsection{Expected biosignature pattern}
 A representative recovery of the injected biosignature pattern is shown in Figure~\ref{fig:surveys_FGKM}.
@@
 
-\subsubsection{Statistical power for a transit survey and sensitivity on astrophysical parameters}\label{sec:results-powergrid}
+\subsubsection{Statistical power and sensitivity on astrophysical parameters}\label{sec:results-powergrid}
 We now investigate the sensitivity of the achieved statistical power of our default transit survey to the a priori unconstrained threshold \gls{NUV} flux $F_\mathrm{NUV, min}$ and the abiogenesis rate $f_\mathrm{life}$.

### 2024-11-16 src/tex/ms.tex
edit discussion, conclusions

@@
 A wide variety of scenarios for the origin of life have been proposed~\citep[e.g.,][]{Baross1985,Brasier2011,Mulkidjanian2012,Fox2013,Deamer2015,Westall2018}.
-While we are still far away from testing any of them, new prospects in the search for conditions favorable to life have opened up by thinking of the origin of life as a planetary phenomenon and identifying global-scale environmental properties that might support pathways to life~\citep{Sasselov2020}.
+While we may still be far away from conclusively testing them, new prospects in the search for conditions favorable to life have opened up by thinking of the origin of life as a planetary phenomenon and identifying global-scale environmental properties that might support pathways to life~\citep{Sasselov2020}.
 In particular, specific planetary conditions are needed to create stockpiles of initial compounds for prebiotic chemistry; and planetary processes are required to trigger the prebiotic synthesis.
-Such planetary conditions can be hypothesized for exoplanets located in the habitable zone of their host star, with persistent liquid water on their surface.
+Such planetary conditions can be hypothesized for exoplanets located in the \gls{HZ} of their host star, with persistent liquid water on their surface.
 For example, the deep-sea hydrothermal (or hydrothermal-sedimentary) context for the origin of life requires a direct contact of an ocean and the planetary mantle/crust~\citep[e.g.,][]{Baross1985}.
@@
 
-The goal of this work is to evaluate the potential of upcoming exoplanet surveys to test the hypothesis that a minimum past NUV flux is required for abiogenesis.
+The goal of this work is to evaluate the potential of potential future exoplanet surveys to test the hypothesis that a minimum past NUV flux is required for abiogenesis.
 We focus on one version of the UV Threshold Hypothesis, the so-called cyanosulfidic scenario, which has been refined to the point where the required threshold flux has been measured to be $F_\mathrm{NUV, min} = \SI{6.8\pm3.6 e10}{\photons\per\centi\meter\squared\per\second\per\nano\meter}$ integrated from \SIrange{200}{280}{\nano\meter} at the surface~\citep{Rimmer2018,Rimmer2021a,Rimmer2023,Ranjan2023a}.
@@
 \label{sec:discussion}
-A key question in the quest to understand the origins of life is which natural processes best explain how living matter spontaneously appears from nonliving matter~\citep{Malaterre2022}.
+A key question in the quest to understand the origins of life is which natural processes best explain how living matter spontaneously appears from nonliving matter~\citep[e.g.,][]{Malaterre2022}.
+Using astronomical methods, this question will likely not be testable for individual planets but only by the presence of patterns in planetary populations.
 The cyanosulfidic scenario~\citep{Patel2015}, in particular its predicted existence of a minimum \gls{NUV} flux required for prebiotic chemistry, seems to offer an opportunity to test an origins of life hypothesis with a statistical transit survey sampling planets with varying \gls{NUV} flux histories.
@@
 
-\subsection{Sampling strategy for testing a predicted minimum \gls{NUV} flux} % corresponds to conclusion point 1
-In Sect.~\ref{sec:results-semianalytical}, we show that testing the hypothesis of a minimum past \gls{NUV} flux required for abiogenesis suffers from `nuisance' parameters that render inference through astronomical observations difficult.
+\subsection{Sampling strategy for testing a \gls{NUV} flux threshold} % corresponds to conclusion point 1
+In Sect.~\ref{sec:results-semianalytical}, we show that testing the UV~Threshold Hypothesis suffers from `nuisance' parameters that render inference through astronomical observations difficult.
 Here, these parameters are the unspecified value of the \gls{NUV} threshold hypothesized to exist under $H_1$, and the unknown probability of detectable life emerging on a habitable planet $f_{\mathrm{life}}$.
-While the history of a planet's received UV flux is difficult to infer~\citep[e.g.,][]{Richey-Yowell2023}, the estimated maximum \gls{NUV} flux that a planet was exposed to may be used as a proxy.
+While the inference of a planet's entire UV flux evolution is difficult~\citep[e.g.,][]{Richey-Yowell2023}, the estimated maximum \gls{NUV} flux that a planet was exposed to may be used as a proxy, at least if one is interested in a minimum threshold flux and makes the assumption that planetary surfaces offer protection against \textit{too high} UV flux.
 Indeed, the distribution of the number of planets with detected biosignature in a particular sample of planets with determined maximum \gls{NUV} values $F_{\mathrm{NUV}}$ depends on both the values of $F_{\mathrm{NUV},min}$ and $f_{\mathrm{life}}$ as shown in equation \ref{eq:semian:likelihoodH1}.
@@
 
-In our semi-analytical analysis (Section~\ref{sec:results-semianalytical}), we project a possible test performed by a future observer equipped with a sample of exoplanets with known past maximum \gls{NUV} exposure for which biosignature detection has been attempted.
+In our semi-analytical analysis (Section~\ref{sec:results-semianalytical}), we project a possible test performed by a future observer equipped with a sample of exoplanets with derived past maximum \gls{NUV} exposure for which biosignature detection has been attempted.
 This is necessarily reductive as this observer will have more knowledge about experimental conditions and will therefore be able to use this information to guide hypothesis testing.
@@
 %The particular finding that our future observer might prefer to prioritize extreme values of this variable in order to maximize statistical power.
-The particular finding that prioritizing extreme values of past \gls{NUV} flux can enhance statistical power clashes with observational constraints, as the composition of the subset of planets that we can observe and for which detection of biosignature can be attempted is not independent from their \gls{NUV} flux history.
-Hence, for our future observer, selectivity and sample size are in conflict.
+The particular finding that prioritizing extreme values of past \gls{NUV} flux can enhance statistical power likely clashes with observational constraints, as the composition of the subset of planets that we can observe and for which detection of biosignature can be attempted is not independent from their \gls{NUV} flux history.
+Hence, for our future observer, selectivity and sample size may be in conflict.
 This trade-off can be quantified in terms of expected evidence yield, which we have done in Section~\ref{sec:results-semianalytical}.
@@
 
-\subsection{Constraining power for the origins of life as a function of biosignature location} % corresponds to conclusion points 2, 5
+%\subsection{Constraining power for the origins of life as a function of biosignature location} % corresponds to conclusion points 2, 5
+\subsection{How planetary context may constrain the UV~Threshold Hypothesis} % corresponds to conclusion points 2, 5
 It comes to no surprise that the success rate of such a test is sensitive to the sample size of the survey and to the occurrence of life on temperate exoplanets.
-As we have shown, the statistical power of this test also depends on the distribution of past \gls{NUV} fluxes in the sample and on the required threshold flux.
+As we have shown, the statistical power of this test also depends on the distribution of past \gls{NUV} fluxes in the sample and on the threshold flux.
 Optimizing the survey to sample a wide range of NUV flux values, particularly at the extremes, can enhance the likelihood of obtaining strong evidence for or against the hypothesis.
 Intermediate values of the threshold \gls{NUV} flux are more likely to yield strong evidence than extreme values, as the dataset under the alternative hypothesis $H_1$ differs more from the null hypothesis in this case while still being sufficiently populated.
-The required threshold flux is, of course, a priori unknown and we cannot influence it.
+The threshold flux is, of course, a priori unknown and we cannot influence it.
 If, however, better theoretical predictions for the required \gls{NUV} flux for abiogenesis become available~\citep{Rimmer2021}, the survey strategy can be further optimized, for instance by targeting planets that are estimated to have received a \gls{NUV} flux slightly below and above this threshold or by applying a bisection algorithm in a sequential survey~\citep{Fields2023}.
@@
 An interesting aspect lies in the distribution of host star properties, as different spectral types probe different past \gls{NUV} flux regimes.
-FGK stars show a narrow distribution of maximum past \gls{NUV} fluxes in the \gls{HZ}, which may limit the constraining power of a survey depending on the (unknown) threshold \gls{NUV} flux.
+FGK stars show a narrow distribution of maximum past \gls{NUV} fluxes in the \gls{HZ}, which may limit the diagnostic power of a survey depending on the (unknown) threshold \gls{NUV} flux.
 With a pure FGK sample, a survey will be more sensitive to the low \gls{NUV} flux end of the hypothesis.
@@
 While old M~dwarfs can be considered low-UV environments, a large fraction of them emit high \gls{NUV} fluxes into their \gls{HZ} during their early stages, in particular later subtypes~\citep{Richey-Yowell2023}.
-an M~dwarf sample will thus help to test the high \gls{NUV} flux end of the UV Threshold Hypothesis; a higher occurrence of biosignatures here would support the hypothesis that a higher NUV flux is beneficial or necessary for life.
+An M~dwarf sample will thus help to test the high \gls{NUV} flux end of the UV Threshold Hypothesis; a higher occurrence of biosignatures here would support the hypothesis that a higher NUV flux is beneficial or necessary for life.
 As our simulations show, the higher and more variable \gls{NUV} fluxes in M~dwarfs increase the likelihood of obtaining strong evidence for or against the hypothesis.
@@
 Our Bioverse simulations that take into account exoplanet demographics, the evolution of habitability and \gls{NUV} fluxes, and observational biases show that not only the likelihood of a conclusive test of the UV~Threshold Hypothesis, but also the likelihood of successful biosignature detection itself is extremely sensitive to the threshold \gls{NUV} flux if the hypothesis is true.
-Even if all biosignatures can be detected and the nominal abiogenesis rate is very high, say $f_\mathrm{life} = 1$, under the condition that prebiotic chemistry requires a minimum \gls{NUV} flux \textit{and} liquid water, the probability of finding life on a randomly selected planet may be very low.
+Even if all biosignatures can be detected and the nominal abiogenesis rate is very high, say $f_\mathrm{life} = 1$, under the condition that prebiotic chemistry requires a minimum \gls{NUV} flux \textit{and} liquid water, if the threshold flux turns out to be high the probability of finding life on a randomly selected planet may be very low.
 As we showed, for high required fluxes the two requirements of simultaneous \gls{HZ} occupancy and sufficient \gls{NUV} flux conspire to diminish the fraction of inhabited planets in the sample.
@@
 
-Unfortunately, our simulations do not yield a clear criterion for the credibility of a biosignature detection based on spectral type of the host star, as both FGK and M~dwarf samples show similar maximum past \gls{NUV} flux distributions.
+Our simulations find no clear criterion for the credibility of a biosignature detection based on spectral type of the host star, as both FGK and M~dwarf samples show similar maximum past \gls{NUV} flux distributions.
 The abiogenesis rate in both samples show similar trends with the threshold \gls{NUV} flux, and the fraction of inhabited planets drops at similar threshold fluxes (see Section~\ref{sec:results-bioverse}).
@@
 \subsection{Overall prospects for testing the UV~Threshold Hypothesis} % corresponds to conclusion point 4
-Our results show that the UV~Threshold Hypothesis is testable with upcoming exoplanet surveys, but that the success of such a test depends on the sample size, the distribution of past \gls{NUV} fluxes, and several unknown astrophysical nuisance parameters.
+Our results show that the UV~Threshold Hypothesis is testable with potential future exoplanet surveys, but that the success of such a test depends on the sample size, the distribution of past \gls{NUV} fluxes, and several unknown astrophysical nuisance parameters.
 Even under idealized conditions, obtaining strong evidence for or against the hypothesis likely requires sample sizes on the order of 100 (see Section~\ref{sec:results-semianalytical}).
@@
 Meaningful tests of the UV~Threshold Hypothesis may thus require a different survey strategy that increases the sample of characterized planets.
-Such concepts, like the Nautilus mission~\citep{Apai2019,Apai2022}, which aims to characterize up to $\sim 1000$ rocky planets via utilizing an innovative telescope design, are only appearing on the horizon.
-Constraints on the origins of life from exoplanet surveys may thus be a long-term endeavor, and the UV~Threshold Hypothesis may not be testable yet with the current and next generation of exoplanet surveys.
-In the meantime, it is important to refine predictions on the role of UV radiation in prebiotic chemistry with both theoretical and experimental work.
+The Nautilus mission~\citep{Apai2019,Apai2022} aims to characterize up to $\sim 1000$ \gls{EEC} via utilizing an innovative telescope design.
+\todo[inline]{DANIEL wants to work on this paragraph}
+%Constraints on the origins of life from exoplanet surveys may thus be a long-term endeavor, and the UV~Threshold Hypothesis may not be testable yet with the current and next generation of exoplanet surveys.
+To prepare for future biosignature surveys, it is important to refine predictions on the role of UV radiation in prebiotic chemistry with both theoretical and experimental work.
 
@@
 By its nature, cyanosulfidic scenario relies on rock surfaces exposed to the planetary atmosphere.
-Water worlds that have their entire planeary surface covered with water contradict this requireent and do not allow for the wet-dry cycling inherent to this origin of life scenario.
+Water worlds that have their entire planeary surface covered by oceans contradict this requireent and do not allow for the wet-dry cycling inherent to this origin of life scenario.
 The competition of tectonic stress with gravitational crustal spreading~\citep{Melosh2011} sets the maximum possible height of mountains, which in the solar system does not exceed $\SI{\sim 20}{\kilo\meter}$.
 Such mountains will be permanently underwater on water worlds.
-Another impediment to wet-dry cycles may be tidal locking of the planet as it stalls stellar tide-induced water movement and diurnal irradiation variability.
+Another impediment to wet-dry cycles may be tidal locking of the planet as it stalls stellar tide-induced water movement and diurnal irradiation variability. CITE!
 
@@
 \label{sec:conclusions}
+We propose that specific origins-of-life scenarios may leave a detectable imprint on the distribution of biosignatures in exoplanet populations.
 We have investigated the potential of upcoming exoplanet surveys to test the hypothesis -- motivated by the cyanosulfidic origins-of-life scenario -- that a minimum past \gls{NUV} flux is required for abiogenesis.
@@
     \item The required sample size for detecting this correlation depends on the abiogenesis rate on temperate exoplanets and the distribution of host star properties in the sample; in particular their maximum past \gls{NUV} fluxes.
-%        A direct imaging mission with the \gls{HWO} will require characterizing XXX planets ...
+    Samples smaller than 50 planets are unlikely to yield conclusive results.
     \item Under the UV~Threshold Hypothesis, the fraction of inhabited planets in a transit survey is sensitive to the threshold \gls{NUV} flux and is expected to drop sharply for required fluxes above a few hundred~\SI{}{\erg\per\second\per\centi\meter\squared}.
@@
 More generally, we found that models of the origins of life provide hypotheses that may be testable with these surveys.
-Conducting realistic survey simulations with representative samples is key to making accurate predictions, developing the optimal survey strategy, and increasing science return.
-Our work highlights the importance of understanding the context in which a biosignature detection is made, which can not only help to assess the credibility of the detection but also to test competing theories of the origins of life on Earth and beyond.
+%Conducting realistic survey simulations with representative samples is key to making accurate predictions, developing the optimal survey strategy, and increasing science return.
+Conducting realistic survey simulations with representative samples is important to identify testable science questions, support trade studies, help define science cases for future missions, and guide further theoretical and experimental work on the origins of life.
+Our work highlights the importance of understanding the context in which a biosignature detection is made, which can not only help to assess the credibility of the detection but also to test competing hypotheses on the origins of life on Earth and beyond.
 

### 2024-11-18 src/tex/ms.tex
draft an abstract

@@
 \begin{abstract}
-    ...
+The origins of life on Earth may provide testable hypotheses about biosignature distributions in exoplanet populations.
+Motivated by the cyanosulfidic origins-of-life scenario, we investigate the hypothesis that a minimum past near-ultraviolet (NUV) flux is necessary for abiogenesis.
+Using a semi-analytical Bayesian framework und the Bioverse survey simulator, we estimate the probability of obtaining strong evidence for or against this UV Threshold Hypothesis with potential future biosignature surveys.
+Our results indicate that a correlation between past NUV flux and current biosignature occurrence is observationally testable, but requires sample sizes of at least 50 planets for meaningful constraints.
+These constraints are critically sensitive to the intrinsic abiogenesis rate, as well as host star properties, particularly maximum past NUV fluxes.
+Surveys targeting extreme NUV flux values and planets orbiting M dwarfs enhance the likelihood of obtaining conclusive results, with sample sizes $\gtrsim$100 providing $\gtrsim$80\% likelihood of strong evidence if abiogenesis rates are high and required NUV fluxes are moderate.
+Conversely, the fraction of inhabited planets sharply decreases for required fluxes exceeding a few hundred erg/s/cm².
+
+Our findings demonstrate the potential of upcoming exoplanet surveys to test origins-of-life hypotheses.
+ Beyond specific scenarios, our work underscores the broader value of realistic survey simulations in identifying testable science questions, optimizing mission strategies, and advancing theoretical and experimental studies of abiogenesis.
+ This approach emphasizes the need to contextualize biosignature detections, enhancing their credibility and offering new approaches to explore the emergence of life on Earth and beyond.
 \end{abstract}

### 2024-11-18 src/tex/acknowledgments_contributions.tex
draft acknowledgments and author contributions

\begin{acknowledgments}
    The authors thank Kevin Heng, Dominik Hintz, and Rhys Seeburger for insightful discussions.
%    We are grateful to ...
%    mention any people who gave comments in early-on meetings
%    We thank the anonymous referees for providing insightful reports that helped to improve this manuscript.
    This material is based upon work supported by the National Aeronautics and Space Administration under Agreement No. 80NSSC21K0593 for the program ``Alien Earths''.
    The results reported herein benefited from collaborations and/or information exchange within NASA’s Nexus for Exoplanet System Science (NExSS) research coordination network sponsored by NASA’s Science Mission Directorate.
    This work has made use of data from the European Space Agency (ESA) mission \gaia\ (\url{https://www.cosmos.esa.int/gaia}), processed by the \gaia\ Data Processing and Analysis Consortium (DPAC, \url{https://www.cosmos.esa.int/web/gaia/dpac/consortium}). Funding for the DPAC has been provided by national institutions, in particular the institutions participating in the \gaia\ Multilateral Agreement.
%    ...
\end{acknowledgments}

\section*{Author contributions}
M.S., D.A., and S.R.\ conceived the project, planned its implementation, and interpreted the results.
M.S.\ developed the planetary evolution component to Bioverse, carried out the hypothesis tests and statistical analyses, and wrote the manuscript.
D.A.\ leads the ``Alien Earths'' program through which this project is funded, helped to guide the strategy of the project, and provided text contributions.
A.A.\ carried out the semi-analytical computations regarding the correlation of past UV flux and biosignature occurrence.
S.R.\ advised on planetary \gls{NUV} flux evolution and the cyanosulfidic scenario of the origins of life.
R.F.\ wrote the initial draft of the Introduction and advised on the evolutionary biology aspects of the project.
K.H.-U.\ contributed to the Bioverse software development and simulations.
%M.S.\ wrote the manuscript; ... and ... provided text contributions.
T.L.\ supported the selection of testable hypotheses and provided text contributions to the initial draft.
S.M.\ advised on the scope of the project and supported the selection of testable hypotheses.
All authors provided comments and suggestions on the manuscript.

%Author affiliations (tbd):
%Martin Schlecker1, Daniel Apai1,2, Antonin Affholder4, Sukrit Ranjan2, Regis Ferriere4, Kevin Hardegree-Ullman1, Tim Lichtenberg3, and Stephane Mazevet5
%1 Steward Observatory and Department of Astronomy, The University of Arizona, Tucson, AZ 85721, USA
%2 Lunar and Planetary Laboratory, The University of Arizona, Tucson, AZ 85721, USA
%3 Kapteyn Astronomical Institute, University of Groningen, PO Box 800, 9700 AV Groningen, The Netherlands
%4 Department of Ecology and Evolutionary Biology, University of Arizona, Tucson, AZ, USA
%5 Observatoire de la Côte d’Azur, 96 Boulevard de l’Observatoire, F-06300 Nice, France

### 2024-11-20 src/tex/ms.tex
draft acknowledgments and author contributions
small edits

@@
 To evaluate the evidence for correctly rejecting the null hypothesis, we employed the Mann-Whitney U test~\citep{Mann1947}.
-This is a non-parametric test to determine whether two independent samples were drawn from a population with the same distribution and is in particular sensitive to one sample being stochastically greater than the other.
+%This is a non-parametric test to determine whether two independent samples were drawn from a population with the same distribution and is in particular sensitive to one sample being stochastically greater than the other.
+This is a non-parametric test of the null hypothesis that two independent samples were drawn from a population with the same distribution and is in particular sensitive to one sample being stochastically greater than the other.
 We used the Mann-Whitney U test to compare the distributions of \gls{NUV} fluxes of planets with and without biosignatures.
-The implementation in Bioverse relies on the \texttt{scipy.stats.mannwhitneyu} function~\citep{Virtanen2020} and returns a two-tailed p-value, for which we set a significance level of $\alpha = 0.05$ to reject the null hypothesis.
+The implementation in Bioverse relies on the \texttt{scipy.stats.mannwhitneyu} function~\citep{Virtanen2020} and returns a p-value, for which we set a significance level of $\alpha = 0.05$ to reject the null hypothesis.
 For every hypothesis test, we repeated randomized survey realizations to estimate the fraction of successful rejections of the null hypothesis, i.e., the statistical power of the survey.
@@
 These host stars, in particular late subtypes, also provide extended periods of increased \gls{NUV} emission that overlap with times when some of these planets occupy the \gls{HZ} (see Figure~\ref{fig:hz_nuv_evo}), our requirement for abiogenesis (compare Equation~\ref{eq:hypothesis-uv}).
-Thus -- under the UV Threshold Hypothesis, most inhabited transiting planets in the sample orbit M~dwarfs.
+Thus -- under the UV Threshold Hypothesis -- most inhabited transiting planets in the sample orbit M~dwarfs.
 
@@
 Green areas show inhabited planets for an abiogenesis rate of $f_\mathrm{life} = \var{f_life}$ and a generic threshold \gls{NUV} flux $F_\mathrm{NUV, min} = \var{NUV_thresh}\,\SI{}{\erg\per\second\per\centi\meter\squared}$.\\
-        Bottom: Recovered biosignature detections ($\checkmark$) and non-detections ($\oslash$) of simulated transit surveys. The dashed line denotes $F_\mathrm{NUV, min}$.}
+        Bottom: Recovered biosignature detections and non-detections of simulated transit surveys. The dashed line denotes $F_\mathrm{NUV, min}$.}
         \label{fig:surveys_FGKM}
@@
         \includegraphics[width=\hsize]{figures/inhabited_aafo_nuv_thresh}
-        \caption{Fraction of inhabited planets for different threshold \gls{NUV} fluxes under the UV~Threshold Hypothesis if the abiogenesis rate is \var{f_life}.
+        \caption{Fraction of inhabited planets for different threshold \gls{NUV} fluxes under the UV~Threshold Hypothesis if the abiogenesis rate is 1.
 Shaded regions denote 90 percent confidence intervals of randomized sample generations, and dashed lines correspond to samples including non-transiting planets.
@@
 \end{figure}
-Figure \ref{fig:inhabited_aafo_nuv_thresh} shows the fraction of inhabited planets under the UV~Threshold Hypothesis for different threshold \gls{NUV} fluxes and for a high abiogenesis rate of \var{f_life}.
+Figure \ref{fig:inhabited_aafo_nuv_thresh} shows the fraction of inhabited planets under the UV~Threshold Hypothesis for different threshold \gls{NUV} fluxes and for a high abiogenesis rate of $f_\mathrm{life} = 1$.
 This fraction decreases sharply with increasing threshold flux, as fewer planets receive sufficient \gls{NUV} flux for abiogenesis.
@@
 Here, the FGK sample is less sensitive.
-The only parameter space region with statistical power above \SI{90}{\percent} lies at abiogenesis rates $f_\mathrm{life} > 0.9$ and threshold \gls{NUV} fluxes of $\sim \SIrange{30}{70}{\erg\per\second\per\centi\meter\squared}$.
 \todo[inline]{Update this with final results from stat. power grid}
+The only parameter space region with statistical power above \SI{90}{\percent} lies at abiogenesis rates $f_\mathrm{life} > 0.9$ and threshold \gls{NUV} fluxes of $\sim \SIrange{200}{400}{\erg\per\second\per\centi\meter\squared}$.
 
@@
 Indeed, the distribution of the number of planets with detected biosignature in a particular sample of planets with determined maximum \gls{NUV} values $F_{\mathrm{NUV}}$ depends on both the values of $F_{\mathrm{NUV},min}$ and $f_{\mathrm{life}}$ as shown in equation \ref{eq:semian:likelihoodH1}.
-Avoiding a loss in statistical power due to nuisance parameters is a common difficulty in statistical testing, and is the motivation behind elaborate test statistics.
+%Avoiding a loss in statistical power due to nuisance parameters is a common difficulty in statistical testing, and is the motivation behind elaborate test statistics.
 
@@
           They may also be more likely to yield biosignature detections under this hypothesis.
+          \todo[inline]{check if new grid still supports this}
 
@@
 
-%\input{acknowledgments_contributions}
+\input{acknowledgments_contributions}
 

### 2024-11-20 src/tex/ms.tex
fix references

@@
 %Test for negative correlations as well?}
-Here, we conduct a theoretical experiment on the UV~Threshold Hypothesis (Sect.~\ref{sec:predictions}) by relating the occurrence of life on an exo-earth candidate with a minimum past quiescent stellar UV flux, focusing on the prebiotically interesting \gls{NUV} range from \SIrange{200}{280}{\nano\meter}.
+Here, we conduct a theoretical experiment on the UV~Threshold Hypothesis by relating the occurrence of life on an exo-earth candidate with a minimum past quiescent stellar UV flux, focusing on the prebiotically interesting \gls{NUV} range from \SIrange{200}{280}{\nano\meter}.
 \todo[inline]{SUKRIT: got a good reference for this?}  %~\citep{RanjanSasselov2016}.?
@@
 we thus consider only rocky planets that may sustain liquid water on their surface, i.e., that occupy their host star's momentary \gls{HZ} during the above period, as well as at the epoch of observation.
-\todo[inline]{Does this need more detailed explanation?}
 To determine \gls{HZ} occupancy, we take into account the evolution of the host star's luminosity and \gls{HZ} boundaries.
@@
 Such mountains will be permanently underwater on water worlds.
-Another impediment to wet-dry cycles may be tidal locking of the planet as it stalls stellar tide-induced water movement and diurnal irradiation variability. CITE!
+Another impediment to wet-dry cycles may be tidal locking of the planet as it stalls stellar tide-induced water movement and diurnal irradiation variability~\citep[e.g.,][]{Ranjan2017b}.
 

### 2024-11-20 src/scripts/surveys_FGKM.py
read actual sample size

@@ -78,7 +78,7 @@ def plot_nuv_distribution(sample, data, fig, ax, spt, plot_fit=False):
 
     # estimate selectivity from average of fitted beta function parameters
     selectivity = np.log10(1 / (np.mean(max_likeli_norm[:2])))
-    save_var_latex("selectivity_{}".format(spt), round(selectivity, 1))
+    save_var_latex("selectivity_{}".format(spt), round(selectivity, 2))
     print(f"selectivity ~ {selectivity:.2f}")
 
     # plot histogram and beta distribution fitted on non-normalized data
@@ -209,7 +209,8 @@ for spt, axlr in zip(["FGK", "M"], [axs_left, axs_right]):
         # axlr[0].legend(loc="center left", title=None)
         leg = axlr[0].legend(
             bbox_to_anchor=(1.07, 0.99),
-            title=f"N = {len(sample)}",
+            # title=f"N = {len(sample)}",
+            title="N = {}".format(read_var_latex('N_nautilus')),
             loc="lower right",
             frameon=False,
         )

### 2024-11-20 src/scripts/bioverse_pipeline.py
no need to specify biosignature

@@ -62,8 +62,8 @@ def inject_nuv_life_correlation(d, f_life=0.1):
 
 
 def inject_biosignature(d, P=1.0):
-    # Randomly assign O2 to some inhabited planets
-    d["has_O2"] = np.logical_and(d["inhabited"], np.random.random(size=len(d)) <= P)
+    # Randomly assign a biosignature to some inhabited planets
+    d["biosig"] = np.logical_and(d["inhabited"], np.random.random(size=len(d)) <= P)
 
     return d
 
@@ -150,7 +150,7 @@ def create_survey_nautilus():
         "M",  # needed for estimates of HZ occupancy
         "S",
         "EEC",
-        "has_O2",
+        "biosig",
         "hz_and_uv",
         "max_nuv",
     ]
@@ -245,7 +245,7 @@ def hypothesis_test(data, testmethod):
     bounds = np.array([[1e-3, 1.0], [10.0, 1e5]])
     bounds_null = np.array([[1e-3, 1.0]])
     features = ("max_nuv",)
-    labels = ("has_O2",)
+    labels = ("biosig",)
 
     h_nuv = Hypothesis(
         h1, bounds, params=params, features=features, labels=labels, log=log
@@ -253,7 +253,7 @@ def hypothesis_test(data, testmethod):
     h_nuv.h_null = Hypothesis(
         h_null,
         bounds_null,
-        params=("f_O2",),
+        params=("f_biosig",),
         features=features,
         labels=labels,
         log=(True,),
@@ -288,7 +288,7 @@ def hypotest_grid(generator, survey, N_grid, fast, testmethod):
     bounds = np.array([[1e-3, 1.0], [10.0, 1e5]])
     bounds_null = np.array([[1e-3, 1.0]])
     features = ("max_nuv",)
-    labels = ("has_O2",)
+    labels = ("biosig",)
 
     h_nuv = Hypothesis(
         h1, bounds, params=params, features=features, labels=labels, log=log
@@ -296,7 +296,7 @@ def hypotest_grid(generator, survey, N_grid, fast, testmethod):
     h_nuv.h_null = Hypothesis(
         h_null,
         bounds_null,
-        params=("f_O2",),
+        params=("f_biosig",),
         features=features,
         labels=labels,
         log=(True,),

### 2024-11-20 src/scripts/bioverse_pipeline.py
prepare pipeline for grid run

@@ -91,7 +91,7 @@ def get_generator_args():
         # 'mr_relation' : 'Wolfgang2016',
         # 'mr_relation' : 'earthlike',
         # past HZ occupancy and NUV flux
-        "deltaT_min": 10.0,  # Myr
+        "deltaT_min": 1.0,  # Myr
         # "NUV_thresh": 100.0,  # erg/s/cm2
     }
     return stars_args, planets_args
@@ -116,9 +116,9 @@ def generate_generator(stars_only=False, **kwargs):
         g.insert_step("classify_planets")
         g.insert_step("scale_height")
         g.insert_step("compute_transit_params")
+        g.insert_step(eec_filter)
         g.insert_step("past_hz_uv")
         # g.insert_step('apply_bias')
-        g.insert_step(eec_filter)
         g.insert_step(inject_nuv_life_correlation)
         g.insert_step(inject_biosignature)
         g.insert_step(limit_sample)
@@ -166,7 +166,8 @@ def create_survey_nautilus():
         "P": 0.000001,
         "S": "5%",
         # "max_nuv": "20%",
-        "max_nuv": "5%",
+        "max_nuv": "5%",  # will it be measured if we remove it from the `precision` dict? No.
+        # "max_nuv": "50.",
     }
 
     # margs['t_ref'] = {'R':1.5}
@@ -303,19 +304,21 @@ def hypotest_grid(generator, survey, N_grid, fast, testmethod):
     )
 
     if fast:
-        N_iter = 4
+        N_iter = 3
+        f_life = np.geomspace(5e-3, 0.999, N_grid)
+        NUV_thresh = np.geomspace(80.0, 500.0, N_grid)
     else:
         N_iter = 16
-    # f_life = np.geomspace(0.1, 1.0, N_grid)
-    # f_life = np.geomspace(1e-3, 1.0, N_grid)
-    f_life = np.geomspace(1e-2, 1.0, N_grid)
-    # f_life = np.geomspace(0.5, 1.0, N_grid)
-    # f_life = (0.9,)  # test 1D hypothesis grid test
-    # f_life = 0.99  # test 1D hypothesis grid test
-    # NUV_thresh = np.geomspace(300.0, 380.0, N_grid)
-    # NUV_thresh = np.geomspace(30.0, 3000.0, N_grid)
-    # NUV_thresh = np.geomspace(200.0, 600.0, N_grid)
[…]

### 2024-11-20 src/tex/ms.tex
update discussion on FGK vs. M dwarfs

@@
 An interesting aspect lies in the distribution of host star properties, as different spectral types probe different past \gls{NUV} flux regimes.
-FGK stars show a narrow distribution of maximum past \gls{NUV} fluxes in the \gls{HZ}, which may limit the diagnostic power of a survey depending on the (unknown) threshold \gls{NUV} flux.
-With a pure FGK sample, a survey will be more sensitive to the low \gls{NUV} flux end of the hypothesis.
-A lack of biosignatures on these planets would support the UV~Threshold Hypothesis.
-Their presence would have little constraining power; it would either suggest that low NUV fluxes are sufficient for abiogenesis or indicate a different abiogenic pathway~\citep[e.g.,][]{Westall2018}.
+FGK stars show a narrow distribution of maximum past \gls{NUV} fluxes in the \gls{HZ}, which may -- depending on the (unknown) threshold \gls{NUV} flux -- limit the diagnostic power of a survey.
+In the case of a pure FGK sample, it will essentially only be sensitive to \gls{NUV} flux thresholds $\sim \SIrange{200}{400}{\erg\per\second\per\centi\meter\squared}$, and the chance of detecting biosignatures diminishes rapidly for higher thresholds within this range (see Figure~\ref{fig:inhabited_aafo_nuv_thresh}).
+Detected biosignatures in an FGK sample would have little constraining power on testing the UV~Threshold Hypothesis; they would either suggest that low NUV fluxes are sufficient for abiogenesis or indicate a different abiogenic pathway~\citep[e.g.,][]{Westall2018}.
+A lack of biosignatures in a larger FGK sample would support the UV~Threshold Hypothesis.
 
 On the other hand, M~dwarfs show a wider distribution of maximum past \gls{NUV} fluxes in their \glspl{HZ}.
-While old M~dwarfs can be considered low-UV environments, a large fraction of them emit high \gls{NUV} fluxes into their \gls{HZ} during their early stages, in particular later subtypes~\citep{Richey-Yowell2023}.
-An M~dwarf sample will thus help to test the high \gls{NUV} flux end of the UV Threshold Hypothesis; a higher occurrence of biosignatures here would support the hypothesis that a higher NUV flux is beneficial or necessary for life.
-As our simulations show, the higher and more variable \gls{NUV} fluxes in M~dwarfs increase the likelihood of obtaining strong evidence for or against the hypothesis.
+While old M~dwarfs can be considered low-UV environments, a significant fraction of them emit high \gls{NUV} fluxes into their \gls{HZ} during their early stages, in particular later subtypes~\citep{Richey-Yowell2023}.
+This will help to test the high \gls{NUV} flux end of the UV Threshold Hypothesis; a higher occurrence of biosignatures here would support the hypothesis that a higher NUV flux is beneficial or necessary for life.
+At the same time, a fraction of host stars in our M~dwarf sample extends it to lower maximum past \gls{NUV} fluxes, enabling tests of the low \gls{NUV} flux end of the hypothesis.
+The higher and more variable \gls{NUV} fluxes in M~dwarfs thus increase the likelihood of obtaining strong evidence for or against the UV~Threshold Hypothesis.
 The combination of a lack of UV radiation today, which makes biosignature gases more detectable~\citep{Segura2005}, and a UV-rich past that may have enabled abiogenesis could make M~dwarfs the preferred targets for biosignature searches.

### 2024-11-20 src/tex/ms.tex
update text based on power grid

@@
 
+Parameter space regions with statistical power above \SI{90}{\percent} lie at abiogenesis rates $f_\mathrm{life} > 0.5$ and mostly at threshold \gls{NUV} fluxes of $\sim \SIrange{200}{400}{\erg\per\second\per\centi\meter\squared}$.
 Notably, the sensitivity of the M~dwarf sample extends into the low \gls{NUV} flux end due to the broader distribution of maximum past \gls{NUV} fluxes in this sample.
-Here, the FGK sample is less sensitive.
-\todo[inline]{Update this with final results from stat. power grid}
-The only parameter space region with statistical power above \SI{90}{\percent} lies at abiogenesis rates $f_\mathrm{life} > 0.9$ and threshold \gls{NUV} fluxes of $\sim \SIrange{200}{400}{\erg\per\second\per\centi\meter\squared}$.
+Here, the FGK sample is barely sensitive.
 
@@
         \includegraphics[width=\hsize]{figures/powergrid}
-%        \includegraphics[width=\hsize]{figures/powergrid.bak.png}
         \caption{Statistical power as a function of threshold \gls{NUV} flux and abiogenesis rate.
 Even for a large sample (here: $N=\var{N_nautilus}$), a high statistical power of the transit survey requires high abiogenesis rates $f_\mathrm{life}$.
-         Intermediate values of $F_\mathrm{NUV, min}$ are more likely to yield strong evidence than extreme values; the sensitivity of the M~dwarf sample extends into the low \gls{NUV} flux end.
+         Intermediate values of $F_\mathrm{NUV, min}$ are more likely to yield strong evidence than extreme values.
+For abiogenesis rates $f_\mathrm{life} \gtrsim \SI{80}{\percent}$, the sensitivity of the M~dwarf sample extends into the low \gls{NUV} flux end.
          }
@@
 This is necessarily reductive as this observer will have more knowledge about experimental conditions and will therefore be able to use this information to guide hypothesis testing.
-For instance, we have made the choice to consider the total number of detected biosignatures as our summary statistic (Equation~\ref{eq:semian:likelihoodH1}), which is not `sufficient' to infer $F_{\mathrm{NUV},min}$ and $f_{life}$ separately.
+For instance, we have made the choice to consider the total number of detected biosignatures as our summary statistic (Equation~\ref{eq:semian:likelihoodH1}), which is not sufficient to infer $F_{\mathrm{NUV},min}$ and $f_{life}$ separately.
 However, by conditioning the Bayes factor to these variables (Equation \ref{eq:bayes_factor}), we calculate the probability distribution of the Bayesian evidence in favor of $H_1$.
@@
 \subsection{How planetary context may constrain the UV~Threshold Hypothesis} % corresponds to conclusion points 2, 5
-It comes to no surprise that the success rate of such a test is sensitive to the sample size of the survey and to the occurrence of life on temperate exoplanets.
+It comes to no surprise that the success rate for testing the UV~Threshold Hypothesis is sensitive to the sample size of the survey and to the occurrence of life on temperate exoplanets.
 As we have shown, the statistical power of this test also depends on the distribution of past \gls{NUV} fluxes in the sample and on the threshold flux.
@@
 The combination of a lack of UV radiation today, which makes biosignature gases more detectable~\citep{Segura2005}, and a UV-rich past that may have enabled abiogenesis could make M~dwarfs the preferred targets for biosignature searches.
-
 These findings underscore the importance of constraining the UV emission profiles of \gls{EEC} host stars throughout their evolutionary stages to assess the viability of M-dwarf planets as testbeds for theories on the origins of life~\citep{Rimmer2021,Ranjan2023a}.
@@
           They may also be more likely to yield biosignature detections under this hypothesis.
-          \todo[inline]{check if new grid still supports this}
 

### 2024-11-21 src/scripts/hz_nuv_evo.py
frame HZ and NUV overlaps

@@ -9,6 +9,7 @@ plotstyle.styleplots()
 import numpy as np
 import matplotlib.pyplot as plt
 import matplotlib.colors as colors
+from matplotlib.patches import Rectangle
 
 from bioverse.util import interpolate_luminosity, interpolate_nuv
 
@@ -18,27 +19,32 @@ NUV_thresh = 300.0
 
 def get_example_planets(d):
     try:
-        with open(
-                paths.data / "pipeline/exampleplanets.dll", "rb"
-        ) as file:
+        with open(paths.data / "pipeline/exampleplanets.dll", "rb") as file:
+            print("Loading example planets from file.")
             sample = dill.load(file)
 
     except FileNotFoundError:
         dd = d.to_pandas()
-        ids = dd[dd.planetID.isin([40278, 578, 71628
-            # 3476,   8688,  15613,  40030,  54874,  56161,  71628,
-            # 79440, 106402, 121746, 149285, 170230, 174441, 180296,
-            # 185852, 189631, 208822, 221669, 224253
-           ])].planetID
-        sample = d[np.isin(d['planetID'], ids)]
+        ids = dd[
+            dd.planetID.isin(
+                [
+                    40278,
+                    578,
+                    71628
+                    # 3476,   8688,  15613,  40030,  54874,  56161,  71628,
+                    # 79440, 106402, 121746, 149285, 170230, 174441, 180296,
+                    # 185852, 189631, 208822, 221669, 224253
+                ]
+            )
+        ].planetID
+        sample = d[np.isin(d["planetID"], ids)]
         sample.evolution = {k: v for k, v in d.evolution.items() if k in ids.values}
 
-        with open(
-                paths.data / "pipeline/exampleplanets.dll", "wb"
-        ) as file:
+        with open(paths.data / "pipeline/exampleplanets.dll", "wb") as file:
             dill.dump(sample, file)
 
     return sample
+
 
 def plot_planets_scatter(ax, d, norm, col=None):
     # for (v, c) in [(False, 'white'), (True, 'C1')]:
@@ -117,7 +123,7 @@ def plot_nuv_evo(fig, ax, log=True, d=None):
 
 def plot_hz_and_nuv(fig, ax, sample, NUV_thresh=300.0, N_sample=2, random_state=44):
[…]

### 2024-11-21 src/tex/ms.tex
update lum/nuv plot and caption

@@
         \caption{Interpolated stellar luminosity evolution (left) and evolution of the \gls{NUV} flux in the \gls{HZ} (right) as a function of host star mass.
-        The scatter plots show age and host star mass of the transiting planets in the synthetic FGK sample; crosses denote the estimated \gls{NUV} values in \citet{Richey-Yowell2019}.
+        Scatter points show age and host star mass of the transiting planets in the synthetic planet sample; crosses denote the estimated \gls{NUV} values in \citet{Richey-Yowell2019}.
 % with \gls{EEC} highlighted in green.
-        A few example tracks for an example threshold flux of $F_\mathrm{NUV, min} = \var{NUV_thresh}\,\SI{}{\erg\per\second\per\centi\meter\squared}$ are shown; extended overlap of \gls{HZ} occupancy and high \gls{NUV} flux (green sections) fulfills our requirement for abiogenesis.
-        Planet~1 is an \gls{EEC} that never receives sufficient \gls{NUV} flux for abiogenesis.
+        We show three evolutionary tracks for a threshold flux of $F_\mathrm{NUV, min} = \var{NUV_thresh}\,\SI{}{\erg\per\second\per\centi\meter\squared}$ that occupy the \gls{HZ} (green sections) and exceed the threshold \gls{NUV} flux (blue sections) at different times.
+        Where these sections overlap (orange rectangles), the requirements for abiogenesis are met and we assign a biosignature detection with probability $f_\mathrm{life}$.
+        Planet~1 is an \gls{EEC} orbiting a K~dwarf that never receives sufficient \gls{NUV} flux for abiogenesis.
         Planet~2 and Planet~3 enter the \gls{HZ} at different times and receive sufficient \gls{NUV} flux for different durations until their respective host star evolves below the threshold.

### 2024-11-21 src/scripts/hz_nuv_evo.py

@@ -30,7 +30,7 @@ def get_example_planets(d):
                 [
                     40278,
                     578,
-                    71628
+                    71628,
                     # 3476,   8688,  15613,  40030,  54874,  56161,  71628,
                     # 79440, 106402, 121746, 149285, 170230, 174441, 180296,
                     # 185852, 189631, 208822, 221669, 224253
@@ -180,6 +180,12 @@ def plot_hz_and_nuv(fig, ax, sample, NUV_thresh=300.0, N_sample=2, random_state=
     ax.plot(
         np.concatenate((np.array([1e-3]), t)), [Mst] + [Mst for tt in t], lw=1, c="gray"
     )
+
+    t = sample.evolution[id]["time"]
+    in_hz = sample.evolution[id]["in_hz"]
+    t_hz = t[in_hz]
+    ax.plot(t_hz, [Mst + offset for tt in t_hz], lw=3, c="C0")
+
     exampleplanets[Mst.values[0]] = t.tolist()[-1]
     print(f"Planet {i+2} has SpT {eecdf[eecdf.planetID == id].SpT.values[0]}")
 
@@ -209,7 +215,7 @@ if __name__ == "__main__":
         d = dill.load(f)
 
     m = np.linspace(0.1, 1.0, num=200)
-    t = np.geomspace(5e-3, 10.0, num=200)
+    t = np.geomspace(1.4e-2, 10.0, num=200)
     M, T = np.meshgrid(m, t)
     interp_lum, extrap_nn = interpolate_luminosity()
 

### 2024-11-21 src/scripts/inhabited-aafo-nuv_thresh.py
re-run inhabited-aafo-nuv_thresh.

@@ -59,12 +59,12 @@ def simulate(transiting=True, debug=False):
 
     # nuv_thresholds = np.geomspace(30.0, 3000.0, n_nuv_thresholds)
     # nuv_thresholds = np.geomspace(200.0, 600.0, n_nuv_thresholds)
-    nuv_thresholds = np.linspace(10.0, 800.0, n_nuv_thresholds)
+    nuv_thresholds = np.linspace(10.0, 600.0, n_nuv_thresholds)
     # f_life = 1.0
 
     generator_kwargs = {
     'f_life' : 1.0,
-    'deltaT_min' : 10., # Myr. Smaller values seems so slightly enhance #inhabited planets in the FGK sample.
+    'deltaT_min' : 1., # Myr. Smaller values seems to slightly enhance inhabited planets in the FGK sample.
 
     # increase sample size
     'f_eta' : 15,
@@ -167,8 +167,7 @@ def main(debug=False):
     # for transiting in [False]:
     # for transiting in [True]:
         simulate(transiting, debug)
-        pass
-    # DEBUG
+        # pass
     plot_results()
 
 

### 2024-11-21 src/tex/ms.tex
add a reproducibility sentence
discuss LIFE, Nautilus; change title

@@
 %\title{Bioverse: Exoplanet Biosignature Detection Under the UV Threshold Hypothesis}
-\title{Bioverse: Exoplanet Biosignature Patterns Under the UV Threshold Hypothesis for the Origin of Life}
+%\title{Bioverse: Exoplanet Biosignature Patterns Under the UV Threshold Hypothesis for the Origin of Life}
+\title{Bioverse: Potentiablly Observable Exoplanet Biosignature Patterns Under the UV Threshold Hypothesis for the Origin of Life}
 % The Role of Ultraviolet Radiation in Abiogenesis: Insights from Exoplanet Survey Simulations
@@
         Top: Host stars of all transiting \glspl{EEC} and inhabited planets in a simulated transit survey.
-        In the FGK sample, most \glspl{EEC} and all inhabited planets orbit K~dwarfs.
+        In the FGK sample, all \glspl{EEC} and all inhabited planets orbit K~dwarfs.
         In an M~dwarf sample of the same size, the fraction of inhabited planets is larger.\\
@@
 The higher and more variable \gls{NUV} fluxes in M~dwarfs thus increase the likelihood of obtaining strong evidence for or against the UV~Threshold Hypothesis.
+
 The combination of a lack of UV radiation today, which makes biosignature gases more detectable~\citep{Segura2005}, and a UV-rich past that may have enabled abiogenesis could make M~dwarfs the preferred targets for biosignature searches.
-These findings underscore the importance of constraining the UV emission profiles of \gls{EEC} host stars throughout their evolutionary stages to assess the viability of M-dwarf planets as testbeds for theories on the origins of life~\citep{Rimmer2021,Ranjan2023a}.
+We note that relevant mission concepts, such as the Large Interferometer for Exoplanets~\citep[\life,][]{Quanz2022}, include M-dwarf systems among their primary targets~\citep{Kammerer2018,Carrion-Gonzalez2023,Boukrouche2024}.
+Our findings underscore the importance of constraining the UV emission profiles of \gls{EEC} host stars throughout their evolutionary stages to assess the viability of M-dwarf planets as testbeds for theories on the origins of life~\citep{Rimmer2021,Ranjan2023a}.
 
@@
 However, we have shown that the impacts from the combined requirements of the UV~Threshold Hypothesis on the fraction of inhabited planets in a sample are comparable in the non-transiting case.
-The likely required sample sizes are larger than the projected sample sizes of most planned missions; for example the \gls{HWO} expects to characterize a sample of only $\sim 25$ Earth analogs~\citep{Mamajek2023,Tuchow2024}.
+%The projected planet yield of most currently planned exoplanet mission concepts are likely insufficient to test the UV~Threshold Hypothesis:
 
-Meaningful tests of the UV~Threshold Hypothesis may thus require a different survey strategy that increases the sample of characterized planets.
-The Nautilus mission~\citep{Apai2019,Apai2022} aims to characterize up to $\sim 1000$ \gls{EEC} via utilizing an innovative telescope design.
-\todo[inline]{DANIEL wants to work on this paragraph}
-%Constraints on the origins of life from exoplanet surveys may thus be a long-term endeavor, and the UV~Threshold Hypothesis may not be testable yet with the current and next generation of exoplanet surveys.
-To prepare for future biosignature surveys, it is important to refine predictions on the role of UV radiation in prebiotic chemistry with both theoretical and experimental work.
+%Meaningful tests of the UV~Threshold Hypothesis may thus require a different survey strategy that increases the sample of characterized planets.
+Given the challenging nature of detecting and characterizing small (Earth-sized) exoplanets, most exoplanet mission concepts currently under development or considered lack the potential for characterizing large enough samples:
+For example, the \gls{HWO} is expected to characterize a sample of $\sim 25$ Earth analogs~\citep{Mamajek2023,Tuchow2024}.
+Depending on the technical design, \life\ is expected to target 25–80 \glspl{EEC}~\citep{Kammerer2018,Quanz2022}, which could be just sufficient to constrain the UV~Threshold Hypothesis.
+One current exception is the Nautilus Space Observatory concept~\citep{Apai2019,Apai2022}.
+Nautilus aims to characterize up to $\sim 1000$ \gls{EEC} via transmission spectroscopy, building on an innovative optical technology.
+To guide the definition of future biosignature surveys, it is important to refine predictions on the role of UV radiation in prebiotic chemistry with both theoretical and experimental work.
 
@@
 \section*{Reproducibility}
+All code required to reproduce our results, figures, and this article itself is available at \url{https://github.com/matiscke/originsoflife}.
 

### 2024-11-22 src/tex/ms.tex
fix a bunch of typos

@@
 %\title{Bioverse: Exoplanet Biosignature Patterns Under the UV Threshold Hypothesis for the Origin of Life}
-\title{Bioverse: Potentiablly Observable Exoplanet Biosignature Patterns Under the UV Threshold Hypothesis for the Origin of Life}
+\title{Bioverse: Potentially Observable Exoplanet Biosignature Patterns Under the UV Threshold Hypothesis for the Origin of Life}
 % The Role of Ultraviolet Radiation in Abiogenesis: Insights from Exoplanet Survey Simulations
@@
 Surveys targeting extreme NUV flux values and planets orbiting M dwarfs enhance the likelihood of obtaining conclusive results, with sample sizes $\gtrsim$100 providing $\gtrsim$80\% likelihood of strong evidence if abiogenesis rates are high and required NUV fluxes are moderate.
-Conversely, the fraction of inhabited planets sharply decreases for required fluxes exceeding a few hundred erg/s/cm².
+Conversely, the fraction of inhabited planets sharply decreases for required fluxes exceeding a few hundred erg/s/cm$^2$.
 
@@
 \subsection{Semi-analytical assessment}\label{sec:results-semianalytical}
-In Section~\ref{sec:met-semianalytical}, we computed the probability for true positive evidence for $H_\mathrm{1}$ and $H_\mathrm{null}$, respectively (Equations~\ref{eq:bayes_factor}, \ref{eq:bayes_factor2}).
+In Section~\ref{sec:met-semianalytical}, we computed the probability for true positive evidence for $H_\mathrm{1}$ and $H_\mathrm{null}$, respectively (Equations~\ref{eq:bayes_factor},~\ref{eq:bayes_factor2}).
 Figure~\ref{fig:semian_true_evidence} shows how these evidences are distributed for sample sizes \var{semian_Nsamp1} and \var{semian_Nsamp2}, and how likely we are to obtain strong evidence ($BF_{H_i, H_j}$ > 10).
@@
 With \gls{HZ} occupancy as a requirement for abiogenesis, and barring selection biases beyond stellar brightness, the host star distribution of inhabited planets in a simulated transit survey is skewed toward later spectral types.
-For a fixed planet sample size, the fraction of inhabited planets is highest for planets orbiting M~dwarfs due to the higher \gls{NUV} fluxes in the \gls{HZ} of these stars (see Figures~\ref{fig:hz_nuv_evo}, \ref{fig:surveys_FGKM}).
+For a fixed planet sample size, the fraction of inhabited planets is highest for planets orbiting M~dwarfs due to the higher \gls{NUV} fluxes in the \gls{HZ} of these stars (see Figures~\ref{fig:hz_nuv_evo},~\ref{fig:surveys_FGKM}).
 Their \gls{NUV} fluxes are generally highest at early times $\lesssim\SI{100}{\mega\year}$.
@@
 \end{figure}
-Figure \ref{fig:inhabited_aafo_nuv_thresh} shows the fraction of inhabited planets under the UV~Threshold Hypothesis for different threshold \gls{NUV} fluxes and for a high abiogenesis rate of $f_\mathrm{life} = 1$.
+Figure~\ref{fig:inhabited_aafo_nuv_thresh} shows the fraction of inhabited planets under the UV~Threshold Hypothesis for different threshold \gls{NUV} fluxes and for a high abiogenesis rate of $f_\mathrm{life} = 1$.
 This fraction decreases sharply with increasing threshold flux, as fewer planets receive sufficient \gls{NUV} flux for abiogenesis.
@@
 While the inference of a planet's entire UV flux evolution is difficult~\citep[e.g.,][]{Richey-Yowell2023}, the estimated maximum \gls{NUV} flux that a planet was exposed to may be used as a proxy, at least if one is interested in a minimum threshold flux and makes the assumption that planetary surfaces offer protection against \textit{too high} UV flux.
-Indeed, the distribution of the number of planets with detected biosignature in a particular sample of planets with determined maximum \gls{NUV} values $F_{\mathrm{NUV}}$ depends on both the values of $F_{\mathrm{NUV},min}$ and $f_{\mathrm{life}}$ as shown in equation \ref{eq:semian:likelihoodH1}.
+Indeed, the distribution of the number of planets with detected biosignature in a particular sample of planets with determined maximum \gls{NUV} values $F_{\mathrm{NUV}}$ depends on both the values of $F_{\mathrm{NUV},min}$ and $f_{\mathrm{life}}$ as shown in equation~\ref{eq:semian:likelihoodH1}.
 %Avoiding a loss in statistical power due to nuisance parameters is a common difficulty in statistical testing, and is the motivation behind elaborate test statistics.
@@
 For instance, we have made the choice to consider the total number of detected biosignatures as our summary statistic (Equation~\ref{eq:semian:likelihoodH1}), which is not sufficient to infer $F_{\mathrm{NUV},min}$ and $f_{life}$ separately.
-However, by conditioning the Bayes factor to these variables (Equation \ref{eq:bayes_factor}), we calculate the probability distribution of the Bayesian evidence in favor of $H_1$.
+However, by conditioning the Bayes factor to these variables (Equation~\ref{eq:bayes_factor}), we calculate the probability distribution of the Bayesian evidence in favor of $H_1$.
 In doing so, we may evaluate how evidence depends on the uncertainty over these unknown parameters in general terms, without assuming which particular test a future observer might choose to actually perform over real data when available.
@@
 This trade-off can be quantified in terms of expected evidence yield, which we have done in Section~\ref{sec:results-semianalytical}.
-Our analysis shows that regardless of selectivity, sample sizes smaller than 50 likely result in inconclusive tests, and that increasing selectivity towards extreme $F_{\mathrm{NUV}}$ offers limited inference gains compared to the uniform case ($s=0$; Figure \ref{fig:semian_selectivity}).
+Our analysis shows that regardless of selectivity, sample sizes smaller than 50 likely result in inconclusive tests, and that increasing selectivity towards extreme $F_{\mathrm{NUV}}$ offers limited inference gains compared to the uniform case ($s=0$; Figure~\ref{fig:semian_selectivity}).
 For larger samples, however, a narrow distribution of $F_{\mathrm{NUV}}$ may prevent inference entirely.
@@
 As we showed, for high required fluxes the two requirements of simultaneous \gls{HZ} occupancy and sufficient \gls{NUV} flux conspire to diminish the fraction of inhabited planets in the sample.
-Taking the infered fluxes from \citet{Richey-Yowell2023} at face value (but taking into account intrinsic scatter), a minimum required \gls{NUV} flux of $\gtrsim\SI{400}{\erg\per\second\per\centi\meter\squared}$ reduces the fraction of inhabited planets to below $\sim\SI{1}{\percent}$.
-This not only calls for a large sample size and a targeted sample selection prefering high expected past \gls{NUV} fluxes, but also highlight the necessity of continued theoretical and experimental research into the role of UV radiation in prebiotic chemistry~\citep{Ranjan2017b,Rimmer2018,Rimmer2021}.
+Taking the inferred fluxes from \citet{Richey-Yowell2023} at face value (but taking into account intrinsic scatter), a minimum required \gls{NUV} flux of $\gtrsim\SI{400}{\erg\per\second\per\centi\meter\squared}$ reduces the fraction of inhabited planets to below $\sim\SI{1}{\percent}$.
+This not only calls for a large sample size and a targeted sample selection preferring high expected past \gls{NUV} fluxes, but also highlight the necessity of continued theoretical and experimental research into the role of UV radiation in prebiotic chemistry~\citep{Ranjan2017b,Rimmer2018,Rimmer2021}.
 
@@
 By its nature, cyanosulfidic scenario relies on rock surfaces exposed to the planetary atmosphere.
-Water worlds that have their entire planeary surface covered by oceans contradict this requireent and do not allow for the wet-dry cycling inherent to this origin of life scenario.
+Water worlds that have their entire planetary surface covered by oceans contradict this requirement and do not allow for the wet-dry cycling inherent to this origin of life scenario.
 The competition of tectonic stress with gravitational crustal spreading~\citep{Melosh2011} sets the maximum possible height of mountains, which in the solar system does not exceed $\SI{\sim 20}{\kilo\meter}$.

### 2024-12-05 src/tex/ms.tex
include ELTs

@@
 \begin{abstract}
-The origins of life on Earth may provide testable hypotheses about biosignature distributions in exoplanet populations.
-Motivated by the cyanosulfidic origins-of-life scenario, we investigate the hypothesis that a minimum past near-ultraviolet (NUV) flux is necessary for abiogenesis.
-Using a semi-analytical Bayesian framework und the Bioverse survey simulator, we estimate the probability of obtaining strong evidence for or against this UV Threshold Hypothesis with potential future biosignature surveys.
+The origin of life on Earth may provide testable hypotheses about biosignature distributions in exoplanet populations.
+Motivated by the cyanosulfidic origins-of-life scenario, we investigate the hypothesis that a minimum near-ultraviolet (NUV) flux is necessary for abiogenesis.
+Using a semi-analytical Bayesian framework and the Bioverse survey simulator, we estimate the probability of obtaining strong evidence for or against this UV Threshold Hypothesis with potential future biosignature surveys.
 Our results indicate that a correlation between past NUV flux and current biosignature occurrence is observationally testable, but requires sample sizes of at least 50 planets for meaningful constraints.
@@
         \caption{Fraction of inhabited planets for different threshold \gls{NUV} fluxes under the UV~Threshold Hypothesis if the abiogenesis rate is 1.
-Shaded regions denote 90 percent confidence intervals of randomized sample generations, and dashed lines correspond to samples including non-transiting planets.
+        %Shaded regions denote 90 percent confidence intervals of randomized sample generations, and dashed lines correspond to samples including non-transiting planets.
         For all samples, the fraction of inhabited planets drops sharply with increasing threshold \gls{NUV} flux due to the combined effects of never receiving sufficient \gls{NUV} flux for abiogenesis or receiving it before entering the \gls{HZ}.}
@@
 The combination of a lack of UV radiation today, which makes biosignature gases more detectable~\citep{Segura2005}, and a UV-rich past that may have enabled abiogenesis could make M~dwarfs the preferred targets for biosignature searches.
-We note that relevant mission concepts, such as the Large Interferometer for Exoplanets~\citep[\life,][]{Quanz2022}, include M-dwarf systems among their primary targets~\citep{Kammerer2018,Carrion-Gonzalez2023,Boukrouche2024}.
+We note that relevant mission concepts, such as the Large Interferometer for Exoplanets~\citep[\life,][]{Quanz2022}, include M-dwarf systems among their primary targets~\citep{Kammerer2018,Carrion-Gonzalez2023}.
 Our findings underscore the importance of constraining the UV emission profiles of \gls{EEC} host stars throughout their evolutionary stages to assess the viability of M-dwarf planets as testbeds for theories on the origins of life~\citep{Rimmer2021,Ranjan2023a}.
@@
 
+Given the challenging nature of detecting and characterizing small (Earth-sized) exoplanets, most exoplanet mission concepts currently under development or considered lack the potential for characterizing large enough samples.
+Ground-based 25--40-meter class extremely large telescopes are expected to have the capabilities to detect biosignatures on exoplanets like Proxima Centauri~b~\citep[e.g.,][]{Wang2017,Hawker2019,Zhang2024,Vaughan2024}.
+\citet{Hardegree-Ullman2024} used Bioverse to determine potential yields for a 10-year direct imaging and high-resolution spectroscopy survey of O$_2$ on the Giant Magellan Telescope (GMT) and on the Extremely Large Telescope (ELT) and found that between 7 and 19 habitable zone Earth-sized planets could be probed for Earth-like oxygen levels.
+Such a sample is too small to test the UV Threshold Hypothesis, but it may be synergistic with other detection methods.
+
 %Meaningful tests of the UV~Threshold Hypothesis may thus require a different survey strategy that increases the sample of characterized planets.
-Given the challenging nature of detecting and characterizing small (Earth-sized) exoplanets, most exoplanet mission concepts currently under development or considered lack the potential for characterizing large enough samples:
-For example, the \gls{HWO} is expected to characterize a sample of $\sim 25$ Earth analogs~\citep{Mamajek2023,Tuchow2024}.
+The \gls{HWO} is expected to characterize a sample of $\sim 25$ Earth analogs~\citep{Mamajek2023,Tuchow2024}.
 Depending on the technical design, \life\ is expected to target 25–80 \glspl{EEC}~\citep{Kammerer2018,Quanz2022}, which could be just sufficient to constrain the UV~Threshold Hypothesis.

### 2024-12-06 src/tex/ms.tex
add an authorlist

@@
 % Author list
-\author{Martin Schlecker}
+\author[0000-0001-8355-2107]{Martin Schlecker}
 %\author[0000-0001-8355-2107]{Martin Schlecker} # issue with ORCID? https://github.com/showyourwork/showyourwork/issues/454
 \affiliation{Steward Observatory, The University of Arizona, Tucson, AZ 85721, USA; \href{mailto:schlecker@arizona.edu}{schlecker@arizona.edu}}
-\author{et al.}
+\author[0000-0003-3714-5855]{D\'{a}niel Apai}
+\affiliation{Steward Observatory, The University of Arizona, Tucson, AZ 85721, USA; \href{mailto:schlecker@arizona.edu}{schlecker@arizona.edu}}
+\affiliation{Lunar and Planetary Laboratory, The University of Arizona, Tucson, AZ 85721, USA}
+\author[0000-0003-3481-0952]{Antonin Affholder}
+\affiliation{Department of Ecology and Evolutionary Biology, University of Arizona, Tucson AZ, USA}
+\author[0000-0002-5147-9053]{Sukrit Ranjan}
+\affiliation{Lunar and Planetary Laboratory, The University of Arizona, Tucson, AZ 85721, USA}
+\affiliation{Blue Marble Space Institute of Science, Seattle, 98104, USA}
+\author[0000-0002-5806-5566]{Régis Ferrière}
+\affiliation{Department of Ecology and Evolutionary Biology, University of Arizona, Tucson AZ, USA}
+\affiliation{Institut de Biologie de l'École Normale Supérieure, ENS, PSL, Paris, France}
+\affiliation{International Research Laboratory for Interdisciplinary Global Environmental Studies (iGLOBES), CNRS, ENS, PSL, University of Arizona, Tucson AZ, USA}
+\author[0000-0003-3702-0382]{Kevin K.\ Hardegree-Ullman}
+\affiliation{Steward Observatory, The University of Arizona, Tucson, AZ 85721, USA; \href{mailto:schlecker@arizona.edu}{schlecker@arizona.edu}}
+\author[0000-0002-3286-7683]{Tim Lichtenberg}
+\affiliation{Kapteyn Astronomical Institute, University of Groningen, PO Box 800, 9700 AV Groningen, The Netherlands}
+\author[0000-0003-3557-6256]{Stéphane Mazevet}
+\affiliation{Observatoire de la Côte d'Azur, Université Côte d'Azur, Nice, France}
 
 \begin{abstract}
-The origin of life on Earth may provide testable hypotheses about biosignature distributions in exoplanet populations.
-Motivated by the cyanosulfidic origins-of-life scenario, we investigate the hypothesis that a minimum near-ultraviolet (NUV) flux is necessary for abiogenesis.
-Using a semi-analytical Bayesian framework and the Bioverse survey simulator, we estimate the probability of obtaining strong evidence for or against this UV Threshold Hypothesis with potential future biosignature surveys.
-Our results indicate that a correlation between past NUV flux and current biosignature occurrence is observationally testable, but requires sample sizes of at least 50 planets for meaningful constraints.
-These constraints are critically sensitive to the intrinsic abiogenesis rate, as well as host star properties, particularly maximum past NUV fluxes.
-Surveys targeting extreme NUV flux values and planets orbiting M dwarfs enhance the likelihood of obtaining conclusive results, with sample sizes $\gtrsim$100 providing $\gtrsim$80\% likelihood of strong evidence if abiogenesis rates are high and required NUV fluxes are moderate.
-Conversely, the fraction of inhabited planets sharply decreases for required fluxes exceeding a few hundred erg/s/cm$^2$.
-
-Our findings demonstrate the potential of upcoming exoplanet surveys to test origins-of-life hypotheses.
- Beyond specific scenarios, our work underscores the broader value of realistic survey simulations in identifying testable science questions, optimizing mission strategies, and advancing theoretical and experimental studies of abiogenesis.
- This approach emphasizes the need to contextualize biosignature detections, enhancing their credibility and offering new approaches to explore the emergence of life on Earth and beyond.
+A wide variety of scenarios for the origin of life have been proposed, each influencing the prevalence and distribution of biosignatures across exoplanet populations.
+This relationship suggests these scenarios can be tested by predicting biosignature distributions and comparing them with empirical data.
+Here, we demonstrate this approach by focusing on the cyanosulfidic origins-of-life scenario and investigating the hypothesis that a minimum near-ultraviolet (NUV) flux is necessary for abiogenesis.
+Using a semi-analytical Bayesian framework and the Bioverse survey simulator, we constrain the probability of obtaining strong evidence for or against this hypothesis with future biosignature surveys.
+Our results indicate that a correlation between past NUV flux and current biosignature occurrence is testable for sample sizes of $\gtrsim$50 planets.
+The diagnostic power of such tests is critically sensitive to the intrinsic abiogenesis rate and host star properties, particularly maximum past NUV fluxes.
+Surveys targeting a wide range of fluxes, including extremes, and planets orbiting M dwarfs enhance the chances of conclusive results, with sample sizes $\gtrsim$100 providing $\gtrsim$80\% likelihood of strong evidence if abiogenesis rates are high and the required NUV fluxes are moderate.
+For required fluxes exceeding a few hundred erg/s/cm$^2$, both the fraction of inhabited planets and the diagnostic power sharply decrease.
+Our findings demonstrate the potential of exoplanet surveys to test origins-of-life hypotheses.
+ Beyond specific scenarios, this work underscores the broader value of realistic survey simulations in identifying testable science questions, optimizing mission strategies, and advancing theoretical and experimental studies of abiogenesis.
+% This approach emphasizes the need to contextualize biosignature detections, enhancing their credibility and offering new approaches to explore the emergence of life on Earth and beyond.
 \end{abstract}
@@
 A wide variety of scenarios for the origin of life have been proposed~\citep[e.g.,][]{Baross1985,Brasier2011,Mulkidjanian2012,Fox2013,Deamer2015,Westall2018}.
-While we may still be far away from conclusively testing them, new prospects in the search for conditions favorable to life have opened up by thinking of the origin of life as a planetary phenomenon and identifying global-scale environmental properties that might support pathways to life~\citep{Sasselov2020}.
+While we may still be far from conclusively testing these scenarios, new prospects in the search for conditions favorable to life have opened up by thinking of the origin of life as a planetary phenomenon and identifying global-scale environmental properties that might support pathways to life~\citep{Sasselov2020}.
 In particular, specific planetary conditions are needed to create stockpiles of initial compounds for prebiotic chemistry; and planetary processes are required to trigger the prebiotic synthesis.
 Such planetary conditions can be hypothesized for exoplanets located in the \gls{HZ} of their host star, with persistent liquid water on their surface.
-For example, the deep-sea hydrothermal (or hydrothermal-sedimentary) context for the origin of life requires a direct contact of an ocean and the planetary mantle/crust~\citep[e.g.,][]{Baross1985}.
+For example, if deep-sea or sedimentary hydrothermalism is required for abiogenesis, then the insulation of an ocean from the planetary crust minerals (e.g., due to high-pressure ices) may reduce or eliminate the chances of life emerging~\citep[e.g.,][]{Baross1985}.
 The alternate scenario of a surface locally subject to wet-dry cycles requires a planetary exposure to mid-range UV irradiation, as a source of energy and an agent of selection in chemical evolution~\citep[e.g.,][]{Deamer2019}.
@@
 Likewise, if UV light is required to get life started, then there is a minimum planetary UV flux requirement to have an inhabited world.
-This requirement is set by competitor thermal processes; if the photoreaction does not move forward at a rate faster than the competitor thermal process(es), then the abiogenesis scenario cannot function.
+This requirement is set by competing thermal processes; if the photoreaction does not move forward at a rate faster than the competitor thermal process(es), then the abiogenesis scenario cannot function.
 On the other hand, abundant UV light vastly in excess of this threshold does not increase the probability of abiogenesis, since once the UV photochemistry is no longer limiting, some other thermal process in the reaction network will be the rate-limiting process instead.
-Therefore, a putative dependence of life on UV light is best encoded as a step function~\citep[e.g.,][]{Ranjan2017c,Rimmer2018,Rimmer2021}.
+Therefore, a putative dependence of life on UV light is best described as a step function~\citep[e.g.,][]{Ranjan2017c,Rimmer2018,Rimmer2021}.
 
-The goal of this work is to evaluate the potential of potential future exoplanet surveys to test the hypothesis that a minimum past NUV flux is required for abiogenesis.
+The goal of this work is to evaluate the potential of future exoplanet surveys to test the hypothesis that a minimum past NUV flux is required for abiogenesis.
 We focus on one version of the UV Threshold Hypothesis, the so-called cyanosulfidic scenario, which has been refined to the point where the required threshold flux has been measured to be $F_\mathrm{NUV, min} = \SI{6.8\pm3.6 e10}{\photons\per\centi\meter\squared\per\second\per\nano\meter}$ integrated from \SIrange{200}{280}{\nano\meter} at the surface~\citep{Rimmer2018,Rimmer2021a,Rimmer2023,Ranjan2023a}.
@@
 
-We first follow a semi-analytical Bayesian analysis to estimate probabilities of obtaining strong evidence for or against this hypothesis.
+We follow a semi-analytical Bayesian analysis to estimate probabilities of obtaining strong evidence for or against this hypothesis.
 Under the UV Threshold Hypothesis ($H_1$), the probability of an exoplanet having detectable biosignatures is zero if the near-ultraviolet (\gls{NUV}) irradiation is less than the threshold, and it is equal to the (unknown) probability of abiogenesis, $f_\mathrm{life}$ if \gls{NUV} exceeds the threshold for a sufficiently long period of time.
-Under the null hypothesis ($H_0$), that probability simply is $f_\mathrm{life}$, that is, it does not correlate with the UV flux.
+Under the null hypothesis ($H_\mathrm{null}$), that probability simply is $f_\mathrm{life}$, that is, it does not correlate with the UV flux.
 \begin{figure*}
@@
 Figure~\ref{fig:hypotheses} shows these hypotheses as derived from the predictions of the cyanosulfidic scenario.
-Given a sample of planets, where for some of them we have convincing biosignature detections but remain agnostic on $f_\mathrm{life}$, we ask what evidence for $H_1$ and $H_0$ we can expect to obtain.
+Given a sample of planets, where for some of them we have convincing biosignature detections but remain agnostic on $f_\mathrm{life}$, we ask what evidence for $H_1$ and $H_\mathrm{null}$ we can expect to obtain.
 
-A real exoplanet survey will be subject to observational biases, sample selection effects, and the underlying demographics of the planet sample.
+A real exoplanet survey will be subject to observational biases and sample selection effects, and will be constrained by the underlying demographics of the planet sample.
 To assess the information gain of a realistic exoplanet survey, we employed Bioverse~\citep{Bixel2021,Hardegree-Ullman2023,Schlecker2024,Hardegree-Ullman2024}, a framework that integrates multiple components including statistically realistic simulations of exoplanet populations, a survey simulation module, and a hypothesis testing module to evaluate the statistical power of different observational strategies.
@@
 %Test for negative correlations as well?}
-Here, we conduct a theoretical experiment on the UV~Threshold Hypothesis by relating the occurrence of life on an exo-earth candidate with a minimum past quiescent stellar UV flux, focusing on the prebiotically interesting \gls{NUV} range from \SIrange{200}{280}{\nano\meter}.
-\todo[inline]{SUKRIT: got a good reference for this?}  %~\citep{RanjanSasselov2016}.?
-Our concrete hypothesis shall be that life only occurs on planets that at some point in their history have received such radiation at a flux exceeding a threshold $F_\mathrm{NUV, min}$.
+Here, we conduct a theoretical experiment on the UV~Threshold Hypothesis by relating the occurrence of life on an exo-earth candidate with a minimum past quiescent stellar UV flux, focusing on the prebiotically interesting \gls{NUV} range from \SIrange{200}{280}{\nano\meter}~\citep{Ranjan2016}.
+%\todo[inline]{SUKRIT: got a good reference for this?}
+Our core hypothesis shall be that life only occurs on planets that at some point in their history have received such radiation at a flux exceeding a threshold $F_\mathrm{NUV, min}$.
 
@@
 We first assessed the expected probabilities of obtaining true negative or true positive evidence for the UV Threshold Hypothesis ($H_1$) above, as well as the probability for misleading or inconclusive evidence, under idealized conditions.
-This serves as a first-order estimate of the information content of a survey, before we take into account impacts from exoplanet demographics, sample selection, and survey strategy.
+This serves as a first-order estimate of the information content of a survey, before we take into account the effects of exoplanet demographics, sample selection, and survey strategy.
 
 Presumably, not all habitable worlds are inhabited and not all inhabited worlds develop detectable biosignatures.
-The fraction of \glspl{EEC} that are both inhabited and harbor detectable biosignatures at the time when we observe them remains speculative; we aggregate them in the unitless parameter $f_\mathrm{life}$.
+The fraction of \glspl{EEC} that are both inhabited and exhibit detectable biosignatures at the time of observation is unknown and may depend on various parameters that we aggregate under a single term $f_\mathrm{life}$.
 Let us consider the probability to detect a biosignature $P(L)$, and let our observable be the inferred past \gls{NUV} flux of the planet $F_\mathrm{NUV}$.
-Under Hypothesis $H_1$ (Equation~\ref{eq:hypothesis-uv}), there exists a special unknown value of $F_\mathrm{NUV}$, noted $F_\mathrm{NUV, min}$ such that
+Under Hypothesis $H_1$, there exists a special unknown value of $F_\mathrm{NUV}$, noted $F_\mathrm{NUV, min}$ such that
 
@@
 where $f_\mathrm{life}$ is the unknown probability of abiogenesis.
-The corresponding null hypothesis is that there exists no such special value of $F_\mathrm{NUV}$ and that
+The corresponding null hypothesis $H_\mathrm{null}$ is that there exists no such special value of $F_\mathrm{NUV}$ and that
 \begin{equation}
@@
 \end{equation}
+In other words, $H_\mathrm{null}$ states that $P(L)$ is independent of $F_\mathrm{NUV}$.
 
-If we now define a sample of size $n$ as $X=\{F_\mathrm{NUV, i},L_i\}_{i \in [1,n]}$ where $L_i$ is equal to 1 if life is detected and 0 otherwise, we can calculate the evidence for hypothesis $H_i$ against $H_j$ through the Bayes factor
+Defining a sample of size $n$ as $X=\{F_\mathrm{NUV, i},L_i\}_{i \in [1,n]}$ where $L_i$ is equal to 1 if life is detected and 0 otherwise, we can calculate the evidence for hypothesis $H_i$ being true against $H_j$ through the Bayes factor
 \begin{equation}
@@
 
-If we define $k=\sum L_i$ and denote $Y$ the random variable that describes it, $H_\mathrm{null}$ represents the likelihood that the number of planets with life in the sample follows the binomial distribution
+Let $Y=\sum_i^n L_i$ be the random variable counting the number of positive life detections in a sample of size $n$.
+Its probability mass function under the null hypothesis $H_\mathrm{null}$ is that of a binomial distribution:
 \begin{equation}
@@
 
-Under $H_1$, $Y$ also follows a binomial distribution, however it is conditioned by $n_{\lambda}=Card(\{F_\mathrm{NUV, i} \text{ if } F_\mathrm{NUV, i}>F_\mathrm{NUV, min}\}_{i \in [1,n]})$ the number of values of $F_\mathrm{NUV}$ in the experiment that exceed $F_\mathrm{NUV, min}$
+Under $H_1$, $Y$ also follows a binomial distribution, however it is conditioned by $n_{\lambda} = n(\{F_\mathrm{NUV, i} \text{ if } F_\mathrm{NUV, i}>F_\mathrm{NUV, min}\}_{i \in [1,n]})$, the number of values of $F_\mathrm{NUV}$ in the experiment that exceed $F_\mathrm{NUV, min}$
 \begin{equation}
@@
 Given a sample of planets, where for some of them we have convincing biosignature detections but remaining agnostic on $f_\mathrm{life}$: What evidence for $H_\mathrm{1}$ and $H_\mathrm{null}$ can we expect to get?
-Our Bayes factor (Equation~\ref{eq:bayes_factor}) is determined by the unknown variables $f_\mathrm{life}$ and $F_\mathrm{NUV, min}$, as well as the number of planets with biosignature detections in the sample $k$.
+The analytical expression for the Bayes factor of this inference problem (Equation~\ref{eq:bayes_factor}) is determined by the unknown variables $f_\mathrm{life}$ and $F_\mathrm{NUV, min}$, as well as by the summary statistic $Y$ (number of biosignature detections).
 To compute the distribution of evidences, we repeatedly generated samples under $H_\mathrm{1}$ and $H_\mathrm{null}$ and computed the Bayes factors $BF_{H_1,H_\mathrm{null}}$ and $BF_{H_\mathrm{null}, H_1}$.
@@
 
-%\subsection{Bayesian hypothesis testing with \bioverse}
+Under a more realistic scenario, the distribution of $n_{\lambda}$ depends on additional planetary properties and their evolution, as well as on observational biases and sample selection effects of the survey.
+We will address these in the following section.
 
@@
     \item \textbf{Hypothesis testing:} We evaluate the likelihood that a given survey would detect a specified demographic trend in the exoplanet population and estimate the precision with which the survey could constrain the parameters of that trend.
-    A common definition of the null hypothesis $H_0$, which is also applied here, is that there is no relationship between the independent variable (here: maximum NUV flux) and the dependent variable (here: biosignature occurrence).
+    A common definition of the null hypothesis $H_\mathrm{null}$, which is also applied here, is that there is no relationship between the independent variable (here: maximum NUV flux) and the dependent variable (here: biosignature occurrence).
     The alternative hypothesis $H_1$ proposes a specific relationship between the independent and dependent variables.
@@
 To determine the diagnostic capability of a given survey, Bioverse runs multiple iterations of the simulated survey and calculates the fraction of realizations that successfully reject the null hypothesis.
-We use this metric, known as the statistical power, to quantify the potential information content of the survey, identify critical design trades, and find strategies that maximize the survey's scientific return.
+We used this metric, known as the statistical power, to quantify the potential information content of the survey, identify critical design trades, and find strategies that maximize the survey's scientific return.
 % TODO: IF INFERENCE: add:
@@
 The lower limit was suggested as a minimum planet size to retain an atmosphere~\citep{Zahnle2017}.
-To generate planet samples larger than what the stellar catalog in combination with these occurrence rates yields, we scaled up the occurrence rates by a constant factor that yields the desired number of planets.
-This was in particular necessary for the FGK sample, where the rate of transiting planets that occupy the \gls{HZ} is low.
-
 For all survey simulations and hypothesis tests, we repeated the above in a Monte Carlo fashion to generate randomized ensembles of synthetic star and planet populations~\citep[][]{Bixel2021}.
@@
 \subsubsection{Habitable zone occupancy and UV flux}\label{sec:met-hz}
-To test the UV~Threshold Hypothesis, we require that life occurs only on planets with sufficient past UV irradiation exceeding the origins of life threshold $F_\mathrm{NUV, min}$.
-Further, we require this flux to have lasted for a minimum duration $\Delta T_\mathrm{min}$ to allow for a sufficient ``origins timescale''~\citep{Rimmer2023}.
-All common Origin of Life scenarios require water as a solvent;
-we thus consider only rocky planets that may sustain liquid water on their surface, i.e., that occupy their host star's momentary \gls{HZ} during the above period, as well as at the epoch of observation.
-To determine \gls{HZ} occupancy, we take into account the evolution of the host star's luminosity and \gls{HZ} boundaries.
+To construct a test the UV~Threshold Hypothesis, we required that life occurs only on planets with sufficient past UV irradiation exceeding the origins of life threshold $F_\mathrm{NUV, min}$.
+Further, we required this flux to have lasted for a minimum duration $\Delta T_\mathrm{min}$ to allow for a sufficient ``origins timescale''~\citep{Rimmer2023}.
+All commonly investigated origins-of-life scenarios require water as a solvent;
+we thus considered only rocky planets that may sustain liquid water on their surface, i.e., that occupy their host star's momentary \gls{HZ} during the above period, as well as at the epoch of observation.
+To determine \gls{HZ} occupancy, we took into account the evolution of the host star's luminosity and \gls{HZ} boundaries.
 
 The \gls{HZ} describes a region around a star where a planet with Earth's atmospheric composition and climate feedbacks can maintain liquid water on its surface~\citep[e.g.,][]{Ramirez2017,Ramirez2018,MolLous2022,Spinelli2023,Tuchow2023}.
-Here, we adopt estimates that define the \gls{HZ} as the region between the runaway greenhouse transition, where the stellar instellation cannot anymore be balanced through infrared cooling to space~\citep{Ingersoll1969}, and the maximum greenhouse limit, corresponding to the maximum distance at which surface temperatures allowing liquid water can be maintained through a CO$_2$ greenhouse effect~\citep{Kasting1991,Kasting1993,Underwood2003,Kopparapu2013,Kopparapu2014}.
-We use the parametrization in \citet{Kopparapu2014} to derive luminosity and planetary mass-dependent edges of the \gls{HZ} $a_\mathrm{inner}$ and $a_\mathrm{outer}$.
+Here, we adopted orbital distance estimates that define the \gls{HZ} as the region between the runaway greenhouse transition, where the stellar instellation cannot anymore be balanced through infrared cooling to space~\citep{Ingersoll1969}, and the maximum greenhouse limit, corresponding to the maximum distance at which surface temperatures allowing liquid water can be maintained through a CO$_2$ greenhouse effect~\citep{Kasting1991,Kasting1993,Underwood2003,Kopparapu2013,Kopparapu2014}.
+We used the parametrization in \citet{Kopparapu2014} to derive luminosity and planetary mass-dependent distance limits of the \gls{HZ} $a_\mathrm{inner}$ and $a_\mathrm{outer}$.
 
@@
 
-For the \gls{NUV} flux, we use the age- and stellar mass-dependent \gls{NUV} fluxes in the \gls{HZ} obtained by \citet{Richey-Yowell2023}, which considers GALEX UV data in the wavelength range of \SIrange{177}{283}{\nano\meter}.
+For the \gls{NUV} flux, we used the age- and stellar mass-dependent \gls{NUV} fluxes in the \gls{HZ} obtained by \citet{Richey-Yowell2023}, which considers GALEX UV data in the wavelength range of \SIrange{177}{283}{\nano\meter}.
 We linearly interpolate in their measured grid, where we convert spectral type to stellar mass using the midpoints of their mass ranges (\SI{0.75}{\Msun} for K~stars, \SI{0.475}{\Msun} for early-type M~stars,  and \SI{0.215}{\Msun} for late-type M~stars).
@@
 To avoid considering short transitional phases, we require this situation to last for a minimum duration $\Delta T_\mathrm{min} \geq \var{deltaT_min}\,\SI{}{\mega\year}$.
-We assigned the development of life to a random fraction $f_\mathrm{life}$ of all temperate planets fulfilling these requirements.
+We assigned the emergence of life to a random fraction $f_\mathrm{life}$ of all temperate planets fulfilling these requirements.
 For the probability of a planet having detectable biosignatures, $P(\mathrm{bio})$, the UV~Threshold Hypothesis then states
@@
             0 , & F_\mathrm{NUV} < F_\mathrm{NUV, min}\\
-            f_\mathrm{life}, & F_\mathrm{NUV} \geq F_\mathrm{NUV, min} \,\mathrm{and\, in\, HZ} \, \mathrm{for} \, \Delta t \geq \var{deltaT_min} \,\SI{}{\mega\year}
+            f_\mathrm{life}, & F_\mathrm{NUV} \geq F_\mathrm{NUV, min} \\
+            & \text{and in HZ for} \, \Delta t \geq \var{deltaT_min} \,\SI{}{\mega\year}
         \end{cases}
@@
 % Actually it is more than just NUV, it is whatever index that integrates NUV and HZ requiremetns into a single scalar variable for which a threshold exist
-A high biosignature detection rate $f_\mathrm{life}$ increases the evidence (cmp. Equation~\ref{eq:bayes_factor}) but we cannot influence it.
+A high biosignature detection rate $f_\mathrm{life}$ increases the evidence (see Equation~\ref{eq:bayes_factor}) but a survey strategy cannot influence it.
 The same is true for $F_\mathrm{NUV, min}$, where again higher values increase the evidence as the binomial distribution for $H_\mathrm{1}$ gets increasingly skewed and shifted away from the one for $H_\mathrm{null}$.
-However, one might `cherry-pick' exoplanets for which a biosignature test is performed based on \textit{a priori} available contextual information~\citep{catling2018exoplanet} in order to maximize the science yield of investing additional resources.
+However, one might select exoplanets for which a biosignature test is performed based on \textit{a priori} available contextual information~\citep{catling2018exoplanet} in order to maximize the science yield of investing additional resources.
 For instance, the distribution of $F_\mathrm{NUV}$ in the planet sample can be influenced by the survey strategy, and a targeted sampling approach could favor extreme values.
@@
 Their \gls{NUV} fluxes are generally highest at early times $\lesssim\SI{100}{\mega\year}$.
-These host stars, in particular late subtypes, also provide extended periods of increased \gls{NUV} emission that overlap with times when some of these planets occupy the \gls{HZ} (see Figure~\ref{fig:hz_nuv_evo}), our requirement for abiogenesis (compare Equation~\ref{eq:hypothesis-uv}).
+These host stars, in particular late subtypes, also provide extended periods of increased \gls{NUV} emission that overlap with times when some of these planets occupy the \gls{HZ} (see Figure~\ref{fig:hz_nuv_evo}), our requirement for abiogenesis (see Equation~\ref{eq:hypothesis-uv}).
 Thus -- under the UV Threshold Hypothesis -- most inhabited transiting planets in the sample orbit M~dwarfs.
@@
 \subsubsection{Selectivity of simulated transit surveys}
-In Section~\ref{sec:results-semianalytical}, we demonstrated that the probability of obtaining true strong evidence for the hypothesis that life only originates on planets with a minimum past \gls{NUV} flux is sensitive to the distribution of sampled past \gls{NUV} fluxes, i.e., the selectivity of the survey (compare Figure~\ref{fig:semian_selectivity}).
+In Section~\ref{sec:results-semianalytical}, we showed that the probability of obtaining true strong evidence for the hypothesis that life only originates on planets with a minimum past \gls{NUV} flux is sensitive to the distribution of sampled past \gls{NUV} fluxes, i.e., the selectivity of the survey (compare Figure~\ref{fig:semian_selectivity}).
 %The maximum \gls{NUV} distribution of our generic transit survey is rather unimodal with a median around \SI{500}{\erg\per\second\per\centi\meter\squared} (see Figure~\ref{fig:surveys_FGKM}), leading to a selectivity of roughly \var{selectivity_transit_volumelim}.
 For both surveys targeting M dwarfs and those targeting FGK dwarfs, the maximum \gls{NUV} distribution is rather unimodal.
-Applying the approach from Sect.~\ref{sec:results-semianalytical} of fitting a Beta function to the distribution, we find rather low selectivities (see Figure~\ref{fig:surveys_FGKM}), which is likely a detriment for statistical hypothesis tests.
+Applying the approach from Sect.~\ref{sec:results-semianalytical} of fitting a Beta function to the distribution, we find rather low selectivities (see Figure~\ref{fig:surveys_FGKM}), which is likely detrimental for statistical hypothesis tests.
 
@@
         \includegraphics[width=\hsize]{figures/inhabited_aafo_nuv_thresh}
-        \caption{Fraction of inhabited planets for different threshold \gls{NUV} fluxes under the UV~Threshold Hypothesis if the abiogenesis rate is 1.
+        \caption{Fraction of inhabited planets for different threshold \gls{NUV} fluxes under the UV~Threshold Hypothesis if the abiogenesis rate $f_\mathrm{life} = 1$.
         %Shaded regions denote 90 percent confidence intervals of randomized sample generations, and dashed lines correspond to samples including non-transiting planets.
@@
 \end{figure}
-Figure~\ref{fig:inhabited_aafo_nuv_thresh} shows the fraction of inhabited planets under the UV~Threshold Hypothesis for different threshold \gls{NUV} fluxes and for a high abiogenesis rate of $f_\mathrm{life} = 1$.
+Figure~\ref{fig:inhabited_aafo_nuv_thresh} shows the fraction of inhabited planets under the UV~Threshold Hypothesis for different threshold \gls{NUV} fluxes and for the limiting case of an abiogenesis rate of $f_\mathrm{life} = 1$.
 This fraction decreases sharply with increasing threshold flux, as fewer planets receive sufficient \gls{NUV} flux for abiogenesis.
@@
 
-\subsubsection{Statistical power and sensitivity on astrophysical parameters}\label{sec:results-powergrid}
+\subsubsection{Statistical power and sensitivity to astrophysical parameters}\label{sec:results-powergrid}
 We now investigate the sensitivity of the achieved statistical power of our default transit survey to the a priori unconstrained threshold \gls{NUV} flux $F_\mathrm{NUV, min}$ and the abiogenesis rate $f_\mathrm{life}$.
@@
 A key question in the quest to understand the origins of life is which natural processes best explain how living matter spontaneously appears from nonliving matter~\citep[e.g.,][]{Malaterre2022}.
-Using astronomical methods, this question will likely not be testable for individual planets but only by the presence of patterns in planetary populations.
-The cyanosulfidic scenario~\citep{Patel2015}, in particular its predicted existence of a minimum \gls{NUV} flux required for prebiotic chemistry, seems to offer an opportunity to test an origins of life hypothesis with a statistical transit survey sampling planets with varying \gls{NUV} flux histories.
+Using astronomical methods, this question will likely not be testable for individual planets but rather for ensembles of planets.
+The cyanosulfidic scenario for the origins of life~\citep{Patel2015}, in particular its predicted existence of a minimum \gls{NUV} flux required for prebiotic chemistry, offers an opportunity to test an origins of life hypothesis with a statistical transit survey sampling planets with varying \gls{NUV} flux histories.
 In the following, we discuss the prospects of testing the UV~Threshold Hypothesis in light of our results.
@@
 \subsection{Sampling strategy for testing a \gls{NUV} flux threshold} % corresponds to conclusion point 1
-In Sect.~\ref{sec:results-semianalytical}, we show that testing the UV~Threshold Hypothesis suffers from `nuisance' parameters that render inference through astronomical observations difficult.
+In Sect.~\ref{sec:results-semianalytical}, we show that testing the UV~Threshold Hypothesis suffers from `nuisance' parameters that hamper inference through astronomical observations.
 Here, these parameters are the unspecified value of the \gls{NUV} threshold hypothesized to exist under $H_1$, and the unknown probability of detectable life emerging on a habitable planet $f_{\mathrm{life}}$.
 While the inference of a planet's entire UV flux evolution is difficult~\citep[e.g.,][]{Richey-Yowell2023}, the estimated maximum \gls{NUV} flux that a planet was exposed to may be used as a proxy, at least if one is interested in a minimum threshold flux and makes the assumption that planetary surfaces offer protection against \textit{too high} UV flux.
-Indeed, the distribution of the number of planets with detected biosignature in a particular sample of planets with determined maximum \gls{NUV} values $F_{\mathrm{NUV}}$ depends on both the values of $F_{\mathrm{NUV},min}$ and $f_{\mathrm{life}}$ as shown in equation~\ref{eq:semian:likelihoodH1}.
+Indeed, the distribution of the number of planets with detected biosignature in a particular sample of planets with inferred maximum \gls{NUV} values $F_{\mathrm{NUV}}$ depends on both the values of $F_{\mathrm{NUV},min}$ and $f_{\mathrm{life}}$ as shown in equation~\ref{eq:semian:likelihoodH1}.
 %Avoiding a loss in statistical power due to nuisance parameters is a common difficulty in statistical testing, and is the motivation behind elaborate test statistics.
@@
 However, by conditioning the Bayes factor to these variables (Equation~\ref{eq:bayes_factor}), we calculate the probability distribution of the Bayesian evidence in favor of $H_1$.
-In doing so, we may evaluate how evidence depends on the uncertainty over these unknown parameters in general terms, without assuming which particular test a future observer might choose to actually perform over real data when available.
+In doing so, we may evaluate how evidence depends on the uncertainty over these unknown parameters in general terms, without assuming which particular test a future observer might actually choose to perform over real data when available.
 From this, we can see that target selection can strongly affect the conclusiveness of a future test of the UV~Threshold Hypothesis.
@@
 For larger samples, however, a narrow distribution of $F_{\mathrm{NUV}}$ may prevent inference entirely.
-We thus argue that selecting a sample with $F_{\mathrm{NUV}}$ distributed uniformly or emphasizing extreme values should -- barren any practical counterarguments -- be considered in any future attempt at testing the UV~Threshold Hypothesis.
+We thus argue that selecting a sample with $F_{\mathrm{NUV}}$ distributed uniformly or emphasizing extreme values should -- barring any practical counterarguments -- be considered in any future attempt at testing the UV~Threshold Hypothesis.
 Since the practical implementation of an exoplanet survey can stand in the way of such a selection, the following discussion focuses on the results of our transit survey simulations with Bioverse.
@@
 While old M~dwarfs can be considered low-UV environments, a significant fraction of them emit high \gls{NUV} fluxes into their \gls{HZ} during their early stages, in particular later subtypes~\citep{Richey-Yowell2023}.
-This will help to test the high \gls{NUV} flux end of the UV Threshold Hypothesis; a higher occurrence of biosignatures here would support the hypothesis that a higher NUV flux is beneficial or necessary for life.
+This will help to test the high \gls{NUV} flux end of the UV Threshold Hypothesis; a higher occurrence of biosignatures here would support the hypothesis that a higher NUV flux is favorable or necessary for life.
 At the same time, a fraction of host stars in our M~dwarf sample extends it to lower maximum past \gls{NUV} fluxes, enabling tests of the low \gls{NUV} flux end of the hypothesis.
@@
 The combination of a lack of UV radiation today, which makes biosignature gases more detectable~\citep{Segura2005}, and a UV-rich past that may have enabled abiogenesis could make M~dwarfs the preferred targets for biosignature searches.
-We note that relevant mission concepts, such as the Large Interferometer for Exoplanets~\citep[\life,][]{Quanz2022}, include M-dwarf systems among their primary targets~\citep{Kammerer2018,Carrion-Gonzalez2023}.
+We note that relevant mission concepts, such as the Large Interferometer for Exoplanets~\citep[\life,][]{Quanz2022,Glauser2024}, include M-dwarf systems among their primary targets~\citep{Kammerer2018,Carrion-Gonzalez2023}.
 Our findings underscore the importance of constraining the UV emission profiles of \gls{EEC} host stars throughout their evolutionary stages to assess the viability of M-dwarf planets as testbeds for theories on the origins of life~\citep{Rimmer2021,Ranjan2023a}.
@@
 Another impediment to wet-dry cycles may be tidal locking of the planet as it stalls stellar tide-induced water movement and diurnal irradiation variability~\citep[e.g.,][]{Ranjan2017b}.
+However, recent dynamical models suggest tidally locked planets to undergo rapid drift of their sub-stellar point~\citep{Revol2024}.
 
@@
 However, these planets will not pollute the below-threshold subsample, as they will not be able to host life under the UV~Threshold Hypothesis.
+Exoplanet surveys focusing on highly irradiated planets offer an opportunity to constrain the typical oxidation state of rocky exoplanets, providing insights into the average composition of their secondary atmospheres~\citep{Lichtenberg2024}.
 

### 2024-12-10 src/tex/ms.tex
revise abstract

--- 2024-12-05 src/tex/ms.tex
@@
 \begin{abstract}
-The origin of life on Earth may provide testable hypotheses about biosignature distributions in exoplanet populations.
-Motivated by the cyanosulfidic origins-of-life scenario, we investigate the hypothesis that a minimum near-ultraviolet (NUV) flux is necessary for abiogenesis.
-Using a semi-analytical Bayesian framework and the Bioverse survey simulator, we estimate the probability of obtaining strong evidence for or against this UV Threshold Hypothesis with potential future biosignature surveys.
-Our results indicate that a correlation between past NUV flux and current biosignature occurrence is observationally testable, but requires sample sizes of at least 50 planets for meaningful constraints.
-These constraints are critically sensitive to the intrinsic abiogenesis rate, as well as host star properties, particularly maximum past NUV fluxes.
-Surveys targeting extreme NUV flux values and planets orbiting M dwarfs enhance the likelihood of obtaining conclusive results, with sample sizes $\gtrsim$100 providing $\gtrsim$80\% likelihood of strong evidence if abiogenesis rates are high and required NUV fluxes are moderate.
-Conversely, the fraction of inhabited planets sharply decreases for required fluxes exceeding a few hundred erg/s/cm$^2$.
-
-Our findings demonstrate the potential of upcoming exoplanet surveys to test origins-of-life hypotheses.
- Beyond specific scenarios, our work underscores the broader value of realistic survey simulations in identifying testable science questions, optimizing mission strategies, and advancing theoretical and experimental studies of abiogenesis.
- This approach emphasizes the need to contextualize biosignature detections, enhancing their credibility and offering new approaches to explore the emergence of life on Earth and beyond.
+A wide variety of scenarios for the origin of life have been proposed, each influencing the prevalence and distribution of biosignatures across exoplanet populations.
+This relationship suggests these scenarios can be tested by predicting biosignature distributions and comparing them with empirical data.
+Here, we demonstrate this approach by focusing on the cyanosulfidic origins-of-life scenario and investigating the hypothesis that a minimum past near-ultraviolet (NUV) flux is necessary for abiogenesis.
+Using a semi-analytical Bayesian framework and the Bioverse survey simulator, we constrain the probability of obtaining strong evidence for or against this hypothesis with future biosignature surveys.
+Our results indicate that a correlation between past NUV flux and current biosignature occurrence is testable for sample sizes of $\gtrsim$50 planets.
+The diagnostic power of such tests is critically sensitive to the intrinsic abiogenesis rate and host star properties, particularly maximum past NUV fluxes.
+Surveys targeting a wide range of fluxes, including extremes, and planets orbiting M dwarfs enhance the chances of conclusive results, with sample sizes $\gtrsim$100 providing $\gtrsim$80\% likelihood of strong evidence if abiogenesis rates are high and required NUV fluxes are moderate.
+For required fluxes exceeding a few hundred erg/s/cm$^2$, both the fraction of inhabited planets and the diagnostic power sharply decrease.
+Our findings demonstrate the potential of exoplanet surveys to test origins-of-life hypotheses.
+ Beyond specific scenarios, this work underscores the broader value of realistic survey simulations in identifying testable science questions, optimizing mission strategies, and advancing theoretical and experimental studies of abiogenesis.
+% This approach emphasizes the need to contextualize biosignature detections, enhancing their credibility and offering new approaches to explore the emergence of life on Earth and beyond.
 \end{abstract}

### 2024-12-11 src/scripts/hz_nuv_evo.py
add a legend

@@ -151,10 +151,10 @@ def plot_hz_and_nuv(fig, ax, sample, NUV_thresh=300.0, N_sample=2, random_state=
         hz_and_uv = in_hz & (nuv > NUV_thresh)
         t_hzuv = t[hz_and_uv]
 
-        ax.plot(t_hz, [Mst + offset for tt in t_hz], lw=3, c="C0")
-        ax.plot(t_nuv, [Mst - offset for tt in t_nuv], lw=3, c="C2")
+        ax.plot(t_hz, [Mst + offset for tt in t_hz], lw=3, c="C0", label="HZ")
+        ax.plot(t_nuv, [Mst - offset for tt in t_nuv], lw=3, c="C2", label="high NUV")
 
-        # plot emptly rectangle around overlapping region
+        # plot empty rectangle around overlapping region
         if len(t_hzuv) > 1:
             # h_off_left = offset*np.abs(np.log10(t_hzuv[0]))
             # h_off_right = offset*np.abs(np.log10(t_hzuv[-1]))
@@ -162,7 +162,7 @@ def plot_hz_and_nuv(fig, ax, sample, NUV_thresh=300.0, N_sample=2, random_state=
             ax.add_patch(
                 Rectangle((t_hzuv[0] - 0.1*t_hzuv[0], Mst.values[0] - 3*offset),
                           t_hzuv[-1] - t_hzuv[0] + 0.2*t_hzuv[-1],  6*offset,
-                fill=False, lw=2, edgecolor="C1", zorder=99)
+                fill=False, lw=2, edgecolor="C1", zorder=99, label="HZ and high NUV")
             )
 
 
@@ -204,6 +204,7 @@ def plot_hz_and_nuv(fig, ax, sample, NUV_thresh=300.0, N_sample=2, random_state=
             fontweight="bold",
         )
 
+
     return ax
 
 
@@ -232,6 +233,11 @@ if __name__ == "__main__":
     sample = get_example_planets(d)
 
     axs[0] = plot_hz_and_nuv(fig, axs[0], sample, NUV_thresh=NUV_thresh)
+    # add a legend covering only the elements 1, 2, and 3 in the axis:
+    axs[0].legend(loc="lower left", framealpha=0.0, handles=axs[0].get_legend_handles_labels()[0][1:4],
+                  bbox_to_anchor=(0.0, 1.0, 1.0, 0.1), ncol=3)
+
+
[…]

### 2024-12-11 src/scripts/semian_true_evidence.py
expand label

@@ -136,7 +136,7 @@ def plot_true_evidence(ax):
     ax.set_ylim(-0.05, 1.05)
     # ax.set_ylabel("$P(\mathrm{evidence} > x)$")
     ax.set_ylabel("$P(BF_{H_i, H_j} > x$)")
-    ax.set_xlabel("$x$")
+    ax.set_xlabel("Evidence threshold $x$")
 
     return ax
 

### 2024-12-11 src/tex/ms.tex
incorporate Regis'/Antonin's comments

--- 2024-12-06 src/tex/ms.tex
@@
 % Author list
-\author[0000-0001-8355-2107]{Martin Schlecker}
+\author{Martin Schlecker}
 %\author[0000-0001-8355-2107]{Martin Schlecker} # issue with ORCID? https://github.com/showyourwork/showyourwork/issues/454
 \affiliation{Steward Observatory, The University of Arizona, Tucson, AZ 85721, USA; \href{mailto:schlecker@arizona.edu}{schlecker@arizona.edu}}
-\author[0000-0003-3714-5855]{D\'{a}niel Apai}
-\affiliation{Steward Observatory, The University of Arizona, Tucson, AZ 85721, USA; \href{mailto:schlecker@arizona.edu}{schlecker@arizona.edu}}
-\affiliation{Lunar and Planetary Laboratory, The University of Arizona, Tucson, AZ 85721, USA}
-\author[0000-0003-3481-0952]{Antonin Affholder}
-\affiliation{Department of Ecology and Evolutionary Biology, University of Arizona, Tucson AZ, USA}
-\author[0000-0002-5147-9053]{Sukrit Ranjan}
-\affiliation{Lunar and Planetary Laboratory, The University of Arizona, Tucson, AZ 85721, USA}
-\affiliation{Blue Marble Space Institute of Science, Seattle, 98104, USA}
-\author[0000-0002-5806-5566]{Régis Ferrière}
-\affiliation{Department of Ecology and Evolutionary Biology, University of Arizona, Tucson AZ, USA}
-\affiliation{Institut de Biologie de l'École Normale Supérieure, ENS, PSL, Paris, France}
-\affiliation{International Research Laboratory for Interdisciplinary Global Environmental Studies (iGLOBES), CNRS, ENS, PSL, University of Arizona, Tucson AZ, USA}
-\author[0000-0003-3702-0382]{Kevin K.\ Hardegree-Ullman}
-\affiliation{Steward Observatory, The University of Arizona, Tucson, AZ 85721, USA; \href{mailto:schlecker@arizona.edu}{schlecker@arizona.edu}}
-\author[0000-0002-3286-7683]{Tim Lichtenberg}
-\affiliation{Kapteyn Astronomical Institute, University of Groningen, PO Box 800, 9700 AV Groningen, The Netherlands}
-\author[0000-0003-3557-6256]{Stéphane Mazevet}
-\affiliation{Observatoire de la Côte d'Azur, Université Côte d'Azur, Nice, France}
+\author{et al.}
 
@@
 This relationship suggests these scenarios can be tested by predicting biosignature distributions and comparing them with empirical data.
-Here, we demonstrate this approach by focusing on the cyanosulfidic origins-of-life scenario and investigating the hypothesis that a minimum near-ultraviolet (NUV) flux is necessary for abiogenesis.
+Here, we demonstrate this approach by focusing on the cyanosulfidic origins-of-life scenario and investigating the hypothesis that a minimum past near-ultraviolet (NUV) flux is necessary for abiogenesis.
 Using a semi-analytical Bayesian framework and the Bioverse survey simulator, we constrain the probability of obtaining strong evidence for or against this hypothesis with future biosignature surveys.
@@
 The diagnostic power of such tests is critically sensitive to the intrinsic abiogenesis rate and host star properties, particularly maximum past NUV fluxes.
-Surveys targeting a wide range of fluxes, including extremes, and planets orbiting M dwarfs enhance the chances of conclusive results, with sample sizes $\gtrsim$100 providing $\gtrsim$80\% likelihood of strong evidence if abiogenesis rates are high and the required NUV fluxes are moderate.
+Surveys targeting a wide range of fluxes, including extremes, and planets orbiting M dwarfs enhance the chances of conclusive results, with sample sizes $\gtrsim$100 providing $\gtrsim$80\% likelihood of strong evidence if abiogenesis rates are high and required NUV fluxes are moderate.
 For required fluxes exceeding a few hundred erg/s/cm$^2$, both the fraction of inhabited planets and the diagnostic power sharply decrease.
@@
 A wide variety of scenarios for the origin of life have been proposed~\citep[e.g.,][]{Baross1985,Brasier2011,Mulkidjanian2012,Fox2013,Deamer2015,Westall2018}.
-While we may still be far from conclusively testing these scenarios, new prospects in the search for conditions favorable to life have opened up by thinking of the origin of life as a planetary phenomenon and identifying global-scale environmental properties that might support pathways to life~\citep{Sasselov2020}.
+While we may still be far away from conclusively testing them, new prospects in the search for conditions favorable to life have opened up by thinking of the origin of life as a planetary phenomenon and identifying global-scale environmental properties that might support pathways to life~\citep{Sasselov2020}.
 In particular, specific planetary conditions are needed to create stockpiles of initial compounds for prebiotic chemistry; and planetary processes are required to trigger the prebiotic synthesis.
@@
 
-We follow a semi-analytical Bayesian analysis to estimate probabilities of obtaining strong evidence for or against this hypothesis.
+We first follow a semi-analytical Bayesian analysis to estimate probabilities of obtaining strong evidence for or against this hypothesis.
 Under the UV Threshold Hypothesis ($H_1$), the probability of an exoplanet having detectable biosignatures is zero if the near-ultraviolet (\gls{NUV}) irradiation is less than the threshold, and it is equal to the (unknown) probability of abiogenesis, $f_\mathrm{life}$ if \gls{NUV} exceeds the threshold for a sufficiently long period of time.
@@
 %Test for negative correlations as well?}
-Here, we conduct a theoretical experiment on the UV~Threshold Hypothesis by relating the occurrence of life on an exo-earth candidate with a minimum past quiescent stellar UV flux, focusing on the prebiotically interesting \gls{NUV} range from \SIrange{200}{280}{\nano\meter}~\citep{Ranjan2016}.
+Here, we conduct a theoretical experiment on the UV~Threshold Hypothesis by relating the occurrence of life on an exo-earth candidate with a minimum past quiescent stellar UV flux, focusing on the prebiotically interesting \gls{NUV} range from \SIrange{200}{280}{\nano\meter}~\citep{RanjanSasselov2016}.
 %\todo[inline]{SUKRIT: got a good reference for this?}
@@
 
-Under $H_1$, $Y$ also follows a binomial distribution, however it is conditioned by $n_{\lambda} = n(\{F_\mathrm{NUV, i} \text{ if } F_\mathrm{NUV, i}>F_\mathrm{NUV, min}\}_{i \in [1,n]})$, the number of values of $F_\mathrm{NUV}$ in the experiment that exceed $F_\mathrm{NUV, min}$
+Under $H_1$, $Y$ also follows a binomial distribution, however it is conditioned by $n_{\lambda}=Card(\{F_\mathrm{NUV, i} \text{ if } F_\mathrm{NUV, i}>F_\mathrm{NUV, min}\}_{i \in [1,n]})$ the number of values of $F_\mathrm{NUV}$ in the experiment that exceed $F_\mathrm{NUV, min}$
 \begin{equation}
@@
 Given a sample of planets, where for some of them we have convincing biosignature detections but remaining agnostic on $f_\mathrm{life}$: What evidence for $H_\mathrm{1}$ and $H_\mathrm{null}$ can we expect to get?
-The analytical expression for the Bayes factor of this inference problem (Equation~\ref{eq:bayes_factor}) is determined by the unknown variables $f_\mathrm{life}$ and $F_\mathrm{NUV, min}$, as well as by the summary statistic $Y$ (number of biosignature detections).
+The analytical expression for the Bayes factor of this infverence problem (Equation~\ref{eq:bayes_factor}) is determined by the unknown variables $f_\mathrm{life}$ and $F_\mathrm{NUV, min}$, as well as by the summary statistic $Y$ (number of biosignature detections).
 To compute the distribution of evidences, we repeatedly generated samples under $H_\mathrm{1}$ and $H_\mathrm{null}$ and computed the Bayes factors $BF_{H_1,H_\mathrm{null}}$ and $BF_{H_\mathrm{null}, H_1}$.
@@
 The lower limit was suggested as a minimum planet size to retain an atmosphere~\citep{Zahnle2017}.
+To generate planet samples larger than what the stellar catalog in combination with these occurrence rates yields, we scaled up the occurrence rates by a constant factor that yields the desired number of planets.
+This was in particular necessary for the FGK sample, where the rate of transiting planets that occupy the \gls{HZ} is low.
+
 For all survey simulations and hypothesis tests, we repeated the above in a Monte Carlo fashion to generate randomized ensembles of synthetic star and planet populations~\citep[][]{Bixel2021}.
@@
 % Actually it is more than just NUV, it is whatever index that integrates NUV and HZ requiremetns into a single scalar variable for which a threshold exist
-A high biosignature detection rate $f_\mathrm{life}$ increases the evidence (see Equation~\ref{eq:bayes_factor}) but a survey strategy cannot influence it.
+A high biosignature detection rate $f_\mathrm{life}$ increases the evidence (see Equation~\ref{eq:bayes_factor}) but we cannot influence it.
 The same is true for $F_\mathrm{NUV, min}$, where again higher values increase the evidence as the binomial distribution for $H_\mathrm{1}$ gets increasingly skewed and shifted away from the one for $H_\mathrm{null}$.
@@
 The combination of a lack of UV radiation today, which makes biosignature gases more detectable~\citep{Segura2005}, and a UV-rich past that may have enabled abiogenesis could make M~dwarfs the preferred targets for biosignature searches.
-We note that relevant mission concepts, such as the Large Interferometer for Exoplanets~\citep[\life,][]{Quanz2022,Glauser2024}, include M-dwarf systems among their primary targets~\citep{Kammerer2018,Carrion-Gonzalez2023}.
+We note that relevant mission concepts, such as the Large Interferometer for Exoplanets~\citep[\life,][]{Quanz2022}, include M-dwarf systems among their primary targets~\citep{Kammerer2018,Carrion-Gonzalez2023}.
 Our findings underscore the importance of constraining the UV emission profiles of \gls{EEC} host stars throughout their evolutionary stages to assess the viability of M-dwarf planets as testbeds for theories on the origins of life~\citep{Rimmer2021,Ranjan2023a}.
@@
 Another impediment to wet-dry cycles may be tidal locking of the planet as it stalls stellar tide-induced water movement and diurnal irradiation variability~\citep[e.g.,][]{Ranjan2017b}.
-However, recent dynamical models suggest tidally locked planets to undergo rapid drift of their sub-stellar point~\citep{Revol2024}.
 
@@
 However, these planets will not pollute the below-threshold subsample, as they will not be able to host life under the UV~Threshold Hypothesis.
-Exoplanet surveys focusing on highly irradiated planets offer an opportunity to constrain the typical oxidation state of rocky exoplanets, providing insights into the average composition of their secondary atmospheres~\citep{Lichtenberg2024}.
 

### 2024-12-11 src/tex/ms.tex
add Tim's comments

@@
 This relationship suggests these scenarios can be tested by predicting biosignature distributions and comparing them with empirical data.
-Here, we demonstrate this approach by focusing on the cyanosulfidic origins-of-life scenario and investigating the hypothesis that a minimum past near-ultraviolet (NUV) flux is necessary for abiogenesis.
+Here, we demonstrate this approach by focusing on the cyanosulfidic origins-of-life scenario and investigating the hypothesis that a minimum near-ultraviolet (NUV) flux is necessary for abiogenesis.
 Using a semi-analytical Bayesian framework and the Bioverse survey simulator, we constrain the probability of obtaining strong evidence for or against this hypothesis with future biosignature surveys.
@@
 The diagnostic power of such tests is critically sensitive to the intrinsic abiogenesis rate and host star properties, particularly maximum past NUV fluxes.
-Surveys targeting a wide range of fluxes, including extremes, and planets orbiting M dwarfs enhance the chances of conclusive results, with sample sizes $\gtrsim$100 providing $\gtrsim$80\% likelihood of strong evidence if abiogenesis rates are high and required NUV fluxes are moderate.
+Surveys targeting a wide range of fluxes, including extremes, and planets orbiting M dwarfs enhance the chances of conclusive results, with sample sizes $\gtrsim$100 providing $\gtrsim$80\% likelihood of strong evidence if abiogenesis rates are high and the required NUV fluxes are moderate.
 For required fluxes exceeding a few hundred erg/s/cm$^2$, both the fraction of inhabited planets and the diagnostic power sharply decrease.
@@
 A wide variety of scenarios for the origin of life have been proposed~\citep[e.g.,][]{Baross1985,Brasier2011,Mulkidjanian2012,Fox2013,Deamer2015,Westall2018}.
-While we may still be far away from conclusively testing them, new prospects in the search for conditions favorable to life have opened up by thinking of the origin of life as a planetary phenomenon and identifying global-scale environmental properties that might support pathways to life~\citep{Sasselov2020}.
+While we may still be far from conclusively testing these scenarios, new prospects in the search for conditions favorable to life have opened up by thinking of the origin of life as a planetary phenomenon and identifying global-scale environmental properties that might support pathways to life~\citep{Sasselov2020}.
 In particular, specific planetary conditions are needed to create stockpiles of initial compounds for prebiotic chemistry; and planetary processes are required to trigger the prebiotic synthesis.
@@
 
-We first follow a semi-analytical Bayesian analysis to estimate probabilities of obtaining strong evidence for or against this hypothesis.
+We follow a semi-analytical Bayesian analysis to estimate probabilities of obtaining strong evidence for or against this hypothesis.
 Under the UV Threshold Hypothesis ($H_1$), the probability of an exoplanet having detectable biosignatures is zero if the near-ultraviolet (\gls{NUV}) irradiation is less than the threshold, and it is equal to the (unknown) probability of abiogenesis, $f_\mathrm{life}$ if \gls{NUV} exceeds the threshold for a sufficiently long period of time.
@@
 %Test for negative correlations as well?}
-Here, we conduct a theoretical experiment on the UV~Threshold Hypothesis by relating the occurrence of life on an exo-earth candidate with a minimum past quiescent stellar UV flux, focusing on the prebiotically interesting \gls{NUV} range from \SIrange{200}{280}{\nano\meter}~\citep{RanjanSasselov2016}.
+Here, we conduct a theoretical experiment on the UV~Threshold Hypothesis by relating the occurrence of life on an exo-earth candidate with a minimum past quiescent stellar UV flux, focusing on the prebiotically interesting \gls{NUV} range from \SIrange{200}{280}{\nano\meter}~\citep{Ranjan2016}.
 %\todo[inline]{SUKRIT: got a good reference for this?}
@@
 Given a sample of planets, where for some of them we have convincing biosignature detections but remaining agnostic on $f_\mathrm{life}$: What evidence for $H_\mathrm{1}$ and $H_\mathrm{null}$ can we expect to get?
-The analytical expression for the Bayes factor of this infverence problem (Equation~\ref{eq:bayes_factor}) is determined by the unknown variables $f_\mathrm{life}$ and $F_\mathrm{NUV, min}$, as well as by the summary statistic $Y$ (number of biosignature detections).
+The analytical expression for the Bayes factor of this inference problem (Equation~\ref{eq:bayes_factor}) is determined by the unknown variables $f_\mathrm{life}$ and $F_\mathrm{NUV, min}$, as well as by the summary statistic $Y$ (number of biosignature detections).
 To compute the distribution of evidences, we repeatedly generated samples under $H_\mathrm{1}$ and $H_\mathrm{null}$ and computed the Bayes factors $BF_{H_1,H_\mathrm{null}}$ and $BF_{H_\mathrm{null}, H_1}$.
@@
 % Actually it is more than just NUV, it is whatever index that integrates NUV and HZ requiremetns into a single scalar variable for which a threshold exist
-A high biosignature detection rate $f_\mathrm{life}$ increases the evidence (see Equation~\ref{eq:bayes_factor}) but we cannot influence it.
+A high biosignature detection rate $f_\mathrm{life}$ increases the evidence (see Equation~\ref{eq:bayes_factor}) but a survey strategy cannot influence it.
 The same is true for $F_\mathrm{NUV, min}$, where again higher values increase the evidence as the binomial distribution for $H_\mathrm{1}$ gets increasingly skewed and shifted away from the one for $H_\mathrm{null}$.

### 2024-12-11 src/tex/acknowledgments_contributions.tex
add an acknowledgement
fix ackn

@@
 \begin{acknowledgments}
+\section*{Acknowledgments}
     The authors thank Kevin Heng, Dominik Hintz, and Rhys Seeburger for insightful discussions.
@@
 Funding for the DPAC has been provided by national institutions, in particular the institutions participating in the \gaia\ Multilateral Agreement.
+    T.L. was supported by the Branco Weiss Foundation, the Netherlands eScience Center (PROTEUS project, NLESC.OEC.2023.017), and the Alfred P. Sloan Foundation (AEThER project, G202114194).
 %    ...

### 2024-12-11 src/tex/ms.tex
fix ackn

@@
 % Author list
-\author{Martin Schlecker}
+\author[0000-0001-8355-2107]{Martin Schlecker}
 %\author[0000-0001-8355-2107]{Martin Schlecker} # issue with ORCID? https://github.com/showyourwork/showyourwork/issues/454
 \affiliation{Steward Observatory, The University of Arizona, Tucson, AZ 85721, USA; \href{mailto:schlecker@arizona.edu}{schlecker@arizona.edu}}
-\author{et al.}
+\author[0000-0003-3714-5855]{D\'{a}niel Apai}
+\affiliation{Steward Observatory, The University of Arizona, Tucson, AZ 85721, USA; \href{mailto:schlecker@arizona.edu}{schlecker@arizona.edu}}
+\affiliation{Lunar and Planetary Laboratory, The University of Arizona, Tucson, AZ 85721, USA}
+\author[0000-0003-3481-0952]{Antonin Affholder}
+\affiliation{Department of Ecology and Evolutionary Biology, University of Arizona, Tucson AZ, USA}
+\author[0000-0002-5147-9053]{Sukrit Ranjan}
+\affiliation{Lunar and Planetary Laboratory, The University of Arizona, Tucson, AZ 85721, USA}
+\affiliation{Blue Marble Space Institute of Science, Seattle, 98104, USA}
+\author[0000-0002-5806-5566]{Régis Ferrière}
+\affiliation{Department of Ecology and Evolutionary Biology, University of Arizona, Tucson AZ, USA}
+\affiliation{Institut de Biologie de l'École Normale Supérieure, ENS, PSL, Paris, France}
+\affiliation{International Research Laboratory for Interdisciplinary Global Environmental Studies (iGLOBES), CNRS, ENS, PSL, University of Arizona, Tucson AZ, USA}
+\author[0000-0003-3702-0382]{Kevin K.\ Hardegree-Ullman}
+\affiliation{Steward Observatory, The University of Arizona, Tucson, AZ 85721, USA; \href{mailto:schlecker@arizona.edu}{schlecker@arizona.edu}}
+\author[0000-0002-3286-7683]{Tim Lichtenberg}
+\affiliation{Kapteyn Astronomical Institute, University of Groningen, PO Box 800, 9700 AV Groningen, The Netherlands}
+\author[0000-0003-3557-6256]{Stéphane Mazevet}
+\affiliation{Observatoire de la Côte d'Azur, Université Côte d'Azur, Nice, France}
 
@@
 Here, we conduct a theoretical experiment on the UV~Threshold Hypothesis by relating the occurrence of life on an exo-earth candidate with a minimum past quiescent stellar UV flux, focusing on the prebiotically interesting \gls{NUV} range from \SIrange{200}{280}{\nano\meter}~\citep{Ranjan2016}.
-%\todo[inline]{SUKRIT: got a good reference for this?}
 Our core hypothesis shall be that life only occurs on planets that at some point in their history have received such radiation at a flux exceeding a threshold $F_\mathrm{NUV, min}$.

### 2024-12-12 src/scripts/hz_nuv_evo.py
tune colors for visibility, remove gray lines

@@ -136,12 +136,12 @@ def plot_hz_and_nuv(fig, ax, sample, NUV_thresh=300.0, N_sample=2, random_state=
     for i, id in enumerate(inhabited.planetID):
         Mst = eecdf[eecdf.planetID == id]["M_st"]
         t = sample.evolution[id]["time"]
-        ax.plot(
-            np.concatenate((np.array([1e-3]), t)),
-            [Mst] + [Mst for tt in t],
-            lw=1,
-            c="gray",
-        )
+        # ax.plot(
+        #     np.concatenate((np.array([1e-3]), t)),
+        #     [Mst] + [Mst for tt in t],
+        #     lw=1,
+        #     c="gray",
+        # )
         in_hz = sample.evolution[id]["in_hz"]
         nuv = sample.evolution[id]["nuv"]
 
@@ -151,8 +151,8 @@ def plot_hz_and_nuv(fig, ax, sample, NUV_thresh=300.0, N_sample=2, random_state=
         hz_and_uv = in_hz & (nuv > NUV_thresh)
         t_hzuv = t[hz_and_uv]
 
-        ax.plot(t_hz, [Mst + offset for tt in t_hz], lw=3, c="C0", label="HZ")
-        ax.plot(t_nuv, [Mst - offset for tt in t_nuv], lw=3, c="C2", label="high NUV")
+        ax.plot(t_hz, [Mst + offset for tt in t_hz], lw=3.5, c="xkcd:pale gold", label="HZ")
+        ax.plot(t_nuv, [Mst - offset for tt in t_nuv], lw=3.5, c="C1", label="high NUV")
 
         # plot empty rectangle around overlapping region
         if len(t_hzuv) > 1:
@@ -160,9 +160,9 @@ def plot_hz_and_nuv(fig, ax, sample, NUV_thresh=300.0, N_sample=2, random_state=
             # h_off_right = offset*np.abs(np.log10(t_hzuv[-1]))
 
             ax.add_patch(
-                Rectangle((t_hzuv[0] - 0.1*t_hzuv[0], Mst.values[0] - 3*offset),
-                          t_hzuv[-1] - t_hzuv[0] + 0.2*t_hzuv[-1],  6*offset,
-                fill=False, lw=2, edgecolor="C1", zorder=99, label="HZ and high NUV")
[…]

### 2024-12-12 src/scripts/powergrid.py
remove flife < 0.01 part

@@ -46,7 +46,8 @@ def plot_powergrid(grid, fig, ax, **kwargs):
 
     # ax.set_xlim(10, 1100)
     ax.set_xlim(50, 1000)
-    ax.set_ylim(bottom=1e-3, top=1.0)
+    # ax.set_ylim(bottom=1e-3, top=1.0)
+    ax.set_ylim(bottom=1e-2, top=1.0)
     return fig, ax
 
 
@@ -71,4 +72,5 @@ for i, (spt, ax) in enumerate(zip(["FGK", "M"], axs)):
 
 fig.colorbar(axs[0].collections[0], cax=cax, label="Statistical power (%)")
 
+fig.show()
 fig.savefig(paths.figures / "powergrid.pdf")

### 2024-12-12 src/tex/ms.tex
change colors of plots

@@
 % with \gls{EEC} highlighted in green.
-        We show three evolutionary tracks for a threshold flux of $F_\mathrm{NUV, min} = \var{NUV_thresh}\,\SI{}{\erg\per\second\per\centi\meter\squared}$ that occupy the \gls{HZ} (green sections) and exceed the threshold \gls{NUV} flux (blue sections) at different times.
-        Where these sections overlap (orange rectangles), the requirements for abiogenesis are met and we assign a biosignature detection with probability $f_\mathrm{life}$.
+        We show three evolutionary tracks for a threshold flux of $F_\mathrm{NUV, min} = \var{NUV_thresh}\,\SI{}{\erg\per\second\per\centi\meter\squared}$ that occupy the \gls{HZ} (yellow sections) and exceed the threshold \gls{NUV} flux (red sections) at different times.
+        Where these sections overlap (white rectangles), the requirements for abiogenesis are met and we assign a biosignature detection with probability $f_\mathrm{life}$.
         Planet~1 is an \gls{EEC} orbiting a K~dwarf that never receives sufficient \gls{NUV} flux for abiogenesis.
@@
         Center: Distribution of inferred maximum past \gls{NUV} flux in transit surveys targeting \glspl{EEC} around FGK and M~stars, respectively.
-The best-fit beta distributions (gray) correspond to selectivities of $s_\mathrm{FGK} = $\var{selectivity_FGK} and $s_\mathrm{M} = $\var{selectivity_M}.
-Green areas show inhabited planets for an abiogenesis rate of $f_\mathrm{life} = \var{f_life}$ and a generic threshold \gls{NUV} flux $F_\mathrm{NUV, min} = \var{NUV_thresh}\,\SI{}{\erg\per\second\per\centi\meter\squared}$.\\
+The best-fit beta distributions correspond to selectivities of $s_\mathrm{FGK} = $\var{selectivity_FGK} and $s_\mathrm{M} = $\var{selectivity_M}.
+Red areas show inhabited planets for an abiogenesis rate of $f_\mathrm{life} = \var{f_life}$ and a generic threshold \gls{NUV} flux $F_\mathrm{NUV, min} = \var{NUV_thresh}\,\SI{}{\erg\per\second\per\centi\meter\squared}$.\\
         Bottom: Recovered biosignature detections and non-detections of simulated transit surveys. The dashed line denotes $F_\mathrm{NUV, min}$.}

### 2024-12-12 src/scripts/inhabited-aafo-nuv_thresh.py

@@ -143,7 +143,7 @@ def plot_results():
 
 
         # Plot FGK stars
-        ax.plot(nuv_thresholds_fgk, mean_fractions_fgk, label=f"FGK host stars ({lbl})", color='C0', **plotkwargs[lbl])
+        ax.plot(nuv_thresholds_fgk, mean_fractions_fgk, label=f"FGK host stars ({lbl})", color='C2', **plotkwargs[lbl])
         ax.fill_between(nuv_thresholds_fgk, lower_bound_fgk, upper_bound_fgk, color=f"{'C0' if transiting else 'white'}", alpha=0.2)
 
         # Plot M stars

### 2024-12-12 src/scripts/nuv_distribution.py

@@ -37,8 +37,8 @@ def plot_nuv_distribution(sample, data, fig, ax, spt):
     # plot histogram and beta distribution fitted on non-normalized data
     x = np.arange(0.0, 1000.0, 5)
     max_likeli = stats.beta.fit(dataa.max_nuv, method="MM")
-    # ax.hist(dataa.max_nuv, density=True, color="C0")
-    ax.hist([dataa.max_nuv[sample.inhabited], dataa.max_nuv[~sample.inhabited]], stacked=True, density=True, color=["C1", "C0"], label=["inhabited", "EEC"])
+    # ax.hist(dataa.max_nuv, density=True, color="C2")
+    ax.hist([dataa.max_nuv[sample.inhabited], dataa.max_nuv[~sample.inhabited]], stacked=True, density=True, color=["C1", "C2"], label=["inhabited", "EEC"])
     ax.plot(
         x,
         stats.beta.pdf(x, *max_likeli[:2], loc=max_likeli[2], scale=max_likeli[3]),

### 2024-12-12 src/scripts/semian_true_evidence.py

@@ -93,15 +93,16 @@ def plot_true_evidence(ax):
 
         ax.plot(
             gtvec,
-            [np.sum(bfh1_kh1 > g) / len(bfh1_kh1) for g in gtvec],
-            label=showlabel + "True evidence for $H_\mathrm{1}$",
-            c="C0",
-            ls=ls,
-        )
-        ax.plot(
-            gtvec,
             [np.sum(bfh0_kh0 > g) / len(bfh1_kh1) for g in gtvec],
             label=showlabel + "True evidence for $H_\mathrm{null}$",
+            c="C2",
+            ls=ls,
+        )
+
+        ax.plot(
+            gtvec,
+            [np.sum(bfh1_kh1 > g) / len(bfh1_kh1) for g in gtvec],
+            label=showlabel + "True evidence for $H_\mathrm{1}$",
             c="C1",
             ls=ls,
         )
@@ -188,7 +189,7 @@ def plot_evidence_sampsize(ax, s_max=500):
     svec = np.arange(10, s_max)
     true_strong_evidence_p = [P_true_evidence(sampsize=s) for s in svec]
 
-    ax.plot(svec, true_strong_evidence_p)
+    ax.plot(svec, true_strong_evidence_p, c="C1")
     ax.set_ylim(-0.05, 1.05)
     ax.set_ylabel("$P(BF_{H_1, H_\mathrm{null}} > 10$)")
     ax.set_xlabel("Planet sample size $n$")

### 2024-12-12 src/scripts/surveys_FGKM.py

@@ -51,7 +51,7 @@ def plot_inhabited_FGKM(d, fig, ax):
         # kind="count",
         hue="Category",
         order=["F", "G", "K", "M"],
-        palette={"EEC": "dimgray", "inhabited": "C0", "non-EEC": "darkgray"},
+        palette={"EEC": "dimgray", "inhabited": "C1", "non-EEC": "darkgray"},
     )  # , hue_order = [True, False])
     # ax.set_yscale("log")
 
@@ -88,13 +88,13 @@ def plot_nuv_distribution(sample, data, fig, ax, spt, plot_fit=False):
     bins = np.linspace(180.0, 610.0, 28)
 
     max_likeli = stats.beta.fit(dataa.max_nuv, method="MM")
-    # ax.hist(dataa.max_nuv, density=True, color="C0")
+    # ax.hist(dataa.max_nuv, density=True, color="C1")
     ax.hist(
         [dataa.max_nuv[sample.inhabited], dataa.max_nuv[~sample.inhabited]],
         histtype="bar",
         stacked=True,
         density=True,
-        color=["C0", "dimgray"],
+        color=["C1", "dimgray"],
         label=["inhabited", "EEC"],
         bins=bins,
     )
@@ -241,4 +241,5 @@ for spt, axlr in zip(["FGK", "M"], [axs_left, axs_right]):
 
 plt.tight_layout()
 plt.subplots_adjust(hspace=0.7)  # Increase the height space between rows
+fig.show()
 fig.savefig(paths.figures / "surveys_FGKM.pdf")

### 2024-12-12 src/tex/ms.tex
tune abstract

@@
 \begin{abstract}
-A wide variety of scenarios for the origin of life have been proposed, each influencing the prevalence and distribution of biosignatures across exoplanet populations.
+A wide variety of scenarios for the origin of life have been proposed, with many influencing the prevalence and distribution of biosignatures across exoplanet populations.
 This relationship suggests these scenarios can be tested by predicting biosignature distributions and comparing them with empirical data.
 Here, we demonstrate this approach by focusing on the cyanosulfidic origins-of-life scenario and investigating the hypothesis that a minimum near-ultraviolet (NUV) flux is necessary for abiogenesis.
-Using a semi-analytical Bayesian framework and the Bioverse survey simulator, we constrain the probability of obtaining strong evidence for or against this hypothesis with future biosignature surveys.
+Using Bayesian modeling and the Bioverse survey simulator, we constrain the probability of obtaining strong evidence for or against this ``UV Threshold Hypothesis'' with future biosignature surveys.
 Our results indicate that a correlation between past NUV flux and current biosignature occurrence is testable for sample sizes of $\gtrsim$50 planets.
 The diagnostic power of such tests is critically sensitive to the intrinsic abiogenesis rate and host star properties, particularly maximum past NUV fluxes.
-Surveys targeting a wide range of fluxes, including extremes, and planets orbiting M dwarfs enhance the chances of conclusive results, with sample sizes $\gtrsim$100 providing $\gtrsim$80\% likelihood of strong evidence if abiogenesis rates are high and the required NUV fluxes are moderate.
+Surveys targeting a wide range of fluxes, and planets orbiting M dwarfs enhance the chances of conclusive results, with sample sizes $\gtrsim$100 providing $\gtrsim$80\% likelihood of strong evidence if abiogenesis rates are high and the required NUV fluxes are moderate.
 For required fluxes exceeding a few hundred erg/s/cm$^2$, both the fraction of inhabited planets and the diagnostic power sharply decrease.
 Our findings demonstrate the potential of exoplanet surveys to test origins-of-life hypotheses.
- Beyond specific scenarios, this work underscores the broader value of realistic survey simulations in identifying testable science questions, optimizing mission strategies, and advancing theoretical and experimental studies of abiogenesis.
+ Beyond specific scenarios, this work underscores the broader value of realistic survey simulations for future observatories (e.g., HWO, LIFE, ELTs, Nautilus) in identifying testable science questions, optimizing mission strategies, and advancing theoretical and experimental studies of abiogenesis.
 % This approach emphasizes the need to contextualize biosignature detections, enhancing their credibility and offering new approaches to explore the emergence of life on Earth and beyond.

### 2024-12-12 src/tex/ms.tex
add oxidation states

@@
 Exoplanet surveys focusing on highly irradiated planets offer an opportunity to constrain the typical oxidation state of rocky exoplanets, providing insights into the average composition of their secondary atmospheres~\citep{Lichtenberg2024}.
+This is particularly relevant for prebiotic worlds, as varying oxidation states significantly perturb the classical habitable zone concept~\citep{Nicholls2024} and also influence surface \gls{UV} levels through changing atmospheric transmission.
+Optimally, the atmospheric composition of young rocky protoplanets will be probed to constrain the possible range of atmospheric and mantle oxidation states during early planetary evolution by future direct imaging concepts~\citep{Cesario2024}.
 

### 2024-12-12 src/tex/ms.tex
use glossary entry for UV

@@
 For example, if deep-sea or sedimentary hydrothermalism is required for abiogenesis, then the insulation of an ocean from the planetary crust minerals (e.g., due to high-pressure ices) may reduce or eliminate the chances of life emerging~\citep[e.g.,][]{Baross1985}.
-The alternate scenario of a surface locally subject to wet-dry cycles requires a planetary exposure to mid-range UV irradiation, as a source of energy and an agent of selection in chemical evolution~\citep[e.g.,][]{Deamer2019}.
-This “UV Threshold Hypothesis” states that UV light in a specific wavelength range played a constructive role in getting life started on Earth~\citep{Ranjan2016,Ranjan2017c,Rimmer2018,Rapf2016}, and it could provide a probabilistic approach to the interpretation of possible future biosignature detections~\citep[e.g.,][]{Catling2018,Walker2018}.
+The alternate scenario of a surface locally subject to wet-dry cycles requires a planetary exposure to mid-range \gls{UV} irradiation, as a source of energy and an agent of selection in chemical evolution~\citep[e.g.,][]{Deamer2019}.
+This “UV Threshold Hypothesis” states that \gls{UV} light in a specific wavelength range played a constructive role in getting life started on Earth~\citep{Ranjan2016,Ranjan2017c,Rimmer2018,Rapf2016}, and it could provide a probabilistic approach to the interpretation of possible future biosignature detections~\citep[e.g.,][]{Catling2018,Walker2018}.
 
@@
 In this case, a testable prediction would be that planets with high-pressure ices do not show biosignatures.
-Likewise, if UV light is required to get life started, then there is a minimum planetary UV flux requirement to have an inhabited world.
+Likewise, if \gls{UV} light is required to get life started, then there is a minimum planetary \gls{UV} flux requirement to have an inhabited world.
 This requirement is set by competing thermal processes; if the photoreaction does not move forward at a rate faster than the competitor thermal process(es), then the abiogenesis scenario cannot function.
-On the other hand, abundant UV light vastly in excess of this threshold does not increase the probability of abiogenesis, since once the UV photochemistry is no longer limiting, some other thermal process in the reaction network will be the rate-limiting process instead.
-Therefore, a putative dependence of life on UV light is best described as a step function~\citep[e.g.,][]{Ranjan2017c,Rimmer2018,Rimmer2021}.
+On the other hand, abundant \gls{UV} light vastly in excess of this threshold does not increase the probability of abiogenesis, since once the \gls{UV} photochemistry is no longer limiting, some other thermal process in the reaction network will be the rate-limiting process instead.
+Therefore, a putative dependence of life on \gls{UV} light is best described as a step function~\citep[e.g.,][]{Ranjan2017c,Rimmer2018,Rimmer2021}.
 
@@
 Under the UV Threshold Hypothesis ($H_1$), the probability of an exoplanet having detectable biosignatures is zero if the near-ultraviolet (\gls{NUV}) irradiation is less than the threshold, and it is equal to the (unknown) probability of abiogenesis, $f_\mathrm{life}$ if \gls{NUV} exceeds the threshold for a sufficiently long period of time.
-Under the null hypothesis ($H_\mathrm{null}$), that probability simply is $f_\mathrm{life}$, that is, it does not correlate with the UV flux.
+Under the null hypothesis ($H_\mathrm{null}$), that probability simply is $f_\mathrm{life}$, that is, it does not correlate with the \gls{UV} flux.
 \begin{figure*}
@@
 %Test for negative correlations as well?}
-Here, we conduct a theoretical experiment on the UV~Threshold Hypothesis by relating the occurrence of life on an exo-earth candidate with a minimum past quiescent stellar UV flux, focusing on the prebiotically interesting \gls{NUV} range from \SIrange{200}{280}{\nano\meter}~\citep{Ranjan2016}.
+Here, we conduct a theoretical experiment on the UV~Threshold Hypothesis by relating the occurrence of life on an exo-earth candidate with a minimum past quiescent stellar \gls{UV} flux, focusing on the prebiotically interesting \gls{NUV} range from \SIrange{200}{280}{\nano\meter}~\citep{Ranjan2016}.
 Our core hypothesis shall be that life only occurs on planets that at some point in their history have received such radiation at a flux exceeding a threshold $F_\mathrm{NUV, min}$.
@@
 \subsubsection{Habitable zone occupancy and UV flux}\label{sec:met-hz}
-To construct a test the UV~Threshold Hypothesis, we required that life occurs only on planets with sufficient past UV irradiation exceeding the origins of life threshold $F_\mathrm{NUV, min}$.
+To construct a test the UV~Threshold Hypothesis, we required that life occurs only on planets with sufficient past \gls{UV} irradiation exceeding the origins of life threshold $F_\mathrm{NUV, min}$.
 Further, we required this flux to have lasted for a minimum duration $\Delta T_\mathrm{min}$ to allow for a sufficient ``origins timescale''~\citep{Rimmer2023}.
@@
 
-For the \gls{NUV} flux, we used the age- and stellar mass-dependent \gls{NUV} fluxes in the \gls{HZ} obtained by \citet{Richey-Yowell2023}, which considers GALEX UV data in the wavelength range of \SIrange{177}{283}{\nano\meter}.
+For the \gls{NUV} flux, we used the age- and stellar mass-dependent \gls{NUV} fluxes in the \gls{HZ} obtained by \citet{Richey-Yowell2023}, which considers GALEX \gls{UV} data in the wavelength range of \SIrange{177}{283}{\nano\meter}.
 We linearly interpolate in their measured grid, where we convert spectral type to stellar mass using the midpoints of their mass ranges (\SI{0.75}{\Msun} for K~stars, \SI{0.475}{\Msun} for early-type M~stars,  and \SI{0.215}{\Msun} for late-type M~stars).
@@
 %\end{equation}
-i.e., no correlation with UV flux.
+i.e., no correlation with \gls{UV} flux.
 % TODO: IF INFERENCE: add:
@@
 Here, these parameters are the unspecified value of the \gls{NUV} threshold hypothesized to exist under $H_1$, and the unknown probability of detectable life emerging on a habitable planet $f_{\mathrm{life}}$.
-While the inference of a planet's entire UV flux evolution is difficult~\citep[e.g.,][]{Richey-Yowell2023}, the estimated maximum \gls{NUV} flux that a planet was exposed to may be used as a proxy, at least if one is interested in a minimum threshold flux and makes the assumption that planetary surfaces offer protection against \textit{too high} UV flux.
+While the inference of a planet's entire \gls{UV} flux evolution is difficult~\citep[e.g.,][]{Richey-Yowell2023}, the estimated maximum \gls{NUV} flux that a planet was exposed to may be used as a proxy, at least if one is interested in a minimum threshold flux and makes the assumption that planetary surfaces offer protection against \textit{too high} UV flux.
 Indeed, the distribution of the number of planets with detected biosignature in a particular sample of planets with inferred maximum \gls{NUV} values $F_{\mathrm{NUV}}$ depends on both the values of $F_{\mathrm{NUV},min}$ and $f_{\mathrm{life}}$ as shown in equation~\ref{eq:semian:likelihoodH1}.
@@
 
-The combination of a lack of UV radiation today, which makes biosignature gases more detectable~\citep{Segura2005}, and a UV-rich past that may have enabled abiogenesis could make M~dwarfs the preferred targets for biosignature searches.
+The combination of a lack of \gls{UV} radiation today, which makes biosignature gases more detectable~\citep{Segura2005}, and a UV-rich past that may have enabled abiogenesis could make M~dwarfs the preferred targets for biosignature searches.
 We note that relevant mission concepts, such as the Large Interferometer for Exoplanets~\citep[\life,][]{Quanz2022,Glauser2024}, include M-dwarf systems among their primary targets~\citep{Kammerer2018,Carrion-Gonzalez2023}.
-Our findings underscore the importance of constraining the UV emission profiles of \gls{EEC} host stars throughout their evolutionary stages to assess the viability of M-dwarf planets as testbeds for theories on the origins of life~\citep{Rimmer2021,Ranjan2023a}.
+Our findings underscore the importance of constraining the \gls{UV} emission profiles of \gls{EEC} host stars throughout their evolutionary stages to assess the viability of M-dwarf planets as testbeds for theories on the origins of life~\citep{Rimmer2021,Ranjan2023a}.
 
@@
 Taking the inferred fluxes from \citet{Richey-Yowell2023} at face value (but taking into account intrinsic scatter), a minimum required \gls{NUV} flux of $\gtrsim\SI{400}{\erg\per\second\per\centi\meter\squared}$ reduces the fraction of inhabited planets to below $\sim\SI{1}{\percent}$.
-This not only calls for a large sample size and a targeted sample selection preferring high expected past \gls{NUV} fluxes, but also highlight the necessity of continued theoretical and experimental research into the role of UV radiation in prebiotic chemistry~\citep{Ranjan2017b,Rimmer2018,Rimmer2021}.
+This not only calls for a large sample size and a targeted sample selection preferring high expected past \gls{NUV} fluxes, but also highlight the necessity of continued theoretical and experimental research into the role of \gls{UV} radiation in prebiotic chemistry~\citep{Ranjan2017b,Rimmer2018,Rimmer2021}.
 
@@
 Nautilus aims to characterize up to $\sim 1000$ \gls{EEC} via transmission spectroscopy, building on an innovative optical technology.
-To guide the definition of future biosignature surveys, it is important to refine predictions on the role of UV radiation in prebiotic chemistry with both theoretical and experimental work.
+To guide the definition of future biosignature surveys, it is important to refine predictions on the role of \gls{UV} radiation in prebiotic chemistry with both theoretical and experimental work.
 
@@
 \subsubsection{Stellar flares}
-Our assumptions on past UV flux neglect the contribution of stellar flares, which may be hypothesized as an alternative source of UV light~\citep{Buccino2007,Ranjan2017c}.
+Our assumptions on past \gls{UV} flux neglect the contribution of stellar flares, which may be hypothesized as an alternative source of \gls{UV} light~\citep{Buccino2007,Ranjan2017c}.
 This concerns mainly ultracool dwarfs, due to their low quiescent emission and high pre-main sequence stellar activity~\citep{Buccino2007,West2008}.
@@
 \subsubsection{Atmosphere transmission}
-We do not take into account absorption of UV radiation by the planetary atmosphere.
+We do not take into account absorption of \gls{UV} radiation by the planetary atmosphere.
 Theoretical work suggests that the atmosphere of prebiotic Earth was largely transparent at \gls{NUV} wavelengths with the only known source of attenuation being Rayleigh scattering~\citep{Ranjan2017,Ranjan2017c}.
-We thus approximated surface UV flux using top-of-atmosphere fluxes.
+We thus approximated surface \gls{UV} flux using top-of-atmosphere fluxes.
 If there are planets in a sample that do not have a transparent atmosphere at \gls{NUV} wavelengths and require higher fluxes for abiogenesis, the fraction of inhabited planets in the sample will be lower.
@@
     \item Under the UV~Threshold Hypothesis, the fraction of inhabited planets in a transit survey is sensitive to the threshold \gls{NUV} flux and is expected to drop sharply for required fluxes above a few hundred~\SI{}{\erg\per\second\per\centi\meter\squared}.
-    \item If the predicted UV~correlation exists, obtaining strong evidence for the hypothesis is likely ($\gtrsim \SI{80}{\percent}$) for sample sizes $\geq 100$ if the abiogenesis rate is high ($\gtrsim \SI{50}{\percent}$) and if no very high \gls{NUV} fluxes are required.
+    \item If the predicted \gls{UV}~correlation exists, obtaining strong evidence for the hypothesis is likely ($\gtrsim \SI{80}{\percent}$) for sample sizes $\geq 100$ if the abiogenesis rate is high ($\gtrsim \SI{50}{\percent}$) and if no very high \gls{NUV} fluxes are required.
     A survey strategy that targets extreme values of inferred past \gls{NUV} irradiation increases the diagnostic power.

### 2024-12-12 src/scripts/inhabited-aafo-nuv_thresh.py
submitted!

@@ -163,11 +163,13 @@ def plot_results():
 
 @timeit
 def main(debug=False):
-    for transiting in [True, False]:
+    # !only simulate if needed! It takes a while...
+
+    # for transiting in [True, False]:
     # for transiting in [False]:
     # for transiting in [True]:
-        simulate(transiting, debug)
-        # pass
+    #     simulate(transiting, debug)
+
     plot_results()
 
 

### 2025-01-23 src/tex/ms.tex
submitted!
address comment #8

@@
 \documentclass[twocolumn,twocolappendix,linenumbers]{aastex631}
+%\documentclass[twocolumn,twocolappendix]{aastex631}
 %\documentclass[modern,linenumbers]{aastex631}
@@
 We follow a semi-analytical Bayesian analysis to estimate probabilities of obtaining strong evidence for or against this hypothesis.
-Under the UV Threshold Hypothesis ($H_1$), the probability of an exoplanet having detectable biosignatures is zero if the near-ultraviolet (\gls{NUV}) irradiation is less than the threshold, and it is equal to the (unknown) probability of abiogenesis, $f_\mathrm{life}$ if \gls{NUV} exceeds the threshold for a sufficiently long period of time.
+Under the UV Threshold Hypothesis ($H_1$), the probability of an exoplanet having detectable biosignatures is zero if the near-ultraviolet (\gls{NUV}) irradiation is less than the threshold, and it is equal to the (unknown) probability of \rev{life emerging and persisting}, $f_\mathrm{life}$ if \gls{NUV} exceeds the threshold for a sufficiently long period of time.
 Under the null hypothesis ($H_\mathrm{null}$), that probability simply is $f_\mathrm{life}$, that is, it does not correlate with the \gls{UV} flux.
@@
 Presumably, not all habitable worlds are inhabited and not all inhabited worlds develop detectable biosignatures.
-The fraction of \glspl{EEC} that are both inhabited and exhibit detectable biosignatures at the time of observation is unknown and may depend on various parameters that we aggregate under a single term $f_\mathrm{life}$.
+The fraction of \glspl{EEC} that are both inhabited and exhibit detectable biosignatures at the time of observation is unknown \rev{and is represented by the term $f_\mathrm{life}$.
+This encompasses the probability of life both emerging and persisting to produce detectable biosignatures.}
 Let us consider the probability to detect a biosignature $P(L)$, and let our observable be the inferred past \gls{NUV} flux of the planet $F_\mathrm{NUV}$.
@@
 To avoid considering short transitional phases, we require this situation to last for a minimum duration $\Delta T_\mathrm{min} \geq \var{deltaT_min}\,\SI{}{\mega\year}$.
-We assigned the emergence of life to a random fraction $f_\mathrm{life}$ of all temperate planets fulfilling these requirements.
+We assigned the emergence \rev{and persistence} of life to a random fraction $f_\mathrm{life}$ of all temperate planets fulfilling these requirements.
 For the probability of a planet having detectable biosignatures, $P(\mathrm{bio})$, the UV~Threshold Hypothesis then states
@@
 
-The expected resulting evidence further depends on the a priori unknown abiogenesis rate $f_\mathrm{life}$ and on the \gls{NUV} flux threshold.
+The expected resulting evidence further depends on the a priori unknown \rev{rate of life's emergence and persistence} $f_\mathrm{life}$ and on the \gls{NUV} flux threshold.
 \begin{figure*}
@@
 In the following, we fix the sample size to \var{N_nautilus} and consider two different survey strategies targeting FGK and M~dwarfs, respectively.
-We further investigate the sensitivity of the survey to the a priori unknown threshold \gls{NUV} flux $F_\mathrm{NUV, min}$ and the abiogenesis rate $f_\mathrm{life}$.
+We further investigate the sensitivity of the survey to the a priori unknown threshold \gls{NUV} flux $F_\mathrm{NUV, min}$ and the \rev{probability of life emerging and persisting} $f_\mathrm{life}$.
 
@@
 \end{figure}
-Figure~\ref{fig:inhabited_aafo_nuv_thresh} shows the fraction of inhabited planets under the UV~Threshold Hypothesis for different threshold \gls{NUV} fluxes and for the limiting case of an abiogenesis rate of $f_\mathrm{life} = 1$.
+Figure~\ref{fig:inhabited_aafo_nuv_thresh} shows the fraction of inhabited planets under the UV~Threshold Hypothesis for different threshold \gls{NUV} fluxes and for the limiting case of \rev{a probability for life's emergence and persistence} of $f_\mathrm{life} = 1$.
 This fraction decreases sharply with increasing threshold flux, as fewer planets receive sufficient \gls{NUV} flux for abiogenesis.
@@
 \subsubsection{Statistical power and sensitivity to astrophysical parameters}\label{sec:results-powergrid}
-We now investigate the sensitivity of the achieved statistical power of our default transit survey to the a priori unconstrained threshold \gls{NUV} flux $F_\mathrm{NUV, min}$ and the abiogenesis rate $f_\mathrm{life}$.
+We now investigate the sensitivity of the achieved statistical power of our default transit survey to the a priori unconstrained threshold \gls{NUV} flux $F_\mathrm{NUV, min}$ and the abiogenesis \rev{and persistence} rate $f_\mathrm{life}$.
 Figure~\ref{fig:powergrid} shows the statistical power as a function of these parameters for a sample size of $N=\var{N_nautilus}$.
 Values of $F_\mathrm{NUV, min}$ that lie between the extrema of the inferred maximum \gls{NUV} flux increase the achieved statistical power of the survey, as in this case the dataset under the alternative hypothesis $H_1$ differs more from the null hypothesis.
-Furthermore, a higher abiogenesis rate $f_\mathrm{life}$ increases the evidence for $H_1$.
+Furthermore, \rev{higher $f_\mathrm{life}$ increase} the evidence for $H_1$.
 
-Parameter space regions with statistical power above \SI{90}{\percent} lie at abiogenesis rates $f_\mathrm{life} > 0.5$ and mostly at threshold \gls{NUV} fluxes of $\sim \SIrange{200}{400}{\erg\per\second\per\centi\meter\squared}$.
+Parameter space regions with statistical power above \SI{90}{\percent} lie at \rev{$f_\mathrm{life} > 0.5$} and mostly at threshold \gls{NUV} fluxes of $\sim \SIrange{200}{400}{\erg\per\second\per\centi\meter\squared}$.
 Notably, the sensitivity of the M~dwarf sample extends into the low \gls{NUV} flux end due to the broader distribution of maximum past \gls{NUV} fluxes in this sample.
@@
         \caption{Statistical power as a function of threshold \gls{NUV} flux and abiogenesis rate.
-Even for a large sample (here: $N=\var{N_nautilus}$), a high statistical power of the transit survey requires high abiogenesis rates $f_\mathrm{life}$.
+Even for a large sample (here: $N=\var{N_nautilus}$), a high statistical power of the transit survey requires high \rev{rates of life emerging and persisting} $f_\mathrm{life}$.
          Intermediate values of $F_\mathrm{NUV, min}$ are more likely to yield strong evidence than extreme values.
-For abiogenesis rates $f_\mathrm{life} \gtrsim \SI{80}{\percent}$, the sensitivity of the M~dwarf sample extends into the low \gls{NUV} flux end.
+\rev{For $f_\mathrm{life} \gtrsim \SI{80}{\percent}$}, the sensitivity of the M~dwarf sample extends into the low \gls{NUV} flux end.
          }
@@
 Our Bioverse simulations that take into account exoplanet demographics, the evolution of habitability and \gls{NUV} fluxes, and observational biases show that not only the likelihood of a conclusive test of the UV~Threshold Hypothesis, but also the likelihood of successful biosignature detection itself is extremely sensitive to the threshold \gls{NUV} flux if the hypothesis is true.
-Even if all biosignatures can be detected and the nominal abiogenesis rate is very high, say $f_\mathrm{life} = 1$, under the condition that prebiotic chemistry requires a minimum \gls{NUV} flux \textit{and} liquid water, if the threshold flux turns out to be high the probability of finding life on a randomly selected planet may be very low.
+Even if all biosignatures can be detected and the nominal \rev{rate of life's emergence and persistence} is very high, say $f_\mathrm{life} = 1$, under the condition that prebiotic chemistry requires a minimum \gls{NUV} flux \textit{and} liquid water, if the threshold flux turns out to be high the probability of finding life on a randomly selected planet may be very low.
 As we showed, for high required fluxes the two requirements of simultaneous \gls{HZ} occupancy and sufficient \gls{NUV} flux conspire to diminish the fraction of inhabited planets in the sample.

### 2025-01-29 src/tex/ms.tex
address comment #1

@@
 
+\rev{\subsubsection{The UV~Threshold Hypothesis as a narrow step function}
+A key aspect of the UV Threshold Hypothesis is the proposed step-function dependence of abiogenesis likelihood on UV flux.
+This approach stems from the constraints governing photochemical pathways, which exhibit a threshold behavior: below a certain flux, competing thermal reactions dominate, preventing abiogenesis, while above the threshold, UV photochemistry proceeds at sufficient rates, and other stochastic processes become rate-limiting.
+
+\citet{Ranjan2017b} speculated that UV photochemistry might be rate-limiting for abiogenesis, particularly on planets orbiting M dwarfs, due to their lower baseline UV fluxes.
+This could delay abiogenesis by orders of magnitude, resulting in a continuous dependence of abiogenesis likelihood on UV flux.
+However, recent studies challenge this view for the cyanosulfidic scenario.
+\citet{Rimmer2021}  calculated photochemical timescales on early Earth at 180-300 hours (7.5-12.5 days), significantly shorter than the ~1000-year timescale for stochastic geological events~\citep{Rimmer2023}.
+Even with 1000x slower photochemistry on M-dwarf planets, abiogenesis would occur within 20-30 years — geologically negligible compared to stochastic processes — supporting the step-function model.
+
+Nonetheless, alternative abiogenesis pathways or combinations of pathways may exhibit continuous or mixed dependencies on UV flux.
+While the step-function formalism is justified for the cyanosulfidic scenario, future work should explore UV dependencies across other scenarios to refine predictions for biosignature distributions and testable hypotheses.
+}
+
 \subsubsection{Existence of an atmosphere-crust interface}

### 2025-01-30 src/tex/ms.tex
#1: edit

@@
 However, recent studies challenge this view for the cyanosulfidic scenario.
-\citet{Rimmer2021}  calculated photochemical timescales on early Earth at 180-300 hours (7.5-12.5 days), significantly shorter than the ~1000-year timescale for stochastic geological events~\citep{Rimmer2023}.
+\citet{Rimmer2021}  calculated photochemical timescales on early Earth at 180-300 hours (7.5-12.5 days), significantly shorter than the timescale for stochastic geological events~\citep{Rimmer2023}.
 Even with 1000x slower photochemistry on M-dwarf planets, abiogenesis would occur within 20-30 years — geologically negligible compared to stochastic processes — supporting the step-function model.

### 2025-01-31 src/tex/ms.tex
expand hypothesis testing part

@@
 \subsubsection{Hypothesis testing}
-To evaluate the evidence for correctly rejecting the null hypothesis, we employed the Mann-Whitney U test~\citep{Mann1947}.
-%This is a non-parametric test to determine whether two independent samples were drawn from a population with the same distribution and is in particular sensitive to one sample being stochastically greater than the other.
-This is a non-parametric test of the null hypothesis that two independent samples were drawn from a population with the same distribution and is in particular sensitive to one sample being stochastically greater than the other.
-We used the Mann-Whitney U test to compare the distributions of \gls{NUV} fluxes of planets with and without biosignatures.
-The implementation in Bioverse relies on the \texttt{scipy.stats.mannwhitneyu} function~\citep{Virtanen2020} and returns a p-value, for which we set a significance level of $\alpha = 0.05$ to reject the null hypothesis.
-For every hypothesis test, we repeated randomized survey realizations to estimate the fraction of successful rejections of the null hypothesis, i.e., the statistical power of the survey.
+%To evaluate the evidence for correctly rejecting the null hypothesis, we employed the Mann-Whitney U test~\citep{Mann1947}.
+\rev{We ought to choose a statistical test that is sensitive to the UV~Threshold Hypothesis, and that could be realistically conducted in a future transit survey (which may include auxiliary information from ground-based observations, archived data, or models).
+Given the available types of data expected from such surveys, our test shall be non-parametric and compare two samples -- planets with and without biosignatures -- to assess whether they are drawn from the same underlying population in terms of their infered historic maximum NUV flux.
+Common options include the Kolmogorov-Smirnov test, the Brunner-Munzel test~\citep{Brunner2000}, and the Mann-Whitney U test~\citep{Mann1947}.
+Due to its availability and suitability for large sample sizes, we chose the Mann-Whitney U test, which evaluates if one sample is stochastically greater than the other.
+Here, we }
+compare the distributions of \gls{NUV} fluxes of planets with and without biosignatures.
+The implementation in Bioverse relies on the \texttt{scipy.stats.mannwhitneyu} function~\citep{Virtanen2020} and returns a p-value \rev{for each test.
+To balance the trade-off between Type I and Type II error risks, we set the significance level to the widely adopted threshold $\alpha = 0.05$.
+%This choice ensures a reasonable trade-off between detecting a potential dependence of biosignatures on NUV flux while controlling for statistical noise.
+To quantify the diagnostic power of the survey, we conducted repeated randomized realizations and calculated the fraction of successful rejections of the null hypothesis, i.e., the statistical power.
+}
 

### 2025-01-31 src/tex/acknowledgments_contributions.tex
address comment #8
add coding font style for bioverse

@@
 %    mention any people who gave comments in early-on meetings
-%    We thank the anonymous referees for providing insightful reports that helped to improve this manuscript.
+%    We thank the anonymous referee for providing insightful reports that helped to improve this manuscript.
     This material is based upon work supported by the National Aeronautics and Space Administration under Agreement No. 80NSSC21K0593 for the program ``Alien Earths''.
@@
 M.S., D.A., and S.R.\ conceived the project, planned its implementation, and interpreted the results.
-M.S.\ developed the planetary evolution component to Bioverse, carried out the hypothesis tests and statistical analyses, and wrote the manuscript.
+M.S.\ developed the planetary evolution component to \bioverse, carried out the hypothesis tests and statistical analyses, and wrote the manuscript.
 D.A.\ leads the ``Alien Earths'' program through which this project is funded, helped to guide the strategy of the project, and provided text contributions.
@@
 R.F.\ wrote the initial draft of the Introduction and advised on the evolutionary biology aspects of the project.
-K.H.-U.\ contributed to the Bioverse software development and simulations.
+K.H.-U.\ contributed to the \bioverse\ software development and simulations.
 %M.S.\ wrote the manuscript; ... and ... provided text contributions.

### 2025-01-31 src/tex/ms.tex
add coding font style for bioverse

@@
 Here, we demonstrate this approach by focusing on the cyanosulfidic origins-of-life scenario and investigating the hypothesis that a minimum near-ultraviolet (NUV) flux is necessary for abiogenesis.
-Using Bayesian modeling and the Bioverse survey simulator, we constrain the probability of obtaining strong evidence for or against this ``UV Threshold Hypothesis'' with future biosignature surveys.
+Using Bayesian modeling and the \bioverse\ survey simulator, we constrain the probability of obtaining strong evidence for or against this ``UV Threshold Hypothesis'' with future biosignature surveys.
 Our results indicate that a correlation between past NUV flux and current biosignature occurrence is testable for sample sizes of $\gtrsim$50 planets.
@@
 A real exoplanet survey will be subject to observational biases and sample selection effects, and will be constrained by the underlying demographics of the planet sample.
-To assess the information gain of a realistic exoplanet survey, we employed Bioverse~\citep{Bixel2021,Hardegree-Ullman2023,Schlecker2024,Hardegree-Ullman2024}, a framework that integrates multiple components including statistically realistic simulations of exoplanet populations, a survey simulation module, and a hypothesis testing module to evaluate the statistical power of different observational strategies.
+To assess the information gain of a realistic exoplanet survey, we employed \bioverse~\citep{Bixel2021,Hardegree-Ullman2023,Schlecker2024,Hardegree-Ullman2024}, a framework that integrates multiple components including statistically realistic simulations of exoplanet populations, a survey simulation module, and a hypothesis testing module to evaluate the statistical power of different observational strategies.
 
@@
 This paper is organized as follows:
-In Section~\ref{sec:methods}, we introduce both our semi-analytical approach and Bioverse simulations for testing the UV Threshold Hypothesis.
+In Section~\ref{sec:methods}, we introduce both our semi-analytical approach and \bioverse\ simulations for testing the UV Threshold Hypothesis.
 Section~\ref{sec:results} presents the results of these experiments for a generic survey as well as for a realistic transit survey.
@@
 
-\subsection{Exoplanet survey simulations with Bioverse}
-To assess the diagnostic power of realistic exoplanet surveys, we employed our survey simulator and hypothesis testing framework Bioverse~\citep{Bixel2021}.
+\subsection{Exoplanet survey simulations with \bioverse}
+To assess the diagnostic power of realistic exoplanet surveys, we employed our survey simulator and hypothesis testing framework \bioverse~\citep{Bixel2021}.
 The general approach is as follows:
@@
     \item \textbf{Survey simulation:} We simulate the detection and characterization of these exoplanets with a hypothetical survey, taking into account the survey's sensitivity, target selection, and observational biases.
-To model the sensitivity of the information gain of a proposed mission to sample selection and survey strategy, we conduct survey simulations with Bioverse using different sample sizes and survey strategies.
+To model the sensitivity of the information gain of a proposed mission to sample selection and survey strategy, we conduct survey simulations with \bioverse\ using different sample sizes and survey strategies.
     \item \textbf{Hypothesis testing:} We evaluate the likelihood that a given survey would detect a specified demographic trend in the exoplanet population and estimate the precision with which the survey could constrain the parameters of that trend.
@@
     The alternative hypothesis $H_1$ proposes a specific relationship between the independent and dependent variables.
-    Bioverse offers either Bayesian model comparison or non-parametric tests to evaluate the evidence for or against the null hypothesis.
+    \bioverse\ offers either Bayesian model comparison or non-parametric tests to evaluate the evidence for or against the null hypothesis.
 \end{enumerate}
 
-To determine the diagnostic capability of a given survey, Bioverse runs multiple iterations of the simulated survey and calculates the fraction of realizations that successfully reject the null hypothesis.
+To determine the diagnostic capability of a given survey, \bioverse\ runs multiple iterations of the simulated survey and calculates the fraction of realizations that successfully reject the null hypothesis.
 We used this metric, known as the statistical power, to quantify the potential information content of the survey, identify critical design trades, and find strategies that maximize the survey's scientific return.
@@
 \subsubsection{Transit survey simulations}
-With the synthetic star and planet samples generated, we used Bioverse's survey module to simulate noisy measurements of key observables with a transit survey.
+With the synthetic star and planet samples generated, we used \bioverse's survey module to simulate noisy measurements of key observables with a transit survey.
 We assumed a hypothetical mission that can target a large planet sample with high photometric precision and conduct a biosignature search on these planets~\citep[e.g.,][]{Apai2019,Apai2022}.
@@
 compare the distributions of \gls{NUV} fluxes of planets with and without biosignatures.
-The implementation in Bioverse relies on the \texttt{scipy.stats.mannwhitneyu} function~\citep{Virtanen2020} and returns a p-value \rev{for each test.
+The implementation in \bioverse\ relies on the \texttt{scipy.stats.mannwhitneyu} function~\citep{Virtanen2020} and returns a p-value \rev{for each test.
 To balance the trade-off between Type I and Type II error risks, we set the significance level to the widely adopted threshold $\alpha = 0.05$.
@@
 
-\subsection{Survey simulations with Bioverse}\label{sec:results-bioverse}
+\subsection{Survey simulations with \bioverse}\label{sec:results-bioverse}
 %\begin{figure}
@@
 We thus argue that selecting a sample with $F_{\mathrm{NUV}}$ distributed uniformly or emphasizing extreme values should -- barring any practical counterarguments -- be considered in any future attempt at testing the UV~Threshold Hypothesis.
-Since the practical implementation of an exoplanet survey can stand in the way of such a selection, the following discussion focuses on the results of our transit survey simulations with Bioverse.
+Since the practical implementation of an exoplanet survey can stand in the way of such a selection, the following discussion focuses on the results of our transit survey simulations with \bioverse.
 
@@
 \subsection{Sensitivity to astrophysical parameters} % corresponds to conclusion point 3
-Our Bioverse simulations that take into account exoplanet demographics, the evolution of habitability and \gls{NUV} fluxes, and observational biases show that not only the likelihood of a conclusive test of the UV~Threshold Hypothesis, but also the likelihood of successful biosignature detection itself is extremely sensitive to the threshold \gls{NUV} flux if the hypothesis is true.
+Our \bioverse\ simulations that take into account exoplanet demographics, the evolution of habitability and \gls{NUV} fluxes, and observational biases show that not only the likelihood of a conclusive test of the UV~Threshold Hypothesis, but also the likelihood of successful biosignature detection itself is extremely sensitive to the threshold \gls{NUV} flux if the hypothesis is true.
 Even if all biosignatures can be detected and the nominal \rev{rate of life's emergence and persistence} is very high, say $f_\mathrm{life} = 1$, under the condition that prebiotic chemistry requires a minimum \gls{NUV} flux \textit{and} liquid water, if the threshold flux turns out to be high the probability of finding life on a randomly selected planet may be very low.
@@
 Even under idealized conditions, obtaining strong evidence for or against the hypothesis likely requires sample sizes on the order of 100 (see Section~\ref{sec:results-semianalytical}).
-This is true for a future transit survey, the specifics of which we have reflected in our Bioverse simulations (see Section~\ref{sec:results-bioverse}).
+This is true for a future transit survey, the specifics of which we have reflected in our \bioverse\ simulations (see Section~\ref{sec:results-bioverse}).
 However, we have shown that the impacts from the combined requirements of the UV~Threshold Hypothesis on the fraction of inhabited planets in a sample are comparable in the non-transiting case.
@@
 Ground-based 25--40-meter class extremely large telescopes are expected to have the capabilities to detect biosignatures on exoplanets like Proxima Centauri~b~\citep[e.g.,][]{Wang2017,Hawker2019,Zhang2024,Vaughan2024}.
-\citet{Hardegree-Ullman2024} used Bioverse to determine potential yields for a 10-year direct imaging and high-resolution spectroscopy survey of O$_2$ on the Giant Magellan Telescope (GMT) and on the Extremely Large Telescope (ELT) and found that between 7 and 19 habitable zone Earth-sized planets could be probed for Earth-like oxygen levels.
+\citet{Hardegree-Ullman2024} used \bioverse\ to determine potential yields for a 10-year direct imaging and high-resolution spectroscopy survey of O$_2$ on the Giant Magellan Telescope (GMT) and on the Extremely Large Telescope (ELT) and found that between 7 and 19 habitable zone Earth-sized planets could be probed for Earth-like oxygen levels.
 Such a sample is too small to test the UV Threshold Hypothesis, but it may be synergistic with other detection methods.
@@
 To this end, we first employed a semi-analytical Bayesian analysis to estimate probabilities of obtaining strong evidence for or against this hypothesis.
-We then used the Bioverse framework to assess the diagnostic power of realistic transit surveys, taking into account exoplanet demographics, time-dependency of habitability and \gls{NUV} fluxes, observational biases, and target selection.
+We then used the \bioverse\ framework to assess the diagnostic power of realistic transit surveys, taking into account exoplanet demographics, time-dependency of habitability and \gls{NUV} fluxes, observational biases, and target selection.
 

### 2025-01-31 src/tex/ms.tex
address #3, different HZs

@@
 We used the parametrization in \citet{Kopparapu2014} to derive luminosity and planetary mass-dependent distance limits of the \gls{HZ} $a_\mathrm{inner}$ and $a_\mathrm{outer}$.
+\rev{The specific boundaries of the habitable zone are known to be sensitive to the star’s luminosity, spectral type, the planet’s mass, and the planet’s atmospheric properties~\citep[e.g.,][]{Pierrehumbert2011,Ramirez2014,Ramirez2017,Ramirez2018b,Koll2019,Ramirez2018,Ramirez2020,Bonati2021,Chaverot2022,Turbet2023}.
+The specific habitable zone limits can therefore be variable.
+Here -- to constrain the complexity of this study -- we adopted a conservative approach that uses the commonly used limits.
+}
 

### 2025-01-31 src/tex/ms.tex
fix step function section

@@
 \rev{\subsubsection{The UV~Threshold Hypothesis as a narrow step function}
-A key aspect of the UV Threshold Hypothesis is the proposed step-function dependence of abiogenesis likelihood on UV flux.
-This approach stems from the constraints governing photochemical pathways, which exhibit a threshold behavior: below a certain flux, competing thermal reactions dominate, preventing abiogenesis, while above the threshold, UV photochemistry proceeds at sufficient rates, and other stochastic processes become rate-limiting.
+}
+\rev{A key aspect of the UV Threshold Hypothesis is the proposed step-function dependence of abiogenesis likelihood on UV flux.
+This approach stems from the constraints governing photochemical pathways, which exhibit a threshold behavior: below a certain flux, competing thermal reactions dominate, preventing abiogenesis, while above the threshold, UV photochemistry proceeds at sufficient rates, and other stochastic processes become rate-limiting.}
 
-\citet{Ranjan2017b} speculated that UV photochemistry might be rate-limiting for abiogenesis, particularly on planets orbiting M dwarfs, due to their lower baseline UV fluxes.
+\rev{\citet{Ranjan2017b} speculated that UV photochemistry might be rate-limiting for abiogenesis, particularly on planets orbiting M dwarfs, due to their lower baseline UV fluxes.
 This could delay abiogenesis by orders of magnitude, resulting in a continuous dependence of abiogenesis likelihood on UV flux.
@@
 Even with 1000x slower photochemistry on M-dwarf planets, abiogenesis would occur within 20-30 years — geologically negligible compared to stochastic processes — supporting the step-function model.
+}
 
-Nonetheless, alternative abiogenesis pathways or combinations of pathways may exhibit continuous or mixed dependencies on UV flux.
+\rev{Nonetheless, alternative abiogenesis pathways or combinations of pathways may exhibit continuous or mixed dependencies on UV flux.
 While the step-function formalism is justified for the cyanosulfidic scenario, future work should explore UV dependencies across other scenarios to refine predictions for biosignature distributions and testable hypotheses.

### 2025-02-01 src/tex/ms.tex
draft something on atm escape

@@
 
+\rev{
+\subsubsection{Other mechanisms regulating habitability and abiogenesis}
+Our study focuses primarily on two factors that may regulate the emergence and persistence of life on a planet: the \gls{NUV} flux and the viability of liquid water, which provide a testable framework for assessing biosignature distributions.
+Of course, planetary habitability and abiogenesis depend on a broader range of physical and chemical conditions.
+For one, a planet must retain an atmosphere capable of supporting liquid water and surface chemistry.
+Atmospheric escape processes have been studied in detail, and their occurrence is supported by planet formation models~\citep[e.g.,][]{Owen2013,Schlichting2014a,Ginzburg2016b,Mordasini2020,Burn2024} as well as exoplanet demographics~\citep[e.g.,][]{Owen2019a,Bergsten2022,Rogers2021a}.
+Population synthesis studies demonstrate that models of atmospheric escape can explain key statistical features of the observed exoplanet population~\citep{Rogers2020,Emsenhuber2021b,Schlecker2021b,Burn2024}.}
+
+\rev{
+Planets around M~dwarfs, in particular, may be subject to significant atmospheric loss due to the extended high-luminosity pre-main-sequence phase of their host stars~\citep[e.g.,][]{Luger2015}.
+%During this period, extreme X-ray and UV irradiation can drive rapid hydrodynamic escape, potentially stripping a planet of its primordial atmosphere before abiogenesis can occur.
+%Even after this early phase, M-dwarf planets remain exposed to intense stellar activity, leading to further erosion of secondary atmospheres if not replenished through volcanic outgassing or other mechanisms.
+Recent works~\citep[e.g.,][]{ParkCoy2024,Luque2024} interpret JWST eclipse measurements as growing evidence for the absence of substantial atmospheres on M-dwarf rocky exoplanets.
+If these planets indeed lack atmospheres, their habitability would be primarily dictated by the rate of atmospheric escape~\citep[e.g.,][]{Owen2020}.
+However, alternative interpretations remain viable: \citet{Ducrot2024} and \citet{Hammond2025} recently demonstrated that current JWST data cannot definitively distinguish between a bare-rock scenario and an atmosphere composed primarily of N$_2$-CO$_2$-H$_2$O.
+This underscores the need for future observations and modeling efforts to constrain atmospheric retention and escape processes more robustly.}
+...
+
+%For instance, atmospheric escape processes may play a critical role not only in maintaining a planet's atmosphere but also in regulating the planet's water inventory and the availability of volatile elements for prebiotic chemistry.
+
+\rev{While not modeled here, the above additional mechanisms may offer directions for future work.}
+
 \section{Conclusions}

### 2025-02-04 src/tex/ms.tex
added some discussion in 4.1 addressing Reviewers 1 comment number 2

@@
 \label{sec:intro}
-A wide variety of scenarios for the origin of life have been proposed~\citep[e.g.,][]{Baross1985,Brasier2011,Mulkidjanian2012,Fox2013,Deamer2015,Westall2018}.
+\rev{While the probability of abiogenesis is unknown and has primarily been explored through statistical arguments~\citep[e.g.,][]{Spiegel2012,Kipping2021,Lingam2024},
+a} wide variety of scenarios for the origin of life have been proposed~\citep[e.g.,][]{Baross1985,Brasier2011,Mulkidjanian2012,Fox2013,Deamer2015,Westall2018}.
 While we may still be far from conclusively testing these scenarios, new prospects in the search for conditions favorable to life have opened up by thinking of the origin of life as a planetary phenomenon and identifying global-scale environmental properties that might support pathways to life~\citep{Sasselov2020}.
@@
 
+\rev{Our Bayes factor analysis quantifies how much evidence future data may provide in favor of a hypothesis against another one (the null hypothesis).
+Typical frequentist (p-value based) tests usually evaluate only the likelihood of the null hypothesis, thus requiring that the experiment be carefully crafted such that rejection of the null hypothesis can be interpreted in favor of a particular studied hypothesis (for instance that samples from coin tosses score a very low likelihood under the hypothesis of a fair coin may indicate that the coin is biased).
+In contrast, Bayes factor analyses such as ours introduce additional constraint to the testing by involving the likelihood of the tested alternate hypothesis (equation \ref{eq:bayes_factor}).
+Thus, under Bayes-factor analyses, evidence for the tested hypothesis is not only measured by how unlikely the null hypothesis is given some observation but rather by how much more likely than the null hypothesis is the alternate hypothesis.
+The most exhaustive form of hypothesis selection test would be performed by using the entire Bayes theorem (equation \ref{eq:BayesTheorem}), computing the \emph{posterior} probability of the tested hypothesis.
+Doing so enables to quantify the information gain from performing some observation (by comparing the posterior with the prior), and to perform inference with consideration for prior information.
+\begin{equation}
+    \label{eq:BayesTheorem}
+    P(H_i|X) = \frac{P(X|H_i)P(H_i)}{P(X)}
+\end{equation}
+However, implementing the Bayes theorem can prove difficult in practice, owing to the (usually) subjective nature of prior distributions.
+Two prior distributions are required for the computation of the posterior distribution: the prior distribution for the considered model and hypothesis, and the prior distribution for the observation.
+While the former requires that at least some amount of consensus is identifiable in the community of scientists that discuss this hypothesis, the latter also requires that either empirical distributions for the prior distribution of the observation be available or that the exhaustive list of mechanistic pathways that could potentially lead to this observation be known and tractable.
+Taking the ratio of the posterior evidence for two alternate hypotheses (i.e.
+of equation \ref{eq:BayesTheorem}) allows this problematic term, the prior probability for the observation, to cancel out (equation \ref{eq:postratio}).
+\begin{equation}
+    \label{eq:postratio}
+    \frac{P(H_1|X)}{P(H_0|X)} = \frac{P(X|H_0)P(H_0)}{P(X|H_0)P(H_0)}
+\end{equation}
+This trades off the opportunity to estimate the posterior probability of a specific hypothesis with the possiblity to narrow down the problem to two specific competing hypotheses.
+Could there be factors other than the past UV flux that would determine the occurrence rate of biosignatures in terrestrial exoplanets?
+Surely, there is a vast array of possible mechanisms that could be considered which could shape the prior probability of an observed sample of biosignatures $X$.
+Truncating the ratio of prior probabilities from equation \ref{eq:postratio} results in the Bayes factor (equation \ref{eq:bayes_factor}), which we used in our semi-analytical analysis (section \ref{sec:results-semianalytical}).
+While the Bayes ratio only paints part of the picture of what a full inference pipeline would look line in a future survey aimed at testing origin of life hypotheses from the observation of biosignatures, it captures an essential part of the yield of a particular strategy: by how much would these observations tip the scale of model selection relative to prior expectations.}
+
+\rev{Our Bayes factor analysis quantifies how much evidence future data may provide in favor of a hypothesis against another one (the null hypothesis).
+Typical frequentist (p-value based) tests usually evaluate only the likelihood of the null hypothesis, thus requiring that the experiment be carefully crafted such that rejection of the null hypothesis can be interpreted in favor of a particular studied hypothesis (for instance that samples from coin tosses score a very low likelihood under the hypothesis of a fair coin may indicate that the coin is biased).
+In contrast, Bayes factor analyses such as ours introduce additional constraint to the testing by involving the likelihood of the tested alternate hypothesis (equation \ref{eq:bayes_factor}).
+Thus, under Bayes-factor analyses, evidence for the tested hypothesis is not only measured by how unlikely the null hypothesis is given some observation but rather by how much more likely than the null hypothesis is the alternate hypothesis.
+The most exhaustive form of hypothesis selection test would be performed by using the entire Bayes theorem (equation \ref{eq:BayesTheorem}), computing the \emph{posterior} probability of the tested hypothesis.
+Doing so enables to quantify the information gain from performing some observation (by comparing the posterior with the prior), and to perform inference with consideration for prior information.
+\begin{equation}
+    \label{eq:BayesTheorem}
+    P(H_i|X) = \frac{P(X|H_i)P(H_i)}{P(X)}
+\end{equation}
+However, implementing the Bayes theorem can prove difficult in practice, owing to the (usually) subjective nature of prior distributions.
+Two prior distributions are required for the computation of the posterior distribution: the prior distribution for the considered model and hypothesis, and the prior distribution for the observation.
+While the former requires that at least some amount of consensus is identifiable in the community of scientists that discuss this hypothesis, the latter also requires that either empirical distributions for the prior distribution of the observation be available or that the exhaustive list of mechanistic pathways that could potentially lead to this observation be known and tractable.
+Taking the ratio of the posterior evidence for two alternate hypotheses (i.e.
+of equation \ref{eq:BayesTheorem}) allows this problematic term, the prior probability for the observation, to cancel out (equation \ref{eq:postratio}).
+\begin{equation}
+    \label{eq:postratio}
+    \frac{P(H_1|X)}{P(H_0|X)} = \frac{P(X|H_0)P(H_0)}{P(X|H_0)P(H_0)}
+\end{equation}
+This trades off the opportunity to estimate the posterior probability of a specific hypothesis with the possiblity to narrow down the problem to two specific competing hypotheses.
+Could there be factors other than the past UV flux that would determine the occurrence rate of biosignatures in terrestrial exoplanets?
+Surely, there is a vast array of possible mechanisms that could be considered which could shape the prior probability of an observed sample of biosignatures $X$.
+Truncating the ratio of prior probabilities from equation \ref{eq:postratio} results in the Bayes factor (equation \ref{eq:bayes_factor}), which we used in our semi-analytical analysis (section \ref{sec:results-semianalytical}).
+While the Bayes ratio only paints part of the picture of what a full inference pipeline would look line in a future survey aimed at testing origin of life hypotheses from the observation of biosignatures, it captures an essential part of the yield of a particular strategy: by how much would these observations tip the scale of model selection relative to prior expectations.}
+
 %We stress that a future observer will likely be able to leverage standard frequentist testing methods which may be more likely to yield conclusive results (whether in favor or in disfavor of $H_1$).
@@
 For one, a planet must retain an atmosphere capable of supporting liquid water and surface chemistry.
-Atmospheric escape processes have been studied in detail, and their occurrence is supported by planet formation models~\citep[e.g.,][]{Owen2013,Schlichting2014a,Ginzburg2016b,Mordasini2020,Burn2024} as well as exoplanet demographics~\citep[e.g.,][]{Owen2019a,Bergsten2022,Rogers2021a}.
+Atmospheric escape processes have been studied in detail, and their occurrence is supported by planet formation models~\citep[e.g.,][]{Owen2013,Schlichting2014,Ginzburg2016,Mordasini2020,Burn2024} as well as exoplanet demographics~\citep[e.g.,][]{Owen2019,Bergsten2022,Rogers2021}.
 Population synthesis studies demonstrate that models of atmospheric escape can explain key statistical features of the observed exoplanet population~\citep{Rogers2020,Emsenhuber2021b,Schlecker2021b,Burn2024}.}
@@
 %Even after this early phase, M-dwarf planets remain exposed to intense stellar activity, leading to further erosion of secondary atmospheres if not replenished through volcanic outgassing or other mechanisms.
-Recent works~\citep[e.g.,][]{ParkCoy2024,Luque2024} interpret JWST eclipse measurements as growing evidence for the absence of substantial atmospheres on M-dwarf rocky exoplanets.
+Recent works~\citep[e.g.,][]{Coy2024,Luque2024} interpret JWST eclipse measurements as growing evidence for the absence of substantial atmospheres on M-dwarf rocky exoplanets.
 If these planets indeed lack atmospheres, their habitability would be primarily dictated by the rate of atmospheric escape~\citep[e.g.,][]{Owen2020}.
@@
 This underscores the need for future observations and modeling efforts to constrain atmospheric retention and escape processes more robustly.}
-...
-
 %For instance, atmospheric escape processes may play a critical role not only in maintaining a planet's atmosphere but also in regulating the planet's water inventory and the availability of volatile elements for prebiotic chemistry.
 
-\rev{While not modeled here, the above additional mechanisms may offer directions for future work.}
+\rev{Beyond atmospheric escape, internal heating from tidal forces or radioactive decay can extend or constrain the limits of planetary surface habitability~\citep[e.g.,][]{Barnes2013,Oosterloo2021}.
+Tidal effects are especially relevant for \gls{HZ} M~dwarf planets, where strong stellar interactions can drive internal heating, loss of water or an atmosphere, or runaway greenhouse conditions.
+Tidal locking may also create extreme climate zones, challenging habitability.}
+\rev{While not modeled here, the factors outlined above may offer directions for future work.}
 

### 2025-02-04 src/tex/ms.tex
added some discussion in 4.1 addressing Reviewers 1 comment number 2

@@
 This encompasses the probability of life both emerging and persisting to produce detectable biosignatures.}
+\rev{Due to our ignorance about its true value or even its order of magnitude, we draw $f_\mathrm{life}$ from a log-uniform prior probability distribution.}
 Let us consider the probability to detect a biosignature $P(L)$, and let our observable be the inferred past \gls{NUV} flux of the planet $F_\mathrm{NUV}$.
@@
 In Section~\ref{sec:met-semianalytical}, we computed the probability for true positive evidence for $H_\mathrm{1}$ and $H_\mathrm{null}$, respectively (Equations~\ref{eq:bayes_factor},~\ref{eq:bayes_factor2}).
-Figure~\ref{fig:semian_true_evidence} shows how these evidences are distributed for sample sizes \var{semian_Nsamp1} and \var{semian_Nsamp2}, and how likely we are to obtain strong evidence ($BF_{H_i, H_j}$ > 10).
-For $n = \var{semian_Nsamp1}$, strong true evidence for $H_\mathrm{1}$ ($H_\mathrm{null}$) can be expected in $\sim \SI{30}{\percent}$ ($\sim \SI{40}{\percent}$) of all random experiments.
+Figure~\ref{fig:semian_true_evidence} shows how these evidences are distributed for sample sizes \var{semian_Nsamp1} and \var{semian_Nsamp2}, and how likely we are to obtain strong evidence ($BF_{H_i, H_j}$ > 10) \rev{in the agnostic case where we draw $f_\mathrm{life}$ from a log-uniform distribution}.
+For $n = \var{semian_Nsamp1}$, strong true evidence for $H_\mathrm{1}$ ($H_\mathrm{null}$) can be \rev{expected in $\sim \SI{3}{\percent}$ ($\sim \SI{6}{\percent}$) of} all random experiments.
 In the majority of cases, the outcome of the survey will be inconclusive.
@@
 \end{figure*}
-The situation improves with larger samples: for $n = \var{semian_Nsamp2}$, \SI{80}{\percent} of random samples permit conclusive inference (strong true evidence) under either $H_\mathrm{1}$ or $H_\mathrm{null}$.
+The situation improves with larger samples: \rev{for $n = \var{semian_Nsamp2}$, \SI{14}{\percent} (\SI{16}{\percent}) of} random samples permit conclusive inference (strong true evidence) \rev{under $H_\mathrm{1}$ ($H_\mathrm{null}$).}
 
@@
 
-So far, we have assumed random, uniform distributions of $f_\mathrm{life}$, $F_\mathrm{NUV, min}$, and $F_\mathrm{NUV}$.
+So far, we have \rev{drawn random values from uniform distributions for} $F_\mathrm{NUV, min}$, and $F_\mathrm{NUV}$\rev{, and from a log-uniform distribution for $f_\mathrm{life}$}.
 % Actually it is more than just NUV, it is whatever index that integrates NUV and HZ requiremetns into a single scalar variable for which a threshold exist
@@
 Left: Sampling distribution for different selectivity parameters $s$.
-Right: Resulting \mbox{P(true strong evidence)}, where $f_\mathrm{life}$ and $F_\mathrm{NUV, min}$ are randomly drawn from a uniform distribution.
+Right: Resulting \mbox{P(true strong evidence)}, where $f_\mathrm{life}$ and $F_\mathrm{NUV, min}$ are randomly drawn from \rev{log-uniform and uniform distributions, respectively}.
 Sampling more extreme values of $F_\mathrm{NUV}$ is more likely to yield strong evidence.}
@@
 
+\appendix
+\section{Alternative priors for the abiogenesis rate}
+\label{sec:appendix-flife-prior}
+\rev{In the main text, we used a log-uniform prior for the probability of life emerging and persisting $f_\mathrm{life}$ in order to reflect our ignorance about this parameter and its order of magnitude.
+We consider this a reasonable choice, as it is agnostic about the scale of $f_\mathrm{life}$ and assigns equal prior probability to all orders of magnitude.
+Nevertheless, we repeated our semi-analytical analysis for the more optimistic choice of a prior that is \textit{uniform} in $f_\mathrm{life}$.
+Figure~\ref{fig:semian_true_evidence_uniform} shows the resulting probabilities of obtaining strong evidence for or against the UV~Threshold Hypothesis under this prior.
+While the resulting trends are qualitatively similar, the uniform prior yields overall higher probabilities of such a conclusive test for all sample sizes.
+}
+
+\begin{figure*}
+    \begin{centering}
+        \includegraphics[width=\hsize]{figures/semian_true_evidence_uniform}
+        \caption{\rev{As Figure~\ref{fig:semian_true_evidence}, but using a uniform prior for the abiogenesis rate $f_\mathrm{life}$.
+        The probabilities of obtaining strong evidence for or against the UV~Threshold Hypothesis are overall higher than for the log-uniform prior.
+        }
+        }
+        \label{fig:semian_true_evidence_uniform}
+    \end{centering}
+\end{figure*}
+
 %! suppress = MissingBibliographystyle

### 2025-02-05 src/tex/ms.tex
Introducing bayes factor in methods

--- 2025-02-04 src/tex/ms.tex
@@
 %First, we apply a Bayes Factor Design Analysis~\citep{Schoenbrodt2018}
-We first assessed the expected probabilities of obtaining true negative or true positive evidence for the UV Threshold Hypothesis ($H_1$) above, as well as the probability for misleading or inconclusive evidence, under idealized conditions.
-This serves as a first-order estimate of the information content of a survey, before we take into account the effects of exoplanet demographics, sample selection, and survey strategy.
+\rev{In order to constrain the evidence yield of potential future surveys aimed at evaluating the} UV Threshold Hypothesis ($H_1$) above, \rev{we first derive a simplified analytical approach.
+This approach ignores the effects of exoplanet demographics and sample selection to focus on the qualitative effects of sample size and the unknown probablity of abiogenesis on the evidence yield of future surveys.
+To do so, we need to frame hypotheses $H_\mathrm{null}$ and $H_1$ as probabilistic models that determine the probability distributions of outcomes in terms of samples of exoplanets for which biosignatures are detected or not.}
+%We first assessed the expected probabilities of obtaining true negative or true positive evidence for the UV Threshold Hypothesis ($H_1$) above, as well as the probability for misleading or inconclusive evidence, under idealized conditions.
+%This serves as a first-order estimate of the information content of a survey, before we take into account the effects of exoplanet demographics, sample selection, and survey strategy.
 
@@
 
-Defining a sample of size $n$ as $X=\{F_\mathrm{NUV, i},L_i\}_{i \in [1,n]}$ where $L_i$ is equal to 1 if life is detected and 0 otherwise, we can calculate the evidence for hypothesis $H_i$ being true against $H_j$ through the Bayes factor
-\begin{equation}
-BF_{H_i,H_j} = \frac{P(X|H_i)}{P(X|H_j)},
-\end{equation}
-with $P(X|H_i)$ and $P(X|H_j)$ likelihoods of obtaining the sample $X$ under either hypothesis.
-
+\rev{Next, we determine the probability distribution of sample outcomes, or likelihood of each hypothesis.}
 Let $Y=\sum_i^n L_i$ be the random variable counting the number of positive life detections in a sample of size $n$.
@@
 \begin{equation}
+    \label{eq:semian:likelihoodHnull}
 P(Y=k|H_\mathrm{null}) = \binom{n}{k}f_\mathrm{life}^k(1-f_\mathrm{life})^{n-k}.
@@
 \end{equation}
+
+\rev{Then, we turn to a metric which measures the information gain of sampling: the Bayes factor \cite{Jeffreys1939}.
+The posterior evidence for the null hypothesis $H_{\mathrm{null}}$ is given by the Bayes' theorem:
+\begin{equation}
+\label{eq:bayesHnull}
+    P(H_{\mathrm{null}}|Y) = \frac{P(Y|H_{\mathrm{null}})P(H_{\mathrm{null}})}{P(Y)}
+\end{equation}
+and the posterior evidence for the NUV hypothesis:
+\begin{equation}
+    \label{eq:bayesHone}
+        P(H_1|Y) = \frac{P(Y|H_1)P(H_1)}{P(Y)}~.
+\end{equation}
+Computing the distributions given by equations \ref{eq:bayesHnull} and \ref{eq:bayesHone} requires the distributions of likelihoods, given in equations \ref{eq:semian:likelihoodHnull} and \ref{eq:semian:likelihoodH1}, as well as the prior distributions for each hypothesis ($P(H_{\mathrm{null}})$ and $P(H_{1})$) and for the sample ($P(H_{1})$).
+}
+\rev{
+Appart from the caveat that the likelihood for $H_{1}$ cannot be expressed analytically due to being conditioned by n_{\lambda} (equation \ref{eq:semian:likelihoodH1}), computing posterior probabilities is challenging due to the unknown priors.
+}
+
+Defining a sample of size $n$ as $X=\{F_\mathrm{NUV, i},L_i\}_{i \in [1,n]}$ where $L_i$ is equal to 1 if life is detected and 0 otherwise, we can calculate the evidence for hypothesis $H_i$ being true against $H_j$ through the Bayes factor
+\begin{equation}
+BF_{H_i,H_j} = \frac{P(X|H_i)}{P(X|H_j)},
+\end{equation}
+with $P(X|H_i)$ and $P(X|H_j)$ likelihoods of obtaining the sample $X$ under either hypothesis.
 

### 2025-02-05 src/tex/ms.tex
oops typo

--- 2025-02-05 src/tex/ms.tex
@@
 \rev{
-Appart from the caveat that the likelihood for $H_{1}$ cannot be expressed analytically due to being conditioned by n_{\lambda} (equation \ref{eq:semian:likelihoodH1}), computing posterior probabilities is challenging due to the unknown priors.
+Apart from the caveat that the likelihood for $H_{1}$ cannot be expressed analytically due to being conditioned by n_{\lambda} (equation \ref{eq:semian:likelihoodH1}), computing posterior probabilities is challenging due to the unknown priors.
 }

### 2025-02-05 src/tex/ms.tex
add statistical constraints to intro

--- 2025-02-05 src/tex/ms.tex
@@
 %First, we apply a Bayes Factor Design Analysis~\citep{Schoenbrodt2018}
-\rev{In order to constrain the evidence yield of potential future surveys aimed at evaluating the} UV Threshold Hypothesis ($H_1$) above, \rev{we first derive a simplified analytical approach.
-This approach ignores the effects of exoplanet demographics and sample selection to focus on the qualitative effects of sample size and the unknown probablity of abiogenesis on the evidence yield of future surveys.
-To do so, we need to frame hypotheses $H_\mathrm{null}$ and $H_1$ as probabilistic models that determine the probability distributions of outcomes in terms of samples of exoplanets for which biosignatures are detected or not.}
-%We first assessed the expected probabilities of obtaining true negative or true positive evidence for the UV Threshold Hypothesis ($H_1$) above, as well as the probability for misleading or inconclusive evidence, under idealized conditions.
-%This serves as a first-order estimate of the information content of a survey, before we take into account the effects of exoplanet demographics, sample selection, and survey strategy.
+We first assessed the expected probabilities of obtaining true negative or true positive evidence for the UV Threshold Hypothesis ($H_1$) above, as well as the probability for misleading or inconclusive evidence, under idealized conditions.
+This serves as a first-order estimate of the information content of a survey, before we take into account the effects of exoplanet demographics, sample selection, and survey strategy.
 
@@
 
-\rev{Next, we determine the probability distribution of sample outcomes, or likelihood of each hypothesis.}
+Defining a sample of size $n$ as $X=\{F_\mathrm{NUV, i},L_i\}_{i \in [1,n]}$ where $L_i$ is equal to 1 if life is detected and 0 otherwise, we can calculate the evidence for hypothesis $H_i$ being true against $H_j$ through the Bayes factor
+\begin{equation}
+BF_{H_i,H_j} = \frac{P(X|H_i)}{P(X|H_j)},
+\end{equation}
+with $P(X|H_i)$ and $P(X|H_j)$ likelihoods of obtaining the sample $X$ under either hypothesis.
+
 Let $Y=\sum_i^n L_i$ be the random variable counting the number of positive life detections in a sample of size $n$.
@@
 \begin{equation}
-    \label{eq:semian:likelihoodHnull}
 P(Y=k|H_\mathrm{null}) = \binom{n}{k}f_\mathrm{life}^k(1-f_\mathrm{life})^{n-k}.
@@
 \end{equation}
-
-\rev{Then, we turn to a metric which measures the information gain of sampling: the Bayes factor \cite{Jeffreys1939}.
-The posterior evidence for the null hypothesis $H_{\mathrm{null}}$ is given by the Bayes' theorem:
-\begin{equation}
-\label{eq:bayesHnull}
-    P(H_{\mathrm{null}}|Y) = \frac{P(Y|H_{\mathrm{null}})P(H_{\mathrm{null}})}{P(Y)}
-\end{equation}
-and the posterior evidence for the NUV hypothesis:
-\begin{equation}
-    \label{eq:bayesHone}
-        P(H_1|Y) = \frac{P(Y|H_1)P(H_1)}{P(Y)}~.
-\end{equation}
-Computing the distributions given by equations \ref{eq:bayesHnull} and \ref{eq:bayesHone} requires the distributions of likelihoods, given in equations \ref{eq:semian:likelihoodHnull} and \ref{eq:semian:likelihoodH1}, as well as the prior distributions for each hypothesis ($P(H_{\mathrm{null}})$ and $P(H_{1})$) and for the sample ($P(H_{1})$).
-}
-\rev{
-Apart from the caveat that the likelihood for $H_{1}$ cannot be expressed analytically due to being conditioned by n_{\lambda} (equation \ref{eq:semian:likelihoodH1}), computing posterior probabilities is challenging due to the unknown priors.
-}
-
-Defining a sample of size $n$ as $X=\{F_\mathrm{NUV, i},L_i\}_{i \in [1,n]}$ where $L_i$ is equal to 1 if life is detected and 0 otherwise, we can calculate the evidence for hypothesis $H_i$ being true against $H_j$ through the Bayes factor
-\begin{equation}
-BF_{H_i,H_j} = \frac{P(X|H_i)}{P(X|H_j)},
-\end{equation}
-with $P(X|H_i)$ and $P(X|H_j)$ likelihoods of obtaining the sample $X$ under either hypothesis.
 
@@
 
-\rev{Our Bayes factor analysis quantifies how much evidence future data may provide in favor of a hypothesis against another one (the null hypothesis).
-Typical frequentist (p-value based) tests usually evaluate only the likelihood of the null hypothesis, thus requiring that the experiment be carefully crafted such that rejection of the null hypothesis can be interpreted in favor of a particular studied hypothesis (for instance that samples from coin tosses score a very low likelihood under the hypothesis of a fair coin may indicate that the coin is biased).
-In contrast, Bayes factor analyses such as ours introduce additional constraint to the testing by involving the likelihood of the tested alternate hypothesis (equation \ref{eq:bayes_factor}).
-Thus, under Bayes-factor analyses, evidence for the tested hypothesis is not only measured by how unlikely the null hypothesis is given some observation but rather by how much more likely than the null hypothesis is the alternate hypothesis.
-The most exhaustive form of hypothesis selection test would be performed by using the entire Bayes theorem (equation \ref{eq:BayesTheorem}), computing the \emph{posterior} probability of the tested hypothesis.
-Doing so enables to quantify the information gain from performing some observation (by comparing the posterior with the prior), and to perform inference with consideration for prior information.
-\begin{equation}
-    \label{eq:BayesTheorem}
-    P(H_i|X) = \frac{P(X|H_i)P(H_i)}{P(X)}
-\end{equation}
-However, implementing the Bayes theorem can prove difficult in practice, owing to the (usually) subjective nature of prior distributions.
-Two prior distributions are required for the computation of the posterior distribution: the prior distribution for the considered model and hypothesis, and the prior distribution for the observation.
-While the former requires that at least some amount of consensus is identifiable in the community of scientists that discuss this hypothesis, the latter also requires that either empirical distributions for the prior distribution of the observation be available or that the exhaustive list of mechanistic pathways that could potentially lead to this observation be known and tractable.
-Taking the ratio of the posterior evidence for two alternate hypotheses (i.e.
-of equation \ref{eq:BayesTheorem}) allows this problematic term, the prior probability for the observation, to cancel out (equation \ref{eq:postratio}).
-\begin{equation}
-    \label{eq:postratio}
-    \frac{P(H_1|X)}{P(H_0|X)} = \frac{P(X|H_0)P(H_0)}{P(X|H_0)P(H_0)}
-\end{equation}
-This trades off the opportunity to estimate the posterior probability of a specific hypothesis with the possiblity to narrow down the problem to two specific competing hypotheses.
-Could there be factors other than the past UV flux that would determine the occurrence rate of biosignatures in terrestrial exoplanets?
-Surely, there is a vast array of possible mechanisms that could be considered which could shape the prior probability of an observed sample of biosignatures $X$.
-Truncating the ratio of prior probabilities from equation \ref{eq:postratio} results in the Bayes factor (equation \ref{eq:bayes_factor}), which we used in our semi-analytical analysis (section \ref{sec:results-semianalytical}).
-While the Bayes ratio only paints part of the picture of what a full inference pipeline would look line in a future survey aimed at testing origin of life hypotheses from the observation of biosignatures, it captures an essential part of the yield of a particular strategy: by how much would these observations tip the scale of model selection relative to prior expectations.}
-
-\rev{Our Bayes factor analysis quantifies how much evidence future data may provide in favor of a hypothesis against another one (the null hypothesis).
-Typical frequentist (p-value based) tests usually evaluate only the likelihood of the null hypothesis, thus requiring that the experiment be carefully crafted such that rejection of the null hypothesis can be interpreted in favor of a particular studied hypothesis (for instance that samples from coin tosses score a very low likelihood under the hypothesis of a fair coin may indicate that the coin is biased).
-In contrast, Bayes factor analyses such as ours introduce additional constraint to the testing by involving the likelihood of the tested alternate hypothesis (equation \ref{eq:bayes_factor}).
-Thus, under Bayes-factor analyses, evidence for the tested hypothesis is not only measured by how unlikely the null hypothesis is given some observation but rather by how much more likely than the null hypothesis is the alternate hypothesis.
-The most exhaustive form of hypothesis selection test would be performed by using the entire Bayes theorem (equation \ref{eq:BayesTheorem}), computing the \emph{posterior} probability of the tested hypothesis.
-Doing so enables to quantify the information gain from performing some observation (by comparing the posterior with the prior), and to perform inference with consideration for prior information.
-\begin{equation}
-    \label{eq:BayesTheorem}
-    P(H_i|X) = \frac{P(X|H_i)P(H_i)}{P(X)}
-\end{equation}
-However, implementing the Bayes theorem can prove difficult in practice, owing to the (usually) subjective nature of prior distributions.
-Two prior distributions are required for the computation of the posterior distribution: the prior distribution for the considered model and hypothesis, and the prior distribution for the observation.
-While the former requires that at least some amount of consensus is identifiable in the community of scientists that discuss this hypothesis, the latter also requires that either empirical distributions for the prior distribution of the observation be available or that the exhaustive list of mechanistic pathways that could potentially lead to this observation be known and tractable.
-Taking the ratio of the posterior evidence for two alternate hypotheses (i.e.
-of equation \ref{eq:BayesTheorem}) allows this problematic term, the prior probability for the observation, to cancel out (equation \ref{eq:postratio}).
-\begin{equation}
-    \label{eq:postratio}
-    \frac{P(H_1|X)}{P(H_0|X)} = \frac{P(X|H_0)P(H_0)}{P(X|H_0)P(H_0)}
-\end{equation}
-This trades off the opportunity to estimate the posterior probability of a specific hypothesis with the possiblity to narrow down the problem to two specific competing hypotheses.
-Could there be factors other than the past UV flux that would determine the occurrence rate of biosignatures in terrestrial exoplanets?
-Surely, there is a vast array of possible mechanisms that could be considered which could shape the prior probability of an observed sample of biosignatures $X$.
-Truncating the ratio of prior probabilities from equation \ref{eq:postratio} results in the Bayes factor (equation \ref{eq:bayes_factor}), which we used in our semi-analytical analysis (section \ref{sec:results-semianalytical}).
-While the Bayes ratio only paints part of the picture of what a full inference pipeline would look line in a future survey aimed at testing origin of life hypotheses from the observation of biosignatures, it captures an essential part of the yield of a particular strategy: by how much would these observations tip the scale of model selection relative to prior expectations.}
-
 %We stress that a future observer will likely be able to leverage standard frequentist testing methods which may be more likely to yield conclusive results (whether in favor or in disfavor of $H_1$).

### 2025-02-05 src/tex/ms.tex
finished editing methods 2

--- 2025-02-05 src/tex/ms.tex
@@
 Apart from the caveat that the likelihood for $H_{1}$ cannot be expressed analytically due to being conditioned by n_{\lambda} (equation \ref{eq:semian:likelihoodH1}), computing posterior probabilities is challenging due to the unknown priors.
+In addition, we are not interested in constraining the posterior distributions for each hypothesis as we are in measuring the information gain of performing a certain observation, regardless of the priors that future observers may adopt.
+The ratio of posterior probabilities (equations \ref{eq:bayesHone} and \ref{eq:bayesHnull}) is the product of the ratio of the prior probabilities and likelihoods of each hypotheses (equation \ref{eq:postratio}).
+\begin{equation}
+    \label{eq:postratio}
+    \frac{P(H_1|X)}{P(H_{\mathrm{null}}|X)} = \frac{P(X|H_{\mathrm{null}})P(H_{\mathrm{null}})}{P(X|H_{\mathrm{null}})P(H_{\mathrm{null}})}~.
+\end{equation}
+By truncating the prior ratio out of equation \ref{eq:postratio} (and substituting likelihoods for their expressions; equations \ref{eq:semian:likelihoodH1}, \ref{eq:semian:likelihoodHnull}), one obtains the Bayes factor \cite{Jeffreys1939}:
 }
-
-Defining a sample of size $n$ as $X=\{F_\mathrm{NUV, i},L_i\}_{i \in [1,n]}$ where $L_i$ is equal to 1 if life is detected and 0 otherwise, we can calculate the evidence for hypothesis $H_i$ being true against $H_j$ through the Bayes factor
-\begin{equation}
-BF_{H_i,H_j} = \frac{P(X|H_i)}{P(X|H_j)},
-\end{equation}
-with $P(X|H_i)$ and $P(X|H_j)$ likelihoods of obtaining the sample $X$ under either hypothesis.
-
-Hence,
-\begin{subequations}
+%\begin{subequations}
 \begin{equation}\label{eq:bayes_factor}
@@
 \end{equation}
-and
-\begin{equation}\label{eq:bayes_factor2}
+\rev{and conversely the Bayes ratio measuring evidence for $H_{\mathrm{null}}$ against $H_1$:}
+\begin{equation}
+    \label{eq:bayes_factor2}
 BF_{H_\mathrm{null},H_1} = \frac{P(Y=k|H_\mathrm{null})}{P(Y=k|H_1)} = \frac{\binom{n}{k}}{\binom{n_\lambda}{k}}(1-f_\mathrm{life})^{n-n_{\lambda}}.
 \end{equation}
-\end{subequations}
+%\end{subequations}
+\rev{
+By focusing on the Bayes ratio, we measure how much the observation of a sample $Y$ tips the scale of evidence for or against hypothesis relative to prior expectations, allowing us to focus on the information yield as a function of sample size.}
 

### 2025-02-05 src/tex/ms.tex
finished editing methods 2

@@
 This encompasses the probability of life both emerging and persisting to produce detectable biosignatures.}
-\rev{Due to our ignorance about its true value or even its order of magnitude, we draw $f_\mathrm{life}$ from a log-uniform prior probability distribution.}
 Let us consider the probability to detect a biosignature $P(L)$, and let our observable be the inferred past \gls{NUV} flux of the planet $F_\mathrm{NUV}$.
@@
 In Section~\ref{sec:met-semianalytical}, we computed the probability for true positive evidence for $H_\mathrm{1}$ and $H_\mathrm{null}$, respectively (Equations~\ref{eq:bayes_factor},~\ref{eq:bayes_factor2}).
-Figure~\ref{fig:semian_true_evidence} shows how these evidences are distributed for sample sizes \var{semian_Nsamp1} and \var{semian_Nsamp2}, and how likely we are to obtain strong evidence ($BF_{H_i, H_j}$ > 10) \rev{in the agnostic case where we draw $f_\mathrm{life}$ from a log-uniform distribution}.
-For $n = \var{semian_Nsamp1}$, strong true evidence for $H_\mathrm{1}$ ($H_\mathrm{null}$) can be \rev{expected in $\sim \SI{3}{\percent}$ ($\sim \SI{6}{\percent}$) of} all random experiments.
+Figure~\ref{fig:semian_true_evidence} shows how these evidences are distributed for sample sizes \var{semian_Nsamp1} and \var{semian_Nsamp2}, and how likely we are to obtain strong evidence ($BF_{H_i, H_j}$ > 10).
+For $n = \var{semian_Nsamp1}$, strong true evidence for $H_\mathrm{1}$ ($H_\mathrm{null}$) can be expected in $\sim \SI{30}{\percent}$ ($\sim \SI{40}{\percent}$) of all random experiments.
 In the majority of cases, the outcome of the survey will be inconclusive.
@@
 \end{figure*}
-The situation improves with larger samples: \rev{for $n = \var{semian_Nsamp2}$, \SI{14}{\percent} (\SI{16}{\percent}) of} random samples permit conclusive inference (strong true evidence) \rev{under $H_\mathrm{1}$ ($H_\mathrm{null}$).}
+The situation improves with larger samples: for $n = \var{semian_Nsamp2}$, \SI{80}{\percent} of random samples permit conclusive inference (strong true evidence) under either $H_\mathrm{1}$ or $H_\mathrm{null}$.
 
@@
 
-So far, we have \rev{drawn random values from uniform distributions for} $F_\mathrm{NUV, min}$, and $F_\mathrm{NUV}$\rev{, and from a log-uniform distribution for $f_\mathrm{life}$}.
+So far, we have assumed random, uniform distributions of $f_\mathrm{life}$, $F_\mathrm{NUV, min}$, and $F_\mathrm{NUV}$.
 % Actually it is more than just NUV, it is whatever index that integrates NUV and HZ requiremetns into a single scalar variable for which a threshold exist
@@
 Left: Sampling distribution for different selectivity parameters $s$.
-Right: Resulting \mbox{P(true strong evidence)}, where $f_\mathrm{life}$ and $F_\mathrm{NUV, min}$ are randomly drawn from \rev{log-uniform and uniform distributions, respectively}.
+Right: Resulting \mbox{P(true strong evidence)}, where $f_\mathrm{life}$ and $F_\mathrm{NUV, min}$ are randomly drawn from a uniform distribution.
 Sampling more extreme values of $F_\mathrm{NUV}$ is more likely to yield strong evidence.}
@@
 
-\appendix
-\section{Alternative priors for the abiogenesis rate}
-\label{sec:appendix-flife-prior}
-\rev{In the main text, we used a log-uniform prior for the probability of life emerging and persisting $f_\mathrm{life}$ in order to reflect our ignorance about this parameter and its order of magnitude.
-We consider this a reasonable choice, as it is agnostic about the scale of $f_\mathrm{life}$ and assigns equal prior probability to all orders of magnitude.
-Nevertheless, we repeated our semi-analytical analysis for the more optimistic choice of a prior that is \textit{uniform} in $f_\mathrm{life}$.
-Figure~\ref{fig:semian_true_evidence_uniform} shows the resulting probabilities of obtaining strong evidence for or against the UV~Threshold Hypothesis under this prior.
-While the resulting trends are qualitatively similar, the uniform prior yields overall higher probabilities of such a conclusive test for all sample sizes.
-}
-
-\begin{figure*}
-    \begin{centering}
-        \includegraphics[width=\hsize]{figures/semian_true_evidence_uniform}
-        \caption{\rev{As Figure~\ref{fig:semian_true_evidence}, but using a uniform prior for the abiogenesis rate $f_\mathrm{life}$.
-        The probabilities of obtaining strong evidence for or against the UV~Threshold Hypothesis are overall higher than for the log-uniform prior.
-        }
-        }
-        \label{fig:semian_true_evidence_uniform}
-    \end{centering}
-\end{figure*}
-
 %! suppress = MissingBibliographystyle

### 2025-02-05 src/tex/ms.tex
edited discusssion (and polished methods)

@@
 In addition, we are not interested in constraining the posterior distributions for each hypothesis as we are in measuring the information gain of performing a certain observation, regardless of the priors that future observers may adopt.
+Previous research aimed at prospectively quantifying the information yield of future exoplanet surveys focused on the Bayes factor, rather than on posterior probabilities \cite{affholder2025interior}.
 The ratio of posterior probabilities (equations \ref{eq:bayesHone} and \ref{eq:bayesHnull}) is the product of the ratio of the prior probabilities and likelihoods of each hypotheses (equation \ref{eq:postratio}).
@@
     \label{eq:postratio}
-    \frac{P(H_1|X)}{P(H_{\mathrm{null}}|X)} = \frac{P(X|H_{\mathrm{null}})P(H_{\mathrm{null}})}{P(X|H_{\mathrm{null}})P(H_{\mathrm{null}})}~.
+    \frac{P(H_1|Y)}{P(H_{\mathrm{null}}|Y)} = \frac{P(Y|H_{\mathrm{null}})P(H_{\mathrm{null}})}{P(Y|H_{\mathrm{null}})P(H_{\mathrm{null}})}~.
 \end{equation}
@@
 
-\rev{Our Bayes factor analysis quantifies how much evidence future data may provide in favor of a hypothesis against another one (the null hypothesis).
-Typical frequentist (p-value based) tests usually evaluate only the likelihood of the null hypothesis, thus requiring that the experiment be carefully crafted such that rejection of the null hypothesis can be interpreted in favor of a particular studied hypothesis (for instance that samples from coin tosses score a very low likelihood under the hypothesis of a fair coin may indicate that the coin is biased).
-In contrast, Bayes factor analyses such as ours introduce additional constraint to the testing by involving the likelihood of the tested alternate hypothesis (equation \ref{eq:bayes_factor}).
-Thus, under Bayes-factor analyses, evidence for the tested hypothesis is not only measured by how unlikely the null hypothesis is given some observation but rather by how much more likely than the null hypothesis is the alternate hypothesis.
-The most exhaustive form of hypothesis selection test would be performed by using the entire Bayes theorem (equation \ref{eq:BayesTheorem}), computing the \emph{posterior} probability of the tested hypothesis.
-Doing so enables to quantify the information gain from performing some observation (by comparing the posterior with the prior), and to perform inference with consideration for prior information.
-\begin{equation}
-    \label{eq:BayesTheorem}
-    P(H_i|X) = \frac{P(X|H_i)P(H_i)}{P(X)}
-\end{equation}
-However, implementing the Bayes theorem can prove difficult in practice, owing to the (usually) subjective nature of prior distributions.
-Two prior distributions are required for the computation of the posterior distribution: the prior distribution for the considered model and hypothesis, and the prior distribution for the observation.
-While the former requires that at least some amount of consensus is identifiable in the community of scientists that discuss this hypothesis, the latter also requires that either empirical distributions for the prior distribution of the observation be available or that the exhaustive list of mechanistic pathways that could potentially lead to this observation be known and tractable.
-Taking the ratio of the posterior evidence for two alternate hypotheses (i.e.
-of equation \ref{eq:BayesTheorem}) allows this problematic term, the prior probability for the observation, to cancel out (equation \ref{eq:postratio}).
-\begin{equation}
-    \label{eq:postratio}
-    \frac{P(H_1|X)}{P(H_0|X)} = \frac{P(X|H_0)P(H_0)}{P(X|H_0)P(H_0)}
-\end{equation}
-This trades off the opportunity to estimate the posterior probability of a specific hypothesis with the possiblity to narrow down the problem to two specific competing hypotheses.
+\rev{
+Our semi-analytical framework focuses on the Bayes ratio as a measure of the evidence that a future observation may bring relative to prior expectations.
+This is useful in order to specifically consider the information yield of an observation, but also trades off the opportunity to estimate the posterior probability of a specific hypothesis with the possibility to narrow down the problem to two specific competing hypotheses.
+Indeed, the posterior probability depends on the prior expectation for a sample $Y$ (see equation \ref{eq:bayesHone}), which may be approximated as the sum of likelihoods of the modeled hypotheses ($P(Y)=P(Y|H_1)P(H_1)+P(Y|H_{\mathrm{null}})P(H_{\mathrm{null}})$) but could in principle depend on any number of unknown processes.
 Could there be factors other than the past UV flux that would determine the occurrence rate of biosignatures in terrestrial exoplanets?
-Surely, there is a vast array of possible mechanisms that could be considered which could shape the prior probability of an observed sample of biosignatures $X$.
-Truncating the ratio of prior probabilities from equation \ref{eq:postratio} results in the Bayes factor (equation \ref{eq:bayes_factor}), which we used in our semi-analytical analysis (section \ref{sec:results-semianalytical}).
-While the Bayes ratio only paints part of the picture of what a full inference pipeline would look line in a future survey aimed at testing origin of life hypotheses from the observation of biosignatures, it captures an essential part of the yield of a particular strategy: by how much would these observations tip the scale of model selection relative to prior expectations.}
+Surely, there is a vast array of possible mechanisms that could be considered which could shape the prior probability of an observed sample of biosignatures $Y$.
+Thus, while the Bayes factor is an essential part of the inference process, and the one relevant to prospective analyses of future evidence yields, future attempts at inference may aim for estimating posterior probabilities rather than evidence ratios.
+Constraining the prior distribution for the probability that life requires exposure to a certain NUV threshold to emerge is beyond the scope of this study, but remains a question that future work will need to address in order to evaluate fully this hypothesis in the prospect of future observation of biosignatures.}
 

### 2025-02-05 src/tex/ms.tex
fix latex issue

@@
 \rev{
-Apart from the caveat that the likelihood for $H_{1}$ cannot be expressed analytically due to being conditioned by n_{\lambda} (equation \ref{eq:semian:likelihoodH1}), computing posterior probabilities is challenging due to the unknown priors.
+Apart from the caveat that the likelihood for $H_{1}$ cannot be expressed analytically due to being conditioned by $n_{\lambda}$ (equation \ref{eq:semian:likelihoodH1}), computing posterior probabilities is challenging due to the unknown priors.
 In addition, we are not interested in constraining the posterior distributions for each hypothesis as we are in measuring the information gain of performing a certain observation, regardless of the priors that future observers may adopt.

### 2025-02-06 src/tex/ms.tex
add an appendix for an optimistic choice of f_life

@@
 
-\rev{Then, we turn to a metric which measures the information gain of sampling: the Bayes factor \cite{Jeffreys1939}.
-The posterior evidence for the null hypothesis $H_{\mathrm{null}}$ is given by the Bayes' theorem:
-\begin{equation}
-\label{eq:bayesHnull}
-    P(H_{\mathrm{null}}|Y) = \frac{P(Y|H_{\mathrm{null}})P(H_{\mathrm{null}})}{P(Y)}
-\end{equation}
-and the posterior evidence for the NUV hypothesis:
-\begin{equation}
-    \label{eq:bayesHone}
-        P(H_1|Y) = \frac{P(Y|H_1)P(H_1)}{P(Y)}~.
-\end{equation}
-Computing the distributions given by equations \ref{eq:bayesHnull} and \ref{eq:bayesHone} requires the distributions of likelihoods, given in equations \ref{eq:semian:likelihoodHnull} and \ref{eq:semian:likelihoodH1}, as well as the prior distributions for each hypothesis ($P(H_{\mathrm{null}})$ and $P(H_{1})$) and for the sample ($P(H_{1})$).
-}
-\rev{
-Apart from the caveat that the likelihood for $H_{1}$ cannot be expressed analytically due to being conditioned by $n_{\lambda}$ (equation \ref{eq:semian:likelihoodH1}), computing posterior probabilities is challenging due to the unknown priors.
-In addition, we are not interested in constraining the posterior distributions for each hypothesis as we are in measuring the information gain of performing a certain observation, regardless of the priors that future observers may adopt.
-Previous research aimed at prospectively quantifying the information yield of future exoplanet surveys focused on the Bayes factor, rather than on posterior probabilities \cite{affholder2025interior}.
-The ratio of posterior probabilities (equations \ref{eq:bayesHone} and \ref{eq:bayesHnull}) is the product of the ratio of the prior probabilities and likelihoods of each hypotheses (equation \ref{eq:postratio}).
-\begin{equation}
-    \label{eq:postratio}
-    \frac{P(H_1|Y)}{P(H_{\mathrm{null}}|Y)} = \frac{P(Y|H_{\mathrm{null}})P(H_{\mathrm{null}})}{P(Y|H_{\mathrm{null}})P(H_{\mathrm{null}})}~.
-\end{equation}
-By truncating the prior ratio out of equation \ref{eq:postratio} (and substituting likelihoods for their expressions; equations \ref{eq:semian:likelihoodH1}, \ref{eq:semian:likelihoodHnull}), one obtains the Bayes factor \cite{Jeffreys1939}:
-}
-%\begin{subequations}
+Hence,
+\begin{subequations}
 \begin{equation}\label{eq:bayes_factor}
@@
 \end{equation}
-\rev{and conversely the Bayes ratio measuring evidence for $H_{\mathrm{null}}$ against $H_1$:}
-\begin{equation}
-    \label{eq:bayes_factor2}
+and
+\begin{equation}\label{eq:bayes_factor2}
 BF_{H_\mathrm{null},H_1} = \frac{P(Y=k|H_\mathrm{null})}{P(Y=k|H_1)} = \frac{\binom{n}{k}}{\binom{n_\lambda}{k}}(1-f_\mathrm{life})^{n-n_{\lambda}}.
 \end{equation}
-%\end{subequations}
-\rev{
-By focusing on the Bayes ratio, we measure how much the observation of a sample $Y$ tips the scale of evidence for or against hypothesis relative to prior expectations, allowing us to focus on the information yield as a function of sample size.}
+\end{subequations}
 
@@
 
-\rev{
-Our semi-analytical framework focuses on the Bayes ratio as a measure of the evidence that a future observation may bring relative to prior expectations.
-This is useful in order to specifically consider the information yield of an observation, but also trades off the opportunity to estimate the posterior probability of a specific hypothesis with the possibility to narrow down the problem to two specific competing hypotheses.
-Indeed, the posterior probability depends on the prior expectation for a sample $Y$ (see equation \ref{eq:bayesHone}), which may be approximated as the sum of likelihoods of the modeled hypotheses ($P(Y)=P(Y|H_1)P(H_1)+P(Y|H_{\mathrm{null}})P(H_{\mathrm{null}})$) but could in principle depend on any number of unknown processes.
-Could there be factors other than the past UV flux that would determine the occurrence rate of biosignatures in terrestrial exoplanets?
-Surely, there is a vast array of possible mechanisms that could be considered which could shape the prior probability of an observed sample of biosignatures $Y$.
-Thus, while the Bayes factor is an essential part of the inference process, and the one relevant to prospective analyses of future evidence yields, future attempts at inference may aim for estimating posterior probabilities rather than evidence ratios.
-Constraining the prior distribution for the probability that life requires exposure to a certain NUV threshold to emerge is beyond the scope of this study, but remains a question that future work will need to address in order to evaluate fully this hypothesis in the prospect of future observation of biosignatures.}
-
-\rev{Our Bayes factor analysis quantifies how much evidence future data may provide in favor of a hypothesis against another one (the null hypothesis).
-Typical frequentist (p-value based) tests usually evaluate only the likelihood of the null hypothesis, thus requiring that the experiment be carefully crafted such that rejection of the null hypothesis can be interpreted in favor of a particular studied hypothesis (for instance that samples from coin tosses score a very low likelihood under the hypothesis of a fair coin may indicate that the coin is biased).
-In contrast, Bayes factor analyses such as ours introduce additional constraint to the testing by involving the likelihood of the tested alternate hypothesis (equation \ref{eq:bayes_factor}).
-Thus, under Bayes-factor analyses, evidence for the tested hypothesis is not only measured by how unlikely the null hypothesis is given some observation but rather by how much more likely than the null hypothesis is the alternate hypothesis.
-The most exhaustive form of hypothesis selection test would be performed by using the entire Bayes theorem (equation \ref{eq:BayesTheorem}), computing the \emph{posterior} probability of the tested hypothesis.
-Doing so enables to quantify the information gain from performing some observation (by comparing the posterior with the prior), and to perform inference with consideration for prior information.
-\begin{equation}
-    \label{eq:BayesTheorem}
-    P(H_i|X) = \frac{P(X|H_i)P(H_i)}{P(X)}
-\end{equation}
-However, implementing the Bayes theorem can prove difficult in practice, owing to the (usually) subjective nature of prior distributions.
-Two prior distributions are required for the computation of the posterior distribution: the prior distribution for the considered model and hypothesis, and the prior distribution for the observation.
-While the former requires that at least some amount of consensus is identifiable in the community of scientists that discuss this hypothesis, the latter also requires that either empirical distributions for the prior distribution of the observation be available or that the exhaustive list of mechanistic pathways that could potentially lead to this observation be known and tractable.
-Taking the ratio of the posterior evidence for two alternate hypotheses (i.e.
-of equation \ref{eq:BayesTheorem}) allows this problematic term, the prior probability for the observation, to cancel out (equation \ref{eq:postratio}).
-\begin{equation}
-    \label{eq:postratio}
-    \frac{P(H_1|X)}{P(H_0|X)} = \frac{P(X|H_0)P(H_0)}{P(X|H_0)P(H_0)}
-\end{equation}
-This trades off the opportunity to estimate the posterior probability of a specific hypothesis with the possiblity to narrow down the problem to two specific competing hypotheses.
-Could there be factors other than the past UV flux that would determine the occurrence rate of biosignatures in terrestrial exoplanets?
-Surely, there is a vast array of possible mechanisms that could be considered which could shape the prior probability of an observed sample of biosignatures $X$.
-Truncating the ratio of prior probabilities from equation \ref{eq:postratio} results in the Bayes factor (equation \ref{eq:bayes_factor}), which we used in our semi-analytical analysis (section \ref{sec:results-semianalytical}).
-While the Bayes ratio only paints part of the picture of what a full inference pipeline would look line in a future survey aimed at testing origin of life hypotheses from the observation of biosignatures, it captures an essential part of the yield of a particular strategy: by how much would these observations tip the scale of model selection relative to prior expectations.}
-
 %We stress that a future observer will likely be able to leverage standard frequentist testing methods which may be more likely to yield conclusive results (whether in favor or in disfavor of $H_1$).

### 2025-02-07 src/scripts/semian_true_evidence.py
flexible n_bootstrap, print out probs

@@ -113,6 +113,9 @@ def plot_true_evidence(ax, nboot=None):
             ls=ls,
         )
 
+        print("N=", s, ":\nProb for strong evidence for H1: ",
+        np.sum(bfh1_kh1 > 10) / len(bfh1_kh1), "\nProb for strong evidence for H0: ", np.sum(bfh0_kh0 > 10) / len(bfh0_kh0))
+
     # ax.vlines(x=[1],ymin=0,ymax=0.9,linestyles='--',colors='grey')
     ax.vlines(x=[10], ymin=0, ymax=1., linestyles="-", lw=1, colors="gray")
     ax.text(10, .95, " strong", va="bottom", ha="left", c="k", alpha=1.0, fontsize=12)
@@ -140,6 +143,7 @@ def plot_true_evidence(ax, nboot=None):
         columnspacing=1.6,
     )
     ax.set_xscale("log")
+    ax.set_xlim(0.5, 1200)
     ax.set_ylim(-0.05, 1.05)
     # ax.set_ylabel("$P(\mathrm{evidence} > x)$")
     ax.set_ylabel("$P(BF_{H_i, H_j} > x$)")
@@ -335,7 +339,7 @@ def plot_selectivity(fig, ax, nboot=int(1e4)):
         arrowprops=dict(arrowstyle="simple", facecolor="C1", connectionstyle="arc3"),
     )
     ax.annotate(
-        xy=(25, 0.15),
+        xy=(45, 0.15),
         text=r"Sample more extreme $F_\mathrm{NUV}$",
         ha="left",
         va="center",
@@ -349,7 +353,7 @@ def plot_selectivity(fig, ax, nboot=int(1e4)):
         arrowprops=dict(arrowstyle="simple", facecolor="C1", connectionstyle="arc3"),
     )
     ax.annotate(
-        xy=(25, -0.15),
+        xy=(45, -0.15),
         text=r"Sample more intermediate $F_\mathrm{NUV}$",
         color="k",
         ha="left",
@@ -359,23 +363,35 @@ def plot_selectivity(fig, ax, nboot=int(1e4)):
 
 
 def main():
+
+
+    nboot = 100000
+    # nboot = 1000  # just for testing
+
+
     # 2-panel figure
     fig, axs = plt.subplots(1, 2, figsize=[13, 4.5])
-    axs[0] = plot_true_evidence(axs[0])
+    axs[0] = plot_true_evidence(axs[0], nboot=nboot)
[…]

### 2025-02-07 src/tex/ms.tex
trim Bayesian methods

@@
 
-Hence,
-\begin{subequations}
+\rev{
+Following a similar approach as in \cite{affholder2025interior}, we aim to quantify the information gain from our sampling procedure by computing the Bayes factors~\citep{Jeffreys1939}
+}
+%\begin{subequations}
 \begin{equation}\label{eq:bayes_factor}
-BF_{H_1,H_\mathrm{null}} = \frac{P(Y=k|H_1)}{P(Y=k|H_\mathrm{null})} = \frac{\binom{n_\lambda}{k}}{\binom{n}{k}}(1-f_\mathrm{life})^{n_{\lambda}-n},
+BF_{H_1,H_\mathrm{null}} = \frac{P(Y=k|H_1)}{P(Y=k|H_\mathrm{null})} = \frac{\binom{n_\lambda}{k}}{\binom{n}{k}}(1-f_\mathrm{life})^{n_{\lambda}-n}
 \end{equation}
 and
-\begin{equation}\label{eq:bayes_factor2}
+\begin{equation}
+    \label{eq:bayes_factor2}
 BF_{H_\mathrm{null},H_1} = \frac{P(Y=k|H_\mathrm{null})}{P(Y=k|H_1)} = \frac{\binom{n}{k}}{\binom{n_\lambda}{k}}(1-f_\mathrm{life})^{n-n_{\lambda}}.
 \end{equation}
-\end{subequations}
-
+%\end{subequations}
 Given a sample of planets, where for some of them we have convincing biosignature detections but remaining agnostic on $f_\mathrm{life}$: What evidence for $H_\mathrm{1}$ and $H_\mathrm{null}$ can we expect to get?
@@
 
+\rev{
+Our semi-analytical framework focuses on the Bayes ratio as a measure of the evidence that a future observation may bring relative to prior expectations.
+This is useful in order to specifically consider the information yield of an observation, but also trades off the opportunity to estimate the posterior probability of a specific hypothesis with the possibility to narrow down the problem to two specific competing hypotheses.
+Indeed, the posterior probability depends on the prior expectation for a sample $Y$ (see equation \ref{eq:bayesHone}), which may be approximated as the sum of likelihoods of the modeled hypotheses ($P(Y)=P(Y|H_1)P(H_1)+P(Y|H_{\mathrm{null}})P(H_{\mathrm{null}})$) but could in principle depend on any number of unknown processes.
+Could there be factors other than the past UV flux that would determine the occurrence rate of biosignatures in terrestrial exoplanets?
+Surely, there is a vast array of possible mechanisms that could be considered which could shape the prior probability of an observed sample of biosignatures $Y$.
+Thus, while the Bayes factor is an essential part of the inference process, and the one relevant to prospective analyses of future evidence yields, future attempts at inference may aim for estimating posterior probabilities rather than evidence ratios.
+Constraining the prior distribution for the probability that life requires exposure to a certain NUV threshold to emerge is beyond the scope of this study, but remains a question that future work will need to address in order to evaluate fully this hypothesis in the prospect of future observation of biosignatures.}
+
+\rev{Our Bayes factor analysis quantifies how much evidence future data may provide in favor of a hypothesis against another one (the null hypothesis).
+Typical frequentist (p-value based) tests usually evaluate only the likelihood of the null hypothesis, thus requiring that the experiment be carefully crafted such that rejection of the null hypothesis can be interpreted in favor of a particular studied hypothesis (for instance that samples from coin tosses score a very low likelihood under the hypothesis of a fair coin may indicate that the coin is biased).
+In contrast, Bayes factor analyses such as ours introduce additional constraint to the testing by involving the likelihood of the tested alternate hypothesis (equation \ref{eq:bayes_factor}).
+Thus, under Bayes-factor analyses, evidence for the tested hypothesis is not only measured by how unlikely the null hypothesis is given some observation but rather by how much more likely than the null hypothesis is the alternate hypothesis.
+The most exhaustive form of hypothesis selection test would be performed by using the entire Bayes theorem (equation \ref{eq:BayesTheorem}), computing the \emph{posterior} probability of the tested hypothesis.
+Doing so enables to quantify the information gain from performing some observation (by comparing the posterior with the prior), and to perform inference with consideration for prior information.
+\begin{equation}
+    \label{eq:BayesTheorem}
+    P(H_i|X) = \frac{P(X|H_i)P(H_i)}{P(X)}
+\end{equation}
+However, implementing the Bayes theorem can prove difficult in practice, owing to the (usually) subjective nature of prior distributions.
+Two prior distributions are required for the computation of the posterior distribution: the prior distribution for the considered model and hypothesis, and the prior distribution for the observation.
+While the former requires that at least some amount of consensus is identifiable in the community of scientists that discuss this hypothesis, the latter also requires that either empirical distributions for the prior distribution of the observation be available or that the exhaustive list of mechanistic pathways that could potentially lead to this observation be known and tractable.
+Taking the ratio of the posterior evidence for two alternate hypotheses (i.e.
+of equation \ref{eq:BayesTheorem}) allows this problematic term, the prior probability for the observation, to cancel out (equation \ref{eq:postratio}).
+\begin{equation}
+    \label{eq:postratio}
+    \frac{P(H_1|X)}{P(H_0|X)} = \frac{P(X|H_0)P(H_0)}{P(X|H_0)P(H_0)}
+\end{equation}
+This trades off the opportunity to estimate the posterior probability of a specific hypothesis with the possiblity to narrow down the problem to two specific competing hypotheses.
+Could there be factors other than the past UV flux that would determine the occurrence rate of biosignatures in terrestrial exoplanets?
+Surely, there is a vast array of possible mechanisms that could be considered which could shape the prior probability of an observed sample of biosignatures $X$.
+Truncating the ratio of prior probabilities from equation \ref{eq:postratio} results in the Bayes factor (equation \ref{eq:bayes_factor}), which we used in our semi-analytical analysis (section \ref{sec:results-semianalytical}).
+While the Bayes ratio only paints part of the picture of what a full inference pipeline would look line in a future survey aimed at testing origin of life hypotheses from the observation of biosignatures, it captures an essential part of the yield of a particular strategy: by how much would these observations tip the scale of model selection relative to prior expectations.}
+
 %We stress that a future observer will likely be able to leverage standard frequentist testing methods which may be more likely to yield conclusive results (whether in favor or in disfavor of $H_1$).

### 2025-02-07 src/tex/ms.tex
streamline discussion on Bayes factors and priors

--- 2025-02-07 src/tex/ms.tex
@@
 
+\rev{
+\subsubsection{Bayesian Evidence and the Influence of Priors}
+Our semi-analytical analysis (Section~\ref{sec:results-semianalytical}) employs Bayes factors to quantify the evidence in favor of or against the UV~Threshold Hypothesis that future observations may provide.
+%In contrast to frequentist p-value tests, which only assess the likelihood of the null hypothesis, this approach explicitly incorporates the likelihood of the alternative hypothesis.
+%However, Bayes factor analysis limits the scope to two competing hypotheses, and it does not estimate posterior probabilities.
+}
+
+\rev{
+Using the Bayes theorem
+%\begin{equation}
+%    \label{eq:BayesTheorem}
+%    P(H_i|X) = \frac{P(X|H_i)P(H_i)}{P(X)}
+%\end{equation}
+to estimate posterior probabilities would allow for a more complete assessment by quantifying information gain and integrating prior knowledge into hypothesis testing.
+In practice, implementation of a full Bayesian inference is often hindered by the subjectivity of prior distributions.
+Here, we thus chose to provide merely a general assessment of the potential evidence yield from future observations.
+In essence, we addressed the question: \textit{By how much does a particular observation tip the scale between the UV~Threshold Hypothesis and the null hypothesis?}
+}
+
 \section{Conclusions}

### 2025-02-12 src/tex/ms.tex
address comment #7

@@
 However, we have shown that the impacts from the combined requirements of the UV~Threshold Hypothesis on the fraction of inhabited planets in a sample are comparable in the non-transiting case.
-%The projected planet yield of most currently planned exoplanet mission concepts are likely insufficient to test the UV~Threshold Hypothesis:
 
@@
 Depending on the technical design, \life\ is expected to target 25–80 \glspl{EEC}~\citep{Kammerer2018,Quanz2022}, which could be just sufficient to constrain the UV~Threshold Hypothesis.
-One current exception is the Nautilus Space Observatory concept~\citep{Apai2019,Apai2022}.
+
+\rev{If the rate of life’s emergence and persistence $f_\mathrm{life}$ is at a \SI{1}{\percent} level or lower, no currently planned exoplanet mission concept has projected sample sizes sufficient to test the UV~Threshold Hypothesis.
+One } exception is the Nautilus Space Observatory concept~\citep{Apai2019,Apai2022}.
 Nautilus aims to characterize up to $\sim 1000$ \gls{EEC} via transmission spectroscopy, building on an innovative optical technology.

### 2025-02-12 src/tex/glossary.tex
use glossary entry for UV
fix NUV typo

@@
 \newglossaryentry{FIR}{name={FIR}, description={Far-Infrared}, first={far-infrared (FIR)}}
-\newglossaryentry{FUV}{name={FUV}, description={Far-Ultraviolet}, first={far-Ultraviolet (FUV)}}
-\newglossaryentry{NUV}{name={NUV}, description={Near-Ultraviolet}, first={near-Ultraviolet (NUV)}}
-\newglossaryentry{EUV}{name={EUV}, description={extreme-Ultraviolet}, first={extreme-Ultraviolet (EUV)}}
+\newglossaryentry{UV}{name={UV}, description={Ultraviolet}, first={ultraviolet (UV)}}
+\newglossaryentry{FUV}{name={FUV}, description={Far-Ultraviolet}, first={far-ultraviolet (FUV)}}
+\newglossaryentry{NUV}{name={NUV}, description={Near-Ultraviolet}, first={near-ultraviolet (NUV)}}
+\newglossaryentry{EUV}{name={EUV}, description={extreme-Ultraviolet}, first={extreme-ultraviolet (EUV)}}
 \newglossaryentry{SED}{name={SED}, description={Spectral energy distribution}, first={spectral energy distribution (SED)}, firstplural={spectral energy distributions (SEDs)}}

### 2025-02-12 src/tex/ms.tex
fix NUV typo

@@
 We follow a semi-analytical Bayesian analysis to estimate probabilities of obtaining strong evidence for or against this hypothesis.
-Under the UV Threshold Hypothesis ($H_1$), the probability of an exoplanet having detectable biosignatures is zero if the near-ultraviolet (\gls{NUV}) irradiation is less than the threshold, and it is equal to the (unknown) probability of \rev{life emerging and persisting}, $f_\mathrm{life}$ if \gls{NUV} exceeds the threshold for a sufficiently long period of time.
+Under the UV Threshold Hypothesis ($H_1$), the probability of an exoplanet having detectable biosignatures is zero if the \rev{\gls{NUV}} irradiation is less than the threshold, and it is equal to the (unknown) probability of \rev{life emerging and persisting}, $f_\mathrm{life}$ if \gls{NUV} exceeds the threshold for a sufficiently long period of time.
 Under the null hypothesis ($H_\mathrm{null}$), that probability simply is $f_\mathrm{life}$, that is, it does not correlate with the \gls{UV} flux.

### 2025-02-14 src/tex/ms.tex
fix a typo and a wrong reference

@@
 \subsubsection{Habitable zone occupancy and UV flux}\label{sec:met-hz}
-To construct a test the UV~Threshold Hypothesis, we required that life occurs only on planets with sufficient past \gls{UV} irradiation exceeding the origins of life threshold $F_\mathrm{NUV, min}$.
+To construct a test \rev{of} the UV~Threshold Hypothesis, we required that life occurs only on planets with sufficient past \gls{UV} irradiation exceeding the origins of life threshold $F_\mathrm{NUV, min}$.
 Further, we required this flux to have lasted for a minimum duration $\Delta T_\mathrm{min}$ to allow for a sufficient ``origins timescale''~\citep{Rimmer2023}.
@@
         \caption{Interpolated stellar luminosity evolution (left) and evolution of the \gls{NUV} flux in the \gls{HZ} (right) as a function of host star mass.
-        Scatter points show age and host star mass of the transiting planets in the synthetic planet sample; crosses denote the estimated \gls{NUV} values in \citet{Richey-Yowell2019}.
+        Scatter points show age and host star mass of the transiting planets in the synthetic planet sample; crosses denote the estimated \gls{NUV} values in \rev{\citet{Richey-Yowell2023}}.
 % with \gls{EEC} highlighted in green.

### 2025-02-14 src/scripts/inhabited-aafo-nuv_thresh.py
add capability for different abiogenesis timescales

@@ -47,7 +47,7 @@ def run_simulation(nuv_thresh, n_planets, star_type, seed, generator_kwargs):
     return fraction
 
 
-def simulate(transiting=True, debug=False):
+def simulate(transiting=True, debug=False, deltatmin=1.):
     if debug:
         n_planets = 100 # maximum number of planets to sample
         n_nuv_thresholds = 5
@@ -64,7 +64,7 @@ def simulate(transiting=True, debug=False):
 
     generator_kwargs = {
     'f_life' : 1.0,
-    'deltaT_min' : 1., # Myr. Smaller values seems to slightly enhance inhabited planets in the FGK sample.
+    'deltaT_min' : deltatmin, # Myr. Smaller values seems to slightly enhance inhabited planets in the FGK sample.
 
     # increase sample size
     'f_eta' : 15,
@@ -109,14 +109,18 @@ def simulate(transiting=True, debug=False):
 
     # Save results to disk
     lbl = 'transiting' if transiting else 'all'
-    with open(paths.data / f"frac-inhabited_fgk_{lbl}.dll", "wb") as f:
+    if deltatmin == 100.:
+        dt = "_dtmin100Myr"
+    else:
+        dt = ""
+    with open(paths.data / f"frac-inhabited_fgk_{lbl}{dt}.dll", "wb") as f:
         dill.dump((results_fgk, nuv_thresholds), f)
 
-    with open(paths.data / f"frac-inhabited_m_{lbl}.dll", "wb") as f:
+    with open(paths.data / f"frac-inhabited_m_{lbl}{dt}.dll", "wb") as f:
         dill.dump((results_m, nuv_thresholds), f)
 
 
-def plot_results():
+def plot_results(deltatmin=1.):
     fig, ax = plt.subplots(figsize=(7, 4.5))
 
     plotkwargs = {
@@ -124,11 +128,15 @@ def plot_results():
     'all' : {'marker' : 's', 'linestyle' : '--', 'alpha' : 0.6, 'linewidth' : 1.5},
     }
 
+    if deltatmin == 100.:
+        dt = "_dtmin100Myr"
+    else:
+        dt = ""
     for transiting, lbl in zip([True, False], ['transiting', 'all']):
-        with open(paths.data / f"frac-inhabited_fgk_{lbl}.dll", "rb") as f:
+        with open(paths.data / f"frac-inhabited_fgk_{lbl}{dt}.dll", "rb") as f:
[…]

### 2025-02-14 src/tex/ms.tex
discuss choice of abiogenesis timescale Delta T_min

@@
 We follow a semi-analytical Bayesian analysis to estimate probabilities of obtaining strong evidence for or against this hypothesis.
-Under the UV Threshold Hypothesis ($H_1$), the probability of an exoplanet having detectable biosignatures is zero if the \rev{\gls{NUV}} irradiation is less than the threshold, and it is equal to the (unknown) probability of \rev{life emerging and persisting}, $f_\mathrm{life}$ if \gls{NUV} exceeds the threshold for a sufficiently long period of time.
+Under the UV Threshold Hypothesis ($H_1$), the probability of an exoplanet having detectable biosignatures is zero if the \rev{\gls{NUV}} irradiation is less than the threshold, and it is equal to the (unknown) probability of \rev{life emerging and persisting}, $f_\mathrm{life}$, if \gls{NUV} exceeds the threshold for a sufficiently long period of time \rev{$\Delta T_\mathrm{min}$.}
 Under the null hypothesis ($H_\mathrm{null}$), that probability simply is $f_\mathrm{life}$, that is, it does not correlate with the \gls{UV} flux.
+\rev{
+While we adopt $\Delta T_\mathrm{min} = \SI{1}{\mega\year}$, we find that increasing this timescale to \SI{100}{\mega\year} reduces the fraction of inhabited planets under $H_1$ by less than 10 \%, suggesting that our results are not highly sensitive to $\Delta T_\mathrm{min}$.
+}
+
 \begin{figure*}

### 2025-02-14 src/tex/ms.tex
NUV -> \gls{NUV}

@@
 
-The goal of this work is to evaluate the potential of future exoplanet surveys to test the hypothesis that a minimum past NUV flux is required for abiogenesis.
+The goal of this work is to evaluate the potential of future exoplanet surveys to test the hypothesis that a minimum past \gls{NUV} flux is required for abiogenesis.
 We focus on one version of the UV Threshold Hypothesis, the so-called cyanosulfidic scenario, which has been refined to the point where the required threshold flux has been measured to be $F_\mathrm{NUV, min} = \SI{6.8\pm3.6 e10}{\photons\per\centi\meter\squared\per\second\per\nano\meter}$ integrated from \SIrange{200}{280}{\nano\meter} at the surface~\citep{Rimmer2018,Rimmer2021a,Rimmer2023,Ranjan2023a}.
@@
     \item \textbf{Hypothesis testing:} We evaluate the likelihood that a given survey would detect a specified demographic trend in the exoplanet population and estimate the precision with which the survey could constrain the parameters of that trend.
-    A common definition of the null hypothesis $H_\mathrm{null}$, which is also applied here, is that there is no relationship between the independent variable (here: maximum NUV flux) and the dependent variable (here: biosignature occurrence).
+    A common definition of the null hypothesis $H_\mathrm{null}$, which is also applied here, is that there is no relationship between the independent variable (here: maximum \gls{NUV} flux) and the dependent variable (here: biosignature occurrence).
     The alternative hypothesis $H_1$ proposes a specific relationship between the independent and dependent variables.
@@
 \rev{We ought to choose a statistical test that is sensitive to the UV~Threshold Hypothesis, and that could be realistically conducted in a future transit survey (which may include auxiliary information from ground-based observations, archived data, or models).
-Given the available types of data expected from such surveys, our test shall be non-parametric and compare two samples -- planets with and without biosignatures -- to assess whether they are drawn from the same underlying population in terms of their infered historic maximum NUV flux.
+Given the available types of data expected from such surveys, our test shall be non-parametric and compare two samples -- planets with and without biosignatures -- to assess whether they are drawn from the same underlying population in terms of their infered historic maximum \gls{NUV} flux.
 Common options include the Kolmogorov-Smirnov test, the Brunner-Munzel test~\citep{Brunner2000}, and the Mann-Whitney U test~\citep{Mann1947}.
@@
 To balance the trade-off between Type I and Type II error risks, we set the significance level to the widely adopted threshold $\alpha = 0.05$.
-%This choice ensures a reasonable trade-off between detecting a potential dependence of biosignatures on NUV flux while controlling for statistical noise.
+%This choice ensures a reasonable trade-off between detecting a potential dependence of biosignatures on \gls{NUV} flux while controlling for statistical noise.
 To quantify the diagnostic power of the survey, we conducted repeated randomized realizations and calculated the fraction of successful rejections of the null hypothesis, i.e., the statistical power.
@@
 So far, we have \rev{drawn random values from uniform distributions for} $F_\mathrm{NUV, min}$, and $F_\mathrm{NUV}$\rev{, and from a log-uniform distribution for $f_\mathrm{life}$}.
-% Actually it is more than just NUV, it is whatever index that integrates NUV and HZ requiremetns into a single scalar variable for which a threshold exist
+% Actually it is more than just NUV, it is whatever index that integrates \gls{NUV} and HZ requiremetns into a single scalar variable for which a threshold exist
 A high biosignature detection rate $f_\mathrm{life}$ increases the evidence (see Equation~\ref{eq:bayes_factor}) but a survey strategy cannot influence it.
@@
 %\end{figure*}
-This leads to a distribution of biosignature detections with detections increasingly occurring above a threshold inferred NUV flux.
+This leads to a distribution of biosignature detections with detections increasingly occurring above a threshold inferred \gls{NUV} flux.
 In this example case, the few biosignature detections in the FGK sample lead to a higher evidence %($dlnZ_\mathrm{FGK} = \var{dlnZ_FGK}$)
@@
 As we have shown, the statistical power of this test also depends on the distribution of past \gls{NUV} fluxes in the sample and on the threshold flux.
-Optimizing the survey to sample a wide range of NUV flux values, particularly at the extremes, can enhance the likelihood of obtaining strong evidence for or against the hypothesis.
+Optimizing the survey to sample a wide range of \gls{NUV} flux values, particularly at the extremes, can enhance the likelihood of obtaining strong evidence for or against the hypothesis.
 Intermediate values of the threshold \gls{NUV} flux are more likely to yield strong evidence than extreme values, as the dataset under the alternative hypothesis $H_1$ differs more from the null hypothesis in this case while still being sufficiently populated.
@@
 While old M~dwarfs can be considered low-UV environments, a significant fraction of them emit high \gls{NUV} fluxes into their \gls{HZ} during their early stages, in particular later subtypes~\citep{Richey-Yowell2023}.
-This will help to test the high \gls{NUV} flux end of the UV Threshold Hypothesis; a higher occurrence of biosignatures here would support the hypothesis that a higher NUV flux is favorable or necessary for life.
+This will help to test the high \gls{NUV} flux end of the UV Threshold Hypothesis; a higher occurrence of biosignatures here would support the hypothesis that a higher \gls{NUV} flux is favorable or necessary for life.
 At the same time, a fraction of host stars in our M~dwarf sample extends it to lower maximum past \gls{NUV} fluxes, enabling tests of the low \gls{NUV} flux end of the hypothesis.

### 2025-02-14 src/tex/ms.tex
reword M star opportunity

@@
 An interesting aspect lies in the distribution of host star properties, as different spectral types probe different past \gls{NUV} flux regimes.
-FGK stars show a narrow distribution of maximum past \gls{NUV} fluxes in the \gls{HZ}, which may -- depending on the (unknown) threshold \gls{NUV} flux -- limit the diagnostic power of a survey.
-In the case of a pure FGK sample, it will essentially only be sensitive to \gls{NUV} flux thresholds $\sim \SIrange{200}{400}{\erg\per\second\per\centi\meter\squared}$, and the chance of detecting biosignatures diminishes rapidly for higher thresholds within this range (see Figure~\ref{fig:inhabited_aafo_nuv_thresh}).
-Detected biosignatures in an FGK sample would have little constraining power on testing the UV~Threshold Hypothesis; they would either suggest that low NUV fluxes are sufficient for abiogenesis or indicate a different abiogenic pathway~\citep[e.g.,][]{Westall2018}.
-A lack of biosignatures in a larger FGK sample would support the UV~Threshold Hypothesis.
+FGK stars \rev{exhibit a relatively} narrow \rev{range} of maximum past \gls{NUV} fluxes in the \gls{HZ}, which may -- depending on the (unknown) threshold \rev{flux} -- limit the diagnostic power of a survey.
+\rev{A} pure FGK sample \rev{would} only be sensitive to \rev{flux} thresholds \rev{in the range of} $\sim\SIrange{200}{400}{\erg\per\second\per\centi\meter\squared}$.
+\rev{A detection of biosignatures in such a sample would likely only suggest that either low \gls{NUV} fluxes are sufficient for abiogenesis or that an alternative abiogenic pathway may be at play~\citep[e.g.,][]{Westall2018}.
+Conversely, a lack of biosignatures in a sufficiently large FGK sample could indicate that the actual threshold is higher than the maximum past flux levels reached by FGK hosts.
+We further note that under the UV~Threshold Hypothesis, the fraction of inhabited planets in an FGK sample declines rapidly for increasing threshold fluxes, as shown in Figure~\ref{fig:inhabited_aafo_nuv_thresh}.
+}
 

### 2025-02-14 src/tex/ms.tex
link back to discussion on priors

@@
 The predicted interplay of \gls{NUV} flux and \gls{HZ} occupancy in enabling abiogenesis via the cyanosulfidic scenario could in principle be used to add or remove credibility from a tentative biosignature detection.
-For example, with a strong belief that this scenario is the only viable one for the origins of life, a biosignature detection on a planet orbiting a strongly UV-radiating star may add credibility to the detection.
+For example, with a strong \rev{prior} belief that this scenario is the only viable one for the origins of life \rev{(see Section~\ref{sec:discussion-bayes})}, a biosignature detection on a planet orbiting a strongly UV-radiating star may add credibility to the detection.
 Conversely, a biosignature detection on a planet estimated to have received very little UV~radiation would increase the likelihood of a false positive detection.
@@
 \rev{
-\subsubsection{Bayesian Evidence and the Influence of Priors}
+\subsubsection{Bayesian Evidence and the Influence of Priors}\label{sec:discussion-bayes}}
 Our semi-analytical analysis (Section~\ref{sec:results-semianalytical}) employs Bayes factors to quantify the evidence in favor of or against the UV~Threshold Hypothesis that future observations may provide.

### 2025-02-14 src/tex/acknowledgments_contributions.tex
acknowledge Chia-Lung Lin and the referee

@@
 \section*{Acknowledgments}
-    The authors thank Kevin Heng, Dominik Hintz, and Rhys Seeburger for insightful discussions.
+    The authors thank Kevin Heng, Dominik Hintz, Chia-Lung Lin, and Rhys Seeburger for insightful discussions.
 %    We are grateful to ...
 %    mention any people who gave comments in early-on meetings
-%    We thank the anonymous referee for providing insightful reports that helped to improve this manuscript.
+    We thank the anonymous referee for providing constructive critical feedback that helped to improve this manuscript.
     This material is based upon work supported by the National Aeronautics and Space Administration under Agreement No. 80NSSC21K0593 for the program ``Alien Earths''.

### 2025-02-17 src/tex/ms.tex
edit photochemistry discussion

@@
 However, recent studies challenge this view for the cyanosulfidic scenario.
-\citet{Rimmer2021}  calculated photochemical timescales on early Earth at 180-300 hours (7.5-12.5 days), significantly shorter than the timescale for stochastic geological events~\citep{Rimmer2023}.
-Even with 1000x slower photochemistry on M-dwarf planets, abiogenesis would occur within 20-30 years — geologically negligible compared to stochastic processes — supporting the step-function model.
+\citet{Rimmer2021}  calculated photochemical timescales on early Earth at 180-300 hours (7.5-12.5 days), significantly shorter than the timescale for stochastic geological events also required for the scenario~\citep{Rimmer2023}.
+Even with 1000x slower photochemistry on M-dwarf planets, prebiotic photochemistry like sulfite photolysis would still occur on a geologically negligible timescale of 20-30 years, meaning that the photochemistry is unlikely to be rate limiting compared to stochastic geological processes, justifying the step-function model.
 }

### 2025-02-18 src/scripts/powergrid.py
adapt fig titles

@@ -68,7 +68,7 @@ def create_powergrid_figure(suffix=""):
 
     # Add figure title for dT 100 case
     if suffix == "_dT100":
-        fig.suptitle('Minimum time in HZ: 100 Myr', y=1.02)
+        fig.suptitle('$\Delta T_\mathrm{min}$ = 100 Myr', horizontalalignment='left', y=1.1)
 
     fig.savefig(paths.figures / f"powergrid{suffix}.pdf")
     plt.close()

### 2025-02-18 src/scripts/surveys_FGKM.py

@@ -279,7 +279,7 @@ def main():
 
             # Add figure title for dT 100 case
             if case['suffix'] == '_dT100':
-                fig.suptitle('Minimum time in HZ: 100 Myr', y=1.02)
+                fig.suptitle('$\Delta T_\mathrm{min}$ = 100 Myr', horizontalalignment='left', y=1.05)
 
             # Remove ylabel and tick labels from right plots
             if spt == "M":

### 2025-02-18 src/tex/ms.tex
include long origins timescale case (appendix) and reference everywhere

@@
 For both surveys targeting M dwarfs and those targeting FGK dwarfs, the maximum \gls{NUV} distribution is rather unimodal.
-Applying the approach from Sect.~\ref{sec:results-semianalytical} of fitting a Beta function to the distribution, we find rather low selectivities (see Figure~\ref{fig:surveys_FGKM}), which is likely detrimental for statistical hypothesis tests.
+Applying the approach from Section~\ref{sec:results-semianalytical} of fitting a Beta function to the distribution, we find rather low selectivities (see Figure~\ref{fig:surveys_FGKM}), which is likely detrimental for statistical hypothesis tests.
 
@@
 \subsection{Sampling strategy for testing a \gls{NUV} flux threshold} % corresponds to conclusion point 1
-In Sect.~\ref{sec:results-semianalytical}, we show that testing the UV~Threshold Hypothesis suffers from `nuisance' parameters that hamper inference through astronomical observations.
+In Section~\ref{sec:results-semianalytical}, we show that testing the UV~Threshold Hypothesis suffers from `nuisance' parameters that hamper inference through astronomical observations.
 Here, these parameters are the unspecified value of the \gls{NUV} threshold hypothesized to exist under $H_1$, and the unknown probability of detectable life emerging on a habitable planet $f_{\mathrm{life}}$.
@@
 We further note that under the UV~Threshold Hypothesis, the fraction of inhabited planets in an FGK sample declines rapidly for increasing threshold fluxes, as shown in Figure~\ref{fig:inhabited_aafo_nuv_thresh}.
+In addition, if the timescale for the emergence of life is $\gg \SI{1}{\mega\year}$, the fraction of inhabited planets in an FGK sample may be negligible (see Section~\ref{sec:long-origins}).
 }
@@
 Our simulations find no clear criterion for the credibility of a biosignature detection based on spectral type of the host star, as both FGK and M~dwarf samples show similar maximum past \gls{NUV} flux distributions.
-The abiogenesis rate in both samples show similar trends with the threshold \gls{NUV} flux, and the fraction of inhabited planets drops at similar threshold fluxes (see Section~\ref{sec:results-bioverse}).
+\rev{While there is a significant preference for M~dwarfs in the case of a long timescale for abiogenesis (see Section \ref{sec:long-origins}
+), the fraction of inhabited planets in the M~dwarf sample drops at similar threshold fluxes as in the FGK sample if the timescale for abiogenesis is short }(see Section~\ref{sec:results-bioverse}).
 A potentially inhabited planet's host star spectral type may thus not be a strong indicator for the credibility of a biosignature detection in the context of the UV~Threshold Hypothesis.
@@
     \item Samples of planets orbiting M~dwarfs overall yield higher chances of successfully testing the UV~Threshold Hypothesis.
-          They may also be more likely to yield biosignature detections under this hypothesis.
+          They may also be more likely to yield biosignature detections under this hypothesis\rev{, in particular if the origins timescale is long}.
 
@@
 
+%\section{Impact of a Longer Abiogenesis Timescale}
+\section{\rev{Impact of a Longer Origins Timescale}}\label{sec:long-origins}
+\rev{
+While evidence from Earth's history suggests that life may have emerged early~\citep{Mojzsis2001, Dodd2017}, there is no strong empirical basis to assume that this timescale is representative of habitable planets in the Universe.
+The emergence of life could be a slow and rare process, possibly requiring much longer periods \citep{Lineweaver2002, Spiegel2012}.
+}
+
+\rev{
+To assess the implications of a more conservative abiogenesis timescale, we consider a scenario in which a planet's minimum time required to be both in the \gls{HZ} and having above-threshold \gls{NUV} fluxes before life can emerge is increased to \(\Delta T_{\text{min}} = 100\) Myr, rather than 1 Myr as assumed above.
+Figures~\ref{fig:inhabited_dT100}, \ref{fig:surveys_dT100}, and \ref{fig:powergrid_dT100} show the results of our \bioverse\ simulations under this assumption.
+}
+
+\rev{
+The overall impact differs significantly between M dwarf and FGK dwarf planets.
+ While M dwarf planets remain largely unaffected, the fraction of inhabited FGK dwarf planets decreases sharply due to the limited overlap between the time of high UV flux and habitable conditions (compare Figure~\ref{fig:hz_nuv_evo}).
+Even under the optimistic assumption that the probability of abiogenesis is unity (\( f_{\text{life}} = 1 \)), the fraction of inhabited planets under the UV Threshold Hypothesis declines rapidly with increasing UV flux thresholds (Figure~\ref{fig:inhabited_dT100}).
+}
+\begin{figure}
+    \centering
+    \includegraphics[width=0.9\linewidth]{figures/inhabited_aafo_nuv_thresh_deltatmin100.pdf}
+    \caption{\rev{Fraction of inhabited planets as a function of NUV threshold flux for \(\Delta T_{\text{min}} = 100\) Myr.
+The decrease in inhabited planets is more pronounced for FGK stars, narrowing the viable range of UV Thresholds compared to shorter origins timescales.}}
+    \label{fig:inhabited_dT100}
+\end{figure}
+\begin{figure*}
+    \centering
+    \includegraphics[width=0.9\linewidth]{figures/surveys_FGKM_dT100.pdf}
+    \caption{\rev{Simulated survey results under the assumption of \(\Delta T_{\text{min}} = 100\) Myr.
+The fraction of inhabited FGK planets is reduced to zero, while the fraction of inhabited M dwarf planets remains largely unaffected.}}
+    \label{fig:surveys_dT100}
+\end{figure*}
+
+\rev{
+The impact of a longer origins timescale on testing the UV Threshold Hypothesis becomes evident when we repeat our example survey ($f_{\text{life}} = 0.8$, $F_{\text{NUV},\text{min}} = \SI{300}{\erg\per\second\per\centi\meter\squared}$) under this assumption (Figure~\ref{fig:surveys_dT100}):
+While the survey of M dwarf planets remains largely unaffected, the fraction of inhabited FGK planets is reduced to zero, making a conclusive test of the UV Threshold Hypothesis impossible.
+}
+\begin{figure*}
+    \centering
+    \includegraphics[width=0.9\linewidth]{figures/powergrid_dT100.pdf}
+    \caption{\rev{Statistical power for different abiogenesis rates and NUV threshold fluxes under the assumption of \(\Delta T_{\text{min}} = 100\) Myr.
+The longer origins timescale makes a test of the UV Threshold Hypothesis impossible for FGK planets, while M dwarf planets remain viable targets.}}
+    \label{fig:powergrid_dT100}
+\end{figure*}
+\rev{
+Figure \ref{fig:powergrid_dT100} demonstrates that this result is independent of the threshold \gls{NUV} flux and the intrinsic abiogenesis rate.
+Given the lack of constraints on the timescale of abiogenesis, these findings highlight the advantages of focusing on M dwarf planets when searching for biosignatures.
+If life emerges slowly, FGK planets may rarely, if ever, reach the inhabited stage, whereas M dwarf planets remain viable targets for future surveys.
+If biosignatures were nevertheless detected on FGK planets despite the constraints imposed by a longer abiogenesis timescale, it would suggest that either the UV Threshold Hypothesis, as formulated here, is incorrect or that the timescale of abiogenesis is shorter than $\mathcal{O}(\SI{100}{\mega\year})$.
+}
+
 %! suppress = MissingBibliographystyle

### 2025-02-25 src/scripts/SEDs.py
plot bb SEDs normalized according to Richey-Yowell et al.

@@ -1,0 +1,127 @@
+# Plots and conversions for Spectral Energy Distributions (SEDs)
+
+import matplotlib.pyplot as plt
+import numpy as np
+from astropy.constants import h, c, k_B
+import astropy.units as u
+import pandas as pd
+
+
+def PlanckFunctionLambda(Temperature, lambdas):
+    iArray = np.zeros(len(lambdas))
+    dlamArray = np.zeros(len(lambdas))
+    for i in range(0, len(lambdas) - 1):
+        dlamArray[i] = lambdas[i + 1] - lambdas[i]
+        iArray[i] = (2.0 * h.value * c.value ** 2 / lambdas[i] ** 5) * (
+                1.0 / (np.exp(h.value * c.value / (lambdas[i] * k_B.value * Temperature)) - 1.0))
+    return iArray, dlamArray
+
+
+def plot_sed(ax, wavelength, flux, label, color):
+    ax.plot(wavelength, flux, label=label, color=color)
+    ax.set_xscale('log')
+    ax.set_yscale('log')
+    ax.set_xlabel('Wavelength (nm)')
+    ax.set_ylabel('Flux (erg/s/cm^2/nm)')
+    ax.legend()
+
+
+def bb_sed(min_wl=100, max_wl=9000):
+    """
+    Plot the spectral energy distributions of different spectral types.
+
+    Parameters
+    ----------
+    min_wl : float
+        Minimum wavelength in nm.
+    max_wl : float
+        Maximum wavelength in nm.
+    """
+
+    wavelengths = np.geomspace(min_wl, max_wl, 100) * u.nm  # Wavelengths from 10 nm to 1000 nm
+    temperatures = {'hot': 20000, 'F-type': 7000, 'G-type': 5800, 'K-type': 4500, 'M-type': 3000}
+
+    fig, ax = plt.subplots()
+    colors = {'hot': 'cyan', 'F-type': 'blue', 'G-type': 'green', 'K-type': 'orange', 'M-type': 'red'}
+
+    for star_type, temp in temperatures.items():
+        flux, _ = PlanckFunctionLambda(temp, wavelengths.to(u.m).value)
+        plot_sed(ax, wavelengths.value, flux, star_type, colors[star_type])
+
+    plt.title('Spectral Energy Distributions of Different Spectral Types')
+    plt.show()
+
+
+# def bb2photons():
+
+
+def load_flux_data(filepath):
[…]

### 2025-03-04 src/scripts/SEDs.py
compute Rimmer et al. threshold in energy flux

@@ -8,6 +8,7 @@ import pandas as pd
 import astropy.constants as const
 import paths
 import cmocean
+from utils import save_var_latex
 
 
 def PlanckFunctionLambda(Temperature, lambdas):
@@ -273,6 +274,47 @@ def plot_number_flux_seds(seds_by_age, wavelengths):
     plt.show()
 
 
+def convert_rimmer_threshold():
+    """Convert the Rimmer et al. (2018) photon flux threshold to energy flux.
+    
+    The threshold was determined using mercury lamps which emit >90% of their energy at 254 nm.
+    Therefore, we assume all photons have wavelength 254 nm for the conversion.
+    
+    Rimmer et al. (2018) threshold:
+    - Photon flux: 6.8±3.6 × 10¹⁰ photons cm⁻² s⁻¹ nm⁻¹ integrated from 200-280nm
+    """
+    # Define threshold and uncertainty (single source for these values)
+    threshold_photon = 6.8e10 * u.photon/u.cm**2/u.s/u.nm  # photons cm⁻² s⁻¹ nm⁻¹
+    uncertainty_photon = 3.6e10 * u.photon/u.cm**2/u.s/u.nm
+    
+    # Integration range
+    delta_lambda = (280 - 200) * u.nm
+    
+    # Total integrated photon flux
+    total_photon_flux = threshold_photon * delta_lambda
+    total_photon_uncertainty = uncertainty_photon * delta_lambda
+    
+    # Set up spectral equivalency at 254 nm
+    wavelength = 254 * u.nm
+    equiv = u.spectral_density(wavelength)
+    
+    # Convert photon flux to energy flux using spectral equivalency
+    energy_flux = total_photon_flux.to(u.erg/u.cm**2/u.s, equivalencies=equiv)
+    energy_uncertainty = total_photon_uncertainty.to(u.erg/u.cm**2/u.s, equivalencies=equiv)
+    
+    # Print results
+    print(f"Rimmer et al. (2018) threshold converted to energy flux:")
+    print(f"Threshold: {energy_flux:.2e}")
+    print(f"Uncertainty: {energy_uncertainty:.2e}")
+    
+    # Save to variables.dat with appropriate rounding
+    # The input has 2 significant figures, so we'll round to 2 significant figures
+    save_var_latex('rimmer_threshold', f"{energy_flux.value:.0f}")
+    save_var_latex('rimmer_uncertainty', f"{energy_uncertainty.value:.0f}")
+    
[…]

### 2025-03-04 src/tex/ms.tex
make bold, add additional variables
motivate energy flux instead of photon number flux

@@
 The goal of this work is to evaluate the potential of future exoplanet surveys to test the hypothesis that a minimum past \gls{NUV} flux is required for abiogenesis.
-We focus on one version of the UV Threshold Hypothesis, the so-called cyanosulfidic scenario, which has been refined to the point where the required threshold flux has been measured to be $F_\mathrm{NUV, min} = \SI{6.8\pm3.6 e10}{\photons\per\centi\meter\squared\per\second\per\nano\meter}$ integrated from \SIrange{200}{280}{\nano\meter} at the surface~\citep{Rimmer2018,Rimmer2021a,Rimmer2023,Ranjan2023a}.
+We focus on one version of the UV Threshold Hypothesis, the so-called cyanosulfidic scenario, which has been refined to the point where the required threshold flux has been measured \rev{to be $\SI{6.8\pm3.6 e10}{\photons\per\centi\meter\squared\per\second\per\nano\meter}$ integrated} from \SIrange{200}{280}{\nano\meter} at the surface~\citep{Rimmer2018,Rimmer2021a,Rimmer2023,Ranjan2023a}.
 %\todo[inline]{SR: This is an interesting number, because it is below what was available on early Earth (so this scenario could have worked on early Earth) but until recently it was below what was thought to be available on \gls{HZ} M-dwarf exoplanets.
@@
 the MUSCLES collaboration and the HAZMAT project, but it's worth keeping an eye on in case it is correct after all.}}
+\rev{While prebiotic photochemistry is fundamentally driven by photon number flux, we choose to frame the problem in terms of energy flux, for which empirical estimates in the \gls{NUV} exist across stellar populations and ages~\citep{Richey-Yowell2023}.
+In these units, the above threshold is approximately \var{rimmer_threshold}$\pm$\var{rimmer_uncertainty}$\,$\SI{}{\erg\per\second\per\centi\meter\squared}.
+However, due to uncertainties in how surface flux relates to top-of-atmosphere flux in the habitable zone, we treat the threshold \gls{NUV} flux as a free parameter.}
 
-We follow a semi-analytical Bayesian analysis to estimate probabilities of obtaining strong evidence for or against this hypothesis.
-Under the UV Threshold Hypothesis ($H_1$), the probability of an exoplanet having detectable biosignatures is zero if the \rev{\gls{NUV}} irradiation is less than the threshold, and it is equal to the (unknown) probability of \rev{life emerging and persisting}, $f_\mathrm{life}$, if \gls{NUV} exceeds the threshold for a sufficiently long period of time \rev{$\Delta T_\mathrm{min}$.}
+We \rev{first} follow a semi-analytical Bayesian analysis to estimate probabilities of obtaining strong evidence for or against \rev{the UV Threshold Hypothesis}.
+Under \rev{this hypothesis} ($H_1$), the probability of an exoplanet having detectable biosignatures is zero if the \rev{\gls{NUV}} irradiation is less than the threshold, and it is equal to the (unknown) probability of \rev{life emerging and persisting}, $f_\mathrm{life}$, if \gls{NUV} exceeds the threshold for a sufficiently long period of time \rev{$\Delta T_\mathrm{min}$.}
 Under the null hypothesis ($H_\mathrm{null}$), that probability simply is $f_\mathrm{life}$, that is, it does not correlate with the \gls{UV} flux.
@@
 \appendix
-\section{Alternative priors for the abiogenesis rate}
+\section{\rev{Alternative priors for the abiogenesis rate}}
 \label{sec:appendix-flife-prior}

### 2025-03-05 src/scripts/SEDs.py
add flux ratios

@@ -1,5 +1,7 @@
 # Plots and conversions for Spectral Energy Distributions (SEDs)
 
+# import matplotlib
+# matplotlib.use('Qt5Agg')  # or 'Qt5Agg', 'Qt4Agg', etc. depending on your setup
 import matplotlib.pyplot as plt
 import numpy as np
 from astropy.constants import h, c, k_B
@@ -381,10 +383,12 @@ def compare_dynamic_range(scaling_factors_by_age, photon_fluxes_by_age, flux_dat
     save_var_latex('photon_flux_ratio', f"{photon_flux_ratio:.2f}")
     save_var_latex('integrated_flux_ratio', f"{integrated_flux_ratio:.2f}")
 
+    return photon_flux_ratio, integrated_flux_ratio
 
-def plot_combined_fluxes(photon_fluxes_by_age, flux_data):
+
+def plot_combined_fluxes(photon_fluxes_by_age, flux_data, photon_flux_ratio, integrated_flux_ratio):
     """Plot photon flux and energy flux side by side."""
-    plt.rcParams.update({'font.size': 14})
+    plt.rcParams.update({'font.size': 16})  # Increase the base font size
     fig, (ax1, ax2) = plt.subplots(1, 2, figsize=(20, 8))
     
     x = np.arange(len(list(photon_fluxes_by_age.values())[0].keys()))
@@ -424,22 +428,55 @@ def plot_combined_fluxes(photon_fluxes_by_age, flux_data):
         ax.spines['top'].set_visible(False)
         ax.spines['right'].set_visible(False)
         ax.set_xticks(x + width*2)
-        ax.set_xticklabels(list(list(photon_fluxes_by_age.values())[0].keys()), fontsize=14)
-        ax.tick_params(axis='y', labelsize=14)
+        ax.set_xticklabels(list(list(photon_fluxes_by_age.values())[0].keys()), fontsize=16)  # Increase x-tick font size
+        ax.tick_params(axis='y', labelsize=16)  # Increase y-tick font size
         ax.set_yscale("log")
-        ax.set_xlabel("Spectral Type", fontsize=16)
+        ax.set_xlabel("Spectral Type", fontsize=16)  # Increase x-label font size
     
     # Specific labels
-    ax1.set_ylabel("Photon Number Flux Density\n(photons/s/cm²/nm)", fontsize=16)
-    ax2.set_ylabel("Integrated Energy Flux\n(erg/s/cm²)", fontsize=16)
[…]

### 2025-03-05 src/tex/ms.tex
add an appendix with figure

@@
 
+\section{\rev{Photon Number Fluxes and Energy Fluxes}}\label{appendix:photon_vs_energy_flux}
+\rev{The \gls{NUV} fluxes we used in our simulations are given in units of energy flux, i.e., \si{\erg\per\second\per\centi\meter\squared}, due to their availability across different spectral types and evolutionary stages of stars.
+Here, we assess the robustness of our conclusions when considering photon number fluxes instead.
+}
+\rev{We first computed blackbody spectral energy distributions (SEDs) for representative effective temperatures corresponding to K-type, early M-type, and late M-type stars, and across the evolutionary stages considered in the main text.
+We then normalized these SEDs to align with the observed UV fluxes from \citet{Richey-Yowell2023} and used them to calculate mean number flux densities (photons \si{\per\centi\meter\squared\per\second\per\nano\meter}) in the \SIrange{200}{280}{\nano\meter} wavelength range.}
+
+\begin{figure*}[htb]
+\centering
+\includegraphics[width=\textwidth]{figures/combined_fluxes.pdf}
+\caption{\rev{Comparison of \gls{NUV} flux evolution in terms of photon number flux and energy flux for K-type, early M-type, and late M-type stars.
+In both regimes, later spectral types exhibit higher maximum values.
+The ratio of the maximum past \gls{NUV} fluxes in K dwarfs and late M dwarfs is slightly larger in the photon number flux regime compared to the energy flux regime, suggesting that our conclusions regarding the impact of host star spectral type on the UV Threshold Hypothesis are conservative.
+}}
+\label{fig:combined_fluxes}
+\end{figure*}
+\rev{We found that the derived flux densities for all considered spectral types exceed the threshold estimated by \citet{Rimmer2018} during earlier stages of stellar evolution (Figure~\ref{fig:combined_fluxes}).
+Notably -- as in the case of energy fluxes -- later spectral types exhibit higher maximum past \gls{NUV} number flux densities.}
+
+\rev{We then calculated the ratio of the maximum past \gls{NUV} fluxes in K dwarfs and late M dwarfs in both photon number flux and energy flux regimes.
+We found that this ``dynamic range'' is slightly larger in the photon number flux regime (\var{photon_flux_ratio}) compared to the energy flux regime (\var{integrated_flux_ratio}).
+This suggests that our use of energy flux may slightly underestimate the variations in \gls{NUV} exposure across different stellar types, rendering our conclusions regarding the impact of host star spectral type on the UV Threshold Hypothesis conservative.
+Our results are thus robust when transitioning between energy flux and photon number flux representations.}
+
 %! suppress = MissingBibliographystyle

### 2025-03-05 src/scripts/SEDs.py
just some eye candy

@@ -204,7 +204,8 @@ def plot_photon_flux(photon_fluxes_by_age):
     ax.annotate('Rimmer et al. 2018\n\nthreshold surface flux',
                 xy=(x[-1] + width*5, threshold),
                 xytext=(10, 0), textcoords='offset points',
-                va='center')
+                va='center',
+                fontsize=10)
     
     # Remove top and right spines
     ax.spines['top'].set_visible(False)
@@ -388,8 +389,8 @@ def compare_dynamic_range(scaling_factors_by_age, photon_fluxes_by_age, flux_dat
 
 def plot_combined_fluxes(photon_fluxes_by_age, flux_data, photon_flux_ratio, integrated_flux_ratio):
     """Plot photon flux and energy flux side by side."""
-    plt.rcParams.update({'font.size': 16})  # Increase the base font size
-    fig, (ax1, ax2) = plt.subplots(1, 2, figsize=(20, 8))
+    plt.rcParams.update({'font.size': 18})  # Increase the base font size
+    fig, (ax1, ax2) = plt.subplots(1, 2, figsize=(15, 6)) 
     
     x = np.arange(len(list(photon_fluxes_by_age.values())[0].keys()))
     width = 0.15
@@ -407,14 +408,15 @@ def plot_combined_fluxes(photon_fluxes_by_age, flux_data, photon_flux_ratio, int
     threshold = 6.8e10
     uncertainty = 3.6e10
     ax1.axhline(y=threshold, color='black', linestyle='--', alpha=0.8)
-    ax1.fill_between([-width, x[-1] + width*5], 
+    ax1.fill_between([-width, x[-1] + width*3.5], 
                     threshold - uncertainty, 
                     threshold + uncertainty, 
                     color='darkgray', alpha=0.4)
     ax1.annotate('Rimmer et al. 2018\n\nthreshold surface flux',
-                xy=(x[-1] + width*5, threshold),
+                xy=(x[-1] + width*3.5, threshold),
                 xytext=(10, 0), textcoords='offset points',
-                va='center')
+                va='center',
+                fontsize=12)
     
     # Right plot: Energy flux
     for i, age in enumerate(ages):
@@ -437,8 +439,6 @@ def plot_combined_fluxes(photon_fluxes_by_age, flux_data, photon_flux_ratio, int
     ax1.set_ylabel("Photon Number Flux Density\n(photons/s/cm²/nm)", fontsize=16)  # Increase y-label font size
[…]

### 2025-03-06 src/tex/ms.tex
fix review bold format

@@
 
-\rev{
-\subsubsection{Bayesian Evidence and the Influence of Priors}\label{sec:discussion-bayes}}
-Our semi-analytical analysis (Section~\ref{sec:results-semianalytical}) employs Bayes factors to quantify the evidence in favor of or against the UV~Threshold Hypothesis that future observations may provide.
+\subsubsection{\rev{Bayesian Evidence and the Influence of Priors}}\label{sec:discussion-bayes}
+\rev{Our semi-analytical analysis (Section~\ref{sec:results-semianalytical}) employs Bayes factors to quantify the evidence in favor of or against the UV~Threshold Hypothesis that future observations may provide.}
 %In contrast to frequentist p-value tests, which only assess the likelihood of the null hypothesis, this approach explicitly incorporates the likelihood of the alternative hypothesis.
 %However, Bayes factor analysis limits the scope to two competing hypotheses, and it does not estimate posterior probabilities.
-}
 

### 2025-03-06 src/tex/ms.tex
move timescale text to methods, point to Appendix

@@
 Under the null hypothesis ($H_\mathrm{null}$), that probability simply is $f_\mathrm{life}$, that is, it does not correlate with the \gls{UV} flux.
-\rev{
-While we adopt $\Delta T_\mathrm{min} = \SI{1}{\mega\year}$, we find that increasing this timescale to \SI{100}{\mega\year} reduces the fraction of inhabited planets under $H_1$ by less than 10 \%, suggesting that our results are not highly sensitive to $\Delta T_\mathrm{min}$.
-}
 
@@
 This encompasses the probability of life both emerging and persisting to produce detectable biosignatures.}
-\rev{Due to our ignorance about its true value or even its order of magnitude, we draw $f_\mathrm{life}$ from a log-uniform prior probability distribution.}
+\rev{Due to our ignorance about its true value or even its order of magnitude, we draw $f_\mathrm{life}$ from a log-uniform prior probability distribution (but see Appendix~\ref{sec:appendix-flife-prior} for a discussion of the impact of different priors).}
 Let us consider the probability to detect a biosignature $P(L)$, and let our observable be the inferred past \gls{NUV} flux of the planet $F_\mathrm{NUV}$.
@@
 Further, we required this flux to have lasted for a minimum duration $\Delta T_\mathrm{min}$ to allow for a sufficient ``origins timescale''~\citep{Rimmer2023}.
+\rev{
+We nominally adopt $\Delta T_\mathrm{min} = \SI{1}{\mega\year}$.
+Under $H_1$, longer origins timescales have minimal impact on a generic transit survey but significantly decrease the fraction of inhabited planets around FGK stars, as explored in Appendix~\ref{sec:long-origins}.
+}
+
 All commonly investigated origins-of-life scenarios require water as a solvent;

### 2025-03-06 src/tex/ms.tex
order Appendices by order of appearence in main text
Section -> Appendix

@@
 We \rev{first} follow a semi-analytical Bayesian analysis to estimate probabilities of obtaining strong evidence for or against \rev{the UV Threshold Hypothesis}.
-Under \rev{this hypothesis} ($H_1$), the probability of an exoplanet having detectable biosignatures is zero if the \rev{\gls{NUV}} irradiation is less than the threshold, and it is equal to the (unknown) probability of \rev{life emerging and persisting}, $f_\mathrm{life}$, if \gls{NUV} exceeds the threshold for a sufficiently long period of time \rev{$\Delta T_\mathrm{min}$.}
+Under \rev{this hypothesis} ($H_1$), the probability of an exoplanet having detectable biosignatures is zero if the \rev{\gls{NUV}} irradiation is less than the threshold, and it is equal to the (unknown) probability of \rev{life emerging and persisting}, $f_\mathrm{life}$, if \gls{NUV} exceeds the threshold for a sufficiently long period of time.
 Under the null hypothesis ($H_\mathrm{null}$), that probability simply is $f_\mathrm{life}$, that is, it does not correlate with the \gls{UV} flux.
@@
 Our simulations find no clear criterion for the credibility of a biosignature detection based on spectral type of the host star, as both FGK and M~dwarf samples show similar maximum past \gls{NUV} flux distributions.
-\rev{While there is a significant preference for M~dwarfs in the case of a long timescale for abiogenesis (see Section \ref{sec:long-origins}
+\rev{While there is a significant preference for M~dwarfs in the case of a long timescale for abiogenesis (see Appendix \ref{sec:long-origins}
 ), the fraction of inhabited planets in the M~dwarf sample drops at similar threshold fluxes as in the FGK sample if the timescale for abiogenesis is short }(see Section~\ref{sec:results-bioverse}).
@@
 \appendix
+
+\section{\rev{Photon Number Fluxes and Energy Fluxes}}\label{appendix:photon_vs_energy_flux}
+\rev{The \gls{NUV} fluxes we used in our simulations are given in units of energy flux, i.e., \si{\erg\per\second\per\centi\meter\squared}, due to their availability across different spectral types and evolutionary stages of stars.
+Here, we assess the robustness of our conclusions when considering photon number fluxes instead.
+}
+\rev{We first computed blackbody spectral energy distributions (SEDs) for representative effective temperatures corresponding to K-type, early M-type, and late M-type stars, and across the evolutionary stages considered in the main text.
+We then normalized these SEDs to align with the observed UV fluxes from \citet{Richey-Yowell2023} and used them to calculate mean number flux densities (photons \si{\per\centi\meter\squared\per\second\per\nano\meter}) in the \SIrange{200}{280}{\nano\meter} wavelength range.}
+
+\begin{figure*}[htb]
+    \centering
+    \includegraphics[width=\textwidth]{figures/combined_fluxes.pdf}
+    \caption{\rev{Comparison of \gls{NUV} flux evolution in terms of photon number flux and energy flux for K-type, early M-type, and late M-type stars.
+    In both regimes, later spectral types exhibit higher maximum values.
+    The ratio of the maximum past \gls{NUV} fluxes in K dwarfs and late M dwarfs is slightly larger in the photon number flux regime compared to the energy flux regime, suggesting that our conclusions regarding the impact of host star spectral type on the UV Threshold Hypothesis are conservative.
+    }}
+    \label{fig:combined_fluxes}
+\end{figure*}
+\rev{We found that the derived flux densities for all considered spectral types exceed the threshold estimated by \citet{Rimmer2018} during earlier stages of stellar evolution (Figure~\ref{fig:combined_fluxes}).
+Notably -- as in the case of energy fluxes -- later spectral types exhibit higher maximum past \gls{NUV} number flux densities.}
+
+\rev{We then calculated the ratio of the maximum past \gls{NUV} fluxes in K dwarfs and late M dwarfs in both photon number flux and energy flux regimes.
+We found that this ``dynamic range'' is slightly larger in the photon number flux regime (\var{photon_flux_ratio}) compared to the energy flux regime (\var{integrated_flux_ratio}).
+This suggests that our use of energy flux may slightly underestimate the variations in \gls{NUV} exposure across different stellar types, rendering our conclusions regarding the impact of host star spectral type on the UV Threshold Hypothesis conservative.
+Our results are thus robust when transitioning between energy flux and photon number flux representations.}
+
+%! suppress = MissingBibliographystyle
+\bibliography{bib,coauthors}
+
 \section{\rev{Alternative priors for the abiogenesis rate}}
@@
     \centering
-    \includegraphics[width=0.9\linewidth]{figures/inhabited_aafo_nuv_thresh_deltatmin100.pdf}
+    \includegraphics[width=0.9\textwidth]{figures/inhabited_aafo_nuv_thresh_deltatmin100.pdf}
     \caption{\rev{Fraction of inhabited planets as a function of NUV threshold flux for \(\Delta T_{\text{min}} = 100\) Myr.
@@
 }
-
-\section{\rev{Photon Number Fluxes and Energy Fluxes}}\label{appendix:photon_vs_energy_flux}
-\rev{The \gls{NUV} fluxes we used in our simulations are given in units of energy flux, i.e., \si{\erg\per\second\per\centi\meter\squared}, due to their availability across different spectral types and evolutionary stages of stars.
-Here, we assess the robustness of our conclusions when considering photon number fluxes instead.
-}
-\rev{We first computed blackbody spectral energy distributions (SEDs) for representative effective temperatures corresponding to K-type, early M-type, and late M-type stars, and across the evolutionary stages considered in the main text.
-We then normalized these SEDs to align with the observed UV fluxes from \citet{Richey-Yowell2023} and used them to calculate mean number flux densities (photons \si{\per\centi\meter\squared\per\second\per\nano\meter}) in the \SIrange{200}{280}{\nano\meter} wavelength range.}
-
-\begin{figure*}[htb]
-\centering
-\includegraphics[width=\textwidth]{figures/combined_fluxes.pdf}
-\caption{\rev{Comparison of \gls{NUV} flux evolution in terms of photon number flux and energy flux for K-type, early M-type, and late M-type stars.
-In both regimes, later spectral types exhibit higher maximum values.
-The ratio of the maximum past \gls{NUV} fluxes in K dwarfs and late M dwarfs is slightly larger in the photon number flux regime compared to the energy flux regime, suggesting that our conclusions regarding the impact of host star spectral type on the UV Threshold Hypothesis are conservative.
-}}
-\label{fig:combined_fluxes}
-\end{figure*}
-\rev{We found that the derived flux densities for all considered spectral types exceed the threshold estimated by \citet{Rimmer2018} during earlier stages of stellar evolution (Figure~\ref{fig:combined_fluxes}).
-Notably -- as in the case of energy fluxes -- later spectral types exhibit higher maximum past \gls{NUV} number flux densities.}
-
-\rev{We then calculated the ratio of the maximum past \gls{NUV} fluxes in K dwarfs and late M dwarfs in both photon number flux and energy flux regimes.
-We found that this ``dynamic range'' is slightly larger in the photon number flux regime (\var{photon_flux_ratio}) compared to the energy flux regime (\var{integrated_flux_ratio}).
-This suggests that our use of energy flux may slightly underestimate the variations in \gls{NUV} exposure across different stellar types, rendering our conclusions regarding the impact of host star spectral type on the UV Threshold Hypothesis conservative.
-Our results are thus robust when transitioning between energy flux and photon number flux representations.}
-
-%! suppress = MissingBibliographystyle
-\bibliography{bib,coauthors}
-
 \end{document}

### 2025-03-08 src/tex/ms.tex
fix a few formatting issues

@@
 However, recent studies challenge this view for the cyanosulfidic scenario.
-\citet{Rimmer2021}  calculated photochemical timescales on early Earth at 180-300 hours (7.5-12.5 days), significantly shorter than the timescale for stochastic geological events also required for the scenario~\citep{Rimmer2023}.
-Even with 1000x slower photochemistry on M-dwarf planets, prebiotic photochemistry like sulfite photolysis would still occur on a geologically negligible timescale of 20-30 years, meaning that the photochemistry is unlikely to be rate limiting compared to stochastic geological processes, justifying the step-function model.
+\citet{Rimmer2021} calculated photochemical timescales on early Earth at 180--300 hours (7.5--12.5 days), significantly shorter than the timescale for stochastic geological events also required for the scenario~\citep{Rimmer2023}.
+Even with 1000x slower photochemistry on M-dwarf planets, prebiotic photochemistry like sulfite photolysis would still occur on a geologically negligible timescale of 20--30 years, meaning that the photochemistry is unlikely to be rate limiting compared to stochastic geological processes, justifying the step-function model.
 }
@@
 \rev{
-To assess the implications of a more conservative abiogenesis timescale, we consider a scenario in which a planet's minimum time required to be both in the \gls{HZ} and having above-threshold \gls{NUV} fluxes before life can emerge is increased to \(\Delta T_{\text{min}} = 100\) Myr, rather than 1 Myr as assumed above.
+To assess the implications of a more conservative abiogenesis timescale, we consider a scenario in which a planet's minimum time required to be both in the \gls{HZ} and having above-threshold \gls{NUV} fluxes before life can emerge is increased to $\Delta T_{\mathrm{min}} = 100$ Myr, rather than 1 Myr as assumed above.
 Figures~\ref{fig:inhabited_dT100}, \ref{fig:surveys_dT100}, and \ref{fig:powergrid_dT100} show the results of our \bioverse\ simulations under this assumption.
@@
  While M dwarf planets remain largely unaffected, the fraction of inhabited FGK dwarf planets decreases sharply due to the limited overlap between the time of high UV flux and habitable conditions (compare Figure~\ref{fig:hz_nuv_evo}).
-Even under the optimistic assumption that the probability of abiogenesis is unity (\( f_{\text{life}} = 1 \)), the fraction of inhabited planets under the UV Threshold Hypothesis declines rapidly with increasing UV flux thresholds (Figure~\ref{fig:inhabited_dT100}).
+Even under the optimistic assumption that the probability of abiogenesis is unity ($ f_{\mathrm{life}} = 1 $), the fraction of inhabited planets under the UV Threshold Hypothesis declines rapidly with increasing UV flux thresholds (Figure~\ref{fig:inhabited_dT100}).
 }
@@
     \includegraphics[width=0.9\textwidth]{figures/inhabited_aafo_nuv_thresh_deltatmin100.pdf}
-    \caption{\rev{Fraction of inhabited planets as a function of NUV threshold flux for \(\Delta T_{\text{min}} = 100\) Myr.
+    \caption{\rev{Fraction of inhabited planets as a function of NUV threshold flux for $\Delta T_{\mathrm{min}} = 100$ Myr.
 The decrease in inhabited planets is more pronounced for FGK stars, narrowing the viable range of UV Thresholds compared to shorter origins timescales.}}
@@
     \includegraphics[width=0.9\linewidth]{figures/surveys_FGKM_dT100.pdf}
-    \caption{\rev{Simulated survey results under the assumption of \(\Delta T_{\text{min}} = 100\) Myr.
+    \caption{\rev{Simulated survey results under the assumption of $\Delta T_{\mathrm{min}} = 100$ Myr.
 The fraction of inhabited FGK planets is reduced to zero, while the fraction of inhabited M dwarf planets remains largely unaffected.}}
@@
 \rev{
-The impact of a longer origins timescale on testing the UV Threshold Hypothesis becomes evident when we repeat our example survey ($f_{\text{life}} = 0.8$, $F_{\text{NUV},\text{min}} = \SI{300}{\erg\per\second\per\centi\meter\squared}$) under this assumption (Figure~\ref{fig:surveys_dT100}):
+The impact of a longer origins timescale on testing the UV Threshold Hypothesis becomes evident when we repeat our example survey ($f_{\mathrm{life}} = 0.8$, $F_{\mathrm{NUV},\mathrm{min}} = \SI{300}{\erg\per\second\per\centi\meter\squared}$) under this assumption (Figure~\ref{fig:surveys_dT100}):
 While the survey of M dwarf planets remains largely unaffected, the fraction of inhabited FGK planets is reduced to zero, making a conclusive test of the UV Threshold Hypothesis impossible.
@@
     \includegraphics[width=0.9\linewidth]{figures/powergrid_dT100.pdf}
-    \caption{\rev{Statistical power for different abiogenesis rates and NUV threshold fluxes under the assumption of \(\Delta T_{\text{min}} = 100\) Myr.
+    \caption{\rev{Statistical power for different abiogenesis rates and NUV threshold fluxes under the assumption of $\Delta T_{\mathrm{min}} = 100$ Myr.
 The longer origins timescale makes a test of the UV Threshold Hypothesis impossible for FGK planets, while M dwarf planets remain viable targets.}}

### 2025-03-11 src/tex/ms.tex
change wording

@@
 
-\rev{If the rate of life’s emergence and persistence $f_\mathrm{life}$ is at a \SI{1}{\percent} level or lower, no currently planned exoplanet mission concept has projected sample sizes sufficient to test the UV~Threshold Hypothesis.
-One } exception is the Nautilus Space Observatory concept~\citep{Apai2019,Apai2022}.
+\rev{If the rate of life’s emergence and persistence $f_\mathrm{life}$ is at a \SI{1}{\percent} level or lower, no currently envisioned future exoplanet mission has projected sample sizes sufficient to test the UV~Threshold Hypothesis.
+One possible} exception is the Nautilus Space Observatory concept~\citep{Apai2019,Apai2022}.
 Nautilus aims to characterize up to $\sim 1000$ \gls{EEC} via transmission spectroscopy, building on an innovative optical technology.

### 2025-03-12 src/tex/ms.tex
justify bb approximation
start stronger

@@
 \label{sec:intro}
-\rev{While the probability of abiogenesis is unknown and has primarily been explored through statistical arguments~\citep[e.g.,][]{Spiegel2012,Kipping2021,Lingam2024},
-a} wide variety of scenarios for the origin of life have been proposed~\citep[e.g.,][]{Baross1985,Brasier2011,Mulkidjanian2012,Fox2013,Deamer2015,Westall2018}.
+\rev{The probability of abiogenesis remains one of the fundamental questions in modern science.
+As of now, it has been primarily explored through statistical arguments~\citep[e.g.,][]{Spiegel2012,Kipping2021,Lingam2024}.
+At the same time, a} wide variety of scenarios for the origin of life have been proposed~\citep[e.g.,][]{Baross1985,Brasier2011,Mulkidjanian2012,Fox2013,Deamer2015,Westall2018}.
 While we may still be far from conclusively testing these scenarios, new prospects in the search for conditions favorable to life have opened up by thinking of the origin of life as a planetary phenomenon and identifying global-scale environmental properties that might support pathways to life~\citep{Sasselov2020}.
@@
 \rev{We first computed blackbody spectral energy distributions (SEDs) for representative effective temperatures corresponding to K-type, early M-type, and late M-type stars, and across the evolutionary stages considered in the main text.
+While blackbody SEDs are not necessarily representative of M-dwarf SEDs in the \gls{NUV}~\citep{Seager2013,Rugheimer2015}, the spectrally resolved evolution of M-dwarf SEDs as a function of age is not known; we therefore adopt this simplified prescription for the purpose of this sensitivity test.
 We then normalized these SEDs to align with the observed UV fluxes from \citet{Richey-Yowell2023} and used them to calculate mean number flux densities (photons \si{\per\centi\meter\squared\per\second\per\nano\meter}) in the \SIrange{200}{280}{\nano\meter} wavelength range.}

### 2025-03-12 src/tex/ms.tex
disentangle fluxes

@@
 The goal of this work is to evaluate the potential of future exoplanet surveys to test the hypothesis that a minimum past \gls{NUV} flux is required for abiogenesis.
-We focus on one version of the UV Threshold Hypothesis, the so-called cyanosulfidic scenario, which has been refined to the point where the required threshold flux has been measured \rev{to be $\SI{6.8\pm3.6 e10}{\photons\per\centi\meter\squared\per\second\per\nano\meter}$ integrated} from \SIrange{200}{280}{\nano\meter} at the surface~\citep{Rimmer2018,Rimmer2021a,Rimmer2023,Ranjan2023a}.
-%\todo[inline]{SR: This is an interesting number, because it is below what was available on early Earth (so this scenario could have worked on early Earth) but until recently it was below what was thought to be available on \gls{HZ} M-dwarf exoplanets.
-So it was thought that identification of biosignatures on M-dwarf planets could therefore falsify the cyanosulfidic scenario, with a potential caveat for transient UV from flares.
-Two recent developments have complicated the picture.
-First, Rimmer et al.
-2018 had an error in their radiative transfer routines.
-Correcting for this error, early M-dwarfs and highly active M-dwarfs emit enough UV to meet the Rimmer et al.
-2018 criterion (Ranjan et al.
-2023).
-%\sout{Second, a recent publication argues that /all/ estimates of M-dwarf UV are underestimates, and that late M-dwarf stars have similar emission to Sunlike stars (Rekhi et al.
-2023).
-I suspect this is incorrect, because it contradicts a lot of work from e.g.
-the MUSCLES collaboration and the HAZMAT project, but it's worth keeping an eye on in case it is correct after all.}}
+We focus on one version of the UV Threshold Hypothesis, the so-called cyanosulfidic scenario, which has been refined to the point where the required threshold flux has been measured \rev{to be $\SI{6.8\pm3.6 e10}{\photons\per\centi\meter\squared\per\second\per\nano\meter}$ integrated} from \SIrange{200}{280}{\nano\meter} at the surface~\citep{Rimmer2018,Rimmer2021a,Rimmer2023,Ranjan2023a}\rev{\footnote{or $\sim$\var{rimmer_threshold}$\pm$\var{rimmer_uncertainty}$\,$\SI{}{\erg\per\second\per\centi\meter\squared} in energy flux units.}}.
 \rev{While prebiotic photochemistry is fundamentally driven by photon number flux, we choose to frame the problem in terms of energy flux, for which empirical estimates in the \gls{NUV} exist across stellar populations and ages~\citep{Richey-Yowell2023}.
-In these units, the above threshold is approximately \var{rimmer_threshold}$\pm$\var{rimmer_uncertainty}$\,$\SI{}{\erg\per\second\per\centi\meter\squared}.
-However, due to uncertainties in how surface flux relates to top-of-atmosphere flux in the habitable zone, we treat the threshold \gls{NUV} flux as a free parameter.}
+Due to the inherent uncertainties concerning the relationship between surface and top-of-atmosphere fluxes, we treat the threshold \gls{NUV} flux as a free parameter.}
 

### 2025-03-12 src/tex/ms.tex
explain why these HZ limits

@@
 Here, we adopted orbital distance estimates that define the \gls{HZ} as the region between the runaway greenhouse transition, where the stellar instellation cannot anymore be balanced through infrared cooling to space~\citep{Ingersoll1969}, and the maximum greenhouse limit, corresponding to the maximum distance at which surface temperatures allowing liquid water can be maintained through a CO$_2$ greenhouse effect~\citep{Kasting1991,Kasting1993,Underwood2003,Kopparapu2013,Kopparapu2014}.
-We used the parametrization in \citet{Kopparapu2014} to derive luminosity and planetary mass-dependent distance limits of the \gls{HZ} $a_\mathrm{inner}$ and $a_\mathrm{outer}$.
-\rev{The specific boundaries of the habitable zone are known to be sensitive to the star’s luminosity, spectral type, the planet’s mass, and the planet’s atmospheric properties~\citep[e.g.,][]{Pierrehumbert2011,Ramirez2014,Ramirez2017,Ramirez2018b,Koll2019,Ramirez2018,Ramirez2020,Bonati2021,Chaverot2022,Turbet2023}.
-The specific habitable zone limits can therefore be variable.
-Here -- to constrain the complexity of this study -- we adopted a conservative approach that uses the commonly used limits.
+\rev{The exact boundaries of the habitable zone are known to be sensitive to the star’s luminosity, spectral type, the planet’s mass, and the planet’s atmospheric properties~\citep[e.g.,][]{Pierrehumbert2011,Ramirez2014,Ramirez2017,Ramirez2018b,Koll2019,Ramirez2018,Ramirez2020,Bonati2021,Chaverot2022,Turbet2023}.
+Here, we adopted the commonly used parametrization in \citet{Kopparapu2014} to derive luminosity and planetary mass-dependent distance limits of the \gls{HZ} $a_\mathrm{inner}$ and $a_\mathrm{outer}$.
 }

### 2025-03-14 src/tex/ms.tex
update affiliation
update biblio

@@
 \affiliation{Steward Observatory, The University of Arizona, Tucson, AZ 85721, USA; \href{mailto:schlecker@arizona.edu}{schlecker@arizona.edu}}
+\affiliation{Caltech/IPAC-NASA Exoplanet Science Institute, 1200 E. California Blvd., MC 100-22, Pasadena, CA 91125, USA}
 \author[0000-0002-3286-7683]{Tim Lichtenberg}
@@
 A real exoplanet survey will be subject to observational biases and sample selection effects, and will be constrained by the underlying demographics of the planet sample.
-To assess the information gain of a realistic exoplanet survey, we employed \bioverse~\citep{Bixel2021,Hardegree-Ullman2023,Schlecker2024,Hardegree-Ullman2024}, a framework that integrates multiple components including statistically realistic simulations of exoplanet populations, a survey simulation module, and a hypothesis testing module to evaluate the statistical power of different observational strategies.
+To assess the information gain of a realistic exoplanet survey, we employed \bioverse~\citep{Bixel2021,Hardegree-Ullman2023,Schlecker2024,Hardegree-Ullman2025}, a framework that integrates multiple components including statistically realistic simulations of exoplanet populations, a survey simulation module, and a hypothesis testing module to evaluate the statistical power of different observational strategies.
 
@@
 Ground-based 25--40-meter class extremely large telescopes are expected to have the capabilities to detect biosignatures on exoplanets like Proxima Centauri~b~\citep[e.g.,][]{Wang2017,Hawker2019,Zhang2024,Vaughan2024}.
-\citet{Hardegree-Ullman2024} used \bioverse\ to determine potential yields for a 10-year direct imaging and high-resolution spectroscopy survey of O$_2$ on the Giant Magellan Telescope (GMT) and on the Extremely Large Telescope (ELT) and found that between 7 and 19 habitable zone Earth-sized planets could be probed for Earth-like oxygen levels.
+\citet{Hardegree-Ullman2025} used \bioverse\ to determine potential yields for a 10-year direct imaging and high-resolution spectroscopy survey of O$_2$ on the Giant Magellan Telescope (GMT) and on the Extremely Large Telescope (ELT) and found that between 7 and 19 habitable zone Earth-sized planets could be probed for Earth-like oxygen levels.
 Such a sample is too small to test the UV Threshold Hypothesis, but it may be synergistic with other detection methods.

### 2025-03-14 src/tex/ms.tex
add information about Zenodo repo

@@
 \section*{Reproducibility}
-All code required to reproduce our results, figures, and this article itself is available at \url{https://github.com/matiscke/originsoflife}.
+All code required to reproduce our results, figures, and this article itself is available at \url{https://github.com/matiscke/originsoflife}\rev{, and the repository has been archived on Zenodo at \url{https://zenodo.org/records/15022723}.}
 

### 2025-03-24 src/tex/ms.tex
did we have to shorten the abstract? remove final sentence.
semi-analytical analysis: lower limit 10^-15

@@
  Beyond specific scenarios, this work underscores the broader value of realistic survey simulations for future observatories (e.g., HWO, LIFE, ELTs, Nautilus) in identifying testable science questions, optimizing mission strategies, and advancing theoretical and experimental studies of abiogenesis.
-% This approach emphasizes the need to contextualize biosignature detections, enhancing their credibility and offering new approaches to explore the emergence of life on Earth and beyond.
 \end{abstract}
@@
 This encompasses the probability of life both emerging and persisting to produce detectable biosignatures.}
-\rev{Due to our ignorance about its true value or even its order of magnitude, we draw $f_\mathrm{life}$ from a log-uniform prior probability distribution (but see Appendix~\ref{sec:appendix-flife-prior} for a discussion of the impact of different priors).}
+\rev{Due to our ignorance about its true value or even its order of magnitude, we draw $f_\mathrm{life}$ from a log-uniform prior probability distribution} \revv{over the range $[10^{-15}, 1]$~\citep[][but see Appendix~\ref{sec:appendix-flife-prior} for a discussion of the impact of different priors]{Balbi2023a}}.
 Let us consider the probability to detect a biosignature $P(L)$, and let our observable be the inferred past \gls{NUV} flux of the planet $F_\mathrm{NUV}$.
@@
 In Section~\ref{sec:met-semianalytical}, we computed the probability for true positive evidence for $H_\mathrm{1}$ and $H_\mathrm{null}$, respectively (Equations~\ref{eq:bayes_factor},~\ref{eq:bayes_factor2}).
-Figure~\ref{fig:semian_true_evidence} shows how these evidences are distributed for sample sizes \var{semian_Nsamp1} and \var{semian_Nsamp2}, and how likely we are to obtain strong evidence ($BF_{H_i, H_j}$ > 10) \rev{in the agnostic case where we draw $f_\mathrm{life}$ from a log-uniform distribution}.
+Figure~\ref{fig:semian_true_evidence} shows how these evidences are distributed for sample sizes \var{semian_Nsamp1} and \var{semian_Nsamp2}, and how likely we are to obtain strong evidence ($BF_{H_i, H_j}$ > 10) \rev{in the agnostic case where we draw $f_\mathrm{life}$ from a log-uniform distribution \revv{over the range $[10^{-15}, 1]$}}.
 For $n = \var{semian_Nsamp1}$, strong true evidence for $H_\mathrm{1}$ ($H_\mathrm{null}$) can be \rev{expected in $\sim \SI{3}{\percent}$ ($\sim \SI{6}{\percent}$) of} all random experiments.
@@
 \label{sec:appendix-flife-prior}
-\rev{In the main text, we used a log-uniform prior for the probability of life emerging and persisting $f_\mathrm{life}$ in order to reflect our ignorance about this parameter and its order of magnitude.
+\rev{In the main text, we used a log-uniform prior \revv{over the range $[10^{-15}, 1]$} for the probability of life emerging and persisting $f_\mathrm{life}$ in order to reflect our ignorance about this parameter and its order of magnitude.
 We consider this a reasonable choice, as it is agnostic about the scale of $f_\mathrm{life}$ and assigns equal prior probability to all orders of magnitude.

### 2025-03-24 src/scripts/semian_true_evidence.py

@@ -73,7 +73,8 @@ def get_evidence(sampsize=100, nboot=100000):
     # pisamp = rng.uniform(0, 1, nboot)
 
     # log-uniform f_life
-    pisamp = 10**rng.uniform(-6, 0, nboot)
+    pisamp = 10**rng.uniform(-15, 0, nboot) # 1e-15 consistent with Balbi&Lingam 2023
+
 
     theta_lsamp = rng.uniform(0, 1, nboot)
     kh1samp = sample_H1(theta, pisamp, theta_lsamp)

### 2025-03-24 src/tex/ms.tex
cite a paper on origins timescale

@@
 Further, we required this flux to have lasted for a minimum duration $\Delta T_\mathrm{min}$ to allow for a sufficient ``origins timescale''~\citep{Rimmer2023}.
-\rev{
-We nominally adopt $\Delta T_\mathrm{min} = \SI{1}{\mega\year}$.
-Under $H_1$, longer origins timescales have minimal impact on a generic transit survey but significantly decrease the fraction of inhabited planets around FGK stars, as explored in Appendix~\ref{sec:long-origins}.
+\revv{While there is no consensus on the minimum timescale required for abiogenesis~\citep[e.g.,][]{Lazcano1996}, we adopt  a nominal value of} \rev{$\Delta T_\mathrm{min} = \SI{1}{\mega\year}$.}
+\rev{Under $H_1$, longer origins timescales have minimal impact on a generic transit survey but significantly decrease the fraction of inhabited planets around FGK stars, as explored in Appendix~\ref{sec:long-origins}.
 }

### 2025-03-25 src/tex/ms.tex
delete biosignature detection rate
compute summary stats for stars

@@
 \affiliation{Blue Marble Space Institute of Science, Seattle, 98104, USA}
-\author[0000-0002-5806-5566]{Régis Ferrière}
+\author[0000-0002-5806-5566]{R\'{e}gis Ferri\`{e}re}
 \affiliation{Department of Ecology and Evolutionary Biology, University of Arizona, Tucson AZ, USA}
@@
 \affiliation{Kapteyn Astronomical Institute, University of Groningen, PO Box 800, 9700 AV Groningen, The Netherlands}
-\author[0000-0003-3557-6256]{Stéphane Mazevet}
-\affiliation{Observatoire de la Côte d'Azur, Université Côte d'Azur, Nice, France}
+\author[0000-0003-3557-6256]{St\'{e}phane Mazevet}
+\affiliation{Observatoire de la C\^{o}te d'Azur, Universit\'{e} C\^{o}te d'Azur, Nice, France}
 
@@
 The alternate scenario of a surface locally subject to wet-dry cycles requires a planetary exposure to mid-range \gls{UV} irradiation, as a source of energy and an agent of selection in chemical evolution~\citep[e.g.,][]{Deamer2019}.
-This “UV Threshold Hypothesis” states that \gls{UV} light in a specific wavelength range played a constructive role in getting life started on Earth~\citep{Ranjan2016,Ranjan2017c,Rimmer2018,Rapf2016}, and it could provide a probabilistic approach to the interpretation of possible future biosignature detections~\citep[e.g.,][]{Catling2018,Walker2018}.
+This "UV Threshold Hypothesis" states that \gls{UV} light in a specific wavelength range played a constructive role in getting life started on Earth~\citep{Ranjan2016,Ranjan2017c,Rimmer2018,Rapf2016}, and it could provide a probabilistic approach to the interpretation of possible future biosignature detections~\citep[e.g.,][]{Catling2018,Walker2018}.
 
@@
 We included stars out to a maximum distance $d_{\max}$ that depends on the required planet sample size.
+\revv{In our FGK sample, we find that K~dwarfs dominate with effective temperatures of $\var{median_T_eff_st_K_FGK}^{+\var{p84_T_eff_st_K_FGK}}_{-\var{p16_T_eff_st_K_FGK}}$~K and masses of $\var{median_M_st_K_FGK}^{+\var{p84_M_st_K_FGK}}_{-\var{p16_M_st_K_FGK}}$~$M_\odot$ (median and 16th/84th percentiles).
+In the M~dwarf sample, the stars have effective temperatures of $\var{median_T_eff_st_M_M}^{+\var{p84_T_eff_st_M_M}}_{-\var{p16_T_eff_st_M_M}}$~K and masses of $\var{median_M_st_M_M}^{+\var{p84_M_st_M_M}}_{-\var{p16_M_st_M_M}}$~$M_\odot$.}
+
 Planets were generated and assigned to the synthetic stars following the occurrence rates and size/orbit distributions of \citet{Bergsten2022}.
@@
 Here, we adopted orbital distance estimates that define the \gls{HZ} as the region between the runaway greenhouse transition, where the stellar instellation cannot anymore be balanced through infrared cooling to space~\citep{Ingersoll1969}, and the maximum greenhouse limit, corresponding to the maximum distance at which surface temperatures allowing liquid water can be maintained through a CO$_2$ greenhouse effect~\citep{Kasting1991,Kasting1993,Underwood2003,Kopparapu2013,Kopparapu2014}.
-\rev{The exact boundaries of the habitable zone are known to be sensitive to the star’s luminosity, spectral type, the planet’s mass, and the planet’s atmospheric properties~\citep[e.g.,][]{Pierrehumbert2011,Ramirez2014,Ramirez2017,Ramirez2018b,Koll2019,Ramirez2018,Ramirez2020,Bonati2021,Chaverot2022,Turbet2023}.
+\rev{The exact boundaries of the habitable zone are known to be sensitive to the star's luminosity, spectral type, the planet's mass, and the planet's atmospheric properties~\citep[e.g.,][]{Pierrehumbert2011,Ramirez2014,Ramirez2017,Ramirez2018b,Koll2019,Ramirez2018,Ramirez2020,Bonati2021,Chaverot2022,Turbet2023}.
 Here, we adopted the commonly used parametrization in \citet{Kopparapu2014} to derive luminosity and planetary mass-dependent distance limits of the \gls{HZ} $a_\mathrm{inner}$ and $a_\mathrm{outer}$.
@@
 % Actually it is more than just NUV, it is whatever index that integrates \gls{NUV} and HZ requiremetns into a single scalar variable for which a threshold exist
-A high biosignature detection rate $f_\mathrm{life}$ increases the evidence (see Equation~\ref{eq:bayes_factor}) but a survey strategy cannot influence it.
+\revv{Higher $f_\mathrm{life}$ increase} the evidence (see Equation~\ref{eq:bayes_factor}) but a survey strategy cannot influence \revv{this parameter}.
 The same is true for $F_\mathrm{NUV, min}$, where again higher values increase the evidence as the binomial distribution for $H_\mathrm{1}$ gets increasingly skewed and shifted away from the one for $H_\mathrm{null}$.
@@
 The \gls{HWO} is expected to characterize a sample of $\sim 25$ Earth analogs~\citep{Mamajek2023,Tuchow2024}.
-Depending on the technical design, \life\ is expected to target 25–80 \glspl{EEC}~\citep{Kammerer2018,Quanz2022}, which could be just sufficient to constrain the UV~Threshold Hypothesis.
+Depending on the technical design, \life\ is expected to target 25--80 \glspl{EEC}~\citep{Kammerer2018,Quanz2022}, which could be just sufficient to constrain the UV~Threshold Hypothesis.
 
-\rev{If the rate of life’s emergence and persistence $f_\mathrm{life}$ is at a \SI{1}{\percent} level or lower, no currently envisioned future exoplanet mission has projected sample sizes sufficient to test the UV~Threshold Hypothesis.
+\rev{If the rate of life's emergence and persistence $f_\mathrm{life}$ is at a \SI{1}{\percent} level or lower, no currently envisioned future exoplanet mission has projected sample sizes sufficient to test the UV~Threshold Hypothesis.
 One possible} exception is the Nautilus Space Observatory concept~\citep{Apai2019,Apai2022}.
@@
 
-\rev{\citet{Ranjan2017b} speculated that UV photochemistry might be rate-limiting for abiogenesis, particularly on planets orbiting M dwarfs, due to their lower baseline UV fluxes.
+\rev{\citet{Ranjan2017b} speculated that UV photochemistry might be rate-limiting for abiogenesis, particularly on planets orbiting M dwarfs, due to their low baseline UV fluxes.
 This could delay abiogenesis by orders of magnitude, resulting in a continuous dependence of abiogenesis likelihood on UV flux.

### 2025-03-25 src/scripts/bioverse_pipeline.py

@@ -389,6 +389,48 @@ def get_file_suffix(params):
     return suffix
 
 
+def compute_stellar_stats(dd, hoststars, file_suffix):
+    """Compute and save statistics about stellar parameters in the sample.
+    
+    Parameters
+    ----------
+    dd : pandas.DataFrame
+        DataFrame containing the stellar data
+    hoststars : str
+        The spectral type of the host stars to consider
+    file_suffix : str
+        Suffix to append to variable names
+    """
+    # Calculate and save statistics for each spectral type
+    for spT in dd.SpT.unique():
+        spT_data = dd[dd.SpT == spT]
+        
+        # Basic counts and fractions
+        save_var_latex(f"N_{spT}_{hoststars}{file_suffix}", len(spT_data))
+        save_var_latex(f"f_{spT}_{hoststars}{file_suffix}", round(len(spT_data)/len(dd), 2))
+        
+        # Stellar parameter statistics
+        for param in ['M_st','T_eff_st']:
+            # Median
+            if param == 'T_eff_st':
+                # Convert temperatures to integers to remove .0
+                save_var_latex(f"median_{param}_{spT}_{hoststars}{file_suffix}", 
+                             int(round(spT_data[param].median(), 0)))
+                # Percentiles as integers for temperatures
+                save_var_latex(f"p16_{param}_{spT}_{hoststars}{file_suffix}", 
+                             int(round(spT_data[param].quantile(0.16), 0)))
+                save_var_latex(f"p84_{param}_{spT}_{hoststars}{file_suffix}", 
+                             int(round(spT_data[param].quantile(0.84), 0)))
+            else:
+                # Keep 3 decimal places for masses
+                save_var_latex(f"median_{param}_{spT}_{hoststars}{file_suffix}", 
+                             round(spT_data[param].median(), 3))
+                save_var_latex(f"p16_{param}_{spT}_{hoststars}{file_suffix}", 
+                             round(spT_data[param].quantile(0.16), 3))
+                save_var_latex(f"p84_{param}_{spT}_{hoststars}{file_suffix}", 
+                             round(spT_data[param].quantile(0.84), 3))
+
+
 def past_uv(
     hoststars="all",
     grid=True,
@@ -439,6 +481,14 @@ def past_uv(
     # Get filename suffix based on special parameters
[…]

### 2025-03-25 src/tex/ms.tex
fix linebreak

@@
 We generated two sets of synthetic exoplanet populations, one for FGK-type stars and one for M-type stars.
-The stellar samples are drawn from the Gaia Catalogue of Nearby Stars~\citep{Smart2021} with a maximum Gaia magnitude of \var{M_G_max} and a maximum stellar mass of \var{M_st_max} \SI{}{\Msun}.
+The stellar samples are drawn from the Gaia Catalogue of Nearby Stars~\citep{Smart2021} with a maximum Gaia magnitude of \var{M_G_max} and a maximum stellar mass of \var{M_st_max}~\SI{}{\Msun}.
 We included stars out to a maximum distance $d_{\max}$ that depends on the required planet sample size.

### 2025-03-25 src/tex/ms.tex
clarify what is t=0

@@
 \subsubsection{Habitable zone occupancy and UV flux}\label{sec:met-hz}
-To construct a test \rev{of} the UV~Threshold Hypothesis, we required that life occurs only on planets with sufficient past \gls{UV} irradiation exceeding the origins of life threshold $F_\mathrm{NUV, min}$.
+To construct a test \rev{of} the UV~Threshold Hypothesis, we required that life occurs only on \revv{\gls{HZ}} planets with sufficient past \gls{UV} irradiation exceeding the origins of life threshold $F_\mathrm{NUV, min}$.
 Further, we required this flux to have lasted for a minimum duration $\Delta T_\mathrm{min}$ to allow for a sufficient ``origins timescale''~\citep{Rimmer2023}.
@@
 \rev{Under $H_1$, longer origins timescales have minimal impact on a generic transit survey but significantly decrease the fraction of inhabited planets around FGK stars, as explored in Appendix~\ref{sec:long-origins}.
+}
+\revv{
+The ``clock'' for the required duration $\Delta T_\mathrm{min}$ starts once the planet resides within the habitable zone and receives NUV flux above the threshold $F_\mathrm{NUV, min}$.
 }

### 2025-03-25 src/tex/ms.tex
discuss pre-main sequence effects; restructure section

@@
 we thus considered only rocky planets that may sustain liquid water on their surface, i.e., that occupy their host star's momentary \gls{HZ} during the above period, as well as at the epoch of observation.
-To determine \gls{HZ} occupancy, we took into account the evolution of the host star's luminosity and \gls{HZ} boundaries.
-
 The \gls{HZ} describes a region around a star where a planet with Earth's atmospheric composition and climate feedbacks can maintain liquid water on its surface~\citep[e.g.,][]{Ramirez2017,Ramirez2018,MolLous2022,Spinelli2023,Tuchow2023}.
@@
 
-To determine \gls{HZ} occupancy, we interpolated the stellar luminosity evolution grid of \citet{Baraffe1998} using a Clough Tocher interpolant~\citep[][see left panel of Figure~\ref{fig:hz_nuv_evo}]{Nielson1983,Alfeld1984} to compute the evolution of the inner (runaway greenhouse) and outer (maximum greenhouse) edges as a function of planet mass and stellar spectral type~\citep{Kopparapu2014}.
+\revv{To determine \gls{HZ} occupancy, we took into account the evolution of the host star's luminosity and \gls{HZ} boundaries.
+Specifically, we} interpolated the stellar luminosity evolution grid of \citet{Baraffe1998} using a Clough Tocher interpolant~\citep[][see left panel of Figure~\ref{fig:hz_nuv_evo}]{Nielson1983,Alfeld1984} to compute the evolution of the inner (runaway greenhouse) and outer (maximum greenhouse) edges as a function of planet mass and stellar spectral type~\citep{Kopparapu2014}.
 Being a local interpolation method, Clough Tocher enables rapid processing while producing a smooth interpolating surface that highlights local trends.
 From this, we get each planet's epochs within and outside the \gls{HZ}.
+
+\revv{We note that in particular for late M dwarfs, the pre-main-sequence phase can reach into the age ranges considered here~\citep{Baraffe2015}.
+This may expose planets to elevated levels of radiation, potentially inducing runaway greenhouse effects, water loss, or atmospheric erosion~\citep{Ramirez2014,Luger2015}.
+As such, the actual window of surface habitability may be shorter than the total time a planet spends in the habitable zone during its main-sequence lifetime.
+We defer a detailed treatment of pre-main-sequence effects to future work.
+}
 

### 2025-03-25 src/tex/ms.tex
throw another figure in (stellar BB)
yolo

@@
 While blackbody SEDs are not necessarily representative of M-dwarf SEDs in the \gls{NUV}~\citep{Seager2013,Rugheimer2015}, the spectrally resolved evolution of M-dwarf SEDs as a function of age is not known; we therefore adopt this simplified prescription for the purpose of this sensitivity test.
-We then normalized these SEDs to align with the observed UV fluxes from \citet{Richey-Yowell2023} and used them to calculate mean number flux densities (photons \si{\per\centi\meter\squared\per\second\per\nano\meter}) in the \SIrange{200}{280}{\nano\meter} wavelength range.}
+We then normalized these SEDs to align with the observed UV fluxes from \citet{Richey-Yowell2023} and used them to calculate mean number flux densities (photons \si{\per\centi\meter\squared\per\second\per\nano\meter}) in the \SIrange{200}{280}{\nano\meter} wavelength range }\revv{(see Figure~\ref{fig:normalized_photon_seds})}.
+
+\begin{figure*}[htb]
+    \centering
+    \includegraphics[width=.9\textwidth]{figures/normalized_photon_seds.pdf}
+    \caption{\revv{Photon flux densities of K-type, early M-type, and late M-type stars at different ages.
+    All flux densities are normalized to the observed \gls{NUV} fluxes from \citet{Richey-Yowell2023}.
+    The green area indicates the \SIrange{200}{280}{\nano\meter} wavelength range used to calculate energy fluxes.
+    }}
+    \label{fig:normalized_photon_seds}
+\end{figure*}
 
@@
 \rev{We found that the derived flux densities for all considered spectral types exceed the threshold estimated by \citet{Rimmer2018} during earlier stages of stellar evolution (Figure~\ref{fig:combined_fluxes}).
-Notably -- as in the case of energy fluxes -- later spectral types exhibit higher maximum past \gls{NUV} number flux densities.}
+Notably -- as in the case of energy fluxes -- later spectral types exhibit higher maximum past \gls{NUV} number flux densities }\revv{due to the steeper rise of the blackbody SEDs at shorter wavelengths}.
 

### 2025-03-25 src/tex/ms.tex
discuss abiogenesis zone

@@
 
-The combination of a lack of \gls{UV} radiation today, which makes biosignature gases more detectable~\citep{Segura2005}, and a UV-rich past that may have enabled abiogenesis could make M~dwarfs the preferred targets for biosignature searches.
+\revv{It was found that the ``abiogenesis zone'', defined as the region around a star where quiescent UV flux is sufficient to drive the prebiotic photochemistry necessary for building RNA precursors, does not necessarily overlap with the liquid water \gls{HZ} for stars cooler than K5 ($T_{\mathrm{eff}} \lesssim \SI{4400}{\kelvin}$), implying that such stars may be less favorable for initiating life via UV-driven chemistry~\citep{Rimmer2018}.
+This conclusion, however, is based on a steady-state view of stellar UV fluxes and habitable zone boundaries.
+In our work, we addressed this limitation by explicitly incorporating stellar evolution, including the high-UV early phase of M~dwarfs.
+Our results suggest that planets orbiting M dwarfs may still be viable sites for abiogenesis under UV-dependent pathways.
+In fact, the} combination of a lack of \gls{UV} radiation today, which makes biosignature gases more detectable~\citep{Segura2005}, and a UV-rich past that may have enabled abiogenesis could make M~dwarfs the preferred targets for biosignature searches.
 We note that relevant mission concepts, such as the Large Interferometer for Exoplanets~\citep[\life,][]{Quanz2022,Glauser2024}, include M-dwarf systems among their primary targets~\citep{Kammerer2018,Carrion-Gonzalez2023}.
@@
 
-\rev{We then calculated the ratio of the maximum past \gls{NUV} fluxes in K dwarfs and late M dwarfs in both photon number flux and energy flux regimes.
+\revv{We also calculated } \rev{the ratio of the maximum past \gls{NUV} fluxes in K dwarfs and late M dwarfs in both photon number flux and energy flux regimes.
 We found that this ``dynamic range'' is slightly larger in the photon number flux regime (\var{photon_flux_ratio}) compared to the energy flux regime (\var{integrated_flux_ratio}).

### 2025-03-25 src/tex/ms.tex
discuss timescales, adapt wording of caveats

@@
 }
+\revv{We note that the photochemical timescale does not include subsequent stages such as assembly into functional biomolecules, protocell formation, or the emergence of self-replicating systems -- each of which may require significantly more time.
+However, in the context of the cyanosulfidic scenario, UV irradiation is needed only for this initial synthetic step, and later stages can proceed under different environmental conditions, potentially long after UV levels have declined.
+Moreover, abiogenesis may require many ``failed trials'', we effectively control for this uncertainty by requiring that both sufficient UV flux and habitable conditions persist for a minimum duration $\Delta T_\mathrm{min}$.
+We further require that the planet remains in the habitable zone at the time of observation, thereby capturing both the origination and persistence phases encapsulated by $f_\mathrm{life}$.
+}
 
-\rev{Nonetheless, alternative abiogenesis pathways or combinations of pathways may exhibit continuous or mixed dependencies on UV flux.
-While the step-function formalism is justified for the cyanosulfidic scenario, future work should explore UV dependencies across other scenarios to refine predictions for biosignature distributions and testable hypotheses.
-}
+\rev{Nonetheless, alternative abiogenesis pathways or combinations of pathways may exhibit continuous or mixed dependencies on UV flux.}
+\revv{Future work} should explore UV dependencies across other scenarios to refine predictions for biosignature distributions and testable hypotheses.
 

### 2025-03-25 src/tex/ms.tex
clarify that step-function f_life is only for the first step of OoL.

@@
 Presumably, not all habitable worlds are inhabited and not all inhabited worlds develop detectable biosignatures.
-The fraction of \glspl{EEC} that are both inhabited and exhibit detectable biosignatures at the time of observation is unknown \rev{and is represented by the term $f_\mathrm{life}$.
-This encompasses the probability of life both emerging and persisting to produce detectable biosignatures.}
+The fraction of \glspl{EEC} that are both inhabited and exhibit detectable biosignatures at the time of observation is unknown \rev{and is represented by }\revv{ $f_\mathrm{life}$.}
+\revv{This term captures the combined probability that life \textit{both} emerges on a planet and persists in a form that produces detectable biosignatures by the time of observation.
+}
 \rev{Due to our ignorance about its true value or even its order of magnitude, we draw $f_\mathrm{life}$ from a log-uniform prior probability distribution} \revv{over the range $[10^{-15}, 1]$~\citep[][but see Appendix~\ref{sec:appendix-flife-prior} for a discussion of the impact of different priors]{Balbi2023a}}.
@@
     P(L|F_\mathrm{NUV},H_1) &=  f_\mathrm{life} \quad \text{if } F_\mathrm{NUV}>F_\mathrm{NUV, min}\\
-    P(L|F_\mathrm{NUV},H_1) &=  0               \quad  \text{otherwise}
+    P(L|F_\mathrm{NUV},H_1) &=  0               \quad  \text{otherwise.}
 \end{align}
+\revv{Here, the step-function dependence on $F_\mathrm{NUV}$ applies only to the initial step of abiogenesis, consistent with the cyanosulfidic scenario in which UV irradiation is required for the synthesis of key prebiotic precursors.
+We do not assume that the persistence of life or the presence of detectable biosignatures depends on UV flux.}
 
-where $f_\mathrm{life}$ is the unknown probability of abiogenesis.
 The corresponding null hypothesis $H_\mathrm{null}$ is that there exists no such special value of $F_\mathrm{NUV}$ and that

### 2025-03-25 src/tex/ms.tex
mention winds

@@
 \rev{
-Planets around M~dwarfs, in particular, may be subject to significant atmospheric loss due to the extended high-luminosity pre-main-sequence phase of their host stars~\citep[e.g.,][]{Luger2015}.
-%During this period, extreme X-ray and UV irradiation can drive rapid hydrodynamic escape, potentially stripping a planet of its primordial atmosphere before abiogenesis can occur.
-%Even after this early phase, M-dwarf planets remain exposed to intense stellar activity, leading to further erosion of secondary atmospheres if not replenished through volcanic outgassing or other mechanisms.
+Planets around M~dwarfs, in particular, may be subject to significant atmospheric loss due to the extended high-luminosity pre-main-sequence phase of their host stars}\revv{, and stellar winds can play a major role in stripping secondary atmospheres ~\citep[e.g.,][]{Luger2015,Dong2018,Garcia-Sage2017}.
 Recent works~\citep[e.g.,][]{Coy2024,Luque2024} interpret JWST eclipse measurements as growing evidence for the absence of substantial atmospheres on M-dwarf rocky exoplanets.

### 2025-03-25 src/tex/ms.tex
grammar.

@@
 However, in the context of the cyanosulfidic scenario, UV irradiation is needed only for this initial synthetic step, and later stages can proceed under different environmental conditions, potentially long after UV levels have declined.
-Moreover, abiogenesis may require many ``failed trials'', we effectively control for this uncertainty by requiring that both sufficient UV flux and habitable conditions persist for a minimum duration $\Delta T_\mathrm{min}$.
+Moreover, abiogenesis may require many ``failed trials''.
+We effectively control for this uncertainty by requiring that both sufficient UV flux and habitable conditions persist for a minimum duration $\Delta T_\mathrm{min}$.
 We further require that the planet remains in the habitable zone at the time of observation, thereby capturing both the origination and persistence phases encapsulated by $f_\mathrm{life}$.

### 2025-03-25 src/scripts/SEDs.py
improve figure readability

@@ -143,7 +143,7 @@ def normalize_seds(flux_data, blackbody_seds, wavelengths, age, temperatures):
 
 
 def plot_seds(seds_by_age, wavelengths):
-    plt.rcParams.update({'font.size': 14})
+    plt.rcParams.update({'font.size': 16})  # Increased base font size
     fig, axes = plt.subplots(3, 2, figsize=(15, 20))
     axes = axes.flatten()
     
@@ -163,14 +163,14 @@ def plot_seds(seds_by_age, wavelengths):
                    flux.to(u.erg/u.s/u.cm**2/u.m).value / 1e-13, 
                    label=st, linewidth=2)
         ax.axvspan(200, 280, alpha=0.2, color='green', label='200-280 nm')
-        ax.set_xlabel("Wavelength (nm)", fontsize=16)
-        ax.set_ylabel("Normalized Energy Flux (erg/s/cm²)", fontsize=16)
-        ax.set_title(f"Normalized Blackbody SEDs - Age: {age} Myr", fontsize=18, pad=15)
+        ax.set_xlabel("Wavelength (nm)", fontsize=18)  # Increased from 16
+        ax.set_ylabel("Normalized Energy Flux (erg/s/cm²)", fontsize=18)  # Increased from 16
+        ax.set_title(f"Normalized Blackbody SEDs - Age: {age} Myr", fontsize=18, pad=15)  # Increased from 18
         ax.set_yscale("log")
-        ax.set_xscale("log")
         ax.set_ylim(y_min, y_max)
-        ax.legend(fontsize=14)
-        ax.tick_params(axis='both', which='major', labelsize=14)
+        ax.legend(fontsize=16)  # Increased from 14
+        ax.tick_params(axis='both', which='major', labelsize=18)  # Increased from 14
+        ax.tick_params(axis='both', which='minor', labelsize=18)  # Added minor tick labels
     
     # Remove the empty subplot
     axes[-1].remove()
@@ -179,7 +179,7 @@ def plot_seds(seds_by_age, wavelengths):
 
 
 def plot_photon_flux(photon_fluxes_by_age):
-    plt.rcParams.update({'font.size': 14})
+    plt.rcParams.update({'font.size': 16})  # Increased base font size
     fig, ax = plt.subplots(figsize=(12, 8))
     
     x = np.arange(len(list(photon_fluxes_by_age.values())[0].keys()))
@@ -205,22 +205,22 @@ def plot_photon_flux(photon_fluxes_by_age):
[…]

### 2025-03-27 src/scripts/bioverse_pipeline.py
fix quantiles

@@ -411,24 +411,30 @@ def compute_stellar_stats(dd, hoststars, file_suffix):
         
         # Stellar parameter statistics
         for param in ['M_st','T_eff_st']:
-            # Median
+            # Calculate median and percentiles
+            median = spT_data[param].median()
+            p16 = spT_data[param].quantile(0.16)
+            p84 = spT_data[param].quantile(0.84)
+            
+            # Calculate uncertainties as differences from median
             if param == 'T_eff_st':
                 # Convert temperatures to integers to remove .0
                 save_var_latex(f"median_{param}_{spT}_{hoststars}{file_suffix}", 
-                             int(round(spT_data[param].median(), 0)))
-                # Percentiles as integers for temperatures
-                save_var_latex(f"p16_{param}_{spT}_{hoststars}{file_suffix}", 
-                             int(round(spT_data[param].quantile(0.16), 0)))
+                             int(round(median, 0)))
+                # Save uncertainties as differences from median
                 save_var_latex(f"p84_{param}_{spT}_{hoststars}{file_suffix}", 
-                             int(round(spT_data[param].quantile(0.84), 0)))
+                             int(round(p84 - median, 0)))
+                save_var_latex(f"p16_{param}_{spT}_{hoststars}{file_suffix}", 
+                             int(round(median - p16, 0)))
             else:
                 # Keep 3 decimal places for masses
                 save_var_latex(f"median_{param}_{spT}_{hoststars}{file_suffix}", 
-                             round(spT_data[param].median(), 3))
-                save_var_latex(f"p16_{param}_{spT}_{hoststars}{file_suffix}", 
-                             round(spT_data[param].quantile(0.16), 3))
+                             round(median, 3))
+                # Save uncertainties as differences from median
                 save_var_latex(f"p84_{param}_{spT}_{hoststars}{file_suffix}", 
-                             round(spT_data[param].quantile(0.84), 3))
+                             round(p84 - median, 3))
+                save_var_latex(f"p16_{param}_{spT}_{hoststars}{file_suffix}", 
+                             round(median - p16, 3))
 
 
 def past_uv(

### 2025-03-27 src/tex/ms.tex
fix a typo, add a citation

@@
 The alternate scenario of a surface locally subject to wet-dry cycles requires a planetary exposure to mid-range \gls{UV} irradiation, as a source of energy and an agent of selection in chemical evolution~\citep[e.g.,][]{Deamer2019}.
-This "UV Threshold Hypothesis" states that \gls{UV} light in a specific wavelength range played a constructive role in getting life started on Earth~\citep{Ranjan2016,Ranjan2017c,Rimmer2018,Rapf2016}, and it could provide a probabilistic approach to the interpretation of possible future biosignature detections~\citep[e.g.,][]{Catling2018,Walker2018}.
+This ``UV Threshold Hypothesis'' states that \gls{UV} light in a specific wavelength range played a constructive role in getting life started on Earth~\citep{Ranjan2016,Ranjan2017c,Rimmer2018,Rapf2016}, and it could provide a probabilistic approach to the interpretation of possible future biosignature detections~\citep[e.g.,][]{Catling2018,Walker2018}.
 
@@
 This fraction decreases sharply with increasing threshold flux, as fewer planets receive sufficient \gls{NUV} flux for abiogenesis.
-Another effect responsible for this drop is that some planets receive the required \gls{NUV} flux only before entering the \gls{HZ} -- this is especially likely for M~dwarfs.
+Another effect responsible for this drop is that some planets receive the required \gls{NUV} flux only before entering the \gls{HZ} -- this is especially likely for M~dwarfs\revv{~\citep{Spinelli2024}}.
 For the FGK sample, the fraction of inhabited planets drops at lower threshold fluxes than for the M dwarf sample.
@@
 On the other hand, M~dwarfs show a wider distribution of maximum past \gls{NUV} fluxes in their \glspl{HZ}.
-While old M~dwarfs can be considered low-UV environments, a significant fraction of them emit high \gls{NUV} fluxes into their \gls{HZ} during their early stages, in particular later subtypes~\citep{Richey-Yowell2023}.
+While old M~dwarfs can be considered low-UV environments, a significant fraction of them emit high \gls{NUV} fluxes into their \gls{HZ} during their early stages, in particular later subtypes\revv{~\citep{Richey-Yowell2023,Spinelli2024}}.
 This will help to test the high \gls{NUV} flux end of the UV Threshold Hypothesis; a higher occurrence of biosignatures here would support the hypothesis that a higher \gls{NUV} flux is favorable or necessary for life.
@@
 \end{figure*}
-\rev{We found that the derived flux densities for all considered spectral types exceed the threshold estimated by \citet{Rimmer2018} during earlier stages of stellar evolution (Figure~\ref{fig:combined_fluxes}).
+\rev{We found that the derived flux densities for all considered spectral types exceed the threshold estimated by \citet{Rimmer2018} during earlier stages of stellar evolution\revv{~\citep[Figure~\ref{fig:combined_fluxes}, also see][]{Spinelli2024}}.
 Notably -- as in the case of energy fluxes -- later spectral types exhibit higher maximum past \gls{NUV} number flux densities }\revv{due to the steeper rise of the blackbody SEDs at shorter wavelengths}.
